---
title: "Severity/Admission Data Explore"
output: html_notebook
---
Clinical Indication Needs dx id to icd code to phecode conversion code, look into old code, there is definitely stuff there.

# Load Packages
```{r}
library(bigrquery) #package to import SQL tables into R
library(tidyverse) #package suite for data manipulation, reading data into R, and plotting (ggplot2)
library(glue) #package to use expressions with strings
library(ggseg)
require(gridExtra)
library(interactions)
library(boot)
library(parameters)
```

# Custom Functions
```{r}
panel_save <- function(plot, filename, extension, folder){
  save_name <- paste0(folder, filename, ".", extension)
  ggsave(save_name, plot = plot, dpi = 300)
}
```

# Get project id
```{r}
bq_auth()
# proj_id <- if (exists('proj_id')) proj_id else bq_projects()[1]
proj_id = "scit605-healthymri-2d73d69f"
```

# Get centiles
```{r}
cent_path <- "/mnt/arcus/lab/users/kafadare/slip_centiles_processed_tables/centiles_all_median_th_mpr_2026-04-24.csv"
cent <- read.csv(cent_path) %>%
  mutate(proc_ord_id_wrong = proc_ord_id, 
         proc_ord_id = sub(".*ses-([0-9]+)procId.*", "\\1", session_id))

pat_id_list <- paste0("'", sapply(strsplit(cent$pat_id, ","),trimws), "'", collapse = ", ")
proc_ord_list <- paste0("'", sapply(strsplit(cent$proc_ord_id, ","),trimws), "'", collapse = ", ")
```

# Define Queries
```{r}
encounter_query <- glue('SELECT
  pat_id,
  encounter_id,
  primary_care_ind,
  specialty_care_ind,
  inpatient_ind,
  ed_ind,
  urgent_care_ind,
  enc_type_name,
  effective_datetime,
  effective_age,
  effective_year,
  length_of_stay,
  inpatient_data_id,
  height_cm,
  weight_kg,
  height_raw,
  weight_oz,
  hosp_admsn_age, 
  hosp_admsn_datetime, 
  hosp_disch_age, 
  hosp_disch_datetime,  
  inp_adm_age, 
  inp_adm_datetime,
  coverage_id
  FROM arcus.encounter
  WHERE pat_id IN ({pat_id_list});')

phecode_query <- glue('SELECT
  *
  FROM lab.phecodes_20260228
  WHERE pat_id IN ({pat_id_list});')

proc_query <- glue('SELECT
  proc.proc_ord_id,
  proc.pat_id,
  proc.encounter_id,
  proc.dx_id,
  icd10.code
  FROM arcus.procedure_order_diagnosis as proc
  LEFT JOIN arcus.diagnosis_icd10 as icd10
  ON proc.dx_id = icd10.dx_id
  WHERE proc.proc_ord_id IN ({proc_ord_list});')

proc_enc_query <- glue('SELECT
  proc.proc_ord_id,
  proc.pat_id,
  proc.encounter_id,
  enc.encounter_id,
  enc.dx_id,
  icd10.code as code10,
  icd9.code as code9
  FROM arcus.procedure_order as proc
  INNER JOIN arcus.encounter_diagnosis as enc
  ON proc.encounter_id = enc.encounter_id
  LEFT JOIN arcus.diagnosis_icd10 as icd10
  ON icd10.dx_id = enc.dx_id
  LEFT JOIN arcus.diagnosis_icd9 as icd9
  ON icd9.dx_id = enc.dx_id
  WHERE proc.proc_ord_id IN ({proc_ord_list});')
```

# Download Tables from Arcus
```{r}
# encounter <- bq_project_query(proj_id, encounter_query) %>% bq_table_download(., page_size=3500)
# print(paste0("arcus.encounter table with 25.11 pat_ids, dimensions: ", dim(encounter)))
# 
# phecode <- bq_project_query(proj_id, phecode_query) %>% bq_table_download(., page_size=3500)
# print(paste0("lab.phecodes table with 25.11 pat_ids, dimensions: ", dim(phecode)))

proc <- bq_project_query(proj_id, proc_query) %>% bq_table_download(., page_size=3500)
print(paste0("arcus.procedure_order_dx table with 25.11 pat_ids, dimensions: ", dim(proc)))

proc_enc <- bq_project_query(proj_id, proc_enc_query) %>% bq_table_download(., page_size=3500)
print(paste0("arcus.proc_encedure_order_dx table with 25.11 pat_ids, dimensions: ", dim(proc_enc)))
```
# Get annotation table for validation
```{r}
indication_icd10 <- read.csv("/mnt/arcus/lab/users/kafadare/indication_files/indication_icd10.csv") %>% select(-X)
indication_icd9 <- read.csv("/mnt/arcus/lab/users/kafadare/indication_files/indication_icd9.csv") %>% select(-X)
```

# Check dimensions for procedure order table
```{r}
# getting dx from procedure_diagnosis
length(unique(cent$proc_ord_id))
sum(cent$proc_ord_id %in% proc$proc_ord_id)/nrow(cent) #57%
sum(is.na(proc$dx_id)) #0
sum(is.na(proc$code)) #312
length(unique(proc$proc_ord_id)) #7642

# getting dx from encounter_diagnosis
sum(cent$proc_ord_id %in% proc_enc$proc_ord_id)/nrow(cent) #86%
sum(is.na(proc_enc$dx_id)) #0
sum(is.na(proc_enc$code10)) #2311
sum(is.na(proc_enc$code10) & is.na(proc_enc$code9)) #3
length(unique(proc_enc$proc_ord_id))

sum(!(proc$proc_ord_id %in% proc_enc$proc_ord_id)) #0
```

# Match based on proc_ord_id
```{r}
indication_validate <- merge(proc_enc %>%
                               rename(enc_icd10_code = code10,
                                      enc_icd9_code = code9), 
                             indication_icd10 %>%
                               rename(annot_icd10_code = icd10_code), by = "proc_ord_id")


# one row per report, with code sets from each pipeline
report_sets <- indication_validate %>%
  group_by(proc_ord_id) %>%
  summarise(
    enc_icd10_code = list(unique(na.omit(enc_icd10_code))),
    annot_icd10_code = list(unique(na.omit(annot_icd10_code))),
    snomed_name = list(unique(na.omit(snomed_name))),
    .groups = "drop"
  ) %>%
  mutate(
    matched_codes = map2(enc_icd10_code, annot_icd10_code, intersect),
    unmatched_p1  = map2(enc_icd10_code, annot_icd10_code, setdiff),
    unmatched_p2  = map2(annot_icd10_code, enc_icd10_code, setdiff),
    n_matched = lengths(matched_codes),
    report_match_status = if_else(
      n_matched > 0,
      "Report has ≥1 matched code",
      "Report has 0 matched codes"
    )
  )

# plot
matched_codes_plot_df <- report_sets %>%
  select(proc_ord_id, matched_codes, ) %>%
  unnest(matched_codes) %>%
  rename(code = matched_codes) %>%
  distinct(proc_ord_id, code) %>%
  count(code, name = "n_reports") %>%
  mutate(code = fct_reorder(code, n_reports))

matched_codes_plot_df <- merge(matched_codes_plot_df, indication_icd10 %>% select(icd10_code, snomed_name) %>% rename(code = icd10_code),
                               by = "code", all.x = TRUE, all.y = FALSE) %>% distinct(snomed_name, .keep_all = TRUE)


ggplot(matched_codes_plot_df, aes(x = n_reports, y = snomed_name)) +
  geom_col() +
  labs(
    x = "Number of reports",
    y = "Matched Snomed Name"
  ) +
  theme_classic()


ind_long <- indication_validate %>%
  select(proc_ord_id, enc_icd10_code, annot_icd10_code, snomed_name) %>%
  pivot_longer(
    cols = c(enc_icd10_code, annot_icd10_code),
    names_to = "ind_source",
    values_to = "code10",
    values_drop_na = TRUE
  ) %>%
  mutate(
    ind_source = recode(
      ind_source,
      enc_icd10_code = "enc_dx_table",
      annot_icd10_code = "annot"
    ),
    code10 = as.character(code10)
  ) %>%
  distinct(proc_ord_id, ind_source, code10, snomed_name)

report_code_status <- ind_long %>%
  count(proc_ord_id, code10, name = "n_source") %>%
  mutate(code_status = if_else(n_source == 2, "Matched", "Unmatched"))



ind_match <- indication_validate %>%
  filter(enc_icd10_code == annot_icd10_code) %>%
  group_by(proc_ord_id)

length(unique(ind_match$proc_ord_id))/ length(unique(indication_validate$proc_ord_id)) # 76.3%

ind_nomatch <- indication_validate %>%
  filter(!(proc_ord_id %in% ind_match$proc_ord_id)) %>%
  group_by(proc_ord_id)
length(unique(ind_nomatch$proc_ord_id))/ length(unique(indication_validate$proc_ord_id)) # 23.6%
```

# Show schematic of matching indications and non-matching indications
```{r}
table(ind_match$snomed_name)

table(ind_nomatch$snomed_name)
ind_match_byname <- ind_match %>%
  group_by(snomed_name) %>%
  summary()

```

# Plot matches (show N in each filtering step)
```{r}


```