---
title: "data-clean-save"
author: "Eren Kafadar"
date: "2025-11-18"
output: html_document
editor_options: 
  chunk_output_type: console
params:
  output_folder : "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/data_clean/"
  centiles_path: "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/centile_csv/slip_median_centiles_041526.csv"
  participants_path: "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/participants_ga_imputed.csv"
  qc_metrics_path: "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/fs7.4.1-clinical-reconall-qc.csv"
  selected_scans_path: "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/slip_selected_scans_041526.csv"
  type: "median"  
  add_qc_flag: "TRUE"
---

# Define params if they do not exist
```{r, echo = FALSE}
if (!exists("params")) {
  params <- list(
  output_folder = "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/data_clean/",
  centiles_path = "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/centile_csv/slip_median_centiles_041526.csv",
  participants_path = "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/participants_ga_imputed.csv",
  qc_metrics_path = "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/fs7.4.1-clinical-reconall-qc.csv",
  selected_scans_path = "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/slip_selected_scans_041526.csv",
  type = "median",
  add_qc_flag = "TRUE"
  )
}
``` 

# Load packages
```{r, echo = FALSE}
#Load libraries
library(dplyr)
library(magrittr)
library(readr)
library(stringr)
library(tidyr)
library(glue)
#For figures/plotting
library(ggplot2)
library(gridExtra)
library(ggseg)
#library(ggpubr)
library(ggcorrplot)
library(cowplot)
library(egg)
library(interactions)
library(growthstandards)
library(gtsummary)
library(svglite)
library(childsds)
```

# Controls
```{r setup, echo = TRUE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)
#set command
save_tables <- TRUE
plot <- TRUE
save_figs <- TRUE

set.seed(42)

output_folder <- params$output_folder
centiles_path <- params$centiles_path
participants_path <- params$participants_path
qc_metrics_path <- params$qc_metrics_path
selected_scans_path <- params$selected_scans_path
type <- params$type
add_qc_flag <- params$add_qc_flag
```

# Define Paths
```{r, echo = TRUE}
knitr::opts_chunk$set(echo = TRUE)

#define paths
code_path <- "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/"
vars_path <-  "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/output/vars_2511_new/"

## output/save paths
if (!dir.exists(output_folder)) dir.create(output_folder)
csv_save_path <- glue("{output_folder}csv/")
if (!dir.exists(csv_save_path)) dir.create(csv_save_path)
plot_save_path <- paste0(output_folder, "plots/")
if (!dir.exists(plot_save_path)) dir.create(plot_save_path)

#load custom data transform and bw/ga qc functions
source(paste0(code_path,"slip_data_utils.R"))
```

# Load Data
```{r, echo = FALSE}
participants <- read.csv(participants_path)
selected_scans <- read.csv(selected_scans_path)
qc_metrics <- read.csv(qc_metrics_path) %>% filter(scan_id %in% selected_scans$scan_id)
centiles <- read.csv(centiles_path)
```

# Load Phenotype Var Names
```{r, echo = FALSE}
full_vars <- read.csv(paste0(vars_path,"all_vars.csv")) %>% pull(x)
global_vars <- read.csv(paste0(vars_path,"global_vars.csv")) %>% pull(x)
vol_vars <- read.csv(paste0(vars_path,"vol_vars.csv")) %>% pull(x)
sa_vars <- read.csv(paste0(vars_path,"sa_vars.csv")) %>% pull(x)
th_vars <- read.csv(paste0(vars_path,"th_vars.csv")) %>% pull(x)
```

# QC inter-variable correlations
```{r, echo = FALSE}
#correlation matrix and visualization
cor_matrix <- cor(qc_metrics[,names(qc_metrics)[!grepl("_id$", names(qc_metrics))]], use = "pairwise.complete.obs")
p_mat <- cor_pmat(qc_metrics[,names(qc_metrics)[!grepl("_id$", names(qc_metrics))]])

ggcorrplot(
  cor_matrix,
  p.mat = p_mat,
  type = "lower",
  insig = "blank",
  lab = TRUE
)
```

# QC metrics prep
```{r}
if(type == "median"){
##group by session-id and take median of euler_mean
qc_metrics <- qc_metrics %>%
  group_by(session_id) %>%
  summarise(median_euler_mean = median(euler_mean),
            .groups = "drop") %>%
  left_join(qc_metrics, by = "session_id") %>%
  rename(participant_id = subject_id)
}
```

# Participant DF prep
double/incongruent BW/GA info is dealt with in previous steps before modeling
```{r, echo = FALSE}
participants_vars_select <- c("pat_id", "study_id", "participant_id", "session_id", "scan_id", "sex", 
                "age_at_scan",
                "height_at_enc", "weight_at_enc", "hc_at_enc", 
                "ga_days_total_qc", "ga_imputed_nomedian_days_total", "ga_imputed_days_total",
                "birth_weight_kg_qc", "birth_length_cm",
                "race", "ethnicity",
                "site", "avg_grade", "useable", "proc_ord_id")

participants <- participants %>%
  select(any_of(participants_vars_select)) %>%
           rename(gestational_age = ga_days_total_qc,
                  ga_bw_imputed = ga_imputed_nomedian_days_total,
                  ga_full_imputed = ga_imputed_days_total,
                  birth_weight_kg = birth_weight_kg_qc,
                  site_hospital_name = site)
```

# check for extra variables in the centiles df
```{r, echo = FALSE}
vars_to_select <- c(full_vars, "participant_id", "session_id", "scan_id", "site", "sex", 
                    "log10_age_days", "age_days", 
                    "dx", "obs_weight")
vars_extra <- names(centiles)[!(names(centiles) %in% vars_to_select)]
vars_extra_print <- vars_extra[!grepl("glasser", vars_extra)]
print(glue("Additional variables in the centiles csv provided (other than glasser): {vars_extra_print}"))
```

# run data cleanup
```{r, echo = FALSE}
print(glue("ALL Centiles DF dimensions before 'cleanup' script: {c(dim(centiles))}"))
vars_notfound <- vars_to_select[!(vars_to_select %in% names(centiles))]
print(glue("Variables not found in the centiles csv: {vars_notfound}"))
#centiles <- centiles %>% select(any_of(vars_to_select))
participants_vars_select_clean <- c("pat_id", "study_id", "participant_id", 
                                    "session_id", "sex", "age_at_scan",
                "height_at_enc", "weight_at_enc", "hc_at_enc", 
                "gestational_age", "ga_bw_imputed", "ga_full_imputed",
                "birth_weight_kg",
                "race", "ethnicity",
                "site_hospital_name", "avg_grade", "useable", "proc_ord_id")
clean_centiles <- slip_centdf_clean(centiles, 
                                    participants, 
                                    qc_metrics, 
                                    type = type, 
                                    add_qc = add_qc_flag,
                                    participants_vars_select = participants_vars_select_clean,
                                    cent_df_vars_select = vars_to_select)
centiles_distinct <- clean_centiles$distinct
centiles_all <- clean_centiles$all
print(glue("DISTINCT Centiles DF dimensions: {c(dim(centiles_distinct))}"))
print(glue("ALL Centiles DF dimensions: {c(dim(centiles_all))}"))

rm(centiles)
```


Need to get median (within 180 days) weight/height from arcus table. Will do this later
<!-- # Weight/ Height Centile & QC -->

<!-- ## Weight/Height Availability (before QC) -->
<!-- ```{r} -->
<!-- print(glue("Number of Data Points with weight at enc {sum(!is.na(centiles_all$weight_at_enc))}")) -->
<!-- print(glue("Number of Data Points with height at enc {sum(!is.na(centiles_all$height_at_enc))}")) -->
<!-- print(glue("Number of Data Points with weight and height at enc {sum(!is.na(centiles_all$height_at_enc) & !is.na(centiles_all$weight_at_enc))}")) -->

<!-- print(glue("Number of Data Points with weight AND height at enc AND gestational age {sum(!is.na(centiles_all$height_at_enc) & !is.na(centiles_all$weight_at_enc) & !is.na(centiles_all$gestational_age))}")) -->
<!-- print(glue("Number of Data Points with weight AND height at enc AND birth weight {sum(!is.na(centiles_all$height_at_enc) & !is.na(centiles_all$weight_at_enc) & !is.na(centiles_all$birth_weight_kg))}")) -->
<!-- ``` -->

<!-- ## Weight/Height Plots -->
<!-- ```{r, echo = FALSE} -->
<!-- ggplot(centiles_all, aes(x = weight_at_enc, y = age_at_scan_years)) + -->
<!--   geom_point(size = 0.5, alpha = 0.2) + -->
<!--   geom_hline(yintercept = 19, color = "#974343", alpha = 0.5) + -->
<!--   labs(x = "Weight at Encounter (kg?)",  -->
<!--        y = "Age At Scan (yrs)") + -->
<!--   theme_classic() -->

<!-- ggplot(centiles_all, aes(x = height_at_enc, y = age_at_scan_years)) + -->
<!--   geom_point(size = 0.5, alpha = 0.2) + -->
<!--   geom_hline(yintercept = 19, color = "#974343", alpha = 0.5) + -->
<!--   labs(x = "Height at Encounter (cm?)",  -->
<!--        y = "Age At Scan (yrs)") + -->
<!--   theme_classic() -->
<!-- ``` -->

<!-- ## Calculate Weight/Height z-score -->
<!-- Using childsds package, who.ref reference charts -->
<!-- ```{r, echo = FALSE} -->
<!-- wc_out <- cdcanthro(centiles_all %>% -->
<!--                    mutate(age_at_scan_months = age_at_scan/30.4375, -->
<!--                           sex = recode(sex, Male = "m", Female = "f")),  -->
<!--                  age = age_at_scan_months, -->
<!--                  wt = weight_at_enc, -->
<!--                  ht = height_at_enc, -->
<!--                  all = FALSE) %>% -->
<!--   select(participant_id, session_id, wap, waz, hap, haz, bmip, bmiz) %>% -->
<!--   rename(weight_cent = wap, weight_zscore = waz, -->
<!--          height_cent = haz, height_zscore = haz, -->
<!--          bmi_cent = bmip, bmi_zscore = bmiz) -->


<!-- centiles_all <- merge(centiles_all, wc_out, by = c("participant_id", "session_id"), -->
<!--                         all.x = TRUE) -->

<!-- ggplot(centiles_all, aes(x = weight_at_enc, y = age_at_scan_years, -->
<!--                          color = weight_cent)) + -->
<!--   geom_point(size = 0.5, alpha = 0.2) + -->
<!--   geom_hline(yintercept = 19, color = "#974343", alpha = 0.5) + -->
<!--   scale_color_gradient(low = "#0595eb",high = "#4fe30e", na.value = "#d73027") + -->
<!--   labs(x = "Weight at Encounter (kg?)",  -->
<!--        y = "Age At Scan (yrs)") + -->
<!--   theme_classic() -->



<!-- # centiles_all <- centiles_all %>% -->
<!-- # mutate(weight_at_enc_sds = case_when( -->
<!-- #     age_at_scan_years<5 ~ sds(value = weight_at_enc, -->
<!-- #                                             age = age_at_scan_years, -->
<!-- #                                             sex = sex, -->
<!-- #                                             item = "weight", -->
<!-- #                                             ref = who.ref, -->
<!-- #                                             type = "SDS", -->
<!-- #                                             male = "Male", -->
<!-- #                                             female = "Female" -->
<!-- #                                             ), -->
<!-- #     age_at_scan_years>5 & age_at_scan_years<19 ~ sds(value = weight_at_enc, -->
<!-- #                                             age = age_at_scan_years, -->
<!-- #                                             sex = sex, -->
<!-- #                                             item = "weight", -->
<!-- #                                             ref = who2007.ref, -->
<!-- #                                             type = "SDS", -->
<!-- #                                             male = "Male", -->
<!-- #                                             female = "Female" -->
<!-- #                                             ), -->
<!-- #     TRUE ~ NA_real_)) -->
<!-- #  -->
<!-- # sum(!is.na(centiles_all$weight_at_enc_sds)) -->
<!-- # sum(!is.na(centiles_all$weight_at_enc)) -->
<!-- ggplot(centiles_all %>% filter(!is.na(weight_at_enc) & is.na(weight_at_enc_sds)),  -->
<!--        aes(x = weight_at_enc, y = age_at_scan_years)) + -->
<!--   geom_point(size = 0.5, alpha = 0.2) + -->
<!--   geom_hline(yintercept = 19, color = "#974343", alpha = 0.5) + -->
<!--   labs(x = "Weight at Encounter (kg?)",  -->
<!--        y = "Age At Scan (yrs)", -->
<!--        title = "Weight Available but SDS NA") + -->
<!--   theme_classic() -->

<!-- ggplot(centiles_all %>% filter(age_at_scan_years > 9.5),  -->
<!--        aes(x = weight_at_enc, y = age_at_scan_years, color = is.na(weight_at_enc_sds))) + -->
<!--   geom_point(size = 0.5, alpha = 0.2) + -->
<!--   geom_hline(yintercept = 19, color = "#974343", alpha = 0.5) + -->
<!--   #scale_color_gradient(low = "#0595eb",high = "#4fe30e", na.value = "#d73027") + -->
<!--   labs(x = "Weight at Encounter (kg?)",  -->
<!--        y = "Age At Scan (yrs)", -->
<!--        title = "Age > 9.5 Years") + -->
<!--   theme_classic() -->


<!-- centiles_all$weight_at_enc_sifelse(centiles_all$age_at_scan_years < 5, -->
<!--                                          sds(value = centiles_all[centiles_all$age_at_scan_years < 5,]$weight_at_enc, -->
<!--                                             age = centiles_all[centiles_all$age_at_scan_years < 5,]$age_at_scan_years, -->
<!--                                             sex = centiles_all[centiles_all$age_at_scan_years < 5,]$sex, -->
<!--                                             item = "weight", -->
<!--                                             ref = who.ref, -->
<!--                                             type = "SDS", -->
<!--                                             male = "Male", -->
<!--                                             female = "Female" -->
<!--                                             ), NA) -->

<!-- centiles_all$weight_at_enc_sds <- ifelse(centiles_all$age_at_scan_years > 5 & centiles_all$age_at_scan_years < 19, -->
<!--                                          sds(value = centiles_all[centiles_all$age_at_scan_years > 5 & centiles_all$age_at_scan_years < 19,]$weight_at_enc, -->
<!--                                             age = centiles_all[centiles_all$age_at_scan_years > 5 & centiles_all$age_at_scan_years < 19,]$age_at_scan_years, -->
<!--                                             sex = centiles_all[centiles_all$age_at_scan_years > 5 & centiles_all$age_at_scan_years < 19,]$sex, -->
<!--                                             item = "weight", -->
<!--                                             ref = who2007.ref, -->
<!--                                             type = "SDS", -->
<!--                                             male = "Male", -->
<!--                                             female = "Female" -->
<!--                                             ), NA) -->


<!-- centiles_all$weight_at_enc_sds <- ifelse(centiles_all$age_at_scan_years < 19, -->
<!--                                          sds(value = centiles_all[centiles_all$age_at_scan_years < 19,]$weight_at_enc, -->
<!--                                             age = centiles_all[centiles_all$age_at_scan_years < 19,]$age_at_scan, -->
<!--                                             sex = centiles_all[centiles_all$age_at_scan_years < 19,]$sex, -->
<!--                                             item = "weight", -->
<!--                                             ref = who2007.ref, -->
<!--                                             type = "SDS", -->
<!--                                             male = "Male", -->
<!--                                             female = "Female" -->
<!--                                             ), NA) -->
<!-- centiles_all$height_at_enc_sds <- sds(value = centiles_all$height_at_enc, -->
<!--                                             age = centiles_all$age_at_scan, -->
<!--                                             sex = centiles_all$sex, -->
<!--                                             item = "height", -->
<!--                                             ref = who.ref, -->
<!--                                             type = "SDS", -->
<!--                                             male = "Male", -->
<!--                                             female = "Female" -->
<!--                                             ) -->
<!-- ``` -->

# Plots with Birth Weight Percentile/Z-score
```{r, echo = FALSE}
ggplot(centiles_distinct, aes(x = bwp_fen, y = gestational_age)) +
  geom_point(size = 0.5, alpha = 0.2) +
  labs(x = "BW Percentile (Fenton)", 
       y = "Gestational Age (days)") +
  theme_classic()

ggplot(centiles_distinct, aes(x = bwz_fen, y = gestational_age)) +
  geom_point(size = 0.5, alpha = 0.2) +
  labs(x = "BW Z-score (Fenton)", 
       y = "Gestational Age (days)") +
  theme_classic()

ggplot(centiles_distinct, aes(x = bwz_fen, y = birth_weight_kg)) +
  geom_point(size = 0.5, alpha = 0.2) +
  labs(x = "BW Z-score (Fenton)", 
       y = "Birth Weight (kg)") +
  theme_classic()

```

# SAVE Data Frames
```{r, echo = FALSE}
if (save_tables == TRUE){
#one row per participant
distinct_filename <- paste0(csv_save_path, "centiles_distinct_", type, "_", Sys.Date(), ".csv")
print(distinct_filename)
write.csv(centiles_distinct, distinct_filename)
#all rows, included repeats of participants at different timepoints/ages
all_filename <- paste0(csv_save_path, "centiles_all_", type, "_", Sys.Date(), ".csv")
print(all_filename)
write.csv(centiles_all, all_filename)
}
```

