---
title: "Severity/Admission Data Explore"
output: html_notebook
---
Clinical Indication Needs dx id to icd code to phecode conversion code, look into old code, there is definitely stuff there.

# Load Packages
```{r}
library(bigrquery) #package to import SQL tables into R
library(tidyverse) #package suite for data manipulation, reading data into R, and plotting (ggplot2)
library(glue) #package to use expressions with strings
library(ggseg)
require(gridExtra)
library(interactions)
library(boot)
library(parameters)
```

# Custom Functions
```{r}
panel_save <- function(plot, filename, extension, folder){
  save_name <- paste0(folder, filename, ".", extension)
  ggsave(save_name, plot = plot, dpi = 300)
}
```

# Get project id
```{r}
bq_auth()
# proj_id <- if (exists('proj_id')) proj_id else bq_projects()[1]
proj_id = "scit605-healthymri-2d73d69f"
```

# Get centiles
```{r}
cent_path <- "/mnt/arcus/lab/users/kafadare/slip_centiles_processed_tables/centiles_all_median_th_mpr_2026-04-24.csv"
cent <- read.csv(cent_path) %>%
  mutate(proc_ord_id_wrong = proc_ord_id, 
         proc_ord_id = sub(".*ses-([0-9]+)procId.*", "\\1", session_id))

pat_id_list <- paste0("'", sapply(strsplit(cent$pat_id, ","),trimws), "'", collapse = ", ")
proc_ord_list <- paste0("'", sapply(strsplit(cent$proc_ord_id, ","),trimws), "'", collapse = ", ")
```

# Define Queries
```{r}
encounter_query <- glue('SELECT
  pat_id,
  encounter_id,
  primary_care_ind,
  specialty_care_ind,
  inpatient_ind,
  ed_ind,
  urgent_care_ind,
  enc_type_name,
  effective_datetime,
  effective_age,
  effective_year,
  length_of_stay,
  inpatient_data_id,
  height_cm,
  weight_kg,
  height_raw,
  weight_oz,
  hosp_admsn_age, 
  hosp_admsn_datetime, 
  hosp_disch_age, 
  hosp_disch_datetime,  
  inp_adm_age, 
  inp_adm_datetime,
  coverage_id
  FROM arcus.encounter
  WHERE pat_id IN ({pat_id_list});')

# phecode_query <- glue('SELECT
#   *
#   FROM lab.phecodes_20260228
#   WHERE pat_id IN ({pat_id_list});')

proc_query <- glue('SELECT
  proc.proc_ord_id,
  proc.pat_id,
  proc.encounter_id,
  proc.dx_id,
  icd10.code
  FROM arcus.procedure_order_diagnosis as proc
  LEFT JOIN arcus.diagnosis_icd10 as icd10
  ON proc.dx_id = icd10.dx_id
  WHERE proc.proc_ord_id IN ({proc_ord_list});')

proc_enc_query <- glue('SELECT
  proc.proc_ord_id,
  proc.pat_id,
  proc.encounter_id,
  enc.encounter_id,
  enc.dx_id,
  icd10.code as code10,
  icd9.code as code9
  FROM arcus.procedure_order as proc
  INNER JOIN arcus.encounter_diagnosis as enc
  ON proc.encounter_id = enc.encounter_id
  LEFT JOIN arcus.diagnosis_icd10 as icd10
  ON icd10.dx_id = enc.dx_id
  LEFT JOIN arcus.diagnosis_icd9 as icd9
  ON icd9.dx_id = enc.dx_id
  WHERE proc.proc_ord_id IN ({proc_ord_list});')
```

# Download Tables from Arcus
```{r}
encounter <- bq_project_query(proj_id, encounter_query) %>% bq_table_download(., page_size=3500)
print(paste0("arcus.encounter table with 25.11 pat_ids, dimensions: ", dim(encounter)))

# phecode <- bq_project_query(proj_id, phecode_query) %>% bq_table_download(., page_size=3500)
# print(paste0("lab.phecodes table with 25.11 pat_ids, dimensions: ", dim(phecode)))

proc <- bq_project_query(proj_id, proc_query) %>% bq_table_download(., page_size=3500)
print(paste0("arcus.procedure_order_dx table with 25.11 pat_ids, dimensions: ", dim(proc)))

proc_enc <- bq_project_query(proj_id, proc_enc_query) %>% bq_table_download(., page_size=3500)
print(paste0("arcus.procedure_order table with 25.11 pat_ids, dimensions: ", dim(proc_enc)))
```

# Get icd codes to snomed conversion lists from arcus
```{r}
# Get snomed tables from arcus for snomed codes in the annotations table
snomed_icd10_query <- glue('SELECT
  icd10.icd10_code,
  icd10.snomed_code,
  icd10.snomed_name
  FROM arcus.icd10_to_snomed AS icd10;')

snomed_icd9_query <- glue('SELECT
  icd9.icd9_code,
  icd9.snomed_code,
  icd9.snomed_name
  FROM arcus.icd9_to_snomed AS icd9;')

snomed_icd10 <- bq_project_query(proj_id, snomed_icd10_query) %>% bq_table_download(., page_size=3500)
print(paste0("arcus.icd10_to_snomed table with 25.11 pat_ids, dimensions: ", dim(snomed_icd10)))

snomed_icd9 <- bq_project_query(proj_id, snomed_icd9_query) %>% bq_table_download(., page_size=3500)
print(paste0("arcus.icd9_to_snomed table with 25.11 pat_ids, dimensions: ", dim(snomed_icd9)))
```

# Get cleaned up annotation table
```{r}
indication_icd10 <- read.csv("/mnt/arcus/lab/users/kafadare/indication_files/indication_icd10.csv") %>% select(-X)
indication_icd9 <- read.csv("/mnt/arcus/lab/users/kafadare/indication_files/indication_icd9.csv") %>% select(-X)
```

# Check availability of diagnosis based on SQL table
```{r}
# getting dx from procedure_diagnosis
length(unique(cent$proc_ord_id))
sum(cent$proc_ord_id %in% proc$proc_ord_id)/nrow(cent) #57%
sum(is.na(proc$dx_id)) #0
sum(is.na(proc$code)) #312
length(unique(proc$proc_ord_id)) #7642

# getting dx from encounter_diagnosis
sum(cent$proc_ord_id %in% proc_enc$proc_ord_id)/nrow(cent) #86%
sum(is.na(proc_enc$dx_id)) #0
sum(is.na(proc_enc$code10)) #2311
sum(is.na(proc_enc$code10) & is.na(proc_enc$code9)) #3
length(unique(proc_enc$proc_ord_id)) #11491

sum(!(proc$proc_ord_id %in% proc_enc$proc_ord_id)) #0
```


# Match based on proc_ord_id
```{r}

indication_icd10 <- merge(indication, snomed_icd10 %>%
                select(-c("snomed_name")), by = "snomed_code", all.x = TRUE, all.y = FALSE)
indication_icd9 <- merge(indication, snomed_icd9 %>%
                           select(-c("snomed_name")),
                    by = "snomed_code", all.x = TRUE, all.y = FALSE)
annot_enc <- merge(proc_enc %>%
                               rename(enc_icd10_code = code10,
                                      enc_icd9_code = code9), 
                             indication_icd10 %>%
                               rename(annot_icd10_code = icd10_code,
                                      annot_snomed_code = snomed_code,
                                      annot_snomed_name = snomed_name), by = "proc_ord_id")

annot_enc <- merge(annot_enc, 
                   snomed_icd10 %>%
                     rename(enc_icd10_code = icd10_code,
                            enc_snomed_code = snomed_code,
                            enc_snomed_name = snomed_name), by = "enc_icd10_code",
                   all.x = TRUE, all.y = FALSE)


# one row per report, looking at codes that match and do not match
# report_sets_icd <- annot_enc %>%
#   group_by(proc_ord_id) %>%
#   summarise(
#     enc_icd10_code = list(unique(na.omit(enc_icd10_code))),
#     annot_icd10_code = list(unique(na.omit(annot_icd10_code))),
#     annot_snomed_name = list(unique(na.omit(annot_snomed_name))),
#     .groups = "drop"
#   ) %>%
#   mutate(
#     matched_codes = map2(enc_icd10_code, annot_icd10_code, intersect),
#     unmatched_enc  = map2(enc_icd10_code, annot_icd10_code, setdiff),
#     unmatched_annot  = map2(annot_icd10_code, enc_icd10_code, setdiff),
#     n_matched = lengths(matched_codes),
#     report_match_status = if_else(
#       n_matched > 0,
#       TRUE,
#       FALSE 
#     )
#   )

report_sets_icd <- annot_enc %>%
  group_by(proc_ord_id) %>%
  summarise(
    enc_icd10_code = list(unique(na.omit(enc_icd10_code))),
    
    annot_dx = list(
      tibble(
        icd10_code = annot_icd10_code,
        snomed_name = annot_snomed_name
      ) %>%
        filter(!is.na(icd10_code)) %>%
        distinct()
    ),
    
    .groups = "drop"
  ) %>%
  mutate(
    matched_dx = map2(
      enc_icd10_code,
      annot_dx,
      ~ .y %>%
        filter(icd10_code %in% .x)
    ),
    
    unmatched_enc = map2(
      enc_icd10_code,
      annot_dx,
      ~ setdiff(.x, .y$icd10_code)
    ),
    
    unmatched_annot = map2(
      annot_dx,
      enc_icd10_code,
      ~ setdiff(.x$icd10_code, .y)
    ),
    
    n_matched = map_int(matched_dx, nrow),
    
    report_match_status = n_matched > 0
  )


sum(report_sets_icd$report_match_status == TRUE)/nrow(report_sets_icd)
```


# Tables of matched and unmatched ICD Codes
```{r}
matched_pair_counts <- report_sets_icd %>%
  select(proc_ord_id, matched_dx) %>%
  unnest(matched_dx) %>%
  count(icd10_code, snomed_name, name = "n_occurrences") %>%
  arrange(desc(n_occurrences))

kableExtra::kable(head(matched_pair_counts, 10))

unmatched_annot_counts <- report_sets_icd %>%
  filter(n_matched == 0) %>%
  select(annot_dx) %>%
  unnest(annot_dx) %>%
  count(icd10_code, snomed_name, name = "n_occurrences") %>%
  arrange(desc(n_occurrences))

kableExtra::kable(head(unmatched_annot_counts, 10))
```

# Patterns of reports with no matches
```{r}
no_match <- report_sets_icd %>%
  filter(n_matched == 0)


# 2. Original annotation SNOMED names and their ICD codes
annot_no_match <- report_sets_icd %>%
  filter(n_matched == 0) %>%
  select(proc_ord_id, annot_dx) %>%
  unnest(annot_dx) %>%
  rename(
    annot_icd10 = icd10_code,
    annot_snomed_name = snomed_name
  ) %>%
  distinct(proc_ord_id, annot_icd10, annot_snomed_name)

enc_no_match <- report_sets_icd %>%
  filter(n_matched == 0) %>%
  select(proc_ord_id, enc_icd10_code) %>%
  unnest(enc_icd10_code) %>%
  rename(enc_icd10 = enc_icd10_code) %>%
  distinct(proc_ord_id, enc_icd10)

enc_no_match_snomed <- enc_no_match %>%
  left_join(
    snomed_icd10 %>%
      select(icd10_code, snomed_name) %>%
      distinct(),
    by = c("enc_icd10" = "icd10_code")
  ) %>%
  rename(
    enc_snomed_name = snomed_name
  )
```