---
title: "lmsz_ga_plotting"
author: "Eren Kafadar"
date: "2026-04-22"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Setup
## Load libraries
```{r}
# LMSz for full sample test GA term
# Load libraries
require(dplyr)
require(magrittr)
require(readr)
require(stringr)
require(tidyr)
require(gamlssTools) # Margaret's package
require(gamlss.add)
require(tibble)
require(glue)
# For figures/plotting
require(ggplot2)
require(patchwork)
# For figures/plotting
require(gridExtra)
require(ggsegDKT)
require(ggseg)
# require(ggpubr)
require(ggcorrplot)
require(cowplot)
require(egg)
require(interactions)
require(growthstandards)
require(gtsummary)
require(svglite)

set.seed(42)
params <- list(
  output_folder = "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/output/lmsz_median_Th-MPR_25.11_GA_splines_v2/",
  centiles_data_path = "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/output/median-Th-MPR-allgrades-25.11/data_clean/centiles_distinct_2026-03-18.csv", 
  raw_data_path = "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11/raw_csv/slip_median_th_mpr_2026-03-17.csv",
  models_path = "/mnt/isilon/bgdlab_processing/braincharts/SLIP/2025_03/code/gamlss/RDS/"
)
save_model = TRUE
reweight = FALSE
```


## set folder paths
```{r}
code_path <- "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/"
vars_path <-  "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/output/vars/"
output_folder <- params$output_folder
centiles_data_path <- params$centiles_data_path
raw_data_path <- params$raw_data_path
models_path <- params$models_path
# output folder
save_path <- params$output_folder
if (!dir.exists(save_path)) dir.create(save_path)

##  output subfolder paths
csv_save_path <-  paste0(save_path, "results_csv/")
if (!dir.exists(csv_save_path)) dir.create(csv_save_path)
plot_save_path <- paste0(save_path, "plots/")
if (!dir.exists(plot_save_path)) dir.create(plot_save_path)
models_save_path <- paste0(save_path,"lmsz_models/")
if (!dir.exists(models_save_path)) dir.create(models_save_path)
```

## Load functions
```{r}
source(glue("{code_path}R_util_lmsz.R"))
source("/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/wp_taki_EK.R")

panel_save <- function(plot, filename, extension, folder, plot_width = 8, plot_height = 10){
  save_name <- paste0(folder, filename, ".", extension)
  ggsave(save_name, plot = plot, dpi = 300, width = plot_width, height = plot_height)
}
```

## Load variables
```{r}

full_vars <- read.csv(paste0(vars_path,"all_vars.csv")) %>% pull(x)
global_vars <- read.csv(paste0(vars_path,"global_vars.csv")) %>% pull(x)
vol_vars <- read.csv(paste0(vars_path,"vol_vars.csv")) %>% pull(x)
sa_vars <- read.csv(paste0(vars_path,"sa_vars.csv")) %>% pull(x)
th_vars <- read.csv(paste0(vars_path,"th_vars.csv")) %>% pull(x)

```


# Load dfs for visualization
```{r}
#out_fans_slip <- read.csv(glue("{csv_save_path}lmsz_centfans_ga_uncorrected.csv"))
df_ga <-  read.csv(glue("{csv_save_path}lmsz_adjusted_centiles_ga_2026-04-22.csv"))
models_ga <- read.csv(glue("{csv_save_path}lmsz_models_ga_2026-04-22.csv"))
#worm_plots <- lapply(list.files(glue("{plot_save_path}worm_plots/"), full.names = TRUE), readRDS)
```

# Define Color Scale for plots
```{r}
low_color = "#4878CF"
mid_color = "white"
high_color = "#D65F5F"
yes_color = "#D28C45"
```

# Prep models_ga for plotting
```{r}
#add label column to be able to plot with ggseg
models_ga$domain <- ifelse(grepl("global", models_ga$feature), "global",
                           ifelse(grepl("GrayVol", models_ga$feature), "vol",
                                  ifelse(grepl("SurfArea", models_ga$feature), "sa",
                                         ifelse(grepl("ThickAvg", models_ga$feature), "th",
                                                ifelse(grepl("aseg", models_ga$feature), "aseg", NA)))))
cols_edit <- grepl("aseg(?!_CC)", models_ga$feature, perl = TRUE)

models_ga$label <- models_ga$feature %>% 
  sub("aparc_GrayVol_", "", .) %>%
  sub("aparc_SurfArea_", "", .) %>%
  sub("aparc_ThickAvg_", "", .) %>%
  sub("aseg_", "", .) %>%
  sub("X", "x", .) %>%
  sub("rd_Ventricle", "rd_ventricle", .) %>%
  sub("th_Ventricle", "th_ventricle", .) %>%
  sub("Thalamus", "Thalamus-Proper", .)

models_ga$label[cols_edit] <- gsub("_", "-", models_ga$label[cols_edit])

models_ga <- models_ga %>%
  mutate(null_model_mu = grepl("~ 0", models_ga$m_model),
         null_model_sigma = grepl("~ 0", models_ga$s_model),
         int_model_mu = grepl("~ 1", models_ga$m_model),
         int_model_sigma = grepl("~ 1", models_ga$s_model),
         ga_mu_spline_flag = grepl("ns\\(ge", m_model),
         ga_sigma_spline_flag = grepl("ns\\(ge", s_model),
         age_mu_spline_flag = grepl("ns\\(adj", m_model),
         age_sigma_spline_flag = grepl("ns\\(adj", s_model))
```

# Merge with Atlas Frames
```{r}
# merge with atlas data
plot_data_dkt <- merge(models_ga, ggseg::dk, by = "label")
plot_data_aseg <- merge(models_ga, ggseg::aseg, by = "label")

# Regions not matched in atlases
models_ga_noplot <- models_ga[!(models_ga$label %in% plot_data_dkt$label) & !(models_ga$label %in% plot_data_aseg$label),
                                    c("label", "m_model", "s_model")]
knitr::kable(models_ga_noplot, 
      title = "Phenotypes not in GGSeg",
      align = "l")
```

# Plot terms on brain maps

## non-NULL model
```{r}
# in Mu
(vol_null_model_mu_dkt_plot <- ggplot(plot_data_dkt %>% filter(domain == "vol")) +
    geom_brain(atlas = dk, 
               position = position_brain(hemi ~ side),
               aes(fill = null_model_mu),
               color = "gray30",
               linewidth = "0.3") +
    scale_fill_manual(values = c("FALSE" = yes_color, "TRUE" = NA)) +
    theme_void()+
    labs(title = "Non-Null Model for Mu, Volume")) +
  theme(legend.position = "none")

(vol_null_model_mu_aseg_plot <- ggplot(plot_data_aseg) +
    geom_brain(atlas = aseg,
               aes(fill = null_model_mu),
               color = "gray30",
               linewidth = "0.3") +
    scale_fill_manual(values = c("FALSE" = yes_color, "TRUE" = NA)) +
    theme_void()+
    labs(title = "Non-Null Model for Mu")) +
  theme(legend.position = "none")

(sa_null_model_mu_dkt_plot <- ggplot(plot_data_dkt %>% filter(domain == "sa")) +
    geom_brain(atlas = dk, 
               position = position_brain(hemi ~ side),
               aes(fill = null_model_mu),
               color = "gray30",
               linewidth = "0.3") +
    scale_fill_manual(values = c("FALSE" = yes_color, "TRUE" = NA)) +
    theme_void()+
    labs(title = "Non-Null Model for Mu, Surface Area")) +
  theme(legend.position = "none")

(th_null_model_mu_dkt_plot <- ggplot(plot_data_dkt %>% filter(domain == "th")) +
    geom_brain(atlas = dk, 
               position = position_brain(hemi ~ side),
               aes(fill = null_model_mu),
               color = "gray30",
               linewidth = "0.3") +
    scale_fill_manual(values = c("FALSE" = yes_color, "TRUE" = NA)) +
    theme_void()+
    labs(title = "Non-Null Model for Mu, Thickness")) +
  theme(legend.position = "none")
```


## intercept model
```{r}
# in Mu
(vol_int_model_mu_dkt_plot <- ggplot(plot_data_dkt %>% filter(domain == "vol")) +
    geom_brain(atlas = dk, 
               position = position_brain(hemi ~ side),
               aes(fill = ifelse(int_model_mu, intercept_mu, 0)),
               color = "gray30") +
    scale_fill_gradient2(low = "#6BA292", mid = "gray", high = yes_color, limits = c(-0.04, 0.04),
                         name = "Intercept Coef") +
    theme_void()+
    labs(title = "Intercept Model for Mu, Volume") +
  theme(legend.position = "bottom"))

(vol_int_model_mu_aseg_plot <- ggplot(plot_data_aseg) +
    geom_brain(atlas = aseg,
               aes(fill = ifelse(int_model_mu, intercept_mu, 0)),
               color = "gray30") +
    scale_fill_gradient2(low = "#6BA292", mid = "gray", high = yes_color, limits = c(-0.04, 0.04),
                         name = "Intercept Coef") +
    theme_void()+
    labs(title = "Intercept Model for Mu") +
  theme(legend.position = "bottom"))

(sa_int_model_mu_dkt_plot <- ggplot(plot_data_dkt %>% filter(domain == "sa")) +
    geom_brain(atlas = dk, 
               position = position_brain(hemi ~ side),
               aes(fill = ifelse(int_model_mu, intercept_mu, 0)),
               color = "gray30") +
    scale_fill_gradient2(low = "#6BA292", mid = "gray", high = yes_color, limits = c(-0.04, 0.04),
                         name = "Intercept Coef") +
    theme_void()+
    labs(title = "Intercept Model for Mu, Surface Area") +
  theme(legend.position = "bottom"))

(th_int_model_mu_dkt_plot <- ggplot(plot_data_dkt %>% filter(domain == "th")) +
    geom_brain(atlas = dk, 
               position = position_brain(hemi ~ side),
               aes(fill = ifelse(int_model_mu, intercept_mu, 0)),
               color = "gray30") +
    scale_fill_gradient2(low = "#6BA292", mid = "gray", high = yes_color, limits = c(-0.04, 0.04),
                         name = "Intercept Coef") +
    theme_void()+
    labs(title = "Intercept Model for Mu, Thickness") +
  theme(legend.position = "bottom"))
```

## GA terms in Mu
```{r}
# in Mu
(vol_ga_term_mu_dkt_plot <- ggplot(plot_data_dkt %>% filter(domain == "vol")) +
    geom_brain(atlas = dk, 
               position = position_brain(hemi ~ side),
               aes(fill = ga_term_mu,
                   color = ga_mu_spline_flag)) +
    scale_fill_manual(values = c("TRUE" = yes_color, "FALSE" = NA)) +
   scale_color_manual(values = c("TRUE" = "black", "FALSE" = "gray"),
                      breaks = c(TRUE),
                      labels = "natural spline for GA") +
    theme_void() +
    labs(title = "GA Term for Mu, Volume") +
  guides(fill = "none") +
  theme(legend.position = "bottom",
        legend.title = element_blank()))

(vol_ga_term_mu_aseg_plot <- ggplot(plot_data_aseg) +
    geom_brain(atlas = aseg, 
               aes(fill = ga_term_mu,
                   color = ga_mu_spline_flag)) +
    scale_fill_manual(values = c("TRUE" = yes_color, "FALSE" = NA)) +
   scale_color_manual(values = c("TRUE" = "black", "FALSE" = "gray"),
                      breaks = c(TRUE),
                      labels = "natural spline for GA") +
    theme_void() +
    labs(title = "GA Term for Mu") +
  guides(fill = "none") +
  theme(legend.position = "bottom",
        legend.title = element_blank()))

(sa_ga_term_mu_dkt_plot <- ggplot(plot_data_dkt %>% filter(domain == "sa")) +
    geom_brain(atlas = dk, 
               position = position_brain(hemi ~ side),
               aes(fill = ga_term_mu,
                   color = ga_mu_spline_flag)) +
    scale_fill_manual(values = c("TRUE" = yes_color, "FALSE" = NA)) +
   scale_color_manual(values = c("TRUE" = "black", "FALSE" = "gray"),
                      breaks = c(TRUE),
                      labels = "natural spline for GA") +
    theme_void() +
    labs(title = "GA Term for Mu, Surface Area") +
  guides(fill = "none") +
  theme(legend.position = "bottom",
        legend.title = element_blank()))

(th_ga_term_mu_dkt_plot <- ggplot(plot_data_dkt %>% filter(domain == "th")) +
    geom_brain(atlas = dk, 
               position = position_brain(hemi ~ side),
               aes(fill = ga_term_mu,
                   color = ga_mu_spline_flag)) +
    scale_fill_manual(values = c("TRUE" = yes_color, "FALSE" = NA)) +
   scale_color_manual(values = c("TRUE" = "black", "FALSE" = "gray"),
                      breaks = c(TRUE),
                      labels = "natural spline for GA") +
    theme_void() +
    labs(title = "GA Term for Mu, Thickness") +
  guides(fill = "none") +
  theme(legend.position = "bottom",
        legend.title = element_blank()))
```

## AGE terms in Mu
```{r}
# in Mu
(vol_age_term_mu_dkt_plot <- ggplot(plot_data_dkt %>% filter(domain == "vol")) +
    geom_brain(atlas = dk, 
               position = position_brain(hemi ~ side),
               aes(fill = age_term_mu,
                   color = age_mu_spline_flag)) +
    scale_fill_manual(values = c("TRUE" = yes_color, "FALSE" = NA)) +
   scale_color_manual(values = c("TRUE" = "black", "FALSE" = "gray"),
                      breaks = c(TRUE),
                      labels = "natural spline for age") +
    theme_void() +
    labs(title = "age Term for Mu, Volume") +
  guides(fill = "none") +
  theme(legend.position = "bottom",
        legend.title = element_blank()))

(vol_age_term_mu_aseg_plot <- ggplot(plot_data_aseg) +
    geom_brain(atlas = aseg, 
               aes(fill = age_term_mu,
                   color = age_mu_spline_flag)) +
    scale_fill_manual(values = c("TRUE" = yes_color, "FALSE" = NA)) +
   scale_color_manual(values = c("TRUE" = "black", "FALSE" = "gray"),
                      breaks = c(TRUE),
                      labels = "natural spline for age") +
    theme_void() +
    labs(title = "age Term for Mu") +
  guides(fill = "none") +
  theme(legend.position = "bottom",
        legend.title = element_blank()))

(sa_age_term_mu_dkt_plot <- ggplot(plot_data_dkt %>% filter(domain == "sa")) +
    geom_brain(atlas = dk, 
               position = position_brain(hemi ~ side),
               aes(fill = age_term_mu,
                   color = age_mu_spline_flag)) +
    scale_fill_manual(values = c("TRUE" = yes_color, "FALSE" = NA)) +
   scale_color_manual(values = c("TRUE" = "black", "FALSE" = "gray"),
                      breaks = c(TRUE),
                      labels = "natural spline for age") +
    theme_void() +
    labs(title = "age Term for Mu, Surface Area") +
  guides(fill = "none") +
  theme(legend.position = "bottom",
        legend.title = element_blank()))

(th_age_term_mu_dkt_plot <- ggplot(plot_data_dkt %>% filter(domain == "th")) +
    geom_brain(atlas = dk, 
               position = position_brain(hemi ~ side),
               aes(fill = age_term_mu,
                   color = age_mu_spline_flag)) +
    scale_fill_manual(values = c("TRUE" = yes_color, "FALSE" = NA)) +
   scale_color_manual(values = c("TRUE" = "black", "FALSE" = "gray"),
                      breaks = c(TRUE),
                      labels = "natural spline for age") +
    theme_void() +
    labs(title = "age Term for Mu, Thickness") +
  guides(fill = "none") +
  theme(legend.position = "bottom",
        legend.title = element_blank()))
```

## SEX terms in Mu
```{r}
# in Mu
(vol_sex_term_mu_dkt_plot <- ggplot(plot_data_dkt %>% filter(domain == "vol")) +
    geom_brain(atlas = dk, 
               position = position_brain(hemi ~ side),
               aes(fill = sex_term_mu)) +
    scale_fill_manual(values = c("TRUE" = yes_color, "FALSE" = NA)) +
    theme_void() +
    labs(title = "sex Term for Mu, Volume") +
  guides(fill = "none") +
  theme(legend.position = "bottom",
        legend.title = element_blank()))

(vol_sex_term_mu_aseg_plot <- ggplot(plot_data_aseg) +
    geom_brain(atlas = aseg, 
               aes(fill = sex_term_mu)) +
    scale_fill_manual(values = c("TRUE" = yes_color, "FALSE" = NA)) +
    theme_void() +
    labs(title = "sex Term for Mu") +
  guides(fill = "none") +
  theme(legend.position = "bottom",
        legend.title = element_blank()))

(sa_sex_term_mu_dkt_plot <- ggplot(plot_data_dkt %>% filter(domain == "sa")) +
    geom_brain(atlas = dk, 
               position = position_brain(hemi ~ side),
               aes(fill = sex_term_mu)) +
    scale_fill_manual(values = c("TRUE" = yes_color, "FALSE" = NA)) +
    theme_void() +
    labs(title = "sex Term for Mu, Surface Area") +
  guides(fill = "none") +
  theme(legend.position = "bottom",
        legend.title = element_blank()))

(th_sex_term_mu_dkt_plot <- ggplot(plot_data_dkt %>% filter(domain == "th")) +
    geom_brain(atlas = dk, 
               position = position_brain(hemi ~ side),
               aes(fill = sex_term_mu)) +
    scale_fill_manual(values = c("TRUE" = yes_color, "FALSE" = NA)) +
    theme_void() +
    labs(title = "sex Term for Mu, Thickness") +
  guides(fill = "none") +
  theme(legend.position = "bottom",
        legend.title = element_blank()))
```

## GA:Age Interaction in Mu
```{r}
# in Mu
(vol_ga.age_mu_interaction_dkt_plot <- ggplot(plot_data_dkt %>% filter(domain == "vol")) +
    geom_brain(atlas = dk, 
               position = position_brain(hemi ~ side),
               aes(fill = ga.age_mu_interaction)) +
    scale_fill_manual(values = c("TRUE" = yes_color, "FALSE" = NA)) +
    theme_void() +
    labs(title = "GA x Age Interaction for Mu, Volume") +
  guides(fill = "none") +
  theme(legend.position = "bottom",
        legend.title = element_blank()))

(vol_ga.age_mu_interaction_aseg_plot <- ggplot(plot_data_aseg) +
    geom_brain(atlas = aseg, 
               aes(fill = ga.age_mu_interaction)) +
    scale_fill_manual(values = c("TRUE" = yes_color, "FALSE" = NA)) +
    theme_void() +
    labs(title = "GA x Age Interaction for Mu") +
  guides(fill = "none") +
  theme(legend.position = "bottom",
        legend.title = element_blank()))

(sa_ga.age_mu_interaction_dkt_plot <- ggplot(plot_data_dkt %>% filter(domain == "sa")) +
    geom_brain(atlas = dk, 
               position = position_brain(hemi ~ side),
               aes(fill = ga.age_mu_interaction)) +
    scale_fill_manual(values = c("TRUE" = yes_color, "FALSE" = NA)) +
    theme_void() +
    labs(title = "GA x Age Interaction for Mu, Surface Area") +
  guides(fill = "none") +
  theme(legend.position = "bottom",
        legend.title = element_blank()))

(th_ga.age_mu_interaction_dkt_plot <- ggplot(plot_data_dkt %>% filter(domain == "th")) +
    geom_brain(atlas = dk, 
               position = position_brain(hemi ~ side),
               aes(fill = ga.age_mu_interaction)) +
    scale_fill_manual(values = c("TRUE" = yes_color, "FALSE" = NA)) +
    theme_void() +
    labs(title = "GA x Age Interaction for Mu, Thickness") +
  guides(fill = "none") +
  theme(legend.position = "bottom",
        legend.title = element_blank()))
```

## GA:Sex Interaction in Mu
```{r}
# in Mu
(vol_ga.sex_mu_interaction_dkt_plot <- ggplot(plot_data_dkt %>% filter(domain == "vol")) +
    geom_brain(atlas = dk, 
               position = position_brain(hemi ~ side),
               aes(fill = ga.sex_mu_interaction)) +
    scale_fill_manual(values = c("TRUE" = yes_color, "FALSE" = NA)) +
    theme_void() +
    labs(title = "GA x Sex Interaction for Mu, Volume") +
  guides(fill = "none") +
  theme(legend.position = "bottom",
        legend.title = element_blank()))

(vol_ga.sex_mu_interaction_aseg_plot <- ggplot(plot_data_aseg) +
    geom_brain(atlas = aseg, 
               aes(fill = ga.sex_mu_interaction)) +
    scale_fill_manual(values = c("TRUE" = yes_color, "FALSE" = NA)) +
    theme_void() +
    labs(title = "GA x Sex Interaction for Mu") +
  guides(fill = "none") +
  theme(legend.position = "bottom",
        legend.title = element_blank()))

(sa_ga.sex_mu_interaction_dkt_plot <- ggplot(plot_data_dkt %>% filter(domain == "sa")) +
    geom_brain(atlas = dk, 
               position = position_brain(hemi ~ side),
               aes(fill = ga.sex_mu_interaction)) +
    scale_fill_manual(values = c("TRUE" = yes_color, "FALSE" = NA)) +
    theme_void() +
    labs(title = "GA x Sex Interaction for Mu, Surface Area") +
  guides(fill = "none") +
  theme(legend.position = "bottom",
        legend.title = element_blank()))

(th_ga.sex_mu_interaction_dkt_plot <- ggplot(plot_data_dkt %>% filter(domain == "th")) +
    geom_brain(atlas = dk, 
               position = position_brain(hemi ~ side),
               aes(fill = ga.sex_mu_interaction)) +
    scale_fill_manual(values = c("TRUE" = yes_color, "FALSE" = NA)) +
    theme_void() +
    labs(title = "GA x Sex Interaction for Mu, Thickness") +
  guides(fill = "none") +
  theme(legend.position = "bottom",
        legend.title = element_blank()))
```


## Age:Sex Interaction in Mu
```{r}
# in Mu
(vol_age.sex_mu_interaction_dkt_plot <- ggplot(plot_data_dkt %>% filter(domain == "vol")) +
    geom_brain(atlas = dk, 
               position = position_brain(hemi ~ side),
               aes(fill = age.sex_mu_interaction)) +
    scale_fill_manual(values = c("TRUE" = yes_color, "FALSE" = NA)) +
    theme_void() +
    labs(title = "Age x Sex Interaction for Mu, Volume") +
  guides(fill = "none") +
  theme(legend.position = "bottom",
        legend.title = element_blank()))

(vol_age.sex_mu_interaction_aseg_plot <- ggplot(plot_data_aseg) +
    geom_brain(atlas = aseg, 
               aes(fill = age.sex_mu_interaction)) +
    scale_fill_manual(values = c("TRUE" = yes_color, "FALSE" = NA)) +
    theme_void() +
    labs(title = "Age x Sex Interaction for Mu") +
  guides(fill = "none") +
  theme(legend.position = "bottom",
        legend.title = element_blank()))

(sa_age.sex_mu_interaction_dkt_plot <- ggplot(plot_data_dkt %>% filter(domain == "sa")) +
    geom_brain(atlas = dk, 
               position = position_brain(hemi ~ side),
               aes(fill = age.sex_mu_interaction)) +
    scale_fill_manual(values = c("TRUE" = yes_color, "FALSE" = NA)) +
    theme_void() +
    labs(title = "Age x Sex Interaction for Mu, Surface Area") +
  guides(fill = "none") +
  theme(legend.position = "bottom",
        legend.title = element_blank()))

(th_age.sex_mu_interaction_dkt_plot <- ggplot(plot_data_dkt %>% filter(domain == "th")) +
    geom_brain(atlas = dk, 
               position = position_brain(hemi ~ side),
               aes(fill = age.sex_mu_interaction)) +
    scale_fill_manual(values = c("TRUE" = yes_color, "FALSE" = NA)) +
    theme_void() +
    labs(title = "Age x Sex Interaction for Mu, Thickness") +
  guides(fill = "none") +
  theme(legend.position = "bottom",
        legend.title = element_blank()))
```


# Trajectory Plots
Each pheno is now saved as a separate csv, so need to re-think how to plot these.
Perhaps choose very specific ones to plot in addition to the global ones.
Maybe the regions with the strongest GA effects ???

And then plot lmsz centiles vs OG centiles for different weeks
And then plot lmsz centiles vs OG centiles against birthweight or birthweight percentile (when it is re-calculated)
## Prep

### Fix scale
```{r}
## Fix scale of out_fans, scale everything back by 1000
#Scale all phenotypes by 1000
out_fans_slip <- out_fans_slip %>% 
  mutate(SLIP_input_fans = SLIP_input_fans*1000,
         LMSz_fans = LMSz_fans*1000) %>%
  mutate(sex_ga = paste0(sex, gestational_age))
```

### Define x axis
```{r}
## Define x axis marks
x_breaks <- log(c(280, 365.25*1+280, 365.25*2+280, 
                  365.25*6+280, 365.25*21+280))
x_labs <- c("0y","1y","2y","6y","21y")
```

### Define plotting functions
```{r}
## plotting loop function
separate_plot_loop <- function(pheno_list, df_plot, out_fans_df, x_breaks, x_labs){
  slip_plots <- list()
  lmsz_plots <- list()
  for (pheno_of_int in pheno_list) {
    df_plot$pheno <- df_plot[[glue("{pheno_of_int}.raw")]]
    df_plot$pheno[is.na(df_plot[[glue("{pheno_of_int}.lmsz")]])] <- NA
    if(grepl("global", pheno_of_int)){
      ylab_text = ""
      title_text = sub("global_", "", pheno_of_int)
    } else if(grepl("GrayVol", pheno_of_int))  {
      ylab_text = "Volume"
      title_text = sub("aparc_GrayVol_", "", pheno_of_int)
    } else if(grepl("SurfArea", pheno_of_int))  {
      ylab_text = "Surface Area"
      title_text = sub("aparc_SurfArea_", "", pheno_of_int)
    } else if(grepl("ThickAvg", pheno_of_int))  {
      ylab_text = "Thickness"
      title_text = sub("aparc_ThickAvg_", "", pheno_of_int)
    } else if(grepl("aseg", pheno_of_int))  {
      ylab_text = "Aseg/Subcortical"
      title_text = sub("aseg_", "", pheno_of_int)
    } else{
      ylab_text = pheno_of_int
      title_text = pheno_of_int
    }
    out_fans_plot <- out_fans_df %>% 
      filter(pheno == pheno_of_int,
             !(centile_cat %in% c("cent_0.25", "cent_0.75")))
    
    # CT data is questionable at 0y. Limit to 2 years
    if (grepl("Thick",pheno_of_int)) {
      df_plot <- df_plot %>% filter(adjusted_age_in_days >= 2*365.25+280)
      
      out_fans_plot <- out_fans_plot %>% filter(adjusted_age_days_log >= log(2*365.25+280))
    }
    
    df_plot$sex_ga <- paste0(df_plot$sex, df_plot$gestational_age)
    
    slip_plots[[pheno_of_int]] <- ggplot(out_fans_plot) +
      geom_point(data = df_plot, aes(y = pheno,
                                     x = adjusted_age_days_log, fill=sex), 
                 alpha=0.5, shape=21, size = 0.5) +
      geom_line(aes(x = adjusted_age_days_log, y = SLIP_input_fans,
                    group = interaction(centile_cat, sex),
                    color = sex,
                    linewidth = line_size,
                    linetype = line_type)) +
      scale_linewidth_identity() +
      scale_linetype_identity() +
      scale_colour_manual(values = c("firebrick","royalblue")) +
      scale_x_continuous(breaks = x_breaks, labels = x_labs) +
      ggtitle(title_text) + theme_classic() +
      xlab("") + ylab(ylab_text) +
      theme(plot.title = element_text(size = 12),
            axis.title = element_text(size = 10),
            axis.text = element_text(size = 8))
    
    lmsz_plots[[pheno_of_int]] <- ggplot(out_fans_plot %>%
                                           filter(gestational_age %in% c(24, 32, 36, 40, 42))) +
      geom_point(data = df_plot %>% 
                   filter(gestational_age %in% c(24, 32, 36, 40, 42)), aes(y = pheno,
                                                                           x = adjusted_age_days_log, color = sex_ga), 
                 alpha=0.5, shape=21, size = 0.5) +
      geom_line(aes(x = adjusted_age_days_log, y = LMSz_fans,
                    group = interaction(centile_cat, sex_ga),
                    color = sex_ga,
                    linewidth = line_size,
                    linetype = line_type)) +
      scale_linewidth_identity() +
      scale_linetype_identity() +
      scale_colour_manual(values = c(
        "#E3AEB7", # lightest pink
        "#D07C8F", # lighter pink
        "#B85C74", # medium light rose
        "#A64A63", # medium dark rose
        "#8A2F4D", # deeper red
        "#7FA6C9", # lightest blue
        "#4F85B5", # lighter blue
        "#2F5F8F", # medium light blue
        "#4C78A8",  # medium dark blue
        "#3F51B5" )) +  # indigo blue
      scale_x_continuous(breaks = x_breaks, labels = x_labs) +
      ggtitle(title_text) + theme_classic() +
      xlab("") + ylab(ylab_text) +
      theme(plot.title = element_text(size = 12),
            axis.title = element_text(size = 10),
            axis.text = element_text(size = 8))
    
    # Equalize Axes
    p_y_min <- min(layer_scales(slip_plots[[pheno_of_int]])$y$range$range[1],
                   layer_scales(lmsz_plots[[pheno_of_int]])$y$range$range[1])
    p_y_max <- max(layer_scales(slip_plots[[pheno_of_int]])$y$range$range[2],
                   layer_scales(lmsz_plots[[pheno_of_int]])$y$range$range[2])
    
    slip_plots[[pheno_of_int]] <- slip_plots[[pheno_of_int]] + ylim(p_y_min,p_y_max)
    lmsz_plots[[pheno_of_int]] <- lmsz_plots[[pheno_of_int]] + ylim(p_y_min,p_y_max)
  }
  all_plots_list <- list()
  all_plots_list$slip_plots <- slip_plots
  all_plots_list$lmsz_plots <- lmsz_plots
  return(all_plots_list)
}

overlaid_plot_loop <- function(pheno_list, df_plot, out_fans_df, x_breaks, x_labs){
  overlap_plots_male <- list()
  overlap_plots_female <- list()
  overlap_plots_avg <- list()
  
  pheno_of_int <- "global_SurfaceArea"
  
  for (pheno_of_int in pheno_list) {
    df_plot$pheno <- df_plot[[glue("{pheno_of_int}.raw")]]
    df_plot$pheno[is.na(df_plot[[glue("{pheno_of_int}.lmsz")]])] <- NA
    if(grepl("global", pheno_of_int)){
      ylab_text = ""
      title_text = sub("global_", "", pheno_of_int)
    } else if(grepl("GrayVol", pheno_of_int))  {
      ylab_text = "Volume"
      title_text = sub("aparc_GrayVol_", "", pheno_of_int)
    } else if(grepl("SurfArea", pheno_of_int))  {
      ylab_text = "Surface Area"
      title_text = sub("aparc_SurfArea_", "", pheno_of_int)
    } else if(grepl("ThickAvg", pheno_of_int))  {
      ylab_text = "Thickness"
      title_text = sub("aparc_ThickAvg_", "", pheno_of_int)
    } else if(grepl("aseg", pheno_of_int))  {
      ylab_text = "Aseg/Subcortical"
      title_text = sub("aseg_", "", pheno_of_int)
    } else{
      ylab_text = pheno_of_int
      title_text = pheno_of_int
    }
    
    out_fans_plot <- out_fans_df %>% 
      filter(pheno == pheno_of_int,
             !(centile_cat %in% c("cent_0.25", "cent_0.75")))
    
    if (grepl("Thick",pheno_of_int)) {
      df_plot <- df_plot %>% filter(adjusted_age_in_days >= 2*365.25+280)
      out_fans_plot <- out_fans_plot %>% filter(adjusted_age_days_log >= log(2*365.25+280))
    }
    
    
    overlap_plots_male[[pheno_of_int]] <- ggplot(out_fans_plot %>% filter(sex == "Male") %>% 
                                                   filter(gestational_age %in% c(24, 32, 36, 40, 42))) +
      geom_point(data = df_plot %>% filter(sex == "Male"), aes(y = pheno,
                                                               x = adjusted_age_days_log), 
                 alpha=1, size = 0.5, color = "gray") +
      geom_line(aes(x = adjusted_age_days_log, y = SLIP_input_fans,
                    group = centile_cat,
                    color = "SLIP",
                    linewidth = line_size,
                    linetype = line_type)) +
      geom_line(aes(x = adjusted_age_days_log, y = LMSz_fans,
                    group = interaction(centile_cat, gestational_age),
                    color = as.factor(gestational_age),
                    linewidth = line_size,
                    linetype = line_type)) +
      scale_linewidth_identity() +
      scale_linetype_identity() +
      scale_x_continuous(breaks = x_breaks, labels = x_labs) +
      ggtitle(paste0(title_text)) + theme_classic() +
      xlab("") + ylab(ylab_text) +
      scale_colour_manual(values = c(
        "#7FA6C9", # lightest blue
        "#4F85B5", # lighter blue
        "#2F5F8F", # medium light blue
        "#4C78A8",  # medium dark blue
        "#3F51B5",  # indigo blue
        "#2F2F2F")) + # black-ish gray
      theme(plot.title = element_text(size = 12),
            axis.title = element_text(size = 10),
            axis.text = element_text(size = 8))
    
    overlap_plots_female[[pheno_of_int]] <- ggplot(out_fans_plot %>% filter(sex == "Female") %>% 
                                                     filter(gestational_age %in% c(24, 32, 36, 40, 42))) +
      geom_point(data = df_plot %>% filter(sex == "Female"), aes(y = pheno,
                                                                 x = adjusted_age_days_log), 
                 alpha=1, size = 0.5, color = "gray") +
      geom_line(aes(x = adjusted_age_days_log, y = SLIP_input_fans,
                    group = centile_cat,
                    color = "SLIP",
                    linewidth = line_size,
                    linetype = line_type)) +
      geom_line(aes(x = adjusted_age_days_log, y = LMSz_fans,
                    group = interaction(centile_cat, gestational_age),
                    color = as.factor(gestational_age),
                    linewidth = line_size,
                    linetype = line_type)) +
      scale_linewidth_identity() +
      scale_linetype_identity() +
      scale_x_continuous(breaks = x_breaks, labels = x_labs) +
      ggtitle(paste0(title_text)) + theme_classic() +
      xlab("") + ylab(ylab_text) +
      scale_colour_manual(values = c(
        "#E3AEB7", # lightest pink
        "#D07C8F", # lighter pink
        "#B85C74", # medium light rose
        "#A64A63", # medium dark rose
        "#8A2F4D",  # deeper red
        "#2F2F2F")) + # black-ish gray
      theme(plot.title = element_text(size = 12),
            axis.title = element_text(size = 10),
            axis.text = element_text(size = 8))
    
    
    # average over 2 sexes
    out_fans_plot_avg <-   out_fans_plot %>% 
      filter(grepl(str_replace_all(pheno_of_int,"SA.","SA.*"),pheno),
             !(centile_cat %in% c("cent_0.25", "cent_0.75"))) %>%
      mutate(SLIP_input_fans = SLIP_input_fans,
             LMSz_fans = LMSz_fans) %>%
      group_by(model, adjusted_age_days_log, gestational_age,
               centile_cat, line_size, line_type) %>%
      summarize(across(contains("fans"), ~mean(.x, na.rm = T)))
    
    overlap_plots_avg[[pheno_of_int]] <- ggplot(out_fans_plot_avg %>% 
                                                  filter(gestational_age %in% c(24, 32, 36, 40, 42))) +
      geom_point(data = df_plot, aes(y = pheno,
                                     x = adjusted_age_days_log), 
                 alpha=1, size = 0.5, color = "gray") +
      geom_line(aes(x = adjusted_age_days_log, y = SLIP_input_fans,
                    group = centile_cat,
                    color = "SLIP",
                    linewidth = line_size,
                    linetype = line_type)) +
      geom_line(aes(x = adjusted_age_days_log, y = LMSz_fans,
                    group = interaction(centile_cat, gestational_age),
                    color = as.factor(gestational_age),
                    linewidth = line_size,
                    linetype = line_type)) +
      scale_linewidth_identity() +
      scale_linetype_identity() +
      scale_x_continuous(breaks = x_breaks, labels = x_labs) +
      ggtitle(paste0(title_text)) + theme_classic() +
      xlab("") + ylab(ylab_text) +
      scale_colour_manual(values = c(
        "#AFC7B5", # lightest green
        "#7FA68F", # lighter green
        "#5F8F73", # medium light green
        "#4C7A5F", # medium dark green
        "#2F5A45",  # deeper green
        "#2F2F2F")) + # black-ish gray
      theme(plot.title = element_text(size = 12),
            axis.title = element_text(size = 10),
            axis.text = element_text(size = 8))
    
  }
  all_plots_list <- list()
  all_plots_list$overlap_male <- overlap_plots_male
  all_plots_list$overlap_female<- overlap_plots_female
  all_plots_list$overlap_avg <- overlap_plots_avg
  return(all_plots_list)
}
```


## Plots

```{r}
## Plot global trajectories
global_pheno_to_plot <- c("global_eTIV","global_VentricleChoroidVol","global_CortexVol",
                          "global_SubCortGrayVol","global_SurfaceArea","global_CerebralWhiteMatterVol",
                          "global_MeanThickness")

#Overlaid
global_plots_overlaid <- list()
global_plots_overlaid <- overlaid_plot_loop(global_pheno_to_plot, 
                                            df_ga, 
                                            out_fans_slip, 
                                            x_breaks, 
                                            x_labs)

#overlaid plots by sex
p_male_title <- ggdraw() + draw_label("Males", angle = 90,
                                      x = 0.5, hjust = 0.5, fontface = "bold")
p_male <- ggarrange(plots =  global_plots_overlaid$overlap_male, 
                    nrow = 2, legend = "none", common.legend = T)
p_female_title <- ggdraw() + draw_label("Females", angle = 90,
                                        x = 0.5, hjust = 0.5, fontface = "bold")
p_female <- ggarrange(plots =  global_plots_overlaid$overlap_female, 
                      nrow = 2, legend = "none", common.legend = T)

p_fits_global_overlaid_sex <- (p_male_title + p_male +
                                 p_female_title + p_female) + 
  plot_layout(ncol = 2, 
              nrow = 2, 
              widths = c(0.07, 1)) +
  plot_annotation(title = "Global, by GA")

p_fits_global_overlaid_sex

#overlaid plots averaged over sex
p_traj_sexavg <- ggarrange(plots =  global_plots_overlaid$overlap_avg, 
                           nrow = 2, legend = "bottom", common.legend = T)

p_traj_collapsed_title <- ggdraw() + draw_label("By GA, Sex Avg", angle = 90, x = 0.5, hjust = 0.5, fontface = "bold")


p_fits_global_overlaid_avg <- (p_traj_collapsed_title  + p_traj_sexavg) + 
  plot_layout(ncol = 2, 
              nrow = 1, 
              widths = c(0.07, 1)) +
  plot_annotation(title = "Global")


#Separate
global_plots <- list()
global_plots <- separate_plot_loop(global_pheno_to_plot, 
                                   df_ga, 
                                   out_fans_slip, 
                                   x_breaks, 
                                   x_labs)


p_slip_title <- ggdraw() + draw_label("Original (SLIP)", angle = 90,
                                      x = 0.5, hjust = 0.5, fontface = "bold")

p_slip <- ggarrange(plots =  global_plots$slip_plots, nrow = 2, legend = "none", common.legend = T)

p_lmsz_title <- ggdraw() + draw_label("GA LMSz", angle = 90,
                                      x = 0.5, hjust = 0.5, fontface = "bold")
p_lmsz <- ggarrange(plots =  global_plots$lmsz_plots, nrow = 2, legend = "bottom", common.legend = T)

p_fits_global_separate <- (p_slip_title + p_slip +
                             p_lmsz_title + p_lmsz) + 
  plot_layout(ncol = 2, 
              nrow = 2, 
              widths = c(0.07, 1)) +
  plot_annotation(title = "Global")

panel_save(p_fits_global_separate, "global_traj_separate", "pdf", plot_save_path, plot_width = 24, plot_height = 16)
panel_save(p_fits_global_overlaid_sex, "global_traj_overlaid_sex", "pdf", plot_save_path, plot_width = 24, plot_height = 16)
panel_save(p_fits_global_overlaid_avg, "global_traj_overlaid_avg", "pdf", plot_save_path, plot_width = 24, plot_height = 16)
```