---
title: "slip_premie_lmsz"
author: "Eren Kafadar"
date: "2025-11-18"
output: html_document
editor_options: 
  chunk_output_type: console
params:
  output_folder: "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/output/lmsz_median_Th-MPR_25.11/"
  centiles_data_path: "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/output/median-Th-MPR-distinct-allgrades-25.11/data_clean/median_Th-MPR_2026-02-04.csv"
  raw_data_path: "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11/raw_csv/median_Th-MPR_2026-02-04.csv"
  models_path: "/mnt/isilon/bgdlab_processing/braincharts/SLIP/2025_03/code/gamlss/RDS/"
---
# Define params if they do not exist
```{r, echo = FALSE}
if (!exists("params")) {
  params <- list(
  output_folder = "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/output/lmsz_median_Th-MPR_25.11/",
  centiles_data_path = "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/output/median-Th-MPR-distinct-allgrades-25.11/data_clean/median_Th-MPR_2026-02-04.csv", 
  raw_data_path = "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11/raw_csv/median_Th-MPR_2026-02-04.csv",
  models_path = "/mnt/isilon/bgdlab_processing/braincharts/SLIP/2025_03/code/gamlss/RDS/"
  )
}
print(params)
``` 

# controls
```{r, echo = TRUE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)

set.seed(42)
#set command
save_tables <- TRUE
plot <- TRUE
save_figs <- TRUE

output_folder <- params$output_folder
centiles_data_path <- params$centiles_data_path
raw_data_path <- params$raw_data_path
models_path <- params$models_path
```

# Load Packages
```{r, echo = FALSE}
# Load libraries
require(dplyr)
require(magrittr)
require(readr)
require(stringr)
require(tidyr)
require(gamlssTools) # Margaret's package
require(gamlss.add)
require(tibble)
require(glue)
# For figures/plotting
require(ggplot2)
# For figures/plotting
require(gridExtra)
require(ggsegDKT)
require(ggseg)
# require(ggpubr)
require(ggcorrplot)
require(cowplot)
require(egg)
require(interactions)
require(growthstandards)
require(gtsummary)
require(svglite)
```

# Define Paths
```{r, echo = TRUE}
# set folder paths
code_path <- "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/"
vars_path <-  "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/output/vars/"

# output folder
save_path <- params$output_folder
if (!dir.exists(save_path)) dir.create(save_path)

##  output subfolder paths
csv_save_path <-  paste0(save_path, "results_csv/")
if (!dir.exists(csv_save_path)) dir.create(csv_save_path)
plot_save_path <- paste0(save_path, "plots/")
if (!dir.exists(plot_save_path)) dir.create(plot_save_path)
```

# Load custom functions
```{r, echo = FALSE}
panel_save <- function(plot, filename, extension, folder){
  save_name <- paste0(folder, filename, ".", extension)
  ggsave(save_name, plot = plot, dpi = 300)
}

source(glue("{code_path}R_util_lmsz.R"))
```

# Load Phenotype Variable Names
```{r, echo = FALSE}
full_vars <- read.csv(paste0(vars_path,"all_vars.csv")) %>% pull(x)
global_vars <- read.csv(paste0(vars_path,"global_vars.csv")) %>% pull(x)
vol_vars <- read.csv(paste0(vars_path,"vol_vars.csv")) %>% pull(x)
sa_vars <- read.csv(paste0(vars_path,"sa_vars.csv")) %>% pull(x)
th_vars <- read.csv(paste0(vars_path,"th_vars.csv")) %>% pull(x)

## remove aseg_CC_Central and "aseg_CC_Mid_Posterior" from the features because there is one NaN value in the centiles causing issues. In the future will edit the functions to deal with this and/or figure out why there is this NaN value.
full_vars <- full_vars[!(full_vars %in% c("aseg_CC_Central", "aseg_CC_Mid_Posterior"))]
```

# Load Clean Data
```{r, echo = FALSE}
df <- read.csv(centiles_data_path) %>% select(-"X") #%>%
  #rename_with(~paste0(.x, ".cent"), all_of(full_vars))
median_th_mpr_flag <- any(grepl(".median", names(df)))
```

# Load Raw Data and merge
```{r, echo = FALSE}
df_raw <- read.csv(raw_data_path) %>% select(any_of(c(full_vars, "participant_id", "session_id", "sex", "age_days",
                                                      "site", "avg_grade", "age_at_scan"))) %>%
  #mutate(across(all_of(full_vars), ~c(scale(.)), .names = "{.col}.raw_normalised")) %>%
  rename_with(~paste0(.x, ".raw"), all_of(full_vars))

shared_names <- intersect(names(df_raw), names(df))

df <- merge(df, df_raw, by = shared_names, all.x = FALSE, all.y = FALSE)
```

# Remove Age > 21
```{r, echo = TRUE}
hist(df$adjusted_age_in_years, breaks = 100)

df <- df %>% filter(adjusted_age_in_years <= 21)

ggplot(df, aes(x = adjusted_age_in_years)) +
  geom_histogram(binwidth = 0.5, fill = "white", color = "darkgray") +
  scale_x_continuous(breaks = seq(min(df$adjusted_age_in_years, na.rm = TRUE),
                                  max(df$adjusted_age_in_years, na.rm = TRUE),
                                  by = 0.5),
    labels = function(x) signif(x, 2)) +
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```
# Z-score Centiles, remove Inf Cells
## Set 0/1 centiles to min/max value to avoid Inf/-Inf
numbers from gamlssTools pkg
```{r}
if(median_th_mpr_flag){
th_vars_median <- paste0(th_vars, ".median")

complete_centile_vars <- c(full_vars, th_vars_median, "bweight_percentile", "weight_at_enc_cent", "height_at_enc_cent")
} else{
  complete_centile_vars <- c(full_vars, "bweight_percentile", "weight_at_enc_cent", "height_at_enc_cent")
}

for (x in complete_centile_vars){
  df[df[,x] == 0 & !is.na(df[,x]), x] <- 0.0000000000000000000000001
  df[df[,x] == 1 & !is.na(df[,x]), x] <- 0.99999999999999994
}
```
## zscore centiles (new df)
```{r, echo = FALSE}
#zscore centiles and other variables of interest
df <- df %>%
  mutate(across(all_of(full_vars), qnorm,
                         .names = "{.col}.zscore"))
#set sex and preterm as factor
df$sex <- as.factor(df$sex)
df$preterm <- as.factor(df$preterm)
zscore_vars_names <- paste0(full_vars, ".zscore")
```

## find Inf cells, make a table of N per phenotype, should be 0
```{r, echo = FALSE}
inf_nan_table <- as.data.frame(t(sapply(df[zscore_vars_names], function(x){
  c(inf = sum(is.infinite(x)), nan = sum(is.nan(x)))
})))
max(inf_nan_table$inf)
max(inf_nan_table$nan)

inf_nan_table %>% filter(inf > 0)
```

# LMSz Growth Charts

## Define LMSz models
```{r, echo = TRUE}
# Define LMSz models to process
# Mean (mu) models
m_models = c('~ adjusted_age_days_log + sex + adjusted_age_days_log*sex',
             '~ adjusted_age_days_log + sex',
             '~ adjusted_age_days_log',
             '~ ns(adjusted_age_days_log, df=2) + sex + ns(adjusted_age_days_log, df=2)*sex',
             '~ ns(adjusted_age_days_log, df=2) + sex',
             '~ ns(adjusted_age_days_log, df=2)',
             '~ ns(adjusted_age_days_log, df=3) + sex + ns(adjusted_age_days_log, df=3)*sex',
             '~ ns(adjusted_age_days_log, df=3) + sex',
             '~ ns(adjusted_age_days_log, df=3)',
             '~ sex',
             '~ 1',
             '~ 0')

# Variance (sigma) models
s_models = c('~ adjusted_age_days_log + sex',
             '~ adjusted_age_days_log',
             '~ sex',
             '~ 0')
```

## Generate LMSz growth charts for VPM and save output
```{r, echo = FALSE}
vpm_save_path <- paste0(save_path,"lmsz_models_vpm/")
df_subset <- df %>% filter(preterm == "VPM")

vpm_lmsz <- lmsz_modeling(df_subset, 
                        m_models, 
                        s_models, 
                        full_vars, 
                        models_path, 
                        vpm_save_path, 
                        save_model = TRUE,
                        reweight = FALSE)

models_vpm <- vpm_lmsz$models
df_vpm <- vpm_lmsz$df
worm_plots_vpm <- vpm_lmsz$wp
rm(vpm_lmsz)
```

### check betas across vars and intercept to see if there are any phenotypes that are wack
```{r, echo = FALSE}
hist(models_vpm$intercept_coef)
hist(models_vpm$intercept_sigma_coef)

hist(models_vpm$age_coef)
hist(models_vpm$age_sigma_coef)

hist(models_vpm$sex_coef)
hist(models_vpm$sex_sigma_coef)
```

### Save Models and Centiles
```{r, echo = FALSE}
write_csv(models_vpm, glue("{csv_save_path}lmsz_models_vpm.csv"))
write_csv(df_vpm, glue("{csv_save_path}lmsz_adjusted_centiles_vpm.csv"))
saveRDS(worm_plots_vpm,glue("{csv_save_path}worm_plots_vpm.rds"))
```

## Generate LMSz growth charts for LPM and save output
```{r, echo = FALSE}
lpm_save_path <- paste0(save_path,"lmsz_models_lpm/")
df_subset <- df %>% filter(preterm == "LPM")

lpm_lmsz <- lmsz_modeling(df_subset, 
                        m_models, 
                        s_models, 
                        full_vars, 
                        models_path, 
                        lpm_save_path, 
                        save_model = TRUE,
                        reweight = FALSE)

models_lpm <- lpm_lmsz$models
df_lpm <- lpm_lmsz$df
worm_plots_lpm <- lpm_lmsz$wp
rm(lpm_lmsz)
```
### check betas across vars and intercept to see if there are any phenotypes that are wack
```{r, echo = FALSE}
hist(models_lpm$intercept_coef)
hist(models_lpm$intercept_sigma_coef)

hist(models_lpm$age_coef)
hist(models_lpm$age_sigma_coef)

hist(models_lpm$sex_coef)
hist(models_lpm$sex_sigma_coef)
```

### Save Models and Centiles
```{r, echo = FALSE}
write_csv(models_lpm, glue("{csv_save_path}lmsz_models_lpm.csv"))
write_csv(df_lpm, glue("{csv_save_path}lmsz_adjusted_centiles_lpm.csv"))
saveRDS(worm_plots_lpm,glue("{csv_save_path}worm_plots_lpm.rds"))
```

# Backtransform Centiles VPM

## Calculate Centile Fans
```{r, echo = FALSE}
##simulation frames, log days age and sex
#between 280 days (40 weeks gestation) and 21 years
sim_list_slip <- list()
sim_list_slip$Male <- expand.grid(adjusted_age_days_log = seq(log(280),log(365.25*21+280),0.01),
                                  sex = c("Male"))
sim_list_slip$Female <- expand.grid(adjusted_age_days_log = seq(log(280),log(365.25*21+280),0.01),
                                    sex = c("Female"))

#| results: false
for (pheno in full_vars) {
  print(pheno)

  #df_vpm[which(is.na(df_vpm$pheno)),glue("{pheno}.raw_normalised")]
  tmp <-  backtransform_centile_fans(pheno, 
                                     glue("{save_path}lmsz_models_vpm/"), 
                                     sim_list_slip, 
                                     df_vpm)
  if (pheno == full_vars[1]) {
    out_fans_slip <- tmp
  } else {
    out_fans_slip <- bind_rows(out_fans_slip, tmp)
  }
}
```

## Save centile fans calculated
```{r, echo = FALSE}
write_csv(out_fans_slip, glue("{csv_save_path}lmsz_centfans_vpm_uncorrected.csv"))
```

# Backtransform Centiles LPM

## Calculate Centile Fans
```{r, echo = FALSE}
##simulation frames, log days age and sex
#between 280 days (40 weeks gestation) and 21 years
sim_list_slip <- list()
sim_list_slip$Male <- expand.grid(adjusted_age_days_log = seq(log(280),log(365.25*21+280),0.01),
                                  sex = c("Male"))
sim_list_slip$Female <- expand.grid(adjusted_age_days_log = seq(log(280),log(365.25*21+280),0.01),
                                    sex = c("Female"))

#| results: false
for (pheno in full_vars) {
  print(pheno)

  #df_vpm[which(is.na(df_vpm$pheno)),glue("{pheno}.raw_normalised")]
  tmp <-  backtransform_centile_fans(pheno, 
                                     glue("{save_path}lmsz_models_lpm/"), 
                                     sim_list_slip, 
                                     df_lpm)
  if (pheno == full_vars[1]) {
    out_fans_slip <- tmp
  } else {
    out_fans_slip <- bind_rows(out_fans_slip, tmp)
  }
}
```

## Save centile fans calculated
```{r, echo = FALSE}
write_csv(out_fans_slip, glue("{csv_save_path}lmsz_centfans_lpm_uncorrected.csv"))
```
