---
title: "slipga_qnorm_df_cmd.Rmd"
author: "Eren Kafadar"
date: "2026-06-02"
output:
  html_document:
    df_print: paged
params:
  output_folder : "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/analysis_results/median-th-mpr/"
  centiles_distinct_path: "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/data_clean/csv/centiles_distinct_median_th_mpr_2026-04-24.csv"
  centiles_all_path: "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/data_clean/csv/centiles_all_median_th_mpr_2026-04-24.csv"
  gamlss_models_path: "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/gamlss_models_rds/"
editor_options: 
  chunk_output_type: console
---
# Define params if they do not exist
```{r params, echo = FALSE}
if (!exists("params")) {
  params <- list(
  output_folder = "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/analysis_results/",
  centiles_distinct_path = "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/data_clean/csv/centiles_distinct_median_th_mpr_2026-04-24.csv",
  centiles_all_path = "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/data_clean/csv/centiles_all_median_th_mpr_2026-04-24.csv",
  gamlss_models_path = "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/gamlss_models_rds/"
  )
}
print(params)
``` 

# Load Packages
```{r packages, echo = FALSE}
# Load libraries
require(dplyr)
require(glue)
require(magrittr)
require(readr)
require(stringr)
require(tidyr)
require(scales)
require(mgcv)
require(pbkrtest)
require(gamlssTools) # Margaret' package
require(tibble)
require(lme4)
require(purrr)
# For figures/plotting
require(ggplot2)
require(gridExtra)
require(ggsegDKT)
require(ggseg)
require(ggpubr)
require(ggcorrplot)
require(cowplot)
require(ggh4x)
require(egg)
require(interactions)
require(growthstandards)
require(gtsummary)
require(svglite)
require(patchwork)

`%nin%` = Negate(`%in%`)
```

# Controls
```{r setup, echo = TRUE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)
# set command
save_tables <- TRUE
plot <- TRUE
save_figs <- FALSE
global_measure_vol <- "global_eTIV"
global_measure_sa <- "global_SurfaceArea"
global_measure_th <- "global_MeanThickness"
set.seed(42)

output_folder <- params$output_folder
centiles_distinct_path <- params$centiles_distinct_path
centiles_all_path <- params$centiles_all_path
gamlss_models_path <- params$gamlss_models_path
```

# Define Paths
```{r, echo = TRUE}
# set folder paths
code_path <- "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/"
vars_path <-  "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/output/vars_2511_new/"

# output folder
output_folder <- params$output_folder
if (!dir.exists(output_folder)) dir.create(output_folder)
##  output subfolder paths
analysis_dfs_folder <- paste0(output_folder, "analysis_dfs/")
if (!dir.exists(analysis_dfs_folder)) dir.create(analysis_dfs_folder)

csv_output_folder <-  paste0(output_folder, "results_csv/")
if (!dir.exists(csv_output_folder)) dir.create(csv_output_folder)

  betas_folder <- paste0(csv_output_folder, "full_betas/")
  if (!dir.exists(betas_folder)) dir.create(betas_folder)
  dkt_folder <- paste0(csv_output_folder, "dkt_betas/")
  if (!dir.exists(dkt_folder)) dir.create(dkt_folder)
  slwindow_folder <- paste0(csv_output_folder, "sliding_window/")
  if (!dir.exists(slwindow_folder)) dir.create(slwindow_folder)
  sp_stat_folder <- paste0(csv_output_folder, "spline_stats/")
  if (!dir.exists(sp_stat_folder)) dir.create(sp_stat_folder)

plot_output_folder <- paste0(output_folder, "plots/")
if (!dir.exists(plot_output_folder)) dir.create(plot_output_folder)

  ggseg_plot_folder <- paste0(plot_output_folder, "ggseg_plots/")
  if (!dir.exists(ggseg_plot_folder)) dir.create(ggseg_output_folder)
  cmd_plot_folder <- paste0(plot_output_folder, "cmd_plots/")
  if (!dir.exists(cmd_plot_folder)) dir.create(cmd_plot_folder)
  modelvis_plot_folder <- paste0(plot_output_folder, "modelvis_plots/")
  if (!dir.exists(modelvis_plot_folder)) dir.create(modelvis_plot_folder)
  slwindow_plot_folder <- paste0(plot_output_folder, "sliding_window_plots/")
  if (!dir.exists(slwindow_plot_folder)) dir.create(slwindow_plot_folder)

```

# Load custom functions
```{r fcts, echo = FALSE}
source(glue("{code_path}sliding_window.R"))
source(glue("{code_path}visualize_model_bartfct.R"))
source(glue("{code_path}vis_spline_noint_EK.R"))

panel_save <- function(plot, filename, extension, folder, plot_width = 8, plot_height = 10){
  save_name <- paste0(folder, filename, ".", extension)
  ggsave(save_name, plot = plot, dpi = 300, width = plot_width, height = plot_height)
}
```

# Color Specifications
```{r, echo = TRUE}
limits = c(-0.3, 0.3)
low_color = "#4878CF" # blue
mid_color = "white"
high_color = "#D65F5F" # red

ga_color = "#6D8299" #muted blue
bw_color = "#709085" # muted green
bwp_color = "#A68A64" # muted beige
```

# Read in data files
```{r, echo = FALSE}
full_vars <- read.csv(paste0(vars_path,"all_vars.csv")) %>% pull(x)
global_vars <- read.csv(paste0(vars_path,"global_vars.csv")) %>% pull(x)
vol_vars <- read.csv(paste0(vars_path,"vol_vars.csv")) %>% pull(x)
sa_vars <- read.csv(paste0(vars_path,"sa_vars.csv")) %>% pull(x)
th_vars <- read.csv(paste0(vars_path,"th_vars.csv")) %>% pull(x)
th_vars_median <- c(paste0(th_vars, ".median"))
namemap <- read.csv(glue("{vars_path}name_mapping.csv"))

cent_distinct <- read.csv(centiles_distinct_path) %>% select(-c("X"))
cent_all <- read.csv(centiles_all_path) %>% select(-c("X"))
  
cent_distinct <-  cent_distinct %>% select(-(grep("_Z", names(cent_distinct)))) # This was for Liza's csvs so I can get their qnorm separately

#determine if there is both mpr and median for thickness values
median_th_mpr_flag <- any(grepl(".median", names(cent_distinct)))

## Define global vars of interest
global_vars_totest <- c("global_eTIV", "global_SubCortGrayVol","global_CortexVol", "global_CerebralWhiteMatterVol", "global_VentricleChoroidVol","global_SurfaceArea", "global_MeanThickness")
```

# Find if any models don't exist, remove from variable list for downstream analysis.
```{r, echo = FALSE}
vars_missing <- full_vars[!(full_vars %in% names(cent_all))]
print(paste0("Phenotype Variables Missing in the Data Provided: ", paste(vars_missing, collapse = ", ")))
full_vars <- full_vars[!(full_vars %in% vars_missing)]
global_vars <- global_vars[!(global_vars %in% vars_missing)]
vol_vars <- vol_vars[!(vol_vars %in% vars_missing)]
sa_vars <- sa_vars[!(sa_vars %in% vars_missing)]
th_vars <- th_vars[!(th_vars %in% vars_missing)]
```

# Find any cells with missing data within the centile scores
```{r, echo = FALSE}
if (median_th_mpr_flag){
check_vars <- c(vol_vars, sa_vars, global_vars, th_vars_median)
check_vars <- check_vars[check_vars != "global_MeanThickness"]
missing_data_sessions <- cent_all[rowSums(is.na(cent_all %>% select(all_of(check_vars)))) > 0, ]$session_id
print(paste0("Number of sessions with any missing centile data: ", length(missing_data_sessions)))
missing_data_sessions_df <- cent_all %>% filter(session_id %in% missing_data_sessions) %>% select(any_of(c("session_id", "participant_id", "adjusted_age_in_years", full_vars)))
} else {
check_vars <- c(full_vars)
missing_data_sessions <- cent_all[rowSums(is.na(cent_all %>% select(all_of(check_vars)))) > 0, ]$session_id
print(paste0("Number of sessions with any missing centile data: ", length(missing_data_sessions)))
missing_data_sessions_df <- cent_all %>% filter(session_id %in% missing_data_sessions) %>% select(any_of(c("session_id", "participant_id", "adjusted_age_in_years", full_vars)))
}
# remove sessions with any missing data (especially for calculating CMDs)
cent_all <- cent_all[!(cent_all$session_id %in% missing_data_sessions),]
```

# Deal with QC Variable Naming Difference
```{r, echo = FALSE}
if(!("median_euler_mean" %in% names(cent_all))){
    cent_all$qc_var <- cent_all$euler_mean
    cent_distinct$qc_var <- cent_distinct$euler_mean
} else if("median_euler_mean" %in% names(cent_all)){
  cent_all$qc_var <- cent_all$median_euler_mean
  cent_distinct$qc_var <- cent_distinct$median_euler_mean
}
```

# Remove BW outliers, and corresponding BWP
```{r}
mu_bw <- mean(cent_distinct$birth_weight_kg, na.rm = TRUE)
sd_bw <- sd(cent_distinct$birth_weight_kg, na.rm = TRUE)

rm_dist = 5

# cent distinct

nrow(cent_distinct %>% filter(
  birth_weight_kg < mu_bw-sd_bw*rm_dist | birth_weight_kg > mu_bw+sd_bw*rm_dist) )

cent_distinct <- cent_distinct %>%
  mutate(birth_weight_kg_inc_out = birth_weight_kg,
    birth_weight_kg = case_when(
        (birth_weight_kg < mu_bw-sd_bw*rm_dist | birth_weight_kg > mu_bw+sd_bw*rm_dist) ~ NA,
        TRUE ~ birth_weight_kg),
    bwp_fen_inc_out = bwp_fen,
    bwp_fen = case_when(
        is.na(birth_weight_kg) ~ NA,
        TRUE ~ bwp_fen))

sum(!is.na(cent_distinct$birth_weight_kg))
range(cent_distinct$birth_weight_kg, na.rm = TRUE)
sum(!is.na(cent_distinct$birth_weight_kg_outrm))
range(cent_distinct$birth_weight_kg_outrm, na.rm = TRUE)

# cent all
nrow(cent_all %>% filter(
  birth_weight_kg < mu_bw-sd_bw*rm_dist | birth_weight_kg > mu_bw+sd_bw*rm_dist) )

cent_all <- cent_all %>%
  mutate(birth_weight_kg_inc_out = birth_weight_kg,
    birth_weight_kg = case_when(
        (birth_weight_kg < mu_bw-sd_bw*rm_dist | birth_weight_kg > mu_bw+sd_bw*rm_dist) ~ NA,
        TRUE ~ birth_weight_kg),
    bwp_fen_inc_out = bwp_fen,
    bwp_fen = case_when(
        (birth_weight_kg < mu_bw-sd_bw*rm_dist | birth_weight_kg > mu_bw+sd_bw*rm_dist) ~ NA,
        TRUE ~ bwp_fen))
```
# Prepare Df for Analysis

## Set 0/1 centiles to min/max value to avoid Inf/-Inf
numbers from gamlssTools pkg
include weight/height centiles here later (when properly calculated and exported from arcus)
```{r}
complete_centile_vars <- c(full_vars, th_vars_median, "global_MeanThickness.median", "bwp_fen")

for (x in complete_centile_vars){
  cent_distinct[cent_distinct[,x] == 0 & !is.na(cent_distinct[,x]), x] <- 0.0000000000000000000000001
  cent_distinct[cent_distinct[,x] == 1 & !is.na(cent_distinct[,x]), x] <- 0.99999999999999994
  
  cent_all[cent_all[,x] == 0 & !is.na(cent_all[,x]), x] <- 0.0000000000000000000000001
  cent_all[cent_all[,x] == 1 & !is.na(cent_all[,x]), x] <- 0.99999999999999994
}
```

## Set marker for >= 2 years (adjusted age) and >= 1 years (unadjusted age) for thickness sensitivity/analyses
```{r}
cent_distinct$age_over2_adj <- ifelse(cent_distinct$age_years_adj >=2, TRUE, FALSE)
cent_distinct$age_over2_adj <- ifelse(cent_distinct$age_years_adj >=2, TRUE, FALSE)
cent_distinct$age_over1_unadj <- ifelse(cent_distinct$age_at_scan_years >=1, TRUE, FALSE)
cent_distinct$age_over1_unadj <- ifelse(cent_distinct$age_at_scan_years >=1, TRUE, FALSE)

cent_all$age_over2_adj <- ifelse(cent_all$age_years_adj >=2, TRUE, FALSE)
cent_all$age_over2_adj <- ifelse(cent_all$age_years_adj >=2, TRUE, FALSE)
cent_all$age_over1_unadj <- ifelse(cent_all$age_at_scan_years >=1, TRUE, FALSE)
cent_all$age_over1_unadj <- ifelse(cent_all$age_at_scan_years >=1, TRUE, FALSE)
```

## zscore centiles (new df)
```{r, echo = FALSE}
# zscore centiles and other variables of interest
if (median_th_mpr_flag) {
df.zscore <- cent_distinct %>%
  mutate(across(all_of(c(full_vars, th_vars_median, "global_MeanThickness.median", 
                         "bwp_fen", "bwp_fen_inc_out")), qnorm)) %>%
  mutate(across(all_of(c("gestational_age", "birth_weight_kg", 
                         "birth_weight_kg_inc_out",
                         "age_at_scan", "age_at_scan_years", "age_days_adj", "age_years_adj",
                         "qc_var", "euler_mean_mpr", "median_euler_mean")), 
                ~as.numeric(scale(.x))))
} else {
  df.zscore <- cent_distinct %>%
  mutate(across(all_of(c(full_vars, "bwp_fen", "bwp_fen_inc_out")), qnorm)) %>%
  mutate(across(all_of(c("gestational_age", "birth_weight_kg", 
                         "birth_weight_kg_inc_out",
                         "age_at_scan", "age_at_scan_years", "age_days_adj", "age_years_adj",
                         "qc_var")), 
                ~as.numeric(scale(.x))))
}

# set sex and preterm as factor
df.zscore$sex <- as.factor(df.zscore$sex)
df.zscore$preterm <- as.factor(df.zscore$preterm)

# zscore centiles and other variables of interest
if (median_th_mpr_flag) {
df.zscore_all <- cent_all %>%
  mutate(across(all_of(c(full_vars, th_vars_median, "bwp_fen", "bwp_fen_inc_out")), qnorm)) %>%
  mutate(across(all_of(c("gestational_age", "birth_weight_kg", 
                         "birth_weight_kg_inc_out",
                         "age_at_scan", "age_at_scan_years", "age_days_adj", "age_years_adj",
                         "qc_var", "euler_mean_mpr", "median_euler_mean")), 
                ~as.numeric(scale(.x))))
} else {
  df.zscore_all <- cent_all %>%
  mutate(across(all_of(c(full_vars, "bwp_fen", "bwp_fen_inc_out")), qnorm)) %>%
  mutate(across(all_of(c("gestational_age", "birth_weight_kg", 
                         "birth_weight_kg_inc_out",
                         "age_at_scan", "age_at_scan_years", "age_days_adj", "age_years_adj",
                         "qc_var")), 
                ~as.numeric(scale(.x))))
}

# set sex and preterm as factor
df.zscore_all$sex <- as.factor(df.zscore_all$sex)
df.zscore_all$preterm <- as.factor(df.zscore_all$preterm)
```

## find Inf cells, sanity check, should be none
```{r, echo = FALSE}
inf_nan_table <- as.data.frame(t(sapply(df.zscore[c(full_vars, "bwp_fen")], function(x){
  c(inf = sum(is.infinite(x)), nan = sum(is.nan(x)))
})))
max(inf_nan_table$inf)
max(inf_nan_table$nan)

inf_nan_table %>% filter(inf > 0)
```

## Store non-normalized distribution stats to convert back for plots
```{r}
stats_cent_all <- cent_all %>%
  dplyr::summarise(across(all_of(c("gestational_age", "birth_weight_kg",
                                   "birth_weight_kg_inc_out",
                         "age_at_scan", "age_at_scan_years", "age_days_adj", "age_years_adj",
                         "qc_var", "euler_mean_mpr", "median_euler_mean")),
                         list(mean = ~ mean(.x, na.rm = TRUE),
                              sd = ~ sd(.x, na.rm = TRUE)),
                         .names = "{.col}__{.fn}")) %>%
  pivot_longer(everything(),
               names_to = c("var", ".value"),
               names_pattern = "^(.*)__(mean|sd)$")

stats_cent_distinct <- cent_distinct %>%
  dplyr::summarise(across(all_of(c("gestational_age", "birth_weight_kg",
                                   "birth_weight_kg_inc_out",
                         "age_at_scan", "age_at_scan_years", "age_days_adj", "age_years_adj",
                         "qc_var", "euler_mean_mpr", "median_euler_mean")),
                         list(mean = ~ mean(.x, na.rm = TRUE),
                              sd = ~ sd(.x, na.rm = TRUE)),
                         .names = "{.col}__{.fn}")) %>%
  pivot_longer(everything(),
               names_to = c("var", ".value"),
               names_pattern = "^(.*)__(mean|sd)$")

mean_lookup_all <- setNames(stats_cent_all$mean, stats_cent_all$var)
sd_lookup_all <- setNames(stats_cent_all$sd, stats_cent_all$var)

mean_lookup_distinct <- setNames(stats_cent_distinct$mean, stats_cent_distinct$var)
sd_lookup_distinct <- setNames(stats_cent_distinct$sd, stats_cent_distinct$var)

if(save_tables){
  save(mean_lookup_distinct, sd_lookup_distinct, file = glue("{analysis_dfs_folder}lookup_stats_distinct.RData"))
  save(mean_lookup_all,sd_lookup_all, file = glue("{analysis_dfs_folder}lookup_stats_all.RData"))
}

```


# CMD on qnorm(centiles)

## calculate CMDs on distinct
```{r, echo = FALSE}
vol_vars_cortical <- vol_vars[grep("aparc", vol_vars)]
vol_vars_subcortical <- vol_vars[grep("aseg", vol_vars)]
names.global <- c("global_SubCortGrayVol","global_CortexVol", "global_CerebralWhiteMatterVol", "global_VentricleChoroidVol")

# CMD Calculation
## global
global.center <- colMeans(df.zscore %>% select(all_of(names.global)))
global.cov_matrix <- cov(df.zscore %>% select(all_of(names.global)))
global.md <- sqrt(mahalanobis(df.zscore %>% select(all_of(names.global)), global.center, global.cov_matrix))
df.zscore$global.cmd <- global.md

## vol
vol.center <- colMeans(df.zscore %>% select(all_of(vol_vars_cortical)))
vol.cov_matrix <- cov(df.zscore %>% select(all_of(vol_vars_cortical)))
vol.md <- sqrt(mahalanobis(df.zscore %>% select(all_of(vol_vars_cortical)), vol.center, vol.cov_matrix))
df.zscore$vol.cmd <- vol.md

## sa
sa.center <- colMeans(df.zscore %>% select(all_of(sa_vars)))
sa.cov_matrix <- cov(df.zscore %>% select(all_of(sa_vars)))
sa.md <- sqrt(mahalanobis(df.zscore %>% select(all_of(sa_vars)), sa.center, sa.cov_matrix))
df.zscore$sa.cmd <- sa.md

## th - calculate above 2 year adjusted, using mpr if the df is median-th-mpr
th.center <- colMeans(df.zscore %>% filter(age_over2_adj == TRUE) %>% select(all_of(th_vars)), na.rm = TRUE)
th.cov_matrix <- cov(df.zscore %>% filter(age_over2_adj == TRUE) %>% select(all_of(th_vars)), use = "complete.obs")
th.md <- sqrt(mahalanobis(df.zscore %>% filter(age_over2_adj == TRUE) %>% select(all_of(th_vars)), th.center, th.cov_matrix))
df.zscore[df.zscore$age_over2_adj == TRUE,"th.cmd"]  <- th.md

## if the df is median-th-mpr th - calculate all ages, using median
if(median_th_mpr_flag){
  th.center_median_full <- colMeans(df.zscore %>% select(all_of(th_vars_median)), na.rm = TRUE)
  th.cov_matrix_median_full <- cov(df.zscore %>% select(all_of(th_vars_median)), use = "complete.obs")
  th.md_median_full <- sqrt(mahalanobis(df.zscore %>% select(all_of(th_vars_median)), th.center_median_full, th.cov_matrix_median_full))
  df.zscore$th.cmd_median_full  <- th.md_median_full
}

## subcort
subcort.center <- colMeans(df.zscore %>% select(all_of(vol_vars_subcortical)))
subcort.cov_matrix <- cov(df.zscore %>% select(all_of(vol_vars_subcortical)))
subcort.md <- sqrt(mahalanobis(df.zscore %>% select(all_of(vol_vars_subcortical)), subcort.center, subcort.cov_matrix))
df.zscore$subcort.cmd <- subcort.md
```

## calculate CMDs on all (rm)
** might need to check how to deal with rm when calculating CMD
```{r, echo = FALSE}
vol_vars_cortical <- vol_vars[grep("aparc", vol_vars)]
vol_vars_subcortical <- vol_vars[grep("aseg", vol_vars)]
names.global <- c("global_SubCortGrayVol","global_CortexVol", "global_CerebralWhiteMatterVol", "global_VentricleChoroidVol")

# CMD Calculation
## global
global.center <- colMeans(df.zscore_all %>% select(all_of(names.global)))
global.cov_matrix <- cov(df.zscore_all %>% select(all_of(names.global)))
global.md <- sqrt(mahalanobis(df.zscore_all %>% select(all_of(names.global)), global.center, global.cov_matrix))
df.zscore_all$global.cmd <- global.md

## vol
vol.center <- colMeans(df.zscore_all %>% select(all_of(vol_vars_cortical)))
vol.cov_matrix <- cov(df.zscore_all %>% select(all_of(vol_vars_cortical)))
vol.md <- sqrt(mahalanobis(df.zscore_all %>% select(all_of(vol_vars_cortical)), vol.center, vol.cov_matrix))
df.zscore_all$vol.cmd <- vol.md

## sa
sa.center <- colMeans(df.zscore_all %>% select(all_of(sa_vars)))
sa.cov_matrix <- cov(df.zscore_all %>% select(all_of(sa_vars)))
sa.md <- sqrt(mahalanobis(df.zscore_all %>% select(all_of(sa_vars)), sa.center, sa.cov_matrix))
df.zscore_all$sa.cmd <- sa.md

## th - calculate above 2 year adjusted, using mpr if the df is median-th-mpr
th.center <- colMeans(df.zscore_all %>% filter(age_over2_adj == TRUE) %>% select(all_of(th_vars)), na.rm = TRUE)
th.cov_matrix <- cov(df.zscore_all %>% filter(age_over2_adj == TRUE) %>% select(all_of(th_vars)), use = "complete.obs")
th.md <- sqrt(mahalanobis(df.zscore_all %>% filter(age_over2_adj == TRUE) %>% select(all_of(th_vars)), th.center, th.cov_matrix))
df.zscore_all[df.zscore_all$age_over2_adj == TRUE,"th.cmd"]  <- th.md

## if the df is median-th-mpr th - calculate all ages, using median
if(median_th_mpr_flag){
  th.center_median_full <- colMeans(df.zscore_all %>% select(all_of(th_vars_median)), na.rm = TRUE)
  th.cov_matrix_median_full <- cov(df.zscore_all %>% select(all_of(th_vars_median)), use = "complete.obs")
  th.md_median_full <- sqrt(mahalanobis(df.zscore_all %>% select(all_of(th_vars_median)), th.center_median_full, th.cov_matrix_median_full))
  df.zscore_all$th.cmd_median_full  <- th.md_median_full
}

## subcort
subcort.center <- colMeans(df.zscore_all %>% select(all_of(vol_vars_subcortical)))
subcort.cov_matrix <- cov(df.zscore_all %>% select(all_of(vol_vars_subcortical)))
subcort.md <- sqrt(mahalanobis(df.zscore_all %>% select(all_of(vol_vars_subcortical)), subcort.center, subcort.cov_matrix))
df.zscore_all$subcort.cmd <- subcort.md
```

## Remove outliers based on SD from the mean

### Document N of outliers removed for Distinct
```{r, echo = FALSE}
sd_remove = 5
print(glue("SD Distance to Remove: {sd_remove}"))

print(glue("Global CMD N removed: {sum((df.zscore$global.cmd < (mean(df.zscore$global.cmd) - sd_remove*sd(df.zscore$global.cmd))) | 
      (df.zscore$global.cmd > (mean(df.zscore$global.cmd) + sd_remove*sd(df.zscore$global.cmd))))}"))

print(glue("Volume CMD N removed: {sum((df.zscore$vol.cmd < (mean(df.zscore$vol.cmd) - sd_remove*sd(df.zscore$vol.cmd))) | 
      (df.zscore$vol.cmd > (mean(df.zscore$vol.cmd) + sd_remove*sd(df.zscore$vol.cmd))))}"))

print(glue("SA CMD N removed: {sum((df.zscore$sa.cmd < (mean(df.zscore$sa.cmd) - sd_remove*sd(df.zscore$sa.cmd))) | 
      (df.zscore$sa.cmd > (mean(df.zscore$sa.cmd) + sd_remove*sd(df.zscore$sa.cmd))))}"))

print(glue("TH CMD, >= 2yrs adj age N removed: {sum((df.zscore$th.cmd < (mean(df.zscore$th.cmd, na.rm = TRUE) - sd_remove*sd(df.zscore$th.cmd, na.rm = TRUE))) | (df.zscore$th.cmd > (mean(df.zscore$th.cmd, na.rm = TRUE) + sd_remove*sd(df.zscore$th.cmd, na.rm = TRUE))), na.rm = TRUE)}"))

print(glue("SubCort CMD N removed: {sum((df.zscore$subcort.cmd < (mean(df.zscore$subcort.cmd) - sd_remove*sd(df.zscore$subcort.cmd))) | 
      (df.zscore$subcort.cmd > (mean(df.zscore$subcort.cmd) + sd_remove*sd(df.zscore$subcort.cmd))))}"))

if(median_th_mpr_flag){
  print(glue("TH CMD, full sample, median N removed: {sum((df.zscore$th.cmd_median_full < (mean(df.zscore$th.cmd_median_full) - sd_remove*sd(df.zscore$th.cmd_median_full))) | 
      (df.zscore$th.cmd_median_full > (mean(df.zscore$th.cmd_median_full) + sd_remove*sd(df.zscore$th.cmd_median_full))))}"))
}
```

### Remove outliers for Distinct
```{r, echo = FALSE}
#sd_remove = 5
#print(glue("SD Distance to Remove: {sd_remove}"))
#sum(is.na(df.zscore$th.cmd))

df.zscore <- df.zscore %>%
  mutate(global.cmd = case_when(
    (global.cmd < (mean(global.cmd) - sd_remove*sd(global.cmd))) | 
      (global.cmd > (mean(global.cmd) + sd_remove*sd(global.cmd))) ~ NA,
    TRUE ~ global.cmd),
    vol.cmd = case_when(
    (vol.cmd < (mean(vol.cmd) - sd_remove*sd(vol.cmd))) | 
      (vol.cmd > (mean(vol.cmd) + sd_remove*sd(vol.cmd))) ~ NA,
    TRUE ~ vol.cmd),
    sa.cmd = case_when(
    (sa.cmd < (mean(sa.cmd) - sd_remove*sd(sa.cmd))) | 
      (sa.cmd > (mean(sa.cmd) + sd_remove*sd(sa.cmd))) ~ NA,
    TRUE ~ sa.cmd),
    th.cmd = case_when(
    (th.cmd < (mean(th.cmd, na.rm = TRUE) - sd_remove*sd(th.cmd, na.rm = TRUE))) | 
      (th.cmd > (mean(th.cmd, na.rm = TRUE) + sd_remove*sd(th.cmd, na.rm = TRUE))) ~ NA,
    TRUE ~ th.cmd),
     subcort.cmd = case_when(
    (subcort.cmd < (mean(subcort.cmd) - sd_remove*sd(subcort.cmd))) | 
      (subcort.cmd > (mean(subcort.cmd) + sd_remove*sd(subcort.cmd))) ~ NA,
    TRUE ~ subcort.cmd) )

if(median_th_mpr_flag){
  df.zscore <- df.zscore %>%
  mutate(th.cmd_median_full = case_when(
    (th.cmd_median_full < (mean(th.cmd_median_full) - sd_remove*sd(th.cmd_median_full))) | 
      (th.cmd_median_full > (mean(th.cmd_median_full) + sd_remove*sd(th.cmd_median_full))) ~ NA,
    TRUE ~ th.cmd_median_full) )
  sum(is.na(df.zscore$th.cmd_median_full))
  
}

#sum(is.na(df.zscore$global.cmd))
#sum(is.na(df.zscore$vol.cmd))
#sum(is.na(df.zscore$sa.cmd))
#sum(is.na(df.zscore$th.cmd))
#sum(is.na(df.zscore$subcort.cmd))
```


### Document N of outliers removed for All (rm)
```{r, echo = FALSE}
sd_remove = 5
print(glue("SD Distance to Remove: {sd_remove}"))

print(glue("Global CMD N removed: {sum((df.zscore_all$global.cmd < (mean(df.zscore_all$global.cmd) - sd_remove*sd(df.zscore_all$global.cmd))) | 
      (df.zscore_all$global.cmd > (mean(df.zscore_all$global.cmd) + sd_remove*sd(df.zscore_all$global.cmd))))}"))

print(glue("Volume CMD N removed: {sum((df.zscore_all$vol.cmd < (mean(df.zscore_all$vol.cmd) - sd_remove*sd(df.zscore_all$vol.cmd))) | 
      (df.zscore_all$vol.cmd > (mean(df.zscore_all$vol.cmd) + sd_remove*sd(df.zscore_all$vol.cmd))))}"))

print(glue("SA CMD N removed: {sum((df.zscore_all$sa.cmd < (mean(df.zscore_all$sa.cmd) - sd_remove*sd(df.zscore_all$sa.cmd))) | 
      (df.zscore_all$sa.cmd > (mean(df.zscore_all$sa.cmd) + sd_remove*sd(df.zscore_all$sa.cmd))))}"))

print(glue("TH CMD, >= 2yrs adj age N removed: {sum((df.zscore_all$th.cmd < (mean(df.zscore_all$th.cmd, na.rm = TRUE) - sd_remove*sd(df.zscore_all$th.cmd, na.rm = TRUE))) | (df.zscore_all$th.cmd > (mean(df.zscore_all$th.cmd, na.rm = TRUE) + sd_remove*sd(df.zscore_all$th.cmd, na.rm = TRUE))), na.rm = TRUE)}"))

print(glue("SubCort CMD N removed: {sum((df.zscore_all$subcort.cmd < (mean(df.zscore_all$subcort.cmd) - sd_remove*sd(df.zscore_all$subcort.cmd))) | 
      (df.zscore_all$subcort.cmd > (mean(df.zscore_all$subcort.cmd) + sd_remove*sd(df.zscore_all$subcort.cmd))))}"))

if(median_th_mpr_flag){
  print(glue("TH CMD, full sample, median N removed: {sum((df.zscore_all$th.cmd_median_full < (mean(df.zscore_all$th.cmd_median_full) - sd_remove*sd(df.zscore_all$th.cmd_median_full))) | 
      (df.zscore_all$th.cmd_median_full > (mean(df.zscore_all$th.cmd_median_full) + sd_remove*sd(df.zscore_all$th.cmd_median_full))))}"))
}
```

### Remove outliers for All (rm)
```{r, echo = FALSE}
#sd_remove = 5
#print(glue("SD Distance to Remove: {sd_remove}"))
#sum(is.na(df.zscore_all$th.cmd))

df.zscore_all <- df.zscore_all %>%
  mutate(global.cmd = case_when(
    (global.cmd < (mean(global.cmd) - sd_remove*sd(global.cmd))) | 
      (global.cmd > (mean(global.cmd) + sd_remove*sd(global.cmd))) ~ NA,
    TRUE ~ global.cmd),
    vol.cmd = case_when(
    (vol.cmd < (mean(vol.cmd) - sd_remove*sd(vol.cmd))) | 
      (vol.cmd > (mean(vol.cmd) + sd_remove*sd(vol.cmd))) ~ NA,
    TRUE ~ vol.cmd),
    sa.cmd = case_when(
    (sa.cmd < (mean(sa.cmd) - sd_remove*sd(sa.cmd))) | 
      (sa.cmd > (mean(sa.cmd) + sd_remove*sd(sa.cmd))) ~ NA,
    TRUE ~ sa.cmd),
    th.cmd = case_when(
    (th.cmd < (mean(th.cmd, na.rm = TRUE) - sd_remove*sd(th.cmd, na.rm = TRUE))) | 
      (th.cmd > (mean(th.cmd, na.rm = TRUE) + sd_remove*sd(th.cmd, na.rm = TRUE))) ~ NA,
    TRUE ~ th.cmd),
     subcort.cmd = case_when(
    (subcort.cmd < (mean(subcort.cmd) - sd_remove*sd(subcort.cmd))) | 
      (subcort.cmd > (mean(subcort.cmd) + sd_remove*sd(subcort.cmd))) ~ NA,
    TRUE ~ subcort.cmd) )

if(median_th_mpr_flag){
  df.zscore_all <- df.zscore_all %>%
  mutate(th.cmd_median_full = case_when(
    (th.cmd_median_full < (mean(th.cmd_median_full) - sd_remove*sd(th.cmd_median_full))) | 
      (th.cmd_median_full > (mean(th.cmd_median_full) + sd_remove*sd(th.cmd_median_full))) ~ NA,
    TRUE ~ th.cmd_median_full) )
  sum(is.na(df.zscore_all$th.cmd_median_full))
  
}

#sum(is.na(df.zscore_all$global.cmd))
#sum(is.na(df.zscore_all$vol.cmd))
#sum(is.na(df.zscore_all$sa.cmd))
#sum(is.na(df.zscore_all$th.cmd))
#sum(is.na(df.zscore_all$subcort.cmd))
```

## Correlate CMDs with each other (using distinct)
```{r, echo = FALSE}
# correlations between CMDs across three centiles
cor_mat <- cor(df.zscore[,c("global.cmd", "vol.cmd", "sa.cmd", "th.cmd", "subcort.cmd")], use = "pairwise.complete.obs")

p_mat <- cor_pmat(df.zscore[,c("global.cmd", "vol.cmd", "sa.cmd", "th.cmd", "subcort.cmd")], use = "pairwise.complete.obs")

(cmd_corrs_plot <- ggcorrplot(
  cor_mat,
  p.mat = p_mat,
  type = "lower",
  insig = "blank",
  lab = TRUE
))

if(save_figs == TRUE){
panel_save(cmd_corrs_plot, "cmd-cor-matrix", "svg", cmd_plot_folder)
}
```

## correlations between CMDs and region-specific centile scores (using distinct)
```{r, echo = FALSE}
# volume
cor_matrix <- cor(df.zscore[,c(vol_vars_cortical, "vol.cmd")], use = "pairwise.complete.obs")
p_mat <- cor_pmat(df.zscore[,c(vol_vars_cortical, "vol.cmd")])

cor_vol.cmd <- cor_matrix[, "vol.cmd" , drop = FALSE]
p_vol.cmd <- p_mat[, "vol.cmd" , drop = FALSE]

ggcorrplot(
  cor_vol.cmd,
  p.mat = p_vol.cmd,
  type = "lower",
  insig = "blank",
  lab = FALSE
)

# sa
cor_matrix <- cor(df.zscore[,c(sa_vars, "sa.cmd")], use = "pairwise.complete.obs")
p_mat <- cor_pmat(df.zscore[,c(sa_vars, "sa.cmd")])

cor_sa.cmd <- cor_matrix[, "sa.cmd" , drop = FALSE]
p_sa.cmd <- p_mat[, "sa.cmd" , drop = FALSE]

ggcorrplot(
  cor_sa.cmd,
  p.mat = p_sa.cmd,
  type = "lower",
  insig = "blank",
  lab = FALSE
)

# th
cor_matrix <- cor(df.zscore[,c(th_vars, "th.cmd")], use = "pairwise.complete.obs")
p_mat <- cor_pmat(df.zscore[,c(th_vars, "th.cmd")])

cor_th.cmd <- cor_matrix[, "th.cmd" , drop = FALSE]
p_th.cmd <- p_mat[, "th.cmd" , drop = FALSE]

ggcorrplot(
  cor_th.cmd,
  p.mat = p_th.cmd,
  type = "lower",
  insig = "blank",
  lab = FALSE
)
```

## CMD Correlations with GA and BW (using distinct)
```{r, echo = FALSE}
vars_bf <- c("gestational_age", "birth_weight_kg", "bwp_fen")
vars_cmd <- c("global.cmd", "vol.cmd", "sa.cmd", "th.cmd", "subcort.cmd")

cmd_bf_plot_df <- df.zscore %>% 
  select(all_of(c(vars_bf, vars_cmd, 
                  "participant_id", "session_id"))) %>%
  pivot_longer(all_of(vars_bf), names_to = "bf_var", values_to = "bf") %>%
  pivot_longer(all_of(vars_cmd), names_to = "cmd_var", values_to = "cmd") %>%
  mutate(bf_var = factor(bf_var, levels = vars_bf),
         cmd_var = factor(cmd_var, levels = vars_cmd))

breaks_ga <- map2_dbl(c(23, 26, 29, 32, 35, 38, 41), "gestational_age", ~ (.x*7 - mean_lookup_distinct[[.y]])/sd_lookup_distinct[[.y]])

breaks_bw <- pretty(df.zscore$birth_weight_kg, n = 5)

breaks_bwp <- qnorm(c(.05,.50,.95))

bf_scales <- list(
  "gestational_age" = scale_x_continuous(
    breaks = breaks_ga, 
    labels = paste0(round(map2_dbl(breaks_ga, "gestational_age", ~ .x*sd_lookup_distinct[[.y]] + mean_lookup_distinct[[.y]])/7))),
  "birth_weight_kg" = scale_x_continuous(
    breaks = breaks_bw, 
    labels = paste0(round(map2_dbl(breaks_bw, "birth_weight_kg", ~ .x*sd_lookup_distinct[[.y]] + mean_lookup_distinct[[.y]]))) ),
  "bwp_fen" = scale_x_continuous(
    breaks = breaks_bwp, 
    labels = paste0(round(pnorm(breaks_bwp)*100))) )

# cmd_scales <- list(
#   "global.cmd" = scale_y_continuous(breaks = c(0, 50, 100), position = "right")
# )

cmd_labels <- c("global.cmd" = "Global", 
                "vol.cmd" = "Volume",
                "sa.cmd" = "Surface Area",
                "th.cmd" = "Thickness",
                "subcort.cmd" = "Subcortical")
bf_labels  <- c("gestational_age" = "Gestational Age (wks)", 
                "birth_weight_kg" = "Birth Weight (kg)",
                "bwp_fen" = "Birth Weight Percentile")


(cmd_bf_plot <- ggplot(cmd_bf_plot_df, aes(x = bf, y = cmd, color = bf_var)) +
  geom_point(alpha = 0.15, size = 0.75) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.7) +
  ggpubr::stat_cor(
    method = "pearson",
    aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")),
    label.x.npc = .4,
    label.y.npc = .95,
    size = 5,
    color = "gray30") +
  scale_color_manual(values = c("gestational_age" = ga_color,
                              "birth_weight_kg" = bw_color,
                              "bwp_fen" = bwp_color)) +
  facet_grid(cmd_var ~ bf_var, scales = "free",
             labeller = labeller(cmd_var = cmd_labels, bf_var = bf_labels),
             switch = "both") +
  ggh4x::facetted_pos_scales(x = bf_scales) +
  labs(x = NULL,
    y = "Centile Mahalanobis Distance") +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        strip.placement = "outside",
        strip.background = element_blank(),
        strip.text.y = element_text(face = "plain", size = 24),
        strip.text.x = element_text(face = "plain", size = 30),
        axis.title.y = element_text(face = "bold", size = 30),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_text(size = 20),
        legend.position = "none",
        panel.spacing.x = unit(0.5, "cm"),
        panel.spacing.y = unit(0.25, "cm")) )

if(save_figs == TRUE){
panel_save(cmd_bf_plot, "cmd-bf-facet-plot", "svg", cmd_plot_folder, plot_width = 14, plot_height = 10)
}
```

# Save df.zscore dfs with CMDs
```{r}
if(save_tables){
  write.csv(df.zscore, glue("{analysis_dfs_folder}df_zscore_distinct.csv"))
  write.csv(df.zscore_all, glue("{analysis_dfs_folder}df_zscore_all.csv"))
}
```