---
title: "slip_geocode"
output: html_document
date: "2026-04-01"
editor_options: 
  chunk_output_type: console
---

# Controls
```{r setup, echo = TRUE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)
save_figs = TRUE
save_tables = TRUE
```

# Load packages
```{r}
#Load packages. Install these packages with install.packages("pkg_name") if they are not installed.
library(bigrquery) #package to import SQL tables into R
library(tidyverse) #package suite for data manipulation, reading data into R, and plotting (ggplot2)
library(glue) #package to use expressions with strings
library(ggseg)
require(gridExtra)
library(interactions)
library(ggcorrplot)

panel_save <- function(plot, filename, extension, folder){
  save_name <- paste0(folder, filename, ".", extension)
  ggsave(save_name, plot = plot, dpi = 300)
}
```

# Define paths
```{r}
save_path <-  "/mnt/arcus/lab/users/kafadare/geo_results/"
if(!dir.exists(save_path)) dir.create(save_path)

plot_output_folder <- paste(save_path, "plots/")
if(!dir.exists(plot_output_folder)) dir.create(plot_output_folder)

csv_output_folder <-  paste0(save_path, "results_csv/")
if (!dir.exists(csv_output_folder)) dir.create(csv_output_folder)

betas_folder <- paste0(csv_output_folder, "full_betas/")
if (!dir.exists(betas_folder)) dir.create(betas_folder)

dkt_folder <- paste0(csv_output_folder, "dkt_betas/")
if (!dir.exists(dkt_folder)) dir.create(dkt_folder)
```

# Load tables from Arcus
```{r}
require(bigrquery)
bq_auth()
proj_id <- if (exists('proj_id')) proj_id else bq_projects()[1]

#Download SWYC table
#query <- glue("SELECT * FROM lab.sorted_geospatial") #the string here is the SQL query
#geo <- bq_project_query(proj_id, query) %>% bq_table_download(., page_size=3500) #assign the name of the dataframe in R here
#print(paste0("Arcus lab.sorted_geospatial table, dimensions: ", dim(geo))) #pastes the dimensions of the table downloaded

#Download SWYC table
query <- glue("SELECT * FROM lab.swyc_normative_scores_20260129") #the string here is the SQL query
swyc <- bq_project_query(proj_id, query) %>% bq_table_download(., page_size=3500) #assign the name of the dataframe in R here
print(paste0("Arcus lab.SWYC table, dimensions: ", dim(swyc))) #pastes the dimensions of the table downloaded

#Download Participants table
query <- glue("SELECT * FROM arcus.patient") #the string here is the SQL query
participants <- bq_project_query(proj_id, query) %>% bq_table_download(., page_size=3500) #assign the name of the dataframe in R here
print(paste0("Arcus arcus.patient table, dimensions: ", dim(participants))) #pastes the dimensions of the table downloaded

```

# Load vars names
```{r}
vars_path <- "/mnt/arcus/lab/users/kafadare/slip_brain_vars/"
full_vars <- read.csv(paste0(vars_path,"all_vars.csv")) %>% pull(x)
global_vars <- read.csv(paste0(vars_path,"global_vars.csv")) %>% pull(x)
vol_vars <- read.csv(paste0(vars_path,"vol_vars.csv")) %>% pull(x)
sa_vars <- read.csv(paste0(vars_path,"sa_vars.csv")) %>% pull(x)
th_vars <- read.csv(paste0(vars_path,"th_vars.csv")) %>% pull(x)
```

# Load centiles and env/geodata tables
```{r}
centiles_distinct_path <- "/mnt/arcus/lab/users/kafadare/slip_centiles_processed_tables/centiles_distinct_2026-03-18.csv"
centiles_distinct <- read.csv(centiles_distinct_path)

centiles_distinct$participant_id_modified <- sub("sub-", "", centiles_distinct$participant_id)

centiles_all_path <- "/mnt/arcus/lab/users/kafadare/slip_centiles_processed_tables/centiles_all_2026-03-18.csv"
centiles_all <- read.csv(centiles_all_path)
centiles_all$participant_id_modified <- sub("sub-", "", centiles_all$participant_id)

centiles_distinct <- centiles_distinct %>% filter(adjusted_age_in_years <= 21)
centiles_all <- centiles_all %>% filter(adjusted_age_in_years <= 21)

env <- readRDS("/mnt/arcus/lab/shared/env_data.rds")
```

# Set 0/1 centiles to min/max value to avoid Inf/-Inf
numbers from gamlssTools pkg
```{r}
complete_centile_vars <- c(full_vars, th_vars_median)

for (x in complete_centile_vars){
  centiles_distinct[centiles_distinct[,x] == 0 & !is.na(centiles_distinct[,x]), x] <- 0.0000000000000000000000001
  centiles_distinct[centiles_distinct[,x] == 1 & !is.na(centiles_distinct[,x]), x] <- 0.99999999999999994
}

for (x in complete_centile_vars){
  centiles_all[centiles_all[,x] == 0 & !is.na(centiles_all[,x]), x] <- 0.0000000000000000000000001
  centiles_all[centiles_all[,x] == 1 & !is.na(centiles_all[,x]), x] <- 0.99999999999999994
}
```

# Explore the env/geodata table
```{r}
names(env)
env <- merge(env, participants %>% select(pat_id, dob, dob_year), by = "pat_id")
#rm(participants)

names(env)
hist(env$ADI)
head(env$year)
#Calculate unadjusted age in days using dob and end date for a particular entry
env$unadjusted_age_in_days_at_end_date <- as.numeric(difftime(env$end_date, as.Date(env$dob), units = "days"))
env$unadjusted_age_in_days_at_start_date <- as.numeric(difftime(env$start_date, as.Date(env$dob), units = "days"))
```
# Make swyc_clean df with one row per patient-form pair
```{r}
swyc$pat_form <- paste0(swyc$pat_id, swyc$form_friendly_name)

dup_df <- swyc %>%
  group_by(pat_form) %>%
  filter(n() > 1)

dup_df_summary <- dup_df %>%
  group_by(pat_form) %>%
  summarise(
    across(c(n_answers, effective_age, age_mo, norm_devo_quotient),
            ~ sd(.x, na.rm = TRUE),
           .names = "sd_{.col}"),
    .groups = "drop"
  )

median(dup_df_summary$sd_n_answers)
median(dup_df_summary$sd_effective_age)
median(dup_df_summary$sd_age_mo)
median(dup_df_summary$sd_norm_devo_quotient)

max(dup_df_summary$sd_n_answers)
max(dup_df_summary$sd_effective_age)
max(dup_df_summary$sd_age_mo)
max(dup_df_summary$sd_norm_devo_quotient)

swyc_clean <- swyc %>%
  group_by(pat_form) %>%
  slice_max(order_by = n_answers, with_ties = TRUE) %>%
  slice_max(order_by = effective_age, with_ties = TRUE) %>%
  distinct(pat_form, .keep_all = TRUE)
```


Ideas/what to do:
For SWYC-based analyses
- Take median up to date/age of SWYC
- Use median SWYC at first pass
- Look at association between env ~ swyc
- See if including env in regression between perinatal ~ swyc changes anything

For  Brain Centile-based analyses
- Take median up to date/age of scan
- Look at association between env ~ centile
- See if including env in regression between perinatal ~ centile changes anything

# Make a birth factors table from arcus.participants
```{r}
bf <- participants %>% select(pat_id, birth_weight_kg, birth_length_cm, gestational_age, gestational_age_num, sex, race, ethnicity)
# convert the string format to days
# Convert different gestational_age_string formats and convert to both 1) number of days AND 2) number of weeks (rounded up/down) AND 3) number of COMPLETED weeks to the best of our knowledge
data <- bf$gestational_age

data <- sub("\\s", "", data)
complete_weeks <- as.numeric(sub("^(.{2}).*", "\\1", data))
indices_plus <- grep("\\+", data)
indices_fraction <- grep("/", data)
indices_decimal <- grep("\\.", data)
days <- rep(NA, length(data))
days[!is.na(data)] <- 3 #add three days by default if no other information is encoded for days
days[indices_fraction] <- as.numeric(sub("^.*?(\\d{1}/\\d{1})$", "\\1", data[indices_fraction]) %>% sub("^([0-9]{1}).*", "\\1", .))
days[indices_plus] <- 3 # add three days if gestational age is entered as [week number]+ (like 40+ of 38+)
days[indices_decimal] <- round(as.numeric(sub("^.*?(\\.\\d+)$", "\\1", data[indices_decimal]))*7)

days_total <- complete_weeks*7 + days
weeks_rounded <- round(days_total/7)

bf$ga_complete_weeks <- complete_weeks
bf$ga_days_total <- days_total
bf$ga_weeks_rounded <- weeks_rounded

# QC BW and GA

# GA: remove anyone encoded as > 44*7 ga days or < 22 weeks (there is a single 43.5 weeker, I don't really want to remove this entry)
# BW: remove anyone with birthweight recorded as > 10kg and < 0.3kg
bf$ga_days_total_qc <- ifelse(bf$ga_days_total < 22*7 | bf$ga_days_total >= 44*7, NA, bf$ga_days_total)
sum(!is.na(bf$ga_days_total) & is.na(bf$ga_days_total_qc)) # 21 removed

bf$ga_num_qc <- ifelse(bf$gestational_age_num < 22 | bf$gestational_age_num >= 44, NA, bf$gestational_age_num)
sum(!is.na(bf$gestational_age_num) & is.na(bf$ga_num_qc)) # 29 removed

bf$birth_weight_kg_qc <- ifelse(bf$birth_weight_kg < 0.3 | bf$birth_weight_kg > 10, NA, bf$birth_weight_kg)
sum(!is.na(bf$birth_weight_kg) & is.na(bf$birth_weight_kg_qc)) # 86 removed
```

# Env with SWYC merging

## Merge env-swyc one patient-form pair per swyc, take age at given form
Merge env with swyc and choose rows with closest swyc-effective-age (-) env-start-date, where swyc-effective-age is after env-start-date
```{r}
names(swyc_clean)
table(swyc_clean$form_friendly_name)


swyc_env_allforms <- merge(swyc_clean %>% select(pat_id, effective_age, form_friendly_name, meets_milestones, norm_devo_quotient, norm_devo_prob, pat_form), env, by = "pat_id")
swyc_env_allforms$swyc_end_date_age_diff <- swyc_env_allforms$effective_age - swyc_env_allforms$unadjusted_age_in_days_at_end_date
swyc_env_allforms$swyc_start_date_age_diff <- swyc_env_allforms$effective_age - swyc_env_allforms$unadjusted_age_in_days_at_start_date

env_vars_of_int <- c("ADI3_EDU", "ADI3_ECON", "ADI3_FINS", "ADI", "Pollution", "Retirement", "University", "SES", "Urban_Core", names(env)[grepl("z_|r_", names(env))])

swyc_env_allforms_median <- swyc_env_allforms %>%
  group_by(pat_id, form_friendly_name) %>%
  mutate(env_limit_row = min(swyc_start_date_age_diff[swyc_start_date_age_diff >= 0])) %>%
  filter(swyc_start_date_age_diff >= env_limit_row) %>%
  summarise(
    n_rows = n(),
    norm_devo_quotient = first(norm_devo_quotient),
    effective_age_swyc = first(effective_age),
    swyc_start_date_age_diff   = first(swyc_start_date_age_diff),
    swyc_end_date_age_diff   = first(swyc_end_date_age_diff),
    dob = first(dob),
    dob_year = first(dob_year),
    pat_form = first(pat_form),
    across(any_of(env_vars_of_int),
           list(
             median = ~ median(.x, na.rm = TRUE),
             n_val = ~ sum(!is.na(.x))
           ),
           .names = "{col}.{fn}"),
    .groups = "drop"
  )

# sanity checks
dim(swyc_env_allforms)
dim(swyc_env_allforms_median)
length(unique(swyc_env_allforms_median$pat_form))
length(unique(swyc_env_allforms$pat_form))
length(unique(swyc_env_allforms_median$pat_id))
length(unique(swyc_env_allforms$pat_id))
```

## Merge swyc and env, take latest swyc date per patient, use median dq
use the date for the latest swyc within a patient to define the endpoint of env variables to take a relevant median
```{r}
env_vars_of_int <- c("ADI3_EDU", "ADI3_ECON", "ADI3_FINS", "ADI", "Pollution", "Retirement", "University", "SES", "Urban_Core", names(env)[grepl("z_|r_", names(env))])

swyc_avg <- swyc_clean %>%
  group_by(pat_id) %>%
  summarise(final_swyc_age = max(effective_age, na.rm = TRUE),
         median_dq = median(norm_devo_quotient, na.rm = TRUE), 
         mean_dq = mean(norm_devo_quotient, na.rm = TRUE))

hist(swyc_avg$final_swyc_age)

swyc_env_avg <- merge(swyc_avg, env, by = "pat_id")

swyc_env_avg$swyc_end_date_age_diff <- swyc_env_avg$final_swyc_age - swyc_env_avg$unadjusted_age_in_days_at_end_date
swyc_env_avg$swyc_start_date_age_diff <- swyc_env_avg$final_swyc_age - swyc_env_avg$unadjusted_age_in_days_at_start_date

swyc_env_avg_median <- swyc_env_avg %>%
  group_by(pat_id) %>%
  mutate(env_limit_row = min(swyc_start_date_age_diff[swyc_start_date_age_diff >= 0])) %>%
  filter(swyc_start_date_age_diff >= env_limit_row) %>%
  summarise(
    n_rows = n(),
    median_dq = first(median_dq),
    final_swyc_age = first(final_swyc_age),
    mean_dq   = first(mean_dq),
    swyc_start_date_age_diff   = first(swyc_start_date_age_diff),
    swyc_end_date_age_diff   = first(swyc_end_date_age_diff),
    dob = first(dob),
    dob_year = first(dob_year),
    across(any_of(env_vars_of_int),
           list(
             median = ~ median(.x, na.rm = TRUE),
             n_val = ~ sum(!is.na(.x))
           ),
           .names = "{col}.{fn}")
  ) %>% ungroup()

# sanity checks
dim(swyc_env_avg)
dim(swyc_env_avg_median)
length(unique(swyc_env_avg$pat_id))
length(unique(swyc_env_avg_median$pat_id))
```

## Add birth factor variables
```{r}
swyc_env_allforms_median <- merge(swyc_env_allforms_median, bf, 
              by = "pat_id")
swyc_env_avg_median <- merge(swyc_env_avg_median, bf, 
              by = "pat_id")
```

# Prepare dfs (z-score)
```{r}
# zscore centiles and other variables of interest
df.zscore_avg <- swyc_env_avg_median %>%
  mutate(across(any_of(c("ga_num_qc", "birth_weight_kg_qc", "ga_complete_weeks", "ga_days_total_qc", "ga_weeks_rounded",
                         "mean_dq", "median_dq", "ADI.median", "ADI3_ECON.median", "ADI3_FINS.median", "ADI3_EDU.median",
                         "Pollution.median", "Retirement.median", "University.median", "SES.median", "Urban_Core.median")), 
                scale))
# set sex and preterm as factor
df.zscore_avg$sex <- as.factor(df.zscore_avg$sex)

# zscore centiles and other variables of interest
df.zscore_allforms <- swyc_env_allforms_median %>%
  mutate(across(any_of(c("ga_num_qc", "birth_weight_kg_qc", "ga_complete_weeks", "ga_days_total_qc", "ga_weeks_rounded",
                         "norm_devo_quotient", "norm_devo_prob", "ADI.median", "ADI3_ECON.median", "ADI3_FINS.median", "ADI3_EDU.median",
                         "Pollution.median", "Retirement.median", "University.median", "SES.median", "Urban_Core.median")), 
                scale))
# set sex and preterm as factor
df.zscore_allforms$sex <- as.factor(df.zscore_allforms$sex)
```

# Look at data post-median/zscore
```{r}
sum(is.na(df$ADI.median))
max(df.zscore$ADI.n_val) #196
patid_max = df.zscore[df.zscore$ADI.n_val == max(df.zscore$ADI.n_val), "pat_id"]
env[env$pat_id == patid_max,]
# just lots of address entries with lots of same end/start dates or dates that are a few days apart
median(df.zscore$ADI.n_val) #6
hist(df.zscore$ADI.n_val, breaks = 100)
```


# Explore swyc ~ env associations

## define vars of interest
```{r}
# vars_exp <- c("ADI.median", "ADI3_ECON.median", "ADI3_FINS.median", "ADI3_EDU.median",
#                          "Pollution.median", "Retirement.median", "University.median", "SES.median", "Urban_Core.median",
#               names(df.zscore)[grep("^z_.*median$", names(df.zscore))])

vars_exp <- c("ADI.median", "ADI3_ECON.median", "ADI3_FINS.median", "ADI3_EDU.median")

```

## Swyc Median across all forms

### Correlations
```{r}
cor_matrix <- cor(df.zscore_avg %>% select(vars_exp, median_dq, birth_weight_kg_qc, ga_days_total_qc), use = "pairwise.complete.obs")
p_matrix <- cor_pmat(df.zscore_avg %>% select(vars_exp, median_dq, birth_weight_kg_qc, ga_days_total_qc), use = "pairwise.complete.obs")

ggcorrplot(
  cor_matrix,
  p.mat = p_matrix,
  type = "lower",
  insig = "pch",
  pch = 4,
  pch.cex = 2,
  sig.level = 0.05,
  lab = TRUE,
  digits = 2
)

# Grid of scatterplots
df.zscore_avg %>% select(vars_exp, median_dq) %>%
  pivot_longer(-median_dq) %>%
  ggplot(aes(x = value, y = median_dq)) +
  geom_point(alpha = 0.1) +
  theme_classic() +
  facet_wrap(~ name, scales = "free")

df.zscore_avg %>% select(vars_exp, birth_weight_kg_qc) %>%
  pivot_longer(-birth_weight_kg_qc) %>%
  ggplot(aes(x = value, y = birth_weight_kg_qc)) +
  geom_point(alpha = 0.1) +
   geom_density_2d(color = "red") +
  theme_classic() +
  facet_wrap(~ name, scales = "free")

df.zscore_avg %>% select(vars_exp, birth_weight_kg_qc) %>%
  pivot_longer(-birth_weight_kg_qc) %>%
  ggplot(aes(x = value, y = birth_weight_kg_qc)) +
  stat_density_2d(geom = "point", aes(size = after_stat(density)), n = 10, contour = FALSE) +
  theme_classic() +
  facet_wrap(~ name, scales = "free")

df.zscore_avg %>% select(vars_exp, ga_days_total_qc) %>%
  pivot_longer(-ga_days_total_qc) %>%
  ggplot(aes(x = value, y = ga_days_total_qc)) +
  geom_point(alpha = 0.1) +
  theme_classic() +
  facet_wrap(~ name, scales = "free")
```

### Partial R
Maybe?
Per Briana permutations for significance testing
```{r}
# Partial R of ADI
# compute partial r^2
model_red <- lm(median_dq ~ ga_days_total_qc, data = df.zscore_avg)
model_full <- lm(median_dq ~ ga_days_total_qc + ADI.median, data = df.zscore_avg)

r2_full <- summary(model_full)$r.squared
r2_red  <- summary(model_red)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(model_full)["ADI.median"])
partial_r <- sign_coef * sqrt(abs(partial_r2))

r2_full <- summary(model_full)$r.squared
r2_red  <- summary(model_red)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(model_full)["ADI.median"])
partial_r <- sign_coef * sqrt(abs(partial_r2))


# partial R of GA
# compute partial r^2
model_red <- lm(median_dq ~ ADI.median, data = df.zscore_avg)
model_full <- lm(median_dq ~ ADI.median + ga_days_total_qc, data = df.zscore_avg)

r2_full <- summary(model_full)$r.squared
r2_red  <- summary(model_red)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(model_full)["ga_days_total_qc"])
partial_r <- sign_coef * sqrt(abs(partial_r2))

r2_full <- summary(model_full)$r.squared
r2_red  <- summary(model_red)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(model_full)["ga_days_total_qc"])
partial_r <- sign_coef * sqrt(abs(partial_r2))
```

## Swyc by Form

### select patients with all forms (9,18,24,30) available
```{r}
df.zscore_9_24_pats <- df.zscore_allforms %>% 
  group_by(pat_id) %>%
  filter(
    any(form_friendly_name == "SWYC : 09 Month" & !is.na(norm_devo_quotient)) &
    any(form_friendly_name == "SWYC : 24 Month" & !is.na(norm_devo_quotient))
        ) %>% ungroup()
  
length(unique(df.zscore_9_24_pats$pat_id))
table(df.zscore_9_24_pats$form_friendly_name)

adi_swyc_9mo <- lm(swyc_9month ~  ADI.median, data = df.zscore_9_24_pats %>% 
  filter(form_friendly_name == "SWYC : 09 Month") %>% 
  rename(swyc_9month = norm_devo_quotient))
summary(adi_swyc_9mo)

adi_swyc_24mo <- lm(swyc_24month ~  ADI.median, data = df.zscore_9_24_pats %>% 
  filter(form_friendly_name == "SWYC : 24 Month") %>% 
  rename(swyc_24month = norm_devo_quotient))
summary(adi_swyc_24mo)

df.zscore_9_24_pats_diff <- df.zscore_9_24_pats %>%
  filter(form_friendly_name %in% c("SWYC : 09 Month", "SWYC : 24 Month")) %>%
  mutate(form_friendly_name = recode(form_friendly_name,
                             "SWYC : 09 Month" = "9mo",
                             "SWYC : 24 Month" = "24mo")) %>%
  select(pat_id, form_friendly_name , ADI.median, norm_devo_quotient) %>%
  tidyr::pivot_wider(names_from = form_friendly_name, values_from = c("ADI.median", 
                                                                      "norm_devo_quotient")) %>%
  mutate(adi_diff_24_minus_9 = `ADI.median_24mo` - `ADI.median_9mo`,
         dq_diff_24_minus_9 = `norm_devo_quotient_24mo` - `norm_devo_quotient_9mo`)

hist(df.zscore_9_24_pats_diff$adi_diff_24_minus_9, breaks = 30)

df_plot <- df.zscore_9_24_pats  %>%
  filter(form_friendly_name %in% c("SWYC : 09 Month", "SWYC : 24 Month")) %>%
  left_join(df.zscore_9_24_pats_diff %>% select(pat_id, adi_diff_24_minus_9, dq_diff_24_minus_9), 
            by = "pat_id") %>%
  mutate(adi_change = ifelse(adi_diff_24_minus_9 == 0, "same", 
  ifelse(adi_diff_24_minus_9 > 0, "higher", "lower")),
         dq_change = ifelse(dq_diff_24_minus_9 == 0, "same", 
  ifelse(dq_diff_24_minus_9 > 0, "higher", "lower")))
 
table(df_plot$adi_change)
table(df_plot$dq_change)
        
ggplot(df_plot,
       aes(x = form_friendly_name, y = ADI.median, group = pat_id, color = adi_change)) +
  geom_line(alpha = 0.3) +
  geom_point(size = 1) +
  theme_classic()

ggplot(df.zscore_9_24_pats_diff,
       aes(x = adi_diff_24_minus_9, y = dq_diff_24_minus_9, color = norm_devo_quotient_24mo)) +
  geom_point(size = 1) +
  theme_classic()
```

### See ADI distribution by form
```{r}
ggplot(df.zscore_allforms, aes(x = ADI.median)) +
  geom_histogram(bins = 30, fill = "#7FA6C9", color = "white") +
  facet_wrap(~form_friendly_name, scales = "free_y") +
  theme_classic()
```

### 24 months
#### Correlations
```{r}
cor_matrix <- cor(df.zscore_allforms %>% filter(form_friendly_name == "SWYC : 24 Month") %>% select(vars_exp, norm_devo_quotient, birth_weight_kg_qc, ga_days_total_qc) %>% rename(swyc_24month = norm_devo_quotient), use = "pairwise.complete.obs")
p_matrix <- cor_pmat(df.zscore_allforms %>% filter(form_friendly_name == "SWYC : 24 Month") %>% select(vars_exp, norm_devo_quotient, birth_weight_kg_qc, ga_days_total_qc) %>% rename(swyc_24month = norm_devo_quotient), use = "pairwise.complete.obs")

ggcorrplot(
  cor_matrix,
  p.mat = p_matrix,
  type = "lower",
  insig = "pch",
  pch = 4,
  pch.cex = 2,
  sig.level = 0.05,
  lab = TRUE,
  digits = 2
)

# Grid of scatterplots
df.zscore_allforms %>% 
  filter(form_friendly_name == "SWYC : 24 Month") %>% 
  select(vars_exp, norm_devo_quotient) %>% 
  rename(swyc_24month = norm_devo_quotient) %>%
  pivot_longer(-swyc_24month) %>%
  ggplot(aes(x = value, y = swyc_24month)) +
  #geom_point(alpha = 0.1) +
  stat_density_2d(geom = "point", aes(size = after_stat(density)), n = 20, contour = FALSE) +
  theme_classic() +
  facet_wrap(~ name, scales = "free")

df.zscore_allforms %>% 
  filter(form_friendly_name == "SWYC : 24 Month") %>% 
  rename(swyc_24month = norm_devo_quotient) %>%
  ggplot(aes(x = effective_age_swyc/30.4, y = ADI.median)) +
  geom_point(alpha = 0.1)

adi_swycage_24mo <- lm(ADI.median ~ effective_age_swyc, data = df.zscore_allforms %>% 
  filter(form_friendly_name == "SWYC : 24 Month"))
summary(adi_swycage_24mo)

adi_swyc_agecov_24mo <- lm(swyc_24month ~  ADI.median + effective_age_swyc, data = df.zscore_allforms %>% 
  filter(form_friendly_name == "SWYC : 24 Month") %>% 
  rename(swyc_24month = norm_devo_quotient))
summary(adi_swyc_agecov_24mo)
```

#### Partial R
Maybe?
Per Briana permutations for significance testing
```{r}
df = df.zscore_allforms %>% filter(form_friendly_name == "SWYC : 24 Month") %>% select(vars_exp, norm_devo_quotient, birth_weight_kg_qc, ga_days_total_qc) %>% rename(swyc_24month = norm_devo_quotient)
# Partial R of ADI
# compute partial r^2
model_red_24mo_adi <- lm(swyc_24month ~ ga_days_total_qc, data = df)
model_full_24mo_adi <- lm(swyc_24month ~ ga_days_total_qc + ADI.median, data = df)

r2_full <- summary(model_full_24mo_adi)$r.squared
r2_red  <- summary(model_red_24mo_adi)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(model_full_24mo_adi)["ADI.median"])
partial_r <- sign_coef * sqrt(abs(partial_r2))

r2_full <- summary(model_full_24mo_adi)$r.squared
r2_red  <- summary(model_red_24mo_adi)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(model_full_24mo_adi)["ADI.median"])
partial_r_adi <- sign_coef * sqrt(abs(partial_r2))


# partial R of GA
# compute partial r^2
model_red_24mo_ga <- lm(swyc_24month ~ ADI.median, data = df)
model_full_24mo_ga <- lm(swyc_24month ~ ADI.median + ga_days_total_qc, data = df)

r2_full <- summary(model_full_24mo_ga)$r.squared
r2_red  <- summary(model_red_24mo_ga)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(model_full_24mo_ga)["ga_days_total_qc"])
partial_r <- sign_coef * sqrt(abs(partial_r2))

r2_full <- summary(model_full_24mo_ga)$r.squared
r2_red  <- summary(model_red_24mo_ga)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(model_full_24mo_ga)["ga_days_total_qc"])
partial_r_24mo_ga <- sign_coef * sqrt(abs(partial_r2))
```


### 9 months
#### Correlations
```{r}
cor_matrix <- cor(df.zscore_allforms %>% filter(form_friendly_name == "SWYC : 09 Month") %>% select(vars_exp, norm_devo_quotient, birth_weight_kg_qc, ga_days_total_qc) %>% rename(swyc_9month = norm_devo_quotient), use = "pairwise.complete.obs")
p_matrix <- cor_pmat(df.zscore_allforms %>% filter(form_friendly_name == "SWYC : 09 Month") %>% select(vars_exp, norm_devo_quotient, birth_weight_kg_qc, ga_days_total_qc) %>% rename(swyc_9month = norm_devo_quotient), use = "pairwise.complete.obs")

ggcorrplot(
  cor_matrix,
  p.mat = p_matrix,
  type = "lower",
  insig = "pch",
  pch = 4,
  pch.cex = 2,
  sig.level = 0.05,
  lab = TRUE,
  digits = 2
)

# Grid of scatterplots
df.zscore_allforms %>% 
  filter(form_friendly_name == "SWYC : 09 Month") %>% 
  select(vars_exp, norm_devo_quotient) %>% 
  rename(swyc_9month = norm_devo_quotient) %>%
  pivot_longer(-swyc_9month) %>%
  ggplot(aes(x = value, y = swyc_9month)) +
  #geom_point(alpha = 0.1) +
  stat_density_2d(geom = "point", aes(size = after_stat(density)), n = 20, contour = FALSE) +
  theme_classic() +
  facet_wrap(~ name, scales = "free")


df.zscore_allforms %>% 
  filter(form_friendly_name == "SWYC : 09 Month") %>% 
  rename(swyc_9month = norm_devo_quotient) %>%
  ggplot(aes(x = effective_age_swyc/30.4, y = ADI.median)) +
  geom_point(alpha = 0.1)

adi_swycage_9mo <- lm(ADI.median ~ effective_age_swyc, data = df.zscore_allforms %>% 
  filter(form_friendly_name == "SWYC : 09 Month"))
summary(adi_swycage_9mo)

adi_swyc_agecov_9mo <- lm(swyc_9month ~  ADI.median + effective_age_swyc, data = df.zscore_allforms %>% 
  filter(form_friendly_name == "SWYC : 09 Month") %>% 
  rename(swyc_9month = norm_devo_quotient))
summary(adi_swyc_agecov_9mo)
```

#### Partial R
Maybe?
Per Briana permutations for significance testing
```{r}
df = df.zscore_allforms %>% filter(form_friendly_name == "SWYC : 09 Month") %>% select(vars_exp, norm_devo_quotient, birth_weight_kg_qc, ga_days_total_qc) %>% rename(swyc_9month = norm_devo_quotient)
# Partial R of ADI
# compute partial r^2
model_red_9mo_adi <- lm(swyc_9month ~ ga_days_total_qc, data = df)
model_full_9mo_adi <- lm(swyc_9month ~ ga_days_total_qc + ADI.median, data = df) 

r2_full <- summary(model_full_9mo_adi)$r.squared
r2_red  <- summary(model_red_9mo_adi)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(model_full_9mo_adi)["ADI.median"])
partial_r_adi <- sign_coef * sqrt(abs(partial_r2))
partial_r_adi

# partial R of GA
# compute partial r^2
model_red_9mo_ga <- lm(swyc_9month ~ ADI.median, data = df)
model_full_9mo_ga <- lm(swyc_9month ~ ADI.median + ga_days_total_qc, data = df)

r2_full <- summary(model_full_9mo_ga)$r.squared
r2_red  <- summary(model_red_9mo_ga)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(model_full_9mo_ga)["ga_days_total_qc"])
partial_r_9mo_ga <- sign_coef * sqrt(abs(partial_r2))
partial_r_9mo_ga
```


### 18 months
#### Correlations
```{r}
cor_matrix <- cor(df.zscore_allforms %>% filter(form_friendly_name == "SWYC : 18 Month") %>% select(vars_exp, norm_devo_quotient, birth_weight_kg_qc, ga_days_total_qc) %>% rename(swyc_18month = norm_devo_quotient), use = "pairwise.complete.obs")
p_matrix <- cor_pmat(df.zscore_allforms %>% filter(form_friendly_name == "SWYC : 18 Month") %>% select(vars_exp, norm_devo_quotient, birth_weight_kg_qc, ga_days_total_qc) %>% rename(swyc_18month = norm_devo_quotient), use = "pairwise.complete.obs")

ggcorrplot(
  cor_matrix,
  p.mat = p_matrix,
  type = "lower",
  insig = "pch",
  pch = 4,
  pch.cex = 2,
  sig.level = 0.05,
  lab = TRUE,
  digits = 2
)

# Grid of scatterplots
df.zscore_allforms %>% 
  filter(form_friendly_name == "SWYC : 18 Month") %>% 
  select(vars_exp, norm_devo_quotient) %>% 
  rename(swyc_18month = norm_devo_quotient) %>%
  pivot_longer(-swyc_18month) %>%
  ggplot(aes(x = value, y = swyc_18month)) +
  #geom_point(alpha = 0.1) +
  stat_density_2d(geom = "point", aes(size = after_stat(density)), n = 20, contour = FALSE) +
  theme_classic() +
  facet_wrap(~ name, scales = "free")
```

#### Partial R
Maybe?
Per Briana permutations for significance testing
```{r}
df = df.zscore_allforms %>% filter(form_friendly_name == "SWYC : 18 Month") %>% select(vars_exp, norm_devo_quotient, birth_weight_kg_qc, ga_days_total_qc) %>% rename(swyc_18month = norm_devo_quotient)
# Partial R of ADI
# compute partial r^2
model_red_18mo_adi <- lm(swyc_18month ~ ga_days_total_qc, data = df)
model_full_18mo_adi <- lm(swyc_18month ~ ga_days_total_qc + ADI.median, data = df)

r2_full <- summary(model_full_18mo_adi)$r.squared
r2_red  <- summary(model_red_18mo_adi)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(model_full_18mo_adi)["ADI.median"])
partial_r <- sign_coef * sqrt(abs(partial_r2))

r2_full <- summary(model_full_18mo_adi)$r.squared
r2_red  <- summary(model_red_18mo_adi)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(model_full_18mo_adi)["ADI.median"])
partial_r_adi <- sign_coef * sqrt(abs(partial_r2))


# partial R of GA
# compute partial r^2
model_red_18mo_ga <- lm(swyc_18month ~ ADI.median, data = df)
model_full_18mo_ga <- lm(swyc_18month ~ ADI.median + ga_days_total_qc, data = df)

r2_full <- summary(model_full_18mo_ga)$r.squared
r2_red  <- summary(model_red_18mo_ga)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(model_full_18mo_ga)["ga_days_total_qc"])
partial_r <- sign_coef * sqrt(abs(partial_r2))

r2_full <- summary(model_full_18mo_ga)$r.squared
r2_red  <- summary(model_red_18mo_ga)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(model_full_18mo_ga)["ga_days_total_qc"])
partial_r_18mo_ga <- sign_coef * sqrt(abs(partial_r2))
```


### 15 months
#### Correlations
```{r}
cor_matrix <- cor(df.zscore_allforms %>% filter(form_friendly_name == "SWYC : 15 Month") %>% select(vars_exp, norm_devo_quotient, birth_weight_kg_qc, ga_days_total_qc) %>% rename(swyc_15month = norm_devo_quotient), use = "pairwise.complete.obs")
p_matrix <- cor_pmat(df.zscore_allforms %>% filter(form_friendly_name == "SWYC : 15 Month") %>% select(vars_exp, norm_devo_quotient, birth_weight_kg_qc, ga_days_total_qc) %>% rename(swyc_15month = norm_devo_quotient), use = "pairwise.complete.obs")

ggcorrplot(
  cor_matrix,
  p.mat = p_matrix,
  type = "lower",
  insig = "pch",
  pch = 4,
  pch.cex = 2,
  sig.level = 0.05,
  lab = TRUE,
  digits = 2
)

# Grid of scatterplots
df.zscore_allforms %>% 
  filter(form_friendly_name == "SWYC : 15 Month") %>% 
  select(vars_exp, norm_devo_quotient) %>% 
  rename(swyc_15month = norm_devo_quotient) %>%
  pivot_longer(-swyc_15month) %>%
  ggplot(aes(x = value, y = swyc_15month)) +
  #geom_point(alpha = 0.1) +
  stat_density_2d(geom = "point", aes(size = after_stat(density)), n = 20, contour = FALSE) +
  theme_classic() +
  facet_wrap(~ name, scales = "free")
```

#### Partial R
Maybe?
Per Briana permutations for significance testing
```{r}
df = df.zscore_allforms %>% filter(form_friendly_name == "SWYC : 15 Month") %>% select(vars_exp, norm_devo_quotient, birth_weight_kg_qc, ga_days_total_qc) %>% rename(swyc_15month = norm_devo_quotient)
# Partial R of ADI
# compute partial r^2
model_red_15mo_adi <- lm(swyc_15month ~ ga_days_total_qc, data = df)
model_full_15mo_adi <- lm(swyc_15month ~ ga_days_total_qc + ADI.median, data = df)

r2_full <- summary(model_full_15mo_adi)$r.squared
r2_red  <- summary(model_red_15mo_adi)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(model_full_15mo_adi)["ADI.median"])
partial_r <- sign_coef * sqrt(abs(partial_r2))

r2_full <- summary(model_full_15mo_adi)$r.squared
r2_red  <- summary(model_red_15mo_adi)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(model_full_15mo_adi)["ADI.median"])
partial_r_adi <- sign_coef * sqrt(abs(partial_r2))


# partial R of GA
# compute partial r^2
model_red_15mo_ga <- lm(swyc_15month ~ ADI.median, data = df)
model_full_15mo_ga <- lm(swyc_15month ~ ADI.median + ga_days_total_qc, data = df)

r2_full <- summary(model_full_15mo_ga)$r.squared
r2_red  <- summary(model_red_15mo_ga)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(model_full_15mo_ga)["ga_days_total_qc"])
partial_r <- sign_coef * sqrt(abs(partial_r2))

r2_full <- summary(model_full_15mo_ga)$r.squared
r2_red  <- summary(model_red_15mo_ga)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(model_full_15mo_ga)["ga_days_total_qc"])
partial_r_15mo_ga <- sign_coef * sqrt(abs(partial_r2))
```



### 30 months
#### Correlations
```{r}
cor_matrix <- cor(df.zscore_allforms %>% filter(form_friendly_name == "SWYC : 30 Month") %>% select(vars_exp, norm_devo_quotient, birth_weight_kg_qc, ga_days_total_qc) %>% rename(swyc_30month = norm_devo_quotient), use = "pairwise.complete.obs")
p_matrix <- cor_pmat(df.zscore_allforms %>% filter(form_friendly_name == "SWYC : 30 Month") %>% select(vars_exp, norm_devo_quotient, birth_weight_kg_qc, ga_days_total_qc) %>% rename(swyc_30month = norm_devo_quotient), use = "pairwise.complete.obs")

ggcorrplot(
  cor_matrix,
  p.mat = p_matrix,
  type = "lower",
  insig = "pch",
  pch = 4,
  pch.cex = 2,
  sig.level = 0.05,
  lab = TRUE,
  digits = 2
)

# Grid of scatterplots
df.zscore_allforms %>% 
  filter(form_friendly_name == "SWYC : 30 Month") %>% 
  select(vars_exp, norm_devo_quotient) %>% 
  rename(swyc_30month = norm_devo_quotient) %>%
  pivot_longer(-swyc_30month) %>%
  ggplot(aes(x = value, y = swyc_30month)) +
  #geom_point(alpha = 0.1) +
  stat_density_2d(geom = "point", aes(size = after_stat(density)), n = 20, contour = FALSE) +
  theme_classic() +
  facet_wrap(~ name, scales = "free")
```

#### Partial R
Maybe?
Per Briana permutations for significance testing
```{r}
df = df.zscore_allforms %>% filter(form_friendly_name == "SWYC : 30 Month") %>% select(vars_exp, norm_devo_quotient, birth_weight_kg_qc, ga_days_total_qc) %>% rename(swyc_30month = norm_devo_quotient)
# Partial R of ADI
# compute partial r^2
model_red_30mo_adi <- lm(swyc_30month ~ ga_days_total_qc, data = df)
model_full_30mo_adi <- lm(swyc_30month ~ ga_days_total_qc + ADI.median, data = df)

r2_full <- summary(model_full_30mo_adi)$r.squared
r2_red  <- summary(model_red_30mo_adi)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(model_full_30mo_adi)["ADI.median"])
partial_r <- sign_coef * sqrt(abs(partial_r2))

r2_full <- summary(model_full_30mo_adi)$r.squared
r2_red  <- summary(model_red_30mo_adi)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(model_full_30mo_adi)["ADI.median"])
partial_r_adi <- sign_coef * sqrt(abs(partial_r2))


# partial R of GA
# compute partial r^2
model_red_30mo_ga <- lm(swyc_30month ~ ADI.median, data = df)
model_full_30mo_ga <- lm(swyc_30month ~ ADI.median + ga_days_total_qc, data = df)

r2_full <- summary(model_full_30mo_ga)$r.squared
r2_red  <- summary(model_red_30mo_ga)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(model_full_30mo_ga)["ga_days_total_qc"])
partial_r <- sign_coef * sqrt(abs(partial_r2))

r2_full <- summary(model_full_30mo_ga)$r.squared
r2_red  <- summary(model_red_30mo_ga)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(model_full_30mo_ga)["ga_days_total_qc"])
partial_r_30mo_ga <- sign_coef * sqrt(abs(partial_r2))
```



# Birthweight vs GA (inc calc bwp)

## Partial R of bw vs ga, at 24 mo SWYC
```{r}
df = df.zscore_allforms %>% filter(form_friendly_name == "SWYC : 24 Month") %>% select(vars_exp, norm_devo_quotient, birth_weight_kg_qc, ga_days_total_qc) %>% rename(swyc_24month = norm_devo_quotient)
# Partial R of BW
# compute partial r^2
model_red_24mo_bw <- lm(swyc_24month ~ ga_days_total_qc, data = df)
model_full_24mo_bw <- lm(swyc_24month ~ ga_days_total_qc + birth_weight_kg_qc, data = df)

r2_full <- summary(model_full_24mo_bw)$r.squared
r2_red  <- summary(model_red_24mo_bw)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(model_full_24mo_bw)["birth_weight_kg_qc"])
partial_r_bw <- sign_coef * sqrt(abs(partial_r2))


# partial R of GA
# compute partial r^2
model_red_24mo_ga <- lm(swyc_24month ~ birth_weight_kg_qc, data = df)
model_full_24mo_ga <- lm(swyc_24month ~ birth_weight_kg_qc + ga_days_total_qc, data = df)

r2_full <- summary(model_full_24mo_ga)$r.squared
r2_red  <- summary(model_red_24mo_ga)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(model_full_24mo_ga)["ga_days_total_qc"])
partial_r_24mo_ga <- sign_coef * sqrt(abs(partial_r2))
```


## Calculate bwp
```{r}

```

# Env with centiles

## Merge env with centiles
choose rows with closest age_at_scan (-) env-start-date, where age_at_scan is after env-start-date (+ diff)
```{r}
env_vars_of_int <- c("ADI3_EDU", "ADI3_ECON", "ADI3_FINS", "ADI", "Pollution", "Retirement", "University", "SES", "Urban_Core", names(env)[grepl("z_|r_", names(env))])

cent_env <- merge(centiles_distinct %>% mutate(pat_id = sub("sub-", "", participant_id)), env, by = "pat_id")

cent_env$scan_end_date_age_diff <- cent_env$age_at_scan - cent_env$unadjusted_age_in_days_at_end_date
cent_env$scan_start_date_age_diff <- cent_env$age_at_scan - cent_env$unadjusted_age_in_days_at_start_date

hist(cent_env$scan_start_date_age_diff)
hist(cent_env$scan_end_date_age_diff)

cent_env_median <- cent_env %>%
  group_by(pat_id) %>%
  mutate(env_limit_row = min(scan_start_date_age_diff[scan_start_date_age_diff >= 0])) %>%
  filter(scan_start_date_age_diff >= env_limit_row) %>%
  summarise(
    n_rows = n(),
    across(all_of(names(centiles_distinct)),
                  ~ first(.x),
                  .names = "{.col}"),
    across(any_of(env_vars_of_int),
           list(
             median = ~ median(.x, na.rm = TRUE),
             n_val = ~ sum(!is.na(.x))
           ),
           .names = "{col}.{fn}")
  )

dim(centiles_distinct)
dim(cent_env)
dim(cent_env_median)

length(unique(centiles_distinct$pat_id))
length(unique(cent_env$pat_id))
length(unique(cent_env_median$pat_id))
```

## Add birth factor variables from arcus.patients
```{r}
cent_env_median <- merge(cent_env_median, participants %>% select(pat_id, gestational_age, gestational_age_num, race, ethnicity) %>% rename(gestational_age_string = gestational_age), 
              by = "pat_id")

# convert the string format to days
# Convert different gestational_age_string formats and convert to both 1) number of days AND 2) number of weeks (rounded up/down) AND 3) number of COMPLETED weeks to the best of our knowledge
data <- cent_env_median$gestational_age_string

data <- sub("\\s", "", data)
complete_weeks <- as.numeric(sub("^(.{2}).*", "\\1", data))
indices_plus <- grep("\\+", data)
indices_fraction <- grep("/", data)
indices_decimal <- grep("\\.", data)
days <- rep(NA, length(data))
days[!is.na(data)] <- 3 #add three days by default if no other information is encoded for days
days[indices_fraction] <- as.numeric(sub("^.*?(\\d{1}/\\d{1})$", "\\1", data[indices_fraction]) %>% sub("^([0-9]{1}).*", "\\1", .))
days[indices_plus] <- 3 # add three days if gestational age is entered as [week number]+ (like 40+ of 38+)
days[indices_decimal] <- round(as.numeric(sub("^.*?(\\.\\d+)$", "\\1", data[indices_decimal]))*7)

days_total <- complete_weeks*7 + days
weeks_rounded <- round(days_total/7)

cent_env_median$ga_complete_weeks <- complete_weeks
cent_env_median$ga_days_total <- days_total
cent_env_median$ga_weeks_rounded <- weeks_rounded

# QC BW and GA

# GA: remove anyone encoded as > 44*7 ga days or < 22 weeks (there is a single 43.5 weeker, I don't really want to remove this entry)
# BW: remove anyone with birthweight recorded as > 10kg and < 0.3kg
cent_env_median$ga_days_total_qc <- ifelse(cent_env_median$ga_days_total < 22*7 | cent_env_median$ga_days_total >= 44*7, NA, cent_env_median$ga_days_total)
sum(!is.na(cent_env_median$ga_days_total) & is.na(cent_env_median$ga_days_total_qc)) # 0 removed

cent_env_median$ga_num_qc <- ifelse(cent_env_median$gestational_age_num < 22 | cent_env_median$gestational_age_num >= 44, NA, cent_env_median$gestational_age_num)
sum(!is.na(cent_env_median$gestational_age_num) & is.na(cent_env_median$ga_num_qc)) # 3 removed

cent_env_median$birth_weight_kg_qc <- ifelse(cent_env_median$birth_weight_kg < 0.3 | cent_env_median$birth_weight_kg > 10, NA, cent_env_median$birth_weight_kg)
sum(!is.na(cent_env_median$birth_weight_kg) & is.na(cent_env_median$birth_weight_kg_qc)) # 0 removed (already QC'ed bw coming from centiles_distinct)
```

## Add swyc median dq
```{r}
cent_env_median <- merge(cent_env_median, swyc_env_avg_median %>% select(pat_id, median_dq, final_swyc_age), 
              by = "pat_id", all.x = TRUE, all.y = FALSE)
```

## Prepare dfs (z-score)
```{r}
# zscore centiles and other variables of interest
df.zscore_cent <- cent_env_median %>%
  mutate(across(all_of(c(complete_centile_vars, "bweight_percentile")), qnorm))  %>%
  mutate(across(any_of(c("ga_num_qc", "birth_weight_kg_qc", "ga_complete_weeks", "ga_days_total_qc", "ga_weeks_rounded", "gestational_age",
                        "median_dq",  "ADI.median", "ADI3_ECON.median", "ADI3_FINS.median", "ADI3_EDU.median",
                         "Pollution.median", "Retirement.median", "University.median", "SES.median", "Urban_Core.median")), 
                scale))
# set sex and preterm as factor
df.zscore_cent$sex <- as.factor(df.zscore_cent$sex)
df.zscore_cent$preterm <- as.factor(df.zscore_cent$preterm)
```

## Look at data post-median/zscore
```{r}
sum(is.na(df.zscore_cent$ADI.median))
max(df.zscore_cent$ADI.n_val) #252
patid_max = df.zscore_cent[df.zscore_cent$ADI.n_val == max(df.zscore_cent$ADI.n_val), "pat_id"]
env[env$pat_id == patid_max,]
# just lots of address entries with lots of same end/start dates or dates that are a few days apart
median(df.zscore_cent$ADI.n_val) #13
hist(df.zscore_cent$ADI.n_val, breaks = 100)
```


## Explore relationships
### define vars of interest
```{r}
# vars_exp <- c("ADI.median", "ADI3_ECON.median", "ADI3_FINS.median", "ADI3_EDU.median",
#                          "Pollution.median", "Retirement.median", "University.median", "SES.median", "Urban_Core.median",
#               names(df.zscore_cent)[grep("^z_.*median$", names(df.zscore_cent))])

vars_exp <- c("ADI.median", "ADI3_ECON.median", "ADI3_FINS.median", "ADI3_EDU.median")

global_vars_totest <- c("global_eTIV", "global_SubCortGrayVol","global_CortexVol", "global_CerebralWhiteMatterVol", "global_VentricleChoroidVol","global_SurfaceArea", "global_MeanThickness")
```

### Correlations
```{r}
cor_matrix <- cor(df.zscore_cent %>% select(vars_exp, global_vars_totest, birth_weight_kg_qc, ga_days_total_qc, median_dq), use = "pairwise.complete.obs")
p_matrix <- cor_pmat(df.zscore_cent %>% select(vars_exp, global_vars_totest, birth_weight_kg_qc, ga_days_total_qc, median_dq), use = "pairwise.complete.obs")

ggcorrplot(
  cor_matrix,
  p.mat = p_matrix,
  type = "lower",
  insig = "blank",
  sig.level = 0.05,
  lab = TRUE,
  digits = 2
)

# Grid of scatterplots
df.zscore_cent %>% select(vars_exp, global_eTIV) %>%
  pivot_longer(-global_eTIV) %>%
  ggplot(aes(x = value, y = global_eTIV)) +
  geom_point(alpha = 0.1) +
  theme_classic() +
  facet_wrap(~ name, scales = "free")
```

### Partial R
Maybe?
Per Briana permutations for significance testing
#### GA
```{r}
# Partial R of ADI
# compute partial r^2
model_red_adi_ga <- lm(global_eTIV ~ ga_days_total_qc, data = df.zscore_cent)
model_full_adi_ga <- lm(global_eTIV ~ ga_days_total_qc + ADI.median, data = df.zscore_cent)

r2_full <- summary(model_full_adi_ga)$r.squared
r2_red  <- summary(model_red_adi_ga)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(model_full_adi_ga)["ADI.median"])
partial_r <- sign_coef * sqrt(abs(partial_r2))

r2_full <- summary(model_full_adi_ga)$r.squared
r2_red  <- summary(model_red_adi_ga)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(model_full_adi_ga)["ADI.median"])
partial_r_adi_ga <- sign_coef * sqrt(abs(partial_r2))

# compute partial r^2 of GA
model_red_ga <- lm(global_eTIV ~ ADI.median, data = df.zscore_cent)
model_full_ga <- lm(global_eTIV ~ ADI.median + ga_days_total_qc, data = df.zscore_cent)

r2_full <- summary(model_full_ga)$r.squared
r2_red  <- summary(model_red_ga)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(model_full_ga)["ga_days_total_qc"])
partial_r <- sign_coef * sqrt(abs(partial_r2))

r2_full <- summary(model_full_ga)$r.squared
r2_red  <- summary(model_red_ga)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(model_full_ga)["ga_days_total_qc"])
partial_r_ga <- sign_coef * sqrt(abs(partial_r2))
```


#### BW
```{r}
# Partial R of ADI
# compute partial r^2
model_red_adi_bw <- lm(global_eTIV ~ birth_weight_kg_qc, data = df.zscore_cent)
model_full_adi_bw <- lm(global_eTIV ~ birth_weight_kg_qc + ADI.median, data = df.zscore_cent)

r2_full <- summary(model_full_adi_bw)$r.squared
r2_red  <- summary(model_red_adi_bw)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(model_full_adi_bw)["ADI.median"])
partial_r <- sign_coef * sqrt(abs(partial_r2))

r2_full <- summary(model_full_adi_bw)$r.squared
r2_red  <- summary(model_red_adi_bw)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(model_full_adi_bw)["ADI.median"])
partial_r_adi_bw <- sign_coef * sqrt(abs(partial_r2)) 

# compute partial r^2 of bw
model_red_bw <- lm(global_eTIV ~ ADI.median, data = df.zscore_cent)
model_full_bw <- lm(global_eTIV ~ ADI.median + birth_weight_kg_qc, data = df.zscore_cent)

r2_full <- summary(model_full_bw)$r.squared
r2_red  <- summary(model_red_bw)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(model_full_bw)["birth_weight_kg_qc"])
partial_r <- sign_coef * sqrt(abs(partial_r2))

r2_full <- summary(model_full)$r.squared
r2_red  <- summary(model_red_bw)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(model_full_bw)["birth_weight_kg_qc"])
partial_r_bw <- sign_coef * sqrt(abs(partial_r2))
```

## Regressions including ADI as covariate

### Global
#### Define global vars of interest
```{r, echo = TRUE}
global_vars_totest <- c("global_eTIV", "global_SubCortGrayVol","global_CortexVol", "global_CerebralWhiteMatterVol", "global_VentricleChoroidVol","global_SurfaceArea",
                        "global_MeanThickness", "aseg_CSF")
```

#### GA
```{r, echo = FALSE}
sum(!is.na(df.zscore_cent$gestational_age) & !is.na(df.zscore_cent$ADI.median))
ga.adi_reg_table_global <- as.data.frame(t(sapply(df.zscore_cent[, which(names(df.zscore_cent) %in% global_vars_totest)], function(x) {
  lm <- lm(x ~ df.zscore_cent$gestational_age + df.zscore_cent$median_euler_mean + df.zscore_cent$ADI.median)
  summary <- summary(lm)
  c(ga_beta = summary$coefficients[2,1], ga_p = summary$coefficients[2,4], ga_se = summary$coefficients[2,2])
})))

ga.adi_reg_table_global$ga_p_adj <- p.adjust(ga.adi_reg_table_global$ga_p, method = "BH")

ga.adi_reg_table_global %>% filter(ga_p_adj < 0.05)
range(ga.adi_reg_table_global %>% filter(ga_p_adj < 0.05) %>% select(ga_beta))
```

#### BW
```{r, echo = FALSE}
sum(!is.na(df.zscore_cent$birth_weight_kg_qc) & !is.na(df.zscore_cent$ADI.median))
bw.adi_reg_table_global <- as.data.frame(t(sapply(df.zscore_cent[, which(names(df.zscore_cent) %in% global_vars_totest)], function(x) {
  lm <- lm(x ~ df.zscore_cent$birth_weight_kg_qc + df.zscore_cent$median_euler_mean + df.zscore_cent$ADI.median)
  summary <- summary(lm)
  c(bw_beta = summary$coefficients[2,1], bw_p = summary$coefficients[2,4], bw_se = summary$coefficients[2,2])
})))

bw.adi_reg_table_global$bw_p_adj <- p.adjust(bw.adi_reg_table_global$bw_p, method = "BH")

bw.adi_reg_table_global %>% filter(bw_p_adj < 0.05)
range(bw.adi_reg_table_global %>% filter(bw_p_adj < 0.05) %>% select(bw_beta))
```

#### Visualize/Figure Panel
```{r, echo = FALSE}
#  Add phenotype column from rownames
ga.adi_reg_table_global$phenotype <- rownames(ga.adi_reg_table_global)
bw.adi_reg_table_global$phenotype <- rownames(bw.adi_reg_table_global)

#  Add model identifiers + rename to common column names
ga.adi_reg_table_global <- ga.adi_reg_table_global %>%
  mutate(model = "Gestational Age",
         beta = ga_beta,
         se   = ga_se,
         p_adj    = ga_p_adj) %>%
    mutate(sig = p_adj < 0.05,
          color = case_when(
            sig & beta > 0 ~ "#c87a7a",   #  muted red
            sig & beta < 0 ~ "#7293b0",   #  muted blue
            TRUE ~ "gray70"))

bw.adi_reg_table_global <- bw.adi_reg_table_global %>%
  mutate(model = "Birthweight",
         beta = bw_beta,
         se   = bw_se,
         p_adj    = bw_p_adj) %>%
    mutate(sig = p_adj < 0.05,
          color = case_when(
            sig & beta > 0 ~ "#c87a7a",   #  muted red
            sig & beta < 0 ~ "#7293b0",   #  muted blue
            TRUE ~ "gray70"))

#  Combine and arrange by specific order of global phenotypes
df <- bind_rows(ga.adi_reg_table_global, bw.adi_reg_table_global) %>%
  mutate(phenotype = factor(phenotype, levels = c("global_eTIV", "global_CortexVol", "global_CerebralWhiteMatterVol",
                                                  "global_SubCortGrayVol", "global_VentricleChoroidVol",
                                                  "global_SurfaceArea", "global_MeanThickness"))) %>% arrange(phenotype)

#  Labeling for phenotypes on plot
phenotype_labs <- c(
 global_eTIV = "Intracranial Volume",
 global_MeanThickness = "Mean Cortical Thickness",
 global_VentricleChoroidVol = "Ventricular Volume",
 global_SurfaceArea = "Total Surface Area",
 global_CortexVol = "Cortical Gray Matter Volume",
 global_CerebralWhiteMatterVol = "Cerebral White Matter Volume",
 global_SubCortGrayVol = "Subcortical Gray Matter Volume"
)

#  Plot
(global_betas_plot <- ggplot(df, aes(x = beta, y = phenotype)) +
    geom_point(aes(shape = model, color = color), size = 3) +
    geom_errorbarh(aes(xmin = beta - se, xmax = beta + se, color = color), height = 0.2) +
    scale_shape_manual(values = c("Gestational Age" = 16, "Birthweight" = 17),
                       name = "") +
    scale_color_identity() +
    scale_x_continuous(limits = c(-0.1, 0.3)) +
    scale_y_discrete(limits = rev, labels = phenotype_labs) +
    geom_vline(xintercept = 0, linetype = "dotted", color = "darkgray") +
    theme_minimal() +
    theme_bw(base_size = 16) +
    theme(
        panel.grid = element_blank(),
        legend.position = "bottom",
        legend.title = element_blank(),
        panel.border = element_blank(),
        axis.text.y  = element_text(color = "black"),
        axis.line.x  = element_line(color = "black"),
        axis.line.y  = element_blank(),
        legend.justification = c("left", "top")
    ) +
    xlab("Standardized Beta Estimate, with QC and ADI covar") +
    ylab(""))

if(save_figs == TRUE){
panel_save(global_betas_plot, "global-betas-gabw", "pdf", plot_output_folder)
}
```


### Regional
#### GA with qc and ADI median covariate

##### Volume

######  simple linear regressions

####### compute regressions
```{r, echo = FALSE}
# Regional -- including cortical and subcortical
ga.qc.adi_vol <- as.data.frame(t(sapply(df.zscore_cent[, which(names(df.zscore_cent) %in% vol_vars)], function(x) {
  lm <- lm(reformulate(c("gestational_age",  "median_euler_mean", "ADI.median"), response = "x"), data = df.zscore_cent)
  summary <- summary(lm)
  c(ga_beta = summary$coefficients[2,1], ga_p = summary$coefficients[2,4], ga_se = summary$coefficients[2,2])
})))

ga.qc.adi_vol$ga_p_adj <- p.adjust(ga.qc.adi_vol$ga_p, method = "BH")
print(paste0("Number of significant IDPs after FDR correction: ", dim(ga.qc.adi_vol %>% filter(ga_p_adj < 0.05))[1]))
```

####### add label column to match to ggseg
```{r, echo = FALSE}
cols_edit <- grepl("aseg(?!_CC)", rownames(ga.qc.adi_vol), perl = TRUE)

ga.qc.adi_vol$label <- rownames(ga.qc.adi_vol) %>% 
  sub("aparc_GrayVol_", "", .) %>% 
  sub("aseg_", "", .) %>%
  sub("X", "x", .) %>%
  sub("rd_Ventricle", "rd_ventricle", .) %>%     sub("th_Ventricle", "th_ventricle", .) %>% sub("Thalamus", "Thalamus-Proper", .)

ga.qc.adi_vol$label[cols_edit] <- gsub("_", "-", ga.qc.adi_vol$label[cols_edit])

ga.qc.adi_vol_sig_MC <- ga.qc.adi_vol %>% filter(ga_p_adj < 0.05)

print(paste0("Beta range of significant IDPs: ", range(ga.qc.adi_vol_sig_MC$ga_beta)))
```

####### save betas
```{r, echo = FALSE}
save_table_full <- ga.qc.adi_vol
name <- "ga.qc.adi_volume.csv"
# Save dkt cortical betas for spin test
if(save_tables == TRUE){
  output_folder <-dkt_folder
  save_table_full$hemisphere <- ifelse(grepl("lh", save_table_full$label), "L", 
                         ifelse(grepl("rh", save_table_full$label), "R", NA))
  save_table_full$label <- sub("^rh_|lh_", "", save_table_full$label)
  filename_all <- paste0(output_folder, name)
  write.csv(save_table_full, filename_all)
}
# Save full betas
if(save_tables == TRUE){
output_folder <- betas_folder
filename <- paste0(output_folder, name)
write.csv(save_table_full, filename)
}
```

####### visualization of betas on surface maps
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(ga.qc.adi_vol_sig_MC, ggseg::dk, by = "label")
plot_data_aseg <- merge(ga.qc.adi_vol_sig_MC, ggseg::aseg, by = "label")

# Regions not matched in atlases
(ga.qc.adi_vol_noplot <- ga.qc.adi_vol_sig_MC[!(ga.qc.adi_vol_sig_MC$label %in% plot_data_dkt$label) & !(ga.qc.adi_vol_sig_MC$label %in% plot_data_aseg$label),])

# plots
(ga.qc.adi_dkt_vol_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = ga_beta)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = limits) +
  theme_void()+
  labs(title = "GA, QC and ADI covar, Volume"))

(ga.qc.adi_aseg_vol_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = ga_beta)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = limits) +
  theme_void() +
  labs(title = "GA, QC and ADI covar, Volume"))

if(save_figs == TRUE){
panel_save(ga.qc.adi_dkt_vol_plot, "ga-qc-adi-dkt-vol", "pdf", plot_output_folder)
panel_save(ga.qc.adi_aseg_vol_plot , "ga-qc-adi-aseg-vol", "pdf", plot_output_folder)
}
```

######  lm with age interaction

####### compute regressions
```{r, echo = FALSE}
# Regional -- including cortical and subcortical
gabyage.qc.adi_vol <- as.data.frame(t(sapply(df.zscore_cent[, which(names(df.zscore_cent) %in% vol_vars)], function(x) {
  lm <- lm(reformulate(c("gestational_age", "adjusted_age_in_years", "median_euler_mean", "ADI.median", "gestational_age:adjusted_age_in_years"), response = "x"), data = df.zscore_cent)
  summary <- summary(lm)
  c(ga_beta = summary$coefficients[2,1], ga_p = summary$coefficients[2,4], ga_se = summary$coefficients[2,2],
   age_beta = summary$coefficients[3,1], age_p = summary$coefficients[3,4], age_se = summary$coefficients[3,2],
   gabyage_beta = summary$coefficients[6,1], gabyage_p = summary$coefficients[6,4], gabyage_se = summary$coefficients[6,2])
})))

gabyage.qc.adi_vol$ga_p_adj <- p.adjust(gabyage.qc.adi_vol$ga_p, method = "BH")
gabyage.qc.adi_vol$gabyage_p_adj <- p.adjust(gabyage.qc.adi_vol$gabyage_p, method = "BH")

print(paste0("Age Main significant results (shouldn't be any): ", dim(gabyage.qc.adi_vol %>% filter(age_p < 0.05))[1]))
print(paste0("Number of significant IDPs - NO correction: ", dim(gabyage.qc.adi_vol %>% filter(gabyage_p < 0.05))[1]))
print(paste0("Number of significant IDPs - AFTER FDR correction: ", dim(gabyage.qc.adi_vol %>% filter(gabyage_p_adj < 0.05))[1]))
```

####### add label column to match to ggseg
```{r, echo = FALSE}
cols_edit <- grepl("aseg(?!_CC)", rownames(gabyage.qc.adi_vol), perl = TRUE)

gabyage.qc.adi_vol$label <- rownames(gabyage.qc.adi_vol) %>% 
  sub("aparc_GrayVol_", "", .) %>% 
  sub("aseg_", "", .) %>%
  sub("X", "x", .) %>%
  sub("rd_Ventricle", "rd_ventricle", .) %>%     sub("th_Ventricle", "th_ventricle", .) %>% sub("Thalamus", "Thalamus-Proper", .)

gabyage.qc.adi_vol$label[cols_edit] <- gsub("_", "-", gabyage.qc.adi_vol$label[cols_edit])

gabyage.qc.adi_vol_sig_MC <- gabyage.qc.adi_vol %>% filter(gabyage_p_adj < 0.05)

print(paste0("Beta range of significant IDPs: ", range(gabyage.qc.adi_vol_sig_MC$gabyage_beta)))
```

####### save betas
```{r, echo = FALSE}
save_table_full <- gabyage.qc.adi_vol
name <- "gabyage.qc.adi_volume.csv"
# Save dkt cortical betas for spin test
if(save_tables == TRUE){
  output_folder <-dkt_folder
  save_table_full$hemisphere <- ifelse(grepl("lh", save_table_full$label), "L", 
                         ifelse(grepl("rh", save_table_full$label), "R", NA))
  save_table_full$label <- sub("^rh_|lh_", "", save_table_full$label)
  filename_all <- paste0(output_folder, name)
  write.csv(save_table_full, filename_all)
}
# Save full betas
if(save_tables == TRUE){
output_folder <- betas_folder
filename <- paste0(output_folder, name)
write.csv(save_table_full, filename)
}
```

####### visualize interaction betas on surface maps
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(gabyage.qc.adi_vol_sig_MC, ggseg::dk, by = "label")
plot_data_aseg <- merge(gabyage.qc.adi_vol_sig_MC, ggseg::aseg, by = "label")

# Regions not matched in atlases
(gabyage.qc.adi_vol_noplot <- gabyage.qc.adi_vol_sig_MC[!(gabyage.qc.adi_vol_sig_MC$label %in% plot_data_dkt$label) & !(gabyage.qc.adi_vol_sig_MC$label %in% plot_data_aseg$label),])

# plots
(gabyage.qc.adi_dkt_vol_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = gabyage_beta)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = limits) +
  theme_void()+
  labs(title = "GAxAge, QC and ADI covar, Volume"))

(gabyage.qc.adi_aseg_vol_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = gabyage_beta)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = limits) +
  theme_void() +
  labs(title = "GAxAge, QC and ADI covar, Volume"))

if(save_figs == TRUE){
panel_save(gabyage.qc.adi_dkt_vol_plot, "gabyage-qc-adi-dkt-vol", "pdf", plot_output_folder)
panel_save(gabyage.qc.adi_aseg_vol_plot , "gabyage-qc-adi-aseg-vol", "pdf", plot_output_folder)
}
```

##### Surface Area

######  simple linear regressions

####### compute regressions
```{r, echo = FALSE}
# Regional -- including cortical and subcortical
ga.qc.adi_sa <- as.data.frame(t(sapply(df.zscore_cent[, which(names(df.zscore_cent) %in% sa_vars)], function(x) {
  lm <- lm(reformulate(c("gestational_age", "median_euler_mean", "ADI.median"), response = "x"), data = df.zscore_cent)
  summary <- summary(lm)
  c(ga_beta = summary$coefficients[2,1], ga_p = summary$coefficients[2,4], ga_se = summary$coefficients[2,2])
})))

ga.qc.adi_sa$ga_p_adj <- p.adjust(ga.qc.adi_sa$ga_p, method = "BH")
print(paste0("Number of significant IDPs after FDR correction: ", dim(ga.qc.adi_sa %>% filter(ga_p_adj < 0.05))[1]))
```

####### add label column to match to ggseg
```{r, echo = FALSE}
ga.qc.adi_sa$label <- rownames(ga.qc.adi_sa) %>% 
  sub("aparc_SurfArea_", "", .)

ga.qc.adi_sa_sig_MC <- ga.qc.adi_sa %>% filter(ga_p_adj < 0.05)

print(paste0("Beta range of significant IDPs: ", range(ga.qc.adi_sa_sig_MC$ga_beta)))
```

####### save betas
```{r, echo = FALSE}
save_table_full <- ga.qc.adi_sa
name <- "ga.qc.adi_surfArea.csv"
# Save dkt cortical betas for spin test
if(save_tables == TRUE){
  output_folder <-dkt_folder
  save_table_full$hemisphere <- ifelse(grepl("lh", save_table_full$label), "L", 
                         ifelse(grepl("rh", save_table_full$label), "R", NA))
  save_table_full$label <- sub("^rh_|lh_", "", save_table_full$label)
  filename_all <- paste0(output_folder, name)
  write.csv(save_table_full, filename_all)
}
# Save full betas
if(save_tables == TRUE){
output_folder <- betas_folder
filename <- paste0(output_folder, name)
write.csv(save_table_full, filename)
}
```

####### visualization of betas on surface maps
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(ga.qc.adi_sa_sig_MC, ggseg::dk, by = "label")

# Regions not matched in atlases
(ga.qc.adi_sa_noplot <- ga.qc.adi_sa_sig_MC[!(ga.qc.adi_sa_sig_MC$label %in% plot_data_dkt$label) & !(ga.qc.adi_sa_sig_MC$label %in% plot_data_aseg$label),])

# plots
(ga.qc.adi_dkt_sa_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = ga_beta)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = limits) +
  theme_void()+
  labs(title = "GA, QC and ADI covar, Surface Area"))

if(save_figs == TRUE){
panel_save(ga.qc.adi_dkt_sa_plot, "ga-qc-adi-dkt-sa", "pdf", plot_output_folder)
}
```

######  lm with age interaction

####### compute regressions
```{r, echo = FALSE}
# Regional -- including cortical and subcortical
gabyage.qc.adi_sa <- as.data.frame(t(sapply(df.zscore_cent[, which(names(df.zscore_cent) %in% sa_vars)], function(x) {
  lm <- lm(reformulate(c("gestational_age", "adjusted_age_in_years", "median_euler_mean", "ADI.median", "gestational_age:adjusted_age_in_years"), response = "x"), data = df.zscore_cent)
  summary <- summary(lm)
  c(ga_beta = summary$coefficients[2,1], ga_p = summary$coefficients[2,4], ga_se = summary$coefficients[2,2],
   age_beta = summary$coefficients[3,1], age_p = summary$coefficients[3,4], age_se = summary$coefficients[3,2],
   gabyage_beta = summary$coefficients[6,1], gabyage_p = summary$coefficients[6,4], gabyage_se = summary$coefficients[6,2])
})))

gabyage.qc.adi_sa$ga_p_adj <- p.adjust(gabyage.qc.adi_sa$ga_p, method = "BH")
gabyage.qc.adi_sa$gabyage_p_adj <- p.adjust(gabyage.qc.adi_sa$gabyage_p, method = "BH")

print(paste0("Age Main significant results (shouldn't be any): ", dim(gabyage.qc.adi_sa %>% filter(age_p < 0.05))[1]))
print(paste0("Number of significant IDPs - NO correction: ", dim(gabyage.qc.adi_sa %>% filter(gabyage_p < 0.05))[1]))
print(paste0("Number of significant IDPs - AFTER FDR correction: ", dim(gabyage.qc.adi_sa %>% filter(gabyage_p_adj < 0.05))[1]))
```

####### add label column to match to ggseg
```{r, echo = FALSE}
gabyage.qc.adi_sa$label <- rownames(gabyage.qc.adi_sa) %>% 
  sub("aparc_SurfArea_", "", .)

gabyage.qc.adi_sa_sig_MC <- gabyage.qc.adi_sa %>% filter(gabyage_p_adj < 0.05)

print(paste0("Beta range of significant IDPs: ", range(gabyage.qc.adi_sa_sig_MC$ga_beta)))
```

####### save betas
```{r, echo = FALSE}
save_table_full <- gabyage.qc.adi_sa
name <- "gabyage.qc.adi_surfArea.csv"
# Save dkt cortical betas for spin test
if(save_tables == TRUE){
  output_folder <-dkt_folder
  save_table_full$hemisphere <- ifelse(grepl("lh", save_table_full$label), "L", 
                         ifelse(grepl("rh", save_table_full$label), "R", NA))
  save_table_full$label <- sub("^rh_|lh_", "", save_table_full$label)
  filename_all <- paste0(output_folder, name)
  write.csv(save_table_full, filename_all)
}
# Save full betas
if(save_tables == TRUE){
output_folder <- betas_folder
filename <- paste0(output_folder, name)
write.csv(save_table_full, filename)
}
```

####### visualize interaction betas on surface maps
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(gabyage.qc.adi_sa_sig_MC, ggseg::dk, by = "label")

# Regions not matched in atlases
(gabyage.qc_sa_noplot <- gabyage.qc.adi_sa_sig_MC[!(gabyage.qc.adi_sa_sig_MC$label %in% plot_data_dkt$label) & !(gabyage.qc.adi_sa_sig_MC$label %in% plot_data_aseg$label),])

# plots
(gabyage.qc_dkt_sa_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = gabyage_beta)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = limits) +
  theme_void()+
  labs(title = "GAxAge, QC and ADI covar, Surface Area"))

if(save_figs == TRUE){
panel_save(gabyage.qc_dkt_sa_plot, "gabyage-qc-adi-dkt-sa", "pdf", plot_output_folder)
}
```

##### Thickness

######  simple linear regressions

####### compute regressions
```{r, echo = FALSE}
sum(!is.na(df.zscore_cent$gestational_age) & !is.na(df.zscore_cent$ADI.median) & !is.na(df.zscore_cent$global_MeanThickness))
# Regional -- including cortical and subcortical
ga.qc.adi_th <- as.data.frame(t(sapply(df.zscore_cent[, which(names(df.zscore_cent) %in% th_vars)], function(x) {
  lm <- lm(reformulate(c("gestational_age", "euler_mean_mpr", "ADI.median"), response = "x"), data = df.zscore_cent)
  summary <- summary(lm)
  c(ga_beta = summary$coefficients[2,1], ga_p = summary$coefficients[2,4], ga_se = summary$coefficients[2,2])
})))

ga.qc.adi_th$ga_p_adj <- p.adjust(ga.qc.adi_th$ga_p, method = "BH")
print(paste0("Number of significant IDPs after FDR correction: ", dim(ga.qc.adi_th %>% filter(ga_p_adj < 0.05))[1]))
```

####### add label column to match to ggseg
```{r, echo = FALSE}
ga.qc.adi_th$label <- rownames(ga.qc.adi_th) %>% 
  sub("aparc_ThickAvg_", "", .)

ga.qc.adi_th_sig_MC <- ga.qc.adi_th %>% filter(ga_p_adj < 0.05)

print(paste0("Beta range of significant IDPs: ", range(ga.qc.adi_th_sig_MC$ga_beta)))
```

####### save betas
```{r, echo = FALSE}
save_table_full <- ga.qc.adi_th
name <- "ga.qc.adi_thick.csv"
# Save dkt cortical betas for spin test
if(save_tables == TRUE){
  output_folder <-dkt_folder
  save_table_full$hemisphere <- ifelse(grepl("lh", save_table_full$label), "L", 
                         ifelse(grepl("rh", save_table_full$label), "R", NA))
  save_table_full$label <- sub("^rh_|lh_", "", save_table_full$label)
  filename_all <- paste0(output_folder, name)
  write.csv(save_table_full, filename_all)
}
# Save full betas
if(save_tables == TRUE){
output_folder <- betas_folder
filename <- paste0(output_folder, name)
write.csv(save_table_full, filename)
}
```

####### visualization of betas on surface maps
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(ga.qc.adi_th_sig_MC, ggseg::dk, by = "label")

# Regions not matched in atlases
(ga.qc.adi_th_noplot <- ga.qc.adi_th_sig_MC[!(ga.qc.adi_th_sig_MC$label %in% plot_data_dkt$label) & !(ga.qc.adi_th_sig_MC$label %in% plot_data_aseg$label),])

# plots
(ga.qc.adi_dkt_th_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = ga_beta)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = limits) +
  theme_void()+
  labs(title = "GA, QC and ADI covar, Cortical Thickness"))

if(save_figs == TRUE){
panel_save(ga.qc.adi_dkt_th_plot, "ga-qc-adi-dkt-th", "pdf", plot_output_folder)
}
```

######  lm with age interaction

####### compute regressions
```{r, echo = FALSE}
# Regional -- including cortical and subcortical
gabyage.qc.adi_th <- as.data.frame(t(sapply(df.zscore_cent[, which(names(df.zscore_cent) %in% th_vars)], function(x) {
  lm <- lm(reformulate(c("gestational_age", "adjusted_age_in_years", "euler_mean_mpr", "ADI.median", "gestational_age:adjusted_age_in_years"), response = "x"), data = df.zscore_cent)
  summary <- summary(lm)
  c(ga_beta = summary$coefficients[2,1], ga_p = summary$coefficients[2,4], ga_se = summary$coefficients[2,2],
   age_beta = summary$coefficients[3,1], age_p = summary$coefficients[3,4], age_se = summary$coefficients[3,2],
   gabyage_beta = summary$coefficients[6,1], gabyage_p = summary$coefficients[6,4], gabyage_se = summary$coefficients[6,2])
})))

gabyage.qc.adi_th$ga_p_adj <- p.adjust(gabyage.qc.adi_th$ga_p, method = "BH")
gabyage.qc.adi_th$gabyage_p_adj <- p.adjust(gabyage.qc.adi_th$gabyage_p, method = "BH")

print(paste0("Age Main significant results (shouldn't be any): ", dim(gabyage.qc.adi_th %>% filter(age_p < 0.05))[1]))
print(paste0("Number of significant IDPs - NO correction: ", dim(gabyage.qc.adi_th %>% filter(gabyage_p < 0.05))[1]))
print(paste0("Number of significant IDPs - AFTER FDR correction: ", dim(gabyage.qc.adi_th %>% filter(gabyage_p_adj < 0.05))[1]))
```

####### add label column to match to ggseg
```{r, echo = FALSE}
gabyage.qc.adi_th$label <- rownames(gabyage.qc.adi_th) %>% 
  sub("aparc_ThickAvg_", "", .)
  
gabyage.qc.adi_th_sig_MC <- gabyage.qc.adi_th %>% filter(gabyage_p_adj < 0.05)

print(paste0("Beta range of significant IDPs: ", range(gabyage.qc.adi_th_sig_MC$gabyage_beta)))
```

####### save betas
```{r, echo = FALSE}
save_table_full <- gabyage.qc.adi_th
name <- "gabyage.qc.adi_thick.csv"
# Save dkt cortical betas for spin test
if(save_tables == TRUE){
  output_folder <-dkt_folder
  save_table_full$hemisphere <- ifelse(grepl("lh", save_table_full$label), "L", 
                         ifelse(grepl("rh", save_table_full$label), "R", NA))
  save_table_full$label <- sub("^rh_|lh_", "", save_table_full$label)
  filename_all <- paste0(output_folder, name)
  write.csv(save_table_full, filename_all)
}
# Save full betas
if(save_tables == TRUE){
output_folder <- betas_folder
filename <- paste0(output_folder, name)
write.csv(save_table_full, filename)
}
```

####### visualize interaction betas on surface maps
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(gabyage.qc.adi_th_sig_MC, ggseg::dk, by = "label")

# Regions not matched in atlases
(gabyage.qc.adi_th_noplot <- gabyage.qc.adi_th_sig_MC[!(gabyage.qc.adi_th_sig_MC$label %in% plot_data_dkt$label) & !(gabyage.qc.adi_th_sig_MC$label %in% plot_data_aseg$label),])

# plots
(gabyage.qc.adi_dkt_th_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = gabyage_beta)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = limits) +
  theme_void()+
  labs(title = "GAxAge, QC and ADI covar, Cortical Thickness"))

if(save_figs == TRUE){
panel_save(gabyage.qc.adi_dkt_th_plot, "gabyage-qc-adi-dkt-th", "pdf", plot_output_folder)
}
```

#### BW with qc and ADI covariate

##### Volume

######  simple linear regressions

####### compute regressions
```{r, echo = FALSE}
sum(!is.na(df.zscore_cent$birth_weight_kg_qc))
# Regional -- including cortical and subcortical
bw.qc.adi_vol <- as.data.frame(t(sapply(df.zscore_cent[, which(names(df.zscore_cent) %in% vol_vars)], function(x) {
  lm <- lm(reformulate(c("birth_weight_kg_qc", "median_euler_mean", "ADI.median"), response = "x"), data = df.zscore_cent)
  summary <- summary(lm)
  c(bw_beta = summary$coefficients[2,1], bw_p = summary$coefficients[2,4], bw_se = summary$coefficients[2,2])
})))

bw.qc.adi_vol$bw_p_adj <- p.adjust(bw.qc.adi_vol$bw_p, method = "BH")
print(paste0("Number of significant IDPs after FDR correction: ", dim(bw.qc.adi_vol %>% filter(bw_p_adj < 0.05))[1]))
```

####### add label column to match to ggseg
```{r, echo = FALSE}
cols_edit <- grepl("aseg(?!_CC)", rownames(bw.qc.adi_vol), perl = TRUE)

bw.qc.adi_vol$label <- rownames(bw.qc.adi_vol) %>% 
  sub("aparc_GrayVol_", "", .) %>% 
  sub("aseg_", "", .) %>%
  sub("X", "x", .) %>%
  sub("rd_Ventricle", "rd_ventricle", .) %>%     sub("th_Ventricle", "th_ventricle", .) %>% sub("Thalamus", "Thalamus-Proper", .)

bw.qc.adi_vol$label[cols_edit] <- gsub("_", "-", bw.qc.adi_vol$label[cols_edit])

bw.qc.adi_vol_sig_MC <- bw.qc.adi_vol %>% filter(bw_p_adj < 0.05)

print(paste0("Beta range of significant IDPs: ", range(bw.qc.adi_vol_sig_MC$bw_beta)))
```

####### save betas
```{r, echo = FALSE}
save_table_full <- bw.qc.adi_vol
name <- "bw.qc.adi_volume.csv"
# Save dkt cortical betas for spin test
if(save_tables == TRUE){
  output_folder <-dkt_folder
  save_table_full$hemisphere <- ifelse(grepl("lh", save_table_full$label), "L", 
                         ifelse(grepl("rh", save_table_full$label), "R", NA))
  save_table_full$label <- sub("^rh_|lh_", "", save_table_full$label)
  filename_all <- paste0(output_folder, name)
  write.csv(save_table_full, filename_all)
}
# Save full betas
if(save_tables == TRUE){
output_folder <- betas_folder
filename <- paste0(output_folder, name)
write.csv(save_table_full, filename)
}
```

####### visualization of betas on surface maps
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(bw.qc.adi_vol_sig_MC, ggseg::dk, by = "label")
plot_data_aseg <- merge(bw.qc.adi_vol_sig_MC, ggseg::aseg, by = "label")

# Regions not matched in atlases
(bw.qc.adi_vol_noplot <- bw.qc.adi_vol_sig_MC[!(bw.qc.adi_vol_sig_MC$label %in% plot_data_dkt$label) & !(bw.qc.adi_vol_sig_MC$label %in% plot_data_aseg$label),])

# plots
(bw.qc.adi_dkt_vol_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = bw_beta)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = limits) +
  theme_void()+
  labs(title = "BW, QC and ADI covar, Volume"))

(bw.qc.adi_aseg_vol_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = bw_beta)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = limits) +
  theme_void() +
  labs(title = "BW, QC and ADI covar, Volume"))

if(save_figs == TRUE){
panel_save(bw.qc.adi_dkt_vol_plot, "bw-qc-adi-dkt-vol", "pdf", plot_output_folder)
panel_save(bw.qc.adi_aseg_vol_plot , "bw-qc-adi-aseg-vol", "pdf", plot_output_folder)
}
```

##### Scatterplot for top betas
```{r, echo = FALSE}
if(dim(bw.qc.adi_vol_sig_MC)[1] > 0){
top_n = if(dim(bw.qc.adi_vol_sig_MC)[1] > 10) {10} else {dim(bw.qc.adi_vol_sig_MC)[1]}
top <- bw.qc.adi_vol_sig_MC[order(bw.qc.adi_vol_sig_MC$bw_beta, decreasing = TRUE)[1:top_n],]
scatterplots_bw.qc.adi_vol <- vector(mode = "list", length = top_n)
for(i in 1:top_n){
  pheno <- rownames(top)[i]
  title <- top$label[i]
  names(scatterplots_bw.qc.adi_vol[i]) <- pheno
  scatterplots_bw.qc.adi_vol[[i]] <- ggplot(df.zscore_cent, aes(x = birth_weight_kg_qc, y = !!sym(pheno))) +
    geom_point() +
    geom_smooth(method = "lm") +
    theme_minimal() +
    labs(x = "Gestational Age (z-scored)", y = "Centile (z-scored)", title = title)
}

nCol <- 2
nRow <- 2
breaks <- c(seq(from = 0, to = top_n , by = (nCol*nRow)), top_n) %>% unique(.)
for(k in 1:(length(breaks)-1)){
  start <- breaks[k]+1
  end <- breaks[k+1]
  do.call("grid.arrange", c(scatterplots_bw.qc.adi_vol[start:end], ncol=nCol))
}
}
```

####  lm with age interaction

##### compute regressions
```{r, echo = FALSE}
# Regional -- including cortical and subcortical
bwbyage.qc.adi_vol <- as.data.frame(t(sapply(df.zscore_cent[, which(names(df.zscore_cent) %in% vol_vars)], function(x) {
  lm <- lm(reformulate(c("birth_weight_kg_qc", "adjusted_age_in_years", "median_euler_mean", "ADI.median", "birth_weight_kg_qc:adjusted_age_in_years"), response = "x"), data = df.zscore_cent)
  summary <- summary(lm)
  c(bw_beta = summary$coefficients[2,1], bw_p = summary$coefficients[2,4], bw_se = summary$coefficients[2,2],
   age_beta = summary$coefficients[3,1], age_p = summary$coefficients[3,4], age_se = summary$coefficients[3,2],
   bwbyage_beta = summary$coefficients[6,1], bwbyage_p = summary$coefficients[6,4], bwbyage_se = summary$coefficients[6,2])
})))

bwbyage.qc.adi_vol$bw_p_adj <- p.adjust(bwbyage.qc.adi_vol$bw_p, method = "BH")
bwbyage.qc.adi_vol$bwbyage_p_adj <- p.adjust(bwbyage.qc.adi_vol$bwbyage_p, method = "BH")

print(paste0("Age Main significant results (shouldn't be any): ", dim(bwbyage.qc.adi_vol %>% filter(age_p < 0.05))[1]))
print(paste0("Number of significant IDPs - NO correction: ", dim(bwbyage.qc.adi_vol %>% filter(bwbyage_p < 0.05))[1]))
print(paste0("Number of significant IDPs - AFTER FDR correction: ", dim(bwbyage.qc.adi_vol %>% filter(bwbyage_p_adj < 0.05))[1]))
```

##### add label column to match to ggseg
```{r, echo = FALSE}
cols_edit <- grepl("aseg(?!_CC)", rownames(bwbyage.qc.adi_vol), perl = TRUE)

bwbyage.qc.adi_vol$label <- rownames(bwbyage.qc.adi_vol) %>% 
  sub("aparc_GrayVol_", "", .) %>% 
  sub("aseg_", "", .) %>%
  sub("X", "x", .) %>%
  sub("rd_Ventricle", "rd_ventricle", .) %>%     sub("th_Ventricle", "th_ventricle", .) %>% sub("Thalamus", "Thalamus-Proper", .)

bwbyage.qc.adi_vol$label[cols_edit] <- gsub("_", "-", bwbyage.qc.adi_vol$label[cols_edit])

bwbyage.qc.adi_vol_sig_MC <- bwbyage.qc.adi_vol %>% filter(bwbyage_p_adj < 0.05)

print(paste0("Beta range of significant IDPs: ", range(bwbyage.qc.adi_vol_sig_MC$bwbyage_beta)))
```

##### save betas
```{r, echo = FALSE}
save_table_full <- bwbyage.qc.adi_vol
name <- "bwbyage.qc.adi_volume.csv"
# Save dkt cortical betas for spin test
if(save_tables == TRUE){
  output_folder <-dkt_folder
  save_table_full$hemisphere <- ifelse(grepl("lh", save_table_full$label), "L", 
                         ifelse(grepl("rh", save_table_full$label), "R", NA))
  save_table_full$label <- sub("^rh_|lh_", "", save_table_full$label)
  filename_all <- paste0(output_folder, name)
  write.csv(save_table_full, filename_all)
}
# Save full betas
if(save_tables == TRUE){
output_folder <- betas_folder
filename <- paste0(output_folder, name)
write.csv(save_table_full, filename)
}
```

##### visualize interaction betas on surface maps
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(bwbyage.qc.adi_vol_sig_MC, ggseg::dk, by = "label")
plot_data_aseg <- merge(bwbyage.qc.adi_vol_sig_MC, ggseg::aseg, by = "label")

# Regions not matched in atlases
(bwbyage.qc.adi_vol_noplot <- bwbyage.qc.adi_vol_sig_MC[!(bwbyage.qc.adi_vol_sig_MC$label %in% plot_data_dkt$label) & !(bwbyage.qc.adi_vol_sig_MC$label %in% plot_data_aseg$label),])

# plots
(bwbyage.qc.adi_dkt_vol_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = bwbyage_beta)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = limits) +
  theme_void()+
  labs(title = "BWxAge, QC and ADI covar, Volume"))

(bwbyage.qc.adi_aseg_vol_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = bwbyage_beta)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = limits) +
  theme_void() +
  labs(title = "BWxAge, QC and ADI covar, Volume"))

if(save_figs == TRUE){
panel_save(bwbyage.qc.adi_dkt_vol_plot, "bwbyage-qc-adi-dkt-vol", "pdf", plot_output_folder)
panel_save(bwbyage.qc.adi_aseg_vol_plot , "bwbyage-qc-adi-aseg-vol", "pdf", plot_output_folder)
}
```

##### visualize interaction plots
```{r, echo = FALSE}
if (dim(bwbyage.qc_vol_sig_MC)[1] > 0){
vars <- rownames(bwbyage.qc_vol_sig_MC[order(-(bwbyage.qc_vol_sig_MC$bwbyage_beta)),])
interaction_plots <-  vector(mode = "list", length = length(vars))
if (length(vars) > 0) {
for (i in 1:length(vars)){
  x <- df.zscore_cent[,grepl(vars[i], names(df.zscore_cent))]
  y_label <- sub("aparc_GrayVol_", "", vars[i])
  y_label <- sub("aseg_", "", vars[i])
  y_label <- sub("X", "", vars[i])
  lm <- lm(x ~ birth_weight_kg_qc*adjusted_age_in_years, df.zscore_cent)
  summary <- summary(lm)
  interaction_plots[[i]] <- interact_plot(lm, pred = birth_weight_kg_qc, modx = adjusted_age_in_years) + ylab(y_label) + ggtitle(paste0("bwbyage.qc beta = ", as.character(signif(summary$coefficients[4,1], 3))))
}
interaction_plots
}
}
```

### Surface Area

####  simple linear regressions

##### compute regressions
```{r, echo = FALSE}
print(glue("Global measure for surface area is {"ADI.median"}"))
sum(!is.na(df.zscore_cent$birth_weight_kg_qc))
# Regional -- including cortical and subcortical
bw.qc.adi_sa <- as.data.frame(t(sapply(df.zscore_cent[, which(names(df.zscore_cent) %in% sa_vars)], function(x) {
  lm <- lm(reformulate(c("birth_weight_kg_qc", "median_euler_mean", "ADI.median"), response = "x"), data = df.zscore_cent)
  summary <- summary(lm)
  c(bw_beta = summary$coefficients[2,1], bw_p = summary$coefficients[2,4], bw_se = summary$coefficients[2,2])
})))

bw.qc.adi_sa$bw_p_adj <- p.adjust(bw.qc.adi_sa$bw_p, method = "BH")
print(paste0("Number of significant IDPs after FDR correction: ", dim(bw.qc.adi_sa %>% filter(bw_p_adj < 0.05))[1]))
```

##### add label column to match to ggseg
```{r, echo = FALSE}
bw.qc.adi_sa$label <- rownames(bw.qc.adi_sa) %>% 
  sub("aparc_SurfArea_", "", .)

bw.qc.adi_sa_sig_MC <- bw.qc.adi_sa %>% filter(bw_p_adj < 0.05)

print(paste0("Beta range of significant IDPs: ", range(bw.qc.adi_sa_sig_MC$bw_beta)))
```

##### save betas
```{r, echo = FALSE}
save_table_full <- bw.qc.adi_sa
name <- "bw.qc.adi_surfArea.csv"
# Save dkt cortical betas for spin test
if(save_tables == TRUE){
  output_folder <-dkt_folder
  save_table_full$hemisphere <- ifelse(grepl("lh", save_table_full$label), "L", 
                         ifelse(grepl("rh", save_table_full$label), "R", NA))
  save_table_full$label <- sub("^rh_|lh_", "", save_table_full$label)
  filename_all <- paste0(output_folder, name)
  write.csv(save_table_full, filename_all)
}
# Save full betas
if(save_tables == TRUE){
output_folder <- betas_folder
filename <- paste0(output_folder, name)
write.csv(save_table_full, filename)
}
```

##### visualization of betas on surface maps
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(bw.qc.adi_sa_sig_MC, ggseg::dk, by = "label")

# Regions not matched in atlases
(bw.qc.adi_sa_noplot <- bw.qc.adi_sa_sig_MC[!(bw.qc.adi_sa_sig_MC$label %in% plot_data_dkt$label) & !(bw.qc.adi_sa_sig_MC$label %in% plot_data_aseg$label),])

# plots
(bw.qc.adi_dkt_sa_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = bw_beta)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = limits) +
  theme_void()+
  labs(title = "BW, QC and ADI covar, Surface Area"))

if(save_figs == TRUE){
panel_save(bw.qc.adi_dkt_sa_plot, "bw-qc-adi-dkt-sa", "pdf", plot_output_folder)
}
```

##### Scatterplot for top betas
```{r, echo = FALSE}
if(dim(bw.qc.adi_sa_sig_MC)[1] > 0){
top_n = if(dim(bw.qc.adi_sa_sig_MC)[1] > 10) {10} else {dim(bw.qc.adi_sa_sig_MC)[1]}
top <- bw.qc.adi_sa_sig_MC[order(bw.qc.adi_sa_sig_MC$bw_beta, decreasing = TRUE)[1:top_n],]
scatterplots_bw.qc.adi_sa <- vector(mode = "list", length = top_n)
for(i in 1:top_n){
  pheno <- rownames(top)[i]
  title <- top$label[i]
  names(scatterplots_bw.qc.adi_sa[i]) <- pheno
  scatterplots_bw.qc.adi_sa[[i]] <- ggplot(df.zscore_cent, aes(x = birth_weight_kg_qc, y = !!sym(pheno))) +
    geom_point() +
    geom_smooth(method = "lm") +
    theme_minimal() +
    labs(x = "Gestational Age (z-scored)", y = "Centile (z-scored)", title = title)
}

nCol <- 2
nRow <- 2
breaks <- c(seq(from = 0, to = top_n , by = (nCol*nRow)), top_n) %>% unique(.)
for(k in 1:(length(breaks)-1)){
  start <- breaks[k]+1
  end <- breaks[k+1]
  do.call("grid.arrange", c(scatterplots_bw.qc.adi_sa[start:end], ncol=nCol))
}
}
```

####  lm with age interaction

##### compute regressions
```{r, echo = FALSE}
# Regional -- including cortical and subcortical
bwbyage.qc.adi_sa <- as.data.frame(t(sapply(df.zscore_cent[, which(names(df.zscore_cent) %in% sa_vars)], function(x) {
  lm <- lm(reformulate(c("birth_weight_kg_qc", "adjusted_age_in_years", "median_euler_mean", "ADI.median", "birth_weight_kg_qc:adjusted_age_in_years"), response = "x"), data = df.zscore_cent)
  summary <- summary(lm)
  c(bw_beta = summary$coefficients[2,1], bw_p = summary$coefficients[2,4], bw_se = summary$coefficients[2,2],
   age_beta = summary$coefficients[3,1], age_p = summary$coefficients[3,4], age_se = summary$coefficients[3,2],
   bwbyage_beta = summary$coefficients[6,1], bwbyage_p = summary$coefficients[6,4], bwbyage_se = summary$coefficients[6,2])
})))

bwbyage.qc.adi_sa$bw_p_adj <- p.adjust(bwbyage.qc.adi_sa$bw_p, method = "BH")
bwbyage.qc.adi_sa$bwbyage_p_adj <- p.adjust(bwbyage.qc.adi_sa$bwbyage_p, method = "BH")

print(paste0("Age Main significant results (shouldn't be any): ", dim(bwbyage.qc.adi_sa %>% filter(age_p < 0.05))[1]))
print(paste0("Number of significant IDPs - NO correction: ", dim(bwbyage.qc.adi_sa %>% filter(bwbyage_p < 0.05))[1]))
print(paste0("Number of significant IDPs - AFTER FDR correction: ", dim(bwbyage.qc.adi_sa %>% filter(bwbyage_p_adj < 0.05))[1]))
```

##### add label column to match to ggseg
```{r, echo = FALSE}
bwbyage.qc.adi_sa$label <- rownames(bwbyage.qc.adi_sa) %>% 
  sub("aparc_SurfArea_", "", .)

bwbyage.qc.adi_sa_sig_MC <- bwbyage.qc.adi_sa %>% filter(bwbyage_p_adj < 0.05)

print(paste0("Beta range of significant IDPs: ", range(bwbyage.qc.adi_sa_sig_MC$bw_beta)))
```

##### save betas
```{r, echo = FALSE}
save_table_full <- bwbyage.qc.adi_sa
name <- "bwbyage.qc.adi_surfArea.csv"
# Save dkt cortical betas for spin test
if(save_tables == TRUE){
  output_folder <-dkt_folder
  save_table_full$hemisphere <- ifelse(grepl("lh", save_table_full$label), "L", 
                         ifelse(grepl("rh", save_table_full$label), "R", NA))
  save_table_full$label <- sub("^rh_|lh_", "", save_table_full$label)
  filename_all <- paste0(output_folder, name)
  write.csv(save_table_full, filename_all)
}
# Save full betas
if(save_tables == TRUE){
output_folder <- betas_folder
filename <- paste0(output_folder, name)
write.csv(save_table_full, filename)
}
```

##### visualize interaction betas on surface maps
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(bwbyage.qc.adi_sa_sig_MC, ggseg::dk, by = "label")

# Regions not matched in atlases
(bwbyage.qc_sa_noplot <- bwbyage.qc.adi_sa_sig_MC[!(bwbyage.qc.adi_sa_sig_MC$label %in% plot_data_dkt$label) & !(bwbyage.qc.adi_sa_sig_MC$label %in% plot_data_aseg$label),])

# plots
(bwbyage.qc_dkt_sa_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = bwbyage_beta)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = limits) +
  theme_void()+
  labs(title = "BWxAge, QC and ADI covar, Surface Area"))

if(save_figs == TRUE){
panel_save(bwbyage.qc_dkt_sa_plot, "bwbyage-qc-adi-dkt-sa", "pdf", plot_output_folder)
}
```

##### visualize interaction plots
```{r, echo = FALSE}
if (dim(bwbyage.qc.adi_sa_sig_MC)[1] > 0){
vars <- rownames(bwbyage.qc.adi_sa_sig_MC[order(-(bwbyage.qc.adi_sa_sig_MC$bwbyage_beta)),])
interaction_plots <-  vector(mode = "list", length = length(vars))
if (length(vars) > 0) {
for (i in 1:length(vars)){
  x <- df.zscore_cent[,grepl(vars[i], names(df.zscore_cent))]
  y_label <- sub("aparc_SurfArea_", "", vars[i])
  lm <- lm(x ~ birth_weight_kg_qc*adjusted_age_in_years, df.zscore_cent)
  summary <- summary(lm)
  interaction_plots[[i]] <- interact_plot(lm, pred = birth_weight_kg_qc, modx = adjusted_age_in_years) + ylab(y_label) + ggtitle(paste0("bwbyage.qc beta = ", as.character(signif(summary$coefficients[4,1], 3))))
}
interaction_plots
}
}
```

### Thickness

####  simple linear regressions

##### compute regressions
```{r, echo = FALSE}
print(glue("Global measure for cortical thickness is {"ADI.median"}"))
sum(!is.na(df.zscore_cent$birth_weight_kg_qc))
# Regional -- including cortical and subcortical
bw.qc.adi_th <- as.data.frame(t(sapply(df.zscore_cent[, which(names(df.zscore_cent) %in% th_vars)], function(x) {
  lm <- lm(reformulate(c("birth_weight_kg_qc", "median_euler_mean", "ADI.median"), response = "x"), data = df.zscore_cent)
  summary <- summary(lm)
  c(bw_beta = summary$coefficients[2,1], bw_p = summary$coefficients[2,4], bw_se = summary$coefficients[2,2])
})))

bw.qc.adi_th$bw_p_adj <- p.adjust(bw.qc.adi_th$bw_p, method = "BH")
print(paste0("Number of significant IDPs after FDR correction: ", dim(bw.qc.adi_th %>% filter(bw_p_adj < 0.05))[1]))
```

##### add label column to match to ggseg
```{r, echo = FALSE}
bw.qc.adi_th$label <- rownames(bw.qc.adi_th) %>% 
  sub("aparc_ThickAvg_", "", .)

bw.qc.adi_th_sig_MC <- bw.qc.adi_th %>% filter(bw_p_adj < 0.05)

print(paste0("Beta range of significant IDPs: ", range(bw.qc.adi_th_sig_MC$bw_beta)))
```

##### save betas
```{r, echo = FALSE}
save_table_full <- bw.qc.adi_th
name <- "bw.qc.adi_thick.csv"
# Save dkt cortical betas for spin test
if(save_tables == TRUE){
  output_folder <-dkt_folder
  save_table_full$hemisphere <- ifelse(grepl("lh", save_table_full$label), "L", 
                         ifelse(grepl("rh", save_table_full$label), "R", NA))
  save_table_full$label <- sub("^rh_|lh_", "", save_table_full$label)
  filename_all <- paste0(output_folder, name)
  write.csv(save_table_full, filename_all)
}
# Save full betas
if(save_tables == TRUE){
output_folder <- betas_folder
filename <- paste0(output_folder, name)
write.csv(save_table_full, filename)
}
```

##### visualization of betas on surface maps
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(bw.qc.adi_th_sig_MC, ggseg::dk, by = "label")

# Regions not matched in atlases
(bw.qc.adi_th_noplot <- bw.qc.adi_th_sig_MC[!(bw.qc.adi_th_sig_MC$label %in% plot_data_dkt$label) & !(bw.qc.adi_th_sig_MC$label %in% plot_data_aseg$label),])

# plots
(bw.qc.adi_dkt_th_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = bw_beta)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = limits) +
  theme_void()+
  labs(title = "BW, QC and ADI covar, Cortical Thickness"))

if(save_figs == TRUE){
panel_save(bw.qc.adi_dkt_th_plot, "bw-qc-adi-dkt-th", "pdf", plot_output_folder)
}
```

##### Scatterplot for top betas
```{r, echo = FALSE}
if(dim(bw.qc.adi_th_sig_MC)[1] > 0) {
top_n = if(dim(bw.qc.adi_th_sig_MC)[1] > 10) {10} else {dim(bw.qc.adi_th_sig_MC)[1]}
top <- bw.qc.adi_th_sig_MC[order(abs(bw.qc.adi_th_sig_MC$bw_beta), decreasing = TRUE)[1:top_n],]
scatterplots_bw.qc.adi_th <- vector(mode = "list", length = top_n)
for(i in 1:top_n){
  print(i)
  pheno <- rownames(top)[i]
  print(pheno)
  title <- top$label[i]
  names(scatterplots_bw.qc.adi_th)[i] <- pheno
  scatterplots_bw.qc.adi_th[[i]] <- ggplot(df.zscore_cent, aes(x = birth_weight_kg_qc, y = !!sym(pheno))) +
    geom_point() +
    geom_smooth(method = "lm") +
    theme_minimal() +
    labs(x = "Gestational Age (z-scored)", y = "Centile (z-scored)", title = title)
}

nCol <- 2
nRow <- 2
breaks <- c(seq(from = 0, to = top_n , by = (nCol*nRow)), top_n) %>% unique(.)
for(k in 1:(length(breaks)-1)){
  start <- breaks[k]+1
  end <- breaks[k+1]
  do.call("grid.arrange", c(scatterplots_bw.qc.adi_th[start:end], ncol=nCol))
}
}
```

####  lm with age interaction

##### compute regressions
```{r, echo = FALSE}
# Regional -- including cortical and subcortical
bwbyage.qc.adi_th <- as.data.frame(t(sapply(df.zscore_cent[, which(names(df.zscore_cent) %in% th_vars)], function(x) {
  lm <- lm(reformulate(c("birth_weight_kg_qc", "adjusted_age_in_years", "median_euler_mean", "ADI.median", "birth_weight_kg_qc:adjusted_age_in_years"), response = "x"), data = df.zscore_cent)
  summary <- summary(lm)
  c(bw_beta = summary$coefficients[2,1], bw_p = summary$coefficients[2,4], bw_se = summary$coefficients[2,2],
   age_beta = summary$coefficients[3,1], age_p = summary$coefficients[3,4], age_se = summary$coefficients[3,2],
   bwbyage_beta = summary$coefficients[6,1], bwbyage_p = summary$coefficients[6,4], bwbyage_se = summary$coefficients[6,2])
})))

bwbyage.qc.adi_th$bw_p_adj <- p.adjust(bwbyage.qc.adi_th$bw_p, method = "BH")
bwbyage.qc.adi_th$bwbyage_p_adj <- p.adjust(bwbyage.qc.adi_th$bwbyage_p, method = "BH")

print(paste0("Age Main significant results (shouldn't be any): ", dim(bwbyage.qc.adi_th %>% filter(age_p < 0.05))[1]))
print(paste0("Number of significant IDPs - NO correction: ", dim(bwbyage.qc.adi_th %>% filter(bwbyage_p < 0.05))[1]))
print(paste0("Number of significant IDPs - AFTER FDR correction: ", dim(bwbyage.qc.adi_th %>% filter(bwbyage_p_adj < 0.05))[1]))
```

##### add label column to match to ggseg
```{r, echo = FALSE}
bwbyage.qc.adi_th$label <- rownames(bwbyage.qc.adi_th) %>% 
  sub("aparc_ThickAvg_", "", .)
  
bwbyage.qc.adi_th_sig_MC <- bwbyage.qc.adi_th %>% filter(bwbyage_p_adj < 0.05)

print(paste0("Beta range of significant IDPs: ", range(bwbyage.qc.adi_th_sig_MC$bwbyage_beta)))
```

##### save betas
```{r, echo = FALSE}
save_table_full <- bwbyage.qc.adi_th
name <- "bwbyage.qc.adi_thick.csv"
# Save dkt cortical betas for spin test
if(save_tables == TRUE){
  output_folder <-dkt_folder
  save_table_full$hemisphere <- ifelse(grepl("lh", save_table_full$label), "L", 
                         ifelse(grepl("rh", save_table_full$label), "R", NA))
  save_table_full$label <- sub("^rh_|lh_", "", save_table_full$label)
  filename_all <- paste0(output_folder, name)
  write.csv(save_table_full, filename_all)
}
# Save full betas
if(save_tables == TRUE){
output_folder <- betas_folder
filename <- paste0(output_folder, name)
write.csv(save_table_full, filename)
}
```

##### visualize interaction betas on surface maps
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(bwbyage.qc.adi_th_sig_MC, ggseg::dk, by = "label")

# Regions not matched in atlases
(bwbyage.qc.adi_th_noplot <- bwbyage.qc.adi_th_sig_MC[!(bwbyage.qc.adi_th_sig_MC$label %in% plot_data_dkt$label) & !(bwbyage.qc.adi_th_sig_MC$label %in% plot_data_aseg$label),])

# plots
(bwbyage.qc.adi_dkt_th_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = bwbyage_beta)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = limits) +
  theme_void()+
  labs(title = "BWxAge, QC and ADI covar, Cortical Thickness"))

if(save_figs == TRUE){
panel_save(bwbyage.qc.adi_dkt_th_plot, "bwbyage-qc-adi-dkt-th", "pdf", plot_output_folder)
}
```

##### visualize interaction plots
```{r, echo = FALSE}
if(dim(bwbyage.qc.adi_th_sig_MC)[1] > 0) {
vars <- rownames(bwbyage.qc.adi_th_sig_MC[order(-(bwbyage.qc.adi_th_sig_MC$bwbyage_beta)),])
interaction_plots <-  vector(mode = "list", length = length(vars))
if (length(vars) > 0) {
for (i in 1:length(vars)){
  x <- df.zscore_cent[,grepl(vars[i], names(df.zscore_cent))]
  y_label <- sub("aparc_ThickAvg_", "", vars[i])
  lm <- lm(x ~ birth_weight_kg_qc*adjusted_age_in_years, df.zscore_cent)
  summary <- summary(lm)
  interaction_plots[[i]] <- interact_plot(lm, pred = birth_weight_kg_qc, modx = adjusted_age_in_years) + ylab(y_label) + ggtitle(paste0("bwbyage.qc.adi beta = ", as.character(signif(summary$coefficients[4,1], 3))))
}
interaction_plots
}
}
```

####  group by age and check bw.qc.adi on thickness effect

##### young

###### compute regressions
```{r, echo = FALSE}
sum(!is.na(df.zscore_cent_young$birth_weight_kg_qc))
# Regional -- including cortical and subcortical
bw.qc.adi_th_young <- as.data.frame(t(sapply(df.zscore_cent_young[, which(names(df.zscore_cent_young) %in% th_vars)], function(x) {
  lm <- lm(reformulate(c("birth_weight_kg_qc", "median_euler_mean", "ADI.median"), response = "x"), data = df.zscore_cent_young)
  summary <- summary(lm)
  c(bw_beta = summary$coefficients[2,1], bw_p = summary$coefficients[2,4], bw_se = summary$coefficients[2,2])
})))

bw.qc.adi_th_young$bw_p_adj <- p.adjust(bw.qc.adi_th_young$bw_p, method = "BH")
print(paste0("Number of significant IDPs after FDR correction: ", dim(bw.qc.adi_th_young %>% filter(bw_p_adj < 0.05))[1]))
```

###### add label column to match to ggseg
```{r, echo = FALSE}
bw.qc.adi_th_young$label <- rownames(bw.qc.adi_th_young) %>% 
  sub("aparc_ThickAvg_", "", .)

bw.qc.adi_th_young_sig_MC <- bw.qc.adi_th_young %>% filter(bw_p_adj < 0.05)

print(paste0("Beta range of significant IDPs: ", range(bw.qc.adi_th_young_sig_MC$bw_beta)))
```

###### save betas
```{r, echo = FALSE}
save_table_full <- bw.qc.adi_th_young
name <- "bw.qc.adi_thick_young.csv"
# Save dkt cortical betas for spin test
if(save_tables == TRUE){
  output_folder <-dkt_folder
  save_table_full$hemisphere <- ifelse(grepl("lh", save_table_full$label), "L", 
                         ifelse(grepl("rh", save_table_full$label), "R", NA))
  save_table_full$label <- sub("^rh_|lh_", "", save_table_full$label)
  filename_all <- paste0(output_folder, name)
  write.csv(save_table_full, filename_all)
}
# Save full betas
if(save_tables == TRUE){
output_folder <- betas_folder
filename <- paste0(output_folder, name)
write.csv(save_table_full, filename)
}
```

###### visualization of betas on surface maps
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(bw.qc.adi_th_young_sig_MC, ggseg::dk, by = "label")

# Regions not matched in atlases
(bw.qc.adi_th_young_noplot <- bw.qc.adi_th_young_sig_MC[!(bw.qc.adi_th_young_sig_MC$label %in% plot_data_dkt$label) & !(bw.qc.adi_th_young_sig_MC$label %in% plot_data_aseg$label),])

# plots
(bw.qc.adiyoung_dkt_th_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = bw_beta)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = limits) +
  theme_void()+
  labs(title = "BW, Young, QC and ADI covar, Cortical Thickness"))

if(save_figs == TRUE){
panel_save(bw.qc.adiyoung_dkt_th_plot, "bw-qc-meanth-young-dkt-th", "pdf", plot_output_folder)
}
```

##### old

###### compute regressions
```{r, echo = FALSE}
sum(!is.na(df.zscore_cent_old$birth_weight_kg_qc))
# Regional -- including cortical and subcortical
bw.qc.adi_th_old <- as.data.frame(t(sapply(df.zscore_cent_old[, which(names(df.zscore_cent_old) %in% th_vars)], function(x) {
  lm <- lm(reformulate(c("birth_weight_kg_qc", "median_euler_mean", "ADI.median"), response = "x"), data = df.zscore_cent_old)
  summary <- summary(lm)
  c(bw_beta = summary$coefficients[2,1], bw_p = summary$coefficients[2,4], bw_se = summary$coefficients[2,2])
})))

bw.qc.adi_th_old$bw_p_adj <- p.adjust(bw.qc.adi_th_old$bw_p, method = "BH")
print(paste0("Number of significant IDPs after FDR correction: ", dim(bw.qc.adi_th_old %>% filter(bw_p_adj < 0.05))[1]))
```

###### add label column to match to ggseg
```{r, echo = FALSE}
bw.qc.adi_th_old$label <- rownames(bw.qc.adi_th_old) %>% 
  sub("aparc_ThickAvg_", "", .)

bw.qc.adi_th_old_sig_MC <- bw.qc.adi_th_old %>% filter(bw_p_adj < 0.05)

print(paste0("Beta range of significant IDPs: ", range(bw.qc.adi_th_old_sig_MC$bw_beta)))
```

###### save betas
```{r, echo = FALSE}
save_table_full <- bw.qc.adi_th_old
name <- "bw.qc.adi_thick_old.csv"
# Save dkt cortical betas for spin test
if(save_tables == TRUE){
  output_folder <-dkt_folder
  save_table_full$hemisphere <- ifelse(grepl("lh", save_table_full$label), "L", 
                         ifelse(grepl("rh", save_table_full$label), "R", NA))
  save_table_full$label <- sub("^rh_|lh_", "", save_table_full$label)
  filename_all <- paste0(output_folder, name)
  write.csv(save_table_full, filename_all)
}
# Save full betas
if(save_tables == TRUE){
output_folder <- betas_folder
filename <- paste0(output_folder, name)
write.csv(save_table_full, filename)
}
```

###### visualization of betas on surface maps
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(bw.qc.adi_th_old_sig_MC, ggseg::dk, by = "label")

# Regions not matched in atlases
(bw.qc.adi_th_old_noplot <- bw.qc.adi_th_old_sig_MC[!(bw.qc.adi_th_old_sig_MC$label %in% plot_data_dkt$label) & !(bw.qc.adi_th_old_sig_MC$label %in% plot_data_aseg$label),])

# plots
(bw.qc.adiold_dkt_th_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = bw_beta)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = limits) +
  theme_void()+
  labs(title = "BW, Old, QC and ADI covar, Cortical Thickness"))
if(save_figs == TRUE){
panel_save(bw.qc.adiold_dkt_th_plot, "bw-qc-meanth-old-dkt-th", "pdf", plot_output_folder)
}
```
