---
title: "slipga_spline_analysis"
author: "Eren Kafadar"
date: "2026-06-02"
output:
  html_document:
    df_print: paged
params:
  output_folder : "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/analysis_results/median-th-mpr/"
  qnorm_centiles_path: "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/analysis_results/median-th-mpr/analysis_dfs/df_zscore_distinct.csv"
  lookup_stats: "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/analysis_results/median-th-mpr/analysis_dfs/lookup_stats_distinct.RData"
  rm : FALSE
editor_options: 
  chunk_output_type: console
---
# Define params if they do not exist
```{r params, echo = FALSE}
if (!exists("params")) {
  params <- list(
  output_folder = "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/analysis_results/",
  qnorm_centiles_path = "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/analysis_results/median-th-mpr/analysis_dfs/df_zscore_distinct.csv",
  lookup_stats = "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/analysis_results/median-th-mpr/analysis_dfs/lookup_stats_distinct.RData",
  rm = FALSE
  )
}
print(params)
``` 

# Load Packages
```{r packages, echo = FALSE}
# Load libraries
require(dplyr)
require(glue)
require(magrittr)
require(readr)
require(stringr)
require(tidyr)
require(scales)
require(mgcv)
require(pbkrtest)
require(gamlssTools) # Margaret' package
require(tibble)
require(lme4)
require(purrr)
# For figures/plotting
require(ggplot2)
require(gridExtra)
require(ggsegDKT)
require(ggseg)
require(ggpubr)
require(ggcorrplot)
require(cowplot)
require(ggh4x)
require(egg)
require(interactions)
require(growthstandards)
require(gtsummary)
require(svglite)
require(patchwork)

`%nin%` = Negate(`%in%`)
```

# Controls
```{r setup, echo = TRUE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)
# set command
save_tables <- TRUE
plot <- TRUE
save_figs <- FALSE
global_measure_vol <- "global_eTIV"
global_measure_sa <- "global_SurfaceArea"
global_measure_th <- "global_MeanThickness"
set.seed(42)

output_folder <- params$output_folder
qnorm_centiles_path <- params$qnorm_centiles_path
lookup_stats <- params$lookup_stats
rm <- params$rm
```

# Define Paths
```{r, echo = TRUE}
# set folder paths
code_path <- "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/"
vars_path <-  "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/output/vars_2511_new/"

# output folder
output_folder <- params$output_folder
if (!dir.exists(output_folder)) dir.create(output_folder)
##  output subfolder paths
analysis_dfs_folder <- paste0(output_folder, "analysis_dfs/")
if (!dir.exists(analysis_dfs_folder)) dir.create(analysis_dfs_folder)

models_folder <- paste0(output_folder, "models/")
if (!dir.exists(models_folder)) dir.create(models_folder)

csv_output_folder <-  paste0(output_folder, "results_csv/")
if (!dir.exists(csv_output_folder)) dir.create(csv_output_folder)

  betas_folder <- paste0(csv_output_folder, "full_betas/")
  if (!dir.exists(betas_folder)) dir.create(betas_folder)
  dkt_folder <- paste0(csv_output_folder, "dkt_betas/")
  if (!dir.exists(dkt_folder)) dir.create(dkt_folder)
  slwindow_folder <- paste0(csv_output_folder, "sliding_window/")
  if (!dir.exists(slwindow_folder)) dir.create(slwindow_folder)
  sp_stat_folder <- paste0(csv_output_folder, "spline_stats/")
  if (!dir.exists(sp_stat_folder)) dir.create(sp_stat_folder)

plot_output_folder <- paste0(output_folder, "plots/")
if (!dir.exists(plot_output_folder)) dir.create(plot_output_folder)

  ggseg_plot_folder <- paste0(plot_output_folder, "ggseg_plots/")
  if (!dir.exists(ggseg_plot_folder)) dir.create(ggseg_output_folder)
  cmd_plot_folder <- paste0(plot_output_folder, "cmd_plots/")
  if (!dir.exists(cmd_plot_folder)) dir.create(cmd_plot_folder)
  modelvis_plot_folder <- paste0(plot_output_folder, "modelvis_plots/")
  if (!dir.exists(modelvis_plot_folder)) dir.create(modelvis_plot_folder)
  slwindow_plot_folder <- paste0(plot_output_folder, "sliding_window_plots/")
  if (!dir.exists(slwindow_plot_folder)) dir.create(slwindow_plot_folder)

```

# Load custom functions
```{r fcts, echo = FALSE}
source(glue("{code_path}sliding_window.R"))
source(glue("{code_path}visualize_model_bartfct.R"))
source(glue("{code_path}vis_spline_noint_EK.R"))
source(glue("{code_path}model_fitting_fcts.R"))
source(glue("{code_path}model_extract_fcts.R"))

panel_save <- function(plot, filename, extension, folder, plot_width = 8, plot_height = 10){
  save_name <- paste0(folder, filename, ".", extension)
  ggsave(save_name, plot = plot, dpi = 300, width = plot_width, height = plot_height)
}
```

# Color Specifications
```{r, echo = TRUE}
limits = c(-0.3, 0.3)
low_color = "#4878CF" # blue
mid_color = "white"
high_color = "#D65F5F" # red

ga_color = "#6D8299" #muted blue
bw_color = "#709085" # muted green
bwp_color = "#A68A64" # muted beige
```

# Read in data files
```{r, echo = FALSE}
full_vars <- read.csv(paste0(vars_path,"all_vars.csv")) %>% pull(x)
global_vars <- read.csv(paste0(vars_path,"global_vars.csv")) %>% pull(x)
vol_vars <- read.csv(paste0(vars_path,"vol_vars.csv")) %>% pull(x)
sa_vars <- read.csv(paste0(vars_path,"sa_vars.csv")) %>% pull(x)
th_vars <- read.csv(paste0(vars_path,"th_vars.csv")) %>% pull(x)
th_vars_median <- c(paste0(th_vars, ".median"))
namemap <- read.csv(glue("{vars_path}name_mapping.csv"))

df.zscore <- read.csv(qnorm_centiles_path)
load(lookup_stats)

#determine if there is both mpr and median for thickness values
median_th_mpr_flag <- any(grepl(".median", names(df.zscore)))

## Define global vars of interest
global_vars_totest <- c("global_eTIV", "global_SubCortGrayVol","global_CortexVol", "global_CerebralWhiteMatterVol", "global_VentricleChoroidVol","global_SurfaceArea", "global_MeanThickness")
global_vars_totest_nothick <- c("global_eTIV", "global_SubCortGrayVol","global_CortexVol", "global_CerebralWhiteMatterVol", "global_VentricleChoroidVol","global_SurfaceArea")
```
# Code to map raw onto z-score
```{r}
#df.zscore %>%  mutate(ga_raw = map2_dbl(gestational_age, "gestational_age", ~ .x*sd_lookup_all[[.y]] + mean_lookup_all[[.y]]))
```


# Global phenotypes, test non-linear vs linear

## GA

```{r}
if(median_th_mpr_flag){
  ga_sp_models_global <- fit_bam_models(df.zscore, global_vars_totest_nothick, 
                                        predictor = "gestational_age",
                                        covariates = c("qc_var"), save = FALSE)
  
  ga_sp_table_global <- extract_bam_features_main(models = ga_sp_models_global, 
                                                  data = df.zscore, 
                                                  predictor = "gestational_age",
                                                  covariates = c("qc_var"))
  
  
  ga_sp_models_meanthick <- fit_bam_models(df.zscore, "global_MeanThickness", 
                                        predictor = "gestational_age",
                                        covariates = c("euler_mean_mpr"), save = FALSE)
  
  ga_sp_table_meanthick <- extract_bam_features_main(models = ga_sp_models_meanthick, 
                                                  data = df.zscore, 
                                                  predictor = "gestational_age",
                                                  covariates = c("euler_mean_mpr"))
  
  ga_sp_models_global <- c(ga_sp_models_global, ga_sp_models_meanthick)
  ga_sp_table_global <- rbind(ga_sp_table_global, ga_sp_table_meanthick)
  
  rm(ga_sp_models_meanthick, ga_sp_table_meanthick)
} else {

  ga_sp_models_global <- fit_bam_models(df.zscore, global_vars_totest, 
                                      predictor = "gestational_age",
                                      covariates = c("qc_var"), save = FALSE)
  
  ga_sp_table_global <- extract_bam_features_main(models = ga_sp_models_global, 
                                                data = df.zscore, 
                                                predictor = "gestational_age",
                                                covariates = c("qc_var"))
}
```
#### fit models
```{r, echo = FALSE}
ga_sp_models_global <- lapply(global_vars_totest, function(x) {
  dat <- df.zscore
  if(grepl("Thick", x)) {
    dat <- df.zscore%>% filter(age_over2_adj == TRUE)
      if(median_th_mpr_flag == TRUE){
         form_og <- as.formula(glue("{x} ~ s(gestational_age) + euler_mean_mpr"))
         form_nolin <- as.formula(glue("{x} ~ s(gestational_age, m = c(2,0)) + gestational_age + euler_mean_mpr"))
        }
      } else {
         form_og <- as.formula(glue("{x} ~ s(gestational_age) + qc_var"))
         form_nolin <- as.formula(glue("{x} ~ s(gestational_age, m = c(2,0)) + gestational_age + qc_var"))
      }
  gam_model_og <- gamm(as.formula(form_og),
                       #random = list(participant_id = ~ 1),
              data = dat)
  gam_model_nolin <- gamm(as.formula(form_nolin),
                          #random = list(participant_id = ~ 1),
              data = dat)
  return(list(full_spline = gam_model_og,
         no_linear = gam_model_nolin)) })

ga_sp_models_global <- setNames(ga_sp_models_global, global_vars_totest)

ga_sp_gam_table_global <- as.data.frame(t(sapply(ga_sp_models_global,function(x) {
  summary_og <- summary(x$full_spline$gam)
  summary_nolin <- summary(x$no_linear$gam)
  c(ga_og_edf = summary_og$s.table[1,1], ga_og_p = summary_og$s.table[1,4],
    ga_nolin_edf = summary_nolin$s.table[1,1], ga_nolin_p = summary_nolin$s.table[1,4])
  })))

rownames(ga_sp_gam_table_global) <- global_vars_totest

ga_sp_gam_table_global$ga_og_p_adj <- p.adjust(ga_sp_gam_table_global$ga_og_p, method = "BH")
ga_sp_gam_table_global$ga_nolin_p_adj <- p.adjust(ga_sp_gam_table_global$ga_nolin_p, method = "BH")

print(paste0("Number of significant IDPs for GA Smooth Sig: ", dim(ga_sp_gam_table_global %>% filter(ga_nolin_p < 0.05))[1]))
print(paste0("Number of significant IDPs for GA Smooth Sig AFTER FDR correction: ", dim(ga_sp_gam_table_global %>% filter(ga_nolin_p_adj < 0.05))[1]))

if(save_tables == TRUE){
  output_folder <- sp_stat_folder
  filename <- paste0(output_folder, "ga_sp_global.csv")
  write.csv(ga_sp_gam_table_global, filename)
}
```

#### Visualize
```{r}
ga_sp_plots_global <- lapply(ga_sp_models_global, function(x) {
  
  modobj = x$full_spline
  
  model_plot <- visualize_spline_model(
  modobj = modobj,
  plot_og_scale = TRUE,
  mean_lookup = mean_lookup_distinct,
  sd_lookup = sd_lookup_distinct,
  pnorm_smooth_var = FALSE,
  smooth_var = "gestational_age",
  plabels = "Gestational Age",
  color_point = ga_color,
  color_line = ga_color)
  
  model_plot <- model_plot + 
    labs(x="Gestational Age (weeks)",
         title = glue("{modobj$gam$terms[[2]]}")) +
    theme(axis.title.y = element_blank())
})

ga_sp_plots_global <- setNames(ga_sp_plots_global, global_vars_totest)

if(save_figs){
saveRDS(ga_sp_plots_global, glue("{modelvis_plot_folder}ga_sp_plots_global.rds"))
}
```

#### Show plots
```{r, echo = TRUE}
wrap_plots(ga_sp_plots_global[1:4], ncol = 2)
wrap_plots(ga_sp_plots_global[5:7], ncol = 2)
```

## BW
#### fit models
```{r, echo = FALSE}
bw_sp_models_global <- lapply(global_vars_totest, function(x) {
  dat <- df.zscore
  if(grepl("Thick", x)) {
    dat <- df.zscore %>% filter(age_over2_adj == TRUE)
      if(median_th_mpr_flag == TRUE){
         form_og <- as.formula(glue("{x} ~ s(birth_weight_kg) + euler_mean_mpr"))
         form_nolin <- as.formula(glue("{x} ~ s(birth_weight_kg, m = c(2,0)) + birth_weight_kg + euler_mean_mpr"))
        }
      } else {
         form_og <- as.formula(glue("{x} ~ s(birth_weight_kg) + qc_var"))
         form_nolin <- as.formula(glue("{x} ~ s(birth_weight_kg, m = c(2,0)) + birth_weight_kg + qc_var"))
      }
  gam_model_og <- gamm(as.formula(form_og),
              data = dat)
  gam_model_nolin <- gamm(as.formula(form_nolin),
              data = dat)
  return(list(full_spline = gam_model_og,
         no_linear = gam_model_nolin)) })

bw_sp_models_global <- setNames(bw_sp_models_global, global_vars_totest)

bw_sp_gam_table_global <- as.data.frame(t(sapply(bw_sp_models_global,function(x) {
  summary_og <- summary(x$full_spline$gam)
  summary_nolin <- summary(x$no_linear$gam)
  c(bw_og_edf = summary_og$s.table[1,1], bw_og_p = summary_og$s.table[1,4],
    bw_nolin_edf = summary_nolin$s.table[1,1], bw_nolin_p = summary_nolin$s.table[1,4])
  })))

rownames(bw_sp_gam_table_global) <- global_vars_totest

bw_sp_gam_table_global$bw_og_p_adj <- p.adjust(bw_sp_gam_table_global$bw_og_p, method = "BH")
bw_sp_gam_table_global$bw_nolin_p_adj <- p.adjust(bw_sp_gam_table_global$bw_nolin_p, method = "BH")

print(paste0("Number of significant IDPs for BW Smooth Sig: ", dim(bw_sp_gam_table_global %>% filter(bw_nolin_p < 0.05))[1]))
print(paste0("Number of significant IDPs for BW Smooth Sig AFTER FDR correction: ", dim(bw_sp_gam_table_global %>% filter(bw_nolin_p_adj < 0.05))[1]))

if(save_tables == TRUE){
  output_folder <- sp_stat_folder
  filename <- paste0(output_folder, "bw_sp_global.csv")
  write.csv(bw_sp_gam_table_global, filename)
}
```

#### Visualize
```{r}
bw_sp_plots_global <- lapply(bw_sp_models_global, function(x) {
  
  modobj = x$full_spline
  
model_plot <- visualize_spline_model(
modobj = modobj,
smooth_var = "birth_weight_kg",
plot_og_scale = TRUE,
mean_lookup = mean_lookup_distinct,
sd_lookup = sd_lookup_distinct,
pnorm_smooth_var = FALSE,
plabels = "Birth Weight",
color_point = bw_color,
color_line = bw_color)
  
  model_plot <- model_plot + 
    labs(x="Birth Weight (kg))",
         title = glue("{modobj$gam$terms[[2]]}")) +
    theme(axis.title.y = element_blank())
})

bw_sp_plots_global <- setNames(bw_sp_plots_global, global_vars_totest)

if(save_figs){
saveRDS(bw_sp_plots_global, glue("{modelvis_plot_folder}bw_sp_plots_global.rds"))
}
```

#### Show plots
```{r, echo = TRUE}
wrap_plots(bw_sp_plots_global[1:4], ncol = 2)
wrap_plots(bw_sp_plots_global[5:7], ncol = 2)
```


## BWP
#### fit models
```{r, echo = FALSE}
bwp_sp_models_global <- lapply(global_vars_totest, function(x) {
  dat <- df.zscore
  if(grepl("Thick", x)) {
    dat <- df.zscore %>% filter(age_over2_adj == TRUE)
      if(median_th_mpr_flag == TRUE){
         form_og <- as.formula(glue("{x} ~ s(bwp_fen) + euler_mean_mpr"))
         form_nolin <- as.formula(glue("{x} ~ s(bwp_fen, m = c(2,0)) + bwp_fen + euler_mean_mpr"))
        }
      } else {
         form_og <- as.formula(glue("{x} ~ s(bwp_fen) + qc_var"))
         form_nolin <- as.formula(glue("{x} ~ s(bwp_fen, m = c(2,0)) + bwp_fen + qc_var"))
      }
  gam_model_og <- gamm(as.formula(form_og),
              data = dat)
  gam_model_nolin <- gamm(as.formula(form_nolin),
              data = dat)
  return(list(full_spline = gam_model_og,
         no_linear = gam_model_nolin)) })

bwp_sp_models_global <- setNames(bwp_sp_models_global, global_vars_totest)

bwp_sp_gam_table_global <- as.data.frame(t(sapply(bwp_sp_models_global,function(x) {
  summary_og <- summary(x$full_spline$gam)
  summary_nolin <- summary(x$no_linear$gam)
  c(bwp_og_edf = summary_og$s.table[1,1], bwp_og_p = summary_og$s.table[1,4],
    bwp_nolin_edf = summary_nolin$s.table[1,1], bwp_nolin_p = summary_nolin$s.table[1,4])
  })))

rownames(bwp_sp_gam_table_global) <- global_vars_totest

bwp_sp_gam_table_global$bwp_og_p_adj <- p.adjust(bwp_sp_gam_table_global$bwp_og_p, method = "BH")
bwp_sp_gam_table_global$bwp_nolin_p_adj <- p.adjust(bwp_sp_gam_table_global$bwp_nolin_p, method = "BH")

print(paste0("Number of significant IDPs for bwp Smooth Sig: ", dim(bwp_sp_gam_table_global %>% filter(bwp_nolin_p < 0.05))[1]))
print(paste0("Number of significant IDPs for bwp Smooth Sig AFTER FDR correction: ", dim(bwp_sp_gam_table_global %>% filter(bwp_nolin_p_adj < 0.05))[1]))

if(save_tables == TRUE){
  output_folder <- sp_stat_folder
  filename <- paste0(output_folder, "bwp_sp_global.csv")
  write.csv(bwp_sp_gam_table_global, filename)
}
```

#### Visualize
```{r}
bwp_sp_plots_global <- lapply(bwp_sp_models_global, function(x) {
  
  modobj = x$full_spline
  
  model_plot <- visualize_spline_model(
  modobj = modobj,
  plot_og_scale = TRUE,
  pnorm_smooth_var = TRUE,
  smooth_var = "bwp_fen",
  plabels = "Birth Weight Percentile",
  color_point = bwp_color,
  color_line = bwp_color)
  
  model_plot <- model_plot + 
    labs(x="Birth Weight Percentile)",
         title = glue("{modobj$gam$terms[[2]]}")) +
    theme(axis.title.y = element_blank())
})

bwp_sp_plots_global <- setNames(bwp_sp_plots_global, global_vars_totest)

if(save_figs){
saveRDS(bwp_sp_plots_global, glue("{modelvis_plot_folder}bwp_sp_plots_global.rds"))
}
```

#### Show plots
```{r, echo = TRUE}
wrap_plots(bwp_sp_plots_global[1:4], ncol = 2)
wrap_plots(bwp_sp_plots_global[5:7], ncol = 2)
```

# Regional phenotypes, test linear vs spline fit

## Volume

### GA

##### fit models
```{r, echo = FALSE}
ga_sp_models_vol <- lapply(vol_vars, function(x) {
  dat <- df.zscore
  if(grepl("Thick", x)) {
    dat <- df.zscore %>% filter(age_over2_adj == TRUE)
      if(median_th_mpr_flag == TRUE){
         form_og <- as.formula(glue("{x} ~ s(gestational_age) + euler_mean_mpr"))
         form_nolin <- as.formula(glue("{x} ~ s(gestational_age, m = c(2,0)) + gestational_age + euler_mean_mpr"))
        }
      } else {
         form_og <- as.formula(glue("{x} ~ s(gestational_age) + qc_var"))
         form_nolin <- as.formula(glue("{x} ~ s(gestational_age, m = c(2,0)) + gestational_age + qc_var"))
      }
  gam_model_og <- gamm(as.formula(form_og),
                       #random = list(participant_id = ~ 1),
              data = dat)
  gam_model_nolin <- gamm(as.formula(form_nolin),
                          #random = list(participant_id = ~ 1),
              data = dat)
  return(list(full_spline = gam_model_og,
         no_linear = gam_model_nolin)) })

ga_sp_models_vol <- setNames(ga_sp_models_vol, vol_vars)

ga_sp_gam_table_vol <- as.data.frame(t(sapply(ga_sp_models_vol,function(x) {
  summary_og <- summary(x$full_spline$gam)
  summary_nolin <- summary(x$no_linear$gam)
  c(ga_og_edf = summary_og$s.table[1,1], ga_og_p = summary_og$s.table[1,4],
    ga_nolin_edf = summary_nolin$s.table[1,1], ga_nolin_p = summary_nolin$s.table[1,4])
  })))

rownames(ga_sp_gam_table_vol) <- vol_vars

ga_sp_gam_table_vol$ga_og_p_adj <- p.adjust(ga_sp_gam_table_vol$ga_og_p, method = "BH")
ga_sp_gam_table_vol$ga_nolin_p_adj <- p.adjust(ga_sp_gam_table_vol$ga_nolin_p, method = "BH")

print(paste0("Number of significant IDPs for GA Smooth Sig: ", dim(ga_sp_gam_table_vol %>% filter(ga_nolin_p < 0.05))[1]))
print(paste0("Number of significant IDPs for GA Smooth Sig AFTER FDR correction: ", dim(ga_sp_gam_table_vol %>% filter(ga_nolin_p_adj < 0.05))[1]))
print(paste0("Number of IDPs with GA smooth sig but not above linear AFTER FDR correction: ", dim(ga_sp_gam_table_vol %>% filter(ga_og_p_adj < 0.05 & ga_nolin_p_adj > 0.05))[1]))

if(save_tables == TRUE){
  output_folder <- sp_stat_folder
  filename <- paste0(output_folder, "ga_sp_vol.csv")
  write.csv(ga_sp_gam_table_vol, filename)
}
```

##### Print Phenotypes
```{r}
print(paste0("Significant IDPs for GA Smooth Sig: ", rownames(ga_sp_gam_table_vol %>% filter(ga_nolin_p_adj < 0.05))))

print(paste0("GA smooth sig but not above linear AFTER FDR correction: ", rownames(ga_sp_gam_table_vol %>% filter(ga_og_p_adj < 0.05 & ga_nolin_p_adj > 0.05))))
```

##### add label column to match to ggseg
```{r, echo = FALSE}
cols_edit <- grepl("aseg(?!_CC)", rownames(ga_sp_gam_table_vol), perl = TRUE)

ga_sp_gam_table_vol$label <- rownames(ga_sp_gam_table_vol) %>% 
  sub("aparc_GrayVol_", "", .) %>% 
  sub("aseg_", "", .) %>%
  sub("(3|4)", "x\\1", .) %>%
  sub("rd_Ventricle", "rd_ventricle", .) %>%     sub("th_Ventricle", "th_ventricle", .) %>% sub("Thalamus", "Thalamus-Proper", .)

ga_sp_gam_table_vol$label[cols_edit] <- gsub("_", "-", ga_sp_gam_table_vol$label[cols_edit])
```

##### visualization on surface maps
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(ga_sp_gam_table_vol, ggseg::dk, by = "label")
plot_data_aseg <- merge(ga_sp_gam_table_vol, ggseg::aseg, by = "label")

# Regions not matched in atlases
(ga_sp_vol_noplot <- ga_sp_gam_table_vol[!(ga_sp_gam_table_vol$label %in% plot_data_dkt$label) & !(ga_sp_gam_table_vol$label %in% plot_data_aseg$label),])

# plots
(ga_sp_dkt_vol_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = ga_og_p_adj < 0.05)) +
  scale_fill_manual(values = c(
    "TRUE" = ga_color,
    "FALSE" = NA)) +
  theme_void()+
  labs(title = "GA, Spline Sig, Volume"))

(ga_splin_dkt_vol_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = ga_nolin_p_adj < 0.05)) +
  scale_fill_manual(values = c(
    "TRUE" = ga_color,
    "FALSE" = NA)) +
  theme_void()+
  labs(title = "GA, Spline Sig over Lin, Volume"))

(ga_sp_aseg_vol_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = ga_og_p_adj < 0.05)) +
  scale_fill_manual(values = c(
    "TRUE" = ga_color,
    "FALSE" = NA)) +
  theme_void() +
  labs(title = "GA, Spline Sig, Volume"))

(ga_splin_aseg_vol_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = ga_nolin_p_adj < 0.05)) +
  scale_fill_manual(values = c(
    "TRUE" = ga_color,
    "FALSE" = NA)) +
  theme_void() +
  labs(title = "GA, Spline Sig over Lin, Volume"))

if(save_figs == TRUE){
panel_save(ga_sp_dkt_vol_plot, "ga-sp-dkt-vol", "svg", ggseg_plot_folder)
panel_save(ga_splin_dkt_vol_plot, "ga-splin-dkt-vol", "svg", ggseg_plot_folder)
panel_save(ga_sp_aseg_vol_plot , "ga-sp-aseg-vol", "svg", ggseg_plot_folder)
panel_save(ga_splin_aseg_vol_plot , "ga-splin-aseg-vol", "svg", ggseg_plot_folder)
}
```

##### Visualize models themselves
```{r}
ga_sp_plots_vol <- lapply(ga_sp_models_vol, function(x) {
  
  modobj = x$full_spline
  
  model_plot <- visualize_spline_model(
  modobj = modobj,
  smooth_var = "gestational_age",
  plabels = "Gestational Age",
  color_point = ga_color,
  color_line = ga_color)
  
  model_plot <- model_plot + 
    labs(x="Gestational Age (weeks)",
         title = glue("{namemap[namemap$key == modobj$gam$terms[[2]], \"plot_names\"]}")) +
    theme(axis.title.y = element_blank()) +
  scale_x_continuous(breaks = sapply(c(24*7, 26*7, 28*7, 30*7, 32*7, 34*7, 36*7, 38*7, 40*7, 42*7), function(x){
    map2_dbl(x, "gestational_age", ~ (.x-mean_lookup_distinct[[.y]])/sd_lookup_distinct[[.y]]) }),
                     labels = c(24, 26, 28, 30, 32, 34, 36, 38, 40, 42)) +
  scale_y_continuous(labels = function(y) signif(pnorm(y)*100, 3))
})

ga_sp_plots_vol <- setNames(ga_sp_plots_vol, vol_vars)

if(save_figs){
saveRDS(ga_sp_plots_vol, glue("{modelvis_plot_folder}ga_sp_plots_vol.rds"))
}
```

##### Show plots
```{r, echo = TRUE}
wrapped_ga_sp_vol <- lapply(split(ga_sp_plots_vol, ceiling(seq_along(ga_sp_plots_vol) / 6)), function(p_list) {
  wrap_plots(p_list, ncol = 2)
})

lapply(wrapped_ga_sp_vol, print)
```


### BW
##### fit models
```{r, echo = FALSE}
bw_sp_models_vol <- lapply(vol_vars, function(x) {
  dat <- df.zscore
  if(grepl("Thick", x)) {
    dat <- df.zscore %>% filter(age_over2_adj == TRUE)
      if(median_th_mpr_flag == TRUE){
         form_og <- as.formula(glue("{x} ~ s(birth_weight_kg) + euler_mean_mpr"))
         form_nolin <- as.formula(glue("{x} ~ s(birth_weight_kg, m = c(2,0)) + birth_weight_kg + euler_mean_mpr"))
        }
      } else {
         form_og <- as.formula(glue("{x} ~ s(birth_weight_kg) + qc_var"))
         form_nolin <- as.formula(glue("{x} ~ s(birth_weight_kg, m = c(2,0)) + birth_weight_kg + qc_var"))
      }
  gam_model_og <- gamm(as.formula(form_og),
              data = dat)
  gam_model_nolin <- gamm(as.formula(form_nolin),
              data = dat)
  return(list(full_spline = gam_model_og,
         no_linear = gam_model_nolin)) })

bw_sp_models_vol <- setNames(bw_sp_models_vol, vol_vars)

bw_sp_gam_table_vol <- as.data.frame(t(sapply(bw_sp_models_vol,function(x) {
  summary_og <- summary(x$full_spline$gam)
  summary_nolin <- summary(x$no_linear$gam)
  c(bw_og_edf = summary_og$s.table[1,1], bw_og_p = summary_og$s.table[1,4],
    bw_nolin_edf = summary_nolin$s.table[1,1], bw_nolin_p = summary_nolin$s.table[1,4])
  })))

rownames(bw_sp_gam_table_vol) <- vol_vars

bw_sp_gam_table_vol$bw_og_p_adj <- p.adjust(bw_sp_gam_table_vol$bw_og_p, method = "BH")
bw_sp_gam_table_vol$bw_nolin_p_adj <- p.adjust(bw_sp_gam_table_vol$bw_nolin_p, method = "BH")

print(paste0("Number of significant IDPs for BW Smooth Sig: ", dim(bw_sp_gam_table_vol %>% filter(bw_nolin_p < 0.05))[1]))
print(paste0("Number of significant IDPs for BW Smooth Sig AFTER FDR correction: ", dim(bw_sp_gam_table_vol %>% filter(bw_nolin_p_adj < 0.05))[1]))
print(paste0("Number of IDPs with BW smooth sig but not above linear AFTER FDR correction: ", dim(bw_sp_gam_table_vol %>% filter(bw_og_p_adj < 0.05 & bw_nolin_p_adj > 0.05))[1]))

if(save_tables == TRUE){
  output_folder <- sp_stat_folder
  filename <- paste0(output_folder, "bw_sp_vol.csv")
  write.csv(bw_sp_gam_table_vol, filename)
}
```

##### Print Phenotypes
```{r}
print(paste0("Significant IDPs for BW Smooth Sig: ", rownames(bw_sp_gam_table_vol %>% filter(bw_nolin_p_adj < 0.05))))

print(paste0("BW smooth sig but not above linear AFTER FDR correction: ", rownames(bw_sp_gam_table_vol %>% filter(bw_og_p_adj < 0.05 & bw_nolin_p_adj > 0.05))))
```

##### add label column to match to ggseg
```{r, echo = FALSE}
cols_edit <- grepl("aseg(?!_CC)", rownames(bw_sp_gam_table_vol), perl = TRUE)

bw_sp_gam_table_vol$label <- rownames(bw_sp_gam_table_vol) %>% 
  sub("aparc_GrayVol_", "", .) %>% 
  sub("aseg_", "", .) %>%
  sub("(3|4)", "x\\1", .) %>%
  sub("rd_Ventricle", "rd_ventricle", .) %>%     sub("th_Ventricle", "th_ventricle", .) %>% sub("Thalamus", "Thalamus-Proper", .)

bw_sp_gam_table_vol$label[cols_edit] <- gsub("_", "-", bw_sp_gam_table_vol$label[cols_edit])
```

##### visualization on surface maps
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(bw_sp_gam_table_vol, ggseg::dk, by = "label")
plot_data_aseg <- merge(bw_sp_gam_table_vol, ggseg::aseg, by = "label")

# Regions not matched in atlases
(bw_sp_vol_noplot <- bw_sp_gam_table_vol[!(bw_sp_gam_table_vol$label %in% plot_data_dkt$label) & !(bw_sp_gam_table_vol$label %in% plot_data_aseg$label),])

# plots
(bw_sp_dkt_vol_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = bw_og_p_adj < 0.05)) +
  scale_fill_manual(values = c(
    "TRUE" = bw_color,
    "FALSE" = NA)) +
  theme_void()+
  labs(title = "BW, Spline Sig, Volume"))

(bw_splin_dkt_vol_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = bw_nolin_p_adj < 0.05)) +
  scale_fill_manual(values = c(
    "TRUE" = bw_color,
    "FALSE" = NA)) +
  theme_void()+
  labs(title = "BW, Spline Sig over Lin, Volume"))

(bw_sp_aseg_vol_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = bw_og_p_adj < 0.05)) +
  scale_fill_manual(values = c(
    "TRUE" = bw_color,
    "FALSE" = NA)) +
  theme_void() +
  labs(title = "BW, Spline Sig, Volume"))

(bw_splin_aseg_vol_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = bw_nolin_p_adj < 0.05)) +
  scale_fill_manual(values = c(
    "TRUE" = bw_color,
    "FALSE" = NA)) +
  theme_void() +
  labs(title = "BW, Spline Sig over Lin, Volume"))

if(save_figs == TRUE){
panel_save(bw_sp_dkt_vol_plot, "bw-sp-dkt-vol", "svg", ggseg_plot_folder)
panel_save(bw_splin_dkt_vol_plot, "bw-splin-dkt-vol", "svg", ggseg_plot_folder)
panel_save(bw_sp_aseg_vol_plot , "bw-sp-aseg-vol", "svg", ggseg_plot_folder)
panel_save(bw_splin_aseg_vol_plot , "bw-splin-aseg-vol", "svg", ggseg_plot_folder)
}
```

##### Visualize
```{r}
bw_sp_plots_vol <- lapply(bw_sp_models_vol, function(x) {
  
  modobj = x$full_spline
  
  # model_plot <- visualize_spline_model(
  # modobj = modobj,
  # smooth_var = "birth_weight_kg",
  # plabels = "Birth Weight",
  # color_point = bw_color,
  # color_line = bw_color)
  
model_plot <- visualize_spline_model(
modobj = modobj,
smooth_var = "birth_weight_kg",
mean_lookup = mean_lookup_distinct,
sd_lookup = sd_lookup_distinct,
pnorm_smooth_var = FALSE,ga
plabels = "Birth Weight",
color_point = bw_color,
color_line = bw_color)
  
  model_plot <- model_plot + 
    labs(x="Birth Weight (kg))",
         title = glue("{namemap[namemap$key == modobj$gam$terms[[2]], \"plot_names\"]}")) +
    theme(axis.title.y = element_blank())
  # scale_x_continuous(labels = function(x) round(map2_dbl(x, "birth_weight_kg", ~ (.x*sd_lookup_distinct[[.y]]) + mean_lookup_distinct[[.y]] ) ) ) +
  # scale_y_continuous(labels = function(y) signif(pnorm(y)*100, 3))
})

bw_sp_plots_vol <- setNames(bw_sp_plots_vol, vol_vars)

if(save_figs){
saveRDS(bw_sp_plots_vol, glue("{modelvis_plot_folder}bw_sp_plots_vol.rds"))
}
```
glue("{modobj$gamglue("{namemap[namemap$key == modobj$gam$terms[[2]], \"plot_names\"]}"))
##### Show plots
```{r, echo = TRUE}
wrapped_bw_sp_vol <- lapply(split(bw_sp_plots_vol, ceiling(seq_along(bw_sp_plots_vol) / 6)), function(p_list) {
  wrap_plots(p_list, ncol = 2)
})

lapply(wrapped_bw_sp_vol, print)
```


### BWP
##### fit models
```{r, echo = FALSE}
bwp_sp_models_vol <- lapply(vol_vars, function(x) {
  dat <- df.zscore
  if(grepl("Thick", x)) {
    dat <- df.zscore %>% filter(age_over2_adj == TRUE)
      if(median_th_mpr_flag == TRUE){
         form_og <- as.formula(glue("{x} ~ s(bwp_fen) + euler_mean_mpr"))
         form_nolin <- as.formula(glue("{x} ~ s(bwp_fen, m = c(2,0)) + bwp_fen + euler_mean_mpr"))
        }
      } else {
         form_og <- as.formula(glue("{x} ~ s(bwp_fen) + qc_var"))
         form_nolin <- as.formula(glue("{x} ~ s(bwp_fen, m = c(2,0)) + bwp_fen + qc_var"))
      }
  gam_model_og <- gamm(as.formula(form_og),
              data = dat)
  gam_model_nolin <- gamm(as.formula(form_nolin),
              data = dat)
  return(list(full_spline = gam_model_og,
         no_linear = gam_model_nolin)) })

bwp_sp_models_vol <- setNames(bwp_sp_models_vol, vol_vars)

bwp_sp_gam_table_vol <- as.data.frame(t(sapply(bwp_sp_models_vol,function(x) {
  summary_og <- summary(x$full_spline$gam)
  summary_nolin <- summary(x$no_linear$gam)
  c(bwp_og_edf = summary_og$s.table[1,1], bwp_og_p = summary_og$s.table[1,4],
    bwp_nolin_edf = summary_nolin$s.table[1,1], bwp_nolin_p = summary_nolin$s.table[1,4])
  })))

rownames(bwp_sp_gam_table_vol) <- vol_vars

bwp_sp_gam_table_vol$bwp_og_p_adj <- p.adjust(bwp_sp_gam_table_vol$bwp_og_p, method = "BH")
bwp_sp_gam_table_vol$bwp_nolin_p_adj <- p.adjust(bwp_sp_gam_table_vol$bwp_nolin_p, method = "BH")

print(paste0("Number of significant IDPs for BWP Smooth Sig: ", dim(bwp_sp_gam_table_vol %>% filter(bwp_nolin_p < 0.05))[1]))
print(paste0("Number of significant IDPs for BWP Smooth Sig AFTER FDR correction: ", dim(bwp_sp_gam_table_vol %>% filter(bwp_nolin_p_adj < 0.05))[1]))
print(paste0("Number of IDPs with BWP smooth sig but not above linear AFTER FDR correction: ", dim(bwp_sp_gam_table_vol %>% filter(bwp_og_p_adj < 0.05 & bwp_nolin_p_adj > 0.05))[1]))

if(save_tables == TRUE){
  output_folder <- sp_stat_folder
  filename <- paste0(output_folder, "bwp_sp_vol.csv")
  write.csv(bwp_sp_gam_table_vol, filename)
}
```

##### Print Phenotypes
```{r}
print(paste0("Significant IDPs for bwp Smooth Sig: ", rownames(bwp_sp_gam_table_vol %>% filter(bwp_nolin_p_adj < 0.05))))

print(paste0("bwp smooth sig but not above linear AFTER FDR correction: ", rownames(bwp_sp_gam_table_vol %>% filter(bwp_og_p_adj < 0.05 & bwp_nolin_p_adj > 0.05))))
```

##### add label column to match to ggseg
```{r, echo = FALSE}
cols_edit <- grepl("aseg(?!_CC)", rownames(bwp_sp_gam_table_vol), perl = TRUE)

bwp_sp_gam_table_vol$label <- rownames(bwp_sp_gam_table_vol) %>% 
  sub("aparc_GrayVol_", "", .) %>% 
  sub("aseg_", "", .) %>%
  sub("(3|4)", "x\\1", .) %>%
  sub("rd_Ventricle", "rd_ventricle", .) %>%     sub("th_Ventricle", "th_ventricle", .) %>% sub("Thalamus", "Thalamus-Proper", .)

bwp_sp_gam_table_vol$label[cols_edit] <- gsub("_", "-", bwp_sp_gam_table_vol$label[cols_edit])
```

##### visualization on surface maps
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(bwp_sp_gam_table_vol, ggseg::dk, by = "label")
plot_data_aseg <- merge(bwp_sp_gam_table_vol, ggseg::aseg, by = "label")

# Regions not matched in atlases
(bwp_sp_vol_noplot <- bwp_sp_gam_table_vol[!(bwp_sp_gam_table_vol$label %in% plot_data_dkt$label) & !(bwp_sp_gam_table_vol$label %in% plot_data_aseg$label),])

# plots
(bwp_sp_dkt_vol_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = bwp_og_p_adj < 0.05)) +
  scale_fill_manual(values = c(
    "TRUE" = bwp_color,
    "FALSE" = NA)) +
  theme_void()+
  labs(title = "BWP, Spline Sig, Volume"))

(bwp_splin_dkt_vol_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = bwp_nolin_p_adj < 0.05)) +
  scale_fill_manual(values = c(
    "TRUE" = bwp_color,
    "FALSE" = NA)) +
  theme_void()+
  labs(title = "BWP, Spline Sig over Lin, Volume"))

(bwp_sp_aseg_vol_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = bwp_og_p_adj < 0.05)) +
  scale_fill_manual(values = c(
    "TRUE" = bwp_color,
    "FALSE" = NA)) +
  theme_void() +
  labs(title = "BWP, Spline Sig, Volume"))

(bwp_splin_aseg_vol_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = bwp_nolin_p_adj < 0.05)) +
  scale_fill_manual(values = c(
    "TRUE" = bwp_color,
    "FALSE" = NA)) +
  theme_void() +
  labs(title = "BWP, Spline Sig over Lin, Volume"))

if(save_figs == TRUE){
panel_save(bwp_sp_dkt_vol_plot, "bwp-sp-dkt-vol", "svg", ggseg_plot_folder)
panel_save(bwp_splin_dkt_vol_plot, "bwp-splin-dkt-vol", "svg", ggseg_plot_folder)
panel_save(bwp_sp_aseg_vol_plot , "bwp-sp-aseg-vol", "svg", ggseg_plot_folder)
panel_save(bwp_splin_aseg_vol_plot , "bwp-splin-aseg-vol", "svg", ggseg_plot_folder)
}
```

##### Visualize
```{r}
bwp_sp_plots_vol <- lapply(bwp_sp_models_vol, function(x) {
  
  modobj = x$full_spline
  
  model_plot <- visualize_spline_model(
  modobj = modobj,
  smooth_var = "bwp_fen",
  plabels = "Birth Weight Percentile",
  color_point = bwp_color,
  color_line = bwp_color)
  
  model_plot <- model_plot + 
    labs(x="Birth Weight (kg))",
         title = glue("{namemap[namemap$key == modobj$gam$terms[[2]], \"plot_names\"]}")) +
    theme(axis.title.y = element_blank()) +
  scale_x_continuous(labels = function(x) signif(pnorm(x)*100, 2)) +
  scale_y_continuous(labels = function(y) signif(pnorm(y)*100, 3))
})

bwp_sp_plots_vol <- setNames(bwp_sp_plots_vol, vol_vars)

if(save_figs){
saveRDS(bwp_sp_plots_vol, glue("{modelvis_plot_folder}bwp_sp_plots_vol.rds"))
}
```

##### Show plots
```{r, echo = TRUE}
wrapped_bwp_sp_vol <- lapply(split(bwp_sp_plots_vol, ceiling(seq_along(bwp_sp_plots_vol) / 6)), function(p_list) {
  wrap_plots(p_list, ncol = 2)
})

lapply(wrapped_bwp_sp_vol, print)
```



## Surface Area

### GA

##### fit models
```{r, echo = FALSE}
ga_sp_models_sa <- lapply(sa_vars, function(x) {
  dat <- df.zscore
  if(grepl("Thick", x)) {
    dat <- df.zscore %>% filter(age_over2_adj == TRUE)
      if(median_th_mpr_flag == TRUE){
         form_og <- as.formula(glue("{x} ~ s(gestational_age) + euler_mean_mpr"))
         form_nolin <- as.formula(glue("{x} ~ s(gestational_age, m = c(2,0)) + gestational_age + euler_mean_mpr"))
        }
      } else {
         form_og <- as.formula(glue("{x} ~ s(gestational_age) + qc_var"))
         form_nolin <- as.formula(glue("{x} ~ s(gestational_age, m = c(2,0)) + gestational_age + qc_var"))
      }
  gam_model_og <- gamm(as.formula(form_og),
                       #random = list(participant_id = ~ 1),
              data = dat)
  gam_model_nolin <- gamm(as.formula(form_nolin),
                          #random = list(participant_id = ~ 1),
              data = dat)
  return(list(full_spline = gam_model_og,
         no_linear = gam_model_nolin)) })

ga_sp_models_sa <- setNames(ga_sp_models_sa, sa_vars)

ga_sp_gam_table_sa <- as.data.frame(t(sapply(ga_sp_models_sa,function(x) {
  summary_og <- summary(x$full_spline$gam)
  summary_nolin <- summary(x$no_linear$gam)
  c(ga_og_edf = summary_og$s.table[1,1], ga_og_p = summary_og$s.table[1,4],
    ga_nolin_edf = summary_nolin$s.table[1,1], ga_nolin_p = summary_nolin$s.table[1,4])
  })))

rownames(ga_sp_gam_table_sa) <- sa_vars

ga_sp_gam_table_sa$ga_og_p_adj <- p.adjust(ga_sp_gam_table_sa$ga_og_p, method = "BH")
ga_sp_gam_table_sa$ga_nolin_p_adj <- p.adjust(ga_sp_gam_table_sa$ga_nolin_p, method = "BH")

print(paste0("Number of significant IDPs for GA Smooth Sig: ", dim(ga_sp_gam_table_sa %>% filter(ga_nolin_p < 0.05))[1]))
print(paste0("Number of significant IDPs for GA Smooth Sig AFTER FDR correction: ", dim(ga_sp_gam_table_sa %>% filter(ga_nolin_p_adj < 0.05))[1]))
print(paste0("Number of IDPs with GA smooth sig but not above linear AFTER FDR correction: ", dim(ga_sp_gam_table_sa %>% filter(ga_og_p_adj < 0.05 & ga_nolin_p_adj > 0.05))[1]))

if(save_tables == TRUE){
  output_folder <- sp_stat_folder
  filename <- paste0(output_folder, "ga_sp_sa.csv")
  write.csv(ga_sp_gam_table_sa, filename)
}
```

##### Print Phenotypes
```{r}
print(paste0("Significant IDPs for GA Smooth Sig: ", rownames(ga_sp_gam_table_sa %>% filter(ga_nolin_p_adj < 0.05))))

print(paste0("GA smooth sig but not above linear AFTER FDR correction: ", rownames(ga_sp_gam_table_sa %>% filter(ga_og_p_adj < 0.05 & ga_nolin_p_adj > 0.05))))
```

##### add label column to match to ggseg
```{r, echo = FALSE}
cols_edit <- grepl("aseg(?!_CC)", rownames(ga_sp_gam_table_sa), perl = TRUE)

ga_sp_gam_table_sa$label <- rownames(ga_sp_gam_table_sa) %>% 
  sub("aparc_Graysa_", "", .) %>% 
  sub("aseg_", "", .) %>%
  sub("(3|4)", "x\\1", .) %>%
  sub("rd_Ventricle", "rd_ventricle", .) %>%     sub("th_Ventricle", "th_ventricle", .) %>% sub("Thalamus", "Thalamus-Proper", .)

ga_sp_gam_table_sa$label[cols_edit] <- gsub("_", "-", ga_sp_gam_table_sa$label[cols_edit])
```

##### visualization on surface maps
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(ga_sp_gam_table_sa, ggseg::dk, by = "label")
plot_data_aseg <- merge(ga_sp_gam_table_sa, ggseg::aseg, by = "label")

# Regions not matched in atlases
(ga_sp_sa_noplot <- ga_sp_gam_table_sa[!(ga_sp_gam_table_sa$label %in% plot_data_dkt$label) & !(ga_sp_gam_table_sa$label %in% plot_data_aseg$label),])

# plots
(ga_sp_dkt_sa_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = ga_og_p_adj < 0.05)) +
  scale_fill_manual(values = c(
    "TRUE" = ga_color,
    "FALSE" = NA)) +
  theme_void()+
  labs(title = "GA, Spline Sig, SA"))

(ga_splin_dkt_sa_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = ga_nolin_p_adj < 0.05)) +
  scale_fill_manual(values = c(
    "TRUE" = ga_color,
    "FALSE" = NA)) +
  theme_void()+
  labs(title = "GA, Spline Sig over Lin, SA"))

(ga_sp_aseg_sa_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = ga_og_p_adj < 0.05)) +
  scale_fill_manual(values = c(
    "TRUE" = ga_color,
    "FALSE" = NA)) +
  theme_void() +
  labs(title = "GA, Spline Sig, SA"))

(ga_splin_aseg_sa_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = ga_nolin_p_adj < 0.05)) +
  scale_fill_manual(values = c(
    "TRUE" = ga_color,
    "FALSE" = NA)) +
  theme_void() +
  labs(title = "GA, Spline Sig over Lin, SA"))

if(save_figs == TRUE){
panel_save(ga_sp_dkt_sa_plot, "ga-sp-dkt-sa", "svg", ggseg_plot_folder)
panel_save(ga_splin_dkt_sa_plot, "ga-splin-dkt-sa", "svg", ggseg_plot_folder)
panel_save(ga_sp_aseg_sa_plot , "ga-sp-aseg-sa", "svg", ggseg_plot_folder)
panel_save(ga_splin_aseg_sa_plot , "ga-splin-aseg-sa", "svg", ggseg_plot_folder)
}
```

##### Visualize models themselves
```{r}
ga_sp_plots_sa <- lapply(ga_sp_models_sa, function(x) {
  
  modobj = x$full_spline
  
  model_plot <- visualize_spline_model(
  modobj = modobj,
  smooth_var = "gestational_age",
  plabels = "Gestational Age",
  color_point = ga_color,
  color_line = ga_color)
  
  model_plot <- model_plot + 
    labs(x="Gestational Age (weeks)",
         title = glue("{namemap[namemap$key == modobj$gam$terms[[2]], \"plot_names\"]}")) +
    theme(axis.title.y = element_blank()) +
  scale_x_continuous(breaks = sapply(c(24*7, 26*7, 28*7, 30*7, 32*7, 34*7, 36*7, 38*7, 40*7, 42*7), function(x){
    map2_dbl(x, "gestational_age", ~ (.x-mean_lookup_distinct[[.y]])/sd_lookup_distinct[[.y]]) }),
                     labels = c(24, 26, 28, 30, 32, 34, 36, 38, 40, 42)) +
  scale_y_continuous(labels = function(y) signif(pnorm(y)*100, 3))
})

ga_sp_plots_sa <- setNames(ga_sp_plots_sa, sa_vars)

if(save_figs){
saveRDS(ga_sp_plots_sa, glue("{modelvis_plot_folder}ga_sp_plots_sa.rds"))
}
```

##### Show plots
```{r, echo = TRUE}
wrapped_ga_sp_sa <- lapply(split(ga_sp_plots_sa, ceiling(seq_along(ga_sp_plots_sa) / 6)), function(p_list) {
  wrap_plots(p_list, ncol = 2)
})

lapply(wrapped_ga_sp_sa, print)
```


### BW
##### fit models
```{r, echo = FALSE}
bw_sp_models_sa <- lapply(sa_vars, function(x) {
  dat <- df.zscore
  if(grepl("Thick", x)) {
    dat <- df.zscore %>% filter(age_over2_adj == TRUE)
      if(median_th_mpr_flag == TRUE){
         form_og <- as.formula(glue("{x} ~ s(birth_weight_kg) + euler_mean_mpr"))
         form_nolin <- as.formula(glue("{x} ~ s(birth_weight_kg, m = c(2,0)) + birth_weight_kg + euler_mean_mpr"))
        }
      } else {
         form_og <- as.formula(glue("{x} ~ s(birth_weight_kg) + qc_var"))
         form_nolin <- as.formula(glue("{x} ~ s(birth_weight_kg, m = c(2,0)) + birth_weight_kg + qc_var"))
      }
  gam_model_og <- gamm(as.formula(form_og),
              data = dat)
  gam_model_nolin <- gamm(as.formula(form_nolin),
              data = dat)
  return(list(full_spline = gam_model_og,
         no_linear = gam_model_nolin)) })

bw_sp_models_sa <- setNames(bw_sp_models_sa, sa_vars)

bw_sp_gam_table_sa <- as.data.frame(t(sapply(bw_sp_models_sa,function(x) {
  summary_og <- summary(x$full_spline$gam)
  summary_nolin <- summary(x$no_linear$gam)
  c(bw_og_edf = summary_og$s.table[1,1], bw_og_p = summary_og$s.table[1,4],
    bw_nolin_edf = summary_nolin$s.table[1,1], bw_nolin_p = summary_nolin$s.table[1,4])
  })))

rownames(bw_sp_gam_table_sa) <- sa_vars

bw_sp_gam_table_sa$bw_og_p_adj <- p.adjust(bw_sp_gam_table_sa$bw_og_p, method = "BH")
bw_sp_gam_table_sa$bw_nolin_p_adj <- p.adjust(bw_sp_gam_table_sa$bw_nolin_p, method = "BH")

print(paste0("Number of significant IDPs for BW Smooth Sig: ", dim(bw_sp_gam_table_sa %>% filter(bw_nolin_p < 0.05))[1]))
print(paste0("Number of significant IDPs for BW Smooth Sig AFTER FDR correction: ", dim(bw_sp_gam_table_sa %>% filter(bw_nolin_p_adj < 0.05))[1]))
print(paste0("Number of IDPs with BW smooth sig but not above linear AFTER FDR correction: ", dim(bw_sp_gam_table_sa %>% filter(bw_og_p_adj < 0.05 & bw_nolin_p_adj > 0.05))[1]))

if(save_tables == TRUE){
  output_folder <- sp_stat_folder
  filename <- paste0(output_folder, "bw_sp_sa.csv")
  write.csv(bw_sp_gam_table_sa, filename)
}
```

##### Print Phenotypes
```{r}
print(paste0("Significant IDPs for BW Smooth Sig: ", rownames(bw_sp_gam_table_sa %>% filter(bw_nolin_p_adj < 0.05))))

print(paste0("BW smooth sig but not above linear AFTER FDR correction: ", rownames(bw_sp_gam_table_sa %>% filter(bw_og_p_adj < 0.05 & bw_nolin_p_adj > 0.05))))
```

##### add label column to match to ggseg
```{r, echo = FALSE}
cols_edit <- grepl("aseg(?!_CC)", rownames(bw_sp_gam_table_sa), perl = TRUE)

bw_sp_gam_table_sa$label <- rownames(bw_sp_gam_table_sa) %>% 
  sub("aparc_Graysa_", "", .) %>% 
  sub("aseg_", "", .) %>%
  sub("(3|4)", "x\\1", .) %>%
  sub("rd_Ventricle", "rd_ventricle", .) %>%     sub("th_Ventricle", "th_ventricle", .) %>% sub("Thalamus", "Thalamus-Proper", .)

bw_sp_gam_table_sa$label[cols_edit] <- gsub("_", "-", bw_sp_gam_table_sa$label[cols_edit])
```

##### visualization on surface maps
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(bw_sp_gam_table_sa, ggseg::dk, by = "label")
plot_data_aseg <- merge(bw_sp_gam_table_sa, ggseg::aseg, by = "label")

# Regions not matched in atlases
(bw_sp_sa_noplot <- bw_sp_gam_table_sa[!(bw_sp_gam_table_sa$label %in% plot_data_dkt$label) & !(bw_sp_gam_table_sa$label %in% plot_data_aseg$label),])

# plots
(bw_sp_dkt_sa_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = bw_og_p_adj < 0.05)) +
  scale_fill_manual(values = c(
    "TRUE" = bw_color,
    "FALSE" = NA)) +
  theme_void()+
  labs(title = "BW, Spline Sig, SA"))

(bw_splin_dkt_sa_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = bw_nolin_p_adj < 0.05)) +
  scale_fill_manual(values = c(
    "TRUE" = bw_color,
    "FALSE" = NA)) +
  theme_void()+
  labs(title = "BW, Spline Sig over Lin, SA"))

(bw_sp_aseg_sa_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = bw_og_p_adj < 0.05)) +
  scale_fill_manual(values = c(
    "TRUE" = bw_color,
    "FALSE" = NA)) +
  theme_void() +
  labs(title = "BW, Spline Sig, SA"))

(bw_splin_aseg_sa_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = bw_nolin_p_adj < 0.05)) +
  scale_fill_manual(values = c(
    "TRUE" = bw_color,
    "FALSE" = NA)) +
  theme_void() +
  labs(title = "BW, Spline Sig over Lin, SA"))

if(save_figs == TRUE){
panel_save(bw_sp_dkt_sa_plot, "bw-sp-dkt-sa", "svg", ggseg_plot_folder)
panel_save(bw_splin_dkt_sa_plot, "bw-splin-dkt-sa", "svg", ggseg_plot_folder)
panel_save(bw_sp_aseg_sa_plot , "bw-sp-aseg-sa", "svg", ggseg_plot_folder)
panel_save(bw_splin_aseg_sa_plot , "bw-splin-aseg-sa", "svg", ggseg_plot_folder)
}
```

##### Visualize
```{r}
bw_sp_plots_sa <- lapply(bw_sp_models_sa, function(x) {
  
  modobj = x$full_spline
  
  # model_plot <- visualize_spline_model(
  # modobj = modobj,
  # smooth_var = "birth_weight_kg",
  # plabels = "Birth Weight",
  # color_point = bw_color,
  # color_line = bw_color)
  
model_plot <- visualize_spline_model(
modobj = modobj,
smooth_var = "birth_weight_kg",
mean_lookup = mean_lookup_distinct,
sd_lookup = sd_lookup_distinct,
pnorm_smooth_var = FALSE,
plabels = "Birth Weight",
color_point = bw_color,
color_line = bw_color)
  
  model_plot <- model_plot + 
    labs(x="Birth Weight (kg))",
         title = glue("{namemap[namemap$key == modobj$gam$terms[[2]], \"plot_names\"]}")) +
    theme(axis.title.y = element_blank())
  # scale_x_continuous(labels = function(x) round(map2_dbl(x, "birth_weight_kg", ~ (.x*sd_lookup_distinct[[.y]]) + mean_lookup_distinct[[.y]] ) ) ) +
  # scale_y_continuous(labels = function(y) signif(pnorm(y)*100, 3))
})

bw_sp_plots_sa <- setNames(bw_sp_plots_sa, sa_vars)

if(save_figs){
saveRDS(bw_sp_plots_sa, glue("{modelvis_plot_folder}bw_sp_plots_sa.rds"))
}
```
glue("{modobj$gamglue("{namemap[namemap$key == modobj$gam$terms[[2]], \"plot_names\"]}"))
##### Show plots
```{r, echo = TRUE}
wrapped_bw_sp_sa <- lapply(split(bw_sp_plots_sa, ceiling(seq_along(bw_sp_plots_sa) / 6)), function(p_list) {
  wrap_plots(p_list, ncol = 2)
})

lapply(wrapped_bw_sp_sa, print)
```


### BWP
##### fit models
```{r, echo = FALSE}
bwp_sp_models_sa <- lapply(sa_vars, function(x) {
  dat <- df.zscore
  if(grepl("Thick", x)) {
    dat <- df.zscore %>% filter(age_over2_adj == TRUE)
      if(median_th_mpr_flag == TRUE){
         form_og <- as.formula(glue("{x} ~ s(bwp_fen) + euler_mean_mpr"))
         form_nolin <- as.formula(glue("{x} ~ s(bwp_fen, m = c(2,0)) + bwp_fen + euler_mean_mpr"))
        }
      } else {
         form_og <- as.formula(glue("{x} ~ s(bwp_fen) + qc_var"))
         form_nolin <- as.formula(glue("{x} ~ s(bwp_fen, m = c(2,0)) + bwp_fen + qc_var"))
      }
  gam_model_og <- gamm(as.formula(form_og),
              data = dat)
  gam_model_nolin <- gamm(as.formula(form_nolin),
              data = dat)
  return(list(full_spline = gam_model_og,
         no_linear = gam_model_nolin)) })

bwp_sp_models_sa <- setNames(bwp_sp_models_sa, sa_vars)

bwp_sp_gam_table_sa <- as.data.frame(t(sapply(bwp_sp_models_sa,function(x) {
  summary_og <- summary(x$full_spline$gam)
  summary_nolin <- summary(x$no_linear$gam)
  c(bwp_og_edf = summary_og$s.table[1,1], bwp_og_p = summary_og$s.table[1,4],
    bwp_nolin_edf = summary_nolin$s.table[1,1], bwp_nolin_p = summary_nolin$s.table[1,4])
  })))

rownames(bwp_sp_gam_table_sa) <- sa_vars

bwp_sp_gam_table_sa$bwp_og_p_adj <- p.adjust(bwp_sp_gam_table_sa$bwp_og_p, method = "BH")
bwp_sp_gam_table_sa$bwp_nolin_p_adj <- p.adjust(bwp_sp_gam_table_sa$bwp_nolin_p, method = "BH")

print(paste0("Number of significant IDPs for BWP Smooth Sig: ", dim(bwp_sp_gam_table_sa %>% filter(bwp_nolin_p < 0.05))[1]))
print(paste0("Number of significant IDPs for BWP Smooth Sig AFTER FDR correction: ", dim(bwp_sp_gam_table_sa %>% filter(bwp_nolin_p_adj < 0.05))[1]))
print(paste0("Number of IDPs with BWP smooth sig but not above linear AFTER FDR correction: ", dim(bwp_sp_gam_table_sa %>% filter(bwp_og_p_adj < 0.05 & bwp_nolin_p_adj > 0.05))[1]))

if(save_tables == TRUE){
  output_folder <- sp_stat_folder
  filename <- paste0(output_folder, "bwp_sp_sa.csv")
  write.csv(bwp_sp_gam_table_sa, filename)
}
```

##### Print Phenotypes
```{r}
print(paste0("Significant IDPs for bwp Smooth Sig: ", rownames(bwp_sp_gam_table_sa %>% filter(bwp_nolin_p_adj < 0.05))))

print(paste0("bwp smooth sig but not above linear AFTER FDR correction: ", rownames(bwp_sp_gam_table_sa %>% filter(bwp_og_p_adj < 0.05 & bwp_nolin_p_adj > 0.05))))
```

##### add label column to match to ggseg
```{r, echo = FALSE}
cols_edit <- grepl("aseg(?!_CC)", rownames(bwp_sp_gam_table_sa), perl = TRUE)

bwp_sp_gam_table_sa$label <- rownames(bwp_sp_gam_table_sa) %>% 
  sub("aparc_Graysa_", "", .) %>% 
  sub("aseg_", "", .) %>%
  sub("(3|4)", "x\\1", .) %>%
  sub("rd_Ventricle", "rd_ventricle", .) %>%     sub("th_Ventricle", "th_ventricle", .) %>% sub("Thalamus", "Thalamus-Proper", .)

bwp_sp_gam_table_sa$label[cols_edit] <- gsub("_", "-", bwp_sp_gam_table_sa$label[cols_edit])
```

##### visualization on surface maps
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(bwp_sp_gam_table_sa, ggseg::dk, by = "label")
plot_data_aseg <- merge(bwp_sp_gam_table_sa, ggseg::aseg, by = "label")

# Regions not matched in atlases
(bwp_sp_sa_noplot <- bwp_sp_gam_table_sa[!(bwp_sp_gam_table_sa$label %in% plot_data_dkt$label) & !(bwp_sp_gam_table_sa$label %in% plot_data_aseg$label),])

# plots
(bwp_sp_dkt_sa_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = bwp_og_p_adj < 0.05)) +
  scale_fill_manual(values = c(
    "TRUE" = bwp_color,
    "FALSE" = NA)) +
  theme_void()+
  labs(title = "BWP, Spline Sig, SA"))

(bwp_splin_dkt_sa_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = bwp_nolin_p_adj < 0.05)) +
  scale_fill_manual(values = c(
    "TRUE" = bwp_color,
    "FALSE" = NA)) +
  theme_void()+
  labs(title = "BWP, Spline Sig over Lin, SA"))

(bwp_sp_aseg_sa_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = bwp_og_p_adj < 0.05)) +
  scale_fill_manual(values = c(
    "TRUE" = bwp_color,
    "FALSE" = NA)) +
  theme_void() +
  labs(title = "BWP, Spline Sig, SA"))

(bwp_splin_aseg_sa_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = bwp_nolin_p_adj < 0.05)) +
  scale_fill_manual(values = c(
    "TRUE" = bwp_color,
    "FALSE" = NA)) +
  theme_void() +
  labs(title = "BWP, Spline Sig over Lin, SA"))

if(save_figs == TRUE){
panel_save(bwp_sp_dkt_sa_plot, "bwp-sp-dkt-sa", "svg", ggseg_plot_folder)
panel_save(bwp_splin_dkt_sa_plot, "bwp-splin-dkt-sa", "svg", ggseg_plot_folder)
panel_save(bwp_sp_aseg_sa_plot , "bwp-sp-aseg-sa", "svg", ggseg_plot_folder)
panel_save(bwp_splin_aseg_sa_plot , "bwp-splin-aseg-sa", "svg", ggseg_plot_folder)
}
```

##### Visualize
```{r}
bwp_sp_plots_sa <- lapply(bwp_sp_models_sa, function(x) {
  
  modobj = x$full_spline
  
  model_plot <- visualize_spline_model(
  modobj = modobj,
  smooth_var = "bwp_fen",
  plabels = "Birth Weight Percentile",
  color_point = bwp_color,
  color_line = bwp_color)
  
  model_plot <- model_plot + 
    labs(x="Birth Weight (kg))",
         title = glue("{namemap[namemap$key == modobj$gam$terms[[2]], \"plot_names\"]}")) +
    theme(axis.title.y = element_blank()) +
  scale_x_continuous(labels = function(x) signif(pnorm(x)*100, 2)) +
  scale_y_continuous(labels = function(y) signif(pnorm(y)*100, 3))
})

bwp_sp_plots_sa <- setNames(bwp_sp_plots_sa, sa_vars)

if(save_figs){
saveRDS(bwp_sp_plots_sa, glue("{modelvis_plot_folder}bwp_sp_plots_sa.rds"))
}
```

##### Show plots
```{r, echo = TRUE}
wrapped_bwp_sp_sa <- lapply(split(bwp_sp_plots_sa, ceiling(seq_along(bwp_sp_plots_sa) / 6)), function(p_list) {
  wrap_plots(p_list, ncol = 2)
})

lapply(wrapped_bwp_sp_sa, print)
```



## Thickness

### GA

##### fit models
```{r, echo = FALSE}
ga_sp_models_th <- lapply(th_vars, function(x) {
  dat <- df.zscore
  if(grepl("Thick", x)) {
    dat <- df.zscore %>% filter(age_over2_adj == TRUE)
      if(median_th_mpr_flag == TRUE){
         form_og <- as.formula(glue("{x} ~ s(gestational_age) + euler_mean_mpr"))
         form_nolin <- as.formula(glue("{x} ~ s(gestational_age, m = c(2,0)) + gestational_age + euler_mean_mpr"))
        }
      } else {
         form_og <- as.formula(glue("{x} ~ s(gestational_age) + qc_var"))
         form_nolin <- as.formula(glue("{x} ~ s(gestational_age, m = c(2,0)) + gestational_age + qc_var"))
      }
  gam_model_og <- gamm(as.formula(form_og),
                       #random = list(participant_id = ~ 1),
              data = dat)
  gam_model_nolin <- gamm(as.formula(form_nolin),
                          #random = list(participant_id = ~ 1),
              data = dat)
  return(list(full_spline = gam_model_og,
         no_linear = gam_model_nolin)) })

ga_sp_models_th <- setNames(ga_sp_models_th, th_vars)

ga_sp_gam_table_th <- as.data.frame(t(sapply(ga_sp_models_th,function(x) {
  summary_og <- summary(x$full_spline$gam)
  summary_nolin <- summary(x$no_linear$gam)
  c(ga_og_edf = summary_og$s.table[1,1], ga_og_p = summary_og$s.table[1,4],
    ga_nolin_edf = summary_nolin$s.table[1,1], ga_nolin_p = summary_nolin$s.table[1,4])
  })))

rownames(ga_sp_gam_table_th) <- th_vars

ga_sp_gam_table_th$ga_og_p_adj <- p.adjust(ga_sp_gam_table_th$ga_og_p, method = "BH")
ga_sp_gam_table_th$ga_nolin_p_adj <- p.adjust(ga_sp_gam_table_th$ga_nolin_p, method = "BH")

print(paste0("Number of significant IDPs for GA Smooth Sig: ", dim(ga_sp_gam_table_th %>% filter(ga_nolin_p < 0.05))[1]))
print(paste0("Number of significant IDPs for GA Smooth Sig AFTER FDR correction: ", dim(ga_sp_gam_table_th %>% filter(ga_nolin_p_adj < 0.05))[1]))
print(paste0("Number of IDPs with GA smooth sig but not above linear AFTER FDR correction: ", dim(ga_sp_gam_table_th %>% filter(ga_og_p_adj < 0.05 & ga_nolin_p_adj > 0.05))[1]))

if(save_tables == TRUE){
  output_folder <- sp_stat_folder
  filename <- paste0(output_folder, "ga_sp_th.csv")
  write.csv(ga_sp_gam_table_th, filename)
}
```

##### Print Phenotypes
```{r}
print(paste0("Significant IDPs for GA Smooth Sig: ", rownames(ga_sp_gam_table_th %>% filter(ga_nolin_p_adj < 0.05))))

print(paste0("GA smooth sig but not above linear AFTER FDR correction: ", rownames(ga_sp_gam_table_th %>% filter(ga_og_p_adj < 0.05 & ga_nolin_p_adj > 0.05))))
```

##### add label column to match to ggseg
```{r, echo = FALSE}
cols_edit <- grepl("aseg(?!_CC)", rownames(ga_sp_gam_table_th), perl = TRUE)

ga_sp_gam_table_th$label <- rownames(ga_sp_gam_table_th) %>% 
  sub("aparc_ThickAvg_", "", .) %>% 
  sub("aseg_", "", .) %>%
  sub("(3|4)", "x\\1", .) %>%
  sub("rd_Ventricle", "rd_ventricle", .) %>%     sub("th_Ventricle", "th_ventricle", .) %>% sub("Thalamus", "Thalamus-Proper", .)

ga_sp_gam_table_th$label[cols_edit] <- gsub("_", "-", ga_sp_gam_table_th$label[cols_edit])
```

##### visualization on surface maps
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(ga_sp_gam_table_th, ggseg::dk, by = "label")

# Regions not matched in atlases
(ga_sp_th_noplot <- ga_sp_gam_table_th[!(ga_sp_gam_table_th$label %in% plot_data_dkt$label),])

# plots
(ga_sp_dkt_th_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = ga_og_p_adj < 0.05)) +
  scale_fill_manual(values = c(
    "TRUE" = ga_color,
    "FALSE" = NA)) +
  theme_void()+
  labs(title = "GA, Spline Sig, th"))

(ga_splin_dkt_th_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = ga_nolin_p_adj < 0.05)) +
  scale_fill_manual(values = c(
    "TRUE" = ga_color,
    "FALSE" = NA)) +
  theme_void()+
  labs(title = "GA, Spline Sig over Lin, th"))

if(save_figs == TRUE){
panel_save(ga_sp_dkt_th_plot, "ga-sp-dkt-th", "svg", ggseg_plot_folder)
panel_save(ga_splin_dkt_th_plot, "ga-splin-dkt-th", "svg", ggseg_plot_folder)
}
```

##### Visualize models themselves
```{r}
ga_sp_plots_th <- lapply(ga_sp_models_th[rownames(ga_sp_gam_table_th %>% filter(ga_og_p_adj < 0.05))], function(x) {
  
  modobj = x$full_spline
  
  model_plot <- visualize_spline_model(
  modobj = modobj,
  plot_og_scale = TRUE,
  mean_lookup = mean_lookup_distinct,
  sd_lookup = sd_lookup_distinct,
  pnorm_smooth_var = FALSE,
  smooth_var = "gestational_age",
  plabels = "Gestational Age",
  color_point = ga_color,
  color_line = ga_color)
  
  model_plot <- model_plot + 
    labs(x="Gestational Age (weeks)",
         title = glue("{namemap[namemap$key == modobj$gam$terms[[2]], \"plot_names\"]}")) +
    theme(axis.title.y = element_blank())
  # scale_x_continuous(breaks = sapply(c(24*7, 26*7, 28*7, 30*7, 32*7, 34*7, 36*7, 38*7, 40*7, 42*7), function(x){
  #   map2_dbl(x, "gestational_age", ~ (.x-mean_lookup_distinct[[.y]])/sd_lookup_distinct[[.y]]) }),
  #                    labels = c(24, 26, 28, 30, 32, 34, 36, 38, 40, 42)) +
  # scale_y_continuous(labels = function(y) signif(pnorm(y)*100, 3))
})

ga_sp_plots_th <- setNames(ga_sp_plots_th, rownames(ga_sp_gam_table_th %>% filter(ga_og_p_adj < 0.05)))

if(save_figs){
saveRDS(ga_sp_plots_th, glue("{modelvis_plot_folder}ga_sp_plots_th.rds"))
}
```

##### Show plots
```{r, echo = TRUE}
wrapped_ga_sp_th <- lapply(split(ga_sp_plots_th, ceiling(seq_along(ga_sp_plots_th) / 6)), function(p_list) {
  wrap_plots(p_list, ncol = 2)
})

lapply(wrapped_ga_sp_th, print)
```


### BW
##### fit models
```{r, echo = FALSE}
bw_sp_models_th <- lapply(th_vars, function(x) {
  dat <- df.zscore
  if(grepl("Thick", x)) {
    dat <- df.zscore %>% filter(age_over2_adj == TRUE)
      if(median_th_mpr_flag == TRUE){
         form_og <- as.formula(glue("{x} ~ s(birth_weight_kg) + euler_mean_mpr"))
         form_nolin <- as.formula(glue("{x} ~ s(birth_weight_kg, m = c(2,0)) + birth_weight_kg + euler_mean_mpr"))
        }
      } else {
         form_og <- as.formula(glue("{x} ~ s(birth_weight_kg) + qc_var"))
         form_nolin <- as.formula(glue("{x} ~ s(birth_weight_kg, m = c(2,0)) + birth_weight_kg + qc_var"))
      }
  gam_model_og <- gamm(as.formula(form_og),
              data = dat)
  gam_model_nolin <- gamm(as.formula(form_nolin),
              data = dat)
  return(list(full_spline = gam_model_og,
         no_linear = gam_model_nolin)) })

bw_sp_models_th <- setNames(bw_sp_models_th, th_vars)

bw_sp_gam_table_th <- as.data.frame(t(sapply(bw_sp_models_th,function(x) {
  summary_og <- summary(x$full_spline$gam)
  summary_nolin <- summary(x$no_linear$gam)
  c(bw_og_edf = summary_og$s.table[1,1], bw_og_p = summary_og$s.table[1,4],
    bw_nolin_edf = summary_nolin$s.table[1,1], bw_nolin_p = summary_nolin$s.table[1,4])
  })))

rownames(bw_sp_gam_table_th) <- th_vars

bw_sp_gam_table_th$bw_og_p_adj <- p.adjust(bw_sp_gam_table_th$bw_og_p, method = "BH")
bw_sp_gam_table_th$bw_nolin_p_adj <- p.adjust(bw_sp_gam_table_th$bw_nolin_p, method = "BH")

print(paste0("Number of significant IDPs for BW Smooth Sig: ", dim(bw_sp_gam_table_th %>% filter(bw_nolin_p < 0.05))[1]))
print(paste0("Number of significant IDPs for BW Smooth Sig AFTER FDR correction: ", dim(bw_sp_gam_table_th %>% filter(bw_nolin_p_adj < 0.05))[1]))
print(paste0("Number of IDPs with BW smooth sig but not above linear AFTER FDR correction: ", dim(bw_sp_gam_table_th %>% filter(bw_og_p_adj < 0.05 & bw_nolin_p_adj > 0.05))[1]))

if(save_tables == TRUE){
  output_folder <- sp_stat_folder
  filename <- paste0(output_folder, "bw_sp_th.csv")
  write.csv(bw_sp_gam_table_th, filename)
}
```

##### Print Phenotypes
```{r}
print(paste0("Significant IDPs for BW Smooth Sig: ", rownames(bw_sp_gam_table_th %>% filter(bw_nolin_p_adj < 0.05))))

print(paste0("BW smooth sig but not above linear AFTER FDR correction: ", rownames(bw_sp_gam_table_th %>% filter(bw_og_p_adj < 0.05 & bw_nolin_p_adj > 0.05))))
```

##### add label column to match to ggseg
```{r, echo = FALSE}
cols_edit <- grepl("aseg(?!_CC)", rownames(bw_sp_gam_table_th), perl = TRUE)

bw_sp_gam_table_th$label <- rownames(bw_sp_gam_table_th) %>% 
  sub("aparc_Grayth_", "", .) %>% 
  sub("aseg_", "", .) %>%
  sub("(3|4)", "x\\1", .) %>%
  sub("rd_Ventricle", "rd_ventricle", .) %>%     sub("th_Ventricle", "th_ventricle", .) %>% sub("Thalamus", "Thalamus-Proper", .)

bw_sp_gam_table_th$label[cols_edit] <- gsub("_", "-", bw_sp_gam_table_th$label[cols_edit])
```

##### visualization on surface maps
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(bw_sp_gam_table_th, ggseg::dk, by = "label")
plot_data_aseg <- merge(bw_sp_gam_table_th, ggseg::aseg, by = "label")

# Regions not matched in atlases
(bw_sp_th_noplot <- bw_sp_gam_table_th[!(bw_sp_gam_table_th$label %in% plot_data_dkt$label) & !(bw_sp_gam_table_th$label %in% plot_data_aseg$label),])

# plots
(bw_sp_dkt_th_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = bw_og_p_adj < 0.05)) +
  scale_fill_manual(values = c(
    "TRUE" = bw_color,
    "FALSE" = NA)) +
  theme_void()+
  labs(title = "BW, Spline Sig, th"))

(bw_splin_dkt_th_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = bw_nolin_p_adj < 0.05)) +
  scale_fill_manual(values = c(
    "TRUE" = bw_color,
    "FALSE" = NA)) +
  theme_void()+
  labs(title = "BW, Spline Sig over Lin, th"))

(bw_sp_aseg_th_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = bw_og_p_adj < 0.05)) +
  scale_fill_manual(values = c(
    "TRUE" = bw_color,
    "FALSE" = NA)) +
  theme_void() +
  labs(title = "BW, Spline Sig, th"))

(bw_splin_aseg_th_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = bw_nolin_p_adj < 0.05)) +
  scale_fill_manual(values = c(
    "TRUE" = bw_color,
    "FALSE" = NA)) +
  theme_void() +
  labs(title = "BW, Spline Sig over Lin, th"))

if(save_figs == TRUE){
panel_save(bw_sp_dkt_th_plot, "bw-sp-dkt-th", "svg", ggseg_plot_folder)
panel_save(bw_splin_dkt_th_plot, "bw-splin-dkt-th", "svg", ggseg_plot_folder)
panel_save(bw_sp_aseg_th_plot , "bw-sp-aseg-th", "svg", ggseg_plot_folder)
panel_save(bw_splin_aseg_th_plot , "bw-splin-aseg-th", "svg", ggseg_plot_folder)
}
```

##### Visualize
```{r}
bw_sp_plots_th <- lapply(bw_sp_models_th, function(x) {
  
  modobj = x$full_spline
  
  # model_plot <- visualize_spline_model(
  # modobj = modobj,
  # smooth_var = "birth_weight_kg",
  # plabels = "Birth Weight",
  # color_point = bw_color,
  # color_line = bw_color)
  
model_plot <- visualize_spline_model(
modobj = modobj,
smooth_var = "birth_weight_kg",
mean_lookup = mean_lookup_distinct,
sd_lookup = sd_lookup_distinct,
pnorm_smooth_var = FALSE,
plabels = "Birth Weight",
color_point = bw_color,
color_line = bw_color)
  
  model_plot <- model_plot + 
    labs(x="Birth Weight (kg))",
         title = glue("{namemap[namemap$key == modobj$gam$terms[[2]], \"plot_names\"]}")) +
    theme(axis.title.y = element_blank())
  # scale_x_continuous(labels = function(x) round(map2_dbl(x, "birth_weight_kg", ~ (.x*sd_lookup_distinct[[.y]]) + mean_lookup_distinct[[.y]] ) ) ) +
  # scale_y_continuous(labels = function(y) signif(pnorm(y)*100, 3))
})

bw_sp_plots_th <- setNames(bw_sp_plots_th, th_vars)

if(save_figs){
saveRDS(bw_sp_plots_th, glue("{modelvis_plot_folder}bw_sp_plots_th.rds"))
}
```
glue("{modobj$gamg lue("{namemap[namemap$key == modobj$gam$terms[[2]], \"plot_names\"]}"))
##### Show plots
```{r, echo = TRUE}
wrapped_bw_sp_th <- lapply(split(bw_sp_plots_th, ceiling(seq_along(bw_sp_plots_th) / 6)), function(p_list) {
  wrap_plots(p_list, ncol = 2)
})

lapply(wrapped_bw_sp_th, print)
```


### BWP
##### fit models
```{r, echo = FALSE}
bwp_sp_models_th <- lapply(th_vars, function(x) {
  dat <- df.zscore
  if(grepl("Thick", x)) {
    dat <- df.zscore %>% filter(age_over2_adj == TRUE)
      if(median_th_mpr_flag == TRUE){
         form_og <- as.formula(glue("{x} ~ s(bwp_fen) + euler_mean_mpr"))
         form_nolin <- as.formula(glue("{x} ~ s(bwp_fen, m = c(2,0)) + bwp_fen + euler_mean_mpr"))
        }
      } else {
         form_og <- as.formula(glue("{x} ~ s(bwp_fen) + qc_var"))
         form_nolin <- as.formula(glue("{x} ~ s(bwp_fen, m = c(2,0)) + bwp_fen + qc_var"))
      }
  gam_model_og <- gamm(as.formula(form_og),
              data = dat)
  gam_model_nolin <- gamm(as.formula(form_nolin),
              data = dat)
  return(list(full_spline = gam_model_og,
         no_linear = gam_model_nolin)) })

bwp_sp_models_th <- setNames(bwp_sp_models_th, th_vars)

bwp_sp_gam_table_th <- as.data.frame(t(sapply(bwp_sp_models_th,function(x) {
  summary_og <- summary(x$full_spline$gam)
  summary_nolin <- summary(x$no_linear$gam)
  c(bwp_og_edf = summary_og$s.table[1,1], bwp_og_p = summary_og$s.table[1,4],
    bwp_nolin_edf = summary_nolin$s.table[1,1], bwp_nolin_p = summary_nolin$s.table[1,4])
  })))

rownames(bwp_sp_gam_table_th) <- th_vars

bwp_sp_gam_table_th$bwp_og_p_adj <- p.adjust(bwp_sp_gam_table_th$bwp_og_p, method = "BH")
bwp_sp_gam_table_th$bwp_nolin_p_adj <- p.adjust(bwp_sp_gam_table_th$bwp_nolin_p, method = "BH")

print(paste0("Number of significant IDPs for BWP Smooth Sig: ", dim(bwp_sp_gam_table_th %>% filter(bwp_nolin_p < 0.05))[1]))
print(paste0("Number of significant IDPs for BWP Smooth Sig AFTER FDR correction: ", dim(bwp_sp_gam_table_th %>% filter(bwp_nolin_p_adj < 0.05))[1]))
print(paste0("Number of IDPs with BWP smooth sig but not above linear AFTER FDR correction: ", dim(bwp_sp_gam_table_th %>% filter(bwp_og_p_adj < 0.05 & bwp_nolin_p_adj > 0.05))[1]))

if(save_tables == TRUE){
  output_folder <- sp_stat_folder
  filename <- paste0(output_folder, "bwp_sp_th.csv")
  write.csv(bwp_sp_gam_table_th, filename)
}
```

##### Print Phenotypes
```{r}
print(paste0("Significant IDPs for bwp Smooth Sig: ", rownames(bwp_sp_gam_table_th %>% filter(bwp_nolin_p_adj < 0.05))))

print(paste0("bwp smooth sig but not above linear AFTER FDR correction: ", rownames(bwp_sp_gam_table_th %>% filter(bwp_og_p_adj < 0.05 & bwp_nolin_p_adj > 0.05))))
```

##### add label column to match to ggseg
```{r, echo = FALSE}
cols_edit <- grepl("aseg(?!_CC)", rownames(bwp_sp_gam_table_th), perl = TRUE)

bwp_sp_gam_table_th$label <- rownames(bwp_sp_gam_table_th) %>% 
  sub("aparc_Grayth_", "", .) %>% 
  sub("aseg_", "", .) %>%
  sub("(3|4)", "x\\1", .) %>%
  sub("rd_Ventricle", "rd_ventricle", .) %>%     sub("th_Ventricle", "th_ventricle", .) %>% sub("Thalamus", "Thalamus-Proper", .)

bwp_sp_gam_table_th$label[cols_edit] <- gsub("_", "-", bwp_sp_gam_table_th$label[cols_edit])
```

##### visualization on surface maps
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(bwp_sp_gam_table_th, ggseg::dk, by = "label")
plot_data_aseg <- merge(bwp_sp_gam_table_th, ggseg::aseg, by = "label")

# Regions not matched in atlases
(bwp_sp_th_noplot <- bwp_sp_gam_table_th[!(bwp_sp_gam_table_th$label %in% plot_data_dkt$label) & !(bwp_sp_gam_table_th$label %in% plot_data_aseg$label),])

# plots
(bwp_sp_dkt_th_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = bwp_og_p_adj < 0.05)) +
  scale_fill_manual(values = c(
    "TRUE" = bwp_color,
    "FALSE" = NA)) +
  theme_void()+
  labs(title = "BWP, Spline Sig, th"))

(bwp_splin_dkt_th_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = bwp_nolin_p_adj < 0.05)) +
  scale_fill_manual(values = c(
    "TRUE" = bwp_color,
    "FALSE" = NA)) +
  theme_void()+
  labs(title = "BWP, Spline Sig over Lin, th"))

(bwp_sp_aseg_th_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = bwp_og_p_adj < 0.05)) +
  scale_fill_manual(values = c(
    "TRUE" = bwp_color,
    "FALSE" = NA)) +
  theme_void() +
  labs(title = "BWP, Spline Sig, th"))

(bwp_splin_aseg_th_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = bwp_nolin_p_adj < 0.05)) +
  scale_fill_manual(values = c(
    "TRUE" = bwp_color,
    "FALSE" = NA)) +
  theme_void() +
  labs(title = "BWP, Spline Sig over Lin, th"))

if(save_figs == TRUE){
panel_save(bwp_sp_dkt_th_plot, "bwp-sp-dkt-th", "svg", ggseg_plot_folder)
panel_save(bwp_splin_dkt_th_plot, "bwp-splin-dkt-th", "svg", ggseg_plot_folder)
panel_save(bwp_sp_aseg_th_plot , "bwp-sp-aseg-th", "svg", ggseg_plot_folder)
panel_save(bwp_splin_aseg_th_plot , "bwp-splin-aseg-th", "svg", ggseg_plot_folder)
}
```

##### Visualize
```{r}
bwp_sp_plots_th <- lapply(bwp_sp_models_th, function(x) {
  
  modobj = x$full_spline
  
  model_plot <- visualize_spline_model(
  modobj = modobj,
  smooth_var = "bwp_fen",
  plabels = "Birth Weight Percentile",
  color_point = bwp_color,
  color_line = bwp_color)
  
  model_plot <- model_plot + 
    labs(x="Birth Weight (kg))",
         title = glue("{namemap[namemap$key == modobj$gam$terms[[2]], \"plot_names\"]}")) +
    theme(axis.title.y = element_blank()) +
  scale_x_continuous(labels = function(x) signif(pnorm(x)*100, 2)) +
  scale_y_continuous(labels = function(y) signif(pnorm(y)*100, 3))
})

bwp_sp_plots_th <- setNames(bwp_sp_plots_th, th_vars)

if(save_figs){
saveRDS(bwp_sp_plots_th, glue("{modelvis_plot_folder}bwp_sp_plots_th.rds"))
}
```

##### Show plots
```{r, echo = TRUE}
wrapped_bwp_sp_th <- lapply(split(bwp_sp_plots_th, ceiling(seq_along(bwp_sp_plots_th) / 6)), function(p_list) {
  wrap_plots(p_list, ncol = 2)
})

lapply(wrapped_bwp_sp_th, print)
```


# Regional phenotypes (non-linear models, with interaction term test)

## Volume

### GA

#### fit models
```{r, echo = FALSE}
gabyage_sp_models_vol <- lapply(vol_vars, function(x) {
  form <- as.formula(glue("{x} ~ s(gestational_age) + s(age_days_adj) + ti(gestational_age, age_days_adj) + qc_var"))
  gam_model <- gamm(as.formula(form),
              data = df.zscore)
  return(gam_model)
  })

gabyage_sp_models_vol <- setNames(gabyage_sp_models_vol, vol_vars)

gabyage_sp_gam_table_vol <- as.data.frame(t(sapply(gabyage_sp_models_vol,function(x) {
  summary <- summary(x$gam)
  c(ga_edf = summary$s.table[1,1], ga_p = summary$s.table[1,4],
    age_edf = summary$s.table[2,1], age_p = summary$s.table[2,4],
    gabyage_edf = summary$s.table[3,1], gabyage_p = summary$s.table[3,4])
  })))

rownames(gabyage_sp_gam_table_vol) <- vol_vars

gabyage_sp_gam_table_vol$gabyage_p_adj <- p.adjust(gabyage_sp_gam_table_vol$gabyage_p, method = "BH")

print(paste0("Number of significant IDPs for GA x Age- NO correction: ", dim(gabyage_sp_gam_table_vol %>% filter(gabyage_p < 0.05))[1]))
print(paste0("Number of significant IDPs for GA x Age- AFTER FDR correction: ", dim(gabyage_sp_gam_table_vol %>% filter(gabyage_p_adj < 0.05))[1]))
print(paste0("Significant IDPs for GA x Age- AFTER FDR correction: ", rownames(gabyage_sp_gam_table_vol %>% filter(gabyage_p_adj < 0.05)) ))

gabyage_sp_vol_sig_pheno <- rownames(gabyage_sp_gam_table_vol %>% filter(gabyage_p_adj < 0.05))
```

#### Visualize models
```{r}
gabyage_sp_vol_plot_pheno <- rownames(gabyage_sp_gam_table_vol %>% filter(ga_edf > 2))
font_size <- 16
theme_set(theme_classic(base_family = "sans",base_size = font_size))
line_size <- 1.5
point_size <- 2

mu_age <- mean(cent_all$age_days_adj, na.rm = TRUE)
sd_age <- sd(cent_all$age_days_adj, na.rm = TRUE)

mu_ga <- mean(cent_all$gestational_age, na.rm = TRUE)
sd_ga <- sd(cent_all$gestational_age, na.rm = TRUE)

gabyage_sp_plots_vol <- lapply(gabyage_sp_models_vol[gabyage_sp_vol_plot_pheno], function(x) {
  
  
  low_age_label <- (quantile(df.zscore[!is.na(df.zscore$gestational_age),
                                       "age_days_adj"],
                             probs = c(.10)) * sd_age + mu_age)/365.25
  
  mid_age_label <- (quantile(df.zscore[!is.na(df.zscore$gestational_age),
                                       "age_days_adj"],
                             probs = c(.50)) * sd_age + mu_age)/365.25

  high_age_label <- (quantile(df.zscore[!is.na(df.zscore$gestational_age),
                                      "age_days_adj"],
                            probs = c(.90)) * sd_age + mu_age)/365.25
  
  model_plot <- visualize_model(
  modobj = x,
  smooth_var = "gestational_age",
  int_var = "age_days_adj",
  plabels = "continuous interaction",
  derivative_plot = F)

  plot_colors <- ggplot_build(model_plot)$plot$scales$get_scales("fill")$palette(1)
  
  model_plot <- model_plot + 
    labs(x="Gestational Age (wks)",
         title = glue("{x$gam$terms[[2]]}"),
                  fill = "Age") +
    theme(axis.title.y = element_blank()) +
  scale_x_continuous(labels = function(x) round(x * sd_ga + mu_ga, 1)) +
  scale_fill_manual(labels = c(paste0("Low: ",
                                      round(low_age_label,1),"yrs"),
                               paste0("Mid: ",
                                      round(mid_age_label,1),"yrs"),
                               paste0("High: ",
                                      round(high_age_label,1),"yrs")), 
                    values = plot_colors) +
   scale_color_gradientn(colors = plot_colors,
    labels = function(x) round((x * sd_age + mu_age)/365.25, 1),
    name = "") +
  scale_y_continuous(labels = function(y) signif(pnorm(y)*100, 3))
  
  return(model_plot)
})

gabyage_sp_plots_vol <- setNames(gabyage_sp_plots_vol, gabyage_sp_vol_plot_pheno)

gabyage_sp_plots_vol_nopt <- lapply(gabyage_sp_plots_vol, function(x){
x$layers[1] <- NULL
x})

# wrap_plots(gabyage_sp_plots_vol_nopt, ncol = 2, guides = "collect") &
#   theme(legend.position = "right")
# 
# wrap_plots(gabyage_sp_plots_vol, ncol = 2, guides = "collect") &
#   theme(legend.position = "right")

wrapped_plot_n <- 4 

plot_batches <- split(gabyage_sp_plots_vol, ceiling(seq_along(gabyage_sp_plots_vol) / wrapped_plot_n))

 wrapped_list <- lapply(plot_batches, function(batch) {
   wrap_plots(batch, ncol = 2)
 })

wrapped_list

plot_batches_nopt <- split(gabyage_sp_plots_vol_nopt, ceiling(seq_along(gabyage_sp_plots_vol_nopt) / wrapped_plot_n))

 wrapped_list_nopt <- lapply(plot_batches_nopt, function(batch) {
   wrap_plots(batch, ncol = 2)
 })

wrapped_list_nopt
```

### BW

#### fit models
```{r, echo = FALSE}
bwbyage_sp_models_vol <- lapply(vol_vars, function(x) {
  form <- as.formula(glue("{x} ~ s(birth_weight_kg) + s(age_days_adj) + ti(birth_weight_kg, age_days_adj) + qc_var"))
  gam_model <- gamm(as.formula(form),
              data = df.zscore)
  return(gam_model)
  })

bwbyage_sp_models_vol <- setNames(bwbyage_sp_models_vol, vol_vars)

bwbyage_sp_gam_table_vol <- as.data.frame(t(sapply(bwbyage_sp_models_vol,function(x) {
  summary <- summary(x$gam)
  c(bw_edf = summary$s.table[1,1], bw_p = summary$s.table[1,4],
    age_edf = summary$s.table[2,1], age_p = summary$s.table[2,4],
    bwbyage_edf = summary$s.table[3,1], bwbyage_p = summary$s.table[3,4])
  })))

rownames(bwbyage_sp_gam_table_vol) <- vol_vars

bwbyage_sp_gam_table_vol$bwbyage_p_adj <- p.adjust(bwbyage_sp_gam_table_vol$bwbyage_p, method = "BH")

print(paste0("Number of significant IDPs for bw x Age- NO correction: ", dim(bwbyage_sp_gam_table_vol %>% filter(bwbyage_p < 0.05))[1]))
print(paste0("Number of significant IDPs for bw x Age- AFTER FDR correction: ", dim(bwbyage_sp_gam_table_vol %>% filter(bwbyage_p_adj < 0.05))[1]))
print(paste0("Significant IDPs for bw x Age- AFTER FDR correction: ", rownames(bwbyage_sp_gam_table_vol %>% filter(bwbyage_p_adj < 0.05)) ))

bwbyage_sp_vol_sig_pheno <- rownames(bwbyage_sp_gam_table_vol %>% filter(bwbyage_p_adj < 0.05))
```

#### Visualize models
```{r}
bwbyage_sp_vol_plot_pheno <- rownames(bwbyage_sp_gam_table_vol %>% filter(bw_edf > 2))
font_size <- 16
theme_set(theme_classic(base_family = "sans",base_size = font_size))
line_size <- 1.5
point_size <- 2
source("/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/visualize_model_bartfct.R")

mu_age <- mean(cent_all$age_days_adj, na.rm = TRUE)
sd_age <- sd(cent_all$age_days_adj, na.rm = TRUE)

mu_bw <- mean(cent_all$birth_weight_kg, na.rm = TRUE)
sd_bw <- sd(cent_all$birth_weight_kg, na.rm = TRUE)

bwbyage_sp_plots_vol <- lapply(bwbyage_sp_models_vol[bwbyage_sp_vol_plot_pheno], function(x) {
  
  
  low_age_label <- (quantile(df.zscore[!is.na(df.zscore$birth_weight_kg),
                                       "age_days_adj"],
                             probs = c(.10)) * sd_age + mu_age)/365.25
  
  mid_age_label <- (quantile(df.zscore[!is.na(df.zscore$birth_weight_kg),
                                       "age_days_adj"],
                             probs = c(.50)) * sd_age + mu_age)/365.25

  high_age_label <- (quantile(df.zscore[!is.na(df.zscore$birth_weight_kg),
                                      "age_days_adj"],
                            probs = c(.90)) * sd_age + mu_age)/365.25
  
  model_plot <- visualize_model(
  modobj = x,
  smooth_var = "birth_weight_kg",
  int_var = "age_days_adj",
  plabels = "continuous interaction",
  derivative_plot = F)

  plot_colors <- ggplot_build(model_plot)$plot$scales$get_scales("fill")$palette(1)
  
  model_plot <- model_plot + 
    labs(x="Birth Weight (kg)",
         title = glue("{x$gam$terms[[2]]}"),
                  fill = "Age") +
    theme(axis.title.y = element_blank()) +
  scale_x_continuous(labels = function(x) round(x * sd_bw + mu_bw, 1)) +
  scale_fill_manual(labels = c(paste0("Low: ",
                                      round(low_age_label,1),"yrs"),
                               paste0("Mid: ",
                                      round(mid_age_label,1),"yrs"),
                               paste0("High: ",
                                      round(high_age_label,1),"yrs")), 
                    values = plot_colors) +
   scale_color_gradientn(colors = plot_colors,
    labels = function(x) round((x * sd_age + mu_age)/365.25, 1),
    name = "") +
  scale_y_continuous(labels = function(y) signif(pnorm(y)*100, 3))
  
  return(model_plot)
})

bwbyage_sp_plots_vol <- setNames(bwbyage_sp_plots_vol, bwbyage_sp_vol_plot_pheno)

bwbyage_sp_plots_vol_nopt <- lapply(bwbyage_sp_plots_vol, function(x){
x$layers[1] <- NULL
x})

wrapped_plot_n <- 4 

plot_batches <- split(bwbyage_sp_plots_vol, ceiling(seq_along(bwbyage_sp_plots_vol) / wrapped_plot_n))

 wrapped_list <- lapply(plot_batches, function(batch) {
   wrap_plots(batch, ncol = 2)
 })

wrapped_list

plot_batches_nopt <- split(bwbyage_sp_plots_vol_nopt, ceiling(seq_along(bwbyage_sp_plots_vol_nopt) / wrapped_plot_n))

 wrapped_list_nopt <- lapply(plot_batches_nopt, function(batch) {
   wrap_plots(batch, ncol = 2)
 })

wrapped_list_nopt
```


## Surface Area
### GA

#### fit models
```{r, echo = FALSE}
gabyage_sp_models_sa <- lapply(sa_vars, function(x) {
  form <- as.formula(glue("{x} ~ s(gestational_age) + s(age_days_adj) + ti(gestational_age, age_days_adj) + qc_var"))
  gam_model <- gamm(as.formula(form),
              data = df.zscore)
  return(gam_model)
  })

gabyage_sp_models_sa <- setNames(gabyage_sp_models_sa, sa_vars)

gabyage_sp_gam_table_sa <- as.data.frame(t(sapply(gabyage_sp_models_sa,function(x) {
  summary <- summary(x$gam)
  c(ga_edf = summary$s.table[1,1], ga_p = summary$s.table[1,4],
    age_edf = summary$s.table[2,1], age_p = summary$s.table[2,4],
    gabyage_edf = summary$s.table[3,1], gabyage_p = summary$s.table[3,4])
  })))

rownames(gabyage_sp_gam_table_sa) <- sa_vars

gabyage_sp_gam_table_sa$gabyage_p_adj <- p.adjust(gabyage_sp_gam_table_sa$gabyage_p, method = "BH")

print(paste0("Number of significant IDPs for GA x Age- NO correction: ", dim(gabyage_sp_gam_table_sa %>% filter(gabyage_p < 0.05))[1]))
print(paste0("Number of significant IDPs for GA x Age- AFTER FDR correction: ", dim(gabyage_sp_gam_table_sa %>% filter(gabyage_p_adj < 0.05))[1]))
print(paste0("Significant IDPs for GA x Age- AFTER FDR correction: ", rownames(gabyage_sp_gam_table_sa %>% filter(gabyage_p_adj < 0.05)) ))

gabyage_sp_sa_sig_pheno <- rownames(gabyage_sp_gam_table_sa %>% filter(gabyage_p_adj < 0.05))
```

#### Visualize models
```{r}
gabyage_sp_sa_plot_pheno <- rownames(gabyage_sp_gam_table_sa %>% filter(ga_edf > 2))
font_size <- 16
theme_set(theme_classic(base_family = "sans",base_size = font_size))
line_size <- 1.5
point_size <- 2
source("/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/visualize_model_bartfct.R")

mu_age <- mean(cent_all$age_days_adj, na.rm = TRUE)
sd_age <- sd(cent_all$age_days_adj, na.rm = TRUE)

mu_ga <- mean(cent_all$gestational_age, na.rm = TRUE)
sd_ga <- sd(cent_all$gestational_age, na.rm = TRUE)

gabyage_sp_plots_sa <- lapply(gabyage_sp_models_sa[gabyage_sp_sa_plot_pheno], function(x) {
  
  
  low_age_label <- (quantile(df.zscore[!is.na(df.zscore$gestational_age),
                                       "age_days_adj"],
                             probs = c(.10)) * sd_age + mu_age)/365.25
  
  mid_age_label <- (quantile(df.zscore[!is.na(df.zscore$gestational_age),
                                       "age_days_adj"],
                             probs = c(.50)) * sd_age + mu_age)/365.25

  high_age_label <- (quantile(df.zscore[!is.na(df.zscore$gestational_age),
                                      "age_days_adj"],
                            probs = c(.90)) * sd_age + mu_age)/365.25
  
  model_plot <- visualize_model(
  modobj = x,
  smooth_var = "gestational_age",
  int_var = "age_days_adj",
  plabels = "continuous interaction",
  derivative_plot = F)

  plot_colors <- ggplot_build(model_plot)$plot$scales$get_scales("fill")$palette(1)
  
  model_plot <- model_plot + 
    labs(x="Gestational Age (wks)",
         title = glue("{x$gam$terms[[2]]}"),
                  fill = "Age") +
    theme(axis.title.y = element_blank()) +
  scale_x_continuous(labels = function(x) round(x * sd_ga + mu_ga, 1)) +
  scale_fill_manual(labels = c(paste0("Low: ",
                                      round(low_age_label,1),"yrs"),
                               paste0("Mid: ",
                                      round(mid_age_label,1),"yrs"),
                               paste0("High: ",
                                      round(high_age_label,1),"yrs")), 
                    values = plot_colors) +
   scale_color_gradientn(colors = plot_colors,
    labels = function(x) round((x * sd_age + mu_age)/365.25, 1),
    name = "") +
  scale_y_continuous(labels = function(y) signif(pnorm(y)*100, 3))
  
  return(model_plot)
})

gabyage_sp_plots_sa <- setNames(gabyage_sp_plots_sa, gabyage_sp_sa_plot_pheno)

gabyage_sp_plots_sa_nopt <- lapply(gabyage_sp_plots_sa, function(x){
x$layers[1] <- NULL
x})

wrapped_plot_n <- 4 

plot_batches <- split(gabyage_sp_plots_sa, ceiling(seq_along(gabyage_sp_plots_sa) / wrapped_plot_n))

 wrapped_list <- lapply(plot_batches, function(batch) {
   wrap_plots(batch, ncol = 2)
 })

wrapped_list

plot_batches_nopt <- split(gabyage_sp_plots_sa_nopt, ceiling(seq_along(gabyage_sp_plots_sa_nopt) / wrapped_plot_n))

 wrapped_list_nopt <- lapply(plot_batches_nopt, function(batch) {
   wrap_plots(batch, ncol = 2)
 })

wrapped_list_nopt

#wrapped_list
```

### BW

#### fit models
```{r, echo = FALSE}
bwbyage_sp_models_sa <- lapply(sa_vars, function(x) {
  form <- as.formula(glue("{x} ~ s(birth_weight_kg) + s(age_days_adj) + ti(birth_weight_kg, age_days_adj) + qc_var"))
  gam_model <- gamm(as.formula(form),
              data = df.zscore)
  return(gam_model)
  })

bwbyage_sp_models_sa <- setNames(bwbyage_sp_models_sa, sa_vars)

bwbyage_sp_gam_table_sa <- as.data.frame(t(sapply(bwbyage_sp_models_sa,function(x) {
  summary <- summary(x$gam)
  c(bw_edf = summary$s.table[1,1], bw_p = summary$s.table[1,4],
    age_edf = summary$s.table[2,1], age_p = summary$s.table[2,4],
    bwbyage_edf = summary$s.table[3,1], bwbyage_p = summary$s.table[3,4])
  })))

rownames(bwbyage_sp_gam_table_sa) <- sa_vars

bwbyage_sp_gam_table_sa$bwbyage_p_adj <- p.adjust(bwbyage_sp_gam_table_sa$bwbyage_p, method = "BH")

print(paste0("Number of significant IDPs for bw x Age- NO correction: ", dim(bwbyage_sp_gam_table_sa %>% filter(bwbyage_p < 0.05))[1]))
print(paste0("Number of significant IDPs for bw x Age- AFTER FDR correction: ", dim(bwbyage_sp_gam_table_sa %>% filter(bwbyage_p_adj < 0.05))[1]))
print(paste0("Significant IDPs for bw x Age- AFTER FDR correction: ", rownames(bwbyage_sp_gam_table_sa %>% filter(bwbyage_p_adj < 0.05)) ))

bwbyage_sp_sa_sig_pheno <- rownames(bwbyage_sp_gam_table_sa %>% filter(bwbyage_p_adj < 0.05))
```

#### Visualize models
```{r}
bwbyage_sp_sa_plot_pheno <- rownames(bwbyage_sp_gam_table_sa %>% filter(bw_edf > 2))
font_size <- 16
theme_set(theme_classic(base_family = "sans",base_size = font_size))
line_size <- 1.5
point_size <- 2
source("/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/visualize_model_bartfct.R")

mu_age <- mean(cent_all$age_days_adj, na.rm = TRUE)
sd_age <- sd(cent_all$age_days_adj, na.rm = TRUE)

mu_bw <- mean(cent_all$birth_weight_kg, na.rm = TRUE)
sd_bw <- sd(cent_all$birth_weight_kg, na.rm = TRUE)

bwbyage_sp_plots_sa <- lapply(bwbyage_sp_models_sa[bwbyage_sp_sa_plot_pheno], function(x) {
  
  
  low_age_label <- (quantile(df.zscore[!is.na(df.zscore$birth_weight_kg),
                                       "age_days_adj"],
                             probs = c(.10)) * sd_age + mu_age)/365.25
  
  mid_age_label <- (quantile(df.zscore[!is.na(df.zscore$birth_weight_kg),
                                       "age_days_adj"],
                             probs = c(.50)) * sd_age + mu_age)/365.25

  high_age_label <- (quantile(df.zscore[!is.na(df.zscore$birth_weight_kg),
                                      "age_days_adj"],
                            probs = c(.90)) * sd_age + mu_age)/365.25
  
  model_plot <- visualize_model(
  modobj = x,
  smooth_var = "birth_weight_kg",
  int_var = "age_days_adj",
  plabels = "continuous interaction",
  derivative_plot = F)

  plot_colors <- ggplot_build(model_plot)$plot$scales$get_scales("fill")$palette(1)
  
  model_plot <- model_plot + 
    labs(x="Birth Weight (kg)",
         title = glue("{x$gam$terms[[2]]}"),
                  fill = "Age") +
    theme(axis.title.y = element_blank()) +
  scale_x_continuous(labels = function(x) round(x * sd_bw + mu_bw, 1)) +
  scale_fill_manual(labels = c(paste0("Low: ",
                                      round(low_age_label,1),"yrs"),
                               paste0("Mid: ",
                                      round(mid_age_label,1),"yrs"),
                               paste0("High: ",
                                      round(high_age_label,1),"yrs")), 
                    values = plot_colors) +
   scale_color_gradientn(colors = plot_colors,
    labels = function(x) round((x * sd_age + mu_age)/365.25, 1),
    name = "") +
  scale_y_continuous(labels = function(y) signif(pnorm(y)*100, 3))
  
  return(model_plot)
})

bwbyage_sp_plots_sa <- setNames(bwbyage_sp_plots_sa, bwbyage_sp_sa_plot_pheno)

bwbyage_sp_plots_sa_nopt <- lapply(bwbyage_sp_plots_sa, function(x){
x$layers[1] <- NULL
x})

wrapped_plot_n <- 4 

plot_batches <- split(bwbyage_sp_plots_sa, ceiling(seq_along(bwbyage_sp_plots_sa) / wrapped_plot_n))

 wrapped_list <- lapply(plot_batches, function(batch) {
   wrap_plots(batch, ncol = 2)
 })

wrapped_list

plot_batches_nopt <- split(bwbyage_sp_plots_sa_nopt, ceiling(seq_along(bwbyage_sp_plots_sa_nopt) / wrapped_plot_n))

 wrapped_list_nopt <- lapply(plot_batches_nopt, function(batch) {
   wrap_plots(batch, ncol = 2)
 })

wrapped_list_nopt
```


## Thickness median full sample
### GA

#### fit models
```{r, echo = FALSE}
if (median_th_mpr_flag == TRUE){
gabyage_sp_models_th_median <- lapply(th_vars_median, function(x) {
  form <- as.formula(glue("{x} ~ s(gestational_age) + s(age_days_adj) + ti(gestational_age, age_days_adj) + qc_var"))
  gam_model <- gamm(as.formula(form),
              data = df.zscore)
  return(gam_model)
  })

gabyage_sp_models_th_median <- setNames(gabyage_sp_models_th_median, th_vars)

gabyage_sp_gam_table_th_median <- as.data.frame(t(sapply(gabyage_sp_models_th_median,function(x) {
  summary <- summary(x$gam)
  c(ga_edf = summary$s.table[1,1], ga_p = summary$s.table[1,4],
    age_edf = summary$s.table[2,1], age_p = summary$s.table[2,4],
    gabyage_edf = summary$s.table[3,1], gabyage_p = summary$s.table[3,4])
  })))

rownames(gabyage_sp_gam_table_th_median) <- th_vars

gabyage_sp_gam_table_th_median$gabyage_p_adj <- p.adjust(gabyage_sp_gam_table_th_median$gabyage_p, method = "BH")

print(paste0("Number of significant IDPs for GA x Age- NO correction: ", dim(gabyage_sp_gam_table_th_median %>% filter(gabyage_p < 0.05))[1]))
print(paste0("Number of significant IDPs for GA x Age- AFTER FDR correction: ", dim(gabyage_sp_gam_table_th_median %>% filter(gabyage_p_adj < 0.05))[1]))
print(paste0("Significant IDPs for GA x Age- AFTER FDR correction: ", rownames(gabyage_sp_gam_table_th_median %>% filter(gabyage_p_adj < 0.05)) ))

gabyage_sp_th_median_sig_pheno <- rownames(gabyage_sp_gam_table_th_median %>% filter(gabyage_p_adj < 0.05))
}
```

#### Visualize models
```{r}
if (median_th_mpr_flag == TRUE){
gabyage_sp_th_median_plot_pheno <- rownames(gabyage_sp_gam_table_th_median %>% filter(ga_edf > 2))
font_size <- 16
theme_set(theme_classic(base_family = "sans",base_size = font_size))
line_size <- 1.5
point_size <- 2
source("/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/visualize_model_bartfct.R")

mu_age <- mean(cent_all$age_days_adj, na.rm = TRUE)
sd_age <- sd(cent_all$age_days_adj, na.rm = TRUE)

mu_ga <- mean(cent_all$gestational_age, na.rm = TRUE)
sd_ga <- sd(cent_all$gestational_age, na.rm = TRUE)

gabyage_sp_plots_th_median <- lapply(gabyage_sp_models_th_median[gabyage_sp_th_median_sig_pheno], function(x) {
  
  
  low_age_label <- (quantile(df.zscore[!is.na(df.zscore$gestational_age),
                                       "age_days_adj"],
                             probs = c(.10)) * sd_age + mu_age)/365.25
  
  mid_age_label <- (quantile(df.zscore[!is.na(df.zscore$gestational_age),
                                       "age_days_adj"],
                             probs = c(.50)) * sd_age + mu_age)/365.25

  high_age_label <- (quantile(df.zscore[!is.na(df.zscore$gestational_age),
                                      "age_days_adj"],
                            probs = c(.90)) * sd_age + mu_age)/365.25
  
  model_plot <- visualize_model(
  modobj = x,
  smooth_var = "gestational_age",
  int_var = "age_days_adj",
  plabels = "continuous interaction",
  derivative_plot = F)

  plot_colors <- ggplot_build(model_plot)$plot$scales$get_scales("fill")$palette(1)
  
  model_plot <- model_plot + 
    labs(x="Gestational Age (wks)",
         title = glue("{x$gam$terms[[2]]}"),
                  fill = "Age") +
    theme(axis.title.y = element_blank()) +
  scale_x_continuous(labels = function(x) round(x * sd_ga + mu_ga, 1)) +
  scale_fill_manual(labels = c(paste0("Low: ",
                                      round(low_age_label,1),"yrs"),
                               paste0("Mid: ",
                                      round(mid_age_label,1),"yrs"),
                               paste0("High: ",
                                      round(high_age_label,1),"yrs")), 
                    values = plot_colors) +
   scale_color_gradientn(colors = plot_colors,
    labels = function(x) round((x * sd_age + mu_age)/365.25, 1),
    name = "") +
  scale_y_continuous(labels = function(y) signif(pnorm(y)*100, 3))
  
  return(model_plot)
})

gabyage_sp_plots_th_median <- setNames(gabyage_sp_plots_th_median, gabyage_sp_th_median_sig_pheno)

gabyage_sp_plots_th_median_nopt <- lapply(gabyage_sp_plots_th_median, function(x){
x$layers[1] <- NULL
x})

wrapped_plot_n <- 4 

plot_batches <- split(gabyage_sp_plots_th_median, ceiling(seq_along(gabyage_sp_plots_th_median) / wrapped_plot_n))

 wrapped_list <- lapply(plot_batches, function(batch) {
   wrap_plots(batch, ncol = 2)
 })

wrapped_list

plot_batches_nopt <- split(gabyage_sp_plots_th_median_nopt, ceiling(seq_along(gabyage_sp_plots_th_median_nopt) / wrapped_plot_n))

 wrapped_list_nopt <- lapply(plot_batches_nopt, function(batch) {
   wrap_plots(batch, ncol = 2)
 })

wrapped_list_nopt
}
```

### BW

#### fit models
```{r, echo = FALSE}
if (median_th_mpr_flag == TRUE){
bwbyage_sp_models_th_median <- lapply(th_vars_median, function(x) {
  form <- as.formula(glue("{x} ~ s(birth_weight_kg) + s(age_days_adj) + ti(birth_weight_kg, age_days_adj) + qc_var"))
  gam_model <- gamm(as.formula(form),
              data = df.zscore)
  return(gam_model)
  })

bwbyage_sp_models_th_median <- setNames(bwbyage_sp_models_th_median, th_vars)

bwbyage_sp_gam_table_th_median <- as.data.frame(t(sapply(bwbyage_sp_models_th_median,function(x) {
  summary <- summary(x$gam)
  c(bw_edf = summary$s.table[1,1], bw_p = summary$s.table[1,4],
    age_edf = summary$s.table[2,1], age_p = summary$s.table[2,4],
    bwbyage_edf = summary$s.table[3,1], bwbyage_p = summary$s.table[3,4])
  })))

rownames(bwbyage_sp_gam_table_th_median) <- th_vars

bwbyage_sp_gam_table_th_median$bwbyage_p_adj <- p.adjust(bwbyage_sp_gam_table_th_median$bwbyage_p, method = "BH")

print(paste0("Number of significant IDPs for bw x Age- NO correction: ", dim(bwbyage_sp_gam_table_th_median %>% filter(bwbyage_p < 0.05))[1]))
print(paste0("Number of significant IDPs for bw x Age- AFTER FDR correction: ", dim(bwbyage_sp_gam_table_th_median %>% filter(bwbyage_p_adj < 0.05))[1]))
print(paste0("Significant IDPs for bw x Age- AFTER FDR correction: ", rownames(bwbyage_sp_gam_table_th_median %>% filter(bwbyage_p_adj < 0.05)) ))

bwbyage_sp_th_median_sig_pheno <- rownames(bwbyage_sp_gam_table_th_median %>% filter(bwbyage_p_adj < 0.05))
}
```

#### Visualize models
```{r}
if (median_th_mpr_flag == TRUE){
bwbyage_sp_th_median_plot_pheno <- rownames(bwbyage_sp_gam_table_th_median %>% filter(bw_edf > 2))
font_size <- 16
theme_set(theme_classic(base_family = "sans",base_size = font_size))
line_size <- 1.5
point_size <- 2
source("/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/visualize_model_bartfct.R")

mu_age <- mean(cent_all$age_days_adj, na.rm = TRUE)
sd_age <- sd(cent_all$age_days_adj, na.rm = TRUE)

mu_bw <- mean(cent_all$birth_weight_kg, na.rm = TRUE)
sd_bw <- sd(cent_all$birth_weight_kg, na.rm = TRUE)

bwbyage_sp_plots_th_median <- lapply(bwbyage_sp_models_th_median[bwbyage_sp_th_median_sig_pheno], function(x) {
  
  
  low_age_label <- (quantile(df.zscore[!is.na(df.zscore$birth_weight_kg),
                                       "age_days_adj"],
                             probs = c(.10)) * sd_age + mu_age)/365.25
  
  mid_age_label <- (quantile(df.zscore[!is.na(df.zscore$birth_weight_kg),
                                       "age_days_adj"],
                             probs = c(.50)) * sd_age + mu_age)/365.25

  high_age_label <- (quantile(df.zscore[!is.na(df.zscore$birth_weight_kg),
                                      "age_days_adj"],
                            probs = c(.90)) * sd_age + mu_age)/365.25
  
  model_plot <- visualize_model(
  modobj = x,
  smooth_var = "birth_weight_kg",
  int_var = "age_days_adj",
  plabels = "continuous interaction",
  derivative_plot = F)

  plot_colors <- ggplot_build(model_plot)$plot$scales$get_scales("fill")$palette(1)
  
  model_plot <- model_plot + 
    labs(x="Birth Weight (kg)",
         title = glue("{x$gam$terms[[2]]}"),
                  fill = "Age") +
    theme(axis.title.y = element_blank()) +
  scale_x_continuous(labels = function(x) round(x * sd_bw + mu_bw, 1)) +
  scale_fill_manual(labels = c(paste0("Low: ",
                                      round(low_age_label,1),"yrs"),
                               paste0("Mid: ",
                                      round(mid_age_label,1),"yrs"),
                               paste0("High: ",
                                      round(high_age_label,1),"yrs")), 
                    values = plot_colors) +
   scale_color_gradientn(colors = plot_colors,
    labels = function(x) round((x * sd_age + mu_age)/365.25, 1),
    name = "") +
  scale_y_continuous(labels = function(y) signif(pnorm(y)*100, 3))
  
  return(model_plot)
})

bwbyage_sp_plots_th_median <- setNames(bwbyage_sp_plots_th_median, bwbyage_sp_th_median_sig_pheno)

bwbyage_sp_plots_th_median_nopt <- lapply(bwbyage_sp_plots_th_median, function(x){
x$layers[1] <- NULL
x})

wrapped_plot_n <- 4 

plot_batches <- split(bwbyage_sp_plots_th_median, ceiling(seq_along(bwbyage_sp_plots_th_median) / wrapped_plot_n))

 wrapped_list <- lapply(plot_batches, function(batch) {
   wrap_plots(batch, ncol = 2)
 })

wrapped_list

plot_batches_nopt <- split(bwbyage_sp_plots_th_median_nopt, ceiling(seq_along(bwbyage_sp_plots_th_median_nopt) / wrapped_plot_n))

 wrapped_list_nopt <- lapply(plot_batches_nopt, function(batch) {
   wrap_plots(batch, ncol = 2)
 })

wrapped_list_nopt
}
```


## Thickness (mpr, df.zscore_overOne overOne sample)
### GA

#### fit models
```{r, echo = FALSE}
gabyage_sp_models_th <- lapply(th_vars, function(x) {
  form <- as.formula(glue("{x} ~ s(gestational_age) + s(age_days_adj) + ti(gestational_age, age_days_adj) + euler_mean_mpr"))
  gam_model <- gamm(as.formula(form),
              data = df.zscore_overOne)
  return(gam_model)
  })

gabyage_sp_models_th <- setNames(gabyage_sp_models_th, th_vars)

gabyage_sp_gam_table_th <- as.data.frame(t(sapply(gabyage_sp_models_th,function(x) {
  summary <- summary(x$gam)
  c(ga_edf = summary$s.table[1,1], ga_p = summary$s.table[1,4],
    age_edf = summary$s.table[2,1], age_p = summary$s.table[2,4],
    gabyage_edf = summary$s.table[3,1], gabyage_p = summary$s.table[3,4])
  })))

rownames(gabyage_sp_gam_table_th) <- th_vars

gabyage_sp_gam_table_th$gabyage_p_adj <- p.adjust(gabyage_sp_gam_table_th$gabyage_p, method = "BH")

print(paste0("Number of significant IDPs for GA x Age- NO correction: ", dim(gabyage_sp_gam_table_th %>% filter(gabyage_p < 0.05))[1]))
print(paste0("Number of significant IDPs for GA x Age- AFTER FDR correction: ", dim(gabyage_sp_gam_table_th %>% filter(gabyage_p_adj < 0.05))[1]))
print(paste0("Significant IDPs for GA x Age- AFTER FDR correction: ", rownames(gabyage_sp_gam_table_th %>% filter(gabyage_p_adj < 0.05)) ))

gabyage_sp_th_sig_pheno <- rownames(gabyage_sp_gam_table_th %>% filter(gabyage_p_adj < 0.05))
```

#### Visualize models
```{r}
gabyage_sp_th_plot_pheno <- rownames(gabyage_sp_gam_table_th %>% filter(ga_edf > 2))
font_size <- 16
theme_set(theme_classic(base_family = "sans",base_size = font_size))
line_size <- 1.5
point_size <- 2
source("/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/visualize_model_bartfct.R")

mu_age <- mean(cent_all$age_days_adj, na.rm = TRUE)
sd_age <- sd(cent_all$age_days_adj, na.rm = TRUE)

mu_ga <- mean(cent_all$gestational_age, na.rm = TRUE)
sd_ga <- sd(cent_all$gestational_age, na.rm = TRUE)

gabyage_sp_plots_th <- lapply(gabyage_sp_models_th[gabyage_sp_th_sig_pheno], function(x) {
  
  
  low_age_label <- (quantile(df.zscore_overOne[!is.na(df.zscore_overOne$gestational_age),
                                       "age_days_adj"],
                             probs = c(.10)) * sd_age + mu_age)/365.25
  
  mid_age_label <- (quantile(df.zscore_overOne[!is.na(df.zscore_overOne$gestational_age),
                                       "age_days_adj"],
                             probs = c(.50)) * sd_age + mu_age)/365.25

  high_age_label <- (quantile(df.zscore_overOne[!is.na(df.zscore_overOne$gestational_age),
                                      "age_days_adj"],
                            probs = c(.90)) * sd_age + mu_age)/365.25
  
  model_plot <- visualize_model(
  modobj = x,
  smooth_var = "gestational_age",
  int_var = "age_days_adj",
  plabels = "continuous interaction",
  derivative_plot = F)

  plot_colors <- ggplot_build(model_plot)$plot$scales$get_scales("fill")$palette(1)
  
  model_plot <- model_plot + 
    labs(x="Gestational Age (wks)",
         title = glue("{x$gam$terms[[2]]}"),
                  fill = "Age") +
    theme(axis.title.y = element_blank()) +
  scale_x_continuous(labels = function(x) round(x * sd_ga + mu_ga, 1)) +
  scale_fill_manual(labels = c(paste0("Low: ",
                                      round(low_age_label,1),"yrs"),
                               paste0("Mid: ",
                                      round(mid_age_label,1),"yrs"),
                               paste0("High: ",
                                      round(high_age_label,1),"yrs")), 
                    values = plot_colors) +
   scale_color_gradientn(colors = plot_colors,
    labels = function(x) round((x * sd_age + mu_age)/365.25, 1),
    name = "") +
  scale_y_continuous(labels = function(y) signif(pnorm(y)*100, 3))
  
  return(model_plot)
})

gabyage_sp_plots_th <- setNames(gabyage_sp_plots_th, gabyage_sp_th_sig_pheno)

gabyage_sp_plots_th_nopt <- lapply(gabyage_sp_plots_th, function(x){
x$layers[1] <- NULL
x})

wrapped_plot_n <- 4 

plot_batches <- split(gabyage_sp_plots_th, ceiling(seq_along(gabyage_sp_plots_th) / wrapped_plot_n))

 wrapped_list <- lapply(plot_batches, function(batch) {
   wrap_plots(batch, ncol = 2)
 })

wrapped_list

plot_batches_nopt <- split(gabyage_sp_plots_th_nopt, ceiling(seq_along(gabyage_sp_plots_th_nopt) / wrapped_plot_n))

 wrapped_list_nopt <- lapply(plot_batches_nopt, function(batch) {
   wrap_plots(batch, ncol = 2)
 })

wrapped_list_nopt
```

### BW

#### fit models
```{r, echo = FALSE}
bwbyage_sp_models_th <- lapply(th_vars, function(x) {
  form <- as.formula(glue("{x} ~ s(birth_weight_kg) + s(age_days_adj) + ti(birth_weight_kg, age_days_adj) + euler_mean_mpr"))
  gam_model <- gamm(as.formula(form),
              data = df.zscore_overOne)
  return(gam_model)
  })

bwbyage_sp_models_th <- setNames(bwbyage_sp_models_th, th_vars)

bwbyage_sp_gam_table_th <- as.data.frame(t(sapply(bwbyage_sp_models_th,function(x) {
  summary <- summary(x$gam)
  c(bw_edf = summary$s.table[1,1], bw_p = summary$s.table[1,4],
    age_edf = summary$s.table[2,1], age_p = summary$s.table[2,4],
    bwbyage_edf = summary$s.table[3,1], bwbyage_p = summary$s.table[3,4])
  })))

rownames(bwbyage_sp_gam_table_th) <- th_vars

bwbyage_sp_gam_table_th$bwbyage_p_adj <- p.adjust(bwbyage_sp_gam_table_th$bwbyage_p, method = "BH")

print(paste0("Number of significant IDPs for bw x Age- NO correction: ", dim(bwbyage_sp_gam_table_th %>% filter(bwbyage_p < 0.05))[1]))
print(paste0("Number of significant IDPs for bw x Age- AFTER FDR correction: ", dim(bwbyage_sp_gam_table_th %>% filter(bwbyage_p_adj < 0.05))[1]))
print(paste0("Significant IDPs for bw x Age- AFTER FDR correction: ", rownames(bwbyage_sp_gam_table_th %>% filter(bwbyage_p_adj < 0.05)) ))

bwbyage_sp_th_sig_pheno <- rownames(bwbyage_sp_gam_table_th %>% filter(bwbyage_p_adj < 0.05))
```

#### Visualize models
```{r}
bwbyage_sp_th_plot_pheno <- rownames(bwbyage_sp_gam_table_th %>% filter(bw_edf > 2))
font_size <- 16
theme_set(theme_classic(base_family = "sans",base_size = font_size))
line_size <- 1.5
point_size <- 2
source("/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/visualize_model_bartfct.R")

mu_age <- mean(cent_all$age_days_adj, na.rm = TRUE)
sd_age <- sd(cent_all$age_days_adj, na.rm = TRUE)

mu_bw <- mean(cent_all$birth_weight_kg, na.rm = TRUE)
sd_bw <- sd(cent_all$birth_weight_kg, na.rm = TRUE)

bwbyage_sp_plots_th <- lapply(bwbyage_sp_models_th[bwbyage_sp_th_sig_pheno], function(x) {
  
  
  low_age_label <- (quantile(df.zscore_overOne[!is.na(df.zscore_overOne$birth_weight_kg),
                                       "age_days_adj"],
                             probs = c(.10)) * sd_age + mu_age)/365.25
  
  mid_age_label <- (quantile(df.zscore_overOne[!is.na(df.zscore_overOne$birth_weight_kg),
                                       "age_days_adj"],
                             probs = c(.50)) * sd_age + mu_age)/365.25

  high_age_label <- (quantile(df.zscore_overOne[!is.na(df.zscore_overOne$birth_weight_kg),
                                      "age_days_adj"],
                            probs = c(.90)) * sd_age + mu_age)/365.25
  
  model_plot <- visualize_model(
  modobj = x,
  smooth_var = "birth_weight_kg",
  int_var = "age_days_adj",
  plabels = "continuous interaction",
  derivative_plot = F)

  plot_colors <- ggplot_build(model_plot)$plot$scales$get_scales("fill")$palette(1)
  
  model_plot <- model_plot + 
    labs(x="Birth Weight (kg)",
         title = glue("{x$gam$terms[[2]]}"),
                  fill = "Age") +
    theme(axis.title.y = element_blank()) +
  scale_x_continuous(labels = function(x) round(x * sd_bw + mu_bw, 1)) +
  scale_fill_manual(labels = c(paste0("Low: ",
                                      round(low_age_label,1),"yrs"),
                               paste0("Mid: ",
                                      round(mid_age_label,1),"yrs"),
                               paste0("High: ",
                                      round(high_age_label,1),"yrs")), 
                    values = plot_colors) +
   scale_color_gradientn(colors = plot_colors,
    labels = function(x) round((x * sd_age + mu_age)/365.25, 1),
    name = "") +
  scale_y_continuous(labels = function(y) signif(pnorm(y)*100, 3))
  
  return(model_plot)
})

bwbyage_sp_plots_th <- setNames(bwbyage_sp_plots_th, bwbyage_sp_th_sig_pheno)

bwbyage_sp_plots_th_nopt <- lapply(bwbyage_sp_plots_th, function(x){
x$layers[1] <- NULL
x})

wrapped_plot_n <- 4 

plot_batches <- split(bwbyage_sp_plots_th, ceiling(seq_along(bwbyage_sp_plots_th) / wrapped_plot_n))

 wrapped_list <- lapply(plot_batches, function(batch) {
   wrap_plots(batch, ncol = 2)
 })

wrapped_list

plot_batches_nopt <- split(bwbyage_sp_plots_th_nopt, ceiling(seq_along(bwbyage_sp_plots_th_nopt) / wrapped_plot_n))

 wrapped_list_nopt <- lapply(plot_batches_nopt, function(batch) {
   wrap_plots(batch, ncol = 2)
 })

wrapped_list_nopt
```

