---
title: "slipga_spline_analysis"
author: "Eren Kafadar"
date: "2026-06-02"
output:
  html_document:
    df_print: paged
params:
  output_folder : "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/analysis_results/median-th-mpr/"
  qnorm_centiles_path: "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/analysis_results/median-th-mpr/analysis_dfs/df_zscore_distinct.csv"
  lookup_stats: "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/analysis_results/median-th-mpr/analysis_dfs/lookup_stats_distinct.RData"
  rm : FALSE
editor_options: 
  chunk_output_type: console
---
# Define params if they do not exist
```{r params, echo = FALSE}
if (!exists("params")) {
  params <- list(
  output_folder = "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/analysis_results/",
  qnorm_centiles_path = "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/analysis_results/median-th-mpr/analysis_dfs/df_zscore_distinct.csv",
  lookup_stats = "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/analysis_results/median-th-mpr/analysis_dfs/lookup_stats_distinct.RData",
  rm = FALSE
  )
}
print(params)
``` 

# Load Packages
```{r packages, echo = FALSE}
# Load libraries
require(dplyr)
require(glue)
require(magrittr)
require(readr)
require(stringr)
require(tidyr)
require(scales)
require(mgcv)
require(pbkrtest)
require(gamlssTools) # Margaret' package
require(tibble)
require(lme4)
require(purrr)
# For figures/plotting
require(ggplot2)
require(gridExtra)
require(ggsegDKT)
require(ggseg)
require(ggpubr)
require(ggcorrplot)
require(cowplot)
require(ggh4x)
require(gratia)
require(egg)
require(interactions)
require(growthstandards)
require(gtsummary)
require(svglite)
require(patchwork)

`%nin%` = Negate(`%in%`)
```

# Controls
```{r setup, echo = TRUE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)
# set command
save_tables <- TRUE
plot <- TRUE
save_figs <- TRUE
save_models <- TRUE
global_measure_vol <- "global_eTIV"
global_measure_sa <- "global_SurfaceArea"
global_measure_th <- "global_MeanThickness"
set.seed(42)

output_folder <- params$output_folder
qnorm_centiles_path <- params$qnorm_centiles_path
lookup_stats <- params$lookup_stats
rm <- params$rm
```

# Define Paths
```{r, echo = TRUE}
# set folder paths
code_path <- "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/"
vars_path <-  "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/output/vars_2511_new/"

# output folder
output_folder <- params$output_folder
if (!dir.exists(output_folder)) dir.create(output_folder)
##  output subfolder paths
analysis_dfs_folder <- paste0(output_folder, "analysis_dfs/")
if (!dir.exists(analysis_dfs_folder)) dir.create(analysis_dfs_folder)

models_folder <- paste0(output_folder, "models/")
if (!dir.exists(models_folder)) dir.create(models_folder)

csv_output_folder <-  paste0(output_folder, "results_csv/")
if (!dir.exists(csv_output_folder)) dir.create(csv_output_folder)

  betas_folder <- paste0(csv_output_folder, "full_betas/")
  if (!dir.exists(betas_folder)) dir.create(betas_folder)
  dkt_folder <- paste0(csv_output_folder, "dkt_betas/")
  if (!dir.exists(dkt_folder)) dir.create(dkt_folder)
  slwindow_folder <- paste0(csv_output_folder, "sliding_window/")
  if (!dir.exists(slwindow_folder)) dir.create(slwindow_folder)
  sp_stat_folder <- paste0(csv_output_folder, "spline_stats/")
  if (!dir.exists(sp_stat_folder)) dir.create(sp_stat_folder)

plot_output_folder <- paste0(output_folder, "plots/")
if (!dir.exists(plot_output_folder)) dir.create(plot_output_folder)

  ggseg_plot_folder <- paste0(plot_output_folder, "ggseg_plots/")
  if (!dir.exists(ggseg_plot_folder)) dir.create(ggseg_output_folder)
  cmd_plot_folder <- paste0(plot_output_folder, "cmd_plots/")
  if (!dir.exists(cmd_plot_folder)) dir.create(cmd_plot_folder)
  modelvis_plot_folder <- paste0(plot_output_folder, "modelvis_plots/")
  if (!dir.exists(modelvis_plot_folder)) dir.create(modelvis_plot_folder)
  slwindow_plot_folder <- paste0(plot_output_folder, "sliding_window_plots/")
  if (!dir.exists(slwindow_plot_folder)) dir.create(slwindow_plot_folder)

```

# Load custom functions
```{r fcts, echo = FALSE}
source(glue("{code_path}sliding_window.R"))
source(glue("{code_path}visualize_model_bartfct.R"))
source(glue("{code_path}vis_spline_noint_EK.R"))
source(glue("{code_path}model_fitting_fcts.R"))
source(glue("{code_path}model_extract_fcts.R"))
source(glue("{code_path}/slip_data_utils.R"))
source(glue("{code_path}/plotting_utils.R"))

panel_save <- function(plot, filename, extension, folder, plot_width = 8, plot_height = 10){
  save_name <- paste0(folder, filename, ".", extension)
  ggsave(save_name, plot = plot, dpi = 300, width = plot_width, height = plot_height)
}
```

# Color Specifications
```{r, echo = TRUE}
limits = c(-0.3, 0.3)
low_color = "#4878CF" # blue
mid_color = "white"
high_color = "#D65F5F" # red

ga_color = "#6D8299" #muted blue
bw_color = "#709085" # muted green
bwp_color = "#A68A64" # muted beige
```

# Read in data files
```{r, echo = FALSE}
full_vars <- read.csv(paste0(vars_path,"all_vars.csv")) %>% pull(x)
global_vars <- read.csv(paste0(vars_path,"global_vars.csv")) %>% pull(x)
vol_vars <- read.csv(paste0(vars_path,"vol_vars.csv")) %>% pull(x)
sa_vars <- read.csv(paste0(vars_path,"sa_vars.csv")) %>% pull(x)
th_vars <- read.csv(paste0(vars_path,"th_vars.csv")) %>% pull(x)
th_vars_median <- c(paste0(th_vars, ".median"))
namemap <- read.csv(glue("{vars_path}name_mapping.csv"))

df.zscore <- read.csv(qnorm_centiles_path)
load(lookup_stats)

#determine if there is both mpr and median for thickness values
median_th_mpr_flag <- any(grepl(".median", names(df.zscore)))

## Define global vars of interest
global_vars_totest <- c("global_eTIV", "global_SubCortGrayVol","global_CortexVol", "global_CerebralWhiteMatterVol", "global_VentricleChoroidVol","global_SurfaceArea", "global_MeanThickness")

global_vars_totest_nothick <- c("global_eTIV", "global_SubCortGrayVol","global_CortexVol", "global_CerebralWhiteMatterVol", "global_VentricleChoroidVol","global_SurfaceArea")
```
# Code to map raw onto z-score
```{r}
#df.zscore %>%  mutate(ga_raw = map2_dbl(gestational_age, "gestational_age", ~ .x*sd_lookup_all[[.y]] + mean_lookup_all[[.y]]))
```

# Set df.zscore with outlier centiles removed
```{r}
df.zscore_full <- df.zscore
rm_sd <- 5
df.zscore <- remove_outliers(df.zscore_full, 
                                   cols = c(full_vars, "bwp_fen"),
                                   limit = rm_sd)
```

# Get N avail for different predictors
```{r}
ga_N <- sum(!is.na(df.zscore$gestational_age))
ga_N_thick <- sum(!is.na(df.zscore$gestational_age) & !is.na(df.zscore$global_MeanThickness))

bw_N <- sum(!is.na(df.zscore$birth_weight_kg))
bw_N_thick <- sum(!is.na(df.zscore$birth_weight_kg) & !is.na(df.zscore$global_MeanThickness))

bwp_N <- sum(!is.na(df.zscore$bwp_fen))
bwp_N_thick <- sum(!is.na(df.zscore$bwp_fen) & !is.na(df.zscore$global_MeanThickness))
```
# Global phenotypes main effects and interactions

## GA

### fit spline models
```{r}
if(median_th_mpr_flag){
  ga_sp_models_eTIV <- fit_bam_models(df.zscore, "global_eTIV", 
                                        predictor = "gestational_age",
                                        covariates = c("qc_var"), save = FALSE)
  
  ga_sp_table_eTIV <- extract_bam_features_main(models = ga_sp_models_eTIV, 
                                                  data = df.zscore, 
                                                  predictor = "gestational_age",
                                                  covariates = c("qc_var"))
  
  ga_sp_models_global <- fit_bam_models(df.zscore,
                                        c("global_SubCortGrayVol","global_CortexVol",
                                          "global_CerebralWhiteMatterVol",
                                          "global_VentricleChoroidVol","global_SurfaceArea"), 
                                        predictor = "gestational_age",
                                        covariates = c("qc_var", "global_eTIV"), save = FALSE)
  
  ga_sp_table_global <- extract_bam_features_main(models = ga_sp_models_global, 
                                                  data = df.zscore, 
                                                  predictor = "gestational_age",
                                                  covariates = c("qc_var", "global_eTIV"))
  
  
  ga_sp_models_meanthick <- fit_bam_models(df.zscore, "global_MeanThickness", 
                                        predictor = "gestational_age",
                                        covariates = c("euler_mean_mpr", "global_eTIV"), save = FALSE)
  
  ga_sp_table_meanthick <- extract_bam_features_main(models = ga_sp_models_meanthick, 
                                                  data = df.zscore, 
                                                  predictor = "gestational_age",
                                                  covariates = c("euler_mean_mpr", "global_eTIV"))
  
  ga_sp_models_global <- c(ga_sp_models_eTIV, ga_sp_models_global, ga_sp_models_meanthick)
  ga_sp_table_global <- rbind(ga_sp_table_eTIV, ga_sp_table_global, ga_sp_table_meanthick)
  
  rm(ga_sp_models_eTIV, ga_sp_table_eTIV, ga_sp_models_meanthick, ga_sp_table_meanthick)
} else {

  ga_sp_models_eTIV <- fit_bam_models(df.zscore, "global_eTIV", 
                                        predictor = "gestational_age",
                                        covariates = c("qc_var"), save = FALSE)
  
  ga_sp_table_eTIV <- extract_bam_features_main(models = ga_sp_models_eTIV, 
                                                  data = df.zscore, 
                                                  predictor = "gestational_age",
                                                  covariates = c("qc_var"))
  
  ga_sp_models_global <- fit_bam_models(df.zscore,
                                        c("global_SubCortGrayVol","global_CortexVol",
                                          "global_CerebralWhiteMatterVol",
                                          "global_VentricleChoroidVol","global_SurfaceArea",
                                          "global_MeanThickness"), 
                                        predictor = "gestational_age",
                                        covariates = c("qc_var", "global_eTIV"), save = FALSE)
  
  ga_sp_table_global <- extract_bam_features_main(models = ga_sp_models, 
                                                  data = df.zscore, 
                                                  predictor = "gestational_age",
                                                  covariates = c("qc_var", "global_eTIV"))
  
  ga_sp_models_global <- c(ga_sp_models_eTIV, ga_sp_models_global)
  ga_sp_table_global <- rbind(ga_sp_table_eTIV, ga_sp_table_global)
  
  rm(ga_sp_models_eTIV, ga_sp_table_eTIV)
}
```

### fit interactions
```{r}
if(median_th_mpr_flag){
  ga_sp_int_models_global <- fit_bam_interactions(df.zscore, global_vars_totest_nothick, 
                                        predictor = "gestational_age",
                                        moderator = "age_days_adj",
                                        covariates = c("qc_var"), save = FALSE)
  
  ga_sp_int_table_global <- extract_bam_features_interaction(models = ga_sp_int_models_global, 
                                                  data = df.zscore, 
                                                  predictor = "gestational_age",
                                                  moderator = "age_days_adj",
                                                  covariates = c("qc_var"))
  
  
  ga_sp_int_models_meanthick <- fit_bam_interactions(df.zscore, "global_MeanThickness", 
                                        predictor = "gestational_age",
                                        moderator = "age_days_adj",
                                        covariates = c("euler_mean_mpr"), save = FALSE)
  
  ga_sp_int_table_meanthick <- extract_bam_features_interaction(models = ga_sp_int_models_meanthick, 
                                                  data = df.zscore, 
                                                  predictor = "gestational_age",
                                                  moderator = "age_days_adj",
                                                  covariates = c("euler_mean_mpr"))
  
  ga_sp_int_models_global <- c(ga_sp_int_models_global, ga_sp_int_models_meanthick)
  ga_sp_int_table_global <- rbind(ga_sp_int_table_global, ga_sp_int_table_meanthick)
  
  rm(ga_sp_int_models_meanthick, ga_sp_int_table_meanthick)
} else {

  ga_sp_int_models_global <- fit_bam_interactions(df.zscore, global_vars_totest, 
                                      predictor = "gestational_age",
                                      moderator = "age_days_adj",
                                      covariates = c("qc_var"), save = FALSE)
  
  ga_sp_int_table_global <- extract_bam_features_interaction(models = ga_sp_int_models_global, 
                                                data = df.zscore, 
                                                predictor = "gestational_age",
                                                moderator = "age_days_adj",
                                                covariates = c("qc_var"))
}
```

## BW

### fit spline models
```{r}
if(median_th_mpr_flag){
  bw_sp_models_global <- fit_bam_models(df.zscore, global_vars_totest_nothick, 
                                        predictor = "birth_weight_kg",
                                        covariates = c("qc_var"), save = FALSE)
  
  bw_sp_table_global <- extract_bam_features_main(models = bw_sp_models_global, 
                                                  data = df.zscore, 
                                                  predictor = "birth_weight_kg",
                                                  covariates = c("qc_var"))
  
  
  bw_sp_models_meanthick <- fit_bam_models(df.zscore, "global_MeanThickness", 
                                        predictor = "birth_weight_kg",
                                        covariates = c("euler_mean_mpr"), save = FALSE)
  
  bw_sp_table_meanthick <- extract_bam_features_main(models = bw_sp_models_meanthick, 
                                                  data = df.zscore, 
                                                  predictor = "birth_weight_kg",
                                                  covariates = c("euler_mean_mpr"))
  
  bw_sp_models_global <- c(bw_sp_models_global, bw_sp_models_meanthick)
  bw_sp_table_global <- rbind(bw_sp_table_global, bw_sp_table_meanthick)
  
  rm(bw_sp_models_meanthick, bw_sp_table_meanthick)
} else {

  bw_sp_models_global <- fit_bam_models(df.zscore, global_vars_totest, 
                                      predictor = "birth_weight_kg",
                                      covariates = c("qc_var"), save = FALSE)
  
  bw_sp_table_global <- extract_bam_features_main(models = bw_sp_models_global, 
                                                data = df.zscore, 
                                                predictor = "birth_weight_kg",
                                                covariates = c("qc_var"))
}
```

### fit interactions
```{r}
if(median_th_mpr_flag){
  bw_sp_int_models_global <- fit_bam_interactions(df.zscore, global_vars_totest_nothick, 
                                        predictor = "birth_weight_kg",
                                        moderator = "age_days_adj",
                                        covariates = c("qc_var"), save = FALSE)
  
  bw_sp_int_table_global <- extract_bam_features_interaction(models = bw_sp_int_models_global, 
                                                  data = df.zscore, 
                                                  predictor = "birth_weight_kg",
                                                  moderator = "age_days_adj",
                                                  covariates = c("qc_var"))
  
  
  bw_sp_int_models_meanthick <- fit_bam_interactions(df.zscore, "global_MeanThickness", 
                                        predictor = "birth_weight_kg",
                                        moderator = "age_days_adj",
                                        covariates = c("euler_mean_mpr"), save = FALSE)
  
  bw_sp_int_table_meanthick <- extract_bam_features_interaction(models = bw_sp_int_models_meanthick, 
                                                  data = df.zscore, 
                                                  predictor = "birth_weight_kg",
                                                  moderator = "age_days_adj",
                                                  covariates = c("euler_mean_mpr"))
  
  bw_sp_int_models_global <- c(bw_sp_int_models_global, bw_sp_int_models_meanthick)
  bw_sp_int_table_global <- rbind(bw_sp_int_table_global, bw_sp_int_table_meanthick)
  
  rm(bw_sp_int_models_meanthick, bw_sp_int_table_meanthick)
} else {

  bw_sp_int_models_global <- fit_bam_interactions(df.zscore, global_vars_totest, 
                                      predictor = "birth_weight_kg",
                                      moderator = "age_days_adj",
                                      covariates = c("qc_var"), save = FALSE)
  
  bw_sp_int_table_global <- extract_bam_features_interaction(models = bw_sp_int_models_global, 
                                                data = df.zscore, 
                                                predictor = "birth_weight_kg",
                                                moderator = "age_days_adj",
                                                covariates = c("qc_var"))
}
```

## BWP

### fit spline models
```{r}
if(median_th_mpr_flag){
  bwp_sp_models_eTIV <- fit_bam_models(df.zscore, "global_eTIV", 
                                        predictor = "bwp_fen",
                                        covariates = c("qc_var"), save = FALSE)
  
  bwp_sp_table_eTIV <- extract_bam_features_main(models = bwp_sp_models_eTIV, 
                                                  data = df.zscore, 
                                                  predictor = "bwp_fen",
                                                  covariates = c("qc_var"))
  
  bwp_sp_models_global <- fit_bam_models(df.zscore,
                                        c("global_SubCortGrayVol","global_CortexVol",
                                          "global_CerebralWhiteMatterVol",
                                          "global_VentricleChoroidVol","global_SurfaceArea"), 
                                        predictor = "bwp_fen",
                                        covariates = c("qc_var", "global_eTIV"), save = FALSE)
  
  bwp_sp_table_global <- extract_bam_features_main(models = bwp_sp_models_global, 
                                                  data = df.zscore, 
                                                  predictor = "bwp_fen",
                                                  covariates = c("qc_var", "global_eTIV"))
  
  
  bwp_sp_models_meanthick <- fit_bam_models(df.zscore, "global_MeanThickness", 
                                        predictor = "bwp_fen",
                                        covariates = c("euler_mean_mpr", "global_eTIV"), save = FALSE)
  
  bwp_sp_table_meanthick <- extract_bam_features_main(models = bwp_sp_models_meanthick, 
                                                  data = df.zscore, 
                                                  predictor = "bwp_fen",
                                                  covariates = c("euler_mean_mpr", "global_eTIV"))
  
  bwp_sp_models_global <- c(bwp_sp_models_eTIV, bwp_sp_models_global, bwp_sp_models_meanthick)
  bwp_sp_table_global <- rbind(bwp_sp_table_eTIV, bwp_sp_table_global, bwp_sp_table_meanthick)
  
  rm(bwp_sp_models_eTIV, bwp_sp_table_eTIV, bwp_sp_models_meanthick, bwp_sp_table_meanthick)
} else {

  bwp_sp_models_eTIV <- fit_bam_models(df.zscore, "global_eTIV", 
                                        predictor = "bwp_fen",
                                        covariates = c("qc_var"), save = FALSE)
  
  bwp_sp_table_eTIV <- extract_bam_features_main(models = bwp_sp_models_eTIV, 
                                                  data = df.zscore, 
                                                  predictor = "bwp_fen",
                                                  covariates = c("qc_var"))
  
  bwp_sp_models_global <- fit_bam_models(df.zscore,
                                        c("global_SubCortGrayVol","global_CortexVol",
                                          "global_CerebralWhiteMatterVol",
                                          "global_VentricleChoroidVol","global_SurfaceArea",
                                          "global_MeanThickness"), 
                                        predictor = "bwp_fen",
                                        covariates = c("qc_var", "global_eTIV"), save = FALSE)
  
  bwp_sp_table_global <- extract_bam_features_main(models = bwp_sp_models, 
                                                  data = df.zscore, 
                                                  predictor = "bwp_fen",
                                                  covariates = c("qc_var", "global_eTIV"))
  
  bwp_sp_models_global <- c(bwp_sp_models_eTIV, bwp_sp_models_global)
  bwp_sp_table_global <- rbind(bwp_sp_table_eTIV, bwp_sp_table_global)
  
  rm(bwp_sp_models_eTIV, bwp_sp_table_eTIV)
}
```

### fit interactions
```{r}
if(median_th_mpr_flag){
  bwp_sp_int_models_global <- fit_bam_interactions(df.zscore, global_vars_totest_nothick, 
                                        predictor = "bwp_fen",
                                        moderator = "age_days_adj",
                                        covariates = c("qc_var"), save = FALSE)
  
  bwp_sp_int_table_global <- extract_bam_features_interaction(models = bwp_sp_int_models_global, 
                                                  data = df.zscore, 
                                                  predictor = "bwp_fen",
                                                  moderator = "age_days_adj",
                                                  covariates = c("qc_var"))
  
  
  bwp_sp_int_models_meanthick <- fit_bam_interactions(df.zscore, "global_MeanThickness", 
                                        predictor = "bwp_fen",
                                        moderator = "age_days_adj",
                                        covariates = c("euler_mean_mpr"), save = FALSE)
  
  bwp_sp_int_table_meanthick <- extract_bam_features_interaction(models = bwp_sp_int_models_meanthick, 
                                                  data = df.zscore, 
                                                  predictor = "bwp_fen",
                                                  moderator = "age_days_adj",
                                                  covariates = c("euler_mean_mpr"))
  
  bwp_sp_int_models_global <- c(bwp_sp_int_models_global, bwp_sp_int_models_meanthick)
  bwp_sp_int_table_global <- rbind(bwp_sp_int_table_global, bwp_sp_int_table_meanthick)
  
  rm(bwp_sp_int_models_meanthick, bwp_sp_int_table_meanthick)
} else {

  bwp_sp_int_models_global <- fit_bam_interactions(df.zscore, global_vars_totest, 
                                      predictor = "bwp_fen",
                                      moderator = "age_days_adj",
                                      covariates = c("qc_var"), save = FALSE)
  
  bwp_sp_int_table_global <- extract_bam_features_interaction(models = bwp_sp_int_models_global, 
                                                data = df.zscore, 
                                                predictor = "bwp_fen",
                                                moderator = "age_days_adj",
                                                covariates = c("qc_var"))
}
```

# Regional phenotypes main effects

## GA
```{r}
for(metric in c("vol", "sa", "th")){

  vars <- get(paste0(metric, "_vars"))

  # covs <- if(metric == "th" && median_th_mpr_flag){
  #   c("euler_mean_mpr")
  # } else {
  #   c("qc_var")
  # }
    
  covs <- if (metric == "th" && median_th_mpr_flag) {
    c("euler_mean_mpr", "global_MeanThickness")
  } else if (metric == "vol") {
    c("qc_var", "global_eTIV")
  } else if (metric == "sa") {
    c("qc_var", "global_SurfaceArea")
  } else if (metric == "th") {
    c("qc_var", "global_MeanThickness")
  } else {
    NULL
  }
  
  suffix <- metric

  assign(paste0("ga_sp_models_", suffix),
         fit_bam_models(df.zscore, 
                        vars,
                        predictor = "gestational_age",
                        covariates = covs, save = FALSE))

  assign(paste0("ga_sp_table_", suffix),
         extract_bam_features_main(
           models = get(paste0("ga_sp_models_", suffix)),
           data = df.zscore,
           predictor = "gestational_age",
           covariates = covs
         ))
}

# Add label column that matches ggseg for plotting

## volume
cols_edit <- grepl("aseg(?!_CC)", rownames(ga_sp_table_vol), perl = TRUE)
ga_sp_table_vol$label <- rownames(ga_sp_table_vol) %>% 
  sub("aparc_GrayVol_", "", .) %>% 
  sub("aseg_", "", .) %>%
  sub("X", "x", .) %>%
  sub("rd_Ventricle", "rd_ventricle", .) %>%     sub("th_Ventricle", "th_ventricle", .) %>% sub("Thalamus", "Thalamus-Proper", .)
ga_sp_table_vol$label[cols_edit] <- gsub("_", "-", ga_sp_table_vol$label[cols_edit])

## surface area
ga_sp_table_sa$label <- rownames(ga_sp_table_sa) %>% 
  sub("aparc_SurfArea_", "", .)

## thickness
ga_sp_table_th$label <- rownames(ga_sp_table_th) %>% 
  sub("aparc_ThickAvg_", "", .)
```

## BW
```{r}
for(metric in c("vol", "sa", "th")){

  vars <- get(paste0(metric, "_vars"))

  covs <- if (metric == "th" && median_th_mpr_flag) {
    c("euler_mean_mpr", "global_MeanThickness")
  } else if (metric == "vol") {
    c("qc_var", "global_eTIV")
  } else if (metric == "sa") {
    c("qc_var", "global_SurfaceArea")
  } else if (metric == "th") {
    c("qc_var", "global_MeanThickness")
  } else {
    NULL
  }

  suffix <- metric

  assign(paste0("bw_sp_models_", suffix),
         fit_bam_models(df.zscore, vars,
                        predictor = "birth_weight_kg",
                        covariates = covs, save = FALSE))

  assign(paste0("bw_sp_table_", suffix),
         extract_bam_features_main(
           models = get(paste0("bw_sp_models_", suffix)),
           data = df.zscore,
           predictor = "birth_weight_kg",
           covariates = covs
         ))
}

# Add label column that matches ggseg for plotting

## volume
cols_edit <- grepl("aseg(?!_CC)", rownames(bw_sp_table_vol), perl = TRUE)
bw_sp_table_vol$label <- rownames(bw_sp_table_vol) %>% 
  sub("aparc_GrayVol_", "", .) %>% 
  sub("aseg_", "", .) %>%
  sub("X", "x", .) %>%
  sub("rd_Ventricle", "rd_ventricle", .) %>%     sub("th_Ventricle", "th_ventricle", .) %>% sub("Thalamus", "Thalamus-Proper", .)
bw_sp_table_vol$label[cols_edit] <- gsub("_", "-", bw_sp_table_vol$label[cols_edit])

## surface area
bw_sp_table_sa$label <- rownames(bw_sp_table_sa) %>% 
  sub("aparc_SurfArea_", "", .)

## thickness
bw_sp_table_th$label <- rownames(bw_sp_table_th) %>% 
  sub("aparc_ThickAvg_", "", .)
```

## BWP
```{r}
for(metric in c("vol", "sa", "th")){

  vars <- get(paste0(metric, "_vars"))

  covs <- if (metric == "th" && median_th_mpr_flag) {
    c("euler_mean_mpr", "global_MeanThickness")
  } else if (metric == "vol") {
    c("qc_var", "global_eTIV")
  } else if (metric == "sa") {
    c("qc_var", "global_SurfaceArea")
  } else if (metric == "th") {
    c("qc_var", "global_MeanThickness")
  } else {
    NULL
  }
  
  suffix <- metric

  assign(paste0("bwp_sp_models_", suffix),
         fit_bam_models(df.zscore, vars,
                        predictor = "bwp_fen",
                        covariates = covs, save = FALSE))

  assign(paste0("bwp_sp_table_", suffix),
         extract_bam_features_main(
           models = get(paste0("bwp_sp_models_", suffix)),
           data = df.zscore,
           predictor = "bwp_fen",
           covariates = covs
         ))
}

# Add label column that matches ggseg for plotting

## volume
cols_edit <- grepl("aseg(?!_CC)", rownames(bwp_sp_table_vol), perl = TRUE)
bwp_sp_table_vol$label <- rownames(bwp_sp_table_vol) %>% 
  sub("aparc_GrayVol_", "", .) %>% 
  sub("aseg_", "", .) %>%
  sub("X", "x", .) %>%
  sub("rd_Ventricle", "rd_ventricle", .) %>%     sub("th_Ventricle", "th_ventricle", .) %>% sub("Thalamus", "Thalamus-Proper", .)
bwp_sp_table_vol$label[cols_edit] <- gsub("_", "-", bwp_sp_table_vol$label[cols_edit])

## surface area
bwp_sp_table_sa$label <- rownames(bwp_sp_table_sa) %>% 
  sub("aparc_SurfArea_", "", .)

## thickness
bwp_sp_table_th$label <- rownames(bwp_sp_table_th) %>% 
  sub("aparc_ThickAvg_", "", .)
```


# Regional phenotypes interactions

## set moderator variable
```{r}
mod_var <- "age_days_adj"
```

## GA
```{r}
for(metric in c("vol", "sa", "th")){

  vars <- get(paste0(metric, "_vars"))

  covs <- if (metric == "th" && median_th_mpr_flag) {
    c("euler_mean_mpr", "global_MeanThickness")
  } else if (metric == "vol") {
    c("qc_var", "global_eTIV")
  } else if (metric == "sa") {
    c("qc_var", "global_SurfaceArea")
  } else if (metric == "th") {
    c("qc_var", "global_MeanThickness")
  } else {
    NULL
  }
  suffix <- metric

  assign(paste0("ga_sp_int_models_", suffix),
         fit_bam_interactions(
           df.zscore,
           phenotypes = vars,
           predictor = "gestational_age",
           moderator = mod_var,
           covariates = covs,
           save = FALSE
         ))

  assign(paste0("ga_sp_int_table_", suffix),
         extract_bam_features_interaction(
           models = get(paste0("ga_sp_int_models_", suffix)),
           data = df.zscore,
           predictor = "gestational_age",
           moderator = mod_var,
           covariates = covs
         ))
}

## volume labels
cols_edit <- grepl("aseg(?!_CC)", rownames(ga_sp_int_table_vol), perl = TRUE)

ga_sp_int_table_vol$label <- rownames(ga_sp_int_table_vol) %>% 
  sub("aparc_GrayVol_", "", .) %>% 
  sub("aseg_", "", .) %>%
  sub("X", "x", .) %>%
  sub("rd_Ventricle", "rd_ventricle", .) %>%
  sub("th_Ventricle", "th_ventricle", .) %>%
  sub("Thalamus", "Thalamus-Proper", .)

ga_sp_int_table_vol$label[cols_edit] <- gsub("_", "-", ga_sp_int_table_vol$label[cols_edit])

## surface area
ga_sp_int_table_sa$label <- rownames(ga_sp_int_table_sa) %>% 
  sub("aparc_SurfArea_", "", .)

## thickness
ga_sp_int_table_th$label <- rownames(ga_sp_int_table_th) %>% 
  sub("aparc_ThickAvg_", "", .)
```

## BW
```{r}
for(metric in c("vol", "sa", "th")){

  vars <- get(paste0(metric, "_vars"))
  
  covs <- if (metric == "th" && median_th_mpr_flag) {
    c("euler_mean_mpr", "global_MeanThickness")
  } else if (metric == "vol") {
    c("qc_var", "global_eTIV")
  } else if (metric == "sa") {
    c("qc_var", "global_SurfaceArea")
  } else if (metric == "th") {
    c("qc_var", "global_MeanThickness")
  } else {
    NULL
  }
  suffix <- metric

  assign(paste0("bw_sp_int_models_", suffix),
         fit_bam_interactions(
           df.zscore,
           phenotypes = vars,
           predictor = "birth_weight_kg",
           moderator = mod_var,
           covariates = covs,
           save = FALSE
         ))

  assign(paste0("bw_sp_int_table_", suffix),
         extract_bam_features_interaction(
           models = get(paste0("bw_sp_int_models_", suffix)),
           data = df.zscore,
           predictor = "birth_weight_kg",
           moderator = mod_var,
           covariates = covs
         ))
}

## volume labels
cols_edit <- grepl("aseg(?!_CC)", rownames(bw_sp_int_table_vol), perl = TRUE)

bw_sp_int_table_vol$label <- rownames(bw_sp_int_table_vol) %>% 
  sub("aparc_GrayVol_", "", .) %>% 
  sub("aseg_", "", .) %>%
  sub("X", "x", .) %>%
  sub("rd_Ventricle", "rd_ventricle", .) %>%
  sub("th_Ventricle", "th_ventricle", .) %>%
  sub("Thalamus", "Thalamus-Proper", .)

bw_sp_int_table_vol$label[cols_edit] <- gsub("_", "-", bw_sp_int_table_vol$label[cols_edit])

## surface area
bw_sp_int_table_sa$label <- rownames(bw_sp_int_table_sa) %>% 
  sub("aparc_SurfArea_", "", .)

## thickness
bw_sp_int_table_th$label <- rownames(bw_sp_int_table_th) %>% 
  sub("aparc_ThickAvg_", "", .)
```

## BWP
```{r}
for(metric in c("vol", "sa", "th")){

  vars <- get(paste0(metric, "_vars"))
  
  covs <- if (metric == "th" && median_th_mpr_flag) {
    c("euler_mean_mpr", "global_MeanThickness")
  } else if (metric == "vol") {
    c("qc_var", "global_eTIV")
  } else if (metric == "sa") {
    c("qc_var", "global_SurfaceArea")
  } else if (metric == "th") {
    c("qc_var", "global_MeanThickness")
  } else {
    NULL
  }

  suffix <- metric

  assign(paste0("bwp_sp_int_models_", suffix),
         fit_bam_interactions(
           df.zscore,
           phenotypes = vars,
           predictor = "bwp_fen",
           moderator = mod_var,
           covariates = covs,
           save = FALSE
         ))

  assign(paste0("bwp_sp_int_table_", suffix),
         extract_bam_features_interaction(
           models = get(paste0("bwp_sp_int_models_", suffix)),
           data = df.zscore,
           predictor = "bwp_fen",
           moderator = mod_var,
           covariates = covs
         ))
}

## volume labels
cols_edit <- grepl("aseg(?!_CC)", rownames(bwp_sp_int_table_vol), perl = TRUE)

bwp_sp_int_table_vol$label <- rownames(bwp_sp_int_table_vol) %>% 
  sub("aparc_GrayVol_", "", .) %>% 
  sub("aseg_", "", .) %>%
  sub("X", "x", .) %>%
  sub("rd_Ventricle", "rd_ventricle", .) %>%
  sub("th_Ventricle", "th_ventricle", .) %>%
  sub("Thalamus", "Thalamus-Proper", .)

bwp_sp_int_table_vol$label[cols_edit] <- gsub("_", "-", bwp_sp_int_table_vol$label[cols_edit])

## surface area
bwp_sp_int_table_sa$label <- rownames(bwp_sp_int_table_sa) %>% 
  sub("aparc_SurfArea_", "", .)

## thickness
bwp_sp_int_table_th$label <- rownames(bwp_sp_int_table_th) %>% 
  sub("aparc_ThickAvg_", "", .)
```

# Plots

## Global Grouped Plots
```{r}
tables_list <- list(
  GA = ga_sp_table_global %>% mutate(phenotype = rownames(.)),
  BWP = bwp_sp_table_global %>% mutate(phenotype = rownames(.))
)

color_vec <- c(GA = ga_color, BWP = bw_color)

## Grouped lolli
ga_bwp_global_partialr2_plot <- plot_grouped_lollipop(
  table_list = tables_list,
  value_col = "r2_full_vs_none",
  p_col = "full_p_adj",
  color_vec = color_vec,
  var_names = global_vars_totest,
  name_mapping = namemap,
  line_width = 2,
  point_size = 10
) +
  labs(x = "Partial R2") +
  theme_minimal(base_family = "Helvetica", 
                base_size = 20) +
  theme(legend.position = "bottom",
        panel.grid.major.x = element_line(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()
        )

ga_bwp_global_partialr2_plot

if(save_figs){
  panel_save(ga_bwp_global_partialr2_plot, "ga_bwp_global_partialr2_eTIVcovar_plot", "svg", plot_output_folder,
             plot_width = 16, plot_height = 10)
}
```


## Brain Maps for main effects partial r2
```{r}
(ga_partialr2_dkt_vol <- plot_brain_map(
  table = ga_sp_table_vol,
  value_col = "r2_full_vs_none",
  atlas = "dk",
  p_col = "full_p_adj",
  p_filter = TRUE,
  title = "GA partial R2 vol",
  limits = c(0, 0.025)
))

(ga_partialr2_aseg_vol <- plot_brain_map(
  table = ga_sp_table_vol,
  value_col = "r2_full_vs_none",
  atlas = "aseg",
  p_col = "full_p_adj",
  p_filter = TRUE,
  title = "GA partial R2 vol",
  limits = c(0, 0.025)
))

(ga_partialr2_sa <- plot_brain_map(
  table = ga_sp_table_sa,
  value_col = "r2_full_vs_none",
  atlas = "dk",
  p_col = "full_p_adj",
  p_filter = TRUE,
  title = "GA partial R2 sa",
  limits = c(0, 0.025)
))

(ga_partialr2_th <- plot_brain_map(
  table = ga_sp_table_th,
  value_col = "r2_full_vs_none",
  atlas = "dk",
  p_col = "full_p_adj",
  p_filter = TRUE,
  title = "GA partial R2 th",
  limits = c(0, 0.025)
))

(bwp_partialr2_dkt_vol <- plot_brain_map(
  table = bwp_sp_table_vol,
  value_col = "r2_full_vs_none",
  atlas = "dk",
  p_col = "full_p_adj",
  p_filter = TRUE,
  title = "bwp partial R2 vol",
  limits = c(0, 0.025)
))

(bwp_partialr2_aseg_vol <- plot_brain_map(
  table = bwp_sp_table_vol,
  value_col = "r2_full_vs_none",
  atlas = "aseg",
  p_col = "full_p_adj",
  p_filter = TRUE,
  title = "bwp partial R2 vol",
  limits = c(0, 0.025)
))

(bwp_partialr2_sa <- plot_brain_map(
  table = bwp_sp_table_sa,
  value_col = "r2_full_vs_none",
  atlas = "dk",
  p_col = "full_p_adj",
  p_filter = TRUE,
  title = "bwp partial R2 sa",
  limits = c(0, 0.025)
))

(bwp_partialr2_th <- plot_brain_map(
  table = bwp_sp_table_th,
  value_col = "r2_full_vs_none",
  atlas = "dk",
  p_col = "full_p_adj",
  p_filter = TRUE,
  title = "bwp partial R2 th",
  limits = c(0, 0.025)
))

plots<- list(ga_partialr2_dkt_vol, ga_partialr2_sa, ga_partialr2_th, ga_partialr2_aseg_vol,
                                      bwp_partialr2_dkt_vol, bwp_partialr2_sa, bwp_partialr2_th, bwp_partialr2_aseg_vol)

plots <- lapply(plots, function(p) {
  p +
    labs(title = NULL) +
    theme(
      legend.position = "bottom"
    )
})

ga_bwp_partialr2_grid <- wrap_plots(plots, ncol = 4) +
  plot_layout(guides = "collect") &
  theme(
    legend.position = "bottom"
  )

if(save_figs){
  panel_save(ga_bwp_partialr2_grid, "ga_bwp_partialr2_grid", "svg", plot_output_folder,
             plot_width = 16, plot_height = 10)
}
```

## Heat Map for Partial R
```{r}
tables_list <- list(
  GA = ga_sp_table_th %>% mutate(phenotype = rownames(.)),
  BWP = bwp_sp_table_th %>% mutate(phenotype = rownames(.))
)

plot_heatmap(tables_list,
             "r2_full_vs_none",
             p_col = "full_p_adj",
             p_filter = TRUE,
             name_mapping = namemap,
             low_col ="#FFF1E9",
             mid_col = "#FFAE7F",
             high_col = "#EB680E")
```

## Brain Maps for main effects average first derivative
```{r}
(ga_avgder_dkt_vol <- plot_brain_map(
  table = ga_sp_table_vol,
  value_col = "avg_first_deriv",
  atlas = "dk",
  p_col = "full_p_adj",
  p_filter = TRUE,
  title = "GA partial R2 vol",
  limits = c(-0.18, 0.18),
  low_color = "#2160B8",
  mid_color = "white",
  high_color = "#B82133",
  na_color = "grey90"
))

(ga_avgder_aseg_vol <- plot_brain_map(
  table = ga_sp_table_vol,
  value_col = "avg_first_deriv",
  atlas = "aseg",
  p_col = "full_p_adj",
  p_filter = TRUE,
  title = "GA partial R2 vol",
  limits = c(0, 0.075)
))

(ga_avgder_sa <- plot_brain_map(
  table = ga_sp_table_sa,
  value_col = "avg_first_derivative",
  atlas = "dk",
  p_col = "full_p_adj",
  p_filter = TRUE,
  title = "GA partial R2 sa",
  limits = c(0, 0.075)
))

(ga_avgder_th <- plot_brain_map(
  table = ga_sp_table_th,
  value_col = "avg_first_derivative",
  atlas = "dk",
  p_col = "full_p_adj",
  p_filter = TRUE,
  title = "GA partial R2 th",
  limits = c(0, 0.075)
))

(bwp_avgder_dkt_vol <- plot_brain_map(
  table = bwp_sp_table_vol,
  value_col = "avg_first_derivative",
  atlas = "dk",
  p_col = "full_p_adj",
  p_filter = TRUE,
  title = "bwp partial R2 vol",
  limits = c(0, 0.075)
))

(bwp_avgder_aseg_vol <- plot_brain_map(
  table = bwp_sp_table_vol,
  value_col = "avg_first_derivative",
  atlas = "aseg",
  p_col = "full_p_adj",
  p_filter = TRUE,
  title = "bwp partial R2 vol",
  limits = c(0, 0.075)
))

(bwp_avgder_sa <- plot_brain_map(
  table = bwp_sp_table_sa,
  value_col = "avg_first_derivative",
  atlas = "dk",
  p_col = "full_p_adj",
  p_filter = TRUE,
  title = "bwp partial R2 sa",
  limits = c(0, 0.075)
))

(bwp_avgder_th <- plot_brain_map(
  table = bwp_sp_table_th,
  value_col = "avg_first_derivative",
  atlas = "dk",
  p_col = "full_p_adj",
  p_filter = TRUE,
  title = "bwp partial R2 th",
  limits = c(0, 0.075)
))

plots<- list(ga_avgder_dkt_vol, ga_avgder_sa, ga_avgder_th, ga_avgder_aseg_vol,
                                      bwp_avgder_dkt_vol, bwp_avgder_sa, bwp_avgder_th, bwp_avgder_aseg_vol)

plots <- lapply(plots, function(p) {
  p +
    labs(title = NULL) +
    theme(
      legend.position = "bottom"
    )
})

ga_bwp_avgder_grid <- wrap_plots(plots, ncol = 4) +
  plot_layout(guides = "collect") &
  theme(
    legend.position = "bottom"
  )

if(save_figs){
  panel_save(ga_bwp_avgder_grid, "ga_bwp_avgder_grid", "svg", plot_output_folder,
             plot_width = 16, plot_height = 10)
}
```

## Grouped for thickness
```{r}
tables_list <- list(
  GA = ga_sp_table_th %>% mutate(phenotype = rownames(.)),
  BWP = bwp_sp_table_th %>% mutate(phenotype = rownames(.))
)

pheno_list <- unique(c(tables_list$GA %>% filter(full_p_adj < 0.05) %>% pull(phenotype),
                       tables_list$BWP %>% filter(full_p_adj < 0.05) %>% pull(phenotype)))

color_vec <- c(GA = ga_color, BWP = bw_color)

## Grouped lollipop
(ga_bwp_th_partialr2_plot <- plot_grouped_lollipop(
  table_list = tables_list,
  value_col = "r2_full_vs_none",
  p_col = "full_p_adj",
  color_vec = color_vec,
  var_names = pheno_list,
  name_mapping = namemap
) +
  labs(x = "Partial R2") +
  theme_minimal(base_family = "Helvetica", 
                base_size = 20) +
  theme(legend.position = "bottom",
        panel.grid.major.x = element_line(),
        panel.grid.major.y = element_line(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()
        )
)

if(save_figs){
  panel_save(ga_bwp_th_partialr2_plot, "ga_bwp_th_partialr2_plot", "svg", plot_output_folder,
             plot_width = 16, plot_height = 10)
}
```

## Brain Maps for interactions
```{r}
(ga_partialr2_dkt_vol <- plot_brain_map(
  table = ga_sp_int_table_vol,
  value_col = "partial_r2",
  atlas = "dk",
  p_col = "int_p_adj",
  p_filter = TRUE,
  title = "GA partial R2 vol",
  limits = c(0, 0.005)
))

(ga_partialr2_aseg_vol <- plot_brain_map(
  table = ga_sp_int_table_vol,
  value_col = "partial_r2",
  atlas = "aseg",
  p_col = "int_p_adj",
  p_filter = TRUE,
  title = "GA partial R2 vol",
  limits = c(0, 0.005)
))

(ga_partialr2_sa <- plot_brain_map(
  table = ga_sp_int_table_sa,
  value_col = "partial_r2",
  atlas = "dk",
  p_col = "int_p_adj",
  p_filter = TRUE,
  title = "GA partial R2 sa",
  limits = c(0, 0.005)
))

(ga_partialr2_th <- plot_brain_map(
  table = ga_sp_int_table_th,
  value_col = "partial_r2",
  atlas = "dk",
  p_col = "int_p_adj",
  p_filter = TRUE,
  title = "GA partial R2 th",
  limits = c(0, 0.005)
))

(bwp_partialr2_dkt_vol <- plot_brain_map(
  table = bwp_sp_int_table_vol,
  value_col = "partial_r2",
  atlas = "dk",
  p_col = "int_p_adj",
  p_filter = TRUE,
  title = "bwp partial R2 vol",
  limits = c(0, 0.005)
))

(bwp_partialr2_aseg_vol <- plot_brain_map(
  table = bwp_sp_int_table_vol,
  value_col = "partial_r2",
  atlas = "aseg",
  p_col = "int_p_adj",
  p_filter = TRUE,
  title = "bwp partial R2 vol",
  limits = c(0, 0.005)
))

(bwp_partialr2_sa <- plot_brain_map(
  table = bwp_sp_int_table_sa,
  value_col = "partial_r2",
  atlas = "dk",
  p_col = "int_p_adj",
  p_filter = TRUE,
  title = "bwp partial R2 sa",
  limits = c(0, 0.005)
))

(bwp_partialr2_th <- plot_brain_map(
  table = bwp_sp_int_table_th,
  value_col = "partial_r2",
  atlas = "dk",
  p_col = "int_p_adj",
  p_filter = TRUE,
  title = "bwp partial R2 th",
  limits = c(0, 0.005)
))

plots<- list(ga_partialr2_dkt_vol, ga_partialr2_sa, ga_partialr2_th, ga_partialr2_aseg_vol,
                                      bwp_partialr2_dkt_vol, bwp_partialr2_sa, bwp_partialr2_th, bwp_partialr2_aseg_vol)

plots <- lapply(plots, function(p) {
  p +
    labs(title = NULL) +
    theme(
      legend.position = "bottom"
    )
})

ga_bwp_int_partialr2_grid <- wrap_plots(plots, ncol = 4) +
  plot_layout(guides = "collect") &
  theme(
    legend.position = "bottom"
  )

if(save_figs){
  panel_save(ga_bwp_int_partialr2_grid, "ga_bwp_int_partialr2_grid", "svg", plot_output_folder,
             plot_width = 16, plot_height = 10)
}
```

## Plots for "regions of change"

### global
```{r}
ga_sig_pheno_global <- rownames(ga_sp_table_global %>% filter(full_p_adj < 0.05) %>% arrange(r2_full_vs_none))
plot_region_of_change(ga_sp_table_global,
                      var_names = ga_sig_pheno_global,
                      name_mapping = namemap,
                      alpha_range = c(0.2, 1),
                      positive_color = "#C95A5A",
                      negative_color = "#5A7BC9") +
  scale_x_continuous(
      #breaks = map2_dbl(c(24*7, 28*7, 32*7, 36*7, 40*7), "gestational_age", ~ (.x - mean_lookup_distinct[[.y]])/(sd_lookup_distinct[[.y]])),
      labels = \(x) round(map2_dbl(x, "gestational_age", ~ (.x*sd_lookup_distinct[[.y]]) + mean_lookup_distinct[[.y]])/7, 1) ) +
  labs(x = "Gestational Age (weeks)")
```

### volume cortical
```{r}
ga_sig_pheno_vol_cortical <- rownames(head(ga_sp_table_vol %>% filter(full_p_adj < 0.05 & !grepl("aseg", rownames(.))) %>% arrange(desc(r2_full_vs_none)), 20))
plot_region_of_change(ga_sp_table_vol,
                      var_names = ga_sig_pheno_vol_cortical,
                      name_mapping = namemap,
                      alpha_range = c(0.2, 1),
                      positive_color = "#C95A5A",
                      negative_color = "#5A7BC9",
                      segment_width = 3,
                      alpha_order = TRUE) +
  scale_x_continuous(
      labels = \(x) round(map2_dbl(x, "gestational_age", ~ (.x*sd_lookup_distinct[[.y]]) + mean_lookup_distinct[[.y]])/7, 1) ) +
  labs(x = "Gestational Age (weeks)")
```

### thickness
```{r}
ga_sig_pheno_th <- rownames(head(ga_sp_table_th %>% filter(full_p_adj < 0.05) %>% arrange(desc(r2_full_vs_none)), 20))
plot_region_of_change(ga_sp_table_th,
                      var_names = ga_sig_pheno_th,
                      name_mapping = namemap,
                      alpha_range = c(0.2, 1),
                      positive_color = "#C95A5A",
                      negative_color = "#5A7BC9",
                      segment_width = 3,
                      alpha_order = TRUE) +
  scale_x_continuous(
      labels = \(x) round(map2_dbl(x, "gestational_age", ~ (.x*sd_lookup_distinct[[.y]]) + mean_lookup_distinct[[.y]])/7, 1) ) +
  labs(x = "Gestational Age (weeks)")
```


### Check models for weird ones
```{r}
visualize_spline_model(
  modobj = ga_sp_models_vol$aparc_GrayVol_rh_pericalcarine$full,
  plot_og_scale = FALSE,
  mean_lookup = mean_lookup_distinct,
  sd_lookup = sd_lookup_distinct,
  pnorm_smooth_var = FALSE,
  smooth_var = "gestational_age",
  plabels = "Gestational Age",
  color_point = ga_color,
  color_line = ga_color)
```

## Model Visualizations

### GA
#### Visualize on z-score scale

##### Global
```{r}
ga_sp_plots_global <- lapply(ga_sp_models_global, function(x) {
  
  modobj = x$full
  
  model_plot <- visualize_spline_model(
  modobj = modobj,
  plot_og_scale = FALSE,
  mean_lookup = mean_lookup_distinct,
  sd_lookup = sd_lookup_distinct,
  pnorm_smooth_var = FALSE,
  smooth_var = "gestational_age",
  plabels = "Gestational Age",
  color_point = ga_color,
  color_line = ga_color)
  
  model_plot <- model_plot + 
    scale_x_continuous(
      breaks = map2_dbl(c(24*7, 28*7, 32*7, 36*7, 40*7), "gestational_age", ~ (.x - mean_lookup_distinct[[.y]])/(sd_lookup_distinct[[.y]])),
      labels = \(x) map2_dbl(x, "gestational_age", ~ (.x*sd_lookup_distinct[[.y]]) + mean_lookup_distinct[[.y]])/7) +
    scale_y_continuous(breaks = qnorm(c(.01, .50, .99)), 
                       labels = c("<=1st",  "50th", ">=99th")) +
    labs(x= "Gestational Age (weeks)",
         title = glue("{modobj$gam$terms[[2]]}")) +
    theme(axis.title.y = element_blank())
})

ga_sp_plots_global <- setNames(ga_sp_plots_global, global_vars_totest)

if(save_figs){
saveRDS(ga_sp_plots_global, glue("{modelvis_plot_folder}ga_sp_plots_global.rds"))
}  
```

##### Choose regions to plot

###### choose regions
```{r}
```

```{r}
pheno_sig_th_ga <- rownames(ga_sp_table_th %>% filter(full_p_adj < 0.05))
ga_sp_plots_th <- lapply(ga_sp_models_th[pheno_sig_th_ga], function(x) {
  
  modobj = x$full
  
  model_plot <- visualize_spline_model(
  modobj = modobj,
  plot_og_scale = FALSE,
  mean_lookup = mean_lookup_distinct,
  sd_lookup = sd_lookup_distinct,
  pnorm_smooth_var = FALSE,
  smooth_var = "gestational_age",
  plabels = "Gestational Age",
  color_point = ga_color,
  color_line = ga_color)
  
  model_plot <- model_plot + 
    scale_x_continuous(
      breaks = map2_dbl(c(24*7, 28*7, 32*7, 36*7, 40*7), "gestational_age", ~ (.x - mean_lookup_distinct[[.y]])/(sd_lookup_distinct[[.y]])),
      labels = \(x) map2_dbl(x, "gestational_age", ~ (.x*sd_lookup_distinct[[.y]]) + mean_lookup_distinct[[.y]])/7) +
    scale_y_continuous(breaks = qnorm(c(.01, .50, .99)), 
                       labels = c("<=1st",  "50th", ">=99th")) +
    labs(x= "Gestational Age (weeks)",
         title = glue("{modobj$gam$terms[[2]]}")) +
    theme(axis.title.y = element_blank())
})

ga_sp_plots_th <- setNames(ga_sp_plots_th, pheno_sig_th_ga)

if(save_figs){
saveRDS(ga_sp_plots_th, glue("{modelvis_plot_folder}ga_sp_plots_th.rds"))
}  
```

#### Show plots
```{r, echo = TRUE}
wrap_plots(ga_sp_plots_global[1:4], ncol = 2)
wrap_plots(ga_sp_plots_global[5:7], ncol = 2)
```

### BW
#### Visualize on z-score scale
```{r}
bw_sp_plots_global <- lapply(bw_sp_models_global, function(x) {
  
  modobj = x$full
  
model_plot <- visualize_spline_model(
modobj = modobj,
smooth_var = "birth_weight_kg",
plot_og_scale = FALSE,
mean_lookup = mean_lookup_distinct,
sd_lookup = sd_lookup_distinct,
pnorm_smooth_var = FALSE,
plabels = "Birth Weight",
color_point = bw_color,
color_line = bw_color)
  
  model_plot <- model_plot + 
    scale_x_continuous(
      breaks = map2_dbl(c(1.5, 3, 4.5, 6), "birth_weight_kg", ~ (.x - mean_lookup_distinct[[.y]])/(sd_lookup_distinct[[.y]])),
      labels = \(x) map2_dbl(x, "birth_weight_kg", ~ (.x*sd_lookup_distinct[[.y]]) + mean_lookup_distinct[[.y]])) +
    scale_y_continuous(breaks = qnorm(c(.01, .50, .99)), 
                       labels = c("<=1st",  "50th", ">=99th")) +
    labs(x="Birth Weight (kg))",
         title = glue("{modobj$gam$terms[[2]]}")) +
    theme(axis.title.y = element_blank())
})

bw_sp_plots_global <- setNames(bw_sp_plots_global, global_vars_totest)

if(save_figs){
saveRDS(bw_sp_plots_global, glue("{modelvis_plot_folder}bw_sp_plots_global.rds"))
}
```

#### Show plots
```{r, echo = TRUE}
wrap_plots(bw_sp_plots_global[1:4], ncol = 2)
wrap_plots(bw_sp_plots_global[5:7], ncol = 2)
```




### BWP

#### Visualize on z-score scale

##### Global
```{r}
bwp_sp_plots_global <- lapply(bwp_sp_models_global, function(x) {
  
  modobj = x$full
  
model_plot <- visualize_spline_model(
modobj = modobj,
smooth_var = "bwp_fen",
plot_og_scale = FALSE,
mean_lookup = mean_lookup_distinct,
sd_lookup = sd_lookup_distinct,
pnorm_smooth_var = FALSE,
plabels = glue("Birth Weight Percentile \n (GA-corrected)"),
color_point = bw_color,
color_line = bw_color)
  
  model_plot <- model_plot + 
    scale_x_continuous(breaks = qnorm(c(.01, .50, .99)), 
                       labels = c("<=1st",  "50th", ">=99th")) +
        scale_y_continuous(breaks = qnorm(c(.01, .50, .99)), 
                       labels = c("<=1st",  "50th", ">=99th")) +
    labs(x = glue("Birth Weight Percentile \n (GA-corrected)"),
         title = glue("{modobj$gam$terms[[2]]}")) +
    theme(axis.title.y = element_blank())
})

bwp_sp_plots_global <- setNames(bwp_sp_plots_global, global_vars_totest)

if(save_figs){
saveRDS(bwp_sp_plots_global, glue("{modelvis_plot_folder}bwp_sp_plots_global.rds"))
}
```

##### Thickness
```{r}
pheno_sig_th_bwp <- rownames(ga_sp_table_th %>% filter(full_p_adj < 0.05))

bwp_sp_plots_th <- lapply(bwp_sp_models_th[pheno_sig_th_bwp], function(x) {
  
  modobj = x$full
  
model_plot <- visualize_spline_model(
modobj = modobj,
smooth_var = "bwp_fen",
plot_og_scale = FALSE,
mean_lookup = mean_lookup_distinct,
sd_lookup = sd_lookup_distinct,
pnorm_smooth_var = FALSE,
plabels = glue("Birth Weight Percentile \n (GA-corrected)"),
color_point = bw_color,
color_line = bw_color)
  
  model_plot <- model_plot + 
    scale_x_continuous(
      labels = \(x) glue("{round(pnorm(x),2)*100}th") ) +
    scale_y_continuous(labels = \(x) glue("{round(pnorm(x),2)*100}th")) +
    labs(x = glue("Birth Weight Percentile \n (GA-corrected)"),
         title = glue("{modobj$gam$terms[[2]]}")) +
    theme(axis.title.y = element_blank())
})

bwp_sp_plots_th <- setNames(bwp_sp_plots_th, pheno_sig_th_bwp)

if(save_figs){
saveRDS(bwp_sp_plots_th, glue("{modelvis_plot_folder}bwp_sp_plots_th.rds"))
}
```


#### Show plots
```{r, echo = TRUE}
wrap_plots(bwp_sp_plots_global[1:4], ncol = 2)
wrap_plots(bwp_sp_plots_global[5:7], ncol = 2)
```




## Visualize Mean Thickness for GA as interaction
```{r}
font_size <- 16
theme_set(theme_classic(base_family = "Helvetica",base_size = font_size))
line_size <- 1.5
point_size <- 2

mu_age <- mean_lookup_distinct["age_days_adj"]
sd_age <- sd_lookup_distinct["age_days_adj"]

mu_ga <- mean_lookup_distinct["gestational_age"]
sd_ga <- sd_lookup_distinct["gestational_age"]
  
  low_age_label <- (quantile(df.zscore[!is.na(df.zscore$gestational_age),
                                       "age_days_adj"],
                             probs = c(.10)) * sd_age + mu_age)/365.25
  
  mid_age_label <- (quantile(df.zscore[!is.na(df.zscore$gestational_age),
                                       "age_days_adj"],
                             probs = c(.50)) * sd_age + mu_age)/365.25

  high_age_label <- (quantile(df.zscore[!is.na(df.zscore$gestational_age),
                                      "age_days_adj"],
                            probs = c(.90)) * sd_age + mu_age)/365.25
  
  model_plot <- visualize_model(
  modobj = ga_sp_int_models_global$global_MeanThickness,
  smooth_var = "gestational_age",
  int_var = "age_days_adj",
  plabels = "continuous interaction",
  derivative_plot = F)

  plot_colors <- ggplot_build(model_plot)$plot$scales$get_scales("fill")$palette(1)
  
  model_plot <- model_plot + 
    labs(x="Gestational Age (wks)",
         title = glue("Mean Thickness"),
                  fill = "Age") +
    theme(axis.title.y = element_blank()) +
  scale_x_continuous(labels = function(x) round((x * sd_ga + mu_ga)/7, 1)) +
  scale_fill_manual(labels = c(paste0("Low: ",
                                      round(low_age_label,1),"yrs"),
                               paste0("Mid: ",
                                      round(mid_age_label,1),"yrs"),
                               paste0("High: ",
                                      round(high_age_label,1),"yrs")), 
                    values = plot_colors) +
   scale_color_gradientn(colors = plot_colors,
    labels = function(x) round((x * sd_age + mu_age)/365.25, 1),
    name = "") +
  scale_y_continuous(labels = function(y) glue("{signif(pnorm(y)*100, 2)}th"))


```

## Grid for GA and BWP global models
```{r}
plots <- list(ga_sp_plots_global$global_eTIV, bwp_sp_plots_global$global_eTIV,
              ga_sp_plots_global$global_VentricleChoroidVol, bwp_sp_plots_global$global_VentricleChoroidVol,
             ga_sp_plots_global$global_CortexVol, bwp_sp_plots_global$global_CortexVol)

ncol <- 2
n <- length(plots)
nrow <- ceiling(n / ncol)

row_i <- ceiling(seq_len(n) / ncol)
col_i <- ((seq_len(n) - 1) %% ncol) + 1

plots <- lapply(plots, function(p){
  p +
    theme(
    axis.title = element_blank(),
    axis.text  = element_blank(),
    axis.line = element_blank(),
    axis.line.x = element_blank(),
    axis.ticks = element_blank(),
    aspect.ratio = 1
  )
})

bottom_idx <- which(row_i == nrow)

plots[bottom_idx] <- lapply(plots[bottom_idx], function(p){
  p + 
    theme(
    axis.text.x = element_text(size = 20),
    axis.ticks.x = element_line(),
    axis.title.x = element_text(size = 20),
    axis.line.x = element_line()
  )
})

left_idx <- which(col_i == 1)

plots[left_idx] <- lapply(plots[left_idx], function(p){
  p + theme(
    axis.text.y = element_text(size = 22),
    axis.ticks.y = element_line(),
    axis.line.x = element_line()
  )
})

ga_bwp_global_models_grid <- wrap_plots(plots, ncol = ncol) &
  plot_layout(guides = "collect")

ga_bwp_global_models_grid

# if(save_figs){
#   panel_save(ga_bwp_global_models_grid, "ga_bwp_global_models_grid", "svg", plot_output_folder,
#              plot_width = 10, plot_height = 24)
# }
```

## Get derivative visualizations for GA and BWP global models
```{r}
ga_etiv_d <- derivatives(ga_sp_models_global$global_eTIV$full)
(ga_etiv_derivative <- draw(ga_etiv_d) + 
  geom_line(color = ga_color, linewidth = 3) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "darkgrey") +
  geom_ribbon(aes(ymin = .lower_ci, ymax = .upper_ci), fill = "grey80", alpha = 0.2) +
  theme_classic(base_size = 15) +
  scale_x_continuous(
      breaks = map2_dbl(c(24*7, 28*7, 32*7, 36*7, 40*7), "gestational_age", ~ (.x - mean_lookup_distinct[[.y]])/(sd_lookup_distinct[[.y]])),
      labels = \(x) map2_dbl(x, "gestational_age", ~ (.x*sd_lookup_distinct[[.y]]) + mean_lookup_distinct[[.y]])/7) +
    labs(x= "Gestational Age (weeks)",
         y = "dY/dX (z)",
         title = "") +
    theme(aspect.ratio = 0.75)
)

bwp_etiv_d <- derivatives(bwp_sp_models_global$global_eTIV$full)
(bwp_etiv_derivative <- draw(bwp_etiv_d) + 
  geom_line(color = bw_color, linewidth = 3) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "darkgrey") +
  geom_ribbon(aes(ymin = .lower_ci, ymax = .upper_ci), fill = "grey80", alpha = 0.2) +
  theme_classic(base_size = 15) +
   scale_x_continuous(
     breaks = qnorm(c(.01, .50, .99)),
      labels = \(x) glue("{round(pnorm(x),2)*100}") ) +
    labs(x = glue("Birth Weight \n (GA-corrected percentile)"),
         y = "dY/dX (z)",
         title = "") +
    theme(aspect.ratio = 0.75)
)

ga_ctx_d <- derivatives(ga_sp_models_global$global_CortexVol$full)
(ga_ctx_derivative <- draw(ga_ctx_d) + 
  geom_line(color = ga_color, linewidth = 3) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "darkgrey") +
  geom_ribbon(aes(ymin = .lower_ci, ymax = .upper_ci), fill = "grey80", alpha = 0.2) +
  theme_classic(base_size = 15) +
  scale_x_continuous(
      breaks = map2_dbl(c(24*7, 28*7, 32*7, 36*7, 40*7), "gestational_age", ~ (.x - mean_lookup_distinct[[.y]])/(sd_lookup_distinct[[.y]])),
      labels = \(x) map2_dbl(x, "gestational_age", ~ (.x*sd_lookup_distinct[[.y]]) + mean_lookup_distinct[[.y]])/7) +
    labs(x= "Gestational Age (weeks)",
         y = "dY/dX (z)",
         title = "")+
    theme(aspect.ratio = 0.75)
)

bwp_ctx_d <- derivatives(bwp_sp_models_global$global_CortexVol$full)
(bwp_ctx_derivative <- draw(bwp_ctx_d) + 
  geom_line(color = bw_color, linewidth = 3) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "darkgrey") +
  geom_ribbon(aes(ymin = .lower_ci, ymax = .upper_ci), fill = "grey80", alpha = 0.2) +
  theme_classic(base_size = 15) +
   scale_x_continuous(
     breaks = qnorm(c(.01, .50, .99)),
      labels = \(x) glue("{round(pnorm(x),2)*100}") ) +
    labs(x = glue("Birth Weight \n (GA-corrected percentile)"),
         y = "dY/dX (z)",
         title = "") +
    theme(aspect.ratio = 0.75)
)

ga_subctx_d <- derivatives(ga_sp_models_global$global_SubCortGrayVol$full)
(ga_subctx_derivative <- draw(ga_subctx_d) + 
  geom_line(color = ga_color, linewidth = 3) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "darkgrey") +
  geom_ribbon(aes(ymin = .lower_ci, ymax = .upper_ci), fill = "grey80", alpha = 0.2) +
  theme_classic(base_size = 15) +
  scale_x_continuous(
      breaks = map2_dbl(c(24*7, 28*7, 32*7, 36*7, 40*7), "gestational_age", ~ (.x - mean_lookup_distinct[[.y]])/(sd_lookup_distinct[[.y]])),
      labels = \(x) map2_dbl(x, "gestational_age", ~ (.x*sd_lookup_distinct[[.y]]) + mean_lookup_distinct[[.y]])/7) +
    labs(x= "Gestational Age (weeks)",
         y = "dY/dX (z)",
         title = "")+
    theme(aspect.ratio = 0.75)
)

bwp_subctx_d <- derivatives(bwp_sp_models_global$global_SubCortGrayVo$full)
(bwp_subctx_derivative <- draw(bwp_subctx_d) + 
  geom_line(color = bw_color, linewidth = 3) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "darkgrey") +
  geom_ribbon(aes(ymin = .lower_ci, ymax = .upper_ci), fill = "grey80", alpha = 0.2) +
  theme_classic(base_size = 15) +
   scale_x_continuous(
     breaks = qnorm(c(.01, .50, .99)),
      labels = \(x) glue("{round(pnorm(x),2)*100}") ) +
    labs(x = glue("Birth Weight \n (GA-corrected percentile)"),
         y = "dY/dX (z)",
         title = "") +
    theme(aspect.ratio = 0.75)
)

ga_vent_d <- derivatives(ga_sp_models_global$global_VentricleChoroidVol$full)
(ga_vent_derivative <- draw(ga_vent_d) + 
  geom_line(color = ga_color, linewidth = 3) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "darkgrey") +
  geom_ribbon(aes(ymin = .lower_ci, ymax = .upper_ci), fill = "grey80", alpha = 0.2) +
  theme_classic(base_size = 15) +
  scale_x_continuous(
      breaks = map2_dbl(c(24*7, 28*7, 32*7, 36*7, 40*7), "gestational_age", ~ (.x - mean_lookup_distinct[[.y]])/(sd_lookup_distinct[[.y]])),
      labels = \(x) map2_dbl(x, "gestational_age", ~ (.x*sd_lookup_distinct[[.y]]) + mean_lookup_distinct[[.y]])/7) +
    labs(x= "Gestational Age (weeks)",
         y = "dY/dX (z)",
         title = "")+
    theme(aspect.ratio = 0.75)
)

bwp_vent_d <- derivatives(bwp_sp_models_global$global_VentricleChoroidVol$full)
(bwp_vent_derivative <- draw(bwp_vent_d) + 
  geom_line(color = bw_color, linewidth = 3) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "darkgrey") +
  geom_ribbon(aes(ymin = .lower_ci, ymax = .upper_ci), fill = "grey80", alpha = 0.2) +
  theme_classic(base_size = 15) +
   scale_x_continuous(
     breaks = qnorm(c(.01, .50, .99)),
      labels = \(x) glue("{round(pnorm(x),2)*100}") ) +
    labs(x = glue("Birth Weight \n (GA-corrected percentile)"),
         y = "dY/dX (z)",
         title = "") +
    theme(aspect.ratio = 0.75)
)
```

## Make figure with plots and derivatives
```{r}
plots_deriv <- list(ga_etiv_derivative, bwp_etiv_derivative,
                    ga_vent_derivative, bwp_vent_derivative,
                    ga_ctx_derivative, bwp_ctx_derivative
                    )

plots_deriv <- lapply(plots_deriv, function(p){
  p + theme_classic(base_size = 20) +
    theme(
    axis.title = element_text(size = 15),
    axis.text  = element_text(size = 12),
    aspect.ratio = 0.7
  )
})

ga_bwp_global_models_deriv <- wrap_plots(plots_deriv,  ncol = 1)

ga_bwp_global_models_fig <- wrap_plots(
  ga_bwp_global_models_grid,
  ga_bwp_global_models_deriv,
  ncol = 2,
  widths = c(2, 1)
)

ga_bwp_global_models_fig

if(save_figs){
  panel_save(ga_bwp_global_models_fig, "ga_bwp_global_models_fig", "svg", plot_output_folder,
             plot_width = 20, plot_height = 22)
}

```
## Grid for GA and BWP thickness models
```{r}
shared_th_sig <- intersect(names(ga_sp_plots_th), names(bwp_sp_plots_th))
ga_sp_table_th %>% filter(rownames(.) %in% shared_th_sig) %>% select(r2_full_vs_none) %>% arrange(desc(r2_full_vs_none))
bwp_sp_table_th %>% filter(rownames(.) %in% shared_th_sig) %>% select(r2_full_vs_none) %>% arrange(desc(r2_full_vs_none))

plots <- list(ga_sp_plots_global$global_CortexVol, bwp_sp_plots_global$global_CortexVol,
             ga_sp_plots_global$global_CerebralWhiteMatterVol, bwp_sp_plots_global$global_CerebralWhiteMatterVol,
             ga_sp_plots_global$global_SubCortGrayVol, bwp_sp_plots_global$global_SubCortGrayVol,
             ga_sp_plots_global$global_VentricleChoroidVol, bwp_sp_plots_global$global_VentricleChoroidVol,
             ga_sp_plots_global$global_SurfaceArea, bwp_sp_plots_global$global_SurfaceArea)

ncol <- 2
n <- length(plots)
nrow <- ceiling(n / ncol)

row_i <- ceiling(seq_len(n) / ncol)
col_i <- ((seq_len(n) - 1) %% ncol) + 1

plots <- lapply(plots, function(p){
  p + theme(
    axis.title = element_blank(),
    axis.text  = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    aspect.ratio = 1
  )
})

bottom_idx <- which(row_i == nrow)

plots[bottom_idx] <- lapply(plots[bottom_idx], function(p){
  p + theme(
    axis.text.x = element_text(),
    axis.ticks.x = element_line(),
    axis.title.x = element_text(),
    axis.line.x = element_line()
  )
})

left_idx <- which(col_i == 1)

plots[left_idx] <- lapply(plots[left_idx], function(p){
  p + theme(
    axis.text.y = element_text(),
    axis.ticks.y = element_line(),
    axis.line.x = element_line()
  )
})

ga_bwp_global_models_grid <- wrap_plots(plots, ncol = ncol) &
  plot_layout(guides = "collect")

ga_bwp_global_models_grid

if(save_figs){
  panel_save(ga_bwp_global_models_grid, "ga_bwp_global_models_grid", "svg", plot_output_folder,
             plot_width = 10, plot_height = 24)
}
```

# Regional phenotypes (non-linear models, with interaction term test)

## Volume

### GA

#### fit models
```{r, echo = FALSE}
gabyage_sp_models_vol <- lapply(vol_vars, function(x) {
  form <- as.formula(glue("{x} ~ s(gestational_age) + s(age_days_adj) + ti(gestational_age, age_days_adj) + qc_var"))
  gam_model <- gamm(as.formula(form),
              data = df.zscore)
  return(gam_model)
  })

gabyage_sp_models_vol <- setNames(gabyage_sp_models_vol, vol_vars)

gabyage_sp_gam_table_vol <- as.data.frame(t(sapply(gabyage_sp_models_vol,function(x) {
  summary <- summary(x$gam)
  c(ga_edf = summary$s.table[1,1], ga_p = summary$s.table[1,4],
    age_edf = summary$s.table[2,1], age_p = summary$s.table[2,4],
    gabyage_edf = summary$s.table[3,1], gabyage_p = summary$s.table[3,4])
  })))

rownames(gabyage_sp_gam_table_vol) <- vol_vars

gabyage_sp_gam_table_vol$gabyage_p_adj <- p.adjust(gabyage_sp_gam_table_vol$gabyage_p, method = "BH")

print(paste0("Number of significant IDPs for GA x Age- NO correction: ", dim(gabyage_sp_gam_table_vol %>% filter(gabyage_p < 0.05))[1]))
print(paste0("Number of significant IDPs for GA x Age- AFTER FDR correction: ", dim(gabyage_sp_gam_table_vol %>% filter(gabyage_p_adj < 0.05))[1]))
print(paste0("Significant IDPs for GA x Age- AFTER FDR correction: ", rownames(gabyage_sp_gam_table_vol %>% filter(gabyage_p_adj < 0.05)) ))

gabyage_sp_vol_sig_pheno <- rownames(gabyage_sp_gam_table_vol %>% filter(gabyage_p_adj < 0.05))
```

#### Visualize models
```{r}
gabyage_sp_vol_plot_pheno <- rownames(gabyage_sp_gam_table_vol %>% filter(ga_edf > 2))
font_size <- 16
theme_set(theme_classic(base_family = "sans",base_size = font_size))
line_size <- 1.5
point_size <- 2

mu_age <- mean(cent_all$age_days_adj, na.rm = TRUE)
sd_age <- sd(cent_all$age_days_adj, na.rm = TRUE)

mu_ga <- mean(cent_all$gestational_age, na.rm = TRUE)
sd_ga <- sd(cent_all$gestational_age, na.rm = TRUE)

gabyage_sp_plots_vol <- lapply(gabyage_sp_models_vol[gabyage_sp_vol_plot_pheno], function(x) {
  
  
  low_age_label <- (quantile(df.zscore[!is.na(df.zscore$gestational_age),
                                       "age_days_adj"],
                             probs = c(.10)) * sd_age + mu_age)/365.25
  
  mid_age_label <- (quantile(df.zscore[!is.na(df.zscore$gestational_age),
                                       "age_days_adj"],
                             probs = c(.50)) * sd_age + mu_age)/365.25

  high_age_label <- (quantile(df.zscore[!is.na(df.zscore$gestational_age),
                                      "age_days_adj"],
                            probs = c(.90)) * sd_age + mu_age)/365.25
  
  model_plot <- visualize_model(
  modobj = x,
  smooth_var = "gestational_age",
  int_var = "age_days_adj",
  plabels = "continuous interaction",
  derivative_plot = F)

  plot_colors <- ggplot_build(model_plot)$plot$scales$get_scales("fill")$palette(1)
  
  model_plot <- model_plot + 
    labs(x="Gestational Age (wks)",
         title = glue("{x$gam$terms[[2]]}"),
                  fill = "Age") +
    theme(axis.title.y = element_blank()) +
  scale_x_continuous(labels = function(x) round(x * sd_ga + mu_ga, 1)) +
  scale_fill_manual(labels = c(paste0("Low: ",
                                      round(low_age_label,1),"yrs"),
                               paste0("Mid: ",
                                      round(mid_age_label,1),"yrs"),
                               paste0("High: ",
                                      round(high_age_label,1),"yrs")), 
                    values = plot_colors) +
   scale_color_gradientn(colors = plot_colors,
    labels = function(x) round((x * sd_age + mu_age)/365.25, 1),
    name = "") +
  scale_y_continuous(labels = function(y) signif(pnorm(y)*100, 3))
  
  return(model_plot)
})

gabyage_sp_plots_vol <- setNames(gabyage_sp_plots_vol, gabyage_sp_vol_plot_pheno)

gabyage_sp_plots_vol_nopt <- lapply(gabyage_sp_plots_vol, function(x){
x$layers[1] <- NULL
x})

# wrap_plots(gabyage_sp_plots_vol_nopt, ncol = 2, guides = "collect") &
#   theme(legend.position = "right")
# 
# wrap_plots(gabyage_sp_plots_vol, ncol = 2, guides = "collect") &
#   theme(legend.position = "right")

wrapped_plot_n <- 4 

plot_batches <- split(gabyage_sp_plots_vol, ceiling(seq_along(gabyage_sp_plots_vol) / wrapped_plot_n))

 wrapped_list <- lapply(plot_batches, function(batch) {
   wrap_plots(batch, ncol = 2)
 })

wrapped_list

plot_batches_nopt <- split(gabyage_sp_plots_vol_nopt, ceiling(seq_along(gabyage_sp_plots_vol_nopt) / wrapped_plot_n))

 wrapped_list_nopt <- lapply(plot_batches_nopt, function(batch) {
   wrap_plots(batch, ncol = 2)
 })

wrapped_list_nopt
```

### BW

#### fit models
```{r, echo = FALSE}
bwbyage_sp_models_vol <- lapply(vol_vars, function(x) {
  form <- as.formula(glue("{x} ~ s(birth_weight_kg) + s(age_days_adj) + ti(birth_weight_kg, age_days_adj) + qc_var"))
  gam_model <- gamm(as.formula(form),
              data = df.zscore)
  return(gam_model)
  })

bwbyage_sp_models_vol <- setNames(bwbyage_sp_models_vol, vol_vars)

bwbyage_sp_gam_table_vol <- as.data.frame(t(sapply(bwbyage_sp_models_vol,function(x) {
  summary <- summary(x$gam)
  c(bw_edf = summary$s.table[1,1], bw_p = summary$s.table[1,4],
    age_edf = summary$s.table[2,1], age_p = summary$s.table[2,4],
    bwbyage_edf = summary$s.table[3,1], bwbyage_p = summary$s.table[3,4])
  })))

rownames(bwbyage_sp_gam_table_vol) <- vol_vars

bwbyage_sp_gam_table_vol$bwbyage_p_adj <- p.adjust(bwbyage_sp_gam_table_vol$bwbyage_p, method = "BH")

print(paste0("Number of significant IDPs for bw x Age- NO correction: ", dim(bwbyage_sp_gam_table_vol %>% filter(bwbyage_p < 0.05))[1]))
print(paste0("Number of significant IDPs for bw x Age- AFTER FDR correction: ", dim(bwbyage_sp_gam_table_vol %>% filter(bwbyage_p_adj < 0.05))[1]))
print(paste0("Significant IDPs for bw x Age- AFTER FDR correction: ", rownames(bwbyage_sp_gam_table_vol %>% filter(bwbyage_p_adj < 0.05)) ))

bwbyage_sp_vol_sig_pheno <- rownames(bwbyage_sp_gam_table_vol %>% filter(bwbyage_p_adj < 0.05))
```

#### Visualize models
```{r}
bwbyage_sp_vol_plot_pheno <- rownames(bwbyage_sp_gam_table_vol %>% filter(bw_edf > 2))
font_size <- 16
theme_set(theme_classic(base_family = "sans",base_size = font_size))
line_size <- 1.5
point_size <- 2
source("/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/visualize_model_bartfct.R")

mu_age <- mean(cent_all$age_days_adj, na.rm = TRUE)
sd_age <- sd(cent_all$age_days_adj, na.rm = TRUE)

mu_bw <- mean(cent_all$birth_weight_kg, na.rm = TRUE)
sd_bw <- sd(cent_all$birth_weight_kg, na.rm = TRUE)

bwbyage_sp_plots_vol <- lapply(bwbyage_sp_models_vol[bwbyage_sp_vol_plot_pheno], function(x) {
  
  
  low_age_label <- (quantile(df.zscore[!is.na(df.zscore$birth_weight_kg),
                                       "age_days_adj"],
                             probs = c(.10)) * sd_age + mu_age)/365.25
  
  mid_age_label <- (quantile(df.zscore[!is.na(df.zscore$birth_weight_kg),
                                       "age_days_adj"],
                             probs = c(.50)) * sd_age + mu_age)/365.25

  high_age_label <- (quantile(df.zscore[!is.na(df.zscore$birth_weight_kg),
                                      "age_days_adj"],
                            probs = c(.90)) * sd_age + mu_age)/365.25
  
  model_plot <- visualize_model(
  modobj = x,
  smooth_var = "birth_weight_kg",
  int_var = "age_days_adj",
  plabels = "continuous interaction",
  derivative_plot = F)

  plot_colors <- ggplot_build(model_plot)$plot$scales$get_scales("fill")$palette(1)
  
  model_plot <- model_plot + 
    labs(x="Birth Weight (kg)",
         title = glue("{x$gam$terms[[2]]}"),
                  fill = "Age") +
    theme(axis.title.y = element_blank()) +
  scale_x_continuous(labels = function(x) round(x * sd_bw + mu_bw, 1)) +
  scale_fill_manual(labels = c(paste0("Low: ",
                                      round(low_age_label,1),"yrs"),
                               paste0("Mid: ",
                                      round(mid_age_label,1),"yrs"),
                               paste0("High: ",
                                      round(high_age_label,1),"yrs")), 
                    values = plot_colors) +
   scale_color_gradientn(colors = plot_colors,
    labels = function(x) round((x * sd_age + mu_age)/365.25, 1),
    name = "") +
  scale_y_continuous(labels = function(y) signif(pnorm(y)*100, 3))
  
  return(model_plot)
})

bwbyage_sp_plots_vol <- setNames(bwbyage_sp_plots_vol, bwbyage_sp_vol_plot_pheno)

bwbyage_sp_plots_vol_nopt <- lapply(bwbyage_sp_plots_vol, function(x){
x$layers[1] <- NULL
x})

wrapped_plot_n <- 4 

plot_batches <- split(bwbyage_sp_plots_vol, ceiling(seq_along(bwbyage_sp_plots_vol) / wrapped_plot_n))

 wrapped_list <- lapply(plot_batches, function(batch) {
   wrap_plots(batch, ncol = 2)
 })

wrapped_list

plot_batches_nopt <- split(bwbyage_sp_plots_vol_nopt, ceiling(seq_along(bwbyage_sp_plots_vol_nopt) / wrapped_plot_n))

 wrapped_list_nopt <- lapply(plot_batches_nopt, function(batch) {
   wrap_plots(batch, ncol = 2)
 })

wrapped_list_nopt
```


## Surface Area
### GA

#### fit models
```{r, echo = FALSE}
gabyage_sp_models_sa <- lapply(sa_vars, function(x) {
  form <- as.formula(glue("{x} ~ s(gestational_age) + s(age_days_adj) + ti(gestational_age, age_days_adj) + qc_var"))
  gam_model <- gamm(as.formula(form),
              data = df.zscore)
  return(gam_model)
  })

gabyage_sp_models_sa <- setNames(gabyage_sp_models_sa, sa_vars)

gabyage_sp_gam_table_sa <- as.data.frame(t(sapply(gabyage_sp_models_sa,function(x) {
  summary <- summary(x$gam)
  c(ga_edf = summary$s.table[1,1], ga_p = summary$s.table[1,4],
    age_edf = summary$s.table[2,1], age_p = summary$s.table[2,4],
    gabyage_edf = summary$s.table[3,1], gabyage_p = summary$s.table[3,4])
  })))

rownames(gabyage_sp_gam_table_sa) <- sa_vars

gabyage_sp_gam_table_sa$gabyage_p_adj <- p.adjust(gabyage_sp_gam_table_sa$gabyage_p, method = "BH")

print(paste0("Number of significant IDPs for GA x Age- NO correction: ", dim(gabyage_sp_gam_table_sa %>% filter(gabyage_p < 0.05))[1]))
print(paste0("Number of significant IDPs for GA x Age- AFTER FDR correction: ", dim(gabyage_sp_gam_table_sa %>% filter(gabyage_p_adj < 0.05))[1]))
print(paste0("Significant IDPs for GA x Age- AFTER FDR correction: ", rownames(gabyage_sp_gam_table_sa %>% filter(gabyage_p_adj < 0.05)) ))

gabyage_sp_sa_sig_pheno <- rownames(gabyage_sp_gam_table_sa %>% filter(gabyage_p_adj < 0.05))
```

#### Visualize models
```{r}
gabyage_sp_sa_plot_pheno <- rownames(gabyage_sp_gam_table_sa %>% filter(ga_edf > 2))
font_size <- 16
theme_set(theme_classic(base_family = "sans",base_size = font_size))
line_size <- 1.5
point_size <- 2
source("/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/visualize_model_bartfct.R")

mu_age <- mean(cent_all$age_days_adj, na.rm = TRUE)
sd_age <- sd(cent_all$age_days_adj, na.rm = TRUE)

mu_ga <- mean(cent_all$gestational_age, na.rm = TRUE)
sd_ga <- sd(cent_all$gestational_age, na.rm = TRUE)

gabyage_sp_plots_sa <- lapply(gabyage_sp_models_sa[gabyage_sp_sa_plot_pheno], function(x) {
  
  
  low_age_label <- (quantile(df.zscore[!is.na(df.zscore$gestational_age),
                                       "age_days_adj"],
                             probs = c(.10)) * sd_age + mu_age)/365.25
  
  mid_age_label <- (quantile(df.zscore[!is.na(df.zscore$gestational_age),
                                       "age_days_adj"],
                             probs = c(.50)) * sd_age + mu_age)/365.25

  high_age_label <- (quantile(df.zscore[!is.na(df.zscore$gestational_age),
                                      "age_days_adj"],
                            probs = c(.90)) * sd_age + mu_age)/365.25
  
  model_plot <- visualize_model(
  modobj = x,
  smooth_var = "gestational_age",
  int_var = "age_days_adj",
  plabels = "continuous interaction",
  derivative_plot = F)

  plot_colors <- ggplot_build(model_plot)$plot$scales$get_scales("fill")$palette(1)
  
  model_plot <- model_plot + 
    labs(x="Gestational Age (wks)",
         title = glue("{x$gam$terms[[2]]}"),
                  fill = "Age") +
    theme(axis.title.y = element_blank()) +
  scale_x_continuous(labels = function(x) round(x * sd_ga + mu_ga, 1)) +
  scale_fill_manual(labels = c(paste0("Low: ",
                                      round(low_age_label,1),"yrs"),
                               paste0("Mid: ",
                                      round(mid_age_label,1),"yrs"),
                               paste0("High: ",
                                      round(high_age_label,1),"yrs")), 
                    values = plot_colors) +
   scale_color_gradientn(colors = plot_colors,
    labels = function(x) round((x * sd_age + mu_age)/365.25, 1),
    name = "") +
  scale_y_continuous(labels = function(y) signif(pnorm(y)*100, 3))
  
  return(model_plot)
})

gabyage_sp_plots_sa <- setNames(gabyage_sp_plots_sa, gabyage_sp_sa_plot_pheno)

gabyage_sp_plots_sa_nopt <- lapply(gabyage_sp_plots_sa, function(x){
x$layers[1] <- NULL
x})

wrapped_plot_n <- 4 

plot_batches <- split(gabyage_sp_plots_sa, ceiling(seq_along(gabyage_sp_plots_sa) / wrapped_plot_n))

 wrapped_list <- lapply(plot_batches, function(batch) {
   wrap_plots(batch, ncol = 2)
 })

wrapped_list

plot_batches_nopt <- split(gabyage_sp_plots_sa_nopt, ceiling(seq_along(gabyage_sp_plots_sa_nopt) / wrapped_plot_n))

 wrapped_list_nopt <- lapply(plot_batches_nopt, function(batch) {
   wrap_plots(batch, ncol = 2)
 })

wrapped_list_nopt

#wrapped_list
```

### BW

#### fit models
```{r, echo = FALSE}
bwbyage_sp_models_sa <- lapply(sa_vars, function(x) {
  form <- as.formula(glue("{x} ~ s(birth_weight_kg) + s(age_days_adj) + ti(birth_weight_kg, age_days_adj) + qc_var"))
  gam_model <- gamm(as.formula(form),
              data = df.zscore)
  return(gam_model)
  })

bwbyage_sp_models_sa <- setNames(bwbyage_sp_models_sa, sa_vars)

bwbyage_sp_gam_table_sa <- as.data.frame(t(sapply(bwbyage_sp_models_sa,function(x) {
  summary <- summary(x$gam)
  c(bw_edf = summary$s.table[1,1], bw_p = summary$s.table[1,4],
    age_edf = summary$s.table[2,1], age_p = summary$s.table[2,4],
    bwbyage_edf = summary$s.table[3,1], bwbyage_p = summary$s.table[3,4])
  })))

rownames(bwbyage_sp_gam_table_sa) <- sa_vars

bwbyage_sp_gam_table_sa$bwbyage_p_adj <- p.adjust(bwbyage_sp_gam_table_sa$bwbyage_p, method = "BH")

print(paste0("Number of significant IDPs for bw x Age- NO correction: ", dim(bwbyage_sp_gam_table_sa %>% filter(bwbyage_p < 0.05))[1]))
print(paste0("Number of significant IDPs for bw x Age- AFTER FDR correction: ", dim(bwbyage_sp_gam_table_sa %>% filter(bwbyage_p_adj < 0.05))[1]))
print(paste0("Significant IDPs for bw x Age- AFTER FDR correction: ", rownames(bwbyage_sp_gam_table_sa %>% filter(bwbyage_p_adj < 0.05)) ))

bwbyage_sp_sa_sig_pheno <- rownames(bwbyage_sp_gam_table_sa %>% filter(bwbyage_p_adj < 0.05))
```

#### Visualize models
```{r}
bwbyage_sp_sa_plot_pheno <- rownames(bwbyage_sp_gam_table_sa %>% filter(bw_edf > 2))
font_size <- 16
theme_set(theme_classic(base_family = "sans",base_size = font_size))
line_size <- 1.5
point_size <- 2
source("/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/visualize_model_bartfct.R")

mu_age <- mean(cent_all$age_days_adj, na.rm = TRUE)
sd_age <- sd(cent_all$age_days_adj, na.rm = TRUE)

mu_bw <- mean(cent_all$birth_weight_kg, na.rm = TRUE)
sd_bw <- sd(cent_all$birth_weight_kg, na.rm = TRUE)

bwbyage_sp_plots_sa <- lapply(bwbyage_sp_models_sa[bwbyage_sp_sa_plot_pheno], function(x) {
  
  
  low_age_label <- (quantile(df.zscore[!is.na(df.zscore$birth_weight_kg),
                                       "age_days_adj"],
                             probs = c(.10)) * sd_age + mu_age)/365.25
  
  mid_age_label <- (quantile(df.zscore[!is.na(df.zscore$birth_weight_kg),
                                       "age_days_adj"],
                             probs = c(.50)) * sd_age + mu_age)/365.25

  high_age_label <- (quantile(df.zscore[!is.na(df.zscore$birth_weight_kg),
                                      "age_days_adj"],
                            probs = c(.90)) * sd_age + mu_age)/365.25
  
  model_plot <- visualize_model(
  modobj = x,
  smooth_var = "birth_weight_kg",
  int_var = "age_days_adj",
  plabels = "continuous interaction",
  derivative_plot = F)

  plot_colors <- ggplot_build(model_plot)$plot$scales$get_scales("fill")$palette(1)
  
  model_plot <- model_plot + 
    labs(x="Birth Weight (kg)",
         title = glue("{x$gam$terms[[2]]}"),
                  fill = "Age") +
    theme(axis.title.y = element_blank()) +
  scale_x_continuous(labels = function(x) round(x * sd_bw + mu_bw, 1)) +
  scale_fill_manual(labels = c(paste0("Low: ",
                                      round(low_age_label,1),"yrs"),
                               paste0("Mid: ",
                                      round(mid_age_label,1),"yrs"),
                               paste0("High: ",
                                      round(high_age_label,1),"yrs")), 
                    values = plot_colors) +
   scale_color_gradientn(colors = plot_colors,
    labels = function(x) round((x * sd_age + mu_age)/365.25, 1),
    name = "") +
  scale_y_continuous(labels = function(y) signif(pnorm(y)*100, 3))
  
  return(model_plot)
})

bwbyage_sp_plots_sa <- setNames(bwbyage_sp_plots_sa, bwbyage_sp_sa_plot_pheno)

bwbyage_sp_plots_sa_nopt <- lapply(bwbyage_sp_plots_sa, function(x){
x$layers[1] <- NULL
x})

wrapped_plot_n <- 4 

plot_batches <- split(bwbyage_sp_plots_sa, ceiling(seq_along(bwbyage_sp_plots_sa) / wrapped_plot_n))

 wrapped_list <- lapply(plot_batches, function(batch) {
   wrap_plots(batch, ncol = 2)
 })

wrapped_list

plot_batches_nopt <- split(bwbyage_sp_plots_sa_nopt, ceiling(seq_along(bwbyage_sp_plots_sa_nopt) / wrapped_plot_n))

 wrapped_list_nopt <- lapply(plot_batches_nopt, function(batch) {
   wrap_plots(batch, ncol = 2)
 })

wrapped_list_nopt
```


## Thickness median full sample
### GA

#### fit models
```{r, echo = FALSE}
if (median_th_mpr_flag == TRUE){
gabyage_sp_models_th_median <- lapply(th_vars_median, function(x) {
  form <- as.formula(glue("{x} ~ s(gestational_age) + s(age_days_adj) + ti(gestational_age, age_days_adj) + qc_var"))
  gam_model <- gamm(as.formula(form),
              data = df.zscore)
  return(gam_model)
  })

gabyage_sp_models_th_median <- setNames(gabyage_sp_models_th_median, th_vars)

gabyage_sp_gam_table_th_median <- as.data.frame(t(sapply(gabyage_sp_models_th_median,function(x) {
  summary <- summary(x$gam)
  c(ga_edf = summary$s.table[1,1], ga_p = summary$s.table[1,4],
    age_edf = summary$s.table[2,1], age_p = summary$s.table[2,4],
    gabyage_edf = summary$s.table[3,1], gabyage_p = summary$s.table[3,4])
  })))

rownames(gabyage_sp_gam_table_th_median) <- th_vars

gabyage_sp_gam_table_th_median$gabyage_p_adj <- p.adjust(gabyage_sp_gam_table_th_median$gabyage_p, method = "BH")

print(paste0("Number of significant IDPs for GA x Age- NO correction: ", dim(gabyage_sp_gam_table_th_median %>% filter(gabyage_p < 0.05))[1]))
print(paste0("Number of significant IDPs for GA x Age- AFTER FDR correction: ", dim(gabyage_sp_gam_table_th_median %>% filter(gabyage_p_adj < 0.05))[1]))
print(paste0("Significant IDPs for GA x Age- AFTER FDR correction: ", rownames(gabyage_sp_gam_table_th_median %>% filter(gabyage_p_adj < 0.05)) ))

gabyage_sp_th_median_sig_pheno <- rownames(gabyage_sp_gam_table_th_median %>% filter(gabyage_p_adj < 0.05))
}
```

#### Visualize models
```{r}
if (median_th_mpr_flag == TRUE){
gabyage_sp_th_median_plot_pheno <- rownames(gabyage_sp_gam_table_th_median %>% filter(ga_edf > 2))
font_size <- 16
theme_set(theme_classic(base_family = "sans",base_size = font_size))
line_size <- 1.5
point_size <- 2
source("/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/visualize_model_bartfct.R")

mu_age <- mean(cent_all$age_days_adj, na.rm = TRUE)
sd_age <- sd(cent_all$age_days_adj, na.rm = TRUE)

mu_ga <- mean(cent_all$gestational_age, na.rm = TRUE)
sd_ga <- sd(cent_all$gestational_age, na.rm = TRUE)

gabyage_sp_plots_th_median <- lapply(gabyage_sp_models_th_median[gabyage_sp_th_median_sig_pheno], function(x) {
  
  
  low_age_label <- (quantile(df.zscore[!is.na(df.zscore$gestational_age),
                                       "age_days_adj"],
                             probs = c(.10)) * sd_age + mu_age)/365.25
  
  mid_age_label <- (quantile(df.zscore[!is.na(df.zscore$gestational_age),
                                       "age_days_adj"],
                             probs = c(.50)) * sd_age + mu_age)/365.25

  high_age_label <- (quantile(df.zscore[!is.na(df.zscore$gestational_age),
                                      "age_days_adj"],
                            probs = c(.90)) * sd_age + mu_age)/365.25
  
  model_plot <- visualize_model(
  modobj = x,
  smooth_var = "gestational_age",
  int_var = "age_days_adj",
  plabels = "continuous interaction",
  derivative_plot = F)

  plot_colors <- ggplot_build(model_plot)$plot$scales$get_scales("fill")$palette(1)
  
  model_plot <- model_plot + 
    labs(x="Gestational Age (wks)",
         title = glue("{x$gam$terms[[2]]}"),
                  fill = "Age") +
    theme(axis.title.y = element_blank()) +
  scale_x_continuous(labels = function(x) round(x * sd_ga + mu_ga, 1)) +
  scale_fill_manual(labels = c(paste0("Low: ",
                                      round(low_age_label,1),"yrs"),
                               paste0("Mid: ",
                                      round(mid_age_label,1),"yrs"),
                               paste0("High: ",
                                      round(high_age_label,1),"yrs")), 
                    values = plot_colors) +
   scale_color_gradientn(colors = plot_colors,
    labels = function(x) round((x * sd_age + mu_age)/365.25, 1),
    name = "") +
  scale_y_continuous(labels = function(y) signif(pnorm(y)*100, 3))
  
  return(model_plot)
})

gabyage_sp_plots_th_median <- setNames(gabyage_sp_plots_th_median, gabyage_sp_th_median_sig_pheno)

gabyage_sp_plots_th_median_nopt <- lapply(gabyage_sp_plots_th_median, function(x){
x$layers[1] <- NULL
x})

wrapped_plot_n <- 4 

plot_batches <- split(gabyage_sp_plots_th_median, ceiling(seq_along(gabyage_sp_plots_th_median) / wrapped_plot_n))

 wrapped_list <- lapply(plot_batches, function(batch) {
   wrap_plots(batch, ncol = 2)
 })

wrapped_list

plot_batches_nopt <- split(gabyage_sp_plots_th_median_nopt, ceiling(seq_along(gabyage_sp_plots_th_median_nopt) / wrapped_plot_n))

 wrapped_list_nopt <- lapply(plot_batches_nopt, function(batch) {
   wrap_plots(batch, ncol = 2)
 })

wrapped_list_nopt
}
```

### BW

#### fit models
```{r, echo = FALSE}
if (median_th_mpr_flag == TRUE){
bwbyage_sp_models_th_median <- lapply(th_vars_median, function(x) {
  form <- as.formula(glue("{x} ~ s(birth_weight_kg) + s(age_days_adj) + ti(birth_weight_kg, age_days_adj) + qc_var"))
  gam_model <- gamm(as.formula(form),
              data = df.zscore)
  return(gam_model)
  })

bwbyage_sp_models_th_median <- setNames(bwbyage_sp_models_th_median, th_vars)

bwbyage_sp_gam_table_th_median <- as.data.frame(t(sapply(bwbyage_sp_models_th_median,function(x) {
  summary <- summary(x$gam)
  c(bw_edf = summary$s.table[1,1], bw_p = summary$s.table[1,4],
    age_edf = summary$s.table[2,1], age_p = summary$s.table[2,4],
    bwbyage_edf = summary$s.table[3,1], bwbyage_p = summary$s.table[3,4])
  })))

rownames(bwbyage_sp_gam_table_th_median) <- th_vars

bwbyage_sp_gam_table_th_median$bwbyage_p_adj <- p.adjust(bwbyage_sp_gam_table_th_median$bwbyage_p, method = "BH")

print(paste0("Number of significant IDPs for bw x Age- NO correction: ", dim(bwbyage_sp_gam_table_th_median %>% filter(bwbyage_p < 0.05))[1]))
print(paste0("Number of significant IDPs for bw x Age- AFTER FDR correction: ", dim(bwbyage_sp_gam_table_th_median %>% filter(bwbyage_p_adj < 0.05))[1]))
print(paste0("Significant IDPs for bw x Age- AFTER FDR correction: ", rownames(bwbyage_sp_gam_table_th_median %>% filter(bwbyage_p_adj < 0.05)) ))

bwbyage_sp_th_median_sig_pheno <- rownames(bwbyage_sp_gam_table_th_median %>% filter(bwbyage_p_adj < 0.05))
}
```

#### Visualize models
```{r}
if (median_th_mpr_flag == TRUE){
bwbyage_sp_th_median_plot_pheno <- rownames(bwbyage_sp_gam_table_th_median %>% filter(bw_edf > 2))
font_size <- 16
theme_set(theme_classic(base_family = "sans",base_size = font_size))
line_size <- 1.5
point_size <- 2
source("/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/visualize_model_bartfct.R")

mu_age <- mean(cent_all$age_days_adj, na.rm = TRUE)
sd_age <- sd(cent_all$age_days_adj, na.rm = TRUE)

mu_bw <- mean(cent_all$birth_weight_kg, na.rm = TRUE)
sd_bw <- sd(cent_all$birth_weight_kg, na.rm = TRUE)

bwbyage_sp_plots_th_median <- lapply(bwbyage_sp_models_th_median[bwbyage_sp_th_median_sig_pheno], function(x) {
  
  
  low_age_label <- (quantile(df.zscore[!is.na(df.zscore$birth_weight_kg),
                                       "age_days_adj"],
                             probs = c(.10)) * sd_age + mu_age)/365.25
  
  mid_age_label <- (quantile(df.zscore[!is.na(df.zscore$birth_weight_kg),
                                       "age_days_adj"],
                             probs = c(.50)) * sd_age + mu_age)/365.25

  high_age_label <- (quantile(df.zscore[!is.na(df.zscore$birth_weight_kg),
                                      "age_days_adj"],
                            probs = c(.90)) * sd_age + mu_age)/365.25
  
  model_plot <- visualize_model(
  modobj = x,
  smooth_var = "birth_weight_kg",
  int_var = "age_days_adj",
  plabels = "continuous interaction",
  derivative_plot = F)

  plot_colors <- ggplot_build(model_plot)$plot$scales$get_scales("fill")$palette(1)
  
  model_plot <- model_plot + 
    labs(x="Birth Weight (kg)",
         title = glue("{x$gam$terms[[2]]}"),
                  fill = "Age") +
    theme(axis.title.y = element_blank()) +
  scale_x_continuous(labels = function(x) round(x * sd_bw + mu_bw, 1)) +
  scale_fill_manual(labels = c(paste0("Low: ",
                                      round(low_age_label,1),"yrs"),
                               paste0("Mid: ",
                                      round(mid_age_label,1),"yrs"),
                               paste0("High: ",
                                      round(high_age_label,1),"yrs")), 
                    values = plot_colors) +
   scale_color_gradientn(colors = plot_colors,
    labels = function(x) round((x * sd_age + mu_age)/365.25, 1),
    name = "") +
  scale_y_continuous(labels = function(y) signif(pnorm(y)*100, 3))
  
  return(model_plot)
})

bwbyage_sp_plots_th_median <- setNames(bwbyage_sp_plots_th_median, bwbyage_sp_th_median_sig_pheno)

bwbyage_sp_plots_th_median_nopt <- lapply(bwbyage_sp_plots_th_median, function(x){
x$layers[1] <- NULL
x})

wrapped_plot_n <- 4 

plot_batches <- split(bwbyage_sp_plots_th_median, ceiling(seq_along(bwbyage_sp_plots_th_median) / wrapped_plot_n))

 wrapped_list <- lapply(plot_batches, function(batch) {
   wrap_plots(batch, ncol = 2)
 })

wrapped_list

plot_batches_nopt <- split(bwbyage_sp_plots_th_median_nopt, ceiling(seq_along(bwbyage_sp_plots_th_median_nopt) / wrapped_plot_n))

 wrapped_list_nopt <- lapply(plot_batches_nopt, function(batch) {
   wrap_plots(batch, ncol = 2)
 })

wrapped_list_nopt
}
```


## Thickness (mpr, df.zscore_overOne overOne sample)
### GA

#### fit models
```{r, echo = FALSE}
gabyage_sp_models_th <- lapply(th_vars, function(x) {
  form <- as.formula(glue("{x} ~ s(gestational_age) + s(age_days_adj) + ti(gestational_age, age_days_adj) + euler_mean_mpr"))
  gam_model <- gamm(as.formula(form),
              data = df.zscore_overOne)
  return(gam_model)
  })

gabyage_sp_models_th <- setNames(gabyage_sp_models_th, th_vars)

gabyage_sp_gam_table_th <- as.data.frame(t(sapply(gabyage_sp_models_th,function(x) {
  summary <- summary(x$gam)
  c(ga_edf = summary$s.table[1,1], ga_p = summary$s.table[1,4],
    age_edf = summary$s.table[2,1], age_p = summary$s.table[2,4],
    gabyage_edf = summary$s.table[3,1], gabyage_p = summary$s.table[3,4])
  })))

rownames(gabyage_sp_gam_table_th) <- th_vars

gabyage_sp_gam_table_th$gabyage_p_adj <- p.adjust(gabyage_sp_gam_table_th$gabyage_p, method = "BH")

print(paste0("Number of significant IDPs for GA x Age- NO correction: ", dim(gabyage_sp_gam_table_th %>% filter(gabyage_p < 0.05))[1]))
print(paste0("Number of significant IDPs for GA x Age- AFTER FDR correction: ", dim(gabyage_sp_gam_table_th %>% filter(gabyage_p_adj < 0.05))[1]))
print(paste0("Significant IDPs for GA x Age- AFTER FDR correction: ", rownames(gabyage_sp_gam_table_th %>% filter(gabyage_p_adj < 0.05)) ))

gabyage_sp_th_sig_pheno <- rownames(gabyage_sp_gam_table_th %>% filter(gabyage_p_adj < 0.05))
```

#### Visualize models
```{r}
gabyage_sp_th_plot_pheno <- rownames(gabyage_sp_gam_table_th %>% filter(ga_edf > 2))
font_size <- 16
theme_set(theme_classic(base_family = "sans",base_size = font_size))
line_size <- 1.5
point_size <- 2
source("/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/visualize_model_bartfct.R")

mu_age <- mean(cent_all$age_days_adj, na.rm = TRUE)
sd_age <- sd(cent_all$age_days_adj, na.rm = TRUE)

mu_ga <- mean(cent_all$gestational_age, na.rm = TRUE)
sd_ga <- sd(cent_all$gestational_age, na.rm = TRUE)

gabyage_sp_plots_th <- lapply(gabyage_sp_models_th[gabyage_sp_th_sig_pheno], function(x) {
  
  
  low_age_label <- (quantile(df.zscore_overOne[!is.na(df.zscore_overOne$gestational_age),
                                       "age_days_adj"],
                             probs = c(.10)) * sd_age + mu_age)/365.25
  
  mid_age_label <- (quantile(df.zscore_overOne[!is.na(df.zscore_overOne$gestational_age),
                                       "age_days_adj"],
                             probs = c(.50)) * sd_age + mu_age)/365.25

  high_age_label <- (quantile(df.zscore_overOne[!is.na(df.zscore_overOne$gestational_age),
                                      "age_days_adj"],
                            probs = c(.90)) * sd_age + mu_age)/365.25
  
  model_plot <- visualize_model(
  modobj = x,
  smooth_var = "gestational_age",
  int_var = "age_days_adj",
  plabels = "continuous interaction",
  derivative_plot = F)

  plot_colors <- ggplot_build(model_plot)$plot$scales$get_scales("fill")$palette(1)
  
  model_plot <- model_plot + 
    labs(x="Gestational Age (wks)",
         title = glue("{x$gam$terms[[2]]}"),
                  fill = "Age") +
    theme(axis.title.y = element_blank()) +
  scale_x_continuous(labels = function(x) round(x * sd_ga + mu_ga, 1)) +
  scale_fill_manual(labels = c(paste0("Low: ",
                                      round(low_age_label,1),"yrs"),
                               paste0("Mid: ",
                                      round(mid_age_label,1),"yrs"),
                               paste0("High: ",
                                      round(high_age_label,1),"yrs")), 
                    values = plot_colors) +
   scale_color_gradientn(colors = plot_colors,
    labels = function(x) round((x * sd_age + mu_age)/365.25, 1),
    name = "") +
  scale_y_continuous(labels = function(y) signif(pnorm(y)*100, 3))
  
  return(model_plot)
})

gabyage_sp_plots_th <- setNames(gabyage_sp_plots_th, gabyage_sp_th_sig_pheno)

gabyage_sp_plots_th_nopt <- lapply(gabyage_sp_plots_th, function(x){
x$layers[1] <- NULL
x})

wrapped_plot_n <- 4 

plot_batches <- split(gabyage_sp_plots_th, ceiling(seq_along(gabyage_sp_plots_th) / wrapped_plot_n))

 wrapped_list <- lapply(plot_batches, function(batch) {
   wrap_plots(batch, ncol = 2)
 })

wrapped_list

plot_batches_nopt <- split(gabyage_sp_plots_th_nopt, ceiling(seq_along(gabyage_sp_plots_th_nopt) / wrapped_plot_n))

 wrapped_list_nopt <- lapply(plot_batches_nopt, function(batch) {
   wrap_plots(batch, ncol = 2)
 })

wrapped_list_nopt
```

### BW

#### fit models
```{r, echo = FALSE}
bwbyage_sp_models_th <- lapply(th_vars, function(x) {
  form <- as.formula(glue("{x} ~ s(birth_weight_kg) + s(age_days_adj) + ti(birth_weight_kg, age_days_adj) + euler_mean_mpr"))
  gam_model <- gamm(as.formula(form),
              data = df.zscore_overOne)
  return(gam_model)
  })

bwbyage_sp_models_th <- setNames(bwbyage_sp_models_th, th_vars)

bwbyage_sp_gam_table_th <- as.data.frame(t(sapply(bwbyage_sp_models_th,function(x) {
  summary <- summary(x$gam)
  c(bw_edf = summary$s.table[1,1], bw_p = summary$s.table[1,4],
    age_edf = summary$s.table[2,1], age_p = summary$s.table[2,4],
    bwbyage_edf = summary$s.table[3,1], bwbyage_p = summary$s.table[3,4])
  })))

rownames(bwbyage_sp_gam_table_th) <- th_vars

bwbyage_sp_gam_table_th$bwbyage_p_adj <- p.adjust(bwbyage_sp_gam_table_th$bwbyage_p, method = "BH")

print(paste0("Number of significant IDPs for bw x Age- NO correction: ", dim(bwbyage_sp_gam_table_th %>% filter(bwbyage_p < 0.05))[1]))
print(paste0("Number of significant IDPs for bw x Age- AFTER FDR correction: ", dim(bwbyage_sp_gam_table_th %>% filter(bwbyage_p_adj < 0.05))[1]))
print(paste0("Significant IDPs for bw x Age- AFTER FDR correction: ", rownames(bwbyage_sp_gam_table_th %>% filter(bwbyage_p_adj < 0.05)) ))

bwbyage_sp_th_sig_pheno <- rownames(bwbyage_sp_gam_table_th %>% filter(bwbyage_p_adj < 0.05))
```

#### Visualize models
```{r}
bwbyage_sp_th_plot_pheno <- rownames(bwbyage_sp_gam_table_th %>% filter(bw_edf > 2))
font_size <- 16
theme_set(theme_classic(base_family = "sans",base_size = font_size))
line_size <- 1.5
point_size <- 2
source("/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/visualize_model_bartfct.R")

mu_age <- mean(cent_all$age_days_adj, na.rm = TRUE)
sd_age <- sd(cent_all$age_days_adj, na.rm = TRUE)

mu_bw <- mean(cent_all$birth_weight_kg, na.rm = TRUE)
sd_bw <- sd(cent_all$birth_weight_kg, na.rm = TRUE)

bwbyage_sp_plots_th <- lapply(bwbyage_sp_models_th[bwbyage_sp_th_sig_pheno], function(x) {
  
  
  low_age_label <- (quantile(df.zscore_overOne[!is.na(df.zscore_overOne$birth_weight_kg),
                                       "age_days_adj"],
                             probs = c(.10)) * sd_age + mu_age)/365.25
  
  mid_age_label <- (quantile(df.zscore_overOne[!is.na(df.zscore_overOne$birth_weight_kg),
                                       "age_days_adj"],
                             probs = c(.50)) * sd_age + mu_age)/365.25

  high_age_label <- (quantile(df.zscore_overOne[!is.na(df.zscore_overOne$birth_weight_kg),
                                      "age_days_adj"],
                            probs = c(.90)) * sd_age + mu_age)/365.25
  
  model_plot <- visualize_model(
  modobj = x,
  smooth_var = "birth_weight_kg",
  int_var = "age_days_adj",
  plabels = "continuous interaction",
  derivative_plot = F)

  plot_colors <- ggplot_build(model_plot)$plot$scales$get_scales("fill")$palette(1)
  
  model_plot <- model_plot + 
    labs(x="Birth Weight (kg)",
         title = glue("{x$gam$terms[[2]]}"),
                  fill = "Age") +
    theme(axis.title.y = element_blank()) +
  scale_x_continuous(labels = function(x) round(x * sd_bw + mu_bw, 1)) +
  scale_fill_manual(labels = c(paste0("Low: ",
                                      round(low_age_label,1),"yrs"),
                               paste0("Mid: ",
                                      round(mid_age_label,1),"yrs"),
                               paste0("High: ",
                                      round(high_age_label,1),"yrs")), 
                    values = plot_colors) +
   scale_color_gradientn(colors = plot_colors,
    labels = function(x) round((x * sd_age + mu_age)/365.25, 1),
    name = "") +
  scale_y_continuous(labels = function(y) signif(pnorm(y)*100, 3))
  
  return(model_plot)
})

bwbyage_sp_plots_th <- setNames(bwbyage_sp_plots_th, bwbyage_sp_th_sig_pheno)

bwbyage_sp_plots_th_nopt <- lapply(bwbyage_sp_plots_th, function(x){
x$layers[1] <- NULL
x})

wrapped_plot_n <- 4 

plot_batches <- split(bwbyage_sp_plots_th, ceiling(seq_along(bwbyage_sp_plots_th) / wrapped_plot_n))

 wrapped_list <- lapply(plot_batches, function(batch) {
   wrap_plots(batch, ncol = 2)
 })

wrapped_list

plot_batches_nopt <- split(bwbyage_sp_plots_th_nopt, ceiling(seq_along(bwbyage_sp_plots_th_nopt) / wrapped_plot_n))

 wrapped_list_nopt <- lapply(plot_batches_nopt, function(batch) {
   wrap_plots(batch, ncol = 2)
 })

wrapped_list_nopt
```

