---
title: "slip_swyc"
output: html_document
date: "2026-02-03"
editor_options: 
  chunk_output_type: console
---
Ideas
- get an average swyc score per participant and correlate with 4 global phenotypes + meanthickness + total SA + domain-based CMD & calculate 4 global phenotypes CMD
- try to figure out how to get a "trajectory" per participant

# Controls
```{r setup, echo = TRUE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)
save_figs = TRUE
save_tables = TRUE
save_processed = TRUE
```

# Load packages
```{r}
#Load packages. Install these packages with install.packages("pkg_name") if they are not installed.
library(bigrquery) #package to import SQL tables into R
library(tidyverse) #package suite for data manipulation, reading data into R, and plotting (ggplot2)
library(glue) #package to use expressions with strings
library(ggseg)
require(gridExtra)
library(interactions)
library(boot)
library(parameters)

panel_save <- function(plot, filename, extension, folder, plot_width = 8, plot_height = 10){
  save_name <- paste0(folder, filename, ".", extension)
  ggsave(save_name, plot = plot, dpi = 300, width = plot_width, height = plot_height)
}

`%nin%` = Negate(`%in%`)
```

# Define paths
```{r}
save_path <-  "/mnt/arcus/lab/users/kafadare/swyc_premie_analysis_results/"
if(!dir.exists(save_path)) dir.create(save_path)

plot_output_folder <- paste0(save_path, "plots/")
if(!dir.exists(plot_output_folder)) dir.create(plot_output_folder)

csv_tables_folder <-  paste0(save_path, "processed_tables/")
if (!dir.exists(csv_tables_folder)) dir.create(csv_tables_folder)

csv_output_folder <-  paste0(save_path, "results_csv/")
if (!dir.exists(csv_output_folder)) dir.create(csv_output_folder)

betas_folder <- paste0(csv_output_folder, "full_betas/")
if (!dir.exists(betas_folder)) dir.create(betas_folder)

dkt_folder <- paste0(csv_output_folder, "dkt_betas/")
if (!dir.exists(dkt_folder)) dir.create(dkt_folder)
```


# Load vars names
```{r}
vars_path <- "/mnt/arcus/lab/users/kafadare/slip_brain_vars/"
full_vars <- read.csv(paste0(vars_path,"all_vars.csv")) %>% pull(x)
global_vars <- read.csv(paste0(vars_path,"global_vars.csv")) %>% pull(x)
vol_vars <- read.csv(paste0(vars_path,"vol_vars.csv")) %>% pull(x)
sa_vars <- read.csv(paste0(vars_path,"sa_vars.csv")) %>% pull(x)
th_vars <- read.csv(paste0(vars_path,"th_vars.csv")) %>% pull(x)
th_vars_median <- paste0(th_vars, ".median")

global_vars_totest <- c("global_eTIV", "global_SubCortGrayVol","global_CortexVol", "global_CerebralWhiteMatterVol", "global_VentricleChoroidVol","global_SurfaceArea", "global_MeanThickness")
```

# Load centiles
```{r}
df.zscore_distinct_path <- "/mnt/arcus/lab/users/kafadare/slip_centiles_processed_tables/df_zscore_distinct.csv"
df.zscore.distinct <- read.csv(df.zscore_distinct_path)

df.zscore_all_path <- "/mnt/arcus/lab/users/kafadare/slip_centiles_processed_tables/df_zscore_all.csv"
df.zscore.all <- read.csv(df.zscore_all_path)

centiles_all_path <- "/mnt/arcus/lab/users/kafadare/slip_centiles_processed_tables/centiles_all_median_th_mpr_2026-04-24.csv"
centiles_all <- read.csv(centiles_all_path)

lookup_distinct <- load("/mnt/arcus/lab/users/kafadare/slip_centiles_processed_tables/lookup_stats_distinct.RData")
lookup_distinct <- load("/mnt/arcus/lab/users/kafadare/slip_centiles_processed_tables/lookup_stats_all.RData")

#lmsz_ga_centiles_path <- "/mnt/arcus/lab/users/kafadare/slip_centiles_processed_tables/lmsz_adjusted_centiles_ga.csv"
#lmsz_ga_centiles <- read.csv(lmsz_GA_centiles_path) %>% select(participant_id, session_id, sex, adjusted_age_in_days, age_at_scan, all_of(full_vars), grep("lmsz", names(.))) %>% mutate(lmsz_subset = "GA")

#lmsz_vpm_centiles_path <- "/mnt/arcus/lab/users/kafadare/slip_centiles_processed_tables/lmsz_adjusted_centiles_vpm.csv"
#lmsz_vpm_centiles <- read.csv(lmsz_vpm_centiles_path) %>% select(participant_id, session_id, sex, age_days, age_at_scan, full_vars, #grep("lmsz", names(.))) %>% mutate(lmsz_subset = "VPM")

#lmsz_lpm_centiles_path <- "/mnt/arcus/lab/users/kafadare/slip_centiles_processed_tables/lmsz_adjusted_centiles_lpm.csv"
#lmsz_lpm_centiles <- read.csv(lmsz_lpm_centiles_path) %>% select(participant_id, session_id, sex, age_days, age_at_scan, full_vars, #grep("lmsz", names(.))) %>% mutate(lmsz_subset = "LPM")

#lmsz_centiles <- bind_rows(lmsz_vpm_centiles, lmsz_lpm_centiles)

#lmsz_centiles <- lmsz_ga_centiles

#rm(lmsz_vpm_centiles, lmsz_lpm_centiles, lmsz_GA_centiles)

#centiles_distinct$pat_id <- sub("sub-", "", centiles_distinct$participant_id)
df.zscore.distinct$pat_id <- sub("sub-", "", df.zscore.distinct$participant_id)
df.zscore.all$pat_id <- sub("sub-", "", df.zscore.all$participant_id)
centiles_all$pat_id <- sub("sub-", "", centiles_all$participant_id)

#determine if there is both mpr and median for thickness values
median_th_mpr_flag <- any(grepl(".median", names(centiles_all)))
```


# Remove BW outliers, and corresponding BWP
```{r}
mu_bw <- mean(centiles_all %>% distinct(participant_id, .keep_all = TRUE) %>% select(birth_weight_kg) %>% pull(.), na.rm = TRUE)
sd_bw <- sd(centiles_all %>% distinct(participant_id, .keep_all = TRUE) %>% select(birth_weight_kg) %>% pull(.), na.rm = TRUE)

rm_dist = 5

nrow(centiles_all %>% filter(
  birth_weight_kg < mu_bw-sd_bw*rm_dist | birth_weight_kg > mu_bw+sd_bw*rm_dist) )

centiles_all <- centiles_all %>%
  mutate(birth_weight_kg_inc_out = birth_weight_kg,
    birth_weight_kg = case_when(
        (birth_weight_kg < mu_bw-sd_bw*rm_dist | birth_weight_kg > mu_bw+sd_bw*rm_dist) ~ NA,
        TRUE ~ birth_weight_kg),
    bwp_fen_inc_out = bwp_fen,
    bwp_fen = case_when(
        (birth_weight_kg < mu_bw-sd_bw*rm_dist | birth_weight_kg > mu_bw+sd_bw*rm_dist) ~ NA,
        TRUE ~ bwp_fen))
```

# Colors
```{r}
ga_color = "#4F7CAC"
bwp_color = "#5D9C74"
swyc_color = "#C77D4A"
bw_color = "#4A8E85"

limits = c(-0.3, 0.3)
low_color = "#4878CF" # blue
mid_color = "white"
high_color = "#D65F5F" # red
```

# Load SWYC using SQL for patients with centiles
```{r}
bq_auth()
#proj_id <- if (exists('proj_id')) proj_id else bq_projects()[1]
proj_id = "scit605-healthymri-2d73d69f"


pat_id_list <- paste0("'", sapply(strsplit(centiles_all$pat_id, ","),trimws), "'", collapse = ", ")

#Download SWYC table
swyc_query <- glue("SELECT * FROM 
              `lab.swyc_normative_scores_2026-04-13`
              WHERE pat_id IN ({pat_id_list});")
swyc <- bq_project_query(proj_id, swyc_query) %>% bq_table_download(., page_size=3500)
print(paste0("Arcus lab.SWYC table, dimensions: ", dim(swyc)))
```

# SWYC Date Range
need to pull this value from arcus at some point
```{r}
encounter_id_list <- paste0("'", sapply(strsplit(swyc$encounter_id, ","),trimws), "'", collapse = ", ")

#Download encounter table to get datetime
enc_query <- glue("SELECT encounter_id, effective_datetime, appt_time FROM 
              `arcus.encounter`
              WHERE encounter_id IN ({encounter_id_list});")
encounter <- bq_project_query(proj_id, enc_query) %>% bq_table_download(., page_size=3500)
print(paste0("Arcus encounter table, dimensions: ", dim(encounter)))

sum(swyc$encounter_id %nin% encounter$encounter_id) #0

#merge

swyc <- merge(swyc, encounter, by = "encounter_id", all.x = TRUE)
sum(is.na(swyc$effective_datetime)) # 0
sum(is.na(swyc$appt_time)) # 0

range(swyc$effective_datetime)
```

# Make swyc_clean df with one row per patient-form pair
Appears to be that there is one subject with different sum_answers (hence different DQ) for the exact same appointment time with different encounter id.One of them is called well visit, the other extended well visit. Will use the value from extended well visit.
```{r}
swyc$pat_form <- paste0(swyc$pat_id, swyc$form_friendly_name)

dup_df <- swyc %>%
  group_by(pat_form) %>%
  filter(n() > 1)

dup_df_summary <- dup_df %>%
  group_by(pat_form) %>%
  summarise(
    across(c(n_answers, effective_age, age_mo, norm_devo_quotient),
            ~ sd(.x, na.rm = TRUE),
           .names = "sd_{.col}"),
    .groups = "drop"
  )

median(dup_df_summary$sd_n_answers)
median(dup_df_summary$sd_effective_age)
median(dup_df_summary$sd_age_mo)
median(dup_df_summary$sd_norm_devo_quotient)

max(dup_df_summary$sd_n_answers)
max(dup_df_summary$sd_effective_age)
max(dup_df_summary$sd_age_mo)
max(dup_df_summary$sd_norm_devo_quotient)

swyc_clean <- swyc %>%
  
  group_by(pat_form) %>%
  slice_max(order_by = n_answers, with_ties = TRUE) %>%
  slice_max(order_by = appt_time, with_ties = TRUE) %>%
  slice(which.max(grepl("EXTENDED", appt_prc_name))) %>%
  ungroup(.)

#check after cleaning up based on n_answers and appt_time (choosing the later time), and "extended" over regular visit (if identical time otherwise)
dup_df_clean <- swyc_clean %>%
  group_by(pat_form) %>%
  filter(n() > 1)

length(unique(swyc_clean$pat_id))
tab_swyc_forms <- table(swyc_clean$form_friendly_name)
data.frame(form = names(tab_swyc_forms), count = as.vector(tab_swyc_forms))
```

# Add GA to SWYC
```{r}
swyc_clean <- merge(swyc_clean, centiles_all %>% 
                      select(pat_id, gestational_age), by = "pat_id",
                    all.x = TRUE, all.y = FALSE)

sum(is.na(swyc_clean$gestational_age))
dim(unique(swyc_clean %>% filter(is.na(gestational_age)) %>% select(pat_id)))

swyc_clean <- swyc_clean %>%
  mutate(ga_complete_weeks = floor(gestational_age/7))
```

# Calculate GA-adjusted DQ for SWYC
```{r}
median_ga <- median(swyc_clean %>% distinct(pat_id, .keep_all = TRUE) %>% select(gestational_age) %>% pull(.), 
                    na.rm = TRUE)
median_ga_term <- median(swyc_clean %>% distinct(pat_id, .keep_all = TRUE) %>% 
                           filter(gestational_age >= 7*37) %>% select(gestational_age) %>% pull(.), 
                         na.rm = TRUE)
swyc_clean$ga_diff_termmedian <- median_ga_term - swyc_clean$gestational_age

swyc_clean$ga_adj_effective_age <- swyc_clean$effective_age - swyc_clean$ga_diff_termmedian
swyc_clean$ga_adj_age_mo <- round(swyc_clean$ga_adj_effective_age/30.437)
#swyc_clean$ga_adj_age_mo <- floor(swyc_clean$ga_adj_effective_age/30.437)
plot(swyc_clean$age_mo, swyc_clean$ga_adj_age_mo)
swyc_clean$ga_adj_norm_devo_quotient <- swyc_clean$norm_devo_age/swyc_clean$ga_adj_age_mo
swyc_clean$ga_adj_norm_devo_quotient_round <- round(swyc_clean$ga_adj_norm_devo_quotient, 5)
plot(swyc_clean$norm_devo_quotient , swyc_clean$ga_adj_norm_devo_quotient)

ggplot(swyc_clean %>% filter(gestational_age<=(32*7)), aes(x = norm_devo_quotient, y = ga_adj_norm_devo_quotient_round, color = ga_complete_weeks)) +
  geom_point() +
  labs(x = "OG DQ",
       y = "GA-ADJ DQ",
       title = "GA <= 32 weeks") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "#B0B0B0")

ggplot(swyc_clean %>% filter(gestational_age >= 37*7), aes(x = norm_devo_quotient, y = ga_adj_norm_devo_quotient_round, color = ga_complete_weeks)) +
  geom_point() +
  labs(x = "OG DQ",
       y = "GA-ADJ DQ",
       title = "GA >= 37 weeks") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "#B0B0B0")

ggplot(swyc_clean, aes(x = norm_devo_quotient, y = ga_adj_norm_devo_quotient_round, color = ga_complete_weeks)) +
  geom_point() +
  labs(x = "OG DQ",
       y = "GA-ADJ DQ",
       title = "All GA") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "#B0B0B0")

cor.test(swyc_clean$ga_adj_norm_devo_quotient, swyc_clean$gestational_age)
cor.test(swyc_clean$norm_devo_quotient, swyc_clean$gestational_age)

sum((swyc_clean$gestational_age < median_ga_term) & (swyc_clean$ga_adj_norm_devo_quotient_round < swyc_clean$norm_devo_quotient) & !is.na(swyc_clean$gestational_age)) # should be 0

# swyc_clean[(swyc_clean$gestational_age < median_ga_term) & (swyc_clean$ga_adj_norm_devo_quotient_round < swyc_clean$norm_devo_quotient) & !is.na(swyc_clean$gestational_age), 
#            c("gestational_age", "ga_complete_weeks", "ga_adj_norm_devo_quotient", "norm_devo_quotient", "ga_adj_age_mo", "age_mo", "norm_devo_age")]
```

# Create Swyc-Avg with median DQ for each subject
```{r}
swyc_avg <- swyc_clean %>%
  group_by(pat_id) %>%
  summarise(n_forms = n(),
    sex = first(sex_abbr),
    ga_complete_weeks = first(ga_complete_weeks),
    gestational_age   = first(gestational_age),
    all_2s_sum = sum(all_2s == TRUE, na.rm = TRUE),
    all_0s_sum = sum(all_0s == TRUE, na.rm = TRUE),
    final_swyc_age = max(effective_age, na.rm = TRUE),
    final_swyc_age_ga_adj = max(ga_adj_effective_age),
    median_dq = median(norm_devo_quotient, na.rm = TRUE), 
    ga_adj_median_dq = median(ga_adj_norm_devo_quotient_round, na.rm = TRUE),
    .groups = "drop")

hist(swyc_avg$median_dq)
hist(swyc_avg$ga_adj_median_dq)
hist(swyc_avg$final_swyc_age/30.437)
```

# Merge with centiles

## Find SWYCs after a scan and take the median among those (using centiles all)
```{r}
swyc_clean_all_cent <- merge(swyc_clean %>% 
                             rename(gestational_age_raw = gestational_age), 
                           centiles_all %>% select(pat_id, age_at_scan),
                           by = "pat_id", all.x = TRUE, all.y = FALSE)

length(unique(swyc_clean_all_cent$pat_id))
length(unique(swyc_clean_all_cent$pat_form))
dim(swyc_clean_all_cent)
swyc_clean_all_cent$scan_swyc_age_diff_days <- swyc_clean_all_cent$age_at_scan - swyc_clean_all_cent$effective_age
ggplot(swyc_clean_all_cent, aes(x = scan_swyc_age_diff_days)) + 
         geom_histogram(bins = 100) + 
         labs(x = "", y = "Count", 
              title = "Days Between Scan and SWYC Form") +
  theme_classic()
swyc_clean_all_cent$scan_before_swyc <- swyc_clean_all_cent$age_at_scan < swyc_clean_all_cent$effective_age
sum(swyc_clean_all_cent$scan_before_swyc) # 995
# if limited scan within two years of swyc
sum(swyc_clean_all_cent$scan_swyc_age_diff_days <= (365.25*2)) # 1706 rows

sum(swyc_clean_all_cent$scan_swyc_age_diff_days <= (365.25*2) & swyc_clean_all_cent$scan_before_swyc) # 995

# within each pat-form, choose scan closest to swyc
swyc_clean_all_cent_subset <- swyc_clean_all_cent %>%
  group_by(pat_form) %>%
  slice_max(order_by = scan_swyc_age_diff_days, with_ties = TRUE) %>%
  ungroup()
```

## Merge swyc avg with distinct centiles
```{r}
swyc_avg_centiles <- merge(swyc_avg %>% 
                             rename(gestational_age_raw = gestational_age) %>%
                             select(-sex), 
                           df.zscore.distinct, 
                     by = "pat_id", all.x = TRUE, all.y = TRUE)

swyc_avg_centiles <- merge(swyc_avg_centiles,
                           centiles_all %>% select(session_id, age_at_scan) %>%
                             rename(age_at_scan_raw = age_at_scan), 
                     by = "session_id", all.x = TRUE, all.y = FALSE)

swyc_avg_centiles$scan_swyc_age_diff_days <- swyc_avg_centiles$age_at_scan_raw - swyc_avg_centiles$final_swyc_age

ggplot(swyc_avg_centiles, aes(x = scan_swyc_age_diff_days)) + 
         geom_histogram(bins = 100) + 
         labs(x = "", y = "Count", 
              title = "Days Between Scan and Final Age Used for SWYC Median") +
  theme_classic()

swyc_avg_centiles <- swyc_avg_centiles %>%
  mutate(scan_swyc_twoyears = abs(scan_swyc_age_diff_days) <= (365.25*2))

sum(swyc_avg_centiles$scan_swyc_twoyears, na.rm = T) # 553
sum(!is.na(swyc_avg_centiles$median_dq)) # 865
```

# Plot Scan-SWYC timing
```{r}
(swyc_distinct_overlap_plot <- ggplot(swyc_avg_centiles %>% 
                                        filter(!is.na(median_dq) & scan_swyc_twoyears), 
                                      aes(y = reorder(pat_id, age_at_scan))) +
  geom_point(aes(x = final_swyc_age, color = "SWYC"), size = 0.5) +
  geom_point(aes(x = age_at_scan_raw, color = "Scan"), size = 0.5) +
  scale_x_continuous(labels = \(x) round(x/30.4)) +
  scale_color_manual( values = c("SWYC" = swyc_color,
                                 "Scan" = "darkgray")
                      ) +
  labs(x = "Age (months)",
       y = "Patients",
       title = "Final SWYC Age Within Two Years of Scan") +
  theme_classic(base_family = "Helvetica", base_size = 14) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        plot.title = element_text(hjust = 0.5),
        legend.position = "bottom")
 )
panel_save(swyc_distinct_overlap_plot, "swyc_distinct_scan_twoyears_overlap_plot", "svg", plot_output_folder)
```

# Prep Data for Analysis

## zscore variables of interest
```{r, echo = FALSE}
# zscore swyc allforms birth factors and DQ
swyc_avg_centiles <- swyc_avg_centiles %>%
  mutate(across(all_of(c("median_dq", "ga_adj_median_dq", "final_swyc_age", "final_swyc_age_ga_adj")),
                ~as.numeric(scale(.x)), .names = "{.col}_zscore"))
# set sex and preterm as factor
swyc_avg_centiles$sex <- as.factor(swyc_avg_centiles$sex)
#swyc_avg_centiles$bwp_fen <- as.numeric(swyc_avg_centiles$bwp_fen)

# keep a column for all median dq available with scans, but set those > 2 years from scan to NULL

swyc_avg_centiles <- swyc_avg_centiles %>%
  mutate(median_dq_full = median_dq,
         ga_adj_median_dq_full = ga_adj_median_dq,
         median_dq_zscore_full = median_dq_zscore,
         ga_adj_median_dq_zscore_full = ga_adj_median_dq_zscore)
swyc_avg_centiles[which(!swyc_avg_centiles$scan_swyc_twoyears),]$median_dq <- NA
swyc_avg_centiles[which(!swyc_avg_centiles$scan_swyc_twoyears),]$ga_adj_median_dq <- NA
swyc_avg_centiles[which(!swyc_avg_centiles$scan_swyc_twoyears),]$median_dq_zscore <- NA
swyc_avg_centiles[which(!swyc_avg_centiles$scan_swyc_twoyears),]$ga_adj_median_dq_zscore <- NA
```
# CMDs

## Correlation Matrix
```{r, echo = FALSE}
# correlations between CMDs across three centiles
cor_mat <- cor(swyc_avg_centiles[,c("global.cmd", "vol.cmd", "sa.cmd", "th.cmd", "subcort.cmd",
                                    "gestational_age", "birth_weight_kg", "bwp_fen",
                                    "median_dq_zscore", "ga_adj_median_dq_zscore",
                                    "median_dq_zscore_full", "ga_adj_median_dq_zscore_full")], 
               use = "pairwise.complete.obs")

p_mat <- cor_pmat(swyc_avg_centiles[,c("global.cmd", "vol.cmd", "sa.cmd", "th.cmd", "subcort.cmd",
                                    "gestational_age", "birth_weight_kg", "bwp_fen",
                                    "median_dq_zscore", "ga_adj_median_dq_zscore",
                                    "median_dq_zscore_full", "ga_adj_median_dq_zscore_full")], 
                  use = "pairwise.complete.obs")

(cmd_corrs_plot <- ggcorrplot(
  cor_mat,
  p.mat = p_mat,
  insig = "blank",
  lab = TRUE
))

if(save_figs == TRUE){
panel_save(cmd_corrs_plot, "cmd-cor-matrix", "svg", plot_output_folder)
}

```

## CMD plots
```{r}
swyc_N <- sum(!is.na(swyc_avg_centiles$median_dq))
# Vol CMD
cor_volCMD_swyc <- cor.test(swyc_avg_centiles$vol.cmd, swyc_avg_centiles$median_dq)
## plot
volCMD_swyc_r <- signif(cor_volCMD_swyc$estimate, 3)
volCMD_swyc_p <- signif(cor_volCMD_swyc$p.value, 3)

(cor_volCMD_swyc_plot <- ggplot(swyc_avg_centiles, aes(median_dq_zscore, vol.cmd)) +
  geom_point(color = swyc_color, alpha = 0.4) +
  geom_smooth(method = "lm", se = TRUE, color = swyc_color) +
  annotate("label",
           x = 2, y = 10,
           label.size = 0,
           label = paste0("r=",volCMD_swyc_r, "; p=", volCMD_swyc_p)) +
  labs(
    x = paste0("Median SWYC DQ, N = ", swyc_N),
    y = "Volume CMD"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    plot.title = element_blank()
  ))

# SA CMD
cor_saCMD_swyc <- cor.test(swyc_avg_centiles$sa.cmd, swyc_avg_centiles$median_dq)
## plot
saCMD_swyc_r <- signif(cor_saCMD_swyc$estimate, 3)
saCMD_swyc_p <- signif(cor_saCMD_swyc$p.value, 3)

(cor_saCMD_swyc_plot <- ggplot(swyc_avg_centiles, aes(median_dq, sa.cmd)) +
  geom_point(color = swyc_color, alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE, color = swyc_color) +
  annotate("label",
           x = 2, y = 10,
           label.size = 0,
           label = paste0("r=",saCMD_swyc_r, "; p=", saCMD_swyc_p)) +
  labs(
    x = paste0("Median SWYC DQ, N = ", swyc_N),
    y = "Surface Area CMD"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    plot.title = element_blank()
  ))

# Thickness CMD
swyc_th_N <- sum(!is.na(swyc_avg_centiles$median_dq) & !is.na(swyc_avg_centiles$th.cmd))
cor_thCMD_swyc <- cor.test(swyc_avg_centiles$th.cmd, swyc_avg_centiles$median_dq)
## plot
thCMD_swyc_r <- signif(cor_thCMD_swyc$estimate, 3)
thCMD_swyc_p <- signif(cor_thCMD_swyc$p.value, 3)

(cor_thCMD_swyc_plot <- ggplot(swyc_avg_centiles, aes(median_dq, th.cmd)) +
  geom_point(color = swyc_color, alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE, color = swyc_color) +
  annotate("label",
           x = 2, y = 10,
           label.size = 0,
           label = paste0("r=",thCMD_swyc_r, "; p=", thCMD_swyc_p)) +
  labs(
    x = paste0("Median SWYC DQ, N = ", swyc_th_N),
    y = "Cortical Thickness CMD"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    plot.title = element_blank()
  ))

#Global CMD
cor_globalCMD_swyc <- cor.test(swyc_avg_centiles$global.cmd, swyc_avg_centiles$median_dq)
## plot
globalCMD_swyc_r <- signif(cor_globalCMD_swyc$estimate, 3)
globalCMD_swyc_p <- signif(cor_globalCMD_swyc$p.value, 3)

(cor_globalCMD_swyc_plot <- ggplot(swyc_avg_centiles, aes(median_dq, global.cmd)) +
  geom_point(color = swyc_color, alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE, color = swyc_color) +
  annotate("label",
           x = 1, y = 7,
           label.size = 0,
           label = paste0("r=",globalCMD_swyc_r, "; p=", globalCMD_swyc_p)) +
  labs(
    x = paste0("Median SWYC DQ, N = ", swyc_N),
    y = "Global CMD"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    plot.title = element_blank()
  ))

#Subcort CMD
cor_subcortCMD_swyc <- cor.test(swyc_avg_centiles$subcort.cmd, swyc_avg_centiles$median_dq)
## plot
subcortCMD_swyc_r <- signif(cor_subcortCMD_swyc$estimate, 3)
subcortCMD_swyc_p <- signif(cor_subcortCMD_swyc$p.value, 3)

(cor_subcortCMD_swyc_plot <- ggplot(swyc_avg_centiles, aes(median_dq, subcort.cmd)) +
  geom_point(color = swyc_color, alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE, color = swyc_color) +
  annotate("label",
           x = 1, y = 10,
           label.size = 0,
           label = paste0("r=",subcortCMD_swyc_r, "; p=", subcortCMD_swyc_p)) +
  labs(
    x = paste0("Median SWYC DQ, N = ", swyc_N),
    y = "Subcortical CMD"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    plot.title = element_blank()
  ))
```



# Linear Regression Global phenotypes (euler mean QC score as covariate)

## GA
```{r, echo = FALSE}
ga_reg_table_global <- as.data.frame(t(sapply(global_vars_totest, function(x) {
  dat <- swyc_avg_centiles
  if (grepl("Thick", x)) {
    dat <- swyc_avg_centiles %>% filter(age_over2_adj == TRUE)
    if(median_th_mpr_flag){
      form <- reformulate(termlabels = c("gestational_age", "euler_mean_mpr"), response = x)
    }
    } else {
     form <- reformulate(termlabels = c("gestational_age", "qc_var"), response = x)
    }
   fit <- lm(form, data = dat)
   summary_fit <- summary(fit)
  c(ga_beta = summary_fit$coefficients["gestational_age", "Estimate"],
    ga_p    = summary_fit$coefficients["gestational_age", "Pr(>|t|)"],
    ga_se   = summary_fit$coefficients["gestational_age", "Std. Error"] )
})))

ga_reg_table_global$ga_p_adj <- p.adjust(ga_reg_table_global$ga_p, method = "BH")

ga_reg_table_global %>% filter(ga_p_adj < 0.05)
range(ga_reg_table_global %>% filter(ga_p_adj < 0.05) %>% select(ga_beta))
```

## BW
```{r, echo = FALSE}
bw_reg_table_global <- as.data.frame(t(sapply(global_vars_totest, function(x) {
  dat <- swyc_avg_centiles
  if (grepl("Thick", x)) {
    dat <- swyc_avg_centiles %>% filter(age_over2_adj == TRUE)
    if(median_th_mpr_flag){
      form <- reformulate(termlabels = c("birth_weight_kg", "euler_mean_mpr"), response = x)
    }
    } else {
     form <- reformulate(termlabels = c("birth_weight_kg", "qc_var"), response = x)
    }
   fit <- lm(form, data = dat)
   summary_fit <- summary(fit)
  c(bw_beta = summary_fit$coefficients["birth_weight_kg", "Estimate"],
    bw_p    = summary_fit$coefficients["birth_weight_kg", "Pr(>|t|)"],
    bw_se   = summary_fit$coefficients["birth_weight_kg", "Std. Error"] )
})))

bw_reg_table_global$bw_p_adj <- p.adjust(bw_reg_table_global$bw_p, method = "BH")

bw_reg_table_global %>% filter(bw_p_adj < 0.05)
range(bw_reg_table_global %>% filter(bw_p_adj < 0.05) %>% select(bw_beta))
```

## BWp
```{r, echo = FALSE}
bwp_reg_table_global <- as.data.frame(t(sapply(global_vars_totest, function(x) {
  dat <- swyc_avg_centiles
  if (grepl("Thick", x)) {
    dat <- swyc_avg_centiles %>% filter(age_over2_adj == TRUE)
    if(median_th_mpr_flag){
      form <- reformulate(termlabels = c("bwp_fen", "euler_mean_mpr"), response = x)
    }
    } else {
     form <- reformulate(termlabels = c("bwp_fen", "qc_var"), response = x)
    }
   fit <- lm(form, data = dat)
   summary_fit <- summary(fit)
  c(bwp_beta = summary_fit$coefficients["bwp_fen", "Estimate"],
    bwp_p    = summary_fit$coefficients["bwp_fen", "Pr(>|t|)"],
    bwp_se   = summary_fit$coefficients["bwp_fen", "Std. Error"] )
})))

bwp_reg_table_global$bwp_p_adj <- p.adjust(bwp_reg_table_global$bwp_p, method = "BH")

bwp_reg_table_global %>% filter(bwp_p_adj < 0.05)
range(bwp_reg_table_global %>% filter(bwp_p_adj < 0.05) %>% select(bwp_beta))
```

## SWYC Median DQ
```{r, echo = FALSE}
swyc_reg_table_global <- as.data.frame(t(sapply(global_vars_totest, function(x) {
  dat <- swyc_avg_centiles
  if (grepl("Thick", x)) {
    dat <- swyc_avg_centiles %>% filter(age_over2_adj == TRUE)
    if(median_th_mpr_flag){
      form <- reformulate(termlabels = c("median_dq_zscore", "euler_mean_mpr"), response = x)
    }
    } else {
     form <- reformulate(termlabels = c("median_dq_zscore", "qc_var"), response = x)
    }
   fit <- lm(form, data = dat)
   summary_fit <- summary(fit)
  c(swyc_beta = summary_fit$coefficients["median_dq_zscore", "Estimate"],
    swyc_p    = summary_fit$coefficients["median_dq_zscore", "Pr(>|t|)"],
    swyc_se   = summary_fit$coefficients["median_dq_zscore", "Std. Error"] )
})))

swyc_reg_table_global$swyc_p_adj <- p.adjust(swyc_reg_table_global$swyc_p, method = "BH")

swyc_reg_table_global %>% filter(swyc_p_adj < 0.05)
range(swyc_reg_table_global %>% filter(swyc_p_adj < 0.05) %>% select(swyc_beta))
```

## Visualize/Figure Panel
```{r, echo = FALSE}
#  Add phenotype column from rownames
ga_reg_table_global$phenotype <- rownames(ga_reg_table_global)
swyc_reg_table_global$phenotype <- rownames(swyc_reg_table_global)
bwp_reg_table_global$phenotype <- rownames(bwp_reg_table_global)

#  Add model identifiers + rename to common column names
ga_reg_table_global <- ga_reg_table_global %>%
  mutate(model = "Gestational Age",
         beta = ga_beta,
         se   = ga_se,
         p_adj    = ga_p_adj) %>%
    mutate(sig = p_adj < 0.05,
          color = case_when(
            sig ~ ga_color,
            TRUE ~ "gray70"))

bwp_reg_table_global <- bwp_reg_table_global %>%
  mutate(model = "Birth Weight Percentile",
         beta = bwp_beta,
         se   = bwp_se,
         p_adj    = bwp_p_adj) %>%
    mutate(sig = p_adj < 0.05,
          color = case_when(
            sig  ~ bwp_color,
            TRUE ~ "gray70"))

swyc_reg_table_global <- swyc_reg_table_global %>%
  mutate(model = "SWYC",
         beta = swyc_beta,
         se   = swyc_se,
         p_adj    = swyc_p_adj) %>%
    mutate(sig = p_adj < 0.05,
          color = case_when(
            sig ~ swyc_color,
            TRUE ~ "gray70"))

#  Combine and arrange by specific order of global phenotypes
df <- bind_rows(ga_reg_table_global, bwp_reg_table_global, swyc_reg_table_global) %>%
  mutate(phenotype = factor(phenotype, levels = c("global_eTIV", "global_CortexVol", "global_CerebralWhiteMatterVol",
                                                  "global_SubCortGrayVol", "global_VentricleChoroidVol",
                                                  "global_SurfaceArea", "global_MeanThickness"))) %>% arrange(phenotype)

#  Labeling for phenotypes on plot
phenotype_labs <- c(
 global_eTIV = "Intracranial Volume",
 global_MeanThickness = "Mean Cortical Thickness",
 global_VentricleChoroidVol = "Ventricular Volume",
 global_SurfaceArea = "Total Surface Area",
 global_CortexVol = "Cortical Gray Matter Volume",
 global_CerebralWhiteMatterVol = "Cerebral White Matter Volume",
 global_SubCortGrayVol = "Subcortical Gray Matter Volume"
)

#  Plot
(global_betas_plot <- ggplot(df, aes(x = beta, y = phenotype)) +
    geom_point(aes(shape = model, color = color), size = 6) +
    geom_errorbarh(aes(xmin = beta - se, xmax = beta + se, color = color), height = 0.1) +
    scale_shape_manual(values = c("Gestational Age" = 16, "Birth Weight Percentile" = 17, "SWYC" = 15),
                       name = "") +
    scale_color_identity() +
    scale_x_continuous(limits = c(-0.2, 0.3)) +
    scale_y_discrete(limits = rev, labels = phenotype_labs) +
    geom_vline(xintercept = 0, linetype = "dotted", color = "darkgray") +
    theme_minimal() +
    theme_bw(base_size = 30) +
    theme(
        panel.grid = element_blank(),
        legend.position = "bottom",
        legend.title = element_blank(),
        panel.border = element_blank(),
        axis.text.y  = element_text(color = "black"),
        axis.line.x  = element_line(color = "black"),
        axis.line.y  = element_blank(),
        legend.justification = c("left", "top")
    ) +
    xlab("Standardized Beta Estimate, QC covar") +
    ylab(""))

if(save_figs == TRUE){
panel_save(global_betas_plot, "global-betas-gabw", "svg", plot_output_folder, plot_width = 15, plot_height = 15)
}
```

# Global Scatterplots for SWYC
```{r}
swyc_N <- sum(!is.na(swyc_avg_centiles$median_dq) & !is.na(swyc_avg_centiles$global_VentricleChoroidVol))

#Ventricles
swyc_vent_beta <- swyc.qc_reg_table_global %>% 
  filter(phenotype == "global_VentricleChoroidVol") %>% 
  select(swyc_beta) %>% pull(swyc_beta)
swyc_vent_p_adj <- swyc.qc_reg_table_global %>% 
  filter(phenotype == "global_VentricleChoroidVol") %>% 
  select(swyc_p_adj) %>% pull(swyc_p_adj)
(swyc_vent_plot <- ggplot(swyc_avg_centiles, aes(x = median_dq, y = global_VentricleChoroidVol)) +
  geom_point(color = "#7293b0", alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE, color = "#7293b0") +
  annotate("label",
           x = 0.5, y = 3.2,
           label = paste0("std beta=", signif(swyc_vent_beta,3), "; p-adj=", signif(swyc_vent_p_adj,3))) +
  labs(
    x = paste0("Median SWYC Score (z-scored), N = ", swyc_N),
    y = "Ventricle Volume Centile (z-scored)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    plot.title = element_blank()
  ))
if(save_figs == TRUE){
panel_save(swyc_vent_plot, "swyc-ventricles-withqc", "pdf", plot_output_folder)
}

#Subcortical GMV
swyc_scgmv_beta <- swyc.qc_reg_table_global %>% 
  filter(phenotype == "global_SubCortGrayVol") %>% 
  select(swyc_beta) %>% pull(swyc_beta)
swyc_scgmv_p_adj <- swyc.qc_reg_table_global %>% 
  filter(phenotype == "global_SubCortGrayVol") %>% 
  select(swyc_p_adj) %>% pull(swyc_p_adj)
(swyc_scgmv_plot <- ggplot(swyc_avg_centiles, aes(x = median_dq, y = global_SubCortGrayVol)) +
  geom_point(color = "#c87a7a", alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE, color = "#c87a7a") +
  annotate("label",
           x = 0.5, y = 3.2,
           label = paste0("std beta=", signif(swyc_scgmv_beta,3), "; p-adj=", signif(swyc_scgmv_p_adj,3))) +
  labs(
    x = paste0("Median SWYC Score (z-scored), N = ", swyc_N),
    y = "Subcortical Gray Volume Centile (z-scored)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    plot.title = element_blank()
  ))

if(save_figs == TRUE){
panel_save(swyc_scgmv_plot, "swyc-scgmv-withqc", "pdf", plot_output_folder)
}

#Cortical GMV
swyc_ctxgmv_beta <- swyc.qc_reg_table_global %>% 
  filter(phenotype == "global_CortexVol") %>% 
  select(swyc_beta) %>% pull(swyc_beta)
swyc_ctxgmv_p_adj <- swyc.qc_reg_table_global %>% 
  filter(phenotype == "global_CortexVol") %>% 
  select(swyc_p_adj) %>% pull(swyc_p_adj)
(swyc_ctxgmv_plot <- ggplot(swyc_avg_centiles, aes(x = median_dq, y = global_CortexVol)) +
  geom_point(color = "#c87a7a", alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE, color = "#c87a7a") +
  annotate("label",
           x = 0.5, y = 3.2,
           label = paste0("std beta=", signif(swyc_ctxgmv_beta,3), "; p-adj=", signif(swyc_ctxgmv_p_adj,3))) +
  labs(
    x = paste0("Median SWYC ctxore (z-ctxored), N = ", swyc_N),
    y = "Cortical Gray Volume Centile (z-ctxored)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    plot.title = element_blank()
  ))

if(save_figs == TRUE){
panel_save(swyc_ctxgmv_plot, "swyc-ctxgmv-withqc", "pdf", plot_output_folder)
}


```

# Regional Variables --- only for SWYC

## Volume with QC covar

### simple linear regressions

#### compute regressions
```{r, echo = FALSE}
# Regional -- including cortical and subcortical
swyc_vol <- as.data.frame(t(sapply(swyc_avg_centiles[, which(names(swyc_avg_centiles) %in% vol_vars)], function(x) {
  lm <- lm(x ~ median_dq_zscore + median_euler_mean, data = swyc_avg_centiles)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2])
})))

swyc_vol$swyc_p_adj <- p.adjust(swyc_vol$swyc_p, method = "BH")
print(paste0("Number of significant IDPs after FDR correction: ", dim(swyc_vol %>% filter(swyc_p_adj < 0.05))[1]))
```

#### add label column to match to ggseg
```{r, echo = FALSE}
cols_edit <- grepl("aseg(?!_CC)", rownames(swyc_vol), perl = TRUE)

swyc_vol$label <- rownames(swyc_vol) %>% 
  sub("aparc_GrayVol_", "", .) %>% 
  sub("aseg_", "", .) %>%
  sub("X", "x", .) %>%
  sub("rd_Ventricle", "rd_ventricle", .) %>% 
  sub("th_Ventricle", "th_ventricle", .) %>% 
  sub("Thalamus", "Thalamus-Proper", .)

swyc_vol$label[cols_edit] <- gsub("_", "-", swyc_vol$label[cols_edit])

swyc_vol_sig_MC <- swyc_vol %>% filter(swyc_p_adj < 0.05)

print(paste0("Beta range of significant IDPs: ", range(swyc_vol_sig_MC$swyc_beta)))
```

#### save betas
```{r, echo = FALSE}
save_table_full <- swyc_vol
name <- "swyc_median_volume.csv"
# Save dkt cortical betas for spin test
if(save_tables == TRUE){
  output_folder <-dkt_folder
  save_table_full$hemisphere <- ifelse(grepl("lh", save_table_full$label), "L", 
                         ifelse(grepl("rh", save_table_full$label), "R", NA))
  save_table_full$label <- sub("^rh_|lh_", "", save_table_full$label)
  filename_all <- paste0(output_folder, name)
  write.csv(save_table_full, filename_all)
}
# Save full betas
if(save_tables == TRUE){
output_folder <- betas_folder
filename <- paste0(output_folder, name)
write.csv(save_table_full, filename)
}
```

#### visualization of betas on surface maps
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(swyc_vol_sig_MC, ggseg::dk, by = "label")
plot_data_aseg <- merge(swyc_vol_sig_MC, ggseg::aseg, by = "label")

# Regions not matched in atlases
(swyc_vol_noplot <- swyc_vol_sig_MC[!(swyc_vol_sig_MC$label %in% plot_data_dkt$label) & !(swyc_vol_sig_MC$label %in% plot_data_aseg$label),])

# plots
(swyc_dkt_vol_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = swyc_beta)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = limits) +
  theme_void()+
  labs(title = "Median SWYC Volume"))

(swyc_aseg_vol_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = swyc_beta)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = limits) +
  theme_void() +
  labs(title = "Median SWYC Volume"))

if(save_figs == TRUE){
panel_save(swyc_dkt_vol_plot, "swyc-median-dkt-vol", "svg", plot_output_folder)
panel_save(swyc_aseg_vol_plot , "swyc-median-aseg-vol", "svg", plot_output_folder)
}
```

### SWYC X age interaction

#### Regression and results
```{r, echo = FALSE}
swycbyage.qc_reg_table_vol <- as.data.frame(t(sapply(swyc_avg_centiles[, which(names(swyc_avg_centiles) %in% vol_vars)], function(x) {
  lm <- lm(x ~ swyc_avg_centiles$median_dq_zscore*swyc_avg_centiles$age_days_adj + swyc_avg_centiles$median_euler_mean)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2],
   age_beta = summary$coefficients[3,1], age_p = summary$coefficients[3,4], age_se = summary$coefficients[3,2],
   swycbyage_beta = summary$coefficients[5,1], swycbyage_p = summary$coefficients[5,4], swycbyage_se = summary$coefficients[5,2])
})))

swycbyage.qc_reg_table_vol$swyc_p_adj <- p.adjust(swycbyage.qc_reg_table_vol$swyc_p, method = "BH")
swycbyage.qc_reg_table_vol$age_p_adj <- p.adjust(swycbyage.qc_reg_table_vol$age_p, method = "BH")
swycbyage.qc_reg_table_vol$swycbyage_p_adj <- p.adjust(swycbyage.qc_reg_table_vol$swycbyage_p, method = "BH")

print(paste0("Age Main significant results: ", dim(swycbyage.qc_reg_table_vol %>% filter(age_p_adj < 0.05))[1]))
print(paste0("Number of significant SWYC x Age IDPs - NO correction: ", dim(swycbyage.qc_reg_table_vol %>% filter(swycbyage_p < 0.05))[1]))
print(paste0("Number of significant SWYC x Age IDPs - AFTER FDR correction: ", dim(swycbyage.qc_reg_table_vol %>% filter(swycbyage_p_adj < 0.05))[1]))

range(swycbyage.qc_reg_table_vol %>% filter(swycbyage_p_adj < 0.05) %>% select(swycbyage_beta))
```


## Surface Area

###  simple linear regressions

#### compute regressions
```{r, echo = FALSE}
sum(!is.na(swyc_avg_centiles$median_dq_zscore))
# Regional -- including cortical and subcortical
swyc_sa <- as.data.frame(t(sapply(swyc_avg_centiles[, which(names(swyc_avg_centiles) %in% sa_vars)], function(x) {
  lm <- lm(x ~ median_dq_zscore + median_euler_mean, data = swyc_avg_centiles)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2])
})))

swyc_sa$swyc_p_adj <- p.adjust(swyc_sa$swyc_p, method = "BH")
print(paste0("Number of significant IDPs after FDR correction: ", dim(swyc_sa %>% filter(swyc_p_adj < 0.05))[1]))
```

#### add label column to match to ggseg
```{r, echo = FALSE}
swyc_sa$label <- rownames(swyc_sa) %>% 
  sub("aparc_SurfArea_", "", .)

swyc_sa_sig_MC <- swyc_sa %>% filter(swyc_p_adj < 0.05)

print(paste0("Beta range of significant IDPs: ", range(swyc_sa_sig_MC$swyc_beta)))
```

#### save betas
```{r, echo = FALSE}
save_table_full <- swyc_sa
name <- "swyc_median_surfArea.csv"
# Save dkt cortical betas for spin test
if(save_tables == TRUE){
  output_folder <-dkt_folder
  save_table_full$hemisphere <- ifelse(grepl("lh", save_table_full$label), "L", 
                         ifelse(grepl("rh", save_table_full$label), "R", NA))
  save_table_full$label <- sub("^rh_|lh_", "", save_table_full$label)
  filename_all <- paste0(output_folder, name)
  write.csv(save_table_full, filename_all)
}
# Save full betas
if(save_tables == TRUE){
output_folder <- betas_folder
filename <- paste0(output_folder, name)
write.csv(save_table_full, filename)
}
```

#### visualization of betas on surface maps
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(swyc_sa_sig_MC, ggseg::dk, by = "label")

# Regions not matched in atlases
(swyc_sa_noplot <- swyc_sa_sig_MC[!(swyc_sa_sig_MC$label %in% plot_data_dkt$label) & !(swyc_sa_sig_MC$label %in% plot_data_aseg$label),])

# plots
(swyc_dkt_sa_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = swyc_beta)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = limits) +
  theme_void()+
  labs(title = "Median SWYC Surface Area"))

if(save_figs == TRUE){
panel_save(swyc_dkt_sa_plot, "swyc-median-dkt-sa", "svg", plot_output_folder)
}
```

### SWYC X age interaction

#### Regression and results
```{r, echo = FALSE}
swycbyage.qc_reg_table_sa <- as.data.frame(t(sapply(swyc_avg_centiles[, which(names(swyc_avg_centiles) %in% sa_vars)], function(x) {
  lm <- lm(x ~ swyc_avg_centiles$median_dq_zscore*swyc_avg_centiles$age_days_adj + swyc_avg_centiles$median_euler_mean)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2],
   age_beta = summary$coefficients[3,1], age_p = summary$coefficients[3,4], age_se = summary$coefficients[3,2],
   swycbyage_beta = summary$coefficients[5,1], swycbyage_p = summary$coefficients[5,4], swycbyage_se = summary$coefficients[5,2])
})))

swycbyage.qc_reg_table_sa$swyc_p_adj <- p.adjust(swycbyage.qc_reg_table_sa$swyc_p, method = "BH")
swycbyage.qc_reg_table_sa$age_p_adj <- p.adjust(swycbyage.qc_reg_table_sa$age_p, method = "BH")
swycbyage.qc_reg_table_sa$swycbyage_p_adj <- p.adjust(swycbyage.qc_reg_table_sa$swycbyage_p, method = "BH")

print(paste0("Age Main significant results: ", dim(swycbyage.qc_reg_table_sa %>% filter(age_p_adj < 0.05))[1]))
print(paste0("Number of significant SWYC x Age IDPs - NO correction: ", dim(swycbyage.qc_reg_table_sa %>% filter(swycbyage_p < 0.05))[1]))
print(paste0("Number of significant SWYC x Age IDPs - AFTER FDR correction: ", dim(swycbyage.qc_reg_table_sa %>% filter(swycbyage_p_adj < 0.05))[1]))

range(swycbyage.qc_reg_table_sa %>% filter(swycbyage_p_adj < 0.05) %>% select(swycbyage_beta))
```

## Thickness

###  simple linear regressions

#### compute regressions
```{r, echo = FALSE}
sum(!is.na(swyc_avg_centiles$aparc_ThickAvg_lh_bankssts))
# Regional -- including cortical and subcortical
swyc_th <- as.data.frame(t(sapply(swyc_avg_centiles[, which(names(swyc_avg_centiles) %in% th_vars)], function(x) {
  lm <- lm(x ~ median_dq_zscore + euler_mean_mpr, data = swyc_avg_centiles)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2])
})))

swyc_th$swyc_p_adj <- p.adjust(swyc_th$swyc_p, method = "BH")
print(paste0("Number of significant IDPs after FDR correction: ", dim(swyc_th %>% filter(swyc_p_adj < 0.05))[1]))
```

#### add label column to match to ggseg
```{r, echo = FALSE}
swyc_th$label <- rownames(swyc_th) %>% 
  sub("aparc_ThickAvg_", "", .)

swyc_th_sig_MC <- swyc_th %>% filter(swyc_p_adj < 0.05)

print(paste0("Beta range of significant IDPs: ", range(swyc_th_sig_MC$swyc_beta)))
```

#### save betas
```{r, echo = FALSE}
save_table_full <- swyc_th
name <- "swyc_thick.csv"
# Save dkt cortical betas for spin test
if(save_tables == TRUE){
  output_folder <-dkt_folder
  save_table_full$hemisphere <- ifelse(grepl("lh", save_table_full$label), "L", 
                         ifelse(grepl("rh", save_table_full$label), "R", NA))
  save_table_full$label <- sub("^rh_|lh_", "", save_table_full$label)
  filename_all <- paste0(output_folder, name)
  write.csv(save_table_full, filename_all)
}
# Save full betas
if(save_tables == TRUE){
output_folder <- betas_folder
filename <- paste0(output_folder, name)
write.csv(save_table_full, filename)
}
```

#### visualization of betas on surface maps
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(swyc_th_sig_MC, ggseg::dk, by = "label")

# Regions not matched in atlases
(swyc_th_noplot <- swyc_th_sig_MC[!(swyc_th_sig_MC$label %in% plot_data_dkt$label) & !(swyc_th_sig_MC$label %in% plot_data_aseg$label),])

# plots
(swyc_dkt_th_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = swyc_beta)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = limits) +
  theme_void()+
  labs(title = "Median SWYC Cortical Thickness"))

if(save_figs == TRUE){
panel_save(swyc_dkt_th_plot, "swyc-median-dkt-th", "svg", plot_output_folder)
}
```

### SWYC X age interaction

#### Regression and results
```{r, echo = FALSE}
swycbyage.qc_reg_table_th <- as.data.frame(t(sapply(swyc_avg_centiles[, which(names(swyc_avg_centiles) %in% th_vars)], function(x) {
  lm <- lm(x ~ swyc_avg_centiles$median_dq_zscore*swyc_avg_centiles$age_days_adj + swyc_avg_centiles$median_euler_mean)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2],
   age_beta = summary$coefficients[3,1], age_p = summary$coefficients[3,4], age_se = summary$coefficients[3,2],
   swycbyage_beta = summary$coefficients[5,1], swycbyage_p = summary$coefficients[5,4], swycbyage_se = summary$coefficients[5,2])
})))

swycbyage.qc_reg_table_th$swyc_p_adj <- p.adjust(swycbyage.qc_reg_table_th$swyc_p, method = "BH")
swycbyage.qc_reg_table_th$age_p_adj <- p.adjust(swycbyage.qc_reg_table_th$age_p, method = "BH")
swycbyage.qc_reg_table_th$swycbyage_p_adj <- p.adjust(swycbyage.qc_reg_table_th$swycbyage_p, method = "BH")

print(paste0("Age Main significant results: ", dim(swycbyage.qc_reg_table_th %>% filter(age_p_adj < 0.05))[1]))
print(paste0("Number of significant SWYC x Age IDPs - NO correction: ", dim(swycbyage.qc_reg_table_th %>% filter(swycbyage_p < 0.05))[1]))
print(paste0("Number of significant SWYC x Age IDPs - AFTER FDR correction: ", dim(swycbyage.qc_reg_table_th %>% filter(swycbyage_p_adj < 0.05))[1]))

range(swycbyage.qc_reg_table_th %>% filter(swycbyage_p_adj < 0.05) %>% select(swycbyage_beta))
```

# Test Mediation

## Get all variable names 

## Brain covars with BW
```{r}
lm_swyc_bw_globalcmd <- lm(median_dq ~ birth_weight_kg + global.cmd, data = swyc_avg_centiles)
lm_swyc_bw_thcmd <- lm(median_dq ~ birth_weight_kg + th.cmd, data = swyc_avg_centiles)
lm_swyc_bw_subcort <- lm(median_dq ~ birth_weight_kg + global_SubCortGrayVol, data = swyc_avg_centiles)
lm_swyc_bw_vent <- lm(median_dq ~ birth_weight_kg + global_VentricleChoroidVol, data = swyc_avg_centiles)
lm_swyc_bw_allcmd <- lm(median_dq ~ birth_weight_kg + global.cmd + vol.cmd + sa.cmd + th.cmd, data = swyc_avg_centiles)

lm_swyc_bw_globalcmd_euler <- lm(median_dq ~ birth_weight_kg + global.cmd + median_euler_mean, data = swyc_avg_centiles)
lm_swyc_bw_thcmd_euler <- lm(median_dq ~ birth_weight_kg + th.cmd + euler_mean_mpr, data = swyc_avg_centiles)
lm_swyc_bw_subcort_euler <- lm(median_dq ~ birth_weight_kg + global_SubCortGrayVol + median_euler_mean, data = swyc_avg_centiles)
lm_swyc_bw_vent_euler <- lm(median_dq ~ birth_weight_kg + global_VentricleChoroidVol + median_euler_mean, data = swyc_avg_centiles)
lm_swyc_bw_allcmd_euler <- lm(median_dq ~ birth_weight_kg + + global.cmd + vol.cmd + sa.cmd + th.cmd + median_euler_mean, data = swyc_avg_centiles)
lm_swyc_bw_euler <- lm(median_dq ~ birth_weight_kg + median_euler_mean, data = swyc_avg_centiles)

#plot
mediation_swycbw <- data.frame(
  model = c("no covar", "global.cmd", "th.cmd", "subcort.vol", "vent.vol", "all.cmd", "global.cmd-euler", "th.cmd-euler", "subcort.vol-euler", "vent.vol-euler", "all.cmd-euler", "only euler"),
  estimate = c(summary(lm_swyc_bw)$coefficients[2, "Estimate"],
               summary(lm_swyc_bw_globalcmd)$coefficients[2, "Estimate"],
               summary(lm_swyc_bw_thcmd)$coefficients[2, "Estimate"],
               summary(lm_swyc_bw_subcort)$coefficients[2, "Estimate"],
               summary(lm_swyc_bw_vent)$coefficients[2, "Estimate"],
               summary(lm_swyc_bw_allcmd)$coefficients[2, "Estimate"],
               summary(lm_swyc_bw_globalcmd_euler)$coefficients[2, "Estimate"],
               summary(lm_swyc_bw_thcmd_euler)$coefficients[2, "Estimate"],
               summary(lm_swyc_bw_subcort_euler)$coefficients[2, "Estimate"],
               summary(lm_swyc_bw_vent_euler)$coefficients[2, "Estimate"],
               summary(lm_swyc_bw_allcmd_euler)$coefficients[2, "Estimate"],
               summary(lm_swyc_bw_euler)$coefficients[2, "Estimate"]),
  se = c(summary(lm_swyc_bw)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bw_globalcmd)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bw_thcmd)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bw_subcort)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bw_vent)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bw_allcmd)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bw_globalcmd_euler)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bw_thcmd_euler)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bw_subcort_euler)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bw_vent_euler)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bw_allcmd_euler)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bw_euler)$coefficients[2, "Std. Error"])
) %>%
  mutate(
    ymin = estimate - se,
    ymax = estimate + se
  )

beta_swyc_bw_nocov = mediation_swycbw[mediation_swycbw$model == "no covar", "estimate"]
sehigh_swyc_bw_nocov = mediation_swycbw[mediation_swycbw$model == "no covar", "ymax"]
selow_swyc_bw_nocov = mediation_swycbw[mediation_swycbw$model == "no covar", "ymin"]

ggplot(mediation_swycbw, aes(x = model, y = estimate)) +
  geom_pointrange(aes(ymin = ymin, ymax = ymax), linewidth = 0.8) +
  geom_hline(yintercept = beta_swyc_bw_nocov, linetype = "dashed", color = "#7A7A7A", linewidth = 0.5) +
  geom_hline(yintercept = sehigh_swyc_bw_nocov, linetype = "dashed", color = "#BDBDBD", linewidth = 0.5) +
  geom_hline(yintercept = selow_swyc_bw_nocov, linetype = "dashed", color = "#BDBDBD", linewidth = 0.5) +
  scale_x_discrete(limits = c("no covar", "global.cmd", "th.cmd", "subcort.vol", "vent.vol", "all.cmd", "global.cmd-euler", "th.cmd-euler", "subcort.vol-euler", "vent.vol-euler", "all.cmd-euler", "only euler")) +
  labs(y = "Beta Estimate with Std Error", x = "LM Model covariates", title = "SWYC-DQ ~ Birth Weight") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
```

## Brain covars with BWp
```{r}
lm_swyc_bwp_subcort_euler <- lm(median_dq_zscore ~ bwp_fen +  global_SubCortGrayVol + median_euler_mean, data = swyc_avg_centiles)
lm_swyc_bwp_vent_euler <- lm(median_dq_zscore ~ bwp_fen + global_VentricleChoroidVol + median_euler_mean, data = swyc_avg_centiles)
lm_swyc_bwp_cortex_euler <- lm(median_dq_zscore ~ bwp_fen + global_CortexVol + median_euler_mean, data = swyc_avg_centiles)
lm_swyc_bwp_cbWM_euler <- lm(median_dq_zscore ~ bwp_fen + global_CerebralWhiteMatterVol + median_euler_mean, data = swyc_avg_centiles)

lm_swyc_bwp_allvol_euler <- lm(median_dq_zscore ~ bwp_fen + global_CortexVol + global_CerebralWhiteMatterVol + global_SubCortGrayVol + global_VentricleChoroidVol + median_euler_mean, data = swyc_avg_centiles)

lm_swyc_bwp_allcmd_euler <- lm(median_dq_zscore ~ bwp_fen + + global.cmd + vol.cmd + sa.cmd + th.cmd + subcort.cmd + median_euler_mean, data = swyc_avg_centiles)
lm_swyc_bwp_euler <- lm(median_dq_zscore ~ bwp_fen + median_euler_mean, data = swyc_avg_centiles)

#plot
mediation_swycbwp <- data.frame(
  model = c("no covar", "subcort.vol-euler", "vent.vol-euler", "cortex.vol-euler", "cbWM.vol-euler", "all.vol-euler", "all.cmd-euler", "only euler"),
  estimate = c(summary(lm_swyc_bwp)$coefficients[2, "Estimate"],
               summary(lm_swyc_bwp_subcort_euler)$coefficients[2, "Estimate"],
               summary(lm_swyc_bwp_vent_euler)$coefficients[2, "Estimate"],
               summary(lm_swyc_bwp_cortex_euler)$coefficients[2, "Estimate"],
               summary(lm_swyc_bwp_cbWM_euler)$coefficients[2, "Estimate"],
               summary(lm_swyc_bwp_allvol_euler)$coefficients[2, "Estimate"],
               summary(lm_swyc_bwp_allcmd_euler)$coefficients[2, "Estimate"],
               summary(lm_swyc_bwp_euler)$coefficients[2, "Estimate"]),
  se = c(summary(lm_swyc_bwp)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bwp_subcort_euler)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bwp_vent_euler)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bwp_cortex_euler)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bwp_cbWM_euler)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bwp_allvol_euler)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bwp_allcmd_euler)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bwp_euler)$coefficients[2, "Std. Error"])
) %>%
  mutate(
    ymin = estimate - se,
    ymax = estimate + se
  )

beta_swyc_bwp_nocov = mediation_swycbwp[mediation_swycbwp$model == "no covar", "estimate"]
sehigh_swyc_bwp_nocov = mediation_swycbwp[mediation_swycbwp$model == "no covar", "ymax"]
selow_swyc_bwp_nocov = mediation_swycbwp[mediation_swycbwp$model == "no covar", "ymin"]

ggplot(mediation_swycbwp, aes(x = model, y = estimate)) +
  geom_pointrange(aes(ymin = ymin, ymax = ymax), linewidth = 0.8) +
  geom_hline(yintercept = beta_swyc_bwp_nocov, linetype = "dashed", color = "#7A7A7A", linewidth = 0.5) +
  geom_hline(yintercept = sehigh_swyc_bwp_nocov, linetype = "dashed", color = "#BDBDBD", linewidth = 0.5) +
  geom_hline(yintercept = selow_swyc_bwp_nocov, linetype = "dashed", color = "#BDBDBD", linewidth = 0.5) +
  scale_x_discrete(limits = c("no covar", "subcort.vol-euler", "vent.vol-euler", "cortex.vol-euler", "cbWM.vol-euler", "all.vol-euler", "all.cmd-euler", "only euler")) +
  labs(y = "Beta Estimate with Std Error", x = "LM Model covariates", title = "SWYC-DQ ~ Birth Weight Percentile") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
```

## Brain covars with GA
```{r}
# lm_swyc_ga <- lm(median_dq_zscore ~ gestational_age, data = swyc_avg_centiles)
# lm_swyc_ga_subcort_euler <- lm(median_dq_zscore ~ gestational_age +  global_SubCortGrayVol + median_euler_mean, data = swyc_avg_centiles)
# lm_swyc_ga_vent_euler <- lm(median_dq_zscore ~ gestational_age + global_VentricleChoroidVol + median_euler_mean, data = swyc_avg_centiles)
# lm_swyc_ga_cortex_euler <- lm(median_dq_zscore ~ gestational_age + global_CortexVol + median_euler_mean, data = swyc_avg_centiles)
# lm_swyc_ga_cbWM_euler <- lm(median_dq_zscore ~ gestational_age + global_CerebralWhiteMatterVol + median_euler_mean, data = swyc_avg_centiles)
# 
# lm_swyc_ga_allvol_euler <- lm(median_dq_zscore ~ gestational_age + global_CortexVol + global_CerebralWhiteMatterVol + global_SubCortGrayVol + global_VentricleChoroidVol + median_euler_mean, data = swyc_avg_centiles)
# 
# lm_swyc_ga_allcmd_euler <- lm(median_dq_zscore ~ gestational_age + + global.cmd + vol.cmd + sa.cmd + th.cmd + subcort.cmd + median_euler_mean, data = swyc_avg_centiles)
# lm_swyc_ga_euler <- lm(median_dq_zscore ~ gestational_age + median_euler_mean, data = swyc_avg_centiles)

lm_swyc_ga <- lm(ga_adj_median_dq_zscore ~ gestational_age, data = swyc_avg_centiles)
lm_swyc_ga_subcort_euler <- lm(ga_adj_median_dq_zscore ~ gestational_age +  global_SubCortGrayVol , data = swyc_avg_centiles)
lm_swyc_ga_vent_euler <- lm(ga_adj_median_dq_zscore ~ gestational_age + global_VentricleChoroidVol , data = swyc_avg_centiles)
lm_swyc_ga_cortex_euler <- lm(ga_adj_median_dq_zscore ~ gestational_age + global_CortexVol , data = swyc_avg_centiles)
lm_swyc_ga_cbWM_euler <- lm(ga_adj_median_dq_zscore ~ gestational_age + global_CerebralWhiteMatterVol , data = swyc_avg_centiles)

lm_swyc_ga_allvol_euler <- lm(ga_adj_median_dq_zscore ~ gestational_age + global_CortexVol + global_CerebralWhiteMatterVol + global_SubCortGrayVol + global_VentricleChoroidVol , data = swyc_avg_centiles)

lm_swyc_ga_allcmd_euler <- lm(ga_adj_median_dq_zscore ~ gestational_age + + global.cmd + vol.cmd + sa.cmd + th.cmd + subcort.cmd , data = swyc_avg_centiles)
lm_swyc_ga_euler <- lm(ga_adj_median_dq_zscore ~ gestational_age , data = swyc_avg_centiles)

#plot
mediation_swycga <- data.frame(
  model = c("no covar", "subcort.vol-euler", "vent.vol-euler", "cortex.vol-euler", "cbWM.vol-euler", "all.vol-euler", "all.cmd-euler", "only euler"),
  estimate = c(summary(lm_swyc_ga)$coefficients[2, "Estimate"],
               summary(lm_swyc_ga_subcort_euler)$coefficients[2, "Estimate"],
               summary(lm_swyc_ga_vent_euler)$coefficients[2, "Estimate"],
               summary(lm_swyc_ga_cortex_euler)$coefficients[2, "Estimate"],
               summary(lm_swyc_ga_cbWM_euler)$coefficients[2, "Estimate"],
               summary(lm_swyc_ga_allvol_euler)$coefficients[2, "Estimate"],
               summary(lm_swyc_ga_allcmd_euler)$coefficients[2, "Estimate"],
               summary(lm_swyc_ga_euler)$coefficients[2, "Estimate"]),
  se = c(summary(lm_swyc_ga)$coefficients[2, "Std. Error"],
               summary(lm_swyc_ga_subcort_euler)$coefficients[2, "Std. Error"],
               summary(lm_swyc_ga_vent_euler)$coefficients[2, "Std. Error"],
               summary(lm_swyc_ga_cortex_euler)$coefficients[2, "Std. Error"],
               summary(lm_swyc_ga_cbWM_euler)$coefficients[2, "Std. Error"],
               summary(lm_swyc_ga_allvol_euler)$coefficients[2, "Std. Error"],
               summary(lm_swyc_ga_allcmd_euler)$coefficients[2, "Std. Error"],
               summary(lm_swyc_ga_euler)$coefficients[2, "Std. Error"])
) %>%
  mutate(
    ymin = estimate - se,
    ymax = estimate + se
  )

beta_swyc_ga_nocov = mediation_swycga[mediation_swycga$model == "no covar", "estimate"]
sehigh_swyc_ga_nocov = mediation_swycga[mediation_swycga$model == "no covar", "ymax"]
selow_swyc_ga_nocov = mediation_swycga[mediation_swycga$model == "no covar", "ymin"]

ggplot(mediation_swycga, aes(x = model, y = estimate)) +
  geom_pointrange(aes(ymin = ymin, ymax = ymax), linewidth = 0.8) +
  geom_hline(yintercept = beta_swyc_ga_nocov, linetype = "dashed", color = "#7A7A7A", linewidth = 0.5) +
  geom_hline(yintercept = sehigh_swyc_ga_nocov, linetype = "dashed", color = "#BDBDBD", linewidth = 0.5) +
  geom_hline(yintercept = selow_swyc_ga_nocov, linetype = "dashed", color = "#BDBDBD", linewidth = 0.5) +
  scale_x_discrete(limits = c("no covar", "subcort.vol-euler", "vent.vol-euler", "cortex.vol-euler", "cbWM.vol-euler", "all.vol-euler", "all.cmd-euler", "only euler")) +
  labs(y = "Beta Estimate with Std Error", x = "LM Model covariates", title = "SWYC-DQ ~ Gestational Age") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
```

## Test via bootstrapping (code from Kevin)
He used it to test if the inclusion of the PFN decreased the effect of PRS_1 on p-factor
From rotation with Ran, who used lavaan to do mediation --- Kevin said maybe this is a more appropriate method

### Gestational Age
```{r}
bootga <- bootstrap_model(lm_swyc_ga, iterations = 100)
bootga$model <- "no_covar"

bootga_globalcmd_euler <- bootstrap_model(lm_swyc_ga_globalcmd_euler, iterations = 100)
bootga_globalcmd_euler$model <- "global.cmd_euler"
bootga_thcmd_euler <- bootstrap_model(lm_swyc_ga_thcmd_euler, iterations = 100)
bootga_thcmd_euler$model <- "th.cmd_euler"
bootga_subcort_euler <- bootstrap_model(lm_swyc_ga_subcort_euler, iterations = 100)
bootga_subcort_euler$model <- "subcort_euler"
bootga_vent_euler <- bootstrap_model(lm_swyc_ga_vent_euler, iterations = 100)
bootga_vent_euler$model <- "vent_euler"

t.test(bootga$gestational_age,bootga_thcmd_euler$gestational_age,paired=FALSE)
t.test(bootga$gestational_age,bootga_globalcmd_euler$gestational_age,paired=FALSE)
t.test(bootga$gestational_age,bootga_subcort_euler$gestational_age,paired=FALSE)
t.test(bootga$gestational_age,bootga_vent_euler$gestational_age,paired=FALSE)

models_tested_ga <- list(global.cmd_euler = bootga_globalcmd_euler %>% select(gestational_age, model),
                        th.cmd_euler = bootga_thcmd_euler %>% select(gestational_age, model),
                        subcort_euler = bootga_subcort_euler %>% select(gestational_age, model),
                        vent_euler = bootga_vent_euler %>% select(gestational_age, model))

bootga_plot_df <- do.call(rbind,
  lapply(names(models_tested_ga), function(facet_name) {
    df <- models_tested_ga[[facet_name]]
    df <- rbind(df, bootga %>% select(gestational_age, model))
    df$facet <- facet_name
    df
    }))

ggplot(bootga_plot_df) +
  geom_density(aes(x = gestational_age, fill = model), alpha = 0.5) +
  facet_wrap(~facet) +
  labs(
    x = "Beta Estimates",
    y = "",
    title = "Bootstrap Betas, Gestational Age "
  ) +
  scale_fill_manual(
    name = "Model",
    values = c("no_covar" = "#7A7A7A", 
               "global.cmd_euler" = "#D08C8A",
               "th.cmd_euler" = "#C9A87C",
               "subcort_euler" = "#A6C48A",
               "vent_euler" = "#90A0B5") ) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5),
    legend.position = "top",
    strip.text = element_blank())

bootga_plot_summary_df <- rbind(do.call(rbind, models_tested_ga), bootga %>% select(gestational_age, model)) %>%
  group_by(model) %>%
  summarize(ga_beta_mean = mean(gestational_age),
            ga_lower_CI = t.test(gestational_age)$conf.int[1],
            ga_upper_CI = t.test(gestational_age)$conf.int[2])

ggplot(bootga_plot_summary_df, aes(x = model, y = ga_beta_mean)) +
  geom_pointrange(aes(ymin = ga_lower_CI, ymax = ga_upper_CI), linewidth = 0.8) +
  scale_x_discrete(limits = c("no_covar", "global.cmd_euler", "th.cmd_euler", "subcort_euler", "vent_euler")) +
  labs(y = "Bootstrap Beta Mean with 95% CI", x = "LM Model covariates", title = "SWYC-DQ ~ Gestational Age") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

hist(bootga$gestational_age)
mean(bootga$gestational_age)
median(bootga$gestational_age)

hist(bootga_globalcmd_euler$gestational_age)
mean(bootga_globalcmd_euler$gestational_age)
median(bootga_globalcmd_euler$gestational_age)
shapiro.test(bootga_globalcmd_euler$gestational_age)
```

### Birth Weight
```{r}
bootbw <- bootstrap_model(lm_swyc_bw, iterations = 100)
bootbw$model <- "no_covar"

bootbw_globalcmd_euler <- bootstrap_model(lm_swyc_bw_globalcmd_euler, iterations = 100)
bootbw_globalcmd_euler$model <- "global.cmd_euler"
bootbw_thcmd_euler <- bootstrap_model(lm_swyc_bw_thcmd_euler, iterations = 100)
bootbw_thcmd_euler$model <- "th.cmd_euler"
bootbw_subcort_euler <- bootstrap_model(lm_swyc_bw_subcort_euler, iterations = 100)
bootbw_subcort_euler$model <- "subcort_euler"
bootbw_vent_euler <- bootstrap_model(lm_swyc_bw_vent_euler, iterations = 100)
bootbw_vent_euler$model <- "vent_euler"

t.test(bootbw$birth_weight_kg,bootbw_thcmd_euler$birth_weight_kg,paired=FALSE)
t.test(bootbw$birth_weight_kg,bootbw_globalcmd_euler$birth_weight_kg,paired=FALSE)
t.test(bootbw$birth_weight_kg,bootbw_subcort_euler$birth_weight_kg,paired=FALSE)
t.test(bootbw$birth_weight_kg,bootbw_vent_euler$birth_weight_kg,paired=FALSE)

models_tested_bw <- list(global.cmd_euler = bootbw_globalcmd_euler %>% select(birth_weight_kg, model),
                        th.cmd_euler = bootbw_thcmd_euler %>% select(birth_weight_kg, model),
                        subcort_euler = bootbw_subcort_euler %>% select(birth_weight_kg, model),
                        vent_euler = bootbw_vent_euler %>% select(birth_weight_kg, model))

bootbw_plot_df <- do.call(rbind,
  lapply(names(models_tested_bw), function(facet_name) {
    df <- models_tested_bw[[facet_name]]
    df <- rbind(df, bootbw %>% select(birth_weight_kg, model))
    df$facet <- facet_name
    df
    }))

ggplot(bootbw_plot_df) +
  geom_density(aes(x = birth_weight_kg, fill = model), alpha = 0.5) +
  facet_wrap(~facet) +
  labs(
    x = "Beta Estimates",
    y = "",
    title = "Bootstrap Betas, Birth Weight "
  ) +
  scale_fill_manual(
    name = "Model",
    values = c("no_covar" = "#7A7A7A", 
               "global.cmd_euler" = "#D08C8A",
               "th.cmd_euler" = "#C9A87C",
               "subcort_euler" = "#A6C48A",
               "vent_euler" = "#90A0B5") ) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5),
    legend.position = "top",
    strip.text = element_blank())
```

### Birth Weight Percentile
```{r}
bootbwp <- bootstrap_model(lm_swyc_bwp, iterations = 100)
bootbwp$model <- "no_covar"

bootbwp_globalcmd_euler <- bootstrap_model(lm_swyc_bwp_globalcmd_euler, iterations = 100)
bootbwp_globalcmd_euler$model <- "global.cmd_euler"
bootbwp_thcmd_euler <- bootstrap_model(lm_swyc_bwp_thcmd_euler, iterations = 100)
bootbwp_thcmd_euler$model <- "th.cmd_euler"
bootbwp_subcort_euler <- bootstrap_model(lm_swyc_bwp_subcort_euler, iterations = 100)
bootbwp_subcort_euler$model <- "subcort_euler"
bootbwp_vent_euler <- bootstrap_model(lm_swyc_bwp_vent_euler, iterations = 100)
bootbwp_vent_euler$model <- "vent_euler"

t.test(bootbwp$bwp_fen,bootbwp_thcmd_euler$bwp_fen,paired=FALSE)
t.test(bootbwp$bwp_fen,bootbwp_globalcmd_euler$bwp_fen,paired=FALSE)
t.test(bootbwp$bwp_fen,bootbwp_subcort_euler$bwp_fen,paired=FALSE)
t.test(bootbwp$bwp_fen,bootbwp_vent_euler$bwp_fen,paired=FALSE)

models_tested_bwp <- list(global.cmd_euler = bootbwp_globalcmd_euler %>% select(bwp_fen, model),
                        th.cmd_euler = bootbwp_thcmd_euler %>% select(bwp_fen, model),
                        subcort_euler = bootbwp_subcort_euler %>% select(bwp_fen, model),
                        vent_euler = bootbwp_vent_euler %>% select(bwp_fen, model))

bootbwp_plot_df <- do.call(rbind,
  lapply(names(models_tested_bwp), function(facet_name) {
    df <- models_tested_bwp[[facet_name]]
    df <- rbind(df, bootbwp %>% select(bwp_fen, model))
    df$facet <- facet_name
    df
    }))

ggplot(bootbwp_plot_df) +
  geom_density(aes(x = bwp_fen, fill = model), alpha = 0.5) +
  facet_wrap(~facet) +
  labs(
    x = "Beta Estimates",
    y = "",
    title = "Bootstrap Betas, Birth Weight Percentile"
  ) +
  scale_fill_manual(
    name = "Model",
    values = c("no_covar" = "#7A7A7A", 
               "global.cmd_euler" = "#D08C8A",
               "th.cmd_euler" = "#C9A87C",
               "subcort_euler" = "#A6C48A",
               "vent_euler" = "#90A0B5") ) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5),
    legend.position = "top",
    strip.text = element_blank())
```

## Test via partial R (code in python from Briana, translated to R) 
[need to check code]
```{r}
# y = standardized outcome
y <- scale(df[[feat]])[, 1]  # scale() returns a matrix; [,1] converts to vector

# full model
X_full <- as.data.frame(scale(df[predictors]))
X_full$const <- 1            # equivalent to sm.add_constant
model_full <- lm(y ~ . + 0, data = X_full)  # +0 because const column added manually

# reduced model (exclude target_cov)
reduced_predictors <- setdiff(predictors, target_cov)
X_reduced <- as.data.frame(scale(df[reduced_predictors]))
X_reduced$const <- 1
model_reduced <- lm(y ~ . + 0, data = X_reduced)

# compute partial r^2
model_red <- lm(median_dq ~ global.cmd + median_euler_mean, data = swyc_avg_centiles)
r2_full <- summary(lm_swyc_ga_globalcmd_euler)$r.squared
r2_red  <- summary(model_red)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(lm_swyc_ga_globalcmd_euler)["gestational_age"])
partial_r <- sign_coef * sqrt(abs(partial_r2))

r2_full <- summary(lm_swyc_ga_globalcmd)$r.squared
r2_red  <- summary(lm_swyc_ga)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(lm_swyc_ga_globalcmd_euler)["global.cmd"])
partial_r <- sign_coef * sqrt(abs(partial_r2))
```
## Test mediation per swyc form (not median dq)

### Regressions for Global
```{r}
vars_totest <- c("global_eTIV", "global_SubCortGrayVol","global_CortexVol", "global_CerebralWhiteMatterVol", "global_VentricleChoroidVol","global_SurfaceArea", "global_MeanThickness") 
swyc_forms_totest <- c("Month-6", "Month-9", "Month-15", "Month-18", "Month-24", "Month-30")

grid <- expand.grid(
  pheno = vars_totest,
  swyc_form = swyc_forms_totest,
  ga_adj = c(TRUE, FALSE),
  stringsAsFactors = FALSE
)

df <- swyc_avg_centiles.swyc.all.cent %>% 
  select(pat_id, gestational_age, birth_weight_kg_qc, vars_totest, 
         form_friendly_name, norm_devo_quotient, ga_adj_norm_devo_quotient_round, 
         median_euler_mean, euler_mean_mpr,complete_centile_vars) %>%
  rename(swyc_form = form_friendly_name) %>%
  mutate(swyc_form = recode(swyc_form,
                      "SWYC : 04 Month" = "Month-4",
                      "SWYC : 06 Month" = "Month-6",
                      "SWYC : 09 Month" = "Month-9",
                      "SWYC : 12 Month" = "Month-12",
                      "SWYC : 15 Month" = "Month-15",
                      "SWYC : 18 Month" = "Month-18",
                      "SWYC : 24 Month" = "Month-24",
                      "SWYC : 30 Month" = "Month-30",
                      "SWYC : 36 Month" = "Month-36",
                      "SWYC : 48 Month" = "Month-48",
                      "SWYC : 60 Month" = "Month-60")) %>%
  pivot_longer(cols = c(ga_adj_norm_devo_quotient_round, norm_devo_quotient),
               names_to = "ga_adj",
               values_to = "dq") %>%
  mutate(ga_adj = grepl("ga_adj", ga_adj))
  
results <- do.call(rbind, lapply(seq_len(nrow(grid)), function(i) {
  
  pheno_i <- grid$pheno[i]
  form_i  <- grid$swyc_form[i]
  ga_i    <- grid$ga_adj[i]
  
  df_sub <- df %>%
    dplyr::filter(swyc_form == form_i, ga_adj == ga_i) %>%
    dplyr::select(all_of(c("dq", pheno_i, "median_euler_mean"))) %>%
    na.omit()
  
  if (nrow(df_sub) < 3) {
    return(data.frame(
      pheno = pheno_i,
      swyc_type = form_i,
      ga_adj = ga_i,
      N = nrow(df_sub),
      beta = NA,
      p = NA,
      se = NA
    ))
  }
  
  fit <- lm(reformulate(c(pheno_i, "median_euler_mean"), response = "dq"),
            data = df_sub)
  
  coef_tab <- summary(fit)$coefficients
  
  data.frame(
    pheno = pheno_i,
    swyc_form = form_i,
    ga_adj = ga_i,
    N = nrow(df_sub),
    beta = coef_tab[pheno_i, "Estimate"],
    p = coef_tab[pheno_i, "Pr(>|t|)"],
    se = coef_tab[pheno_i, "Std. Error"],
    p_adj_pheno = p.adjust(coef_tab[pheno_i, "Pr(>|t|)"], method = "bonferroni", n = length(vars_totest)),
    p_adj_phenoform = p.adjust(coef_tab[pheno_i, "Pr(>|t|)"], method = "bonferroni", n = length(vars_totest)*length(swyc_forms_totest))
  )
}))

# arrange by swyc form and add columns for plotting
results <- results %>%
  mutate(swyc_form = factor(swyc_form, levels = c("Month-6", "Month-9", "Month-15", "Month-18", "Month-24", "Month-30"))) %>% arrange(swyc_form) %>%
  mutate(sig = p_adj_phenoform < 0.05,
          color = case_when(
            sig & beta > 0 ~ "#c87a7a",   #  muted red
            sig & beta < 0 ~ "#7293b0",   #  muted blue
            TRUE ~ "gray70"),
         y_labels = paste(swyc_form, N, sep = ", N="))


(swyc_allforms_global_plot <- ggplot(results, aes(x = beta, y = swyc_form, 
                                                  color = ga_adj, alpha = p_adj_pheno <= 0.05)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbarh(aes(xmin = beta - se, xmax = beta + se),
                 height = 0.2,
                 position = position_dodge(width = 0.5)) +
  facet_wrap(~pheno, scales = "free_x") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "#B0B0B0") +
  theme_classic() +
  theme(
        panel.grid = element_blank(),
        legend.position = "bottom",
        legend.title = element_blank(),
        panel.border = element_blank(),
        axis.text.y  = element_text(color = "black"),
        axis.line.x  = element_line(color = "black"),
        axis.line.y  = element_blank(),
        legend.justification = c("left", "top")
    ) +
    xlab("Std Beta DQ ~ Global Centiles, Across Forms > N = 40") +
    ylab(""))


# if(save_figs == TRUE){
# panel_save(swyc_bf_all_plot, "swyc-bf-all", "pdf", plot_output_folder)
# panel_save(swyc_bf_median_plot, "swyc-bf-median", "pdf", plot_output_folder)
# panel_save(swyc_bf_allforms_ga, "swyc-bf-allforms-ga", "pdf", plot_output_folder)
# }
```

### Regressions for CMD
```{r}
vars_totest <- c("global.cmd", "vol.cmd", "sa.cmd", "th.cmd") 
swyc_forms_totest <- c("Month-6", "Month-9", "Month-15", "Month-18", "Month-24", "Month-30")

cmd.grid <- expand.grid(
  pheno = vars_totest,
  swyc_form = swyc_forms_totest,
  ga_adj = c(TRUE, FALSE),
  stringsAsFactors = FALSE
)

df <- swyc_avg_centiles.swyc.all.cent %>% 
  select(pat_id, gestational_age, birth_weight_kg_qc, vars_totest, 
         form_friendly_name, norm_devo_quotient, ga_adj_norm_devo_quotient_round, 
         median_euler_mean, euler_mean_mpr,complete_centile_vars,
         global.cmd, vol.cmd, sa.cmd, th.cmd) %>%
  rename(swyc_form = form_friendly_name) %>%
  mutate(swyc_form = recode(swyc_form,
                      "SWYC : 04 Month" = "Month-4",
                      "SWYC : 06 Month" = "Month-6",
                      "SWYC : 09 Month" = "Month-9",
                      "SWYC : 12 Month" = "Month-12",
                      "SWYC : 15 Month" = "Month-15",
                      "SWYC : 18 Month" = "Month-18",
                      "SWYC : 24 Month" = "Month-24",
                      "SWYC : 30 Month" = "Month-30",
                      "SWYC : 36 Month" = "Month-36",
                      "SWYC : 48 Month" = "Month-48",
                      "SWYC : 60 Month" = "Month-60")) %>%
  pivot_longer(cols = c(ga_adj_norm_devo_quotient_round, norm_devo_quotient),
               names_to = "ga_adj",
               values_to = "dq") %>%
  mutate(ga_adj = grepl("ga_adj", ga_adj))
  
cmd.results <- do.call(rbind, lapply(seq_len(nrow(cmd.grid)), function(i) {
  
  pheno_i <- cmd.grid$pheno[i]
  form_i  <- cmd.grid$swyc_form[i]
  ga_i    <- cmd.grid$ga_adj[i]
  
  df_sub <- df %>%
    dplyr::filter(swyc_form == form_i, ga_adj == ga_i) %>%
    dplyr::select(all_of(c("dq", pheno_i, "median_euler_mean"))) %>%
    na.omit()
  
  if (nrow(df_sub) < 3) {
    return(data.frame(
      pheno = pheno_i,
      swyc_type = form_i,
      ga_adj = ga_i,
      N = nrow(df_sub),
      beta = NA,
      p = NA,
      se = NA
    ))
  }
  
  fit <- lm(reformulate(c(pheno_i, "median_euler_mean"), response = "dq"),
            data = df_sub)
  
  coef_tab <- summary(fit)$coefficients
  
  data.frame(
    pheno = pheno_i,
    swyc_form = form_i,
    ga_adj = ga_i,
    N = nrow(df_sub),
    beta = coef_tab[pheno_i, "Estimate"],
    p = coef_tab[pheno_i, "Pr(>|t|)"],
    se = coef_tab[pheno_i, "Std. Error"],
    p_adj_pheno = p.adjust(coef_tab[pheno_i, "Pr(>|t|)"], method = "bonferroni", n = length(vars_totest)),
    p_adj_phenoform = p.adjust(coef_tab[pheno_i, "Pr(>|t|)"], method = "bonferroni", n = length(vars_totest)*length(swyc_forms_totest))
  )
}))


# arrange by swyc form and add columns for plotting
cmd.results <- cmd.results %>%
  mutate(swyc_form = factor(swyc_form, levels = c("Month-6", "Month-9", "Month-15", "Month-18", "Month-24", "Month-30"))) %>% arrange(swyc_form) %>%
  mutate(sig = p_adj_phenoform < 0.05,
          color = case_when(
            sig & beta > 0 ~ "#c87a7a",   #  muted red
            sig & beta < 0 ~ "#7293b0",   #  muted blue
            TRUE ~ "gray70"),
         y_labels = paste(swyc_form, pheno, ga_adj, N, sep = "_"))


(swyc_allforms_cmd_plot <- ggplot(cmd.results, aes(x = beta, y = swyc_form, 
                                                  color = ga_adj, alpha = p_adj_pheno <= 0.05)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_text(aes(label = N),
          vjust = -0.5) +
  geom_errorbarh(aes(xmin = beta - se, xmax = beta + se),
                 height = 0.2,
                 position = position_dodge(width = 0.5)) +
  facet_wrap(~pheno, scales = "free_x") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "#B0B0B0") +
  scale_alpha_manual(values = c("FALSE" = 0.35, "TRUE" = 0.8),
    labels = c("FALSE" = "nonsig",
               "TRUE" = "p < 0.05"), 
    name = "Bonferroni corrected across phenotypes (N = 4)") +
  scale_color_manual( values = c("FALSE" = "#4F85B5", "TRUE" = "#B34F6B"),
    labels = c("FALSE" = "Original DQ",
               "TRUE" = "GA-ADJ DQ"), name = "DQ CALCULATION") +
  theme_classic() +
  theme(
        panel.grid = element_blank(),
        legend.position = "bottom",
        panel.border = element_blank(),
        axis.text.y  = element_text(color = "black"),
        axis.line.x  = element_line(color = "black"),
        axis.line.y  = element_blank(),
        legend.justification = c("left", "top")
    ) +
    xlab("Std Beta DQ ~ CMDs, Across Forms > N = 40") +
    ylab(""))


# if(save_figs == TRUE){
# panel_save(swyc_bf_all_plot, "swyc-bf-all", "pdf", plot_output_folder)
# panel_save(swyc_bf_median_plot, "swyc-bf-median", "pdf", plot_output_folder)
# panel_save(swyc_bf_allforms_ga, "swyc-bf-allforms-ga", "pdf", plot_output_folder)
# }
```
# Data Availability - All (includes repeated scan measures) (repeat analyses in "temporally restricted" sample)

## First-Pass plots for scan-swyc timing
```{r}
centiles_all$participant_id_modified <- sub("sub-", "", centiles_all$participant_id)
swyc_subset_all <- swyc %>% filter(pat_id %in% centiles_all$participant_id_modified)
swyc_subset_all <- merge(swyc_subset_all, centiles_all %>% select(c("participant_id_modified", "age_at_scan", "gestational_age", "birth_weight_kg")), 
                     by.x = "pat_id", by.y = "participant_id_modified", all.x = TRUE, all.y = FALSE)
swyc_subset_all$age_diff <- swyc_subset_all$age_at_scan - swyc_subset_all$effective_age

if (save_processed == TRUE){
  write.csv(swyc_subset_all, glue("{csv_tables_folder}swyc_overlap_centiles_all_2511_gabwinc.csv"))
}

hist(swyc_subset_all$age_diff)
summary(swyc_subset_all$age_diff)
sum(swyc_subset_all$age_diff <= 0)
(swyc_overlap_plot <- ggplot(swyc_subset_all, aes(y = pat_id, group = pat_id)) +
  geom_point(aes(x = effective_age, color = form_friendly_name), size = 0.2) +
  geom_line(aes(x = effective_age, color = form_friendly_name),size = 0.1) + 
  geom_point(data = swyc_subset_all %>% 
              filter(age_at_scan <= max(effective_age) | age_at_scan >= min(effective_age)), 
            aes(x = age_at_scan), color = "black", size = 0.1, alpha = 0.5) +
  theme(axis.text.y  = element_blank())) 

(swyc_overlap_scanb4swyc_plot <- ggplot(swyc_subset_all %>% filter(age_diff <= 0),  aes(y = pat_id, group = pat_id)) +
  geom_point(aes(x = effective_age, color = form_friendly_name), size = 0.2) +
  geom_line(aes(x = effective_age, color = form_friendly_name),size = 0.1) + 
  geom_point(aes(x = age_at_scan), color = "black", size = 0.1, alpha = 0.5) +
  theme(axis.text.y  = element_blank())) 
```

## Determine availability of SWYC based on scan date

### adjust age at scan and effective age (swyc) for GA and convert to months
```{r}
# adjust effective age and age at scan for prematurity
swyc_subset_all$effective_age_months <- swyc_subset_all$effective_age/30.437
swyc_subset_all$adj_effective_age_months <- ifelse(!is.na(swyc_subset_all$gestational_age),
                                                 swyc_subset_all$effective_age_months + swyc_subset_all$gestational_age/4.35,
                                                 swyc_subset_all$effective_age_months + 40/4.35)

swyc_subset_all$age_at_scan_months <- swyc_subset_all$age_at_scan/30.437
swyc_subset_all$adj_age_at_scan_months <- ifelse(!is.na(swyc_subset_all$gestational_age),
                                                 swyc_subset_all$age_at_scan_months + swyc_subset_all$gestational_age/4.35,
                                                 swyc_subset_all$age_at_scan_months + 40/4.35)

swyc_subset_all$age_diff_months <- swyc_subset_all$age_at_scan_months - swyc_subset_all$adj_effective_age_months
```

### swyc within 10% of scan in either direction
```{r, echo = FALSE}
swyc_subset_all_matched <- swyc_subset_all %>% filter(abs(age_diff_months) < 0.1*age_at_scan_months)
length(unique(swyc_subset_all_matched$pat_id))
dim(swyc_subset_all_matched)

(swyc_all_matched_plot <- ggplot(swyc_subset_all_matched, aes(y = pat_id, group = pat_id)) +
  geom_point(aes(x = adj_effective_age_months, color = form_friendly_name), size = 0.2) +
  geom_line(aes(x = adj_effective_age_months, color = form_friendly_name),size = 0.1) + 
  geom_point(aes(x = adj_age_at_scan_months), color = "black", size = 0.1, alpha = 0.5) +
  theme(axis.text.y  = element_blank(),
        panel.grid = element_blank(),
        panel.background = element_blank()))
```



### scan before SWYC, take average of SWYCs
```{r, echo = FALSE}
swyc_subset_all_matched2 <- swyc_subset_all %>% filter(age_diff_months < 0)
length(unique(swyc_subset_all_matched2$pat_id))
dim(swyc_subset_all_matched2)
dim(swyc_subset_all_matched2[duplicated(swyc_subset_all_matched2) | duplicated(swyc_subset_all_matched2, fromLast = TRUE), ])
swyc_subset_all_matched2 <- swyc_subset_all_matched2[!duplicated(swyc_subset_all_matched2),]
dim(swyc_subset_all_matched2)


(swyc_all_matched2_plot <- ggplot(swyc_subset_all_matched2, aes(y = pat_id, group = pat_id)) +
  geom_point(aes(x = adj_effective_age_months, color = form_friendly_name), size = 0.2) +
  geom_line(aes(x = adj_effective_age_months, color = form_friendly_name),size = 0.1) + 
  geom_point(aes(x = adj_age_at_scan_months), color = "black", size = 0.1, alpha = 0.5) +
  theme(axis.text.y  = element_blank(),
        panel.grid = element_blank(),
        panel.background = element_blank()))

# take average of SWYCs
swyc_subset_matched2_avg <- swyc_subset_all_matched2 %>% 
  group_by(pat_id) %>%
  summarise(
    mean_dq = mean(norm_devo_quotient),
    median_dq = median(norm_devo_quotient)
  )
dim(swyc_subset_matched2_avg)
```

### scan closest to SWYC, remove if > 24 months, choose the earliest scan
```{r, echo = FALSE}
swyc_subset_all_matched3 <- swyc_subset_all %>% 
  group_by(pat_id) %>%
  filter(age_diff_months == min(age_diff_months) & age_diff_months <= 24) %>%
  filter(age_at_scan == min(age_at_scan)) %>%
  ungroup() %>%
  mutate(pat_age = paste0(pat_id, effective_age)) %>%
  all(pat_age, .keep_all = TRUE)

length(unique(swyc_subset_all_matched3$pat_id))
dim(swyc_subset_all_matched3)
dim(swyc_subset_all_matched3[duplicated(swyc_subset_all_matched3) | duplicated(swyc_subset_all_matched3, fromLast = TRUE), ])
swyc_subset_all_matched3 <- swyc_subset_all_matched3[!duplicated(swyc_subset_all_matched3),]
dim(swyc_subset_all_matched3)

(swyc_all_matched3_plot <- ggplot(swyc_subset_all_matched3, aes(y = pat_id, group = pat_id)) +
  geom_point(aes(x = adj_effective_age_months, color = form_friendly_name), size = 0.2) +
  geom_line(aes(x = adj_effective_age_months, color = form_friendly_name),size = 0.1) + 
  geom_point(aes(x = adj_age_at_scan_months), color = "black", size = 0.1, alpha = 0.5) +
  theme(axis.text.y  = element_blank(),
        panel.grid = element_blank(),
        panel.background = element_blank()))
hist(swyc_subset_all_matched3$age_diff_months)
```

# Repeat Analyses with Date-Matched Subset

## Prep Data for Analysis

### Merge swyc and centiles df
```{r}
swyc_subset_matched_merged <- merge(swyc_subset_all_matched3, centiles_all %>% select(all_of(c("participant_id_modified",
                                                                                                   "age_at_scan", "preterm",
                                                                        "birth_weight_kg", "bwp_fen",
                                                                        "sex", "adjusted_age_in_days","median_euler_mean",
                                                                        full_vars))), 
                     by.x = c("pat_id", "age_at_scan"), by.y = c("participant_id_modified", "age_at_scan"), all.x = TRUE, all.y = FALSE)
```

### keep one row per participant id (for now)
```{r}
swyc_subset_matched_merged <- swyc_subset_matched_merged %>%
  all(pat_id, .keep_all = TRUE)
```

### zscore variables of interest
```{r, echo = FALSE}
# zscore centiles and other variables of interest
df_matched.zscore <- swyc_subset_matched_merged %>%
  mutate(across(all_of(c(full_vars, "bwp_fen")), qnorm)) %>%
  mutate(across(all_of(c("gestational_age", "birth_weight_kg", "adjusted_age_in_days", 
                         "median_euler_mean", "norm_devo_quotient")), scale))
# set sex and preterm as factor
df_matched.zscore$sex <- as.factor(df_matched.zscore$sex)
df_matched.zscore$preterm <- as.factor(df_matched.zscore$preterm)
```

### find Inf cells, make a table of N per phenotype, then convert to NaN
```{r, echo = FALSE}
inf_nan_table <- as.data.frame(t(sapply(df_matched.zscore[c(full_vars, "bwp_fen")], function(x){
  c(inf = sum(is.infinite(x)), nan = sum(is.nan(x)))
})))
max(inf_nan_table$inf)
max(inf_nan_table$nan)

inf_nan_table %>% filter(inf > 0)
```

### set Inf cells to NaN
```{r, echo = FALSE}
df_matched.zscore <- df_matched.zscore %>% 
  mutate(across(all_of(c(full_vars, "bwp_fen")), ~ replace(., is.infinite(.), NaN)))
```

## Global variables

### Define global vars of interest
```{r, echo = TRUE}
global_vars_totest <- c("global_eTIV", "global_SubCortGrayVol","global_CortexVol", "global_CerebralWhiteMatterVol", "global_VentricleChoroidVol","global_SurfaceArea", "global_MeanThickness")
```

### SWYC ~ global without QC
```{r, echo = FALSE}
swyc_matched_reg_table_global <- as.data.frame(t(sapply(df_matched.zscore[, which(names(df_matched.zscore) %in% global_vars_totest)], function(x) {
  lm <- lm(x ~ df_matched.zscore$norm_devo_quotient)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2])
})))

swyc_matched_reg_table_global$swyc_p_adj <- p.adjust(swyc_matched_reg_table_global$swyc_p, method = "BH")

swyc_matched_reg_table_global %>% filter(swyc_p_adj < 0.05)
range(swyc_matched_reg_table_global %>% filter(swyc_p_adj < 0.05) %>% select(swyc_beta))
```

### SWYC ~ global with QC
```{r, echo = FALSE}
swyc.qc_reg_table_global <- as.data.frame(t(sapply(df_matched.zscore[, which(names(df_matched.zscore) %in% global_vars_totest)], function(x) {
  lm <- lm(x ~ df_matched.zscore$norm_devo_quotient + df_matched.zscore$median_euler_mean)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2])
})))

swyc.qc_reg_table_global$swyc_p_adj <- p.adjust(swyc.qc_reg_table_global$swyc_p, method = "BH")

swyc.qc_reg_table_global %>% filter(swyc_p_adj < 0.05)
range(swyc.qc_reg_table_global %>% filter(swyc_p_adj < 0.05) %>% select(swyc_beta))
```

### Visualize/Figure Panel without QC
```{r, echo = FALSE}
#  Add phenotype column from rownames
swyc_matched_reg_table_global$phenotype <- rownames(swyc_matched_reg_table_global)

#  Add model identifiers + rename to common column names
swyc_matched_reg_table_global <- swyc_matched_reg_table_global %>%
  mutate(model = "SWYC",
         beta = swyc_beta,
         se   = swyc_se,
         p_adj    = swyc_p_adj) %>%
    mutate(sig = p_adj < 0.05,
          color = case_when(
            sig & beta > 0 ~ "#c87a7a",   #  muted red
            sig & beta < 0 ~ "#7293b0",   #  muted blue
            TRUE ~ "gray70"))

#  Combine and arrange by specific order of global phenotypes
df <- swyc_matched_reg_table_global %>%
  mutate(phenotype = factor(phenotype, levels = c("global_eTIV", "global_CortexVol", "global_CerebralWhiteMatterVol",
                                                  "global_SubCortGrayVol", "global_VentricleChoroidVol",
                                                  "global_SurfaceArea", "global_MeanThickness"))) %>% arrange(phenotype)

#  Labeling for phenotypes on plot
phenotype_labs <- c(
 global_eTIV = "Intracranial Volume",
 global_MeanThickness = "Mean Cortical Thickness",
 global_VentricleChoroidVol = "Ventricular Volume",
 global_SurfaceArea = "Total Surface Area",
 global_CortexVol = "Cortical Gray Matter Volume",
 global_CerebralWhiteMatterVol = "Cerebral White Matter Volume",
 global_SubCortGrayVol = "Subcortical Gray Matter Volume"
)

#  Plot
(global_betas_noqc_plot <- ggplot(df, aes(x = beta, y = phenotype)) +
    geom_point(aes(shape = model, color = color), size = 3) +
    geom_errorbarh(aes(xmin = beta - se, xmax = beta + se, color = color), height = 0.2) +
    scale_shape_manual(values = c("SWYC" = 8),
                       name = "") +
    scale_color_identity() +
    scale_x_continuous(limits = c(-0.25, 0.35)) +
    scale_y_discrete(limits = rev, labels = phenotype_labs) +
    geom_vline(xintercept = 0, linetype = "dotted", color = "darkgray") +
    theme_minimal() +
    theme_bw(base_size = 16) +
    theme(
        panel.grid = element_blank(),
        legend.position = "bottom",
        legend.title = element_blank(),
        panel.border = element_blank(),
        axis.text.y  = element_text(color = "black"),
        axis.line.x  = element_line(color = "black"),
        axis.line.y  = element_blank(),
        legend.justification = c("left", "top")
    ) +
    xlab("Standardized Beta Estimate, without QC covar") +
    ylab(""))
```

### Visualize/Figure Panel with QC
```{r, echo = FALSE}
#  Add phenotype column from rownames
swyc.qc_reg_table_global$phenotype <- rownames(swyc.qc_reg_table_global)

swyc.qc_reg_table_global <- swyc.qc_reg_table_global %>%
  mutate(model = "SWYC",
         beta = swyc_beta,
         se   = swyc_se,
         p_adj    = swyc_p_adj) %>%
    mutate(sig = p_adj < 0.05,
          color = case_when(
            sig & beta > 0 ~ "#c87a7a",   #  muted red
            sig & beta < 0 ~ "#7293b0",   #  muted blue
            TRUE ~ "gray70"))

#  Combine and arrange by specific order of global phenotypes
df <- swyc.qc_reg_table_global %>%
  mutate(phenotype = factor(phenotype, levels = c("global_eTIV", "global_CortexVol", "global_CerebralWhiteMatterVol",
                                                  "global_SubCortGrayVol", "global_VentricleChoroidVol",
                                                  "global_SurfaceArea", "global_MeanThickness"))) %>% arrange(phenotype)

#  Labeling for phenotypes on plot
phenotype_labs <- c(
 global_eTIV = "Intracranial Volume",
 global_MeanThickness = "Mean Cortical Thickness",
 global_VentricleChoroidVol = "Ventricular Volume",
 global_SurfaceArea = "Total Surface Area",
 global_CortexVol = "Cortical Gray Matter Volume",
 global_CerebralWhiteMatterVol = "Cerebral White Matter Volume",
 global_SubCortGrayVol = "Subcortical Gray Matter Volume"
)

#  Plot
(global_betas_withqc_plot <- ggplot(df, aes(x = beta, y = phenotype)) +
    geom_point(aes(shape = model, color = color), size = 3) +
    geom_errorbarh(aes(xmin = beta - se, xmax = beta + se, color = color), height = 0.2) +
    scale_shape_manual(values = c("SWYC" = 8),
                       name = "") +
    scale_color_identity() +
    scale_x_continuous(limits = c(-0.25, 0.35)) +
    scale_y_discrete(limits = rev, labels = phenotype_labs) +
    geom_vline(xintercept = 0, linetype = "dotted", color = "darkgray") +
    theme_minimal() +
    theme_bw(base_size = 16) +
    theme(
        panel.grid = element_blank(),
        legend.position = "bottom",
        legend.title = element_blank(),
        panel.border = element_blank(),
        axis.text.y  = element_text(color = "black"),
        axis.line.x  = element_line(color = "black"),
        axis.line.y  = element_blank(),
        legend.justification = c("left", "top")
    ) +
    xlab("Standardized Beta Estimate, with QC covar") +
    ylab(""))
```

### Individual Scatterplots for SWYC
```{r}
swyc_matched_N <- sum(!is.na(df_matched.zscore$norm_devo_quotient) & !is.na(df_matched.zscore$global_VentricleChoroidVol))

#Ventricles
swyc_matched_vent_beta <- swyc.qc_reg_table_global %>% 
  filter(phenotype == "global_VentricleChoroidVol") %>% 
  select(swyc_beta) %>% pull(swyc_beta)
swyc_matched_vent_p_adj <- swyc.qc_reg_table_global %>% 
  filter(phenotype == "global_VentricleChoroidVol") %>% 
  select(swyc_p_adj) %>% pull(swyc_p_adj)
(swyc_matched_vent_plot <- ggplot(df_matched.zscore, aes(x = norm_devo_quotient, y = global_VentricleChoroidVol)) +
  geom_point(color = "#7293b0", alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE, color = "#7293b0") +
  annotate("label",
           x = 0.5, y = 3.2,
           label = paste0("std beta=", signif(swyc_matched_vent_beta,3), "; p-adj=", signif(swyc_matched_vent_p_adj,3))) +
  labs(
    x = paste0("Median SWYC Score (z-scored), N = ", swyc_matched_N),
    y = "Ventricle Volume Centile (z-scored)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    plot.title = element_blank()
  ))

#Subcortical GMV
swyc_matched_scgmv_beta <- swyc.qc_reg_table_global %>% 
  filter(phenotype == "global_SubCortGrayVol") %>% 
  select(swyc_beta) %>% pull(swyc_beta)
swyc_matched_scgmv_p_adj <- swyc.qc_reg_table_global %>% 
  filter(phenotype == "global_SubCortGrayVol") %>% 
  select(swyc_p_adj) %>% pull(swyc_p_adj)
(swyc_matched_scgmv_plot <- ggplot(df_matched.zscore, aes(x = norm_devo_quotient, y = global_SubCortGrayVol)) +
  geom_point(color = "#c87a7a", alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE, color = "#c87a7a") +
  annotate("label",
           x = 0.5, y = 3.2,
           label = paste0("std beta=", signif(swyc_matched_scgmv_beta,3), "; p-adj=", signif(swyc_matched_scgmv_p_adj,3))) +
  labs(
    x = paste0("Median SWYC Score (z-scored), N = ", swyc_matched_N),
    y = "Subcortical Gray Volume Centile (z-scored)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    plot.title = element_blank()
  ))


#Cortical GMV
swyc_matched_ctxgmv_beta <- swyc.qc_reg_table_global %>% 
  filter(phenotype == "global_CortexVol") %>% 
  select(swyc_beta) %>% pull(swyc_beta)
swyc_matched_ctxgmv_p_adj <- swyc.qc_reg_table_global %>% 
  filter(phenotype == "global_CortexVol") %>% 
  select(swyc_p_adj) %>% pull(swyc_p_adj)
(swyc_matched_ctxgmv_plot <- ggplot(df_matched.zscore, aes(x = norm_devo_quotient, y = global_CortexVol)) +
  geom_point(color = "#c87a7a", alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE, color = "#c87a7a") +
  annotate("label",
           x = 0.5, y = 3.2,
           label = paste0("std beta=", signif(swyc_matched_ctxgmv_beta,3), "; p-adj=", signif(swyc_matched_ctxgmv_p_adj,3))) +
  labs(
    x = paste0("Median SWYC ctxore (z-ctxored), N = ", swyc_matched_N),
    y = "Cortical Gray Volume Centile (z-ctxored)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    plot.title = element_blank()
  ))

#Total Surface Area
swyc_matched_sa_beta <- swyc.qc_reg_table_global %>% 
  filter(phenotype == "global_SurfaceArea") %>% 
  select(swyc_beta) %>% pull(swyc_beta)
swyc_matched_sa_p_adj <- swyc.qc_reg_table_global %>% 
  filter(phenotype == "global_SurfaceArea") %>% 
  select(swyc_p_adj) %>% pull(swyc_p_adj)
(swyc_matched_sa_plot <- ggplot(df_matched.zscore, aes(x = norm_devo_quotient, y = global_SurfaceArea)) +
  geom_point(color = "#c87a7a", alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE, color = "#c87a7a") +
  annotate("label",
           x = 0.5, y = 3.2,
           label = paste0("std beta=", signif(swyc_matched_sa_beta,3), "; p-adj=", signif(swyc_matched_sa_p_adj,3))) +
  labs(
    x = paste0("Median SWYC ctxore (z-scored), N = ", swyc_matched_N),
    y = "Total Surface Area Centile (z-scored)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    plot.title = element_blank()
  ))
```

## CMDs

### Calculate CMDs
```{r}
vol_vars_cortical <- vol_vars[grep("aparc", vol_vars)]
length_names <- length(c(vol_vars_cortical, sa_vars, th_vars))
names.global <- c("global_SubCortGrayVol","global_CortexVol", "global_CerebralWhiteMatterVol", "global_VentricleChoroidVol")
# Ensure all names are in the same order
names_base <- names(df_matched.zscore)[grep("aparc_GrayVol", names(df_matched.zscore))] %>% sub("aparc_GrayVol_","",.) %>%
  gsub("_Z", "", .)
names.vol <- paste0("aparc_GrayVol_",names_base)
names.sa <- paste0("aparc_SurfArea_",names_base)
names.th <- paste0("aparc_ThickAvg_",names_base)
# CMD Calculation
## vol
vol.center <- colMeans(df_matched.zscore %>% select(all_of(names.vol)))
vol.cov_matrix <- cov(df_matched.zscore %>% select(all_of(names.vol)))
vol.md <- sqrt(mahalanobis(df_matched.zscore %>% select(all_of(names.vol)), vol.center, vol.cov_matrix))
df_matched.zscore$vol.cmd <- vol.md
## sa
sa.center <- colMeans(df_matched.zscore %>% select(all_of(names.sa)), na.rm = TRUE)
sa.cov_matrix <- cov(df_matched.zscore %>% select(all_of(names.sa)), use = "complete.obs")
sa.md <- sqrt(mahalanobis(df_matched.zscore %>% select(all_of(names.sa)), sa.center, sa.cov_matrix))
df_matched.zscore$sa.cmd <- sa.md
## th
th.center <- colMeans(df_matched.zscore %>% select(all_of(names.th)), na.rm = TRUE)
th.cov_matrix <- cov(df_matched.zscore %>% select(all_of(names.th)), use = "complete.obs")
th.md <- sqrt(mahalanobis(df_matched.zscore %>% select(all_of(names.th)), th.center, th.cov_matrix))
df_matched.zscore$th.cmd <- th.md
## global
global.center <- colMeans(df_matched.zscore %>% select(all_of(names.global)), na.rm = TRUE)
global.cov_matrix <- cov(df_matched.zscore %>% select(all_of(names.global)), use = "complete.obs")
global.md <- sqrt(mahalanobis(df_matched.zscore %>% select(all_of(names.global)), global.center, global.cov_matrix))
df_matched.zscore$global.cmd <- global.md
```

### Cor.test CMDs
```{r, echo = FALSE}
# Exploratory
cor.test(df_matched.zscore$vol.cmd, df_matched.zscore$gestational_age)
cor.test(df_matched.zscore$vol.cmd, df_matched.zscore$birth_weight_kg)
cor.test(df_matched.zscore$vol.cmd, df_matched.zscore$norm_devo_quotient)

cor.test(df_matched.zscore$sa.cmd, df_matched.zscore$gestational_age)
cor.test(df_matched.zscore$sa.cmd, df_matched.zscore$birth_weight_kg)
cor.test(df_matched.zscore$sa.cmd, df_matched.zscore$norm_devo_quotient)

cor.test(df_matched.zscore$th.cmd, df_matched.zscore$gestational_age)
cor.test(df_matched.zscore$th.cmd, df_matched.zscore$birth_weight_kg)
cor.test(df_matched.zscore$th.cmd, df_matched.zscore$bwp_fen)
cor.test(df_matched.zscore$th.cmd, df_matched.zscore$norm_devo_quotient)

cor.test(df_matched.zscore$global.cmd, df_matched.zscore$gestational_age)
cor.test(df_matched.zscore$global.cmd, df_matched.zscore$birth_weight_kg)
cor.test(df_matched.zscore$global.cmd, df_matched.zscore$bwp_fen)
cor.test(df_matched.zscore$global.cmd, df_matched.zscore$norm_devo_quotient)
```

### CMD plots
```{r}
# Vol CMD
cor_volCMD_swyc.matched <- cor.test(df_matched.zscore$vol.cmd, df_matched.zscore$norm_devo_quotient)
## plot
volCMD_swyc.matched_r <- signif(cor_volCMD_swyc.matched$estimate, 3)
volCMD_swyc.matched_p <- signif(cor_volCMD_swyc.matched$p.value, 3)

(cor_volCMD_swyc.matched_plot <- ggplot(df_matched.zscore, aes(norm_devo_quotient, vol.cmd)) +
  geom_point(color = "#8C7FAF", alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE, color = "#8C7FAF") +
  annotate("label",
           x = 2, y = 10,
           label.size = 0,
           label = paste0("r=",volCMD_swyc.matched_r, "; p=", volCMD_swyc.matched_p)) +
  labs(
    x = paste0("SWYC DQ (age within 10% of scan), N = ", bw_N),
    y = "Volume CMD"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    plot.title = element_blank()
  ))

# SA CMD
cor_saCMD_swyc.matched <- cor.test(df_matched.zscore$sa.cmd, df_matched.zscore$norm_devo_quotient)
## plot
saCMD_swyc.matched_r <- signif(cor_saCMD_swyc.matched$estimate, 3)
saCMD_swyc.matched_p <- signif(cor_saCMD_swyc.matched$p.value, 3)

(cor_saCMD_swyc.matched_plot <- ggplot(df_matched.zscore, aes(norm_devo_quotient, sa.cmd)) +
  geom_point(color = "#8C7FAF", alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE, color = "#8C7FAF") +
  annotate("label",
           x = 2, y = 10,
           label.size = 0,
           label = paste0("r=",saCMD_swyc.matched_r, "; p=", saCMD_swyc.matched_p)) +
  labs(
    x = paste0("SWYC DQ (age within 10% of scan), N = ", bw_N),
    y = "Surface Area CMD"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    plot.title = element_blank()
  ))

# Thickness CMD
cor_thCMD_swyc.matched <- cor.test(df_matched.zscore$th.cmd, df_matched.zscore$norm_devo_quotient)
## plot
thCMD_swyc.matched_r <- signif(cor_thCMD_swyc.matched$estimate, 3)
thCMD_swyc.matched_p <- signif(cor_thCMD_swyc.matched$p.value, 3)

(cor_thCMD_swyc.matched_plot <- ggplot(df_matched.zscore, aes(norm_devo_quotient, th.cmd)) +
  geom_point(color = "#8C7FAF", alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE, color = "#8C7FAF") +
  annotate("label",
           x = 2, y = 10,
           label.size = 0,
           label = paste0("r=",thCMD_swyc.matched_r, "; p=", thCMD_swyc.matched_p)) +
  labs(
    x = paste0("SWYC DQ (age within 10% of scan), N = ", bw_N),
    y = "Cortical Thickness CMD"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    plot.title = element_blank()
  ))

#Global CMD
cor_globalCMD_swyc.matched <- cor.test(df_matched.zscore$global.cmd, df_matched.zscore$norm_devo_quotient)
## plot
globalCMD_swyc.matched_r <- signif(cor_globalCMD_swyc.matched$estimate, 3)
globalCMD_swyc.matched_p <- signif(cor_globalCMD_swyc.matched$p.value, 3)

(cor_globalCMD_swyc.matched_plot <- ggplot(df_matched.zscore, aes(norm_devo_quotient, global.cmd)) +
  geom_point(color = "#8C7FAF", alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE, color = "#8C7FAF") +
  annotate("label",
           x = 2, y = 10,
           label.size = 0,
           label = paste0("r=",globalCMD_swyc.matched_r, "; p=", globalCMD_swyc.matched_p)) +
  labs(
    x = paste0("SWYC DQ (age within 10% of scan), N = ", bw_N),
    y = "Global CMD"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    plot.title = element_blank()
  ))
```

# Regional Variables

## Set limits for ggseg visualization
```{r}
limits <- c(-0.3, 0.3)
low_color = "#4878CF"
mid_color = "white"
high_color = "#D65F5F"
```

## Volume

### simple linear regressions

#### compute regressions
```{r, echo = FALSE}
# Regional -- including cortical and subcortical
swyc_matched_vol <- as.data.frame(t(sapply(df_matched.zscore[, which(names(df_matched.zscore) %in% vol_vars)], function(x) {
  lm <- lm(x ~ norm_devo_quotient, data = df_matched.zscore)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2])
})))

swyc_matched_vol$swyc_p_adj <- p.adjust(swyc_matched_vol$swyc_p, method = "BH")
print(paste0("Number of significant IDPs after FDR correction: ", dim(swyc_matched_vol %>% filter(swyc_p_adj < 0.05))[1]))
```

#### add label column to match to ggseg
```{r, echo = FALSE}
cols_edit <- grepl("aseg(?!_CC)", rownames(swyc_matched_vol), perl = TRUE)

swyc_matched_vol$label <- rownames(swyc_matched_vol) %>% 
  sub("aparc_GrayVol_", "", .) %>% 
  sub("aseg_", "", .) %>%
  sub("X", "x", .) %>%
  sub("rd_Ventricle", "rd_ventricle", .) %>% 
  sub("th_Ventricle", "th_ventricle", .) %>% 
  sub("Thalamus", "Thalamus-Proper", .)

swyc_matched_vol$label[cols_edit] <- gsub("_", "-", swyc_matched_vol$label[cols_edit])

swyc_matched_vol_sig_MC <- swyc_matched_vol %>% filter(swyc_p_adj < 0.05)

print(paste0("Beta range of significant IDPs: ", range(swyc_matched_vol_sig_MC$swyc_beta)))
```

#### save betas
```{r, echo = FALSE}
save_table_full <- swyc_matched_vol
name <- "swyc_matched_median_volume.csv"
# Save dkt cortical betas for spin test
if(save_tables == TRUE){
  output_folder <-dkt_folder
  save_table_full$hemisphere <- ifelse(grepl("lh", save_table_full$label), "L", 
                         ifelse(grepl("rh", save_table_full$label), "R", NA))
  save_table_full$label <- sub("^rh_|lh_", "", save_table_full$label)
  filename_all <- paste0(output_folder, name)
  write.csv(save_table_full, filename_all)
}
# Save full betas
if(save_tables == TRUE){
output_folder <- betas_folder
filename <- paste0(output_folder, name)
write.csv(save_table_full, filename)
}
```

#### visualization of betas on surface maps
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(swyc_matched_vol_sig_MC, ggseg::dk, by = "label")
plot_data_aseg <- merge(swyc_matched_vol_sig_MC, ggseg::aseg, by = "label")

# Regions not matched in atlases
(swyc_matched_vol_noplot <- swyc_matched_vol_sig_MC[!(swyc_matched_vol_sig_MC$label %in% plot_data_dkt$label) & !(swyc_matched_vol_sig_MC$label %in% plot_data_aseg$label),])

# plots
(swyc_matched_dkt_vol_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = swyc_beta)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = limits) +
  theme_void()+
  labs(title = "SWYC Volume"))

(swyc_matched_aseg_vol_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = swyc_beta)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = limits) +
  theme_void() +
  labs(title = "SWYC Volume"))

if(save_figs == TRUE){
panel_save(swyc_matched_dkt_vol_plot, "swyc_matched_median-dkt-vol", "pdf", plot_output_folder)
panel_save(swyc_matched_aseg_vol_plot , "swyc_matched_median-aseg-vol", "pdf", plot_output_folder)
}
```

#### Scatterplot for top betas
```{r, echo = FALSE}
if (dim(swyc_matched_vol_sig_MC)[1] > 0){
top_n = if(dim(swyc_matched_vol_sig_MC)[1] > 10) {10} else {dim(swyc_matched_vol_sig_MC)[1]}
top <- swyc_matched_vol_sig_MC[order(swyc_matched_vol_sig_MC$swyc_beta, decreasing = TRUE)[1:top_n],]
scatterplots_swyc_matched_vol <- vector(mode = "list", length = top_n)
for(i in 1:top_n){
  pheno <- rownames(top)[i]
  title <- top$label[i]
  names(scatterplots_swyc_matched_vol[i]) <- pheno
  scatterplots_swyc_matched_vol[[i]] <- ggplot(df_matched.zscore, aes(x = norm_devo_quotient, y = !!sym(pheno))) +
    geom_point() +
    geom_smooth(method = "lm") +
    theme_minimal() +
    labs(x = "SWYC (z-scored)", y = "Centile (z-scored)", title = title)
}

nCol <- 2
nRow <- 2
breaks <- c(seq(from = 0, to = top_n , by = (nCol*nRow)), top_n) %>% unique(.)
for(k in 1:(length(breaks)-1)){
  start <- breaks[k]+1
  end <- breaks[k+1]
  do.call("grid.arrange", c(scatterplots_swyc_matched_vol[start:end], ncol=nCol))
}
}
```

### SWYC X age interaction

#### Regression and results
```{r, echo = FALSE}
swycbyage.qc_reg_table_vol <- as.data.frame(t(sapply(df_matched.zscore[, which(names(df_matched.zscore) %in% vol_vars)], function(x) {
  lm <- lm(x ~ df_matched.zscore$norm_devo_quotient*df_matched.zscore$adjusted_age_in_days + df_matched.zscore$median_euler_mean)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2],
   age_beta = summary$coefficients[3,1], age_p = summary$coefficients[3,4], age_se = summary$coefficients[3,2],
   swycbyage_beta = summary$coefficients[5,1], swycbyage_p = summary$coefficients[5,4], swycbyage_se = summary$coefficients[5,2])
})))

swycbyage.qc_reg_table_vol$swyc_p_adj <- p.adjust(swycbyage.qc_reg_table_vol$swyc_p, method = "BH")
swycbyage.qc_reg_table_vol$age_p_adj <- p.adjust(swycbyage.qc_reg_table_vol$age_p, method = "BH")
swycbyage.qc_reg_table_vol$swycbyage_p_adj <- p.adjust(swycbyage.qc_reg_table_vol$swycbyage_p, method = "BH")

print(paste0("Age Main significant results: ", dim(swycbyage.qc_reg_table_vol %>% filter(age_p_adj < 0.05))[1]))
print(paste0("Number of significant SWYC x Age IDPs - NO correction: ", dim(swycbyage.qc_reg_table_vol %>% filter(swycbyage_p < 0.05))[1]))
print(paste0("Number of significant SWYC x Age IDPs - AFTER FDR correction: ", dim(swycbyage.qc_reg_table_vol %>% filter(swycbyage_p_adj < 0.05))[1]))

range(swycbyage.qc_reg_table_vol %>% filter(swycbyage_p_adj < 0.05) %>% select(swycbyage_beta))

if(dim(swycbyage.qc_reg_table_vol %>% filter(age_p_adj < 0.05))[1] > 0){
  age_reg_table_vol <- as.data.frame(t(sapply(df_matched.zscore[, which(names(df_matched.zscore) %in% vol_vars)], function(x) {
  lm <- lm(x ~ df_matched.zscore$adjusted_age_in_days)
  summary <- summary(lm)
  c(age_beta = summary$coefficients[2,1], age_p = summary$coefficients[2,4], age_se = summary$coefficients[2,2])
})))
  age_reg_table_vol$age_p_adj <- p.adjust(age_reg_table_vol$age_p, method = "BH")
  print(paste0("Age Main significant results (no other covariates): ", dim(age_reg_table_vol %>% filter(age_p_adj < 0.05))[1]))
  
  
  full_age_reg_table_vol <- as.data.frame(t(sapply(df_matched.zscore_fullcentall[, which(names(df_matched.zscore_fullcentall) %in% vol_vars)], function(x) {
  lm <- lm(x ~ df_matched.zscore_fullcentall$adjusted_age_in_days)
  summary <- summary(lm)
  c(age_beta = summary$coefficients[2,1], age_p = summary$coefficients[2,4], age_se = summary$coefficients[2,2])
})))
  full_age_reg_table_vol$age_p_adj <- p.adjust(full_age_reg_table_vol$age_p, method = "BH")
  print(paste0("Age Main significant results full all sample (no other covariates): ", dim(full_age_reg_table_vol %>% filter(age_p_adj < 0.05))[1]))
}
```

## Surface Area

###  simple linear regressions

#### compute regressions
```{r, echo = FALSE}
sum(!is.na(df_matched.zscore$norm_devo_quotient))
# Regional -- including cortical and subcortical
swyc_matched_sa <- as.data.frame(t(sapply(df_matched.zscore[, which(names(df_matched.zscore) %in% sa_vars)], function(x) {
  lm <- lm(x ~ norm_devo_quotient, data = df_matched.zscore)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2])
})))

swyc_matched_sa$swyc_p_adj <- p.adjust(swyc_matched_sa$swyc_p, method = "BH")
print(paste0("Number of significant IDPs after FDR correction: ", dim(swyc_matched_sa %>% filter(swyc_p_adj < 0.05))[1]))
```

#### add label column to match to ggseg
```{r, echo = FALSE}
swyc_matched_sa$label <- rownames(swyc_matched_sa) %>% 
  sub("aparc_SurfArea_", "", .)

swyc_matched_sa_sig_MC <- swyc_matched_sa %>% filter(swyc_p_adj < 0.05)

print(paste0("Beta range of significant IDPs: ", range(swyc_matched_sa_sig_MC$swyc_beta)))
```

#### save betas
```{r, echo = FALSE}
save_table_full <- swyc_matched_sa
name <- "swyc_matched_median_surfArea.csv"
# Save dkt cortical betas for spin test
if(save_tables == TRUE){
  output_folder <-dkt_folder
  save_table_full$hemisphere <- ifelse(grepl("lh", save_table_full$label), "L", 
                         ifelse(grepl("rh", save_table_full$label), "R", NA))
  save_table_full$label <- sub("^rh_|lh_", "", save_table_full$label)
  filename_all <- paste0(output_folder, name)
  write.csv(save_table_full, filename_all)
}
# Save full betas
if(save_tables == TRUE){
output_folder <- betas_folder
filename <- paste0(output_folder, name)
write.csv(save_table_full, filename)
}
```

#### visualization of betas on surface maps
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(swyc_matched_sa_sig_MC, ggseg::dk, by = "label")

# Regions not matched in atlases
(swyc_matched_sa_noplot <- swyc_matched_sa_sig_MC[!(swyc_matched_sa_sig_MC$label %in% plot_data_dkt$label) & !(swyc_matched_sa_sig_MC$label %in% plot_data_aseg$label),])

# plots
(swyc_matched_dkt_sa_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = swyc_beta)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = limits) +
  theme_void()+
  labs(title = "SWYC Surface Area"))

if(save_figs == TRUE){
panel_save(swyc_matched_dkt_sa_plot, "swyc_matched_median-dkt-sa", "pdf", plot_output_folder)
}
```

#### Scatterplot for top betas
```{r, echo = FALSE}
if (dim(swyc_matched_sa_sig_MC)[1] > 0){
top_n = if(dim(swyc_matched_sa_sig_MC)[1] > 10) {10} else {dim(swyc_matched_sa_sig_MC)[1]}
top <- swyc_matched_sa_sig_MC[order(swyc_matched_sa_sig_MC$swyc_beta, decreasing = TRUE)[1:top_n],]
scatterplots_swyc_matched_sa <- vector(mode = "list", length = top_n)
for(i in 1:top_n){
  pheno <- rownames(top)[i]
  title <- top$label[i]
  names(scatterplots_swyc_matched_sa[i]) <- pheno
  scatterplots_swyc_matched_sa[[i]] <- ggplot(df_matched.zscore, aes(x = norm_devo_quotient, y = !!sym(pheno))) +
    geom_point() +
    geom_smooth(method = "lm") +
    theme_minimal() +
    labs(x = "SWYC (z-scored)", y = "Centile (z-scored)", title = title)
}

nCol <- 2
nRow <- 2
breaks <- c(seq(from = 0, to = top_n , by = (nCol*nRow)), top_n) %>% unique(.)
for(k in 1:(length(breaks)-1)){
  start <- breaks[k]+1
  end <- breaks[k+1]
  do.call("grid.arrange", c(scatterplots_swyc_matched_sa[start:end], ncol=nCol))
}
}
```

### SWYC X age interaction

#### Regression and results
```{r, echo = FALSE}
swycbyage.qc_reg_table_sa <- as.data.frame(t(sapply(df_matched.zscore[, which(names(df_matched.zscore) %in% sa_vars)], function(x) {
  lm <- lm(x ~ df_matched.zscore$norm_devo_quotient*df_matched.zscore$adjusted_age_in_days + df_matched.zscore$median_euler_mean)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2],
   age_beta = summary$coefficients[3,1], age_p = summary$coefficients[3,4], age_se = summary$coefficients[3,2],
   swycbyage_beta = summary$coefficients[5,1], swycbyage_p = summary$coefficients[5,4], swycbyage_se = summary$coefficients[5,2])
})))

swycbyage.qc_reg_table_sa$swyc_p_adj <- p.adjust(swycbyage.qc_reg_table_sa$swyc_p, method = "BH")
swycbyage.qc_reg_table_sa$age_p_adj <- p.adjust(swycbyage.qc_reg_table_sa$age_p, method = "BH")
swycbyage.qc_reg_table_sa$swycbyage_p_adj <- p.adjust(swycbyage.qc_reg_table_sa$swycbyage_p, method = "BH")

print(paste0("Age Main significant results: ", dim(swycbyage.qc_reg_table_sa %>% filter(age_p_adj < 0.05))[1]))
print(paste0("Number of significant SWYC x Age IDPs - NO correction: ", dim(swycbyage.qc_reg_table_sa %>% filter(swycbyage_p < 0.05))[1]))
print(paste0("Number of significant SWYC x Age IDPs - AFTER FDR correction: ", dim(swycbyage.qc_reg_table_sa %>% filter(swycbyage_p_adj < 0.05))[1]))

range(swycbyage.qc_reg_table_sa %>% filter(swycbyage_p_adj < 0.05) %>% select(swycbyage_beta))

if(dim(swycbyage.qc_reg_table_sa %>% filter(age_p_adj < 0.05))[1] > 0){
  age_reg_table_sa <- as.data.frame(t(sapply(df_matched.zscore[, which(names(df_matched.zscore) %in% sa_vars)], function(x) {
  lm <- lm(x ~ df_matched.zscore$adjusted_age_in_days)
  summary <- summary(lm)
  c(age_beta = summary$coefficients[2,1], age_p = summary$coefficients[2,4], age_se = summary$coefficients[2,2])
})))
  age_reg_table_sa$age_p_adj <- p.adjust(age_reg_table_sa$age_p, method = "BH")
  print(paste0("Age Main significant results (no other covariates): ", dim(age_reg_table_sa %>% filter(age_p_adj < 0.05))[1]))
  
  
  full_age_reg_table_sa <- as.data.frame(t(sapply(df_matched.zscore_fullcentall[, which(names(df_matched.zscore_fullcentall) %in% sa_vars)], function(x) {
  lm <- lm(x ~ df_matched.zscore_fullcentall$adjusted_age_in_days)
  summary <- summary(lm)
  c(age_beta = summary$coefficients[2,1], age_p = summary$coefficients[2,4], age_se = summary$coefficients[2,2])
})))
  full_age_reg_table_sa$age_p_adj <- p.adjust(full_age_reg_table_sa$age_p, method = "BH")
  print(paste0("Age Main significant results full all sample (no other covariates): ", dim(full_age_reg_table_sa %>% filter(age_p_adj < 0.05))[1]))
}
```

## Thickness

###  simple linear regressions

#### compute regressions
```{r, echo = FALSE}
sum(!is.na(df_matched.zscore$aparc_ThickAvg_lh_bankssts))
# Regional -- including cortical and subcortical
swyc_matched_th <- as.data.frame(t(sapply(df_matched.zscore[, which(names(df_matched.zscore) %in% th_vars)], function(x) {
  lm <- lm(x ~ norm_devo_quotient, data = df_matched.zscore)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2])
})))

swyc_matched_th$swyc_p_adj <- p.adjust(swyc_matched_th$swyc_p, method = "BH")
print(paste0("Number of significant IDPs after FDR correction: ", dim(swyc_matched_th %>% filter(swyc_p_adj < 0.05))[1]))
```

#### add label column to match to ggseg
```{r, echo = FALSE}
swyc_matched_th$label <- rownames(swyc_matched_th) %>% 
  sub("aparc_ThickAvg_", "", .)

swyc_matched_th_sig_MC <- swyc_matched_th %>% filter(swyc_p_adj < 0.05)

print(paste0("Beta range of significant IDPs: ", range(swyc_matched_th_sig_MC$swyc_beta)))
```

#### save betas
```{r, echo = FALSE}
save_table_full <- swyc_matched_th
name <- "swyc_matched_thick.csv"
# Save dkt cortical betas for spin test
if(save_tables == TRUE){
  output_folder <-dkt_folder
  save_table_full$hemisphere <- ifelse(grepl("lh", save_table_full$label), "L", 
                         ifelse(grepl("rh", save_table_full$label), "R", NA))
  save_table_full$label <- sub("^rh_|lh_", "", save_table_full$label)
  filename_all <- paste0(output_folder, name)
  write.csv(save_table_full, filename_all)
}
# Save full betas
if(save_tables == TRUE){
output_folder <- betas_folder
filename <- paste0(output_folder, name)
write.csv(save_table_full, filename)
}
```

#### visualization of betas on surface maps
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(swyc_matched_th_sig_MC, ggseg::dk, by = "label")

# Regions not matched in atlases
(swyc_matched_th_noplot <- swyc_matched_th_sig_MC[!(swyc_matched_th_sig_MC$label %in% plot_data_dkt$label) & !(swyc_matched_th_sig_MC$label %in% plot_data_aseg$label),])

# plots
(swyc_matched_dkt_th_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = swyc_beta)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = limits) +
  theme_void()+
  labs(title = "SWYC Cortical Thickness"))

if(save_figs == TRUE){
panel_save(swyc_matched_dkt_th_plot, "swyc_matched_median-dkt-th", "pdf", plot_output_folder)
}
```

#### Scatterplot for top betas
```{r, echo = FALSE}
if (dim(swyc_matched_th_sig_MC)[1] > 0){
top_n = if(dim(swyc_matched_th_sig_MC)[1] > 10) {10} else {dim(swyc_matched_th_sig_MC)[1]}
top <- swyc_matched_th_sig_MC[order(abs(swyc_matched_th_sig_MC$swyc_beta), decreasing = TRUE)[1:top_n],]
scatterplots_swyc_matched_th <- vector(mode = "list", length = top_n)
for(i in 1:top_n){
  print(i)
  pheno <- rownames(top)[i]
  print(pheno)
  title <- top$label[i]
  names(scatterplots_swyc_matched_th)[i] <- pheno
  scatterplots_swyc_matched_th[[i]] <- ggplot(df_matched.zscore, aes(x = norm_devo_quotient, y = !!sym(pheno))) +
    geom_point() +
    geom_smooth(method = "lm") +
    theme_minimal() +
    labs(x = "SWYC (z-scored)", y = "Centile (z-scored)", title = title)
}

nCol <- 2
nRow <- 2
breaks <- c(seq(from = 0, to = top_n , by = (nCol*nRow)), top_n) %>% unique(.)
for(k in 1:(length(breaks)-1)){
  start <- breaks[k]+1
  end <- breaks[k+1]
  do.call("grid.arrange", c(scatterplots_swyc_matched_th[start:end], ncol=nCol))
}
}
```

### SWYC X age interaction

#### Regression and results
```{r, echo = FALSE}
swycbyage.qc_reg_table_th <- as.data.frame(t(sapply(df_matched.zscore[, which(names(df_matched.zscore) %in% th_vars)], function(x) {
  lm <- lm(x ~ df_matched.zscore$norm_devo_quotient*df_matched.zscore$adjusted_age_in_days + df_matched.zscore$median_euler_mean)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2],
   age_beta = summary$coefficients[3,1], age_p = summary$coefficients[3,4], age_se = summary$coefficients[3,2],
   swycbyage_beta = summary$coefficients[5,1], swycbyage_p = summary$coefficients[5,4], swycbyage_se = summary$coefficients[5,2])
})))

swycbyage.qc_reg_table_th$swyc_p_adj <- p.adjust(swycbyage.qc_reg_table_th$swyc_p, method = "BH")
swycbyage.qc_reg_table_th$age_p_adj <- p.adjust(swycbyage.qc_reg_table_th$age_p, method = "BH")
swycbyage.qc_reg_table_th$swycbyage_p_adj <- p.adjust(swycbyage.qc_reg_table_th$swycbyage_p, method = "BH")

print(paste0("Age Main significant results: ", dim(swycbyage.qc_reg_table_th %>% filter(age_p_adj < 0.05))[1]))
print(paste0("Number of significant SWYC x Age IDPs - NO correction: ", dim(swycbyage.qc_reg_table_th %>% filter(swycbyage_p < 0.05))[1]))
print(paste0("Number of significant SWYC x Age IDPs - AFTER FDR correction: ", dim(swycbyage.qc_reg_table_th %>% filter(swycbyage_p_adj < 0.05))[1]))

range(swycbyage.qc_reg_table_th %>% filter(swycbyage_p_adj < 0.05) %>% select(swycbyage_beta))
```

## Test Mediation

### Brain covars with bw
```{r}
sum(!is.na(df_matched.zscore$birth_weight_kg))

lm_swyc_matched_bw_matched <- lm(norm_devo_quotient ~ birth_weight_kg, data = df_matched.zscore)
lm_swyc_matched_bw_matched_stats <- summary(lm_swyc_matched_bw_matched)
lm_swyc_matched_bw_thcmd <- lm(norm_devo_quotient ~ birth_weight_kg + th.cmd, data = df_matched.zscore)
lm_swyc_matched_bw_ctxvol <- lm(norm_devo_quotient ~ birth_weight_kg + global_CortexVol, data = df_matched.zscore)
lm_swyc_matched_bw_thcmd_stats <- summary(lm_swyc_matched_bw_thcmd)
lm_swyc_matched_bw_ctxvol_stats <- summary(lm_swyc_matched_bw_ctxvol)
lm_swyc_matched_bw_matched_stats
lm_swyc_matched_bw_thcmd_stats
lm_swyc_matched_bw_ctxvol_stats

#plot

df_plot_bw <- data.frame(
  model = c("no covar BW", "thickness cmd covar", "ctxvol covar"),
  estimate = c(summary(lm_swyc_matched_bw_matched)$coefficients[2, "Estimate"],
               summary(lm_swyc_matched_bw_thcmd)$coefficients[2, "Estimate"],
               summary(lm_swyc_matched_bw_ctxvol)$coefficients[2, "Estimate"]),
  se = c(summary(lm_swyc_matched_bw_matched)$coefficients[2, "Std. Error"],
         summary(lm_swyc_matched_bw_thcmd)$coefficients[2, "Std. Error"],
         summary(lm_swyc_matched_bw_ctxvol)$coefficients[2, "Std. Error"])
) %>%
  mutate(
    ymin = estimate - se,
    ymax = estimate + se
  )

ggplot(df_plot_bw, aes(x = model, y = estimate)) +
  geom_pointrange(aes(ymin = ymin, ymax = ymax), linewidth = 0.8) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_classic()

```

### Brain covars with bw
```{r}
sum(!is.na(df_matched.zscore$gestational_age))

lm_swyc_matched_ga_matched <- lm(norm_devo_quotient ~ gestational_age, data = df_matched.zscore)
lm_swyc_matched_ga_matched_stats <- summary(lm_swyc_matched_ga_matched)
lm_swyc_matched_ga_thcmd <- lm(norm_devo_quotient ~ gestational_age + th.cmd, data = df_matched.zscore)
lm_swyc_matched_ga_ctxvol <- lm(norm_devo_quotient ~ gestational_age + global_CortexVol, data = df_matched.zscore)
lm_swyc_matched_ga_thcmd_stats <- summary(lm_swyc_matched_ga_thcmd)
lm_swyc_matched_ga_ctxvol_stats <- summary(lm_swyc_matched_ga_ctxvol)
lm_swyc_matched_ga_matched_stats
lm_swyc_matched_ga_thcmd_stats
lm_swyc_matched_ga_ctxvol_stats

#plot

df_plot_ga <- data.frame(
  model = c("no covar ga", "thickness cmd covar", "ctxvol covar"),
  estimate = c(summary(lm_swyc_matched_ga_matched)$coefficients[2, "Estimate"],
               summary(lm_swyc_matched_ga_thcmd)$coefficients[2, "Estimate"],
               summary(lm_swyc_matched_ga_ctxvol)$coefficients[2, "Estimate"]),
  se = c(summary(lm_swyc_matched_ga_matched)$coefficients[2, "Std. Error"],
         summary(lm_swyc_matched_ga_thcmd)$coefficients[2, "Std. Error"],
         summary(lm_swyc_matched_ga_ctxvol)$coefficients[2, "Std. Error"])
) %>%
  mutate(
    ymin = estimate - se,
    ymax = estimate + se
  )

ggplot(df_plot_ga, aes(x = model, y = estimate)) +
  geom_pointrange(aes(ymin = ymin, ymax = ymax), linewidth = 0.8) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_classic()

```


# LMSz Centiles

## Prep SWYC-LMSz centiles data

### merge and take z-scores of centiles
```{r}
lmsz_centiles$participant_id_modified <- sub("sub-", "", lmsz_centiles$participant_id)

swyc_lmsz <- merge(swyc_subset_avg %>% select(all_of(c("pat_id", "mean_dq", "median_dq", "preterm",
                                                   "birth_weight_kg","gestational_age", "bwp_fen",
                                                   "sex", "adjusted_age_in_days", "median_euler_mean"))),
                   lmsz_centiles, by.x = c("pat_id", "sex"), by.y = c("participant_id_modified", "sex"), all.x = FALSE, all.y = FALSE)
#swyc_lmsz <- merge(swyc_lmsz, swyc_avg_centiles %>% select(pat_id, grep("cmd", names(.))), 
                   #by = "pat_id", all.x = TRUE, all.y = FALSE)

table(swyc_lmsz$preterm)
# zscore centiles and other variables of interest
lmsz_vars <- names(swyc_lmsz)[grep(".lmsz", names(swyc_lmsz))]
swyc_avg_centiles_lmsz <- swyc_lmsz %>%
  mutate(across(all_of(c(full_vars, lmsz_vars, "bwp_fen")), qnorm)) %>%
  mutate(across(all_of(c("gestational_age", "birth_weight_kg", 
                         #"adjusted_age_in_days", 
                         "age_at_scan",
                         "median_euler_mean", "mean_dq", "median_dq")), scale))
# set sex and preterm as factor
swyc_avg_centiles_lmsz$sex <- as.factor(swyc_avg_centiles_lmsz$sex)
swyc_avg_centiles_lmsz$preterm <- as.factor(swyc_avg_centiles_lmsz$preterm)
```

### find Inf cells, make a table of N per phenotype, then convert to NaN
```{r, echo = FALSE}
inf_nan_table <- as.data.frame(t(sapply(swyc_avg_centiles_lmsz[c(full_vars, "bwp_fen")], function(x){
  c(inf = sum(is.infinite(x)), nan = sum(is.nan(x)))
})))
max(inf_nan_table$inf)
max(inf_nan_table$nan)

inf_nan_table %>% filter(inf > 0)
```

## Global Variables

### Run regressions with original and lmsz centiles
```{r}
swyc_orig_reg_table_global <- as.data.frame(t(sapply(swyc_avg_centiles_lmsz[, which(names(swyc_avg_centiles_lmsz) %in% global_vars_totest)], function(x) {
  lm <- lm(x ~ swyc_avg_centiles_lmsz$median_dq)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2])
})))

swyc_orig_reg_table_global$swyc_p_adj <- p.adjust(swyc_orig_reg_table_global$swyc_p, method = "BH")

global_vars_totest_lmsz <- paste0(global_vars_totest, ".lmsz")

swyc_lmsz_reg_table_global <- as.data.frame(t(sapply(swyc_avg_centiles_lmsz[, which(names(swyc_avg_centiles_lmsz) %in% global_vars_totest_lmsz)], function(x) {
  lm <- lm(x ~ swyc_avg_centiles_lmsz$median_dq)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2])
})))

swyc_lmsz_reg_table_global$swyc_p_adj <- p.adjust(swyc_lmsz_reg_table_global$swyc_p, method = "BH")
```

### Calculate SE bars and combine into one df
```{r}
#orig centiles
swyc_orig_reg_table_global$swyc_lower <- (swyc_orig_reg_table_global$swyc_beta - 1.96*swyc_orig_reg_table_global$swyc_se)
swyc_orig_reg_table_global$swyc_upper <- (swyc_orig_reg_table_global$swyc_beta + 1.96*swyc_orig_reg_table_global$swyc_se)
swyc_orig_reg_table_global <- swyc_orig_reg_table_global %>% mutate(type = "orig")

#lmsz centiles
swyc_lmsz_reg_table_global$swyc_lower <- (swyc_lmsz_reg_table_global$swyc_beta - 1.96*swyc_lmsz_reg_table_global$swyc_se)
swyc_lmsz_reg_table_global$swyc_upper <- (swyc_lmsz_reg_table_global$swyc_beta + 1.96*swyc_lmsz_reg_table_global$swyc_se)
swyc_lmsz_reg_table_global <- swyc_lmsz_reg_table_global %>% mutate(type = "lmsz")

#combine tables
comb_swyc_lmsz_reg_table_global <- rbind(swyc_orig_reg_table_global, swyc_lmsz_reg_table_global)
comb_swyc_lmsz_reg_table_global$type <- as.factor(comb_swyc_lmsz_reg_table_global$type)
comb_swyc_lmsz_reg_table_global$label <- rownames(comb_swyc_lmsz_reg_table_global) %>% sub(".lmsz", "", .)
```

### Plot betas from both models with SE bars
```{r}
ggplot(comb_swyc_lmsz_reg_table_global, aes(y = label, x = swyc_beta, color = type)) +
  geom_point(size = 2, position = position_dodge(width = 0.3)) +
  geom_errorbar(aes(xmin = swyc_upper, xmax = swyc_lower),
                width = 0,
                position = position_dodge(width = 0.3)) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 15, margin = margin(r = 10)),  # space labels
        axis.title.y = element_text(margin = margin(r = 15)),             # space title
        panel.grid.major.x = element_line(), # x major grids visible
        panel.grid.major.y = element_blank(), # y grids gone
        panel.grid.minor.x = element_blank() ) + # x minor grids gone
  scale_y_discrete(expand = expansion(mult = c(0.1, 0.1))) +
  scale_color_manual(values = c("orig" = "grey", "lmsz" = "#B56576")) +
  labs(x = "Beta Estimate", y = "Phenotype", color = "Centile Model") + 
  geom_vline(xintercept = 0, linetype = "dashed", color = "black")
```


## Volume Variables

### Run regressions with original and lmsz centiles
```{r}
swyc_orig_reg_table_vol <- as.data.frame(t(sapply(swyc_avg_centiles_lmsz[, which(names(swyc_avg_centiles_lmsz) %in% vol_vars)], function(x) {
  lm <- lm(x ~ swyc_avg_centiles_lmsz$median_dq)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2])
})))

swyc_orig_reg_table_vol$swyc_p_adj <- p.adjust(swyc_orig_reg_table_vol$swyc_p, method = "BH")

vol_vars_lmsz <- paste0(vol_vars, ".lmsz")

swyc_lmsz_reg_table_vol <- as.data.frame(t(sapply(swyc_avg_centiles_lmsz[, which(names(swyc_avg_centiles_lmsz) %in% vol_vars_lmsz)], function(x) {
  lm <- lm(x ~ swyc_avg_centiles_lmsz$median_dq)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2])
})))

swyc_lmsz_reg_table_vol$swyc_p_adj <- p.adjust(swyc_lmsz_reg_table_vol$swyc_p, method = "BH")
```

### Calculate SE bars and combine into one df
```{r}
#orig centiles
swyc_orig_reg_table_vol$swyc_lower <- (swyc_orig_reg_table_vol$swyc_beta - 1.96*swyc_orig_reg_table_vol$swyc_se)
swyc_orig_reg_table_vol$swyc_upper <- (swyc_orig_reg_table_vol$swyc_beta + 1.96*swyc_orig_reg_table_vol$swyc_se)
swyc_orig_reg_table_vol <- swyc_orig_reg_table_vol %>% mutate(type = "orig")

#lmsz centiles
swyc_lmsz_reg_table_vol$swyc_lower <- (swyc_lmsz_reg_table_vol$swyc_beta - 1.96*swyc_lmsz_reg_table_vol$swyc_se)
swyc_lmsz_reg_table_vol$swyc_upper <- (swyc_lmsz_reg_table_vol$swyc_beta + 1.96*swyc_lmsz_reg_table_vol$swyc_se)
swyc_lmsz_reg_table_vol <- swyc_lmsz_reg_table_vol %>% mutate(type = "lmsz")

#combine tables
comb_swyc_lmsz_reg_table_vol <- rbind(swyc_orig_reg_table_vol, swyc_lmsz_reg_table_vol)
comb_swyc_lmsz_reg_table_vol$type <- as.factor(comb_swyc_lmsz_reg_table_vol$type)
comb_swyc_lmsz_reg_table_vol$label <- rownames(comb_swyc_lmsz_reg_table_vol) %>% sub(".lmsz", "", .)
```

### Plot betas from both models with SE bars
```{r}
# Right Hemisphere
ggplot(comb_swyc_lmsz_reg_table_vol %>% filter(grepl("aparc", label) & grepl("rh", label)), aes(y = sub("aparc_GrayVol_rh_", "", label), x = swyc_beta, color = type)) +
  geom_point(size = 2, position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(xmin = swyc_upper, xmax = swyc_lower),
                width = 0,
                position = position_dodge(width = 0.5)) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 10, margin = margin(r = 10)),  # space labels
        axis.title.y = element_text(margin = margin(r = 15)),             # space title
        panel.grid.major.x = element_line(), # x major grids visible
        panel.grid.major.y = element_blank(), # y grids gone
        panel.grid.minor.x = element_blank() ) + # x minor grids gone
  scale_y_discrete(expand = expansion(mult = c(0.1, 0.1))) +
  scale_color_manual(values = c("orig" = "grey", "lmsz" = "#B56576")) +
  labs(x = "Beta Estimate", y = "Right Cortical Vol", color = "Centile Model") + 
  geom_vline(xintercept = 0, linetype = "dashed", color = "black")

#Left Hemisphere
ggplot(comb_swyc_lmsz_reg_table_vol %>% filter(grepl("aparc", label) & grepl("lh", label)), aes(y = sub("aparc_GrayVol_lh_", "", label), x = swyc_beta, color = type)) +
  geom_point(size = 2, position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(xmin = swyc_upper, xmax = swyc_lower),
                width = 0,
                position = position_dodge(width = 0.5)) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 10, margin = margin(r = 10)),  # space labels
        axis.title.y = element_text(margin = margin(r = 15)),             # space title
        panel.grid.major.x = element_line(), # x major grids visible
        panel.grid.major.y = element_blank(), # y grids gone
        panel.grid.minor.x = element_blank() ) + # x minor grids gone
  scale_y_discrete(expand = expansion(mult = c(0.1, 0.1))) +
  scale_color_manual(values = c("orig" = "grey", "lmsz" = "#B56576")) +
  labs(x = "Beta Estimate", y = "Left Cortical Vol", color = "Centile Model") + 
  geom_vline(xintercept = 0, linetype = "dashed", color = "black")

#Subcortical
ggplot(comb_swyc_lmsz_reg_table_vol %>% filter(grepl("aseg", label)), aes(y = sub("aseg_", "", label), x = swyc_beta, color = type)) +
  geom_point(size = 2, position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(xmin = swyc_upper, xmax = swyc_lower),
                width = 0,
                position = position_dodge(width = 0.5)) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 10, margin = margin(r = 10)),  # space labels
        axis.title.y = element_text(margin = margin(r = 15)),             # space title
        panel.grid.major.x = element_line(), # x major grids visible
        panel.grid.major.y = element_blank(), # y grids gone
        panel.grid.minor.x = element_blank() ) + # x minor grids gone
  scale_y_discrete(expand = expansion(mult = c(0.1, 0.1))) +
  scale_color_manual(values = c("orig" = "grey", "lmsz" = "#B56576")) +
  labs(x = "Beta Estimate", y = "Subcortical Vol", color = "Centile Model") + 
  geom_vline(xintercept = 0, linetype = "dashed", color = "black")


```

## Surface Area Variables

### Run regressions with original and lmsz centiles
```{r}
swyc_orig_reg_table_sa <- as.data.frame(t(sapply(swyc_avg_centiles_lmsz[, which(names(swyc_avg_centiles_lmsz) %in% sa_vars)], function(x) {
  lm <- lm(x ~ swyc_avg_centiles_lmsz$median_dq)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2])
})))

swyc_orig_reg_table_sa$swyc_p_adj <- p.adjust(swyc_orig_reg_table_sa$swyc_p, method = "BH")

sa_vars_lmsz <- paste0(sa_vars, ".lmsz")

swyc_lmsz_reg_table_sa <- as.data.frame(t(sapply(swyc_avg_centiles_lmsz[, which(names(swyc_avg_centiles_lmsz) %in% sa_vars_lmsz)], function(x) {
  lm <- lm(x ~ swyc_avg_centiles_lmsz$median_dq)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2])
})))

swyc_lmsz_reg_table_sa$swyc_p_adj <- p.adjust(swyc_lmsz_reg_table_sa$swyc_p, method = "BH")
```

### Calculate SE bars and combine into one df
```{r}
#orig centiles
swyc_orig_reg_table_sa$swyc_lower <- (swyc_orig_reg_table_sa$swyc_beta - 1.96*swyc_orig_reg_table_sa$swyc_se)
swyc_orig_reg_table_sa$swyc_upper <- (swyc_orig_reg_table_sa$swyc_beta + 1.96*swyc_orig_reg_table_sa$swyc_se)
swyc_orig_reg_table_sa <- swyc_orig_reg_table_sa %>% mutate(type = "orig")

#lmsz centiles
swyc_lmsz_reg_table_sa$swyc_lower <- (swyc_lmsz_reg_table_sa$swyc_beta - 1.96*swyc_lmsz_reg_table_sa$swyc_se)
swyc_lmsz_reg_table_sa$swyc_upper <- (swyc_lmsz_reg_table_sa$swyc_beta + 1.96*swyc_lmsz_reg_table_sa$swyc_se)
swyc_lmsz_reg_table_sa <- swyc_lmsz_reg_table_sa %>% mutate(type = "lmsz")

#combine tables
comb_swyc_lmsz_reg_table_sa <- rbind(swyc_orig_reg_table_sa, swyc_lmsz_reg_table_sa)
comb_swyc_lmsz_reg_table_sa$type <- as.factor(comb_swyc_lmsz_reg_table_sa$type)
comb_swyc_lmsz_reg_table_sa$label <- rownames(comb_swyc_lmsz_reg_table_sa) %>% sub(".lmsz", "", .)
```

### Plot betas from both models with SE bars
```{r}
# Right Hemisphere
ggplot(comb_swyc_lmsz_reg_table_sa %>% filter(grepl("rh", label)), aes(y = sub("aparc_SurfArea_rh_", "", label), x = swyc_beta, color = type)) +
  geom_point(size = 2, position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(xmin = swyc_upper, xmax = swyc_lower),
                width = 0,
                position = position_dodge(width = 0.5)) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 10, margin = margin(r = 10)),  # space labels
        axis.title.y = element_text(margin = margin(r = 15)),             # space title
        panel.grid.major.x = element_line(), # x major grids visible
        panel.grid.major.y = element_blank(), # y grids gone
        panel.grid.minor.x = element_blank() ) + # x minor grids gone
  scale_y_discrete(expand = expansion(mult = c(0.1, 0.1))) +
  scale_color_manual(values = c("orig" = "grey", "lmsz" = "#B56576")) +
  labs(x = "Beta Estimate", y = "Right Cortical SA", color = "Centile Model") + 
  geom_vline(xintercept = 0, linetype = "dashed", color = "black")

#Left Hemisphere
ggplot(comb_swyc_lmsz_reg_table_sa %>% filter(grepl("lh", label)), aes(y = sub("aparc_SurfArea_lh_", "", label), x = swyc_beta, color = type)) +
  geom_point(size = 2, position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(xmin = swyc_upper, xmax = swyc_lower),
                width = 0,
                position = position_dodge(width = 0.5)) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 10, margin = margin(r = 10)),  # space labels
        axis.title.y = element_text(margin = margin(r = 15)),             # space title
        panel.grid.major.x = element_line(), # x major grids visible
        panel.grid.major.y = element_blank(), # y grids gone
        panel.grid.minor.x = element_blank() ) + # x minor grids gone
  scale_y_discrete(expand = expansion(mult = c(0.1, 0.1))) +
  scale_color_manual(values = c("orig" = "grey", "lmsz" = "#B56576")) +
  labs(x = "Beta Estimate", y = "Left Cortical SA", color = "Centile Model") + 
  geom_vline(xintercept = 0, linetype = "dashed", color = "black")

```

## Cortical Thickness Variables

### Run regressions with original and lmsz centiles
```{r}
swyc_orig_reg_table_th <- as.data.frame(t(sapply(swyc_avg_centiles_lmsz[, which(names(swyc_avg_centiles_lmsz) %in% th_vars)], function(x) {
  lm <- lm(x ~ swyc_avg_centiles_lmsz$median_dq)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2])
})))

swyc_orig_reg_table_th$swyc_p_adj <- p.adjust(swyc_orig_reg_table_th$swyc_p, method = "BH")

th_vars_lmsz <- paste0(th_vars, ".lmsz")

swyc_lmsz_reg_table_th <- as.data.frame(t(sapply(swyc_avg_centiles_lmsz[, which(names(swyc_avg_centiles_lmsz) %in% th_vars_lmsz)], function(x) {
  lm <- lm(x ~ swyc_avg_centiles_lmsz$median_dq)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2])
})))

swyc_lmsz_reg_table_th$swyc_p_adj <- p.adjust(swyc_lmsz_reg_table_th$swyc_p, method = "BH")
```

### Calculate SE bars and combine into one df
```{r}
#orig centiles
swyc_orig_reg_table_th$swyc_lower <- (swyc_orig_reg_table_th$swyc_beta - 1.96*swyc_orig_reg_table_th$swyc_se)
swyc_orig_reg_table_th$swyc_upper <- (swyc_orig_reg_table_th$swyc_beta + 1.96*swyc_orig_reg_table_th$swyc_se)
swyc_orig_reg_table_th <- swyc_orig_reg_table_th %>% mutate(type = "orig")

#lmsz centiles
swyc_lmsz_reg_table_th$swyc_lower <- (swyc_lmsz_reg_table_th$swyc_beta - 1.96*swyc_lmsz_reg_table_th$swyc_se)
swyc_lmsz_reg_table_th$swyc_upper <- (swyc_lmsz_reg_table_th$swyc_beta + 1.96*swyc_lmsz_reg_table_th$swyc_se)
swyc_lmsz_reg_table_th <- swyc_lmsz_reg_table_th %>% mutate(type = "lmsz")

#combine tables
comb_swyc_lmsz_reg_table_th <- rbind(swyc_orig_reg_table_th, swyc_lmsz_reg_table_th)
comb_swyc_lmsz_reg_table_th$type <- as.factor(comb_swyc_lmsz_reg_table_th$type)
comb_swyc_lmsz_reg_table_th$label <- rownames(comb_swyc_lmsz_reg_table_th) %>% sub(".lmsz", "", .)
```

### Plot betas from both models with SE bars
```{r}
# Right Hemisphere
ggplot(comb_swyc_lmsz_reg_table_th %>% filter(grepl("rh", label)), aes(y = sub("aparc_ThickAvg_rh_", "", label), x = swyc_beta, color = type)) +
  geom_point(size = 2, position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(xmin = swyc_upper, xmax = swyc_lower),
                width = 0,
                position = position_dodge(width = 0.5)) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 10, margin = margin(r = 10)),  # space labels
        axis.title.y = element_text(margin = margin(r = 15)),             # space title
        panel.grid.major.x = element_line(), # x major grids visible
        panel.grid.major.y = element_blank(), # y grids gone
        panel.grid.minor.x = element_blank() ) + # x minor grids gone
  scale_y_discrete(expand = expansion(mult = c(0.1, 0.1))) +
  scale_color_manual(values = c("orig" = "grey", "lmsz" = "#B56576")) +
  labs(x = "Beta Estimate", y = "Right Cortical Thick", color = "Centile Model") + 
  geom_vline(xintercept = 0, linetype = "dashed", color = "black")

#Left Hemisphere
ggplot(comb_swyc_lmsz_reg_table_th %>% filter(grepl("lh", label)), aes(y = sub("aparc_ThickAvg_lh_", "", label), x = swyc_beta, color = type)) +
  geom_point(size = 2, position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(xmin = swyc_upper, xmax = swyc_lower),
                width = 0,
                position = position_dodge(width = 0.5)) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 10, margin = margin(r = 10)),  # space labels
        axis.title.y = element_text(margin = margin(r = 15)),             # space title
        panel.grid.major.x = element_line(), # x major grids visible
        panel.grid.major.y = element_blank(), # y grids gone
        panel.grid.minor.x = element_blank() ) + # x minor grids gone
  scale_y_discrete(expand = expansion(mult = c(0.1, 0.1))) +
  scale_color_manual(values = c("orig" = "grey", "lmsz" = "#B56576")) +
  labs(x = "Beta Estimate", y = "Left Cortical Thick", color = "Centile Model") + 
  geom_vline(xintercept = 0, linetype = "dashed", color = "black")
```
