---
title: "slip_premie_lmsz_plotting"
author: "Eren Kafadar"
date: "2026-01-29"
output:
  html_document:
    df_print: paged
params:
  output_folder: "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/output/lmsz_median_Th-MPR_25.11/"
  centiles_data_path: "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/output/median-Th-MPR-distinct-allgrades-25.11/data_clean/median_Th-MPR_2026-02-04.csv"
  raw_data_path: "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11/raw_csv/median_Th-MPR_2026-02-04.csv"

---
# Define params if they do not exist
```{r define params, echo = FALSE}
if (!exists("params")) {
  params <- list(
   output_folder = "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/output/lmsz_median_Th-MPR_25.11/",
  centiles_data_path = "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/output/median-Th-MPR-distinct-allgrades-25.11/data_clean/median_Th-MPR_2026-02-04.csv",
  raw_data_path = "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11/raw_csv/median_Th-MPR_2026-02-04.csv"
  )
}
print(params)
``` 

# Controls
```{r setup, echo = TRUE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)
# set command
save_tables <- TRUE
plot <- TRUE
save_figs <- TRUE
set.seed(42)

output_folder <- params$output_folder
centiles_data_path <- params$centiles_data_path
raw_data_path <- params$raw_data_path
```

# Load Packages
```{r, echo = FALSE}
# Load libraries
require(dplyr)
require(magrittr)
require(readr)
require(stringr)
require(tidyr)
require(gamlssTools) # Margaret's package
require(tibble)
require(glue)
# For figures/plotting
require(ggplot2)
# For figures/plotting
require(gridExtra)
require(ggsegDKT)
require(ggseg)
# require(ggpubr)
require(ggcorrplot)
require(cowplot)
require(egg)
require(patchwork)
require(interactions)
require(growthstandards)
require(gtsummary)
require(svglite)
```

# Define Paths
```{r, echo = TRUE}
# set folder paths
code_path <- "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/"
vars_path <-  "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/output/vars/"

##  output subfolder paths
csv_save_path <-  paste0(output_folder, "results_csv/")
if (!dir.exists(csv_save_path)) dir.create(csv_save_path)
plot_save_path <- paste0(output_folder, "plots/")
if (!dir.exists(plot_save_path)) dir.create(plot_save_path)
```

# Load custom functions
```{r, echo = FALSE}
panel_save <- function(plot, filename, extension, folder, plot_width = 8, plot_height = 10){
  save_name <- paste0(folder, filename, ".", extension)
  ggsave(save_name, plot = plot, dpi = 300, width = plot_width, height = plot_height)
}

source(glue("{code_path}R_util_lmsz.R"))
```

# Load Phenotype Variable Names
```{r, echo = FALSE}
full_vars <- read.csv(paste0(vars_path,"all_vars.csv")) %>% pull(x)
global_vars <- read.csv(paste0(vars_path,"global_vars.csv")) %>% pull(x)
vol_vars <- read.csv(paste0(vars_path,"vol_vars.csv")) %>% pull(x)
sa_vars <- read.csv(paste0(vars_path,"sa_vars.csv")) %>% pull(x)
th_vars <- read.csv(paste0(vars_path,"th_vars.csv")) %>% pull(x)

## remove aseg_CC_Central and "aseg_CC_Mid_Posterior" from the features because there is one NaN value in the centiles causing issues. In the future will edit the functions to deal with this and/or figure out why there is this NaN value.
full_vars <- full_vars[!(full_vars %in% c("aseg_CC_Central", "aseg_CC_Mid_Posterior"))]
```
# Load Clean Data
```{r, echo = FALSE}
df <- read.csv(centiles_data_path) %>% select(-"X")
```

# Load Raw Data and merge
```{r, echo = FALSE}
df_raw <- read.csv(raw_data_path) %>% select(all_of(c(full_vars, "participant_id", "session_id", "sex", "age_days",
                                                      "site", "avg_grade", "age_at_scan"))) %>%
  #mutate(across(all_of(full_vars), ~c(scale(.)), .names = "{.col}.raw_normalised")) %>%
  rename_with(~paste0(.x, ".raw"), all_of(full_vars))

shared_names <- intersect(names(df_raw), names(df))

df <- merge(df, df_raw, by = shared_names, all.x = FALSE, all.y = FALSE)
```

# Z-score Centiles, remove Inf Cells

## zscore centiles (new df)
```{r, echo = FALSE}
#zscore centiles and other variables of interest
df <- df %>%
  mutate(across(all_of(full_vars), qnorm,
                         .names = "{.col}.zscore"))
#set sex and preterm as factor
df$sex <- as.factor(df$sex)
df$preterm <- as.factor(df$preterm)
zscore_vars_names <- paste0(full_vars, ".zscore")
raw_vars_names <- paste0(full_vars, ".raw")
```

## find Inf cells, make a table of N per phenotype, then convert to NaN
```{r, echo = FALSE}
inf_nan_table <- as.data.frame(t(sapply(df[zscore_vars_names], function(x){
  c(inf = sum(is.infinite(x)), nan = sum(is.nan(x)))
})))
max(inf_nan_table$inf)
max(inf_nan_table$nan)

inf_nan_table %>% filter(inf > 0)
```

## set Inf cells to NaN
```{r, echo = FALSE}
df <- df %>% mutate(across(all_of(zscore_vars_names), ~ replace(., is.infinite(.), NaN)))
```

# Load LMSz dfs and model paths
```{r, echo = FALSE}
models_vpm <- read.csv(glue("{csv_save_path}lmsz_models_vpm.csv"))
df_vpm <- read.csv(glue("{csv_save_path}lmsz_adjusted_centiles_vpm.csv"))
models_vpm_path <- glue("{output_folder}lmsz_models_vpm/")

models_lpm <- read.csv(glue("{csv_save_path}lmsz_models_lpm.csv"))
df_lpm <- read.csv(glue("{csv_save_path}lmsz_adjusted_centiles_lpm.csv"))
models_lpm_path <- glue("{output_folder}lmsz_models_lpm/")

#fix columns of type logical to avoid plotting errors
vpm_col_fix <- models_vpm %>% select(names(models_vpm)[grepl("coef", names(models_vpm))] & is.logical) %>% colnames()
models_vpm[,vpm_col_fix] <- NA_real_

lpm_col_fix <- models_lpm %>% select(names(models_lpm)[grepl("coef", names(models_lpm))] & is.logical) %>% colnames()
models_lpm[,lpm_col_fix] <- NA_real_

## Load previously calculated centile fans
out_fans_slip <- read.csv(glue("{csv_save_path}lmsz_centfans_vpm_uncorrected.csv"))
```

# Visualize Model Output on Brain Maps

## Plots specifications
```{r, echo = TRUE}
#limits = c(-3, 3)
low_color = "#4878CF"
mid_color = "white"
high_color = "#D65F5F"
```

## add label column to match to ggseg
```{r, echo = FALSE}
#models vpm
models_vpm$domain <- ifelse(grepl("global", models_vpm$feature), "global",
                            ifelse(grepl("GrayVol", models_vpm$feature), "vol",
                            ifelse(grepl("SurfArea", models_vpm$feature), "sa",
                                   ifelse(grepl("ThickAvg", models_vpm$feature), "th",
                                          ifelse(grepl("aseg", models_vpm$feature), "aseg", NA)))))
cols_edit <- grepl("aseg(?!_CC)", models_vpm$feature, perl = TRUE)

models_vpm$label <- models_vpm$feature %>% 
  sub("aparc_GrayVol_", "", .) %>%
  sub("aparc_SurfArea_", "", .) %>%
  sub("aparc_ThickAvg_", "", .) %>%
  sub("aseg_", "", .) %>%
  sub("X", "x", .) %>%
  sub("rd_Ventricle", "rd_ventricle", .) %>%
  sub("th_Ventricle", "th_ventricle", .) %>%
  sub("Thalamus", "Thalamus-Proper", .)
  
models_vpm$label[cols_edit] <- gsub("_", "-", models_vpm$label[cols_edit])

#models lpm
models_lpm$domain <- ifelse(grepl("global", models_lpm$feature), "global",
                            ifelse(grepl("GrayVol", models_lpm$feature), "vol",
                            ifelse(grepl("SurfArea", models_lpm$feature), "sa",
                                   ifelse(grepl("ThickAvg", models_lpm$feature), "th",
                                          ifelse(grepl("aseg", models_lpm$feature), "aseg", NA)))))
cols_edit <- grepl("aseg(?!_CC)", models_lpm$feature, perl = TRUE)

models_lpm$label <- models_lpm$feature %>% 
  sub("aparc_GrayVol_", "", .) %>%
  sub("aparc_SurfArea_", "", .) %>%
  sub("aparc_ThickAvg_", "", .) %>%
  sub("aseg_", "", .) %>%
  sub("X", "x", .) %>%
  sub("rd_Ventricle", "rd_ventricle", .) %>%
  sub("th_Ventricle", "th_ventricle", .) %>%
  sub("Thalamus", "Thalamus-Proper", .)
  
models_lpm$label[cols_edit] <- gsub("_", "-", models_lpm$label[cols_edit])
```

## Plot mu intercept value of all phenotypes

### VPM
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(models_vpm, ggseg::dk, by = "label")
plot_data_aseg <- merge(models_vpm, ggseg::aseg, by = "label")

# Regions not matched in atlases
#print(models_vpm_noplot <- models_vpm[!(models_vpm$label %in% plot_data_dkt$label) & !(models_vpm$label %in% plot_data_aseg$label),"label"], n = 100)

# plots
(vpm_vol_mu_intercept_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "vol")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = intercept_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-2.5, 2.5)) +
  theme_void()+
  labs(title = "VPM Volume Mu Intercept"))

(vpm_vol_mu_intercept_aseg_plot <- ggplot(plot_data_aseg %>% filter(domain == "aseg")) +
  geom_brain(atlas = aseg,
             aes(fill = intercept_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-2.5, 2.5)) +
  theme_void() +
  labs(title = "VPM Volume Mu Intercept"))

(vpm_sa_mu_intercept_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "sa")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = intercept_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-2.5, 2.5)) +
  theme_void()+
  labs(title = "VPM SA Mu Intercept"))

(vpm_th_mu_intercept_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "th")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = intercept_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-2.5, 2.5)) +
  theme_void()+
  labs(title = "VPM Thickness Mu Intercept"))

if(save_figs == TRUE){
panel_save(vpm_vol_mu_intercept_dk_plot, "vpm-vol-mu-intercept-dk", "pdf", plot_save_path)
panel_save(vpm_vol_mu_intercept_aseg_plot, "vpm-vol-mu-intercept-aseg", "pdf", plot_save_path)
panel_save(vpm_sa_mu_intercept_dk_plot, "vpm-sa-mu-intercept-dk", "pdf", plot_save_path)
panel_save(vpm_th_mu_intercept_dk_plot, "vpm-th-mu-intercept-dk", "pdf", plot_save_path)
}
```

### LPM
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(models_lpm, ggseg::dk, by = "label")
plot_data_aseg <- merge(models_lpm, ggseg::aseg, by = "label")

# Regions not matched in atlases
#print(models_lpm_noplot <- models_lpm[!(models_lpm$label %in% plot_data_dkt$label) & !(models_lpm$label %in% plot_data_aseg$label),"label"], n = 100)

# plots
(lpm_vol_mu_intercept_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "vol")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = intercept_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-2.5, 2.5)) +
  theme_void()+
  labs(title = "LPM Volume Mu Intercept"))

(lpm_vol_mu_intercept_aseg_plot <- ggplot(plot_data_aseg %>% filter(domain == "aseg")) +
  geom_brain(atlas = aseg,
             aes(fill = intercept_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-2.5, 2.5)) +
  theme_void() +
  labs(title = "LPM Volume Mu Intercept"))

(lpm_sa_mu_intercept_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "sa")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = intercept_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-2.5, 2.5)) +
  theme_void()+
  labs(title = "LPM SA Mu Intercept"))

(lpm_th_mu_intercept_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "th")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = intercept_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-2.5, 2.5)) +
  theme_void()+
  labs(title = "LPM Thickness Mu Intercept"))

if(save_figs == TRUE){
panel_save(lpm_vol_mu_intercept_dk_plot, "lpm-vol-mu-intercept-dk", "pdf", plot_save_path)
panel_save(lpm_vol_mu_intercept_aseg_plot, "lpm-vol-mu-intercept-aseg", "pdf", plot_save_path)
panel_save(lpm_sa_mu_intercept_dk_plot, "lpm-sa-mu-intercept-dk", "pdf", plot_save_path)
panel_save(lpm_th_mu_intercept_dk_plot, "lpm-th-mu-intercept-dk", "pdf", plot_save_path)
}
```

## Plot coefficient of age in mu

### VPM
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(models_vpm, ggseg::dk, by = "label")
plot_data_aseg <- merge(models_vpm, ggseg::aseg, by = "label")

# Regions not matched in atlases
#print(models_vpm_noplot <- models_vpm[!(models_vpm$label %in% plot_data_dkt$label) & !(models_vpm$label %in% plot_data_aseg$label),"label"], n = 100)

# plots
(vpm_vol_mu_age_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "vol")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = age_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-0.35, 0.35)) +
  theme_void()+
  labs(title = "VPM Volume Mu age"))

(vpm_vol_mu_age_aseg_plot <- ggplot(plot_data_aseg %>% filter(domain == "aseg")) +
  geom_brain(atlas = aseg,
             aes(fill = age_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-0.35, 0.35)) +
  theme_void() +
  labs(title = "VPM Volume Mu age"))

(vpm_sa_mu_age_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "sa")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = age_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-0.35, 0.35)) +
  theme_void()+
  labs(title = "VPM SA Mu age"))

(vpm_th_mu_age_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "th")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = age_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-0.35, 0.35)) +
  theme_void()+
  labs(title = "VPM Thickness Mu age"))

if(save_figs == TRUE){
panel_save(vpm_vol_mu_age_dk_plot, "vpm-vol-mu-age-dk", "pdf", plot_save_path)
panel_save(vpm_vol_mu_age_aseg_plot, "vpm-vol-mu-age-aseg", "pdf", plot_save_path)
panel_save(vpm_sa_mu_age_dk_plot, "vpm-sa-mu-age-dk", "pdf", plot_save_path)
panel_save(vpm_th_mu_age_dk_plot, "vpm-th-mu-age-dk", "pdf", plot_save_path)
}
```

### LPM
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(models_lpm, ggseg::dk, by = "label")
plot_data_aseg <- merge(models_lpm, ggseg::aseg, by = "label")

# Regions not matched in atlases
#print(models_lpm_noplot <- models_lpm[!(models_lpm$label %in% plot_data_dkt$label) & !(models_lpm$label %in% plot_data_aseg$label),"label"], n = 100)

# plots
(lpm_vol_mu_age_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "vol")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = age_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-0.35, 0.35)) +
  theme_void()+
  labs(title = "LPM Volume Mu age"))

(lpm_vol_mu_age_aseg_plot <- ggplot(plot_data_aseg %>% filter(domain == "aseg")) +
  geom_brain(atlas = aseg,
             aes(fill = age_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-0.35, 0.35)) +
  theme_void() +
  labs(title = "LPM Volume Mu age"))

(lpm_sa_mu_age_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "sa")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = age_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-0.35, 0.35)) +
  theme_void()+
  labs(title = "LPM SA Mu age"))

(lpm_th_mu_age_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "th")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = age_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-0.35, 0.35)) +
  theme_void()+
  labs(title = "LPM Thickness Mu age"))

if(save_figs == TRUE){
panel_save(lpm_vol_mu_age_dk_plot, "lpm-vol-mu-age-dk", "pdf", plot_save_path)
panel_save(lpm_vol_mu_age_aseg_plot, "lpm-vol-mu-age-aseg", "pdf", plot_save_path)
panel_save(lpm_sa_mu_age_dk_plot, "lpm-sa-mu-age-dk", "pdf", plot_save_path)
panel_save(lpm_th_mu_age_dk_plot, "lpm-th-mu-age-dk", "pdf", plot_save_path)
}
```

## Plot coefficient of sex in mu

### VPM
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(models_vpm, ggseg::dk, by = "label")
plot_data_aseg <- merge(models_vpm, ggseg::aseg, by = "label")

# plots
(vpm_vol_mu_sex_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "vol")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = sex_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-0.4, 0.4)) +
  theme_void()+
  labs(title = "VPM Volume Mu sex"))

(vpm_vol_mu_sex_aseg_plot <- ggplot(plot_data_aseg %>% filter(domain == "aseg")) +
  geom_brain(atlas = aseg,
             aes(fill = sex_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-0.4, 0.4)) +
  theme_void() +
  labs(title = "VPM Volume Mu sex"))

(vpm_sa_mu_sex_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "sa")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = sex_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-0.4, 0.4)) +
  theme_void()+
  labs(title = "VPM SA Mu sex"))

(vpm_th_mu_sex_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "th")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = sex_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-0.4, 0.4)) +
  theme_void()+
  labs(title = "VPM Thickness Mu sex"))

if(save_figs == TRUE){
panel_save(vpm_vol_mu_sex_dk_plot, "vpm-vol-mu-sex-dk", "pdf", plot_save_path)
panel_save(vpm_vol_mu_sex_aseg_plot, "vpm-vol-mu-sex-aseg", "pdf", plot_save_path)
panel_save(vpm_sa_mu_sex_dk_plot, "vpm-sa-mu-sex-dk", "pdf", plot_save_path)
panel_save(vpm_th_mu_sex_dk_plot, "vpm-th-mu-sex-dk", "pdf", plot_save_path)
}
```

### LPM
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(models_lpm, ggseg::dk, by = "label")
plot_data_aseg <- merge(models_lpm, ggseg::aseg, by = "label")

# Regions not matched in atlases
#print(models_lpm_noplot <- models_lpm[!(models_lpm$label %in% plot_data_dkt$label) & !(models_lpm$label %in% plot_data_aseg$label),"label"], n = 100)

# plots
(lpm_vol_mu_sex_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "vol")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = sex_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-0.4, 0.4)) +
  theme_void()+
  labs(title = "LPM Volume Mu sex"))

(lpm_vol_mu_sex_aseg_plot <- ggplot(plot_data_aseg %>% filter(domain == "aseg")) +
  geom_brain(atlas = aseg,
             aes(fill = sex_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-0.4, 0.4)) +
  theme_void() +
  labs(title = "LPM Volume Mu sex"))

(lpm_sa_mu_sex_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "sa")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = sex_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-0.4, 0.4)) +
  theme_void()+
  labs(title = "LPM SA Mu sex"))

(lpm_th_mu_sex_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "th")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = sex_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-0.4, 0.4)) +
  theme_void()+
  labs(title = "LPM Thickness Mu sex"))

if(save_figs == TRUE){
panel_save(lpm_vol_mu_sex_dk_plot, "lpm-vol-mu-sex-dk", "pdf", plot_save_path)
panel_save(lpm_vol_mu_sex_aseg_plot, "lpm-vol-mu-sex-aseg", "pdf", plot_save_path)
panel_save(lpm_sa_mu_sex_dk_plot, "lpm-sa-mu-sex-dk", "pdf", plot_save_path)
panel_save(lpm_th_mu_sex_dk_plot, "lpm-th-mu-sex-dk", "pdf", plot_save_path)
}
```

## Plot intercept in sigma

### VPM
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(models_vpm, ggseg::dk, by = "label")
plot_data_aseg <- merge(models_vpm, ggseg::aseg, by = "label")

# Regions not matched in atlases
#print(models_vpm_noplot <- models_vpm[!(models_vpm$label %in% plot_data_dkt$label) & !(models_vpm$label %in% plot_data_aseg$label),"label"], n = 100)

# plots
(vpm_vol_sigma_intercept_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "vol")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = intercept_sigma_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-1.75, 1.75)) +
  theme_void()+
  labs(title = "VPM Volume sigma Intercept"))

(vpm_vol_sigma_intercept_aseg_plot <- ggplot(plot_data_aseg %>% filter(domain == "aseg")) +
  geom_brain(atlas = aseg,
             aes(fill = intercept_sigma_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-1.75, 1.75)) +
  theme_void() +
  labs(title = "VPM Volume sigma Intercept"))

(vpm_sa_sigma_intercept_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "sa")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = intercept_sigma_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-1.75, 1.75)) +
  theme_void()+
  labs(title = "VPM SA sigma Intercept"))

(vpm_th_sigma_intercept_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "th")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = intercept_sigma_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-1.75, 1.75)) +
  theme_void()+
  labs(title = "VPM Thickness sigma Intercept"))

if(save_figs == TRUE){
panel_save(vpm_vol_sigma_intercept_dk_plot, "vpm-vol-sigma-intercept-dk", "pdf", plot_save_path)
panel_save(vpm_vol_sigma_intercept_aseg_plot, "vpm-vol-sigma-intercept-aseg", "pdf", plot_save_path)
panel_save(vpm_sa_sigma_intercept_dk_plot, "vpm-sa-sigma-intercept-dk", "pdf", plot_save_path)
panel_save(vpm_th_sigma_intercept_dk_plot, "vpm-th-sigma-intercept-dk", "pdf", plot_save_path)
}
```

### LPM
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(models_lpm, ggseg::dk, by = "label")
plot_data_aseg <- merge(models_lpm, ggseg::aseg, by = "label")

# Regions not matched in atlases
#print(models_lpm_noplot <- models_lpm[!(models_lpm$label %in% plot_data_dkt$label) & !(models_lpm$label %in% plot_data_aseg$label),"label"], n = 100)

# plots
(lpm_vol_sigma_intercept_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "vol")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = intercept_sigma_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-1.75, 1.75)) +
  theme_void()+
  labs(title = "LPM Volume sigma Intercept"))

(lpm_vol_sigma_intercept_aseg_plot <- ggplot(plot_data_aseg %>% filter(domain == "aseg")) +
  geom_brain(atlas = aseg,
             aes(fill = intercept_sigma_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-1.75, 1.75)) +
  theme_void() +
  labs(title = "LPM Volume sigma Intercept"))

(lpm_sa_sigma_intercept_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "sa")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = intercept_sigma_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-1.75, 1.75)) +
  theme_void()+
  labs(title = "LPM SA sigma Intercept"))

(lpm_th_sigma_intercept_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "th")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = intercept_sigma_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-1.75, 1.75)) +
  theme_void()+
  labs(title = "LPM Thickness sigma Intercept"))

if(save_figs == TRUE){
panel_save(lpm_vol_sigma_intercept_dk_plot, "lpm-vol-sigma-intercept-dk", "pdf", plot_save_path)
panel_save(lpm_vol_sigma_intercept_aseg_plot, "lpm-vol-sigma-intercept-aseg", "pdf", plot_save_path)
panel_save(lpm_sa_sigma_intercept_dk_plot, "lpm-sa-sigma-intercept-dk", "pdf", plot_save_path)
panel_save(lpm_th_sigma_intercept_dk_plot, "lpm-th-sigma-intercept-dk", "pdf", plot_save_path)
}
```

## Plot coefficient of age in sigma

### VPM
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(models_vpm, ggseg::dk, by = "label")
plot_data_aseg <- merge(models_vpm, ggseg::aseg, by = "label")

# Regions not matched in atlases
#print(models_vpm_noplot <- models_vpm[!(models_vpm$label %in% plot_data_dkt$label) & !(models_vpm$label %in% plot_data_aseg$label),"label"], n = 100)

# plots
(vpm_vol_sigma_age_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "vol")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = age_sigma_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-0.25, 0.25)) +
  theme_void()+
  labs(title = "VPM Volume sigma age"))

(vpm_vol_sigma_age_aseg_plot <- ggplot(plot_data_aseg %>% filter(domain == "aseg")) +
  geom_brain(atlas = aseg,
             aes(fill = age_sigma_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-0.25, 0.25)) +
  theme_void() +
  labs(title = "VPM Volume sigma age"))

(vpm_sa_sigma_age_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "sa")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = age_sigma_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-0.25, 0.25)) +
  theme_void()+
  labs(title = "VPM SA sigma age"))

(vpm_th_sigma_age_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "th")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = age_sigma_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-0.25, 0.25)) +
  theme_void()+
  labs(title = "VPM Thickness sigma age"))

if(save_figs == TRUE){
panel_save(vpm_vol_sigma_age_dk_plot, "vpm-vol-sigma-age-dk", "pdf", plot_save_path)
panel_save(vpm_vol_sigma_age_aseg_plot, "vpm-vol-sigma-age-aseg", "pdf", plot_save_path)
panel_save(vpm_sa_sigma_age_dk_plot, "vpm-sa-sigma-age-dk", "pdf", plot_save_path)
panel_save(vpm_th_sigma_age_dk_plot, "vpm-th-sigma-age-dk", "pdf", plot_save_path)
}
```

### LPM
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(models_lpm, ggseg::dk, by = "label")
plot_data_aseg <- merge(models_lpm, ggseg::aseg, by = "label")

# Regions not matched in atlases
#print(models_lpm_noplot <- models_lpm[!(models_lpm$label %in% plot_data_dkt$label) & !(models_lpm$label %in% plot_data_aseg$label),"label"], n = 100)

# plots
(lpm_vol_sigma_age_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "vol")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = age_sigma_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-0.25, 0.25)) +
  theme_void()+
  labs(title = "LPM Volume sigma age"))

(lpm_vol_sigma_age_aseg_plot <- ggplot(plot_data_aseg %>% filter(domain == "aseg")) +
  geom_brain(atlas = aseg,
             aes(fill = age_sigma_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-0.25, 0.25)) +
  theme_void() +
  labs(title = "LPM Volume sigma age"))

(lpm_sa_sigma_age_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "sa")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = age_sigma_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-0.25, 0.25)) +
  theme_void()+
  labs(title = "LPM SA sigma age"))

(lpm_th_sigma_age_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "th")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = age_sigma_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-0.25, 0.25)) +
  theme_void()+
  labs(title = "LPM Thickness sigma age"))

if(save_figs == TRUE){
panel_save(lpm_vol_sigma_age_dk_plot, "lpm-vol-sigma-age-dk", "pdf", plot_save_path)
panel_save(lpm_vol_sigma_age_aseg_plot, "lpm-vol-sigma-age-aseg", "pdf", plot_save_path)
panel_save(lpm_sa_sigma_age_dk_plot, "lpm-sa-sigma-age-dk", "pdf", plot_save_path)
panel_save(lpm_th_sigma_age_dk_plot, "lpm-th-sigma-age-dk", "pdf", plot_save_path)
}
```

## Plot coefficient of sex in sigma

### VPM
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(models_vpm, ggseg::dk, by = "label")
plot_data_aseg <- merge(models_vpm, ggseg::aseg, by = "label")

# Regions not matched in atlases
#print(models_vpm_noplot <- models_vpm[!(models_vpm$label %in% plot_data_dkt$label) & !(models_vpm$label %in% plot_data_aseg$label),"label"], n = 100)

# plots
(vpm_vol_sigma_sex_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "vol")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = sex_sigma_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-0.5, 0.5)) +
  theme_void()+
  labs(title = "VPM Volume sigma sex"))

(vpm_vol_sigma_sex_aseg_plot <- ggplot(plot_data_aseg %>% filter(domain == "aseg")) +
  geom_brain(atlas = aseg,
             aes(fill = sex_sigma_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-0.5, 0.5)) +
  theme_void() +
  labs(title = "VPM Volume sigma sex"))

(vpm_sa_sigma_sex_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "sa")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = sex_sigma_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-0.5, 0.5)) +
  theme_void()+
  labs(title = "VPM SA sigma sex"))

(vpm_th_sigma_sex_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "th")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = sex_sigma_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-0.5, 0.5)) +
  theme_void()+
  labs(title = "VPM Thickness sigma sex"))

if(save_figs == TRUE){
panel_save(vpm_vol_sigma_sex_dk_plot, "vpm-vol-sigma-sex-dk", "pdf", plot_save_path)
panel_save(vpm_vol_sigma_sex_aseg_plot, "vpm-vol-sigma-sex-aseg", "pdf", plot_save_path)
panel_save(vpm_sa_sigma_sex_dk_plot, "vpm-sa-sigma-sex-dk", "pdf", plot_save_path)
panel_save(vpm_th_sigma_sex_dk_plot, "vpm-th-sigma-sex-dk", "pdf", plot_save_path)
}
```

### LPM
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(models_lpm, ggseg::dk, by = "label")
plot_data_aseg <- merge(models_lpm, ggseg::aseg, by = "label")

# Regions not matched in atlases
#print(models_lpm_noplot <- models_lpm[!(models_lpm$label %in% plot_data_dkt$label) & !(models_lpm$label %in% plot_data_aseg$label),"label"], n = 100)

# plots
(lpm_vol_sigma_sex_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "vol")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = sex_sigma_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-0.5, 0.5)) +
  theme_void()+
  labs(title = "LPM Volume sigma sex"))

(lpm_vol_sigma_sex_aseg_plot <- ggplot(plot_data_aseg %>% filter(domain == "aseg")) +
  geom_brain(atlas = aseg,
             aes(fill = sex_sigma_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-0.5, 0.5)) +
  theme_void() +
  labs(title = "LPM Volume sigma sex"))

(lpm_sa_sigma_sex_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "sa")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = sex_sigma_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-0.5, 0.5)) +
  theme_void()+
  labs(title = "LPM SA sigma sex"))

(lpm_th_sigma_sex_dk_plot <- ggplot(plot_data_dkt %>% filter(domain == "th")) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = sex_sigma_coef)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = c(-0.5, 0.5)) +
  theme_void()+
  labs(title = "LPM Thickness sigma sex"))

if(save_figs == TRUE){
panel_save(lpm_vol_sigma_sex_dk_plot, "lpm-vol-sigma-sex-dk", "pdf", plot_save_path)
panel_save(lpm_vol_sigma_sex_aseg_plot, "lpm-vol-sigma-sex-aseg", "pdf", plot_save_path)
panel_save(lpm_sa_sigma_sex_dk_plot, "lpm-sa-sigma-sex-dk", "pdf", plot_save_path)
panel_save(lpm_th_sigma_sex_dk_plot, "lpm-th-sigma-sex-dk", "pdf", plot_save_path)
}
```

# Trajectory Plots Prep

## Fix scale of out_fans, scale everything back by 1000
```{r, echo = FALSE}
#Scale all phenotypes by 1000
out_fans_slip <- out_fans_slip %>% 
                          mutate(SLIP_input_fans = SLIP_input_fans*1000,
                                 LMSz_fans = LMSz_fans*1000)
```

## Define x axis marks
```{r, echo = FALSE}
x_breaks <- log(c(280, 365.25*1+280, 365.25*2+280, 
                  365.25*6+280, 365.25*21+280))
x_labs <- c("0y","1y","2y","6y","21y")
```

## plotting loop function
```{r, echo = FALSE}
separate_plot_loop <- function(pheno_list, df_plot, out_fans_df, x_breaks, x_labs, mult){
  slip_plots <- list()
  lmsz_plots <- list()
    for (pheno_of_int in pheno_list) {
    df_plot$pheno <- df_plot[[glue("{pheno_of_int}.raw")]]
    df_plot$pheno[is.na(df_plot[[glue("{pheno_of_int}.lmsz")]])] <- NA
    if(grepl("global", pheno_of_int)){
      ylab_text = ""
      title_text = sub("global_", "", pheno_of_int)
    } else if(grepl("GrayVol", pheno_of_int))  {
      ylab_text = "Volume"
      title_text = sub("aparc_GrayVol_", "", pheno_of_int)
    } else if(grepl("SurfArea", pheno_of_int))  {
      ylab_text = "Surface Area"
      title_text = sub("aparc_SurfArea_", "", pheno_of_int)
    } else if(grepl("ThickAvg", pheno_of_int))  {
      ylab_text = "Thickness"
      title_text = sub("aparc_ThickAvg_", "", pheno_of_int)
    } else if(grepl("aseg", pheno_of_int))  {
      ylab_text = "Aseg/Subcortical"
      title_text = sub("aseg_", "", pheno_of_int)
    } else{
      ylab_text = pheno_of_int
      title_text = pheno_of_int
    }
    out_fans_plot <- out_fans_df %>% 
                      filter(pheno == pheno_of_int,
                             !(centile_cat %in% c("cent_0.25", "cent_0.75"))) %>%
                      mutate(SLIP_input_fans = SLIP_input_fans*mult,
                             LMSz_fans = LMSz_fans*mult)
    # CT data is questionable at 0y. Limit to 2 years
    if (grepl("Thick",pheno_of_int)) {
      df_plot <- df_plot %>% filter(age_days >= 2*365.25+280)
      out_fans_plot <- out_fans_plot %>% filter(adjusted_age_days_log >= log(2*365.25+280))
    }
    slip_plots[[pheno_of_int]] <- ggplot(out_fans_plot) +
                                      geom_point(data = df_plot, aes(y = pheno,
                                                 x = adjusted_age_days_log, fill=sex), 
                                                 alpha=0.5, shape=21, size = 0.5) +
                                      geom_line(aes(x = adjusted_age_days_log, y = SLIP_input_fans,
                                                    group = interaction(centile_cat, sex),
                                                    color = sex,
                                                    linewidth = line_size,
                                                    linetype = line_type)) +
                                      scale_linewidth_identity() +
                                      scale_linetype_identity() +
                                      scale_colour_manual(values = c("firebrick","royalblue")) +
                                      scale_x_continuous(breaks = x_breaks, labels = x_labs) +
                                      ggtitle(title_text) + theme_classic() +
                                      xlab("") + ylab(ylab_text) +
                                    theme(plot.title = element_text(size = 12),
                                          axis.title = element_text(size = 10),
                                          axis.text = element_text(size = 8))
  
    lmsz_plots[[pheno_of_int]] <- ggplot(out_fans_plot) +
                                      geom_point(data = df_plot, aes(y = pheno,
                                                 x = adjusted_age_days_log, fill=sex), 
                                                 alpha=0.5, shape=21, size = 0.5) +
                                      geom_line(aes(x = adjusted_age_days_log, y = LMSz_fans,
                                                    group = interaction(centile_cat, sex),
                                                    color = sex,
                                                    linewidth = line_size,
                                                    linetype = line_type)) +
                                      scale_linewidth_identity() +
                                      scale_linetype_identity() +
                                      scale_colour_manual(values = c("firebrick","royalblue")) +
                                      scale_x_continuous(breaks = x_breaks, labels = x_labs) +
                                      ggtitle(title_text) + theme_classic() +
                                      xlab("") + ylab(ylab_text) +
                                    theme(plot.title = element_text(size = 12),
                                          axis.title = element_text(size = 10),
                                          axis.text = element_text(size = 8))
    
    # Equalize Axes
    p_y_min <- min(layer_scales(slip_plots[[pheno_of_int]])$y$range$range[1],
                   layer_scales(lmsz_plots[[pheno_of_int]])$y$range$range[1])
    p_y_max <- max(layer_scales(slip_plots[[pheno_of_int]])$y$range$range[2],
                   layer_scales(lmsz_plots[[pheno_of_int]])$y$range$range[2])
  
    slip_plots[[pheno_of_int]] <- slip_plots[[pheno_of_int]] + ylim(p_y_min,p_y_max)
    lmsz_plots[[pheno_of_int]] <- lmsz_plots[[pheno_of_int]] + ylim(p_y_min,p_y_max)
  }
  all_plots_list <- list()
  all_plots_list$slip_plots <- slip_plots
  all_plots_list$lmsz_plots <- lmsz_plots
  return(all_plots_list)
}

overlaid_plot_loop <- function(pheno_list, df_plot, out_fans_df, x_breaks, x_labs, mult){
  overlap_plots_male <- list()
  overlap_plots_female <- list()
  overlap_plots_avg <- list()
  for (pheno_of_int in pheno_list) {
  df_plot$pheno <- df_plot[[glue("{pheno_of_int}.raw")]]
  df_plot$pheno[is.na(df_plot[[glue("{pheno_of_int}.lmsz")]])] <- NA
    if(grepl("global", pheno_of_int)){
    ylab_text = ""
    title_text = sub("global_", "", pheno_of_int)
  } else if(grepl("GrayVol", pheno_of_int))  {
    ylab_text = "Volume"
    title_text = sub("aparc_GrayVol_", "", pheno_of_int)
  } else if(grepl("SurfArea", pheno_of_int))  {
    ylab_text = "Surface Area"
    title_text = sub("aparc_SurfArea_", "", pheno_of_int)
  } else if(grepl("ThickAvg", pheno_of_int))  {
    ylab_text = "Thickness"
    title_text = sub("aparc_ThickAvg_", "", pheno_of_int)
  } else if(grepl("aseg", pheno_of_int))  {
    ylab_text = "Aseg/Subcortical"
    title_text = sub("aseg_", "", pheno_of_int)
  } else{
    ylab_text = pheno_of_int
    title_text = pheno_of_int
  }
  
  out_fans_plot <- out_fans_df %>% 
                    filter(pheno == pheno_of_int,
                           !(centile_cat %in% c("cent_0.25", "cent_0.75"))) %>%
                    mutate(SLIP_input_fans = SLIP_input_fans*mult,
                           LMSz_fans = LMSz_fans*mult)
  if (grepl("Thick",pheno_of_int)) {
    df_plot <- df_plot %>% filter(age_days >= 2*365.25+280)
    out_fans_plot <- out_fans_plot %>% filter(adjusted_age_days_log >= log(2*365.25+280))
  }
  overlap_plots_male[[pheno_of_int]] <- ggplot(out_fans_plot %>% filter(sex == "Male")) +
                                    geom_point(data = df_plot %>% filter(sex == "Male"), aes(y = pheno,
                                                 x = adjusted_age_days_log), 
                                                 alpha=0.5, shape=21, size = 0.5, color = "royalblue") +
                                    geom_line(aes(x = adjusted_age_days_log, y = SLIP_input_fans,
                                                  group = centile_cat,
                                                  color = "SLIP",
                                                  linewidth = line_size,
                                                  linetype = line_type)) +
                                    geom_line(aes(x = adjusted_age_days_log, y = LMSz_fans,
                                                  group = centile_cat,
                                                  color = "LMSz",
                                                  linewidth = line_size,
                                                  linetype = line_type)) +
                                    scale_linewidth_identity() +
                                    scale_linetype_identity() +
                                    scale_x_continuous(breaks = x_breaks, labels = x_labs) +
                                    ggtitle(title_text) + theme_classic() +
                                    xlab("") + ylab(ylab_text) +
                                    scale_colour_manual(values = c("#566fa3", "grey")) +
                                    theme(plot.title = element_text(size = 12),
                                          axis.title = element_text(size = 10),
                                          axis.text = element_text(size = 8))
  
  overlap_plots_female[[pheno_of_int]] <- ggplot(out_fans_plot %>% filter(sex == "Female")) +
                                      geom_point(data = df_plot %>% filter(sex == "Female"), aes(y = pheno,
                                                 x = adjusted_age_days_log), 
                                                 alpha=0.5, shape=21, size = 0.5, color = "firebrick") +
                                      geom_line(aes(x = adjusted_age_days_log, y = SLIP_input_fans,
                                                    group = centile_cat,
                                                    color = "SLIP",
                                                    linewidth = line_size,
                                                    linetype = line_type)) +
                                      geom_line(aes(x = adjusted_age_days_log, y = LMSz_fans,
                                                    group = centile_cat,
                                                    color = "LMSz",
                                                    linewidth = line_size,
                                                    linetype = line_type)) +
                                      scale_linewidth_identity() +
                                      scale_linetype_identity() +
                                      scale_x_continuous(breaks = x_breaks, labels = x_labs) +
                                      ggtitle(pheno_of_int) + theme_classic() +
                                      xlab("") + ylab(ylab_text) +
                                      scale_colour_manual(values = c("#E396B7", "grey")) +
                                      theme(plot.title = element_text(size = 12),
                                            axis.title = element_text(size = 10),
                                            axis.text = element_text(size = 8))
  out_fans_plot_avg <-   out_fans_plot %>% 
                          filter(grepl(str_replace_all(pheno_of_int,"SA.","SA.*"),pheno),
                                 !(centile_cat %in% c("cent_0.25", "cent_0.75"))) %>%
                          mutate(SLIP_input_fans = SLIP_input_fans*mult,
                                 LMSz_fans = LMSz_fans*mult) %>%
                          group_by(model, adjusted_age_days_log, 
                                   centile_cat, line_size, line_type) %>%
                          summarize(across(contains("fans"), ~mean(.x, na.rm = T)))
  
  overlap_plots_avg[[pheno_of_int]] <- ggplot(out_fans_plot_avg) +
                                    geom_point(data = df_plot, aes(y = pheno,
                                                 x = adjusted_age_days_log), 
                                                 alpha=0.5, shape=21, size = 0.5, color = "#5FA052") +
                                    geom_line(aes(x = adjusted_age_days_log, y = SLIP_input_fans,
                                                  group = centile_cat,
                                                  color = "SLIP",
                                                  linewidth = line_size,
                                                  linetype = line_type)) +
                                    geom_line(aes(x = adjusted_age_days_log, y = LMSz_fans,
                                                  group = centile_cat,
                                                  color = "LMSz",
                                                  linewidth = line_size,
                                                  linetype = line_type)) +
                                    scale_linewidth_identity() +
                                    scale_linetype_identity() +
                                    scale_x_continuous(breaks = x_breaks, labels = x_labs) +
                                    ggtitle(title_text) + theme_classic() +
                                    xlab("") + ylab(ylab_text) + labs(colour = "Model") +
                                    scale_colour_manual(values = c("#5FA052", "grey")) +
                                    theme(plot.title = element_text(size = 12),
                                          axis.title = element_text(size = 10),
                                          axis.text = element_text(size = 8))
  
  }
  all_plots_list <- list()
  all_plots_list$overlap_male <- overlap_plots_male
  all_plots_list$overlap_female<- overlap_plots_female
  all_plots_list$overlap_avg <- overlap_plots_avg
  return(all_plots_list)
}
```

# Trajectory Plots VPM

## Plot global trajectories
```{r, echo = FALSE, fig.show='hide'}
global_pheno_to_plot <- c("global_eTIV","global_VentricleChoroidVol","global_CortexVol",
                       "global_SubCortGrayVol","global_SurfaceArea","global_CerebralWhiteMatterVol",
                       "global_MeanThickness")

#Separate
global_plots <- list()
global_plots <- separate_plot_loop(global_pheno_to_plot, 
                                   df_vpm, 
                                   out_fans_slip, 
                                   x_breaks, 
                                   x_labs, 
                                   mult = 1)
p_slip_title <- ggdraw() + draw_label("Original (SLIP)", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_slip <- ggarrange(plots =  global_plots$slip_plots, nrow = 2, legend = "none", common.legend = T)

p_lmsz_title <- ggdraw() + draw_label("VPM LMSz", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_lmsz <- ggarrange(plots =  global_plots$lmsz_plots, nrow = 2, legend = "bottom", common.legend = T)

p_fits_global_vpm_separate <- (p_slip_title + p_slip +
             p_lmsz_title + p_lmsz) + 
          plot_layout(ncol = 2, 
                      nrow = 2, 
                      widths = c(0.07, 1)) +
         plot_annotation(title = "Global, VPM")

#Overlaid
global_plots_overlaid <- list()
global_plots_overlaid <- overlaid_plot_loop(global_pheno_to_plot, 
                                   df_vpm, 
                                   out_fans_slip, 
                                   x_breaks, 
                                   x_labs, 
                                   mult = 1)

#overlaid plots by sex
p_male_title <- ggdraw() + draw_label("Males", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_male <- ggarrange(plots =  global_plots_overlaid$overlap_male, 
                nrow = 2, legend = "none", common.legend = T)
p_female_title <- ggdraw() + draw_label("Females", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_female <- ggarrange(plots =  global_plots_overlaid$overlap_female, 
                nrow = 2, legend = "none", common.legend = T)

p_fits_global_vpm_overlaid_sex <- (p_male_title + p_male +
             p_female_title + p_female) + 
          plot_layout(ncol = 2, 
                      nrow = 2, 
                      widths = c(0.07, 1)) +
         plot_annotation(title = "Global, VPM")

#overlaid plots averaged over sex
p_traj_sexavg <- ggarrange(plots =  global_plots_overlaid$overlap_avg, 
                nrow = 2, legend = "bottom", common.legend = T)

p_traj_collapsed_title <- ggdraw() + draw_label("VPM Sex Avg Trajectory", angle = 90, x = 0.5, hjust = 0.5, fontface = "bold")


p_fits_global_vpm_overlaid_avg <- (p_traj_collapsed_title  + p_traj_sexavg) + 
          plot_layout(ncol = 2, 
                      nrow = 1, 
                      widths = c(0.07, 1)) +
         plot_annotation(title = "Global, VPM")
```

## Plot age in mu VPM trajectories
```{r, echo = FALSE, fig.show='hide'}
pheno_vpm_age_mu <- models_vpm %>% filter(grepl("age", m_model)) %>% select(feature) %>% pull(feature)
if(length(pheno_vpm_age_mu != 0)) {
#Separate plots
age_mu_vpm_plots <- list()
age_mu_vpm_plots <- separate_plot_loop(pheno_vpm_age_mu, 
                                   df_vpm, 
                                   out_fans_slip, 
                                   x_breaks, 
                                   x_labs, 
                                   mult = 1)

p_slip_title <- ggdraw() + draw_label("Original (SLIP)", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_slip <- ggarrange(plots =  age_mu_vpm_plots$slip_plots, nrow = 3, legend = "none", common.legend = T)

p_lmsz_title <- ggdraw() + draw_label("VPM LMSz", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_lmsz <- ggarrange(plots =  age_mu_vpm_plots$lmsz_plots, nrow = 3, legend = "bottom", common.legend = T)

p_fits_age_mu_vpm_separate <- (p_slip_title + p_slip +
             p_lmsz_title + p_lmsz) + 
          plot_layout(ncol = 2, 
                      nrow = 2, 
                      widths = c(0.07, 1)) +
         plot_annotation(title = "Age in Mu, VPM")


#Overlaid
age_mu_vpm_plots_overlaid <- list()
age_mu_vpm_plots_overlaid <- overlaid_plot_loop(pheno_vpm_age_mu, 
                                   df_vpm, 
                                   out_fans_slip, 
                                   x_breaks, 
                                   x_labs, 
                                   mult = 1)

#overlaid plots by sex
p_male_title <- ggdraw() + draw_label("Males", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_male <- ggarrange(plots =  age_mu_vpm_plots_overlaid$overlap_male, 
                nrow = 3, legend = "none", common.legend = T)
p_female_title <- ggdraw() + draw_label("Females", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_female <- ggarrange(plots =  age_mu_vpm_plots_overlaid$overlap_female, 
                nrow = 3, legend = "none", common.legend = T)

p_fits_age_mu_vpm_overlaid_sex <- (p_male_title + p_male +
             p_female_title + p_female) + 
          plot_layout(ncol = 2, 
                      nrow = 2, 
                      widths = c(0.07, 1)) +
         plot_annotation(title = "Age in Mu, VPM")

#overlaid plots averaged over sex
p_traj_sexavg <- ggarrange(plots =  age_mu_vpm_plots_overlaid$overlap_avg, 
                nrow = 3, legend = "bottom", common.legend = T)

p_traj_collapsed_title <- ggdraw() + draw_label("Sex Avg Trajectory", angle = 90, x = 0.5, hjust = 0.5, fontface = "bold")


p_fits_age_mu_vpm_overlaid_avg <- (p_traj_collapsed_title  + p_traj_sexavg) + 
          plot_layout(ncol = 2, 
                      nrow = 1, 
                      widths = c(0.07, 1)) +
         plot_annotation(title = "Age in Mu, VPM")

} else {
  print(glue("No phenotypes have age in mu for VPM-LMSz models: {length(pheno_vpm_age_mu)}."))
}
```

## Plot age spline in mu VPM trajectories
```{r, echo = FALSE, fig.show='hide'}
pheno_vpm_agespline_mu <- models_vpm %>% filter(grepl("ns", m_model)) %>% select(feature) %>% pull(feature)
if(length(pheno_vpm_agespline_mu != 0)) {
#Separate plots
agespline_mu_vpm_plots <- list()
agespline_mu_vpm_plots <- separate_plot_loop(pheno_vpm_agespline_mu, 
                                   df_vpm, 
                                   out_fans_slip, 
                                   x_breaks, 
                                   x_labs, 
                                   mult = 1)

p_slip_title <- ggdraw() + draw_label("Original (SLIP)", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_slip <- ggarrange(plots =  agespline_mu_vpm_plots$slip_plots, nrow = 3, legend = "none", common.legend = T)

p_lmsz_title <- ggdraw() + draw_label("VPM LMSz", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_lmsz <- ggarrange(plots =  agespline_mu_vpm_plots$lmsz_plots, nrow = 3, legend = "bottom", common.legend = T)

p_fits_agespline_mu_vpm_separate <- (p_slip_title + p_slip +
             p_lmsz_title + p_lmsz) + 
          plot_layout(ncol = 2, 
                      nrow = 2, 
                      widths = c(0.07, 1)) +
         plot_annotation(title = "Age Spline in Mu, VPM")


#Overlaid
agespline_mu_vpm_plots_overlaid <- list()
agespline_mu_vpm_plots_overlaid <- overlaid_plot_loop(pheno_vpm_agespline_mu, 
                                   df_vpm, 
                                   out_fans_slip, 
                                   x_breaks, 
                                   x_labs, 
                                   mult = 1)

#overlaid plots by sex
p_male_title <- ggdraw() + draw_label("Males", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_male <- ggarrange(plots =  agespline_mu_vpm_plots_overlaid$overlap_male, 
                nrow = 3, legend = "none", common.legend = T)
p_female_title <- ggdraw() + draw_label("Females", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_female <- ggarrange(plots =  agespline_mu_vpm_plots_overlaid$overlap_female, 
                nrow = 3, legend = "none", common.legend = T)

p_fits_agespline_mu_vpm_overlaid_sex <- (p_male_title + p_male +
             p_female_title + p_female) + 
          plot_layout(ncol = 2, 
                      nrow = 2, 
                      widths = c(0.07, 1)) +
         plot_annotation(title = "Age in Mu, VPM")

#overlaid plots averaged over sex
p_traj_sexavg <- ggarrange(plots =  agespline_mu_vpm_plots_overlaid$overlap_avg, 
                nrow = 3, legend = "bottom", common.legend = T)

p_traj_collapsed_title <- ggdraw() + draw_label("Sex Avg Trajectory", angle = 90, x = 0.5, hjust = 0.5, fontface = "bold")


p_fits_agespline_mu_vpm_overlaid_avg <- (p_traj_collapsed_title  + p_traj_sexavg) + 
          plot_layout(ncol = 2, 
                      nrow = 1, 
                      widths = c(0.07, 1)) +
         plot_annotation(title = "Age in Mu, VPM")

} else {
  print(glue("No phenotypes have age spline in mu for VPM-LMSz models: {length(pheno_vpm_agespline_mu)}."))
}
```

## Plot sex in mu VPM trajectories
```{r, echo = FALSE, fig.show='hide'}
pheno_vpm_sex_mu <- models_vpm %>% filter(grepl("sex", m_model)) %>% select(feature) %>% pull(feature)
if(length(pheno_vpm_sex_mu != 0)) {
#Separate plots
sex_mu_vpm_plots <- list()
sex_mu_vpm_plots <- separate_plot_loop(pheno_vpm_sex_mu, 
                                   df_vpm, 
                                   out_fans_slip, 
                                   x_breaks, 
                                   x_labs, 
                                   mult = 1)

p_slip_title <- ggdraw() + draw_label("Original (SLIP)", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_slip <- ggarrange(plots =  sex_mu_vpm_plots$slip_plots, nrow = 3, legend = "none", common.legend = T)

p_lmsz_title <- ggdraw() + draw_label("VPM LMSz", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_lmsz <- ggarrange(plots =  sex_mu_vpm_plots$lmsz_plots, nrow = 3, legend = "bottom", common.legend = T)

p_fits_sex_mu_vpm_separate <- (p_slip_title + p_slip +
             p_lmsz_title + p_lmsz) + 
          plot_layout(ncol = 2, 
                      nrow = 2, 
                      widths = c(0.07, 1)) +
         plot_annotation(title = "Sex in Mu, VPM")


#Overlaid
sex_mu_vpm_plots_overlaid <- list()
sex_mu_vpm_plots_overlaid <- overlaid_plot_loop(pheno_vpm_sex_mu, 
                                   df_vpm, 
                                   out_fans_slip, 
                                   x_breaks, 
                                   x_labs, 
                                   mult = 1)

#overlaid plots by sex
p_male_title <- ggdraw() + draw_label("Males", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_male <- ggarrange(plots =  sex_mu_vpm_plots_overlaid$overlap_male, 
                nrow = 3, legend = "none", common.legend = T)
p_female_title <- ggdraw() + draw_label("Females", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_female <- ggarrange(plots =  sex_mu_vpm_plots_overlaid$overlap_female, 
                nrow = 3, legend = "none", common.legend = T)

p_fits_sex_mu_vpm_overlaid_sex <- (p_male_title + p_male +
             p_female_title + p_female) + 
          plot_layout(ncol = 2, 
                      nrow = 2, 
                      widths = c(0.07, 1)) +
         plot_annotation(title = "Sex in Mu, VPM")

#overlaid plots averaged over sex
p_traj_sexavg <- ggarrange(plots =  sex_mu_vpm_plots_overlaid$overlap_avg, 
                nrow = 3, legend = "bottom", common.legend = T)

p_traj_collapsed_title <- ggdraw() + draw_label("Sex Avg Trajectory", angle = 90, x = 0.5, hjust = 0.5, fontface = "bold")


p_fits_sex_mu_vpm_overlaid_avg <- (p_traj_collapsed_title  + p_traj_sexavg) + 
          plot_layout(ncol = 2, 
                      nrow = 1, 
                      widths = c(0.07, 1)) +
         plot_annotation(title = "Sex in Mu, VPM")

} else {
  print(glue("No phenotypes have sex in mu for VPM-LMSz models: {length(pheno_vpm_sex_mu)}."))
}
```

## Plot age in sigma VPM trajectories
```{r, echo = FALSE, fig.show='hide'}
pheno_vpm_age_sigma <- models_vpm %>% filter(grepl("age", s_model)) %>% select(feature) %>% pull(feature)

if(length(pheno_vpm_age_sigma != 0)) {
  
#Separate plots
age_sigma_vpm_plots <- list()
age_sigma_vpm_plots <- separate_plot_loop(pheno_vpm_age_sigma, 
                                   df_vpm, 
                                   out_fans_slip, 
                                   x_breaks, 
                                   x_labs, 
                                   mult = 1)

p_slip_title <- ggdraw() + draw_label("Original (SLIP)", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_slip <- ggarrange(plots =  age_sigma_vpm_plots$slip_plots, nrow = 5, legend = "none", common.legend = T)

p_lmsz_title <- ggdraw() + draw_label("VPM LMSz", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_lmsz <- ggarrange(plots =  age_sigma_vpm_plots$lmsz_plots, nrow = 5, legend = "bottom", common.legend = T)

p_fits_age_sigma_vpm_separate <- (p_slip_title + p_slip +
             p_lmsz_title + p_lmsz) + 
          plot_layout(ncol = 2, 
                      nrow = 2, 
                      widths = c(0.07, 1)) +
         plot_annotation(title = "Age in Sigma, VPM")


#Overlaid
age_sigma_vpm_plots_overlaid <- list()
age_sigma_vpm_plots_overlaid <- overlaid_plot_loop(pheno_vpm_age_sigma, 
                                   df_vpm, 
                                   out_fans_slip, 
                                   x_breaks, 
                                   x_labs, 
                                   mult = 1)

#overlaid plots by sex
p_male_title <- ggdraw() + draw_label("Males", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_male <- ggarrange(plots =  age_sigma_vpm_plots_overlaid$overlap_male, 
                nrow = 5, legend = "none", common.legend = T)
p_female_title <- ggdraw() + draw_label("Females", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_female <- ggarrange(plots =  age_sigma_vpm_plots_overlaid$overlap_female, 
                nrow = 5, legend = "none", common.legend = T)

p_fits_age_sigma_vpm_overlaid_sex <- (p_male_title + p_male +
             p_female_title + p_female) + 
          plot_layout(ncol = 2, 
                      nrow = 2, 
                      widths = c(0.07, 1)) +
         plot_annotation(title = "Age in Sigma, VPM")

#overlaid plots averaged over sex
p_traj_sexavg <- ggarrange(plots =  age_sigma_vpm_plots_overlaid$overlap_avg, 
                nrow = 5, legend = "bottom", common.legend = T)

p_traj_collapsed_title <- ggdraw() + draw_label("Sex Avg Trajectory", angle = 90, x = 0.5, hjust = 0.5, fontface = "bold")


p_fits_age_sigma_vpm_overlaid_avg <- (p_traj_collapsed_title  + p_traj_sexavg) + 
          plot_layout(ncol = 2, 
                      nrow = 1, 
                      widths = c(0.07, 1)) +
         plot_annotation(title = "Age in Sigma, VPM")

} else {
  print(glue("No phenotypes have age in sigma for VPM-LMSz models: {length(pheno_vpm_age_sigma)}."))
}
```


## Plot sex in sigma VPM trajectories
```{r sex in sigma VPM, echo = FALSE, fig.show='hide'}
pheno_vpm_sex_sigma <- models_vpm %>% filter(grepl("sex", s_model)) %>% select(feature) %>% pull(feature)

if(length(pheno_vpm_sex_sigma != 0)) {
  
#Separate plots
sex_sigma_vpm_plots <- list()
sex_sigma_vpm_plots <- separate_plot_loop(pheno_vpm_sex_sigma, 
                                   df_vpm, 
                                   out_fans_slip, 
                                   x_breaks, 
                                   x_labs, 
                                   mult = 1)

p_slip_title <- ggdraw() + draw_label("Original (SLIP)", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_slip <- ggarrange(plots =  sex_sigma_vpm_plots$slip_plots, nrow = 4, legend = "none", common.legend = T)

p_lmsz_title <- ggdraw() + draw_label("VPM LMSz", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_lmsz <- ggarrange(plots =  sex_sigma_vpm_plots$lmsz_plots, nrow = 4, legend = "bottom", common.legend = T)

p_fits_sex_sigma_vpm_separate <- (p_slip_title + p_slip +
             p_lmsz_title + p_lmsz) + 
          plot_layout(ncol = 2, 
                      nrow = 2, 
                      widths = c(0.07, 1)) +
         plot_annotation(title = "Sex in Sigma, VPM")


#Overlaid
sex_sigma_vpm_plots_overlaid <- list()
sex_sigma_vpm_plots_overlaid <- overlaid_plot_loop(pheno_vpm_sex_sigma, 
                                   df_vpm, 
                                   out_fans_slip, 
                                   x_breaks, 
                                   x_labs, 
                                   mult = 1)

#overlaid plots by sex
p_male_title <- ggdraw() + draw_label("Males", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_male <- ggarrange(plots =  sex_sigma_vpm_plots_overlaid$overlap_male, 
                nrow = 4, legend = "none", common.legend = T)
p_female_title <- ggdraw() + draw_label("Females", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_female <- ggarrange(plots =  sex_sigma_vpm_plots_overlaid$overlap_female, 
                nrow = 4, legend = "none", common.legend = T)

p_fits_sex_sigma_vpm_overlaid_sex <- (p_male_title + p_male +
             p_female_title + p_female) + 
          plot_layout(ncol = 2, 
                      nrow = 2, 
                      widths = c(0.07, 1)) +
         plot_annotation(title = "Sex in Sigma, VPM")

#overlaid plots averaged over sex
p_traj_sexavg <- ggarrange(plots =  sex_sigma_vpm_plots_overlaid$overlap_avg, 
                nrow = 4, legend = "bottom", common.legend = T)

p_traj_collapsed_title <- ggdraw() + draw_label("Sex Avg Trajectory", angle = 90, x = 0.5, hjust = 0.5, fontface = "bold")


p_fits_sex_sigma_vpm_overlaid_avg <- (p_traj_collapsed_title  + p_traj_sexavg) + 
          plot_layout(ncol = 2, 
                      nrow = 1, 
                      widths = c(0.07, 1)) +
         plot_annotation(title = "Sex in Sigma, VPM")

} else {
  print(glue("No phenotypes have sex in sigma for VPM-LMSz models: {length(pheno_vpm_sex_sigma)}."))
}
```

## Plot age:sex interaction in mu VPM trajectories
```{r, echo = FALSE, fig.show='hide'}
pheno_vpm_agesex_mu <- models_vpm %>% filter(grepl("\\*", m_model)) %>% select(feature) %>% pull(feature)

if(length(pheno_vpm_agesex_mu != 0)) {
  #Separate plots
  agesex_mu_vpm_plots <- list()
  agesex_mu_vpm_plots <- separate_plot_loop(pheno_vpm_agesex_mu, 
                                     df_vpm, 
                                     out_fans_slip, 
                                     x_breaks, 
                                     x_labs, 
                                     mult = 1)
  
  p_slip_title <- ggdraw() + draw_label("Original (SLIP)", angle = 90,
                                    x = 0.5, hjust = 0.5, fontface = "bold")
  p_slip <- ggarrange(plots =  agesex_mu_vpm_plots$slip_plots, nrow = 1, legend = "none", common.legend = T)
  
  p_lmsz_title <- ggdraw() + draw_label("VPM LMSz", angle = 90,
                                    x = 0.5, hjust = 0.5, fontface = "bold")
  p_lmsz <- ggarrange(plots =  agesex_mu_vpm_plots$lmsz_plots, nrow = 1, legend = "bottom", common.legend = T)
  
  p_fits_agesex_mu_vpm_separate <- (p_slip_title + p_slip +
               p_lmsz_title + p_lmsz) + 
            plot_layout(ncol = 2, 
                        nrow = 2, 
                        widths = c(0.07, 1)) +
         plot_annotation(title = "Age*Sex in Mu, VPM")
  
  
  #Overlaid
  agesex_mu_vpm_plots_overlaid <- list()
  agesex_mu_vpm_plots_overlaid <- overlaid_plot_loop(pheno_vpm_agesex_mu, 
                                     df_vpm, 
                                     out_fans_slip, 
                                     x_breaks, 
                                     x_labs, 
                                     mult = 1)
  
  #overlaid plots by sex
  p_male_title <- ggdraw() + draw_label("Males", angle = 90,
                                    x = 0.5, hjust = 0.5, fontface = "bold")
  p_male <- ggarrange(plots =  agesex_mu_vpm_plots_overlaid$overlap_male, 
                  nrow = 1, legend = "none", common.legend = T)
  p_female_title <- ggdraw() + draw_label("Females", angle = 90,
                                    x = 0.5, hjust = 0.5, fontface = "bold")
  p_female <- ggarrange(plots =  agesex_mu_vpm_plots_overlaid$overlap_female, 
                  nrow = 1, legend = "none", common.legend = T)
  
  p_fits_agesex_mu_vpm_overlaid_sex <- (p_male_title + p_male +
               p_female_title + p_female) + 
            plot_layout(ncol = 2, 
                        nrow = 2, 
                        widths = c(0.07, 1)) +
         plot_annotation(title = "Age*Sex in Mu, VPM")
  
  #overlaid plots averaged over sex
  p_traj_sexavg <- ggarrange(plots =  agesex_mu_vpm_plots_overlaid$overlap_avg, 
                  nrow = 1, legend = "bottom", common.legend = T)
  
  p_traj_collapsed_title <- ggdraw() + draw_label("Sex Avg Trajectory", angle = 90, x = 0.5, hjust = 0.5, fontface = "bold")
  
  
  p_fits_agesex_mu_vpm_overlaid_avg <- (p_traj_collapsed_title  + p_traj_sexavg) + 
            plot_layout(ncol = 2, 
                        nrow = 1, 
                        widths = c(0.07, 1)) +
         plot_annotation(title = "Age*Sex in Mu, VPM")
} else {
  print(glue("No phenotypes have age*sex in mu for VPM-LMSz models: {length(pheno_vpm_agesex_mu)}."))
}
```

# Trajectory Plots LPM

## Plot global trajectories
```{r, echo = FALSE, fig.show='hide'}
global_pheno_to_plot <- c("global_eTIV","global_VentricleChoroidVol","global_CortexVol",
                       "global_SubCortGrayVol","global_SurfaceArea","global_CerebralWhiteMatterVol",
                       "global_MeanThickness")

#Separate
global_plots <- list()
global_plots <- separate_plot_loop(global_pheno_to_plot, 
                                   df_lpm, 
                                   out_fans_slip, 
                                   x_breaks, 
                                   x_labs, 
                                   mult = 1)
p_slip_title <- ggdraw() + draw_label("Original (SLIP)", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_slip <- ggarrange(plots =  global_plots$slip_plots, nrow = 2, legend = "none", common.legend = T)

p_lmsz_title <- ggdraw() + draw_label("LPM LMSz", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_lmsz <- ggarrange(plots =  global_plots$lmsz_plots, nrow = 2, legend = "bottom", common.legend = T)

p_fits_global_lpm_separate <- (p_slip_title + p_slip +
             p_lmsz_title + p_lmsz) + 
          plot_layout(ncol = 2, 
                      nrow = 2, 
                      widths = c(0.07, 1)) +
         plot_annotation(title = "Global, LPM")

#Overlaid
global_plots_overlaid <- list()
global_plots_overlaid <- overlaid_plot_loop(global_pheno_to_plot, 
                                   df_lpm, 
                                   out_fans_slip, 
                                   x_breaks, 
                                   x_labs, 
                                   mult = 1)

#overlaid plots by sex
p_male_title <- ggdraw() + draw_label("Males", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_male <- ggarrange(plots =  global_plots_overlaid$overlap_male, 
                nrow = 2, legend = "none", common.legend = T)
p_female_title <- ggdraw() + draw_label("Females", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_female <- ggarrange(plots =  global_plots_overlaid$overlap_female, 
                nrow = 2, legend = "none", common.legend = T)

p_fits_global_lpm_overlaid_sex <- (p_male_title + p_male +
             p_female_title + p_female) + 
          plot_layout(ncol = 2, 
                      nrow = 2, 
                      widths = c(0.07, 1)) +
         plot_annotation(title = "Global, LPM")

#overlaid plots averaged over sex
p_traj_sexavg <- ggarrange(plots =  global_plots_overlaid$overlap_avg, 
                nrow = 2, legend = "bottom", common.legend = T)

p_traj_collapsed_title <- ggdraw() + draw_label("Sex Avg Trajectory", angle = 90, x = 0.5, hjust = 0.5, fontface = "bold")


p_fits_global_lpm_overlaid_avg <- (p_traj_collapsed_title  + p_traj_sexavg) + 
          plot_layout(ncol = 2, 
                      nrow = 1, 
                      widths = c(0.07, 1)) +
         plot_annotation(title = "Global, LPM")
```

## Plot age in mu lpm trajectories
```{r, echo = FALSE, fig.show='hide'}
pheno_lpm_age_mu <- models_lpm %>% filter(grepl("age", m_model)) %>% select(feature) %>% pull(feature)
if(length(pheno_lpm_age_mu != 0)) {

  #Separate plots
  age_mu_lpm_plots <- list()
  age_mu_lpm_plots <- separate_plot_loop(pheno_lpm_age_mu, 
                                     df_lpm, 
                                     out_fans_slip, 
                                     x_breaks, 
                                     x_labs, 
                                     mult = 1)
  
  p_slip_title <- ggdraw() + draw_label("Original (SLIP)", angle = 90,
                                    x = 0.5, hjust = 0.5, fontface = "bold")
  p_slip <- ggarrange(plots =  age_mu_lpm_plots$slip_plots, nrow = 1, legend = "none", common.legend = T)
  
  p_lmsz_title <- ggdraw() + draw_label("LPM LMSz", angle = 90,
                                    x = 0.5, hjust = 0.5, fontface = "bold")
  p_lmsz <- ggarrange(plots =  age_mu_lpm_plots$lmsz_plots, nrow = 1, legend = "bottom", common.legend = T)
  
  p_fits_age_mu_lpm_separate <- (p_slip_title + p_slip +
               p_lmsz_title + p_lmsz) + 
            plot_layout(ncol = 2, 
                        nrow = 2, 
                        widths = c(0.07, 1)) +
         plot_annotation(title = "Age in Mu, LPM")
  
  
  #Overlaid
  age_mu_lpm_plots_overlaid <- list()
  age_mu_lpm_plots_overlaid <- overlaid_plot_loop(pheno_lpm_age_mu, 
                                     df_lpm, 
                                     out_fans_slip, 
                                     x_breaks, 
                                     x_labs, 
                                     mult = 1)
  
  #overlaid plots by sex
  p_male_title <- ggdraw() + draw_label("Males", angle = 90,
                                    x = 0.5, hjust = 0.5, fontface = "bold")
  p_male <- ggarrange(plots =  age_mu_lpm_plots_overlaid$overlap_male, 
                  nrow = 1, legend = "none", common.legend = T)
  p_female_title <- ggdraw() + draw_label("Females", angle = 90,
                                    x = 0.5, hjust = 0.5, fontface = "bold")
  p_female <- ggarrange(plots =  age_mu_lpm_plots_overlaid$overlap_female, 
                  nrow = 1, legend = "none", common.legend = T)
  
  p_fits_age_mu_lpm_overlaid_sex <- (p_male_title + p_male +
               p_female_title + p_female) + 
            plot_layout(ncol = 2, 
                        nrow = 2, 
                        widths = c(0.07, 1)) +
         plot_annotation(title = "Age in Mu, LPM")
  
  #overlaid plots averaged over sex
  p_traj_sexavg <- ggarrange(plots =  age_mu_lpm_plots_overlaid$overlap_avg, 
                  nrow = 3, legend = "bottom", common.legend = T)
  
  p_traj_collapsed_title <- ggdraw() + draw_label("Sex Avg Trajectory", angle = 90, x = 0.5, hjust = 0.5, fontface = "bold")
  
  
  p_fits_age_mu_lpm_overlaid_avg <- (p_traj_collapsed_title  + p_traj_sexavg) + 
            plot_layout(ncol = 2, 
                        nrow = 1, 
                        widths = c(0.07, 1)) +
         plot_annotation(title = "Age in Mu, LPM")
} else {
  print(glue("No phenotypes have age in mu for LPM-LMSz models: {length(pheno_lpm_age_mu)}."))
}
```

## Plot age spline in mu lpm trajectories
```{r age spline lpm, echo = FALSE, fig.show='hide'}
pheno_lpm_agespline_mu <- models_lpm %>% filter(grepl("ns", m_model)) %>% select(feature) %>% pull(feature)
if(length(pheno_lpm_agespline_mu != 0)) {
#Separate plots
agespline_mu_lpm_plots <- list()
agespline_mu_lpm_plots <- separate_plot_loop(pheno_lpm_agespline_mu, 
                                   df_lpm, 
                                   out_fans_slip, 
                                   x_breaks, 
                                   x_labs, 
                                   mult = 1)

p_slip_title <- ggdraw() + draw_label("Original (SLIP)", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_slip <- ggarrange(plots =  agespline_mu_lpm_plots$slip_plots, nrow = 3, legend = "none", common.legend = T)

p_lmsz_title <- ggdraw() + draw_label("LPM LMSz", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_lmsz <- ggarrange(plots =  agespline_mu_lpm_plots$lmsz_plots, nrow = 3, legend = "bottom", common.legend = T)

p_fits_agespline_mu_lpm_separate <- (p_slip_title + p_slip +
             p_lmsz_title + p_lmsz) + 
          plot_layout(ncol = 2, 
                      nrow = 2, 
                      widths = c(0.07, 1)) +
         plot_annotation(title = "Age Spline in Mu, lpm")


#Overlaid
agespline_mu_lpm_plots_overlaid <- list()
agespline_mu_lpm_plots_overlaid <- overlaid_plot_loop(pheno_lpm_agespline_mu, 
                                   df_lpm, 
                                   out_fans_slip, 
                                   x_breaks, 
                                   x_labs, 
                                   mult = 1)

#overlaid plots by sex
p_male_title <- ggdraw() + draw_label("Males", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_male <- ggarrange(plots =  agespline_mu_lpm_plots_overlaid$overlap_male, 
                nrow = 3, legend = "none", common.legend = T)
p_female_title <- ggdraw() + draw_label("Females", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_female <- ggarrange(plots =  agespline_mu_lpm_plots_overlaid$overlap_female, 
                nrow = 3, legend = "none", common.legend = T)

p_fits_agespline_mu_lpm_overlaid_sex <- (p_male_title + p_male +
             p_female_title + p_female) + 
          plot_layout(ncol = 2, 
                      nrow = 2, 
                      widths = c(0.07, 1)) +
         plot_annotation(title = "Age in Mu, LPM")

#overlaid plots averaged over sex
p_traj_sexavg <- ggarrange(plots =  agespline_mu_lpm_plots_overlaid$overlap_avg, 
                nrow = 3, legend = "bottom", common.legend = T)

p_traj_collapsed_title <- ggdraw() + draw_label("Sex Avg Trajectory", angle = 90, x = 0.5, hjust = 0.5, fontface = "bold")


p_fits_agespline_mu_lpm_overlaid_avg <- (p_traj_collapsed_title  + p_traj_sexavg) + 
          plot_layout(ncol = 2, 
                      nrow = 1, 
                      widths = c(0.07, 1)) +
         plot_annotation(title = "Age in Mu, LPM")

} else {
  print(glue("No phenotypes have age spline in mu for LPM-LMSz models: {length(pheno_lpm_agespline_mu)}."))
}
```

## Plot sex in mu lpm trajectories
```{r sex in mu lpm, echo = FALSE, fig.show='hide'}
pheno_lpm_sex_mu <- models_lpm %>% filter(grepl("sex", m_model)) %>% select(feature) %>% pull(feature)
print(length(pheno_lpm_sex_mu))
  if(length(pheno_lpm_sex_mu != 0)) {
  
  #Separate plots
  sex_mu_lpm_plots <- list()
  sex_mu_lpm_plots <- separate_plot_loop(pheno_lpm_sex_mu, 
                                     df_lpm, 
                                     out_fans_slip, 
                                     x_breaks, 
                                     x_labs, 
                                     mult = 1)
  
  p_slip_title <- ggdraw() + draw_label("Original (SLIP)", angle = 90,
                                    x = 0.5, hjust = 0.5, fontface = "bold")
  p_slip <- ggarrange(plots =  sex_mu_lpm_plots$slip_plots, nrow = 1, legend = "none", common.legend = T)
  
  p_lmsz_title <- ggdraw() + draw_label("LPM LMSz", angle = 90,
                                    x = 0.5, hjust = 0.5, fontface = "bold")
  p_lmsz <- ggarrange(plots =  sex_mu_lpm_plots$lmsz_plots, nrow = 1, legend = "bottom", common.legend = T)
  
  p_fits_sex_mu_lpm_separate <- (p_slip_title + p_slip +
               p_lmsz_title + p_lmsz) + 
            plot_layout(ncol = 2, 
                        nrow = 2, 
                        widths = c(0.07, 1)) +
         plot_annotation(title = "Sex in Mu, LPM")
  
  
  #Overlaid
  sex_mu_lpm_plots_overlaid <- list()
  sex_mu_lpm_plots_overlaid <- overlaid_plot_loop(pheno_lpm_sex_mu, 
                                     df_lpm, 
                                     out_fans_slip, 
                                     x_breaks, 
                                     x_labs, 
                                     mult = 1)
  
  #overlaid plots by sex
  p_male_title <- ggdraw() + draw_label("Males", angle = 90,
                                    x = 0.5, hjust = 0.5, fontface = "bold")
  p_male <- ggarrange(plots =  sex_mu_lpm_plots_overlaid$overlap_male, 
                  nrow = 3, legend = "none", common.legend = T)
  p_female_title <- ggdraw() + draw_label("Females", angle = 90,
                                    x = 0.5, hjust = 0.5, fontface = "bold")
  p_female <- ggarrange(plots =  sex_mu_lpm_plots_overlaid$overlap_female, 
                  nrow = 3, legend = "none", common.legend = T)
  
  p_fits_sex_mu_lpm_overlaid_sex <- (p_male_title + p_male +
               p_female_title + p_female) + 
            plot_layout(ncol = 2, 
                        nrow = 2, 
                        widths = c(0.07, 1)) +
         plot_annotation(title = "Sex in Mu, LPM")
  
  #overlaid plots averaged over sex
  p_traj_sexavg <- ggarrange(plots =  sex_mu_lpm_plots_overlaid$overlap_avg, 
                  nrow = 3, legend = "bottom", common.legend = T)
  
  p_traj_collapsed_title <- ggdraw() + draw_label("Sex Avg Trajectory", angle = 90, x = 0.5, hjust = 0.5, fontface = "bold")
  
  
  p_fits_sex_mu_lpm_overlaid_avg <- (p_traj_collapsed_title  + p_traj_sexavg) + 
            plot_layout(ncol = 2, 
                        nrow = 1, 
                        widths = c(0.07, 1)) +
         plot_annotation(title = "Sex in Mu, LPM")

  } else {
  print(glue("No phenotypes have sex in mu for LPM-LMSz models: {length(pheno_lpm_sex_mu)}."))
}
```

## Plot age in sigma lpm trajectories
```{r, echo = FALSE, fig.show='hide'}
pheno_lpm_age_sigma <- models_lpm %>% filter(grepl("age", s_model)) %>% select(feature) %>% pull(feature)

if(length(pheno_lpm_age_sigma != 0)) {

#Separate plots
age_sigma_lpm_plots <- list()
age_sigma_lpm_plots <- separate_plot_loop(pheno_lpm_age_sigma, 
                                   df_lpm, 
                                   out_fans_slip, 
                                   x_breaks, 
                                   x_labs, 
                                   mult = 1)

p_slip_title <- ggdraw() + draw_label("Original (SLIP)", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_slip <- ggarrange(plots =  age_sigma_lpm_plots$slip_plots, nrow = 3, legend = "none", common.legend = T)

p_lmsz_title <- ggdraw() + draw_label("LPM LMSz", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_lmsz <- ggarrange(plots =  age_sigma_lpm_plots$lmsz_plots, nrow = 3, legend = "bottom", common.legend = T)

p_fits_age_sigma_lpm_separate <- (p_slip_title + p_slip +
             p_lmsz_title + p_lmsz) + 
          plot_layout(ncol = 2, 
                      nrow = 2, 
                      widths = c(0.07, 1)) +
         plot_annotation(title = "Age in Sigma, LPM")


#Overlaid
age_sigma_lpm_plots_overlaid <- list()
age_sigma_lpm_plots_overlaid <- overlaid_plot_loop(pheno_lpm_age_sigma, 
                                   df_lpm, 
                                   out_fans_slip, 
                                   x_breaks, 
                                   x_labs, 
                                   mult = 1)

#overlaid plots by sex
p_male_title <- ggdraw() + draw_label("Males", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_male <- ggarrange(plots =  age_sigma_lpm_plots_overlaid$overlap_male, 
                nrow = 3, legend = "none", common.legend = T)
p_female_title <- ggdraw() + draw_label("Females", angle = 90,
                                  x = 0.5, hjust = 0.5, fontface = "bold")
p_female <- ggarrange(plots =  age_sigma_lpm_plots_overlaid$overlap_female, 
                nrow = 3, legend = "none", common.legend = T)

p_fits_age_sigma_lpm_overlaid_sex <- (p_male_title + p_male +
             p_female_title + p_female) + 
          plot_layout(ncol = 2, 
                      nrow = 2, 
                      widths = c(0.07, 1)) +
         plot_annotation(title = "Age in Sigma, LPM")

#overlaid plots averaged over sex
p_traj_sexavg <- ggarrange(plots =  age_sigma_lpm_plots_overlaid$overlap_avg, 
                nrow = 3, legend = "bottom", common.legend = T)

p_traj_collapsed_title <- ggdraw() + draw_label("Sex Avg Trajectory", angle = 90, x = 0.5, hjust = 0.5, fontface = "bold")


p_fits_age_sigma_lpm_overlaid_avg <- (p_traj_collapsed_title  + p_traj_sexavg) + 
          plot_layout(ncol = 2, 
                      nrow = 1, 
                      widths = c(0.07, 1)) +
         plot_annotation(title = "Age in Sigma, LPM")

} else {
  print(glue("No phenotypes have age in sigma for LPM-LMSz models: {length(pheno_lpm_age_sigma)}."))
}
```


## Plot sex in sigma lpm trajectories
```{r, echo = FALSE, fig.show='hide'}
pheno_lpm_sex_sigma <- models_lpm %>% filter(grepl("sex", s_model)) %>% select(feature) %>% pull(feature)

if(length(pheno_lpm_sex_sigma != 0)) {

  #Separate plots
  sex_sigma_lpm_plots <- list()
  sex_sigma_lpm_plots <- separate_plot_loop(pheno_lpm_sex_sigma, 
                                     df_lpm, 
                                     out_fans_slip, 
                                     x_breaks, 
                                     x_labs, 
                                     mult = 1)
  
  p_slip_title <- ggdraw() + draw_label("Original (SLIP)", angle = 90,
                                    x = 0.5, hjust = 0.5, fontface = "bold")
  p_slip <- ggarrange(plots =  sex_sigma_lpm_plots$slip_plots, nrow = 1, legend = "none", common.legend = T)
  
  p_lmsz_title <- ggdraw() + draw_label("LPM LMSz", angle = 90,
                                    x = 0.5, hjust = 0.5, fontface = "bold")
  p_lmsz <- ggarrange(plots =  sex_sigma_lpm_plots$lmsz_plots, nrow = 1, legend = "bottom", common.legend = T)
  
  p_fits_sex_sigma_lpm_separate <- (p_slip_title + p_slip +
               p_lmsz_title + p_lmsz) + 
            plot_layout(ncol = 2, 
                        nrow = 2, 
                        widths = c(0.07, 1)) +
         plot_annotation(title = "Sex in Sigma, LPM")
  
  
  #Overlaid
  sex_sigma_lpm_plots_overlaid <- list()
  sex_sigma_lpm_plots_overlaid <- overlaid_plot_loop(pheno_lpm_sex_sigma, 
                                     df_lpm, 
                                     out_fans_slip, 
                                     x_breaks, 
                                     x_labs, 
                                     mult = 1)
  
  #overlaid plots by sex
  p_male_title <- ggdraw() + draw_label("Males", angle = 90,
                                    x = 0.5, hjust = 0.5, fontface = "bold")
  p_male <- ggarrange(plots =  sex_sigma_lpm_plots_overlaid$overlap_male, 
                  nrow = 1, legend = "none", common.legend = T)
  p_female_title <- ggdraw() + draw_label("Females", angle = 90,
                                    x = 0.5, hjust = 0.5, fontface = "bold")
  p_female <- ggarrange(plots =  sex_sigma_lpm_plots_overlaid$overlap_female, 
                  nrow = 1, legend = "none", common.legend = T)
  
  p_fits_sex_sigma_lpm_overlaid_sex <- (p_male_title + p_male +
               p_female_title + p_female) + 
            plot_layout(ncol = 2, 
                        nrow = 2, 
                        widths = c(0.07, 1)) +
         plot_annotation(title = "Sex in Sigma, LPM")
  
  #overlaid plots averaged over sex
  p_traj_sexavg <- ggarrange(plots =  sex_sigma_lpm_plots_overlaid$overlap_avg, 
                  nrow = 1, legend = "bottom", common.legend = T)
  
  p_traj_collapsed_title <- ggdraw() + draw_label("Sex Avg Trajectory", angle = 90, x = 0.5, hjust = 0.5, fontface = "bold")
  
  
  p_fits_sex_sigma_lpm_overlaid_avg <- (p_traj_collapsed_title  + p_traj_sexavg) + 
            plot_layout(ncol = 2, 
                        nrow = 1, 
                        widths = c(0.07, 1)) +
         plot_annotation(title = "Sex in Sigma, LPM")

} else {
  print(glue("No phenotypes have sex in sigma for LPM-LMSz models: {length(pheno_lpm_sex_sigma)}."))
}
```

# Save trajectory plots
```{r, echo = FALSE, fig.height = 12, fig.width = 16}
p_fits_global_vpm_separate
p_fits_global_vpm_overlaid_sex
p_fits_global_vpm_overlaid_avg
p_fits_global_lpm_separate
p_fits_global_lpm_overlaid_sex
p_fits_global_lpm_overlaid_avg
if(save_figs == TRUE){
panel_save(p_fits_global_vpm_separate, "global_traj_slip_vpm_separate", "pdf", plot_save_path, plot_width = 12, plot_height = 8)
panel_save(p_fits_global_vpm_overlaid_sex, "global_traj_slip_vpm_overlaid_sex", "pdf", plot_save_path, plot_width = 12, plot_height = 8)
panel_save(p_fits_global_vpm_overlaid_avg, "global_traj_slip_vpm_overlaid_avg", "pdf", plot_save_path, plot_width = 12, plot_height = 8)
panel_save(p_fits_global_lpm_separate, "global_traj_slip_lpm_separate", "pdf", plot_save_path, plot_width = 12, plot_height = 8)
panel_save(p_fits_global_lpm_overlaid_sex, "global_traj_slip_lpm_overlaid_sex", "pdf", plot_save_path, plot_width = 12, plot_height = 8)
panel_save(p_fits_global_lpm_overlaid_avg, "global_traj_slip_lpm_overlaid_avg", "pdf", plot_save_path, plot_width = 12, plot_height = 8)
}

if(exists("p_fits_age_mu_vpm_separate")){
p_fits_age_mu_vpm_separate
p_fits_age_mu_vpm_overlaid_sex
p_fits_age_mu_vpm_overlaid_avg
if(save_figs == TRUE){
  panel_save(p_fits_age_mu_vpm_separate, "age_mu_traj_slip_vpm_separate", "pdf", plot_save_path, plot_width = 12, plot_height = 8)
  panel_save(p_fits_age_mu_vpm_overlaid_sex, "age_mu_traj_slip_vpm_overlaid_sex", "pdf", plot_save_path, plot_width = 12, plot_height = 8)
  panel_save(p_fits_age_mu_vpm_overlaid_avg, "age_mu_traj_slip_vpm_overlaid_avg", "pdf", plot_save_path, plot_width = 10, plot_height = 8)
}
}

if(exists("p_fits_sex_mu_vpm_separate")){
p_fits_sex_mu_vpm_separate
p_fits_sex_mu_vpm_overlaid_sex
p_fits_sex_mu_vpm_overlaid_avg
if(save_figs == TRUE){
  panel_save(p_fits_sex_mu_vpm_separate, "sex_mu_traj_slip_vpm_separate", "pdf", plot_save_path, plot_width = 6, plot_height = 8)
  panel_save(p_fits_sex_mu_vpm_overlaid_sex, "sex_mu_traj_slip_vpm_overlaid_sex", "pdf", plot_save_path, plot_width = 6, plot_height = 8)
  panel_save(p_fits_sex_mu_vpm_overlaid_avg, "sex_mu_traj_slip_vpm_overlaid_avg", "pdf", plot_save_path, plot_width = 6, plot_height = 6)
}
}

if(exists("p_fits_age_sigma_vpm_separate")){
  p_fits_age_sigma_vpm_separate
  p_fits_age_sigma_vpm_overlaid_sex
  p_fits_age_sigma_vpm_overlaid_avg
  if(save_figs == TRUE){
  panel_save(p_fits_age_sigma_vpm_separate, "age_sigma_traj_slip_vpm_separate", "pdf", plot_save_path, plot_width = 20, plot_height = 16)
panel_save(p_fits_age_sigma_vpm_overlaid_sex, "age_sigma_traj_slip_vpm_overlaid_sex", "pdf", plot_save_path, plot_width = 20, plot_height = 16)
panel_save(p_fits_age_sigma_vpm_overlaid_avg, "age_sigma_traj_slip_vpm_overlaid_avg", "pdf", plot_save_path, plot_width = 18, plot_height = 14)
  }
}


if(exists("p_fits_sex_sigma_vpm_separate")){
  p_fits_sex_sigma_vpm_separate
  p_fits_sex_sigma_vpm_overlaid_sex
  p_fits_sex_sigma_vpm_overlaid_avg
  if(save_figs == TRUE){
  panel_save(p_fits_sex_sigma_vpm_separate, "sex_sigma_traj_slip_vpm_separate", "pdf", plot_save_path, plot_width = 18, plot_height = 14)
  panel_save(p_fits_sex_sigma_vpm_overlaid_sex, "sex_sigma_traj_slip_vpm_overlaid_sex", "pdf", plot_save_path, plot_width = 18, plot_height = 14)
  panel_save(p_fits_sex_sigma_vpm_overlaid_avg, "sex_sigma_traj_slip_vpm_overlaid_avg", "pdf", plot_save_path, plot_width = 18, plot_height = 12)
  }
}

if(exists("p_fits_agesex_mu_vpm_separate")){
print(p_fits_agesex_mu_vpm_separate)
print(p_fits_agesex_mu_vpm_overlaid_sex)
print(p_fits_agesex_mu_vpm_overlaid_avg)
if(save_figs == TRUE){
  panel_save(p_fits_agesex_mu_vpm_separate, "agesex_mu_traj_slip_vpm_separate", "pdf", plot_save_path, plot_width = 6, plot_height = 4)
  panel_save(p_fits_agesex_mu_vpm_overlaid_sex, "agesex_mu_traj_slip_vpm_overlaid_sex", "pdf", plot_save_path, plot_width = 6, plot_height = 4)
  panel_save(p_fits_agesex_mu_vpm_overlaid_avg, "agesex_mu_traj_slip_vpm_overlaid_avg", "pdf", plot_save_path, plot_width = 4, plot_height = 4)
  }
}

if(exists("p_fits_agespline_mu_vpm_separate")){
print(p_fits_agespline_mu_vpm_separate)
print(p_fits_agespline_mu_vpm_overlaid_sex)
print(p_fits_agespline_mu_vpm_overlaid_avg)
if(save_figs == TRUE){
  panel_save(p_fits_agespline_mu_vpm_separate, "agespline_mu_traj_slip_vpm_separate", "pdf", plot_save_path, plot_width = 6, plot_height = 4)
  panel_save(p_fits_agespline_mu_vpm_overlaid_sex, "agespline_mu_traj_slip_vpm_overlaid_sex", "pdf", plot_save_path, plot_width = 6, plot_height = 4)
  panel_save(p_fits_agespline_mu_vpm_overlaid_avg, "agespline_mu_traj_slip_vpm_overlaid_avg", "pdf", plot_save_path, plot_width = 4, plot_height = 4)
  }
}

if(exists("p_fits_age_mu_lpm_separate")){
print(p_fits_age_mu_lpm_separate)
print(p_fits_age_mu_lpm_overlaid_sex)
print(p_fits_age_mu_lpm_overlaid_avg)
if(save_figs == TRUE){
  panel_save(p_fits_age_mu_lpm_separate, "age_mu_traj_slip_lpm_separate", "pdf", plot_save_path, plot_width = 12, plot_height = 8)
  panel_save(p_fits_age_mu_lpm_overlaid_sex, "age_mu_traj_slip_lpm_overlaid_sex", "pdf", plot_save_path, plot_width = 12, plot_height = 8)
  panel_save(p_fits_age_mu_lpm_overlaid_avg, "age_mu_traj_slip_lpm_overlaid_avg", "pdf", plot_save_path, plot_width = 12, plot_height = 8)
  }
}

if(exists("p_fits_agespline_mu_lpm_separate")){
print(p_fits_agespline_mu_lpm_separate)
print(p_fits_agespline_mu_lpm_overlaid_sex)
print(p_fits_agespline_mu_lpm_overlaid_avg)
if(save_figs == TRUE){
  panel_save(p_fits_agespline_mu_lpm_separate, "agespline_mu_traj_slip_lpm_separate", "pdf", plot_save_path, plot_width = 6, plot_height = 4)
  panel_save(p_fits_agespline_mu_lpm_overlaid_sex, "agespline_mu_traj_slip_lpm_overlaid_sex", "pdf", plot_save_path, plot_width = 6, plot_height = 4)
  panel_save(p_fits_agespline_mu_lpm_overlaid_avg, "agespline_mu_traj_slip_lpm_overlaid_avg", "pdf", plot_save_path, plot_width = 4, plot_height = 4)
  }
}

if(exists("p_fits_sex_mu_lpm_separate")){
print(p_fits_sex_mu_lpm_separate)
print(p_fits_sex_mu_lpm_overlaid_sex)
print(p_fits_sex_mu_lpm_overlaid_avg)
if(save_figs == TRUE){
  panel_save(p_fits_sex_mu_lpm_separate, "sex_mu_traj_slip_lpm_separate", "pdf", plot_save_path, plot_width = 12, plot_height = 8)
  panel_save(p_fits_sex_mu_lpm_overlaid_sex, "sex_mu_traj_slip_lpm_overlaid_sex", "pdf", plot_save_path, plot_width = 12, plot_height = 8)
  panel_save(p_fits_sex_mu_lpm_overlaid_avg, "sex_mu_traj_slip_lpm_overlaid_avg", "pdf", plot_save_path, plot_width = 12, plot_height = 8)
  }
}

if(exists("p_fits_age_sigma_lpm_separate")){
print(p_fits_age_sigma_lpm_separate)
print(p_fits_age_sigma_lpm_overlaid_sex)
print(p_fits_age_sigma_lpm_overlaid_avg)
if(save_figs == TRUE){
  panel_save(p_fits_age_sigma_lpm_separate, "age_sigma_traj_slip_lpm_separate", "pdf", plot_save_path, plot_width = 14, plot_height = 8)
  panel_save(p_fits_age_sigma_lpm_overlaid_sex, "age_sigma_traj_slip_lpm_overlaid_sex", "pdf", plot_save_path, plot_width = 14, plot_height = 8)
  panel_save(p_fits_age_sigma_lpm_overlaid_avg, "age_sigma_traj_slip_lpm_overlaid_avg", "pdf", plot_save_path, plot_width = 16, plot_height = 8)
  }
}

if(exists("p_fits_sex_sigma_lpm_separate")){
print(p_fits_sex_sigma_lpm_separate)
print(p_fits_sex_sigma_lpm_overlaid_sex)
print(p_fits_sex_sigma_lpm_overlaid_avg)
if(save_figs == TRUE){
  panel_save(p_fits_sex_sigma_lpm_separate, "sex_sigma_traj_slip_lpm_separate", "pdf", plot_save_path, plot_width = 16, plot_height = 8)
  panel_save(p_fits_sex_sigma_lpm_overlaid_sex, "sex_sigma_traj_slip_lpm_overlaid_sex", "pdf", plot_save_path, plot_width = 16, plot_height = 8)
  panel_save(p_fits_sex_sigma_lpm_overlaid_avg, "sex_sigma_traj_slip_lpm_overlaid_avg", "pdf", plot_save_path, plot_width = 18, plot_height = 8)
  }
}
```

# Plot bigger fits (for visibility)
```{r, echo = FALSE, fig.height = 16, fig.width = 20}
p_fits_age_sigma_vpm_separate
p_fits_age_sigma_vpm_overlaid_sex
p_fits_age_sigma_vpm_overlaid_avg

p_fits_sex_sigma_vpm_separate
p_fits_sex_sigma_vpm_overlaid_sex
p_fits_sex_sigma_vpm_overlaid_avg
```