---
title: "Spline Analysis: Covariate-Structure Comparison"
author: "Eren Kafadar"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
params:
  output_folder: "/mnt/arcus/lab/users/kafadare/swyc_premie_analysis_results/all_variants_consolidated/"
  lookup_stats: "/mnt/arcus/lab/users/kafadare/slip_centiles_processed_tables/lookup_stats_distinct.RData"
  vars_path: "/mnt/arcus/lab/users/kafadare/slip_brain_vars/"
  code_path: "/mnt/arcus/lab/users/kafadare/slip_premie_wip/"
  variant_to_visualize: "eTIV"
editor_options:
  chunk_output_type: console
---

Reads the merged long-format outputs of `run_spline_analysis_variant.R` +
`merge_spline_analysis_variants.R` and compares partial R2 / RESI across
covariate-structure variants. Does not fit any models itself.

# Define params if they do not exist
```{r params, echo = FALSE}
if (!exists("params")) {
  params <- list(
    output_folder = "/mnt/arcus/lab/users/kafadare/swyc_premie_analysis_results/all_variants_consolidated/",
    lookup_stats = "/mnt/arcus/lab/users/kafadare/slip_centiles_processed_tables/lookup_stats_distinct.RData",
    vars_path = "/mnt/arcus/lab/users/kafadare/slip_brain_vars/",
    code_path = "/mnt/arcus/lab/users/kafadare/slip_premie_wip/",
    variant_to_visualize = "eTIV"
  )
}
print(params)
```

# Load packages
```{r packages, echo = FALSE}
require(dplyr)
require(glue)
require(purrr)
require(tibble)
require(tidyr)
require(ggplot2)
require(patchwork)
require(ggseg)
require(ggsegDKT)
require(scales)

knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

# Load functions
```{r fcts, echo = FALSE}
# Model/interaction curve visualization -- deliberately model_vis_fcts.R only.
# vis_spline_noint_EK.R / visualize_model_bartfct.R define near-identical
# visualize_spline_model()/visualize_model() functions (used by the original
# slipga_spline_analysis_*.Rmd files) but are NOT sourced here.
source(glue("{params$code_path}model_vis_fcts.R"))

# plot_grouped_lollipop(), plot_brain_map()
source(glue("{params$code_path}plotting_utils.R"))
```

# Paths
```{r paths, echo = TRUE}
stats_folder  <- paste0(params$output_folder, "results_csv/")
models_folder <- paste0(params$output_folder, "models/")

plot_output_folder <- paste0(params$output_folder, "comparison_plots/")
if (!dir.exists(plot_output_folder)) dir.create(plot_output_folder, recursive = TRUE)
```

# Load data
```{r load-data, echo = TRUE}
main_stats <- read.csv(file.path(stats_folder, "main_effects_stats_all_variants.csv"))
resi_stats <- read.csv(file.path(stats_folder, "resi_stats_all_variants.csv"))
int_stats  <- read.csv(file.path(stats_folder, "interaction_stats_all_variants.csv"))

all_models <- readRDS(file.path(models_folder, "all_variants_models.rds"))

load(params$lookup_stats)  # mean_lookup_distinct / sd_lookup_distinct, for the GA axis conversion
namemap <- read.csv(glue("{params$vars_path}name_mapping.csv"))
```

# Color Specifications
```{r colors, echo = TRUE}
# ga_color / bwp_color match the base hues used in the original spline Rmds,
# with the mislabeling corrected there -- bw_color ("#709085") was being reused
# for BWP plots even though a distinct bwp_color existed. Only GA and BWP are
# plotted going forward (birth_weight_kg was dropped as a predictor), so
# bw_color isn't used here at all.
ga_color  <- "#6D8299"  # muted blue-grey  -- Gestational Age
bwp_color <- "#A68A64"  # muted tan/gold   -- Birth Weight Percentile

# One color family per predictor, with a distinct, consistent sub-color per
# covariate_group variant (same variant = same shade everywhere in this
# document, regardless of which pair it's being compared against).
variant_order <- c("nocovar", "eTIV", "ADI", "eTIV_ADI", "wh", "eTIV_wh",
                    "neonatal_dx", "eTIV_neonatal_dx")

make_variant_family <- function(base_color, variants) {
  setNames(colorRampPalette(c("#F0F0F0", base_color))(length(variants)), variants)
}

predictor_colors <- list(
  gestational_age = make_variant_family(ga_color, variant_order),
  bwp_fen         = make_variant_family(bwp_color, variant_order)
)
```

# Covariate-structure comparison pairs
```{r pairs, echo = TRUE}
# Each pair is (reference, +extra-covariate). Extend this list if new variants
# are added to run_spline_analysis_variant.R.
covariate_pairs <- list(
  c("nocovar", "eTIV"),
  c("nocovar", "ADI"),
  c("nocovar", "wh"),
  c("nocovar", "neonatal_dx"),
  c("eTIV", "eTIV_ADI"),
  c("eTIV", "eTIV_wh"),
  c("eTIV", "eTIV_neonatal_dx")
)

predictors <- c("gestational_age", "bwp_fen")

global_groups <- c("global_self", "global_other", "global_meanthick")
cmd_groups    <- c("global_cmd", "vol_cmd", "sa_cmd", "th_cmd", "subcort_cmd")
```

# Axis convention helpers (kept as in the original spline Rmds)
```{r axis-helpers, echo = TRUE}
ga_x_scale <- function() {
  scale_x_continuous(
    breaks = map2_dbl(c(24*7, 28*7, 32*7, 36*7, 40*7), "gestational_age",
                       ~ (.x - mean_lookup_distinct[[.y]]) / sd_lookup_distinct[[.y]]),
    labels = \(x) round(map2_dbl(x, "gestational_age",
                       ~ (.x * sd_lookup_distinct[[.y]]) + mean_lookup_distinct[[.y]]) / 7, 1)
  )
}

bwp_x_scale <- function() {
  scale_x_continuous(breaks = qnorm(c(.01, .50, .99)), labels = c("<=1st", "50th", ">=99th"))
}

outcome_y_scale <- function() {
  scale_y_continuous(breaks = qnorm(c(.01, .50, .99)), labels = c("<=1st", "50th", ">=99th"))
}

predictor_x_scale <- function(predictor) {
  if (predictor == "gestational_age") ga_x_scale() else bwp_x_scale()
}

predictor_label <- function(predictor) {
  if (predictor == "gestational_age") "Gestational Age (weeks)" else "Birth Weight Percentile\n(GA-corrected)"
}
```

# ggseg label helper
```{r ggseg-label-fct, echo = TRUE}
# plot_brain_map() joins on a `label` column matching ggseg::dk/ggseg::aseg,
# but main_stats/resi_stats only carry the raw FreeSurfer `phenotype` name
# (run_spline_analysis_variant.R doesn't add a label column). Reproduces the
# per-region-type cleanup the original Rmds applied right after stat
# extraction (e.g. `ga_sp_table_vol$label <- rownames(ga_sp_table_vol) %>% ...`).
add_ggseg_label <- function(df, region_type = c("vol", "sa", "th")) {
  region_type <- match.arg(region_type)

  if (region_type == "vol") {
    cols_edit <- grepl("aseg(?!_CC)", df$phenotype, perl = TRUE)
    df$label <- df$phenotype %>%
      sub("aparc_GrayVol_", "", .) %>%
      sub("aseg_", "", .) %>%
      sub("X", "x", .) %>%
      sub("rd_Ventricle", "rd_ventricle", .) %>%
      sub("th_Ventricle", "th_ventricle", .) %>%
      sub("Thalamus", "Thalamus-Proper", .)
    df$label[cols_edit] <- gsub("_", "-", df$label[cols_edit])
  } else if (region_type == "sa") {
    df$label <- sub("aparc_SurfArea_", "", df$phenotype)
  } else {
    df$label <- sub("aparc_ThickAvg_", "", df$phenotype)
  }

  df
}
```

# Grouped Lollipop Plots (Global & CMD, Partial R2 & RESI)
```{r lollipop-fcts, echo = TRUE}
make_pair_table_list <- function(df, this_predictor, groups, pair) {
  sub <- df %>% filter(predictor == this_predictor, group %in% groups, variant %in% pair)
  split(sub, sub$variant)[pair]
}

plot_pair_lollipop <- function(df, this_predictor, groups, pair, value_col, p_col, set_label) {
  table_list <- make_pair_table_list(df, this_predictor, groups, pair)

  plot_grouped_lollipop(
    table_list = table_list,
    value_col = value_col,
    p_col = p_col,
    color_vec = predictor_colors[[this_predictor]][pair],
    name_mapping = namemap
  ) +
    labs(
      title = glue("{set_label} -- {predictor_label(this_predictor)}: {pair[1]} vs {pair[2]}"),
      x = value_col
    ) +
    theme_minimal()
}
```

```{r lollipop-plots, results = 'asis', fig.width = 9, fig.height = 6}
for (this_predictor in predictors) {
  for (pair in covariate_pairs) {

    print(plot_pair_lollipop(main_stats, this_predictor, global_groups, pair,
                              "r2_full_vs_none", "full_p_adj", "Global vars (Partial R2)"))
    print(plot_pair_lollipop(resi_stats, this_predictor, global_groups, pair,
                              "RESI", "p_adj", "Global vars (RESI)"))
    print(plot_pair_lollipop(main_stats, this_predictor, cmd_groups, pair,
                              "r2_full_vs_none", "full_p_adj", "CMD (Partial R2)"))
    print(plot_pair_lollipop(resi_stats, this_predictor, cmd_groups, pair,
                              "RESI", "p_adj", "CMD (RESI)"))
  }
}
```

# Brain Maps (Vol/SA/TH, Partial R2 & RESI, GA top / BWP bottom)
```{r brainmap-fcts, echo = TRUE}
# Difference map: pair[2]'s value minus pair[1]'s value, restricted to regions
# significant (p < .05) in at least one of the two variants being compared.
build_diff_table <- function(df, this_predictor, group, pair, value_col, p_col, region_type) {
  d1 <- df %>% filter(predictor == this_predictor, group == !!group, variant == pair[1])
  d2 <- df %>% filter(predictor == this_predictor, group == !!group, variant == pair[2])

  merged <- inner_join(
    d1 %>% select(phenotype, val1 = all_of(value_col), p1 = all_of(p_col)),
    d2 %>% select(phenotype, val2 = all_of(value_col), p2 = all_of(p_col)),
    by = "phenotype"
  )

  merged$diff  <- merged$val2 - merged$val1
  merged$min_p <- pmin(merged$p1, merged$p2)

  add_ggseg_label(merged, region_type)
}

plot_pair_brainmap <- function(df, this_predictor, group, pair, value_col, p_col, region_type,
                               limits, low_color = "#2160B8", mid_color = "white", high_color = "#B82133") {
  diff_table <- build_diff_table(df, this_predictor, group, pair, value_col, p_col, region_type)

  plot_brain_map(
    table = diff_table,
    value_col = "diff",
    atlas = "dk",
    p_col = "min_p",
    p_filter = TRUE,
    p_thresh = 0.05,
    title = glue("{region_type} {value_col} diff: {pair[2]} - {pair[1]}"),
    limits = limits,
    low_color = low_color, mid_color = mid_color, high_color = high_color
  )
}

# GA (top) / BWP (bottom) arrangement is just row-major wrap_plots with GA's
# plots listed first, matching how the original Rmds built these grids.
plot_pair_brainmap_grid <- function(df, group, pair, value_col, p_col, region_type, limits) {
  plots <- lapply(predictors, function(p) {
    plot_pair_brainmap(df, p, group, pair, value_col, p_col, region_type, limits) +
      labs(subtitle = predictor_label(p))
  })
  wrap_plots(plots, ncol = 1) & plot_layout(guides = "collect")
}
```

```{r brainmap-plots, results = 'asis', fig.width = 6, fig.height = 10}
region_types <- c(vol = "vol", sa = "sa", th = "th")

for (pair in covariate_pairs) {
  for (region_type in region_types) {

    print(
      plot_pair_brainmap_grid(main_stats, region_type, pair,
                               "r2_full_vs_none", "full_p_adj", region_type,
                               limits = c(-0.025, 0.025))
    )
    print(
      plot_pair_brainmap_grid(resi_stats, region_type, pair,
                               "RESI", "p_adj", region_type,
                               limits = c(-0.2, 0.2))
    )
  }
}
```

# Model Visualization
```{r modelvis-fcts, echo = TRUE}
get_main_model <- function(variant, this_predictor, group, phenotype) {
  run_key <- paste(this_predictor, group, sep = "_")
  all_models[[variant]][[run_key]]$main[[phenotype]]$full
}

plot_main_model <- function(variant, this_predictor, group, phenotype) {
  visualize_spline_model(
    modobj = get_main_model(variant, this_predictor, group, phenotype),
    smooth_var = this_predictor,
    plot_og_scale = FALSE,
    mean_lookup = mean_lookup_distinct,
    sd_lookup = sd_lookup_distinct,
    pnorm_smooth_var = FALSE,
    plabels = phenotype,
    color_point = if (this_predictor == "gestational_age") ga_color else bwp_color,
    color_line  = if (this_predictor == "gestational_age") ga_color else bwp_color
  ) +
    predictor_x_scale(this_predictor) +
    outcome_y_scale() +
    labs(x = predictor_label(this_predictor), title = phenotype) +
    theme(axis.title.y = element_blank())
}
```

## Significant global & CMD models
```{r modelvis-global-cmd, results = 'asis', fig.width = 10, fig.height = 8}
variant_to_visualize <- params$variant_to_visualize

for (this_predictor in predictors) {
  sig_pheno <- main_stats %>%
    filter(variant == variant_to_visualize, predictor == this_predictor,
           group %in% c(global_groups, cmd_groups), full_p_adj < 0.05)

  if (nrow(sig_pheno) == 0) next

  plots <- Map(function(ph, grp) plot_main_model(variant_to_visualize, this_predictor, grp, ph),
               sig_pheno$phenotype, sig_pheno$group)

  print(wrap_plots(plots, ncol = 2) & plot_layout(guides = "collect"))
}
```

## Top 5 RESI regions within vol / sa / th
```{r modelvis-top5-resi, results = 'asis', fig.width = 10, fig.height = 8}
for (this_predictor in predictors) {
  for (region_type in region_types) {

    top5 <- resi_stats %>%
      filter(variant == variant_to_visualize, predictor == this_predictor,
             group == region_type, p_adj < 0.05) %>%
      arrange(desc(RESI)) %>%
      head(5)

    if (nrow(top5) == 0) next

    plots <- lapply(top5$phenotype, function(ph) plot_main_model(variant_to_visualize, this_predictor, region_type, ph))
    print(wrap_plots(plots, ncol = 2) & plot_layout(guides = "collect"))
  }
}
```

## Significant interactions
```{r modelvis-sig-fct, echo = TRUE}
# `int_stats` is the union of columns across all three interaction_types
# (bind_rows() fills whatever a given row's type didn't produce with NA), so
# a row is "significant" if:
#  - continuous: its own int_p_adj < .05
#  - ordered:    ANY of its int1_p_adj, int2_p_adj, ... (per-level difference
#                smooths vs. the reference level) is < .05
#  - categorical: ANY of its per-level <level>_p_adj columns is < .05 -- there
#                is no int_p_adj/pred_p_adj at all for categorical rows (no
#                bare s(predictor) or ti() term exists in that formula), so
#                those two columns are excluded from the level-column search
#                to avoid a name collision with a moderator level literally
#                called "int1" or "pred" (the known tradeoff of matching by
#                column name instead of a stored, unambiguous type tag)
p_adj_cols     <- grep("_p_adj$", names(int_stats), value = TRUE)
ordered_cols   <- grep("^int[0-9]+_p_adj$", p_adj_cols, value = TRUE)
categorical_cols <- setdiff(p_adj_cols, c("pred_p_adj", "int_p_adj", ordered_cols))

any_below <- function(df, cols, thresh = 0.05) {
  if (length(cols) == 0) return(rep(FALSE, nrow(df)))
  mat <- as.matrix(df[, cols, drop = FALSE])
  apply(mat < thresh, 1, any, na.rm = TRUE)
}

int_stats$any_sig <- NA
cont_idx <- int_stats$interaction_type == "continuous"
ord_idx  <- int_stats$interaction_type == "ordered"
cat_idx  <- int_stats$interaction_type == "categorical"

int_stats$any_sig[cont_idx] <- int_stats$int_p_adj[cont_idx] < 0.05
int_stats$any_sig[ord_idx]  <- any_below(int_stats[ord_idx, ], ordered_cols)
int_stats$any_sig[cat_idx]  <- any_below(int_stats[cat_idx, ], categorical_cols)
```

```{r modelvis-interactions, results = 'asis', fig.width = 9, fig.height = 6}
# visualize_model_int() already dispatches on the moderator's actual R class
# (factor/ordered/numeric), so the same plotting call below works unchanged
# across all three interaction_types -- only the significance filter above
# needed to branch.
sig_int <- int_stats %>%
  filter(variant == variant_to_visualize, any_sig)

for (i in seq_len(nrow(sig_int))) {
  row <- sig_int[i, ]
  run_key <- paste(row$predictor, row$group, sep = "_")
  modobj <- all_models[[row$variant]][[run_key]]$interactions[[row$moderator_label]][[row$phenotype]]

  p <- visualize_model_int(
    modobj = modobj,
    smooth_var = row$predictor,
    int_var = row$moderator,
    plabels = glue("{row$phenotype} x {row$moderator_label}"),
    derivative_plot = FALSE
  ) +
    predictor_x_scale(row$predictor) +
    outcome_y_scale() +
    labs(x = predictor_label(row$predictor))

  print(p)
}
```
