---
title: "Spline Analysis: Covariate-Structure Comparison"
author: "Eren Kafadar"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
params:
  output_folder: "/mnt/arcus/lab/users/kafadare/slip_premie_results_fxfalse/"
  lookup_stats: "/mnt/arcus/lab/users/kafadare/slip_centiles_processed_tables/lookup_stats_distinct.RData"
  vars_path: "/mnt/arcus/lab/users/kafadare/slip_brain_vars/"
  code_path: "/mnt/arcus/lab/users/kafadare/slip_premie_wip/"
  cent_path: "/mnt/arcus/lab/users/kafadare/slip_centiles_processed_tables/df_zscore_distinct.csv"
  model_vis: TRUE
  save_plots: TRUE
  variant_to_visualize: "eTIV"
editor_options:
  chunk_output_type: console
---

Reads the merged long-format outputs of `run_spline_analysis_variant.R` +
`merge_spline_analysis_variants.R` and compares partial R2 / RESI across
covariate-structure variants. Does not fit any models itself.

# Define params if they do not exist
```{r params, echo = FALSE}
if (!exists("params")) {
  params <- list(
    output_folder = "/mnt/arcus/lab/users/kafadare/slip_premie_results_fxfalse/",
    lookup_stats = "/mnt/arcus/lab/users/kafadare/slip_centiles_processed_tables/lookup_stats_distinct.RData",
    vars_path = "/mnt/arcus/lab/users/kafadare/slip_brain_vars/",
    code_path = "/mnt/arcus/lab/users/kafadare/slip_premie_wip/",
    cent_path = "/mnt/arcus/lab/users/kafadare/slip_centiles_processed_tables/df_zscore_distinct.csv",
    model_vis = FALSE,
    save_plots = TRUE,
    variant_to_visualize = "eTIV"
  )
}
print(params)
```

# Load packages
```{r packages, echo = FALSE}
require(dplyr)
require(glue)
require(purrr)
require(tibble)
require(tidyr)
require(ggplot2)
require(patchwork)
require(ggseg)
require(ggsegDKT)
require(scales)
require(svglite)  # needed by ggsave() for .svg output

knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

# Load functions
```{r fcts, echo = FALSE}
source(glue("{params$code_path}model_vis_fcts.R"))
source(glue("{params$code_path}plotting_utils.R"))
```

# Paths
```{r paths, echo = TRUE}
stats_folder  <- paste0(params$output_folder, "results_csv/")
models_folder <- paste0(params$output_folder, "models/")

plot_output_folder <- paste0(params$output_folder, "plots/")
if (!dir.exists(plot_output_folder)) dir.create(plot_output_folder, recursive = TRUE)

spin_output_folder <- paste0(params$output_folder, "spin_tables/")
if (!dir.exists(spin_output_folder)) dir.create(spin_output_folder, recursive = TRUE)
```

# Load data
```{r load-data, echo = TRUE}
main_stats <- read.csv(paste0(stats_folder, "main_effects_stats_all_variants.csv"))
resi_stats <- read.csv(paste0(stats_folder, "resi_stats_all_variants.csv"))
int_stats  <- read.csv(paste0(stats_folder, "interaction_stats_all_variants.csv"))

if (params$model_vis == TRUE){
  variant_to_visualize <- params$variant_to_visualize
  models_to_load <- paste0(models_folder, paste0(variant_to_visualize, "_models.rds"))
  models <- readRDS(models_to_load)
}

load(params$lookup_stats)  # mean_lookup_distinct / sd_lookup_distinct, for the GA axis conversion
namemap <- read.csv(glue("{params$vars_path}name_mapping.csv"))
df.zscore <- read.csv(params$cent_path)
```

# Plotting Specifications
```{r colors, echo = TRUE}
# One dark + one medium-dark hue per predictor -- replaces the old per-variant
# color ramp (make_variant_family()/predictor_colors), which was only ever
# used to shade a pair of things being compared. Two fixed shades per
# predictor cover that: the "reference" item in a pair gets the medium shade,
# the "extra"/emphasis item gets the dark shade.
predictor_hues <- list(
  gestational_age = c(dark = "#0B3D42", medium = "#3E7C89"),  # dark / medium petrol blue
  bwp_fen         = c(dark = "#1B4332", medium = "#52796F")   # dark / medium forest green
)

# Kept for the Model Visualization section, which colors points/lines by
# predictor directly (not through a pair comparison) -- single source of
# truth now instead of independently-set constants.
ga_color  <- predictor_hues$gestational_age[["dark"]]
bwp_color <- predictor_hues$bwp_fen[["dark"]]
bw_color = "green" # muted green -- Birth Weight (birth_weight_kg dropped as a predictor; unused)

# For lollipop plots colored by effect direction (average first derivative
# sign) instead of by predictor/variant identity.
deriv_pos_color <- "#B85C5C"  # muted red  -- positive average derivative
deriv_neg_color <- "#5C7CB8"  # muted blue -- negative average derivative

font_size <- 16
theme_set(theme_classic(base_family = "Helvetica",base_size = font_size))
line_size <- 1.5
point_size <- 2

# limits = c(-0.3, 0.3)
# low_color = "#4878CF" # blue
# mid_color = "white"
# high_color = "#D65F5F" # red
```

# Covariate-structure comparison pairs
```{r pairs, echo = TRUE}
# Each pair is (reference, +extra-covariate). Extend this list if new variants
# are added to run_spline_analysis_variant.R.
covariate_pairs <- list(
  c("nocovar", "eTIV"),
  #c("nocovar", "ADI"),
  #c("nocovar", "wh"),
  c("eTIV", "eTIV_ADI"),
  c("eTIV", "eTIV_wh")
)

predictors <- c("gestational_age", "bwp_fen")

global_groups <- c("global_self", "global_other", "global_meanthick")
cmd_groups    <- c("global_cmd", "vol_cmd", "sa_cmd", "th_cmd", "subcort_cmd")
```

# Axis convention helpers
```{r axis-helpers, echo = TRUE}
ga_x_scale <- function() {
  scale_x_continuous(
    breaks = map2_dbl(c(24*7, 28*7, 32*7, 36*7, 40*7), "gestational_age",
                       ~ (.x - mean_lookup_distinct[[.y]]) / sd_lookup_distinct[[.y]]),
    labels = \(x) round(map2_dbl(x, "gestational_age",
                       ~ (.x * sd_lookup_distinct[[.y]]) + mean_lookup_distinct[[.y]]) / 7, 1)
  )
}

bwp_x_scale <- function() {
  scale_x_continuous(breaks = qnorm(c(.01, .50, .99)), labels = c("<=1st", "50th", ">=99th"))
}

outcome_y_scale <- function() {
  scale_y_continuous(breaks = qnorm(c(.01, .50, .99)), labels = c("<=1st", "50th", ">=99th"))
}

predictor_x_scale <- function(predictor) {
  if (predictor == "gestational_age") ga_x_scale() else bwp_x_scale()
}

predictor_label <- function(predictor) {
  if (predictor == "gestational_age") "Gestational Age (weeks)" else "Birth Weight Percentile\n(GA-corrected)"
}
```

# ggseg label helper
```{r ggseg-label-fct, echo = TRUE}
# plot_brain_map() joins on a `label` column matching ggseg::dk/ggseg::aseg,
# but main_stats/resi_stats only carry the raw FreeSurfer `phenotype` name
# (run_spline_analysis_variant.R doesn't add a label column). Reproduces the
# per-region-type cleanup the original Rmds applied right after stat
# extraction (e.g. `ga_sp_table_vol$label <- rownames(ga_sp_table_vol) %>% ...`).
add_ggseg_label <- function(df, region_type = c("vol", "sa", "th")) {
  region_type <- match.arg(region_type)

  if (region_type == "vol") {
    cols_edit <- grepl("aseg(?!_CC)", df$phenotype, perl = TRUE)
    df$label <- df$phenotype %>%
      sub("aparc_GrayVol_", "", .) %>%
      sub("aseg_", "", .) %>%
      sub("X", "x", .) %>%
      sub("rd_Ventricle", "rd_ventricle", .) %>%
      sub("th_Ventricle", "th_ventricle", .) %>%
      sub("Thalamus", "Thalamus-Proper", .)
    df$label[cols_edit] <- gsub("_", "-", df$label[cols_edit])
  } else if (region_type == "sa") {
    df$label <- sub("aparc_SurfArea_", "", df$phenotype)
  } else {
    df$label <- sub("aparc_ThickAvg_", "", df$phenotype)
  }

  df
}
```

# Grouped Lollipop Plots
```{r lollipop-fcts, echo = TRUE}
# fixed_col/varying_col name which axis is held constant vs. compared, so the
# same filter+split logic covers "same predictor, compare variants" and "same
# variant, compare predictors" without two near-duplicate functions.
make_pair_table_list <- function(df, fixed_col, fixed_value, groups, varying_col, pair) {
  sub <- df %>%
    filter(.data[[fixed_col]] == fixed_value, group %in% groups, .data[[varying_col]] %in% pair)
  split(sub, sub[[varying_col]])[pair]
}

# avg_first_deriv comes straight out of extract_bam_features_main() (see
# model_extract_fcts.R), so it's already in main_stats -- only main_stats has
# it (resi_stats doesn't), so the derivative-sign-colored plots below only
# work for Partial R2 right now. To use them with RESI (or anything else),
# join deriv_sign in from main_stats by phenotype/predictor/variant/group first.
main_stats$deriv_sign <- ifelse(main_stats$avg_first_deriv >= 0, "Positive", "Negative")
```

Each plot below is a standalone block for one specific comparison -- not a loop that regenerates every predictor x variant x group x stat combination, since most of those aren't actually used downstream.

## GA vs BW within eTIV, colored by average-derivative sign (Partial R2)
```{r lollipop-etiv-ga-vs-bw-deriv, fig.width = 9, fig.height = 6}
this_variant <- "eTIV"
pair         <- c("gestational_age", "bwp_fen")
groups       <- global_groups
df           <- main_stats            # swap to resi_stats for RESI (see note above re: deriv_sign)
value_col    <- "r2_full_vs_none"     # swap to "RESI" for RESI
p_col        <- "full_p_adj"          # swap to "p_adj" for RESI

table_list <- make_pair_table_list(df, "variant", this_variant, groups, "predictor", pair)

p <- plot_lollipop(
  table_list = table_list,
  value_col = value_col,
  p_col = p_col,
  color_col = "deriv_sign",
  color_vec = c(Positive = deriv_pos_color, Negative = deriv_neg_color),
  name_mapping = namemap
) +
  labs(title = glue("{this_variant} -- Gestational Age vs Birth Weight Percentile (Partial R2)"), x = value_col) +
  theme_minimal()

# Add-on: mark which dodge position is GA vs BW, since color no longer
# identifies the predictor. plot_lollipop() doesn't expose the df it
# actually plots (post name_mapping / reorder), so this rebuilds just enough
# of it here to place text at the right y / x / dodge position -- a one-off,
# not worth a function for.
label_df <- dplyr::bind_rows(table_list, .id = "predictor")
if (!is.null(namemap)) {
  label_df <- merge(label_df, namemap, by.x = "phenotype", by.y = "key", all.x = TRUE)
  label_df$plot_label <- ifelse(is.na(label_df$plot_names), label_df$phenotype, label_df$plot_names)
} else {
  label_df$plot_label <- label_df$phenotype
}
label_df$predictor_short <- ifelse(label_df$predictor == "gestational_age", "GA", "BW")

p <- p +
  geom_text(
    data = label_df,
    aes(y = plot_label, x = .data[[value_col]], group = predictor, label = predictor_short),
    position = position_dodge(width = 0.6),
    hjust = -0.3, size = 3, color = "grey30"
  )

print(p)

if (params$save_plots == TRUE) {
  ggsave(glue("{plot_output_folder}lollipop_{this_variant}_ga_vs_bw_derivsign_{value_col}.svg"), p, width = 9, height = 6)
}
```

## GA only, colored by average-derivative sign (Partial R2)
```{r lollipop-etiv-ga-only-deriv, fig.width = 9, fig.height = 6}
this_variant <- "eTIV"
groups       <- global_groups
df           <- main_stats
value_col    <- "r2_full_vs_none"
p_col        <- "full_p_adj"

ga_df <- df %>% filter(predictor == "gestational_age", variant == this_variant, group %in% groups)

p_ga <- plot_lollipop(
  table_list = ga_df,
  value_col = value_col,
  p_col = p_col,
  color_col = "deriv_sign",
  color_vec = c(Positive = deriv_pos_color, Negative = deriv_neg_color),
  name_mapping = namemap
) +
  labs(title = glue("{this_variant} -- Gestational Age (Partial R2)"), x = value_col) +
  theme_minimal()

print(p_ga)

if (params$save_plots == TRUE) {
  ggsave(glue("{plot_output_folder}lollipop_{this_variant}_ga_only_derivsign_{value_col}.svg"), p_ga, width = 9, height = 6)
}
```

## BW only, colored by average-derivative sign (Partial R2)
```{r lollipop-etiv-bw-only-deriv, fig.width = 9, fig.height = 6}
this_variant <- "eTIV"
groups       <- global_groups
df           <- main_stats
value_col    <- "r2_full_vs_none"
p_col        <- "full_p_adj"

bw_df <- df %>% filter(predictor == "bwp_fen", variant == this_variant, group %in% groups)

p_bw <- plot_lollipop(
  table_list = bw_df,
  value_col = value_col,
  p_col = p_col,
  color_col = "deriv_sign",
  color_vec = c(Positive = deriv_pos_color, Negative = deriv_neg_color),
  name_mapping = namemap
) +
  labs(title = glue("{this_variant} -- Birth Weight Percentile (Partial R2)"), x = value_col) +
  theme_minimal()

print(p_bw)

if (params$save_plots == TRUE) {
  ggsave(glue("{plot_output_folder}lollipop_{this_variant}_bw_only_derivsign_{value_col}.svg"), p_bw, width = 9, height = 6)
}
```
# Brain Maps (Vol/SA/TH, Partial R2 & RESI, GA top / BWP bottom)

## define region types
```{r}
region_types <- c("vol", "sa", "th")
```

## single variant brain map plots
```{r}
# Single-variant brain map: plots df[[value_col]] directly for one `variant`
# (e.g. "nocovar"), instead of the pair-wise difference build_diff_table() computes.
# For region_type == "vol", `atlas` picks which subset of the "vol" group's
# phenotypes to plot -- DK cortical volumes (aparc_GrayVol_*) or ASEG subcortical
# volumes (aseg_*) -- since both live under the same "vol" group in
# main_stats/resi_stats and add_ggseg_label() already knows how to label either.
build_variant_table <- function(df, this_predictor, group, variant, value_col, p_col, region_type, atlas = "dk") {
  d <- df %>%
    filter(predictor == this_predictor, group == !!group, variant == !!variant) %>%
    select(phenotype, val = all_of(value_col), p = all_of(p_col))

  if (region_type == "vol") {
    d <- if (atlas == "aseg") {
      d %>% filter(grepl("^aseg_", phenotype))
    } else {
      d %>% filter(grepl("^aparc_GrayVol_", phenotype))
    }
  }

  add_ggseg_label(d, region_type)
}

plot_variant_brainmap <- function(df, this_predictor, group, variant, value_col, p_col, region_type,
                                   atlas = "dk", limits = NULL, title = "", fill_name = "value",
                                   low_color = "#F2F0EC", mid_color = "#D99A9A", high_color = "#D95C72") {
  table <- build_variant_table(df, this_predictor, group, variant, value_col, p_col, region_type, atlas)

  plot_brain_map(
    table = table,
    value_col = "val",
    atlas = atlas,
    p_col = "p",
    p_filter = TRUE,
    p_thresh = 0.05,
    title = title,
    fill_name = fill_name,
    limits = limits,
    low_color = low_color, mid_color = mid_color, high_color = high_color
  )
}

# Legend title shown on the shared colorbar -- "value_col" is the raw column
# name (r2_full_vs_none / RESI); readers should see the stat name instead.
stat_label <- function(value_col) if (value_col == "RESI") "RESI" else "Partial R2"

# Short text panels used as the shared grid's column headers (horizontal) and
# row headers (vertical), since the brain map panels themselves carry no title.
col_header_plot <- function(label) {
  ggplot() + theme_void() +
    annotate("text", x = 0, y = 0, label = label, size = 5, fontface = "bold") +
    xlim(-1, 1) + ylim(-1, 1)
}

row_header_plot <- function(label) {
  ggplot() + theme_void() +
    annotate("text", x = 0, y = 0, label = label, angle = 90, size = 5, fontface = "bold") +
    xlim(-1, 1) + ylim(-1, 1)
}

row_label_short <- function(predictor) if (predictor == "gestational_age") "GA" else "BWP"

# Columns of the shared grid: DK cortical volume, DK surface area, DK cortical
# thickness, ASEG subcortical volume.
brainmap_cols <- list(
  list(header = "Vol", region_type = "vol", atlas = "dk"),
  list(header = "SA",  region_type = "sa",  atlas = "dk"),
  list(header = "CTh", region_type = "th",  atlas = "dk"),
  list(header = "Vol", region_type = "vol", atlas = "aseg")
)

# One page per variant/stat combination: rows are GA (top) / BWP (bottom),
# columns are brainmap_cols, all sharing one color scale ("limits") collected
# into a single colorbar placed at the bottom of the page.
plot_variant_brainmap_page <- function(df, variant, value_col, p_col, limits) {
  fill_name <- stat_label(value_col)

  panels <- c(
    list(plot_spacer()),
    lapply(brainmap_cols, function(col) col_header_plot(col$header))
  )

  for (this_predictor in predictors) {
    panels <- c(panels, list(row_header_plot(row_label_short(this_predictor))))
    panels <- c(panels, lapply(brainmap_cols, function(col) {
      plot_variant_brainmap(df, this_predictor, col$region_type, variant, value_col, p_col,
                             col$region_type, atlas = col$atlas, limits = limits, fill_name = fill_name)
    }))
  }

  wrap_plots(panels, ncol = 5, nrow = 1 + length(predictors),
             widths = c(0.15, 1, 1, 1, 1), heights = c(0.15, rep(1, length(predictors))),
             guides = "collect") &
    theme(legend.position = "bottom")
}
```

## Spin test export
```{r spin-export-fct, echo = TRUE}
# Reuses build_variant_table() above (filter + add_ggseg_label()) instead of
# re-deriving the ggseg label / hemisphere split by hand the way
# slipga_linearanalyses_new.Rmd's "save betas" chunk did for each table.
save_stat_for_spin <- function(df, this_predictor, this_variant, group, value_col, p_col,
                                region_type, atlas = "dk", output_folder) {

  table <- build_variant_table(df, this_predictor, group, this_variant, value_col, p_col, region_type, atlas)

  # Spin tests want hemisphere as its own column and a bare region name in
  # `label`, not the ggseg-style "lh_"/"rh_" prefix -- same convention
  # slipga_linearanalyses_new.Rmd applied by hand for each table it saved.
  table$hemisphere <- dplyr::case_when(
    grepl("^lh_", table$label) ~ "L",
    grepl("^rh_", table$label) ~ "R",
    TRUE ~ NA_character_
  )
  table$label <- sub("^lh_|^rh_", "", table$label)

  names(table)[names(table) == "val"] <- value_col
  names(table)[names(table) == "p"]   <- p_col

  filename <- glue("{output_folder}{this_predictor}_{this_variant}_{group}_{value_col}_spin.csv")
  write.csv(table, filename, row.names = FALSE)
  invisible(table)
}
```

```{r spin-export-run, echo = TRUE}
# One call per (predictor, variant, group, value_col) combination needed for
# a given spin test -- add calls here as new ones come up rather than
# exporting every combination main_stats/resi_stats could produce.
save_stat_for_spin(main_stats, "gestational_age", "eTIV", "vol", "r2_full_vs_none", "full_p_adj",
                    region_type = "vol", atlas = "dk", output_folder = spin_output_folder)
```

```{r}
# Shared color-scale limits for the single-variant ("nocovar") brain map grids --
# fixed here (rather than per region_type) since anatomy is now a column within
# one page and all four columns on a page share one colorbar.
r2_limits   <- c(0, 0.022)
resi_limits <- c(0, 0.15)
```

### no covar
```{r}
p_nocovar_r2 <- plot_variant_brainmap_page(main_stats, "eTIV", "r2_full_vs_none", "full_p_adj",
                                            limits = r2_limits)
p_nocovar_resi <- plot_variant_brainmap_page(resi_stats, "eTIV", "RESI", "p_adj",
                                              limits = resi_limits)

print(p_nocovar_r2)
print(p_nocovar_resi)

if (params$save_plots == TRUE) {
  ggsave(glue("{plot_output_folder}brainmap_nocovar_r2.svg"),
         p_nocovar_r2, width = 16, height = 6)
  ggsave(glue("{plot_output_folder}brainmap_nocovar_resi.svg"),
         p_nocovar_resi, width = 16, height = 6)
}
```

### eTIV covar
```{r}
p_eTIV_r2 <- plot_variant_brainmap_page(main_stats, "eTIV", "r2_full_vs_none", "full_p_adj",
                                            limits = r2_limits)
p_eTIV_resi <- plot_variant_brainmap_page(resi_stats, "eTIV", "RESI", "p_adj",
                                              limits = resi_limits)

print(p_eTIV_r2)
print(p_eTIV_resi)

if (params$save_plots == TRUE) {
  ggsave(glue("{plot_output_folder}brainmap_etiv_r2.svg"),
         p_eTIV_r2, width = 16, height = 6)
  ggsave(glue("{plot_output_folder}brainmap_etiv_resi.svg"),
         p_eTIV_resi, width = 16, height = 6)
}
```

## difference plots

### helper functions
```{r brainmap-fcts, echo = TRUE}
# Difference map: pair[2]'s value minus pair[1]'s value, restricted to regions
# significant (p < .05) in at least one of the two variants being compared.
# Same "vol" group / atlas split as build_variant_table() -- aseg_* phenotypes
# (ASEG atlas) and aparc_GrayVol_* phenotypes (DK atlas) both live under group == "vol".
build_diff_table <- function(df, this_predictor, group, pair, value_col, p_col, region_type, atlas = "dk") {
  d1 <- df %>% filter(predictor == this_predictor, group == !!group, variant == pair[1])
  d2 <- df %>% filter(predictor == this_predictor, group == !!group, variant == pair[2])

  merged <- inner_join(
    d1 %>% select(phenotype, val1 = all_of(value_col), p1 = all_of(p_col)),
    d2 %>% select(phenotype, val2 = all_of(value_col), p2 = all_of(p_col)),
    by = "phenotype"
  )

  merged$diff  <- merged$val2 - merged$val1
  merged$min_p <- pmin(merged$p1, merged$p2)

  if (region_type == "vol") {
    merged <- if (atlas == "aseg") {
      merged %>% filter(grepl("^aseg_", phenotype))
    } else {
      merged %>% filter(grepl("^aparc_GrayVol_", phenotype))
    }
  }

  add_ggseg_label(merged, region_type)
}

plot_pair_brainmap <- function(df, this_predictor, group, pair, value_col, p_col, region_type,
                                atlas = "dk", limits, title = "", fill_name = "value",
                                low_color = "#2160B8", mid_color = "white", high_color = "#B82133") {
  diff_table <- build_diff_table(df, this_predictor, group, pair, value_col, p_col, region_type, atlas)

  plot_brain_map(
    table = diff_table,
    value_col = "diff",
    atlas = atlas,
    p_col = "min_p",
    p_filter = TRUE,
    p_thresh = 0.05,
    title = title,
    fill_name = fill_name,
    limits = limits,
    low_color = low_color, mid_color = mid_color, high_color = high_color
  )
}

# One grid per comparison: rows are GA (top) / BWP (bottom), columns are
# brainmap_cols (Vol DK, SA DK, CTh DK, Vol ASEG), sharing one color scale
# collected into a single colorbar at the bottom -- same layout as
# plot_variant_brainmap_page(), reusing its col_header_plot/row_header_plot/
# brainmap_cols/row_label_short/stat_label helpers.
plot_pair_brainmap_page <- function(df, pair, value_col, p_col, limits) {
  fill_name <- stat_label(value_col)

  panels <- c(
    list(plot_spacer()),
    lapply(brainmap_cols, function(col) col_header_plot(col$header))
  )

  for (this_predictor in predictors) {
    panels <- c(panels, list(row_header_plot(row_label_short(this_predictor))))
    panels <- c(panels, lapply(brainmap_cols, function(col) {
      plot_pair_brainmap(df, this_predictor, col$region_type, pair, value_col, p_col,
                          col$region_type, atlas = col$atlas, limits = limits, fill_name = fill_name)
    }))
  }

  (wrap_plots(panels, ncol = 5, nrow = 1 + length(predictors),
              widths = c(0.15, 1, 1, 1, 1), heights = c(0.15, rep(1, length(predictors))),
              guides = "collect") &
     theme(legend.position = "bottom")) +
    plot_annotation(title = glue("{pair[2]} - {pair[1]}"))
}
```

### find appropriate color range for plots
```{r}
# ---- Find shared colorbar limits for the difference brain map grids ----
# Mirrors exactly what plot_pair_brainmap_page() ends up plotting: for every
# (pair, predictor, brainmap_cols column) combination, build the diff table
# and keep only the regions that pass the same p_filter (min_p < 0.05) the
# plots use -- so the range reflects what's actually shown, not values that
# get dropped as non-significant and never appear on the map.
diff_value_range <- function(df, value_col, p_col, p_thresh = 0.05) {
  vals <- unlist(lapply(covariate_pairs, function(pair) {
    unlist(lapply(predictors, function(this_predictor) {
      unlist(lapply(brainmap_cols, function(col) {
        t <- build_diff_table(df, this_predictor, col$region_type, pair,
                               value_col, p_col, col$region_type, atlas = col$atlas)
        t$diff[t$min_p < p_thresh]
      }))
    }))
  }))
  range(vals, na.rm = TRUE)
}

r2_diff_range   <- diff_value_range(main_stats, "r2_full_vs_none", "full_p_adj")
resi_diff_range <- diff_value_range(resi_stats, "RESI", "p_adj")

# Symmetric versions, since plot_brain_map()'s mid_color sits at mean(limits)
# rather than at 0 -- for a diverging (blue-white-red) diff scale centered on
# zero, use these instead of the raw min/max above.
r2_diff_limits_sym   <- c(-1, 1) * max(abs(r2_diff_range))
resi_diff_limits_sym <- c(-1, 1) * max(abs(resi_diff_range))

print(list(r2_diff_range = r2_diff_range, resi_diff_range = resi_diff_range,
           r2_diff_limits_sym = r2_diff_limits_sym, resi_diff_limits_sym = resi_diff_limits_sym))
```

```{r}
# Shared color-scale limits for the difference brain map grids.
diff_r2_limits   <- c(-0.065, 0.065)
diff_resi_limits <- c(-0.24, 0.24)
```

### plot and save
```{r brainmap-plots, results = 'asis', fig.width = 16, fig.height = 6}
for (pair in covariate_pairs[4]) {
  pair_label <- paste(pair, collapse = "_vs_")

  p_pair_r2 <- plot_pair_brainmap_page(main_stats, pair, "r2_full_vs_none", "full_p_adj",
                                        limits = diff_r2_limits)
  p_pair_resi <- plot_pair_brainmap_page(resi_stats, pair, "RESI", "p_adj",
                                          limits = diff_resi_limits)

  print(p_pair_r2)
  print(p_pair_resi)

  if (params$save_plots == TRUE) {
    ggsave(glue("{plot_output_folder}brainmap_diff_{pair_label}_r2.svg"),
           p_pair_r2, width = 16, height = 6)
    ggsave(glue("{plot_output_folder}brainmap_diff_{pair_label}_resi.svg"),
           p_pair_resi, width = 16, height = 6)
  }
}
```


# Significant interactions
```{r modelvis-sig-fct, echo = TRUE}
int_stats$ interaction_type = "continuous"
# `int_stats` is the union of columns across all three interaction_types
# (bind_rows() fills whatever a given row's type didn't produce with NA), so
# a row is "significant" if:
#  - continuous: its own int_p_adj < .05
#  - ordered:    ANY of its int1_p_adj, int2_p_adj, ... (per-level difference
#                smooths vs. the reference level) is < .05
#  - categorical: ANY of its per-level <level>_p_adj columns is < .05 -- there
#                is no int_p_adj/pred_p_adj at all for categorical rows (no
#                bare s(predictor) or ti() term exists in that formula), so
#                those two columns are excluded from the level-column search
#                to avoid a name collision with a moderator level literally
#                called "int1" or "pred" (the known tradeoff of matching by
#                column name instead of a stored, unambiguous type tag)
if (params$model_vis == TRUE) {
  p_adj_cols     <- grep("_p_adj$", names(int_stats), value = TRUE)
  ordered_cols   <- grep("^int[0-9]+_p_adj$", p_adj_cols, value = TRUE)
  categorical_cols <- setdiff(p_adj_cols, c("pred_p_adj", "int_p_adj", ordered_cols))
  
  any_below <- function(df, cols, thresh = 0.05) {
    if (length(cols) == 0) return(rep(FALSE, nrow(df)))
    mat <- as.matrix(df[, cols, drop = FALSE])
    apply(mat < thresh, 1, any, na.rm = TRUE)
  }
  
  int_stats$any_sig <- NA
  cont_idx <- int_stats$interaction_type == "continuous"
  ord_idx  <- int_stats$interaction_type == "ordered"
  cat_idx  <- int_stats$interaction_type == "categorical"
  
  int_stats$any_sig[cont_idx] <- int_stats$int_p_adj[cont_idx] < 0.05
  int_stats$any_sig[ord_idx]  <- any_below(int_stats[ord_idx, ], ordered_cols)
  int_stats$any_sig[cat_idx]  <- any_below(int_stats[cat_idx, ], categorical_cols)
}
```

## Interactions across different covariate models
```{r}
interactions_summary <- int_stats %>%
  filter(predictor %in% c("gestational_age", "bwp_fen")) %>%
  group_by(predictor, variant) %>%
  summarise(
    n_sig = sum(any_sig, na.rm = TRUE),
    phenotypes = list(phenotype[any_sig]),
    .groups = "drop"
  )

View(
interactions_summary %>%
  mutate(
    phenotypes = sapply(
      phenotypes,
      \(x) paste(x, collapse = "\n")
    )
  )
)
```

# Model Visualization
```{r modelvis-fcts, echo = TRUE}
if (params$model_vis == TRUE) {
  get_main_model <- function(variant, this_predictor, group, phenotype) {
    run_key <- paste(this_predictor, group, sep = "_")
    models[[run_key]]$main[[phenotype]]$full
  }
  
  plot_main_model <- function(variant, this_predictor, group, phenotype) {
    visualize_spline_model(
      modobj = get_main_model(variant, this_predictor, group, phenotype),
      smooth_var = this_predictor,
      plot_og_scale = FALSE,
      mean_lookup = mean_lookup_distinct,
      sd_lookup = sd_lookup_distinct,
      pnorm_smooth_var = FALSE,
      plabels = phenotype,
      color_point = if (this_predictor == "gestational_age") ga_color else bwp_color,
      color_line  = if (this_predictor == "gestational_age") ga_color else bwp_color
    ) +
      predictor_x_scale(this_predictor) +
      outcome_y_scale() +
      labs(x = predictor_label(this_predictor), title = phenotype) +
      theme(axis.title.y = element_blank())
  }
}
```

## Significant global & CMD models
```{r modelvis-global-cmd, results = 'asis', fig.width = 10, fig.height = 8}
if (params$model_vis == TRUE) {
  for (this_predictor in predictors) {
    sig_pheno <- main_stats %>%
      filter(predictor %in% c("gestational_age")) %>%
      filter(variant == variant_to_visualize, predictor == this_predictor,
             group %in% c(global_groups, cmd_groups), full_p_adj < 0.05)
  
    if (nrow(sig_pheno) == 0) next
  
    plots_ga <- Map(function(ph, grp) plot_main_model(variant_to_visualize, this_predictor, grp, ph),
                 sig_pheno$phenotype, sig_pheno$group)
    
    print(plots_ga)
  
    #print(wrap_plots(plots, ncol = 2) & plot_layout(guides = "collect"))
  }
}

if (params$model_vis == TRUE) {
  for (this_predictor in predictors) {
    sig_pheno <- main_stats %>%
      filter(predictor %in% c("bwp_fen")) %>%
      filter(variant == variant_to_visualize, predictor == this_predictor,
             group %in% c(global_groups, cmd_groups), full_p_adj < 0.05)
  
    if (nrow(sig_pheno) == 0) next
  
    plots_bwp <- Map(function(ph, grp) plot_main_model(variant_to_visualize, this_predictor, grp, ph),
                 sig_pheno$phenotype, sig_pheno$group)
    
    print(plots_bwp)
  
    #print(wrap_plots(plots, ncol = 2) & plot_layout(guides = "collect"))
  }
}
```

## Top 5 RESI regions within vol / sa / th
```{r modelvis-top5-resi, results = 'asis', fig.width = 10, fig.height = 8}
if (params$model_vis == TRUE) {
  for (this_predictor in predictors) {
    for (region_type in region_types) {
  
      top5 <- resi_stats %>%
        filter(predictor %in% c("gestational_age", "bwp_fen")) %>%
        filter(variant == variant_to_visualize, predictor == this_predictor,
               group == region_type, p_adj < 0.05) %>%
        arrange(desc(RESI)) %>%
        head(5)
  
      if (nrow(top5) == 0) next
  
      plots <- lapply(top5$phenotype, function(ph) plot_main_model(variant_to_visualize, this_predictor, region_type, ph))
      print(wrap_plots(plots, ncol = 2) & plot_layout(guides = "collect"))
    }
  }
}
```

## Significant Interactions

### calculate plotting age labels
```{r}
mu_age <- mean_lookup_distinct["age_days_adj"]
sd_age <- sd_lookup_distinct["age_days_adj"]

mu_ga <- mean_lookup_distinct["gestational_age"]
sd_ga <- sd_lookup_distinct["gestational_age"]
  
low_age_label <- (quantile(df.zscore[!is.na(df.zscore$gestational_age),
                                       "age_days_adj"],
                             probs = c(.10)) * sd_age + mu_age)/365.25
  
mid_age_label <- (quantile(df.zscore[!is.na(df.zscore$gestational_age),
                                       "age_days_adj"],
                             probs = c(.50)) * sd_age + mu_age)/365.25

high_age_label <- (quantile(df.zscore[!is.na(df.zscore$gestational_age),
                                      "age_days_adj"],
                            probs = c(.90)) * sd_age + mu_age)/365.25
```

### Show interaction models
```{r modelvis-interactions, results = 'asis', fig.width = 9, fig.height = 6}
# visualize_model_int() already dispatches on the moderator's actual R class
# (factor/ordered/numeric), so the same plotting call below works unchanged
# across all three interaction_types -- only the significance filter above
# needed to branch.
if (params$model_vis == TRUE) {
  sig_int <- int_stats %>%
    filter(predictor %in% c("gestational_age", "bwp_fen")) %>%
    filter(variant == variant_to_visualize, any_sig)
  
  for (i in seq_len(nrow(sig_int))) {
    row <- sig_int[i, ]
    run_key <- paste(row$predictor, row$group, sep = "_")
    modobj <- models[[run_key]]$interactions[[row$moderator_label]][[row$phenotype]]
  
    p <- visualize_model_int(
      modobj = modobj,
      smooth_var = row$predictor,
      int_var = row$moderator,
      plabels = glue("{row$phenotype} x {row$moderator_label}"),
      derivative_plot = FALSE
    ) +
      predictor_x_scale(row$predictor) +
      outcome_y_scale() +
      labs(x = predictor_label(row$predictor))
    
    plot_colors <- ggplot_build(p)$plot$scales$get_scales("fill")$palette(1)
    
    p <- p +
  scale_fill_manual(labels = c(paste0("Low: ",
                                      round(low_age_label,1),"yrs"),
                               paste0("Mid: ",
                                      round(mid_age_label,1),"yrs"),
                               paste0("High: ",
                                      round(high_age_label,1),"yrs")), 
                    values = plot_colors) +
   scale_color_gradientn(colors = plot_colors,
    labels = function(x) round((x * sd_age + mu_age)/365.25, 1),
    name = "")
  
    print(p)
  }
}
```