---
title: "slip_swyc"
output: html_document
date: "2026-02-03"
editor_options: 
  chunk_output_type: console
---
Ideas
- get an average swyc score per participant and correlate with 4 global phenotypes + meanthickness + total SA + domain-based CMD & calculate 4 global phenotypes CMD
- try to figure out how to get a "trajectory" per participant

# Controls
```{r setup, echo = TRUE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)
save_figs = TRUE
save_tables = TRUE
save_processed = TRUE
```

# Load packages
```{r}
#Load packages. Install these packages with install.packages("pkg_name") if they are not installed.
library(bigrquery) #package to import SQL tables into R
require(dplyr)
require(glue)
require(magrittr)
require(readr)
require(stringr)
require(tidyr)
require(scales)
require(mgcv)
require(pbkrtest)
require(gamlssTools) # Margaret' package
require(tibble)
require(lme4)
require(purrr)
# For figures/plotting
require(ggplot2)
require(gridExtra)
require(ggsegDKT)
require(ggseg)
require(ggpubr)
require(ggeffects)
require(ggcorrplot)
require(cowplot)
require(ggh4x)
require(gratia)
require(egg)
require(interactions)
require(growthstandards)
require(gtsummary)
require(svglite)
require(patchwork)
```

# Define paths
```{r}
code_path <- "/mnt/arcus/lab/users/kafadare/slip_premie_wip/"

save_path <-  "/mnt/arcus/lab/users/kafadare/swyc_premie_analysis_results/"
if(!dir.exists(save_path)) dir.create(save_path)

plot_output_folder <- paste0(save_path, "plots/")
if(!dir.exists(plot_output_folder)) dir.create(plot_output_folder)

csv_tables_folder <-  paste0(save_path, "processed_tables/")
if (!dir.exists(csv_tables_folder)) dir.create(csv_tables_folder)

csv_output_folder <-  paste0(save_path, "results_csv/")
if (!dir.exists(csv_output_folder)) dir.create(csv_output_folder)

betas_folder <- paste0(csv_output_folder, "full_betas/")
if (!dir.exists(betas_folder)) dir.create(betas_folder)

dkt_folder <- paste0(csv_output_folder, "dkt_betas/")
if (!dir.exists(dkt_folder)) dir.create(dkt_folder)
```

# Load custom functions
```{r}
source(glue("{code_path}sliding_window.R"))
source(glue("{code_path}visualize_model_bartfct.R"))
source(glue("{code_path}vis_spline_noint_EK.R"))
source(glue("{code_path}model_fitting_fcts.R"))
source(glue("{code_path}model_extract_fcts.R"))
source(glue("{code_path}/slip_data_utils.R"))
source(glue("{code_path}/plotting_utils.R"))

panel_save <- function(plot, filename, extension, folder, plot_width = 8, plot_height = 10){
  save_name <- paste0(folder, filename, ".", extension)
  ggsave(save_name, plot = plot, dpi = 300, width = plot_width, height = plot_height)
}

`%nin%` = Negate(`%in%`)
```

# Load data 
```{r}
swyc_all_path <- "/mnt/arcus/lab/users/kafadare/swyc_premie_analysis_results/processed_tables/swyc_allforms.csv"
swyc_avg_path <- "/mnt/arcus/lab/users/kafadare/swyc_premie_analysis_results/processed_tables/swyc_avg.csv"

swyc_all <- read.csv(swyc_all_path) %>% select(-X)
swyc_avg <- read.csv(swyc_avg_path) %>% select(-X)
```

# Colors
```{r}
ga_color = "#6D8299"
bw_color = "#709085"
bwp_color = "#A68A64"
swyc_color = "#9D73BA"


limits = c(-0.3, 0.3)
low_color = "#4878CF" # blue
mid_color = "white"
high_color = "#D65F5F" # red
```

# Prep Data for Analysis

## zscore variables of interest
```{r, echo = FALSE}
swyc_all[swyc_all[,"bwp_fen"] == 0 & !is.na(swyc_all[,"bwp_fen"]),"bwp_fen"] <- 0.0000000000000000000000001
swyc_all[swyc_all[,"bwp_fen"] == 1 & !is.na(swyc_all[,"bwp_fen"]), "bwp_fen"] <- 0.99999999999999994
  
swyc_avg[swyc_avg[,"bwp_fen"] == 0 & !is.na(swyc_avg[,"bwp_fen"]), "bwp_fen"] <- 0.0000000000000000000000001
swyc_avg[swyc_avg[,"bwp_fen"] == 1 & !is.na(swyc_avg[,"bwp_fen"]), "bwp_fen"] <- 0.99999999999999994

# zscore swyc avg birth factors and DQ
swyc_avg <- swyc_avg %>%
  mutate(across(all_of(c("median_dq", "ga_adj_median_dq", 
                         "final_swyc_age", "final_swyc_age_ga_adj",
                         "gestational_age", "birth_weight_kg",
                         "final_ADI", "final_ADI3_FINS", "final_ADI3_ECON", "final_ADI3_EDU" )),
                ~as.numeric(scale(.x)), .names = "{.col}_zscore"),
         bwp_fen_zscore = qnorm(bwp_fen))

# set sex as factor
swyc_avg$sex <- as.factor(swyc_avg$sex)

# Add numeric column for form month
swyc_all <- swyc_all %>%
  mutate(
    form_month_num = as.numeric(sub("SWYC : ", "", form_friendly_name) %>%
      sub(" Month", "", .) %>% sub("0", "", .)),
    form_name = sub(" Month", "_Month", form_friendly_name) %>% sub("SWYC : ", "", .)) %>%
  select(!form_friendly_name) %>%
  mutate(
    pat_form = paste0(pat_id, "_", form_name)
  )

# zscore swyc all forms birth factors and DQ
swyc_all <- swyc_all %>%
  group_by(form_name) %>%
  mutate(across(all_of(c("norm_devo_quotient", "ga_adj_norm_devo_quotient", 
                         "effective_age", "ga_adj_effective_age",
                         "gestational_age", "birth_weight_kg",
                         "ADI.median", "ADI3_FINS.median", "ADI3_ECON.median", "ADI3_EDU.median" )),
                ~as.numeric(scale(.x)), .names = "{.col}_zscore"),
         bwp_fen_zscore = qnorm(bwp_fen)) %>%
  ungroup()

# set sex as factor
swyc_all$sex <- as.factor(swyc_all$sex_abbr)

# change form name order factor levels
form_order <- c("04_Month",
                "06_Month", 
                "09_Month", 
                "12_Month",
                "15_Month",
                "18_Month", 
                "24_Month", 
                "30_Month",
                "36_Month",
                "48_Month",
                "60_Month")

swyc_all <- swyc_all %>%
  mutate(
    form_name = factor(
      form_name,
      levels = form_order,
      ordered = TRUE
    )
  )
      
swyc_all_wide <- swyc_all %>%
  pivot_wider(
    id_cols = c(pat_id, gestational_age, birth_weight_kg, bwp_fen),
    names_from = form_name,
    values_from = c(norm_devo_quotient, ga_adj_norm_devo_quotient, effective_age, ga_adj_effective_age,
                    ADI.median, ADI3_FINS.median, ADI3_ECON.median, ADI3_EDU.median),
    names_glue = "{.value}_{form_name}"
  )

dq_vars <- names(swyc_all_wide)[grepl("norm_devo", names(swyc_all_wide))]
adi_vars <- names(swyc_all_wide)[grepl("ADI", names(swyc_all_wide))]
vars_tozscore <- names(swyc_all_wide)[grepl("norm_devo|ADI|effective_age", names(swyc_all_wide))]

# zscore swyc avg birth factors and DQ wide format
swyc_all_wide <- swyc_all_wide %>%
  mutate(across(all_of(c(vars_tozscore,
                         "gestational_age", "birth_weight_kg")),
                ~as.numeric(scale(.x)), .names = "{.col}_zscore"),
         bwp_fen_zscore = qnorm(bwp_fen))
```

# Select which SWYC forms to investigate (based on available N)
```{r}
tab_swyc_forms <- table(swyc_all$form_name)
data.frame(form = names(tab_swyc_forms), count = as.vector(tab_swyc_forms))

forms_totest <- c("06_Month", 
                       "09_Month",
                       "15_Month",
                       "18_Month",
                       "24_Month",
                       "30_Month")

og_dq_vars_totest <- c("06_Month", 
                       "09_Month",
                       "15_Month",
                       "18_Month",
                       "24_Month",
                       "30_Month")

ga_adj_dq_vars_totest <- paste0("ga_adj_", og_dq_vars_totest)
```

# Explore relationships between SWYC, birth factors, and ADI

## SWYC and birth factors

### Simple Correlation Matrix
```{r}
og_cor_mat <- cor(swyc_all_wide[,c(og_dq_vars_totest, "gestational_age", "birth_weight_kg", "bwp_fen")], use = "pairwise.complete.obs")

og_p_mat <- cor_pmat(swyc_all_wide[,c(og_dq_vars_totest, "gestational_age", "birth_weight_kg", "bwp_fen")], use = "pairwise.complete.obs")

(swyc_ogDQ_forms_bf_corrs_plot <- ggcorrplot(
  og_cor_mat,
  p.mat = og_p_mat,
  type = "full",
  insig = "blank",
  lab = TRUE
))

ga_adj_cor_mat <- cor(swyc_all_wide[,c(ga_adj_dq_vars_totest, "gestational_age", "birth_weight_kg", "bwp_fen")], use = "pairwise.complete.obs")

ga_adj_p_mat <- cor_pmat(swyc_all_wide[,c(ga_adj_dq_vars_totest, "gestational_age", "birth_weight_kg", "bwp_fen")], use = "pairwise.complete.obs")

(swyc_ga_adjDQ_forms_bf_corrs_plot <- ggcorrplot(
  ga_adj_cor_mat,
  p.mat = ga_adj_p_mat,
  type = "full",
  insig = "blank",
  lab = TRUE
))

if(save_figs == TRUE){
  panel_save(swyc_ogDQ_forms_bf_corrs_plot, "swyc_ogDQ_birth_cor_matrix", "svg", plot_output_folder)
  panel_save(swyc_ga_adjDQ_forms_bf_corrs_plot, "swyc_ga_adjDQ_birth_cor_matrix", "svg", plot_output_folder)
}

```

### Correlations Grouped (not matrix)
```{r}
bf_vars <- c(
  "gestational_age",
  "birth_weight_kg",
  "bwp_fen"
)

ogDQ_bf_cor_results <- swyc_all %>%
  filter(form_name %in% forms_totest) %>%
  pivot_longer(
    cols = all_of(bf_vars),
    names_to = "bf_type",
    values_to = "bf_score"
  ) %>%
  group_by(form_name, bf_type) %>%
  group_modify(~{
    ct <- cor.test(
      .x$norm_devo_quotient,
      .x$bf_score,
      method = "pearson"
    )
    tibble(
      n = nrow(.x),
      r = unname(ct$estimate),
      lower = ct$conf.int[1],
      upper = ct$conf.int[2],
      p = ct$p.value
    )
  }) %>%
  ungroup()

(swyc_ogDQ_forms_bf_corrs_plot <- ggplot(ogDQ_bf_cor_results,
       aes(x = r,
           y = form_name,
           colour = bf_type)) +
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "grey60") +
  geom_errorbarh(
    aes(xmin = lower, xmax = upper),
    height = 0.2
  ) +
  geom_point(size = 3) +
  facet_wrap(~bf_type) +
  scale_x_continuous(
    limits = c(-1, 1),
    breaks = seq(-1, 1, 0.25)
  ) +
  labs(
    x = "Pearson correlation (r)",
    y = NULL,
    title = "OG DQ",
    colour = "bf measure"
  ) +
  theme_classic())

ga_adjDQ_bf_cor_results <- swyc_all %>%
  filter(form_name %in% forms_totest) %>%
  pivot_longer(
    cols = all_of(bf_vars),
    names_to = "bf_type",
    values_to = "bf_score"
  ) %>%
  group_by(form_name, bf_type) %>%
  group_modify(~{
    ct <- cor.test(
      .x$ga_adj_norm_devo_quotient,
      .x$bf_score,
      method = "pearson"
    )
    tibble(
      n = nrow(.x),
      r = unname(ct$estimate),
      lower = ct$conf.int[1],
      upper = ct$conf.int[2],
      p = ct$p.value
    )
  }) %>%
  ungroup()

(swyc_ga_adjDQ_forms_bf_corrs_plot <- ggplot(ga_adjDQ_bf_cor_results,
       aes(x = r,
           y = form_name,
           colour = bf_type)) +
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "grey60") +
  geom_errorbarh(
    aes(xmin = lower, xmax = upper),
    height = 0.2
  ) +
  geom_point(size = 3) +
  facet_wrap(~bf_type) +
  scale_x_continuous(
    limits = c(-1, 1),
    breaks = seq(-1, 1, 0.25)
  ) +
  labs(
    x = "Pearson correlation (r)",
    y = NULL,
    title = "GA-ADJ DQ",
    colour = "bf measure"
  ) +
  theme_classic())

if(save_figs == TRUE){
  panel_save(swyc_ogDQ_forms_bf_corrs_plot, "swyc_ogDQ_bf_cors", "svg", plot_output_folder)
  panel_save(swyc_ga_adjDQ_forms_bf_corrs_plot, "swyc_ga_adjDQ_bf_cors", "svg", plot_output_folder)
}
```

### Test linear relationships, with form type interaction
```{r}
swyc_ga_ogdq_lm <- lm(norm_devo_quotient_zscore ~ gestational_age_zscore*form_name, 
                      data = swyc_all %>% filter(form_name %in% forms_totest))

summary(swyc_ga_ogdq_lm)

swyc_ga_ga_adjdq_lm <- lm(ga_adj_norm_devo_quotient_zscore ~ gestational_age_zscore*form_name, 
                      data = swyc_all %>% filter(form_name %in% forms_totest))

summary(swyc_ga_ga_adjdq_lm)

swyc_bwp_ogdq_lm <- lm(norm_devo_quotient_zscore ~ bwp_fen_zscore*form_name, 
                      data = swyc_all %>% filter(form_name %in% forms_totest))

summary(swyc_bwp_ogdq_lm)

swyc_bwp_ga_adjdq_lm <- lm(ga_adj_norm_devo_quotient_zscore ~ bwp_fen_zscore*form_name, 
                      data = swyc_all %>% filter(form_name %in% forms_totest))

summary(swyc_bwp_ga_adjdq_lm)
```

### Plot form type X GA linear interaction
```{r}
# Get predicted values across GA z-score range for each form
pred <- ggpredict(
  swyc_ga_ogdq_lm,
  terms = c(
    "gestational_age_zscore",
    "form_name"
  )
)

# Plot interaction
(ogdq_ga_form_interaction_plot <- ggplot(pred,
       aes(x = x,
           y = predicted,
           colour = group)) +
  geom_line(linewidth = 1) +
  geom_ribbon(
    aes(ymin = conf.low, ymax = conf.high, fill = group),
    alpha = 0.2,
    colour = NA
  ) +
  labs(
    x = "Gestational age z-score",
    y = "Predicted normalized OG DQ",
    colour = "SWYC form",
    fill = "SWYC form"
  ) +
  theme_classic())

pred <- ggpredict(
  swyc_ga_ga_adjdq_lm,
  terms = c(
    "gestational_age_zscore",
    "form_name"
  )
)

# Plot interaction
(ga_adjdq_ga_form_interaction_plot <- ggplot(pred,
       aes(x = x,
           y = predicted,
           colour = group)) +
  geom_line(linewidth = 1) +
  geom_ribbon(
    aes(ymin = conf.low, ymax = conf.high, fill = group),
    alpha = 0.2,
    colour = NA
  ) +
  labs(
    x = "Gestational age z-score",
    y = "Predicted normalized GA-ADJ DQ",
    colour = "SWYC form",
    fill = "SWYC form"
  ) +
  theme_classic())

if(save_figs == TRUE){
  panel_save(ogdq_ga_form_interaction_plot, "ogdq_ga_form_interaction_plot", "svg", plot_output_folder)
  panel_save(ga_adjdq_ga_form_interaction_plot, "ga_adjdq_ga_form_interaction_plot", "svg", plot_output_folder)
}
```


### Test spline between GA and DQ

#### with OG DQ 

##### fit the no interaction model
```{r}
ga_ogdq_model <- fit_bam_models(swyc_avg, "median_dq_zscore", 
                                        predictor = "gestational_age_zscore",
                                 save = FALSE)
  
tab <- extract_bam_table(models = ga_ogdq_model, 
                                                  predictor ="gestational_age_zscore")
# Add the rest here manually, because bam() won't run with just an intercept (m_red crashes inside the function)

models = ga_ogdq_model
predictor ="gestational_age_zscore"
data = swyc_avg 
covariates = NULL
random_effect = NULL

feats <- lapply(names(models), function(ph){
    
    m_full  <- models[[ph]]$full
    m_nolin <- models[[ph]]$nolin
    
    # rebuild reduced model WITHOUT predictor entirely
    covs <- if(!is.null(covariates)) paste(covariates, collapse = " + ") else ""
    re <- if(!is.null(random_effect)) paste0("s(", random_effect, ", bs='re')") else ""
    
    rhs <- paste(c(covs, re)[c(covs, re) != ""], collapse = " + ")

    f_red <- as.formula(paste(ph, "~ 1", 
                              if(rhs != "") paste("+", rhs)))
    
    f_onlylin <- as.formula(paste(ph, "~ 1 + ", predictor,
                                  if(rhs != "") paste("+", rhs)))
    
    # subset data to only include predictor is not NA to fit reduced model
    data_fit <- data %>% filter(!is.na(get(predictor)) & !is.na(get(ph)))
    #m_red <- mgcv:bam(f_red, data = data_fit, method = "fREML", discrete = TRUE)
    m_onlylin <- mgcv::bam(f_onlylin, data = data_fit, method = "fREML", discrete = TRUE)
    
    # partial R2: full predictor contribution vs no predictor model
    #r2_full_vs_none <- 1 - deviance(m_full)/deviance(m_red)
    #r2_full_vs_none <- (summary(m_full)$r.sq - summary(m_red)$r.sq) /  (1 - summary(m_red)$r.sq)
    
    # partial R2: spline contribution over linear model
    r2_spline_vs_linear <- 1 - deviance(m_full)/deviance(m_onlylin)
    #r2_spline_vs_linear <- (summary(m_full)$r.sq - summary(m_onlylin)$r.sq) /  (1 - summary(m_onlylin)$r.sq)
    
    # average first derivative
    d <- derivatives(m_full)
    avg_deriv <- mean(d$.derivative)
    
    # region of effect (CI crossing rule)
    d <- d %>%
      mutate (sig = !(0 >.lower_ci & 0 < .upper_ci))
    region <- if(sum(d$sig) > 0) c(min(d[d$sig == T, predictor]), max(d[d$sig == T, predictor])) else c(NA, NA)
    
    c(
      #r2_full_vs_none = r2_full_vs_none,
      r2_spline_vs_linear = r2_spline_vs_linear,
      avg_first_deriv = avg_deriv,
      region_start = region[1],
      region_end = region[2]
    )
  })
  
ga_ogdq_table <- cbind(tab, do.call(rbind, feats))
  
```

##### print no interaction model table
```{r}
knitr::kable(t(ga_ogdq_table))

```

##### visualize no interaction model
```{r}
ga_ogdq_plot <- visualize_spline_model(
  modobj = ga_ogdq_model$median_dq_zscore$full,
  plot_og_scale = FALSE,
  pnorm_smooth_var = FALSE,
  smooth_var = "gestational_age_zscore",
  plabels = "Gestational Age",
  color_point = ga_color,
  color_line = ga_color) + 
    scale_x_continuous(
      breaks = map2_dbl(c(24*7, 28*7, 32*7, 36*7, 40*7), "gestational_age", ~ (.x - mean(swyc_avg[,.y], na.rm =T))/(sd(swyc_avg[,.y], na.rm = T))),
      labels = \(x) round(map2_dbl(x, "gestational_age", ~ (.x*sd(swyc_avg[,.y], na.rm = T) + mean(swyc_avg[,.y], na.rm = T))/7))) +
    # scale_y_continuous(breaks = qnorm(c(.01, .50, .99)), 
    #                    labels = c("<=1st",  "50th", ">=99th")) +
    labs(x= "Gestational Age (weeks)",
         y = "OG DQ (z-score)",
         title = "")

ga_ogdq_plot

if(save_figs){
saveRDS(ga_ogdq_plot, glue("{plot_output_folder}ga_ogdq_spline_model_plot.rds"))
}

```

##### fit the model with form interaction (form as numerical with month)
```{r} 
ga_ogdq_formint_model <- fit_bam_interactions(data = swyc_all,
                                        phenotypes = "norm_devo_quotient_zscore", 
                                        predictor = "gestational_age_zscore",
                                        moderator = "form_month_num",
                                        covariates = NULL, save = FALSE)
  
ga_ogdq_formint_table <- extract_bam_features_interaction(models = ga_ogdq_formint_model, 
                                                  data = swyc_all, 
                                                  predictor = "gestational_age_zscore",
                                                  moderator = "form_month_num")

knitr::kable(t(ga_ogdq_formint_table))
```


##### fit the model with form interaction (form as ordered)
```{r} 
ga_ogdq_formint_categorical_model <- fit_bam_interactions(swyc_all %>% 
                                                            filter(form_name %in% forms_totest), 
                                        "norm_devo_quotient_zscore", 
                                        predictor = "gestational_age_zscore",
                                        moderator = "form_name",
                                        interaction = "ordered",
                                        covariates = NULL, 
                                        save = FALSE)
  
ga_ogdq_formint_categorical_table <- extract_bam_features_interaction(
                                                        models = ga_ogdq_formint_categorical_model,
                                                          data = swyc_all %>% 
                                                            filter(form_name %in% forms_totest),
                                                          predictor = "gestational_age_zscore",
                                                          moderator = "form_name",
                                                       interaction_type = "ordered")
  
knitr::kable(t(ga_ogdq_formint_categorical_table))
```

##### significance testing using log likelihood test for interaction term
shouldnt use fREML if comparing log ratio using ANOVA, using ML to fit models
```{r}
m1 <- fit_bam_interactions(swyc_all %>% 
                                                            filter(form_name %in% forms_totest), 
                                        "norm_devo_quotient_zscore", 
                                        predictor = "gestational_age_zscore",
                                        moderator = "form_name",
                                        interaction = "ordered",
                                        covariates = NULL, 
                                        save = FALSE,
                                        method = "ML")

m1 <- m1$norm_devo_quotient_zscore

m0 <- fit_bam_models(swyc_all %>% 
                       filter(form_name %in% forms_totest), "norm_devo_quotient_zscore", 
                     predictor = "gestational_age_zscore",
                                 save = FALSE,
                                method = "ML")

m0 <- m0$norm_devo_quotient_zscore$full

anova(m0, m1, test = "Chisq")
```

##### visualize form interaction model (continuous)
```{r}
font_size <- 16
theme_set(theme_classic(base_family = "Helvetica",base_size = font_size))
line_size <- 1.5
point_size <- 2

ga_ogdq_formint_plot <- visualize_model(
  modobj = ga_ogdq_formint_model$norm_devo_quotient_zscore,
  smooth_var = "gestational_age_zscore",
  int_var = "form_month_num",
  plabels = "continuous interaction",
  derivative_plot = F) + 
    labs(x="Gestational Age (wks)",
         title = "",
                  fill = "Form Month") +
    theme(axis.title.y = element_blank()) +
      scale_x_continuous(
      breaks = map2_dbl(c(24*7, 28*7, 32*7, 36*7, 40*7), "gestational_age", ~ (.x - mean(swyc_avg[,.y], na.rm =T))/(sd(swyc_avg[,.y], na.rm = T))),
      labels = \(x) round(map2_dbl(x, "gestational_age", ~ (.x*sd(swyc_avg[,.y], na.rm = T) + mean(swyc_avg[,.y], na.rm = T))/7)))
    # scale_y_continuous(breaks = qnorm(c(.01, .50, .99)), 
    #                    labels = c("<=1st",  "50th", ">=99th")) +
    
ga_ogdq_formint_plot
```


##### visualize form interaction model (categorical)
```{r}
font_size <- 16
theme_set(theme_classic(base_family = "Helvetica",base_size = font_size))
line_size <- 1.5
point_size <- 2

ga_ogdq_formint_categorical_plot <- visualize_model_int(
  modobj = ga_ogdq_formint_categorical_model$norm_devo_quotient_zscore,
  smooth_var = "gestational_age_zscore",
  int_var = "form_name",
  plabels = "categorical",
  derivative_plot = F) + 
    labs(x="Gestational Age (wks)",
         title = "",
         color = "Form Month") +
    theme(axis.title.y = element_blank()) +
      scale_x_continuous(
      breaks = map2_dbl(c(24*7, 28*7, 32*7, 36*7, 40*7), "gestational_age", ~ (.x - mean(swyc_avg[,.y], na.rm =T))/(sd(swyc_avg[,.y], na.rm = T))),
      labels = \(x) round(map2_dbl(x, "gestational_age", ~ (.x*sd(swyc_avg[,.y], na.rm = T) + mean(swyc_avg[,.y], na.rm = T))/7)))
    # scale_y_continuous(breaks = qnorm(c(.01, .50, .99)), 
    #                    labels = c("<=1st",  "50th", ">=99th")) +
    
ga_ogdq_formint_categorical_plot
```


#### with GA-ADJ DQ 

##### fit the no interaction model
```{r}
ga_ga_adjdq_model <- fit_bam_models(swyc_avg, "ga_adj_median_dq_zscore", 
                                        predictor = "gestational_age_zscore",
                                 save = FALSE)
  
tab <- extract_bam_table(models = ga_ga_adjdq_model, 
                                                  predictor ="gestational_age_zscore")
# Add the rest here manually, because bam() won't run with just an intercept (m_red crashes inside the function)

models = ga_ga_adjdq_model
predictor ="gestational_age_zscore"
data = swyc_avg 
covariates = NULL
random_effect = NULL

feats <- lapply(names(models), function(ph){
    
    m_full  <- models[[ph]]$full
    m_nolin <- models[[ph]]$nolin
    
    # rebuild reduced model WITHOUT predictor entirely
    covs <- if(!is.null(covariates)) paste(covariates, collapse = " + ") else ""
    re <- if(!is.null(random_effect)) paste0("s(", random_effect, ", bs='re')") else ""
    
    rhs <- paste(c(covs, re)[c(covs, re) != ""], collapse = " + ")

    f_red <- as.formula(paste(ph, "~ 1", 
                              if(rhs != "") paste("+", rhs)))
    
    f_onlylin <- as.formula(paste(ph, "~ 1 + ", predictor,
                                  if(rhs != "") paste("+", rhs)))
    
    # subset data to only include predictor is not NA to fit reduced model
    data_fit <- data %>% filter(!is.na(get(predictor)) & !is.na(get(ph)))
    #m_red <- mgcv:bam(f_red, data = data_fit, method = "fREML", discrete = TRUE)
    m_onlylin <- mgcv::bam(f_onlylin, data = data_fit, method = "fREML", discrete = TRUE)
    
    # partial R2: full predictor contribution vs no predictor model
    #r2_full_vs_none <- 1 - deviance(m_full)/deviance(m_red)
    #r2_full_vs_none <- (summary(m_full)$r.sq - summary(m_red)$r.sq) /  (1 - summary(m_red)$r.sq)
    
    # partial R2: spline contribution over linear model
    r2_spline_vs_linear <- 1 - deviance(m_full)/deviance(m_onlylin)
    #r2_spline_vs_linear <- (summary(m_full)$r.sq - summary(m_onlylin)$r.sq) /  (1 - summary(m_onlylin)$r.sq)
    
    # average first derivative
    d <- derivatives(m_full)
    avg_deriv <- mean(d$.derivative)
    
    # region of effect (CI crossing rule)
    d <- d %>%
      mutate (sig = !(0 >.lower_ci & 0 < .upper_ci))
    region <- if(sum(d$sig) > 0) c(min(d[d$sig == T, predictor]), max(d[d$sig == T, predictor])) else c(NA, NA)
    
    c(
      #r2_full_vs_none = r2_full_vs_none,
      r2_spline_vs_linear = r2_spline_vs_linear,
      avg_first_deriv = avg_deriv,
      region_start = region[1],
      region_end = region[2]
    )
  })
  
ga_ga_adjdq_table <- cbind(tab, do.call(rbind, feats))
  
```

##### print no interaction model table
```{r}
knitr::kable(t(ga_ga_adjdq_table))
```

##### visualize no interaction model
```{r}
ga_ga_adjdq_plot <- visualize_spline_model(
  modobj = ga_ga_adjdq_model$ga_adj_median_dq_zscore$full,
  plot_og_scale = FALSE,
  pnorm_smooth_var = FALSE,
  smooth_var = "gestational_age_zscore",
  plabels = "Gestational Age",
  color_point = ga_color,
  color_line = ga_color) + 
    scale_x_continuous(
      breaks = map2_dbl(c(24*7, 28*7, 32*7, 36*7, 40*7), "gestational_age", ~ (.x - mean(swyc_avg[,.y], na.rm =T))/(sd(swyc_avg[,.y], na.rm = T))),
      labels = \(x) round(map2_dbl(x, "gestational_age", ~ (.x*sd(swyc_avg[,.y], na.rm = T) + mean(swyc_avg[,.y], na.rm = T))/7))) +
    # scale_y_continuous(breaks = qnorm(c(.01, .50, .99)), 
    #                    labels = c("<=1st",  "50th", ">=99th")) +
    labs(x= "Gestational Age (weeks)",
         y = "GA-ADJ DQ (z-score)",
         title = "")

ga_ga_adjdq_plot

if(save_figs){
saveRDS(ga_ga_adjdq_plot, glue("{plot_output_folder}ga_ga_adjdq_spline_model_plot.rds"))
}

```

##### fit the model with form interaction (form as numerical with month)
```{r} 
ga_ga_adjdq_formint_model <- fit_bam_interactions(data = swyc_all,
                                        phenotypes = "norm_devo_quotient_zscore", 
                                        predictor = "gestational_age_zscore",
                                        moderator = "form_month_num",
                                        covariates = NULL, save = FALSE)
  
ga_ga_adjdq_formint_table <- extract_bam_features_interaction(models = ga_ga_adjdq_formint_model, 
                                                  data = swyc_all, 
                                                  predictor = "gestational_age_zscore",
                                                  moderator = "form_month_num")

knitr::kable(t(ga_ga_adjdq_formint_table))
  
```


##### fit the model with form interaction (form as categorical/ordered)
```{r} 
ga_ga_adjdq_formint_categorical_model <- fit_bam_interactions(swyc_all %>% 
                                                            filter(form_name %in% forms_totest), 
                                        "ga_adj_norm_devo_quotient_zscore", 
                                        predictor = "gestational_age_zscore",
                                        moderator = "form_name",
                                        interaction = "ordered",
                                        covariates = NULL, save = FALSE)
  
ga_ga_adjdq_formint_categorical_table <-  extract_bam_features_interaction(
                                                        models = ga_ga_adjdq_formint_categorical_model,
                                                          data = swyc_all %>% 
                                                            filter(form_name %in% forms_totest),
                                                          predictor = "gestational_age_zscore",
                                                          moderator = "form_name",
                                                       interaction_type = "ordered")

knitr::kable(t(ga_ga_adjdq_formint_categorical_table))
```

##### significance testing using log likelihood test for interaction term
shouldnt use fREML if comparing log ratio using ANOVA, using ML to fit models
```{r}
m1 <- fit_bam_interactions(swyc_all %>% 
                                                            filter(form_name %in% forms_totest), 
                                        "ga_adj_norm_devo_quotient_zscore", 
                                        predictor = "gestational_age_zscore",
                                        moderator = "form_name",
                                        interaction = "ordered",
                                        covariates = NULL, 
                                        save = FALSE,
                                        method = "ML")

m1 <- m1$ga_adj_norm_devo_quotient_zscore

m0 <- fit_bam_models(swyc_all %>% 
                       filter(form_name %in% forms_totest), "ga_adj_norm_devo_quotient_zscore", 
                     predictor = "gestational_age_zscore",
                                 save = FALSE,
                                method = "ML")

m0 <- m0$ga_adj_norm_devo_quotient_zscore$full

anova(m0, m1, test = "Chisq")
```

##### visualize form interaction model (categorical)
```{r}
font_size <- 16
theme_set(theme_classic(base_family = "Helvetica",base_size = font_size))
line_size <- 1.5
point_size <- 2

ga_ga_adjdq_formint_categorical_plot <- visualize_model_int(
  modobj = ga_ga_adjdq_formint_categorical_model$ga_adj_norm_devo_quotient_zscore,
  smooth_var = "gestational_age_zscore",
  int_var = "form_name",
  plabels = "categorical",
  derivative_plot = F) + 
    labs(x="Gestational Age (wks)",
         title = "",
         color = "Form Month") +
    theme(axis.title.y = element_blank()) +
      scale_x_continuous(
      breaks = map2_dbl(c(24*7, 28*7, 32*7, 36*7, 40*7), "gestational_age", ~ (.x - mean(swyc_avg[,.y], na.rm =T))/(sd(swyc_avg[,.y], na.rm = T))),
      labels = \(x) round(map2_dbl(x, "gestational_age", ~ (.x*sd(swyc_avg[,.y], na.rm = T) + mean(swyc_avg[,.y], na.rm = T))/7)))
    # scale_y_continuous(breaks = qnorm(c(.01, .50, .99)), 
    #                    labels = c("<=1st",  "50th", ">=99th")) +
    
ga_ga_adjdq_formint_categorical_plot
```

## SWYC and ADI

### Simple Correlations
```{r}
adi_vars <- c(
  "ADI.median",
  "ADI3_FINS.median",
  "ADI3_EDU.median",
  "ADI3_ECON.median"
)

ogDQ_adi_cor_results <- swyc_all %>%
  filter(form_friendly_name %in% forms_totest) %>%
  pivot_longer(
    cols = all_of(adi_vars),
    names_to = "ADI_type",
    values_to = "ADI_score"
  ) %>%
  group_by(form_friendly_name, ADI_type) %>%
  group_modify(~{
    ct <- cor.test(
      .x$norm_devo_quotient,
      .x$ADI_score,
      method = "pearson"
    )
    tibble(
      n = nrow(.x),
      r = unname(ct$estimate),
      lower = ct$conf.int[1],
      upper = ct$conf.int[2],
      p = ct$p.value
    )
  }) %>%
  ungroup()

(swyc_ogDQ_forms_adi_corrs_plot <- ggplot(ogDQ_adi_cor_results,
       aes(x = r,
           y = form_friendly_name,
           colour = ADI_type)) +
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "grey60") +
  geom_errorbarh(
    aes(xmin = lower, xmax = upper),
    height = 0.2
  ) +
  geom_point(size = 3) +
  facet_wrap(~ADI_type) +
  scale_x_continuous(
    limits = c(-1, 1),
    breaks = seq(-1, 1, 0.25)
  ) +
  labs(
    x = "Pearson correlation (r)",
    y = NULL,
    title = "OG DQ",
    colour = "ADI measure"
  ) +
  theme_classic())

ga_adjDQ_adi_cor_results <- swyc_all %>%
  filter(form_friendly_name %in% forms_totest) %>%
  pivot_longer(
    cols = all_of(adi_vars),
    names_to = "ADI_type",
    values_to = "ADI_score"
  ) %>%
  group_by(form_friendly_name, ADI_type) %>%
  group_modify(~{
    ct <- cor.test(
      .x$ga_adj_norm_devo_quotient,
      .x$ADI_score,
      method = "pearson"
    )
    tibble(
      n = nrow(.x),
      r = unname(ct$estimate),
      lower = ct$conf.int[1],
      upper = ct$conf.int[2],
      p = ct$p.value
    )
  }) %>%
  ungroup()

(swyc_ga_adjDQ_forms_adi_corrs_plot <- ggplot(ga_adjDQ_adi_cor_results,
       aes(x = r,
           y = form_friendly_name,
           colour = ADI_type)) +
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "grey60") +
  geom_errorbarh(
    aes(xmin = lower, xmax = upper),
    height = 0.2
  ) +
  geom_point(size = 3) +
  facet_wrap(~ADI_type) +
  scale_x_continuous(
    limits = c(-1, 1),
    breaks = seq(-1, 1, 0.25)
  ) +
  labs(
    x = "Pearson correlation (r)",
    y = NULL,
    title = "GA-ADJ DQ",
    colour = "ADI measure"
  ) +
  theme_classic())

if(save_figs == TRUE){
  panel_save(swyc_ogDQ_forms_adi_corrs_plot, "swyc_ogDQ_adi_cors", "svg", plot_output_folder)
  panel_save(swyc_ga_adjDQ_forms_adi_corrs_plot, "swyc_ga_adjDQ_adi_cors", "svg", plot_output_folder)
}
```

