---
title: "slip_swyc"
output: html_document
date: "2026-02-03"
editor_options: 
  chunk_output_type: console
---
Ideas
- get an average swyc score per participant and correlate with 4 global phenotypes + meanthickness + total SA + domain-based CMD & calculate 4 global phenotypes CMD
- try to figure out how to get a "trajectory" per participant

# Controls
```{r setup, echo = TRUE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)
save_figs = TRUE
save_tables = TRUE
save_processed = TRUE
```

# Load packages
```{r}
#Load packages. Install these packages with install.packages("pkg_name") if they are not installed.
library(bigrquery) #package to import SQL tables into R
require(dplyr)
require(glue)
require(magrittr)
require(readr)
require(stringr)
require(tidyr)
require(scales)
require(mgcv)
require(pbkrtest)
require(gamlssTools) # Margaret' package
require(tibble)
require(lme4)
require(purrr)
# For figures/plotting
require(ggplot2)
require(gridExtra)
require(ggsegDKT)
require(ggseg)
require(ggpubr)
require(ggeffects)
require(ggcorrplot)
require(cowplot)
require(ggh4x)
require(gratia)
require(egg)
require(interactions)
require(growthstandards)
require(gtsummary)
require(svglite)
require(patchwork)
```

# Define paths
```{r}
code_path <- "/mnt/arcus/lab/users/kafadare/slip_premie_wip/"
vars_path <-  "/mnt/arcus/lab/users/kafadare/slip_brain_vars/"

save_path <-  "/mnt/arcus/lab/users/kafadare/swyc_premie_centile_analysis_results/"
if(!dir.exists(save_path)) dir.create(save_path)

plot_output_folder <- paste0(save_path, "plots/")
if(!dir.exists(plot_output_folder)) dir.create(plot_output_folder)

csv_tables_folder <-  paste0(save_path, "processed_tables/")
if (!dir.exists(csv_tables_folder)) dir.create(csv_tables_folder)

csv_output_folder <-  paste0(save_path, "results_csv/")
if (!dir.exists(csv_output_folder)) dir.create(csv_output_folder)

spin_folder <- paste0(csv_output_folder, "spin_tables/")
if (!dir.exists(spin_folder)) dir.create(spin_folder)
```

# Load custom functions
```{r}
source(glue("{code_path}sliding_window.R"))
source(glue("{code_path}model_fitting_fcts.R"))
source(glue("{code_path}model_extract_fcts.R"))
source(glue("{code_path}model_vis_fcts.R"))
source(glue("{code_path}/slip_data_utils.R"))
source(glue("{code_path}/plotting_utils.R"))

panel_save <- function(plot, filename, extension, folder, plot_width = 8, plot_height = 10){
  save_name <- paste0(folder, filename, ".", extension)
  ggsave(save_name, plot = plot, dpi = 300, width = plot_width, height = plot_height)
}

`%nin%` = Negate(`%in%`)
```

# Colors
```{r}
predictor_hues <- list(
  gestational_age = c(dark = "#0B3D42", medium = "#3E7C89"),  # dark / medium petrol blue
  bwp_fen         = c(dark = "#1B4332", medium = "#52796F")   # dark / medium forest green
)

# Kept for the Model Visualization section, which colors points/lines by
# predictor directly (not through a pair comparison) -- single source of
# truth now instead of independently-set constants.
ga_color  <- predictor_hues$gestational_age[["medium"]]
bwp_color <- predictor_hues$bwp_fen[["medium"]]
swyc_color = "#9D73BA"

# For lollipop plots colored by effect direction (average first derivative
# sign) instead of by predictor/variant identity.
deriv_pos_color <- "#B85C5C"  # muted red  -- positive average derivative
deriv_neg_color <- "#5C7CB8"  # muted blue -- negative average derivative

font_size <- 16
theme_set(theme_classic(base_family = "Helvetica",base_size = font_size))
line_size <- 1.5
point_size <- 2
```

# Load centiles
```{r}
cent_avg_path <- "/mnt/arcus/lab/users/kafadare/swyc_premie_analysis_results/processed_tables/swyc_avgtwoyrs_formintwo_cent.csv"
cent_avg <- read.csv(cent_avg_path)

cent_all_path <- "/mnt/arcus/lab/users/kafadare/swyc_premie_analysis_results/processed_tables/swyc_allforms_cent.csv"
cent_all <- read.csv(cent_all_path)

load("/mnt/arcus/lab/users/kafadare/slip_centiles_processed_tables/lookup_stats_distinct.RData")

#determine if there is both mpr and median for thickness values
median_th_mpr_flag <- any(grepl(".median", names(cent_all)))


#lmsz_ga_centiles_path <- "/mnt/arcus/lab/users/kafadare/slip_centiles_processed_tables/lmsz_adjusted_centiles_ga.csv"
#lmsz_ga_centiles <- read.csv(lmsz_GA_centiles_path) %>% select(participant_id, session_id, sex, adjusted_age_in_days, age_at_scan, all_of(full_vars), grep("lmsz", names(.))) %>% mutate(lmsz_subset = "GA")

#lmsz_vpm_centiles_path <- "/mnt/arcus/lab/users/kafadare/slip_centiles_processed_tables/lmsz_adjusted_centiles_vpm.csv"
#lmsz_vpm_centiles <- read.csv(lmsz_vpm_centiles_path) %>% select(participant_id, session_id, sex, age_days, age_at_scan, full_vars, #grep("lmsz", names(.))) %>% mutate(lmsz_subset = "VPM")

#lmsz_lpm_centiles_path <- "/mnt/arcus/lab/users/kafadare/slip_centiles_processed_tables/lmsz_adjusted_centiles_lpm.csv"
#lmsz_lpm_centiles <- read.csv(lmsz_lpm_centiles_path) %>% select(participant_id, session_id, sex, age_days, age_at_scan, full_vars, #grep("lmsz", names(.))) %>% mutate(lmsz_subset = "LPM")

#lmsz_centiles <- bind_rows(lmsz_vpm_centiles, lmsz_lpm_centiles)

#lmsz_centiles <- lmsz_ga_centiles

#rm(lmsz_vpm_centiles, lmsz_lpm_centiles, lmsz_GA_centiles)

# df.zscore.distinct$pat_id <- sub("sub-", "", df.zscore.distinct$participant_id)
# df.zscore.all$pat_id <- sub("sub-", "", df.zscore.all$participant_id)
# centiles_all$pat_id <- sub("sub-", "", centiles_all$participant_id)
```

# Calculate swyc-avg for all shared pat-id (N form >= 2)
```{r}
## Take SWYC DQ Median Within SWYC forms that are <= 2 years from Scan
swyc_avg_full <- cent_all %>%
  group_by(pat_id) %>%
  filter(n() >= 2) %>%
  summarise(n_forms = n(),
            all_2s_sum = sum(all_2s == TRUE, na.rm = TRUE),
            all_0s_sum = sum(all_0s == TRUE, na.rm = TRUE),
            final_swyc_age = max(effective_age, na.rm = TRUE),
            final_swyc_age_ga_adj = max(ga_adj_effective_age),
            median_dq = median(norm_devo_quotient, na.rm = TRUE), 
            ga_adj_median_dq = median(ga_adj_norm_devo_quotient, na.rm = TRUE),
            final_swyc_date = effective_datetime[which.max(effective_age)],
            proc_ord_id = first(proc_ord_id),
            .groups = "drop")

cent_avg_full <- merge(cent_all %>% distinct(proc_ord_id, .keep_all = TRUE) %>%
  select(-c("form_friendly_name", "pat_form", "norm_devo_quotient",
            "ga_adj_norm_devo_quotient", "effective_age", "ga_adj_effective_age",
            "all_2s", "sum_answers", "all_0s", "effective_datetime", "scan_swyc_age_diff_days", 
            "scan_before_swyc")), swyc_avg_full, by = "proc_ord_id", all.x = FALSE, all.y = TRUE)

```
# Load vars
```{r}
full_vars <- read.csv(paste0(vars_path,"all_vars.csv")) %>% pull(x)
global_vars <- read.csv(paste0(vars_path,"global_vars.csv")) %>% pull(x)
vol_vars <- read.csv(paste0(vars_path,"vol_vars.csv")) %>% pull(x)
sa_vars <- read.csv(paste0(vars_path,"sa_vars.csv")) %>% pull(x)
th_vars <- read.csv(paste0(vars_path,"th_vars.csv")) %>% pull(x)
th_vars_median <- c(paste0(th_vars, ".median"))
namemap <- read.csv(glue("{vars_path}name_mapping.csv"))

## Define global vars of interest
global_vars_totest <- c("global_eTIV", "global_SubCortGrayVol","global_CortexVol", "global_CerebralWhiteMatterVol", "global_VentricleChoroidVol","global_SurfaceArea", "global_MeanThickness")

global_vars_totest_nothick <- c("global_eTIV", "global_SubCortGrayVol","global_CortexVol", "global_CerebralWhiteMatterVol", "global_VentricleChoroidVol","global_SurfaceArea")
```

# Find if any models don't exist, remove from variable list for downstream analysis.
```{r, echo = FALSE}
vars_missing <- full_vars[!(full_vars %in% names(cent_all))]
print(paste0("Phenotype Variables Missing in the Data Provided: ", paste(vars_missing, collapse = ", ")))
full_vars <- full_vars[!(full_vars %in% vars_missing)]
global_vars <- global_vars[!(global_vars %in% vars_missing)]
vol_vars <- vol_vars[!(vol_vars %in% vars_missing)]
sa_vars <- sa_vars[!(sa_vars %in% vars_missing)]
th_vars <- th_vars[!(th_vars %in% vars_missing)]
```

# Prep Data for Analysis

## zscore swyc-related variables and re-zscore non-centile variables

### convert GA, BW, etc to original raw values using original stats
```{r}
cent_all <- cent_all %>%
  mutate(across(all_of(c("gestational_age", 
                         "birth_weight_kg",
                         "birth_weight_kg_inc_out",
                         "age_at_scan", 
                         "age_at_scan_years", 
                         "age_days_adj",
                         "age_years_adj",
                         "qc_var", 
                         "euler_mean_mpr", 
                         "median_euler_mean")),
                     ~.x*sd_lookup_distinct[[cur_column()]] + mean_lookup_distinct[[cur_column()]]))
```

### zscore swyc avg dataframes
```{r, echo = FALSE}
# zscore centiles and other variables of interest
if (median_th_mpr_flag) {
df.zscore_avg <- cent_avg %>%
  mutate(across(all_of(c("gestational_age", 
                         "birth_weight_kg",
                         "birth_weight_kg_inc_out",
                         "age_at_scan", 
                         "age_at_scan_years", 
                         "age_days_adj",
                         "age_years_adj",
                         "qc_var", 
                         "euler_mean_mpr", 
                         "median_euler_mean",
                         "median_dq", 
                         "ga_adj_median_dq",
                         "ADI.median", 
                         "ADI3_FINS.median", 
                         "ADI3_ECON.median", 
                         "ADI3_EDU.median")), 
                ~as.numeric(scale(.x))))
} else {
  df.zscore_avg <- cent_avg %>%
  mutate(across(all_of(c("gestational_age", 
                         "birth_weight_kg",
                         "birth_weight_kg_inc_out",
                         "age_at_scan", 
                         "age_at_scan_years", 
                         "age_days_adj",
                         "age_years_adj",
                         "qc_var",
                         "median_dq", 
                         "ga_adj_median_dq",
                         "ADI.median", 
                         "ADI3_FINS.median", 
                         "ADI3_ECON.median", 
                         "ADI3_EDU.median")), 
                ~as.numeric(scale(.x))))
}

# set sex and preterm as factor
df.zscore_avg$sex <- as.factor(df.zscore_avg$sex)
df.zscore_avg$preterm <- as.factor(df.zscore_avg$preterm)

if (median_th_mpr_flag) {
df.zscore_avg_full <- cent_avg_full %>%
  mutate(across(all_of(c("gestational_age", 
                         "birth_weight_kg",
                         "birth_weight_kg_inc_out",
                         "age_at_scan", 
                         "age_at_scan_years", 
                         "age_days_adj",
                         "age_years_adj",
                         "qc_var", 
                         "euler_mean_mpr", 
                         "median_euler_mean",
                         "median_dq", 
                         "ga_adj_median_dq",
                         "ADI.median", 
                         "ADI3_FINS.median", 
                         "ADI3_ECON.median", 
                         "ADI3_EDU.median")), 
                ~as.numeric(scale(.x))))
} else {
  df.zscore_avg_full <- cent_avg_full %>%
  mutate(across(all_of(c("gestational_age", 
                         "birth_weight_kg",
                         "birth_weight_kg_inc_out",
                         "age_at_scan", 
                         "age_at_scan_years", 
                         "age_days_adj",
                         "age_years_adj",
                         "qc_var",
                         "median_dq", 
                         "ga_adj_median_dq",
                         "ADI.median", 
                         "ADI3_FINS.median", 
                         "ADI3_ECON.median", 
                         "ADI3_EDU.median")), 
                ~as.numeric(scale(.x))))
}

# set sex and preterm as factor
df.zscore_avg_full$sex <- as.factor(df.zscore_avg_full$sex)
df.zscore_avg_full$preterm <- as.factor(df.zscore_avg_full$preterm)
```

### zscore dataframe with all swyc forms
```{r}
# Add numeric column for form month
cent_all <- cent_all %>%
  mutate(
    form_month_num = as.numeric(sub("SWYC : ", "", form_friendly_name) %>%
      sub(" Month", "", .) %>% sub("0", "", .)),
    form_name = sub(" Month", "_Month", form_friendly_name) %>% sub("SWYC : ", "", .)) %>%
  select(!form_friendly_name) %>%
  mutate(
    pat_form = paste0(pat_id, "_", form_name)
  )

# zscore swyc all forms birth factors and DQ
cent_all <- cent_all %>%
  group_by(form_name) %>%
  mutate(across(all_of(c("norm_devo_quotient", "ga_adj_norm_devo_quotient", 
                         "effective_age", "ga_adj_effective_age",
                         "gestational_age", 
                         "birth_weight_kg",
                         "birth_weight_kg_inc_out",
                         "age_at_scan", 
                         "age_at_scan_years", 
                         "age_days_adj",
                         "age_years_adj",
                         "qc_var", 
                         "euler_mean_mpr", 
                         "median_euler_mean",
                         "ADI.median", 
                         "ADI3_FINS.median", 
                         "ADI3_ECON.median", 
                         "ADI3_EDU.median")),
                ~as.numeric(scale(.x)), .names = "{.col}_zscore")) %>%
  ungroup()

# set sex as factor
cent_all$sex <- as.factor(cent_all$sex)

# change form name order factor levels
form_order <- c("04_Month",
                "06_Month", 
                "09_Month", 
                "12_Month",
                "15_Month",
                "18_Month", 
                "24_Month", 
                "30_Month",
                "36_Month",
                "48_Month",
                "60_Month")

cent_all <- cent_all %>%
  mutate(
    form_name = factor(
      form_name,
      levels = form_order,
      ordered = TRUE
    )
  )
      
cent_all_wide <- cent_all %>%
  pivot_wider(
    id_cols = c(pat_id, gestational_age, birth_weight_kg, bwp_fen),
    names_from = form_name,
    values_from = c(norm_devo_quotient, ga_adj_norm_devo_quotient, effective_age, ga_adj_effective_age,
                    ADI.median, ADI3_FINS.median, ADI3_ECON.median, ADI3_EDU.median),
    names_glue = "{.value}_{form_name}"
  )

dq_vars <- names(cent_all_wide)[grepl("norm_devo", names(cent_all_wide))]
adi_vars <- names(cent_all_wide)[grepl("ADI", names(cent_all_wide))]
vars_tozscore <- names(cent_all_wide)[grepl("norm_devo|ADI|effective_age", names(cent_all_wide))]

# zscore swyc avg birth factors and DQ wide format
cent_all_wide <- cent_all_wide %>%
  mutate(across(all_of(c(vars_tozscore,
                         "gestational_age", "birth_weight_kg")),
                ~as.numeric(scale(.x)), .names = "{.col}_zscore"))
```

# Select which SWYC forms to investigate (based on available N)
```{r}
tab_swyc_forms <- table(cent_all$form_name)
data.frame(form = names(tab_swyc_forms), count = as.vector(tab_swyc_forms))

forms_totest <- c("06_Month", 
                       "09_Month",
                       "15_Month",
                       "18_Month",
                       "24_Month",
                       "30_Month")

og_dq_vars_totest <- c("06_Month", 
                       "09_Month",
                       "15_Month",
                       "18_Month",
                       "24_Month",
                       "30_Month")

ga_adj_dq_vars_totest <- paste0("ga_adj_", og_dq_vars_totest)
```

# CMDs

## Define CMD vars
```{r}
cmd_vars <- c("global.cmd", "vol.cmd", "sa.cmd", "th.cmd", "subcort.cmd")
```

## Correlation Matrix with median SWYC
```{r, echo = FALSE}
# correlations between CMDs across three centiles
cor_mat <- cor(df.zscore_avg[,c("global.cmd", "vol.cmd", "sa.cmd", "th.cmd", "subcort.cmd",
                                    "gestational_age", "birth_weight_kg",
                                "bwp_fen",
                                    "median_dq", "ga_adj_median_dq")], 
               use = "pairwise.complete.obs")

p_mat <- cor_pmat(df.zscore_avg[,c("global.cmd", "vol.cmd", "sa.cmd", "th.cmd", "subcort.cmd",
                                    "gestational_age", "birth_weight_kg",
                                "bwp_fen",
                                    "median_dq", "ga_adj_median_dq")], 
               use = "pairwise.complete.obs")


(cmd_corrs_plot <- ggcorrplot(
  cor_mat,
  p.mat = p_mat,
  insig = "blank",
  lab = TRUE
))

if(save_figs == TRUE){
panel_save(cmd_corrs_plot, "cmd-cor-matrix", "svg", plot_output_folder)
}

```



## Correlations with All SWYC Forms
```{r}
ogDQ_cmd_cor_results <- cent_all %>%
  filter(form_name %in% forms_totest) %>%
  pivot_longer(
    cols = all_of(cmd_vars),
    names_to = "cmd_type",
    values_to = "cmd_score"
  ) %>%
  group_by(form_name, cmd_type) %>%
  group_modify(~{
    ct <- cor.test(
      .x$norm_devo_quotient,
      .x$cmd_score,
      method = "pearson"
    )
    tibble(
      n = nrow(.x),
      r = unname(ct$estimate),
      lower = ct$conf.int[1],
      upper = ct$conf.int[2],
      p = ct$p.value
    )
  }) %>%
  ungroup()

(swyc_ogDQ_forms_cmd_corrs_plot <- ggplot(ogDQ_cmd_cor_results,
       aes(x = r,
           y = form_name,
           colour = cmd_type)) +
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "grey60") +
  geom_errorbarh(
    aes(xmin = lower, xmax = upper),
    height = 0.2
  ) +
  geom_point(size = 3) +
  facet_wrap(~cmd_type) +
  scale_x_continuous(
    limits = c(-1, 1),
    breaks = seq(-1, 1, 0.25)
  ) +
  labs(
    x = "Pearson correlation (r)",
    y = NULL,
    title = "OG DQ",
    colour = "cmd measure"
  ) +
  theme_classic())

adjdq_cmd_cor_results <- cent_all %>%
  filter(form_name %in% forms_totest) %>%
  pivot_longer(
    cols = all_of(cmd_vars),
    names_to = "cmd_type",
    values_to = "cmd_score"
  ) %>%
  group_by(form_name, cmd_type) %>%
  group_modify(~{
    ct <- cor.test(
      .x$ga_adj_norm_devo_quotient,
      .x$cmd_score,
      method = "pearson"
    )
    tibble(
      n = nrow(.x),
      r = unname(ct$estimate),
      lower = ct$conf.int[1],
      upper = ct$conf.int[2],
      p = ct$p.value
    )
  }) %>%
  ungroup()

(swyc_adjdq_forms_cmd_corrs_plot <- ggplot(adjdq_cmd_cor_results,
       aes(x = r,
           y = form_name,
           colour = cmd_type)) +
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "grey60") +
  geom_errorbarh(
    aes(xmin = lower, xmax = upper),
    height = 0.2
  ) +
  geom_point(size = 3) +
  facet_wrap(~cmd_type) +
  scale_x_continuous(
    limits = c(-1, 1),
    breaks = seq(-1, 1, 0.25)
  ) +
  labs(
    x = "Pearson correlation (r)",
    y = NULL,
    title = "GA-ADJ DQ",
    colour = "cmd measure"
  ) +
  theme_classic())

if(save_figs == TRUE){
  panel_save(swyc_ogDQ_forms_cmd_corrs_plot, "swyc_ogDQ_cmd_cors", "svg", plot_output_folder)
  panel_save(swyc_adjdq_forms_cmd_corrs_plot, "swyc_adjdq_cmd_cors", "svg", plot_output_folder)
}
```




## Spline with OG DQ

### fit spline models to median
```{r}
ogdq_sp_models_cmd <- fit_bam_models(df.zscore_avg, cmd_vars, 
                                        predictor = "median_dq",
                                        covariates = c("qc_var", "global_eTIV"), save = FALSE)
  
ogdq_sp_table_cmd <- extract_bam_features_main(models = ogdq_sp_models_cmd, 
                                                  data = df.zscore_avg, 
                                                  predictor = "median_dq",
                                                  covariates = c("qc_var", "global_eTIV"))

knitr::kable(t(ogdq_sp_table_cmd))

ogdq_sp_table_cmd <- ogdq_sp_table_cmd %>% mutate(signed_partial_r2 = r2_full_vs_none*sign(avg_first_deriv),
                                                  phenotype = rownames(.))
```

### lollipop, signed partial R
```{r}
df           <- ogdq_sp_table_cmd
value_col    <- "signed_partial_r2"
p_col        <- "full_p_adj"

p <- plot_lollipop(
  table_list = df,
  value_col = value_col,
  p_col = p_col,
  color = swyc_color,
  name_mapping = namemap
) +
  labs(title = glue("SWYC Median OG DQ (Partial R2)"), x = value_col) +
  theme_minimal()

p

if (params$save_plots == TRUE) {
  ggsave(glue("{plot_output_folder}lollipop_swyc_cmd.svg"), p, width = 9, height = 6)
}
```

### visualize model with median dq
```{r}
ogdq_sp_plots_cmd <- lapply(ogdq_sp_models_cmd, function(x) {
  
  modobj = x$full
  
  model_plot <- visualize_spline_model(
  modobj = modobj,
  plot_og_scale = FALSE,
  mean_lookup = mean_lookup_distinct,
  sd_lookup = sd_lookup_distinct,
  pnorm_smooth_var = FALSE,
  smooth_var = "median_dq",
  plabels = "OG DQ",
  color_point = swyc_color,
  color_line = swyc_color)
  
  model_plot <- model_plot + 
    labs(x= "OG SWYC (z-scored)",
         title = glue("{modobj$terms[[2]]}")) +
    theme(axis.title.y = element_blank())
})

ogdq_sp_plots_cmd  <- setNames(ogdq_sp_plots_cmd , cmd_vars)

if(save_figs){
ggsave(glue("{plot_output_folder}swyc_ogdq_volcmd_model.svg"), ogdq_sp_plots_cmd$vol.cmd, width = 9, height = 6)
}
```

### fit interactions with form type
```{r}
ogdq_sp_int_models_cmd <- fit_bam_interactions(cent_all %>% filter(form_name %in% forms_totest), 
                                               cmd_vars, 
                                        predictor = "norm_devo_quotient_zscore",
                                        moderator = "form_name",
                                        interaction = "ordered",
                                        covariates = c("qc_var"), 
                                        save = FALSE)

ogdq_sp_int_table_cmd <- extract_bam_features_interaction(models = ogdq_sp_int_models_cmd,
                                                          data = cent_all %>% 
                                                            filter(form_name %in% forms_totest),
                                                          predictor = "norm_devo_quotient_zscore",
                                                          moderator = "form_name",
                                                          covariates = c("qc_var"),
                                                       interaction_type = "ordered")

knitr::kable(t(ogdq_sp_int_table_cmd))
```

### test significance by log ratio test (ML to fit)
```{r}
ogdq_int_lrt_cmd <- test_interaction_lrt(data = cent_all %>% filter(form_name %in% forms_totest), 
                                 phenotypes = cmd_vars,
                                predictor = "norm_devo_quotient_zscore",
                                moderator = "form_name",
                                interaction = "ordered",
                                covariates = c("qc_var"), 
                                save = FALSE,
                                 method = "ML")

knitr::kable((ogdq_int_lrt_cmd))
```

### visualize form interaction model (ordered) ** I think the significant interactions are due to data, and not meaningful?
```{r}
font_size <- 16
theme_set(theme_classic(base_family = "Helvetica",base_size = font_size))
line_size <- 1.5
point_size <- 2

ogdq_sp_int_plots_cmd <- lapply(ogdq_sp_int_models_cmd, function(x) {
  
  model_plot <- visualize_model_int(
  modobj = x,
  smooth_var = "norm_devo_quotient_zscore",
  int_var = "form_name",
  plabels = "ordered",
  derivative_plot = F) + 
    labs(x="OG DQ Median (z-scored)",
         title = glue("{x$terms[[2]]}"),
         color = "Form Month") +
    theme(axis.title.y = element_blank())
    
})

ogdq_sp_int_plots_cmd <- setNames(ogdq_sp_int_plots_cmd, cmd_vars)


if(save_figs){
ggsave(glue("{plot_output_folder}swyc_ogdq_volcmd_model.svg"), ogdq_sp_plots_cmd$vol.cmd, width = 9, height = 6)
}
```


## Spline with GA-ADJ DQ

### fit spline models
```{r}
adjdq_sp_models_cmd <- fit_bam_models(df.zscore_avg, cmd_vars, 
                                        predictor = "ga_adj_median_dq",
                                        covariates = c("qc_var"), save = FALSE)
  
adjdq_sp_table_cmd <- extract_bam_features_main(models = adjdq_sp_models_cmd, 
                                                  data = df.zscore_avg, 
                                                  predictor = "ga_adj_median_dq",
                                                  covariates = c("qc_var"))

knitr::kable(t(adjdq_sp_table_cmd))
```

### fit interactions with form type
```{r}
adjdq_sp_int_models_cmd <- fit_bam_interactions(cent_all %>% filter(form_name %in% forms_totest), 
                                               cmd_vars, 
                                        predictor = "ga_adj_norm_devo_quotient_zscore",
                                        moderator = "form_name",
                                        interaction = "ordered",
                                        covariates = c("qc_var"), 
                                        save = FALSE)

adjdq_sp_int_table_cmd <- extract_bam_features_interaction(models = adjdq_sp_int_models_cmd,
                                                          data = cent_all %>% 
                                                            filter(form_name %in% forms_totest),
                                                          predictor = "ga_adj_norm_devo_quotient_zscore",
                                                          moderator = "form_name",
                                                          covariates = c("qc_var"),
                                                       interaction_type = "ordered")

knitr::kable(t(adjdq_sp_int_table_cmd))
```

## Compare R^2 between OG and GA-ADJ
```{r}
tables_list <- list(
  OG = ogdq_sp_table_cmd %>% mutate(phenotype = rownames(.)),
  GA_ADJ = adjdq_sp_table_cmd %>% mutate(phenotype = rownames(.))
)

color_vec <- c(OG = swyc_color, GA_ADJ = ga_adj_color)

## Main effects: Original DQ vs GA-adjusted DQ
plot_lollipop(
  table_list = tables_list,
  value_col = "r2_full_vs_none",
  color_vec = color_vec,
  p_col = "full_p_adj"
) +
  labs(
    x = expression("Partial " * R^2),
    color = NULL
  )
```


## RESI plots for median DQ
```{r}
ogdq_resi_table_cmd <- extract_bam_resi(models = ogdq_sp_models_cmd,
                                           term = "s(median_dq)")

adjdq_resi_table_cmd <- extract_bam_resi(models = adjdq_sp_models_cmd,
                                           term = "s(ga_adj_median_dq)")
tables_list <- list(
  OG = ogdq_resi_table_cmd %>% mutate(phenotype = rownames(.)),
  GA_ADJ = adjdq_resi_table_cmd %>% mutate(phenotype = rownames(.))
)

color_vec <- c(OG = swyc_color, GA_ADJ = ga_adj_color)

## Main effects: Original DQ vs GA-adjusted DQ
plot_lollipop(
  table_list = tables_list,
  value_col = "RESI",
  color_vec = color_vec,
  p_col = "p_adj"
) +
  labs(
    x = "RESI from F-statistic",
    color = NULL
  )
```

## Fixed DF spline for RESI
```{r}
ogdq_sp_fx_models_cmd <- fit_bam_models(df.zscore_avg, cmd_vars, 
                                        predictor = "median_dq",
                                        covariates = c("qc_var"), 
                                        save = FALSE,
                                        fx = TRUE,
                                        k = 10)
  
ogdq_sp_fx_table_cmd <- extract_bam_features_main(models = ogdq_sp_fx_models_cmd, 
                                                  data = df.zscore_avg, 
                                                  predictor = "median_dq",
                                                  covariates = c("qc_var"))

ogdq_fx_resi_table_cmd <- extract_bam_resi(models = ogdq_sp_fx_models_cmd,
                                           term = "s(median_dq)")


tables_list <- list(
  PENALIZED = ogdq_resi_table_cmd %>% mutate(phenotype = rownames(.)),
  FIXED = ogdq_fx_resi_table_cmd %>% mutate(phenotype = rownames(.))
)

color_vec <- c(PENALIZED = "pink", FIXED = "red")

## Main effects: Original DQ vs GA-adjusted DQ
plot_lollipop(
  table_list = tables_list,
  value_col = "RESI",
  color_vec = color_vec,
  p_col = "p_adj"
) +
  labs(
    x = "RESI from F-statistic",
    color = NULL
  )
```

## Plot DF for fixed vs penalized
```{r}
tables_list <- list(
  PENALIZED = ogdq_sp_table_cmd %>% mutate(phenotype = rownames(.)),
  FIXED = ogdq_sp_fx_table_cmd %>% mutate(phenotype = rownames(.))
)

color_vec <- c(PENALIZED = "pink", FIXED = "red")

## Main effects: Original DQ vs GA-adjusted DQ
plot_lollipop(
  table_list = tables_list,
  value_col = "full_edf",
  color_vec = color_vec,
  p_col = "p_adj"
) +
  labs(
    x = "EDF",
    color = NULL
  )
```

## Visualize models for fixed DF
```{r}
ogdq_sp_fx_plots_cmd <- lapply(ogdq_sp_fx_models_cmd, function(x) {
  
  modobj = x$full
  
  model_plot <- visualize_spline_model(
  modobj = modobj,
  plot_og_scale = FALSE,
  mean_lookup = mean_lookup_distinct,
  sd_lookup = sd_lookup_distinct,
  pnorm_smooth_var = FALSE,
  smooth_var = "median_dq",
  plabels = "OG DQ",
  color_point = swyc_color,
  color_line = swyc_color)
  
  model_plot <- model_plot + 
    labs(x= "OG SWYC (z-scored)",
         title = glue("{modobj$terms[[2]]}")) +
    theme(axis.title.y = element_blank())
})

ogdq_sp_fx_plots_cmd  <- setNames(ogdq_sp_fx_plots_cmd , cmd_vars)

wrap_plots(ogdq_sp_fx_plots_cmd[1:5], ncol = 2)
```

# Global Variables

## Correlation Matrix with median SWYC
```{r, echo = FALSE}
# correlations between globals across three centiles
cor_mat <- cor(df.zscore_avg[,c(global_vars_totest,
                                    "gestational_age", "birth_weight_kg",
                                "bwp_fen",
                                    "median_dq", "ga_adj_median_dq")], 
               use = "pairwise.complete.obs")

p_mat <- cor_pmat(df.zscore_avg[,c(global_vars_totest,
                                    "gestational_age", "birth_weight_kg",
                                "bwp_fen",
                                    "median_dq", "ga_adj_median_dq")], 
               use = "pairwise.complete.obs")


(global_corrs_plot <- ggcorrplot(
  cor_mat,
  p.mat = p_mat,
  insig = "blank",
  lab = TRUE
))

if(save_figs == TRUE){
panel_save(global_corrs_plot, "global-cor-matrix", "svg", plot_output_folder)
}

```



## Correlations with All SWYC Forms
```{r}
ogDQ_global_cor_results <- cent_all %>%
  filter(form_name %in% forms_totest) %>%
  pivot_longer(
    cols = all_of(global_vars_totest),
    names_to = "global_type",
    values_to = "global_score"
  ) %>%
  group_by(form_name, global_type) %>%
  group_modify(~{
    ct <- cor.test(
      .x$norm_devo_quotient,
      .x$global_score,
      method = "pearson"
    )
    tibble(
      n = nrow(.x),
      r = unname(ct$estimate),
      lower = ct$conf.int[1],
      upper = ct$conf.int[2],
      p = ct$p.value
    )
  }) %>%
  ungroup()

(swyc_ogDQ_forms_global_corrs_plot <- ggplot(ogDQ_global_cor_results,
       aes(x = r,
           y = form_name,
           colour = global_type)) +
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "grey60") +
  geom_errorbarh(
    aes(xmin = lower, xmax = upper),
    height = 0.2
  ) +
  geom_point(size = 3) +
  facet_wrap(~global_type) +
  scale_x_continuous(
    limits = c(-1, 1),
    breaks = seq(-1, 1, 0.25)
  ) +
  labs(
    x = "Pearson correlation (r)",
    y = NULL,
    title = "OG DQ",
    colour = "global measure"
  ) +
  theme_classic())

adjdq_global_cor_results <- cent_all %>%
  filter(form_name %in% forms_totest) %>%
  pivot_longer(
    cols = all_of(global_vars_totest),
    names_to = "global_type",
    values_to = "global_score"
  ) %>%
  group_by(form_name, global_type) %>%
  group_modify(~{
    ct <- cor.test(
      .x$ga_adj_norm_devo_quotient,
      .x$global_score,
      method = "pearson"
    )
    tibble(
      n = nrow(.x),
      r = unname(ct$estimate),
      lower = ct$conf.int[1],
      upper = ct$conf.int[2],
      p = ct$p.value
    )
  }) %>%
  ungroup()

(swyc_adjdq_forms_global_corrs_plot <- ggplot(adjdq_global_cor_results,
       aes(x = r,
           y = form_name,
           colour = global_type)) +
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "grey60") +
  geom_errorbarh(
    aes(xmin = lower, xmax = upper),
    height = 0.2
  ) +
  geom_point(size = 3) +
  facet_wrap(~global_type) +
  scale_x_continuous(
    limits = c(-1, 1),
    breaks = seq(-1, 1, 0.25)
  ) +
  labs(
    x = "Pearson correlation (r)",
    y = NULL,
    title = "GA-ADJ DQ",
    colour = "global measure"
  ) +
  theme_classic())

if(save_figs == TRUE){
  panel_save(swyc_ogDQ_forms_global_corrs_plot, "swyc_ogDQ_global_cors", "svg", plot_output_folder)
  panel_save(swyc_adjdq_forms_global_corrs_plot, "swyc_adjdq_global_cors", "svg", plot_output_folder)
}
```

## Splines with OG DQ

### fit spline models with median DQ
```{r}
ogdq_sp_models_eTIV<- fit_bam_models(df.zscore_avg,
                            c("global_eTIV"), 
                                      predictor = "median_dq",
                                      covariates = c("qc_var"), save = FALSE)

ogdq_sp_table_eTIV <- extract_bam_features_main(models = ogdq_sp_models_eTIV, 
                                                data = df.zscore_avg, 
                                                predictor = "median_dq",
                                                covariates = c("qc_var"))

ogdq_sp_models_global <- fit_bam_models(df.zscore_avg,
                            c("global_SubCortGrayVol","global_CortexVol",
                                        "global_CerebralWhiteMatterVol",
                                        "global_VentricleChoroidVol","global_SurfaceArea"), 
                                      predictor = "median_dq",
                                      covariates = c("qc_var", "global_eTIV"), save = FALSE)

ogdq_sp_table_global <- extract_bam_features_main(models = ogdq_sp_models_global, 
                                                data = df.zscore_avg, 
                                                predictor = "median_dq",
                                                covariates = c("qc_var", "global_eTIV"))


ogdq_sp_models_meanthick <- fit_bam_models(df.zscore_avg, "global_MeanThickness", 
                                      predictor = "median_dq",
                                      covariates = c("euler_mean_mpr", "global_eTIV"), save = FALSE)

ogdq_sp_table_meanthick <- extract_bam_features_main(models = ogdq_sp_models_meanthick, 
                                                data = df.zscore_avg, 
                                                predictor = "median_dq",
                                                covariates = c("euler_mean_mpr", "global_eTIV"))

ogdq_sp_models_global <- c(ogdq_sp_models_eTIV, ogdq_sp_models_global, ogdq_sp_models_meanthick)
ogdq_sp_table_global <- rbind(ogdq_sp_table_eTIV, ogdq_sp_table_global, ogdq_sp_table_meanthick)

rm(ogdq_sp_models_meanthick, ogdq_sp_table_meanthick, ogdq_sp_models_eTIV, ogdq_sp_table_eTIV)
  
knitr::kable(t(ogdq_sp_table_global[1:4,]))
knitr::kable(t(ogdq_sp_table_global[5:7,]))
```


### Show lollipop plot for OG DQ Global
```{r}
p <- plot_lollipop(
  table_list = ogdq_sp_table_global %>% mutate(phenotype = rownames(.), 
                                               signed_partial_r2 = r2_full_vs_none*sign(avg_first_deriv)),
  value_col = "signed_partial_r2",
  p_col = "full_p_adj",
  color = swyc_color,
  name_mapping = namemap
) +
  labs(
    x = expression("Signed Partial " * R^2),
    color = swyc_color
  )

ggsave(glue("{plot_output_folder}lollipop_swyc_global.svg"), p, width = 9, height = 6)
```

### visualize no interaction model
```{r}
ogdq_sp_plots_global <- lapply(ogdq_sp_models_global, function(x) {
  
  modobj = x$full
  
  model_plot <- visualize_spline_model(
  modobj = modobj,
  plot_og_scale = FALSE,
  mean_lookup = mean_lookup_distinct,
  sd_lookup = sd_lookup_distinct,
  pnorm_smooth_var = FALSE,
  smooth_var = "median_dq",
  plabels = "OG DQ",
  color_point = swyc_color,
  color_line = swyc_color)
  
  model_plot <- model_plot + 
    labs(x= "OG SWYC (z-scored)",
         title = glue("{modobj$terms[[2]]}")) +
    theme(axis.title.y = element_blank())
})

ogdq_sp_plots_global  <- setNames(ogdq_sp_plots_global , global_vars_totest)

ggsave(glue("{plot_output_folder}swyc_ogdq_vent_model.svg"), ogdq_sp_plots_global$global_VentricleChoroidVol, width = 9, height = 6)
ggsave(glue("{plot_output_folder}swyc_ogdq_subcortgray_model.svg"), ogdq_sp_plots_global$global_SubCortGrayVol, width = 9, height = 6)
```

### fit interactions with form type
```{r}
ogdq_sp_int_models_global <- fit_bam_interactions(cent_all %>% filter(form_name %in% forms_totest),
                            c("global_eTIV","global_SubCortGrayVol","global_CortexVol",
                                        "global_CerebralWhiteMatterVol",
                                        "global_VentricleChoroidVol","global_SurfaceArea"), 
                                        predictor = "norm_devo_quotient_zscore",
                                      moderator = "form_name",
                                      interaction = "ordered",
                                      covariates = c("qc_var"), save = FALSE)

ogdq_sp_int_table_global <- extract_bam_features_interaction(models = ogdq_sp_int_models_global,
                                                          data = cent_all %>% 
                                                            filter(form_name %in% forms_totest),
                                                          predictor = "norm_devo_quotient_zscore",
                                                          moderator = "form_name",
                                                          covariates = c("qc_var"),
                                                       interaction_type = "ordered")

ogdq_sp_int_models_meanthick <- fit_bam_interactions(cent_all %>% filter(form_name %in% forms_totest),
                            "global_MeanThickness", 
                                        predictor = "norm_devo_quotient_zscore",
                                      moderator = "form_name",
                                      interaction = "ordered",
                                      covariates = c("euler_mean_mpr"), save = FALSE)

ogdq_sp_int_table_meanthick <- extract_bam_features_interaction(models = ogdq_sp_int_models_meanthick,
                                                          data = cent_all %>% 
                                                            filter(form_name %in% forms_totest),
                                                          predictor = "norm_devo_quotient_zscore",
                                                          moderator = "form_name",
                                                          covariates = c("qc_var"),
                                                       interaction_type = "ordered")

ogdq_sp_int_models_global <- c(ogdq_sp_int_models_global, ogdq_sp_int_models_meanthick)
ogdq_sp_int_table_global <- rbind(ogdq_sp_int_table_global, ogdq_sp_int_table_meanthick)

rm(ogdq_sp_int_models_meanthick, ogdq_sp_int_table_meanthick)
  
knitr::kable(t(ogdq_sp_int_table_global[1:4,]))
knitr::kable(t(ogdq_sp_int_table_global[5:7,]))
```

### significance testing using log likelihood of interaction term
```{r}
ogdq_int_lrt_global <- test_interaction_lrt(data = cent_all %>% filter(form_name %in% forms_totest), 
                                 phenotypes = c("global_eTIV","global_SubCortGrayVol","global_CortexVol",
                                        "global_CerebralWhiteMatterVol",
                                        "global_VentricleChoroidVol","global_SurfaceArea"),
                                predictor = "norm_devo_quotient_zscore",
                                moderator = "form_name",
                                interaction = "ordered",
                                covariates = c("qc_var"), 
                                save = FALSE,
                                 method = "ML")

knitr::kable((ogdq_int_lrt_global))
```
Non-significant form interactions for global phenotypes

## Splines with GA-ADJ DQ

### fit spline models
```{r}
adjdq_sp_models_global <- fit_bam_models(df.zscore_avg,
                            c("global_eTIV","global_SubCortGrayVol","global_CortexVol",
                                        "global_CerebralWhiteMatterVol",
                                        "global_VentricleChoroidVol","global_SurfaceArea"), 
                                      predictor = "ga_adj_median_dq",
                                      covariates = c("qc_var"), save = FALSE)

adjdq_sp_table_global <- extract_bam_features_main(models = adjdq_sp_models_global, 
                                                data = df.zscore_avg, 
                                                predictor = "ga_adj_median_dq",
                                                covariates = c("qc_var"))


adjdq_sp_models_meanthick <- fit_bam_models(df.zscore_avg, "global_MeanThickness", 
                                      predictor = "ga_adj_median_dq",
                                      covariates = c("euler_mean_mpr"), save = FALSE)

adjdq_sp_table_meanthick <- extract_bam_features_main(models = adjdq_sp_models_meanthick, 
                                                data = df.zscore_avg, 
                                                predictor = "ga_adj_median_dq",
                                                covariates = c("euler_mean_mpr"))

adjdq_sp_models_global <- c(adjdq_sp_models_global, adjdq_sp_models_meanthick)
adjdq_sp_table_global <- rbind(adjdq_sp_table_global, adjdq_sp_table_meanthick)

rm(adjdq_sp_models_meanthick, adjdq_sp_table_meanthick)
  
knitr::kable(t(adjdq_sp_table_global[1:4,]))
knitr::kable(t(adjdq_sp_table_global[5:7,]))
```

### visualize no interaction model
```{r}
adjdq_sp_plots_global <- lapply(adjdq_sp_models_global, function(x) {
  
  modobj = x$full
  
  model_plot <- visualize_spline_model(
  modobj = modobj,
  plot_og_scale = FALSE,
  mean_lookup = mean_lookup_distinct,
  sd_lookup = sd_lookup_distinct,
  pnorm_smooth_var = FALSE,
  smooth_var = "ga_adj_median_dq",
  plabels = "GA-ADJ DQ",
  color_point = swyc_color,
  color_line = swyc_color)
  
  model_plot <- model_plot + 
    scale_y_continuous(breaks = qnorm(c(.01, .50, .99)),
                       labels = c("<=1st",  "50th", ">=99th")) +
    labs(x= "GA_ADJ SWYC (z-scored)",
         title = glue("{modobj$terms[[2]]}")) +
    theme(axis.title.y = element_blank())
})

adjdq_sp_plots_global  <- setNames(adjdq_sp_plots_global , global_vars_totest)

wrap_plots(adjdq_sp_plots_global[1:4], ncol = 2)
wrap_plots(adjdq_sp_plots_global[5:7], ncol = 2)

if(save_figs){
saveRDS(adjdq_sp_plots_global, glue("{plot_output_folder}adjdq_sp_modelvis_plots_global.rds"))
}
```

### fit interactions with form type
```{r}
adjdq_sp_int_models_global <- fit_bam_interactions(cent_all %>% filter(form_name %in% forms_totest),
                            c("global_eTIV","global_SubCortGrayVol","global_CortexVol",
                                        "global_CerebralWhiteMatterVol",
                                        "global_VentricleChoroidVol","global_SurfaceArea"), 
                                        predictor = "ga_adj_norm_devo_quotient_zscore",
                                      moderator = "form_name",
                                      interaction = "ordered",
                                      covariates = c("qc_var"), save = FALSE)

adjdq_sp_int_table_global <- extract_bam_features_interaction(models = adjdq_sp_int_models_global,
                                                          data = cent_all %>% 
                                                            filter(form_name %in% forms_totest),
                                                          predictor = "ga_adj_norm_devo_quotient_zscore",
                                                          moderator = "form_name",
                                                          covariates = c("qc_var"),
                                                       interaction_type = "ordered")

adjdq_sp_int_models_meanthick <- fit_bam_interactions(cent_all %>% filter(form_name %in% forms_totest),
                            "global_MeanThickness", 
                                        predictor = "ga_adj_norm_devo_quotient_zscore",
                                      moderator = "form_name",
                                      interaction = "ordered",
                                      covariates = c("euler_mean_mpr"), save = FALSE)

adjdq_sp_int_table_meanthick <- extract_bam_features_interaction(models = adjdq_sp_int_models_meanthick,
                                                          data = cent_all %>% 
                                                            filter(form_name %in% forms_totest),
                                                          predictor = "ga_adj_norm_devo_quotient_zscore",
                                                          moderator = "form_name",
                                                          covariates = c("qc_var"),
                                                       interaction_type = "ordered")

adjdq_sp_int_models_global <- c(adjdq_sp_int_models_global, adjdq_sp_int_models_meanthick)
adjdq_sp_int_table_global <- rbind(adjdq_sp_int_table_global, adjdq_sp_int_table_meanthick)

rm(adjdq_sp_int_models_meanthick, adjdq_sp_int_table_meanthick)
  
knitr::kable(t(adjdq_sp_int_table_global[1:4,]))
knitr::kable(t(adjdq_sp_int_table_global[5:7,]))
```

### visualize form interaction model (ordered)
```{r}
font_size <- 16
theme_set(theme_classic(base_family = "Helvetica",base_size = font_size))
line_size <- 1.5
point_size <- 2

adjdq_sp_int_plots_global <- lapply(adjdq_sp_int_models_global, function(x) {
  
  model_plot <- visualize_model_int(
  modobj = x,
  smooth_var = "ga_adj_norm_devo_quotient_zscore",
  int_var = "form_name",
  plabels = "ordered",
  derivative_plot = F) + 
    labs(x="GA-ADJ DQ Median (z-scored)",
         title = glue("{x$terms[[2]]}"),
         color = "Form Month") +
    theme(axis.title.y = element_blank()) +
    scale_y_continuous(breaks = qnorm(c(.01, .50, .99)),
                       labels = c("<=1st",  "50th", ">=99th"))
    
})

adjdq_sp_int_plots_global <- setNames(adjdq_sp_int_plots_global, global_vars_totest)

wrap_plots(adjdq_sp_int_plots_global[1:4], ncol = 2, guides = "collect")
wrap_plots(adjdq_sp_int_plots_global[5:7], ncol = 2, guides = "collect")

if(save_figs){
saveRDS(plot_output_folder, glue("{plot_output_folder}adjdq_sp_int_modelvis_plots_global.rds"))
}
```

## Compare R^2 between OG and GA-ADJ
```{r}
tables_list <- list(
  OG = ogdq_sp_table_global %>% mutate(phenotype = rownames(.)),
  GA_ADJ = adjdq_sp_table_global %>% mutate(phenotype = rownames(.))
)

color_vec <- c(OG = swyc_color, GA_ADJ = ga_adj_color)

## Main effects: Original DQ vs GA-adjusted DQ
plot_lollipop(
  table_list = tables_list,
  value_col = "r2_full_vs_none",
  color_vec = color_vec,
  p_col = "full_p_adj",
  name_mapping = namemap
) +
  labs(
    x = expression("Partial " * R^2),
    color = NULL
  )


## Ordered form interactions
plot_lollipop(
  table_list = list(
    ogdq_sp_int_table_global,
    adjdq_sp_int_table_global
  ),
  value_col = "partial_r2",
  group_names = c("Original DQ", "GA-adjusted DQ"),
  color_vec = c(swyc_color, ga_color),
  p_col = "int_p_adj",
  name_mapping = name_mapping
) +
  labs(
    x = expression("Partial " * R^2 ~ "(Interaction)"),
    color = NULL
)
```


## RESI plots for median DQ
```{r}
ogdq_resi_table_global <- extract_bam_resi(models = ogdq_sp_models_global,
                                           term = "s(median_dq)")

adjdq_resi_table_global <- extract_bam_resi(models = adjdq_sp_models_global,
                                           term = "s(ga_adj_median_dq)")
tables_list <- list(
  OG = ogdq_resi_table_global %>% mutate(phenotype = rownames(.)),
  GA_ADJ = adjdq_resi_table_global %>% mutate(phenotype = rownames(.))
)

color_vec <- c(OG = swyc_color, GA_ADJ = ga_adj_color)

## Main effects: Original DQ vs GA-adjusted DQ
plot_lollipop(
  table_list = tables_list,
  value_col = "RESI",
  color_vec = color_vec,
  p_col = "p_adj"
) +
  labs(
    x = "RESI from F-statistic",
    color = NULL
  )
```

# Regional Variables

## ggseg helpers
### ggseg label helper
```{r ggseg-label-fct, echo = TRUE}
# plot_brain_map() joins on a `label` column matching ggseg::dk/ggseg::aseg,
# but main_stats/resi_stats only carry the raw FreeSurfer `phenotype` name
# (run_spline_analysis_variant.R doesn't add a label column). Reproduces the
# per-region-type cleanup the original Rmds applied right after stat
# extraction (e.g. `ga_sp_table_vol$label <- rownames(ga_sp_table_vol) %>% ...`).
add_ggseg_label <- function(df, region_type = c("vol", "sa", "th")) {
  region_type <- match.arg(region_type)

  if (region_type == "vol") {
    cols_edit <- grepl("aseg(?!_CC)", df$phenotype, perl = TRUE)
    df$label <- df$phenotype %>%
      sub("aparc_GrayVol_", "", .) %>%
      sub("aseg_", "", .) %>%
      sub("X", "x", .) %>%
      sub("rd_Ventricle", "rd_ventricle", .) %>%
      sub("th_Ventricle", "th_ventricle", .) %>%
      sub("Thalamus", "Thalamus-Proper", .)
    df$label[cols_edit] <- gsub("_", "-", df$label[cols_edit])
  } else if (region_type == "sa") {
    df$label <- sub("aparc_SurfArea_", "", df$phenotype)
  } else {
    df$label <- sub("aparc_ThickAvg_", "", df$phenotype)
  }

  df
}
```


### helper functions for plotting
```{r}
build_ggseg_table <- function(df, group, value_col, p_col, region_type, atlas = "dk") {
  d <-  df %>%
    filter(group == !!group) %>%
    select(phenotype, val = all_of(value_col), p = all_of(p_col))

  if (region_type == "vol") {
    d <- if (atlas == "aseg") {
      d %>% filter(grepl("^aseg_", phenotype))
    } else {
      d %>% filter(grepl("^aparc_GrayVol_", phenotype))
    }
  }

  add_ggseg_label(d, region_type)
}

plot_variant_brainmap <- function(df, group, value_col, p_col, region_type,
                                   atlas = "dk", limits = NULL, title = "", fill_name = "value",
                                   low_color = "#F2F0EC", mid_color = "#D99A9A", high_color = "#D95C72") {
  
  table <- build_ggseg_table(df, group, value_col, p_col, region_type, atlas)

  plot_brain_map(
    table = table,
    value_col = "val",
    atlas = atlas,
    p_col = "p",
    p_filter = TRUE,
    p_thresh = 0.05,
    title = title,
    fill_name = fill_name,
    limits = limits,
    low_color = low_color, mid_color = mid_color, high_color = high_color
  )
}

# Legend title shown on the shared colorbar -- "value_col" is the raw column
# name (r2_full_vs_none / RESI); readers should see the stat name instead.
stat_label <- function(value_col) if (value_col == "RESI") "RESI" else "Partial R2"

# Short text panels used as the shared grid's column headers (horizontal) and
# row headers (vertical), since the brain map panels themselves carry no title.
col_header_plot <- function(label) {
  ggplot() + theme_void() +
    annotate("text", x = 0, y = 0, label = label, size = 5, fontface = "bold") +
    xlim(-1, 1) + ylim(-1, 1)
}

row_header_plot <- function(label) {
  ggplot() + theme_void() +
    annotate("text", x = 0, y = 0, label = label, angle = 90, size = 5, fontface = "bold") +
    xlim(-1, 1) + ylim(-1, 1)
}

# Columns of the shared grid: DK cortical volume, DK surface area, DK cortical
# thickness, ASEG subcortical volume.
brainmap_cols <- list(
  list(header = "Vol", region_type = "vol", atlas = "dk"),
  list(header = "SA",  region_type = "sa",  atlas = "dk"),
  list(header = "CTh", region_type = "th",  atlas = "dk"),
  list(header = "Vol", region_type = "vol", atlas = "aseg")
)
```

## Main Effects
### OG DQ fit main effects
```{r}
for(metric in c("vol", "sa", "th")){

  vars <- get(paste0(metric, "_vars"))
    
  covs <- if (metric == "th" && median_th_mpr_flag) {
    c("euler_mean_mpr", "global_MeanThickness")
  } else if (metric == "vol") {
    c("qc_var", "global_eTIV")
  } else if (metric == "sa") {
    c("qc_var", "global_SurfaceArea")
  } else if (metric == "th") {
    c("qc_var", "global_MeanThickness")
  } else {
    NULL
  }
  
  suffix <- metric

  assign(paste0("ogdq_sp_models_", suffix),
         fit_bam_models(df.zscore_avg, 
                        vars,
                        predictor = "median_dq",
                        covariates = covs, save = FALSE))

  assign(paste0("ogdq_sp_table_", suffix),
         extract_bam_features_main(
           models = get(paste0("ogdq_sp_models_", suffix)),
           data = df.zscore_avg,
           predictor = "median_dq",
           covariates = covs
         ))
  
    assign(paste0("ogdq_resi_table_", suffix),
         extract_bam_resi(
           models = get(paste0("ogdq_sp_models_", suffix)),
           term = "s(median_dq)"
         ))
}

ogdq_sp_tables_regional <- rbind(ogdq_sp_table_vol %>% mutate(group = "vol"),
                                 ogdq_sp_table_sa %>% mutate(group = "sa"),
                                 ogdq_sp_table_th %>% mutate(group = "th")) %>%
  mutate(phenotype = rownames(.),
         signed_partial_r2 = r2_full_vs_none*sign(avg_first_deriv))


```

### GGSeg plot with og-dq, signed partial r2
```{r}
df = ogdq_sp_tables_regional %>% mutate(signed_partial_r2 = r2_full_vs_none*sign(avg_first_deriv))
value_col = "signed_partial_r2"
p_col = "full_p_adj"
range(df %>% filter(group %in% c("vol", "sa", "th")) %>% pull(value_col))
signed_limits = c(-0.12, 0.12)

fill_name <- "Signed Partial R2"

  panels <- c(
    list(plot_spacer()),
    lapply(brainmap_cols, function(col) col_header_plot(col$header))
  )
  panels <- c(panels, list(row_header_plot("SWYC Median")))
  panels <- c(panels, lapply(brainmap_cols, function(col) {
    plot_variant_brainmap(df, col$region_type, value_col, p_col,
                             col$region_type, atlas = col$atlas, limits = signed_limits, fill_name = fill_name,
                            low_color = "#4775AD", mid_color = "white", high_color = "#B84A58")
  }))

  
ogdq_signedR2_regions_plot <- wrap_plots(panels, ncol = 5, nrow = 2,
             widths = c(0.15, 1, 1, 1, 1), heights = c(0.15, 1),
             guides = "collect") &
    theme(legend.position = "bottom")
  
ggsave(glue("{plot_output_folder}ggseg_ogdq_median_signedr2.svg"), ogdq_signedR2_regions_plot, width = 16, height = 6)

```

### GA-ADJ DQ fit main effects
```{r}
for(metric in c("vol", "sa", "th")){

  vars <- get(paste0(metric, "_vars"))
    
  covs <- if (metric == "th" && median_th_mpr_flag) {
    c("euler_mean_mpr")
  } else if (metric == "vol") {
    c("qc_var")
  } else if (metric == "sa") {
    c("qc_var")
  } else if (metric == "th") {
    c("qc_var")
  } else {
    NULL
  }
  
  suffix <- metric

  assign(paste0("adjdq_sp_models_", suffix),
         fit_bam_models(df.zscore_avg, 
                        vars,
                        predictor = "ga_adj_median_dq",
                        covariates = covs, save = FALSE))

  assign(paste0("adjdq_sp_table_", suffix),
         extract_bam_features_main(
           models = get(paste0("adjdq_sp_models_", suffix)),
           data = df.zscore_avg,
           predictor = "ga_adj_median_dq",
           covariates = covs
         ))
  
   assign(paste0("adjdq_resi_table_", suffix),
         extract_bam_resi(
           models = get(paste0("adjdq_sp_models_", suffix)),
           term = "s(ga_adj_median_dq)"
         ))
}

# Add label column that matches ggseg for plotting

## volume
cols_edit <- grepl("aseg(?!_CC)", rownames(adjdq_sp_table_vol), perl = TRUE)
adjdq_sp_table_vol$label <- rownames(adjdq_sp_table_vol) %>% 
  sub("aparc_GrayVol_", "", .) %>% 
  sub("aseg_", "", .) %>%
  sub("X", "x", .) %>%
  sub("rd_Ventricle", "rd_ventricle", .) %>%     sub("th_Ventricle", "th_ventricle", .) %>% sub("Thalamus", "Thalamus-Proper", .)
adjdq_sp_table_vol$label[cols_edit] <- gsub("_", "-", adjdq_sp_table_vol$label[cols_edit])

## surface area
adjdq_sp_table_sa$label <- rownames(adjdq_sp_table_sa) %>% 
  sub("aparc_SurfArea_", "", .)

## thickness
adjdq_sp_table_th$label <- rownames(adjdq_sp_table_th) %>% 
  sub("aparc_ThickAvg_", "", .)


## volume
cols_edit <- grepl("aseg(?!_CC)", rownames(adjdq_resi_table_vol), perl = TRUE)
adjdq_resi_table_vol$label <- rownames(adjdq_resi_table_vol) %>% 
  sub("aparc_GrayVol_", "", .) %>% 
  sub("aseg_", "", .) %>%
  sub("X", "x", .) %>%
  sub("rd_Ventricle", "rd_ventricle", .) %>%     sub("th_Ventricle", "th_ventricle", .) %>% sub("Thalamus", "Thalamus-Proper", .)
adjdq_resi_table_vol$label[cols_edit] <- gsub("_", "-", adjdq_resi_table_vol$label[cols_edit])

## surface area
adjdq_resi_table_sa$label <- rownames(adjdq_resi_table_sa) %>% 
  sub("aparc_SurfArea_", "", .)

## thickness
adjdq_resi_table_th$label <- rownames(adjdq_resi_table_th) %>% 
  sub("aparc_ThickAvg_", "", .)
```


## Interaction with Form Type
### OG DQ fit interactions
```{r}
for(metric in c("vol", "sa", "th")){

  vars <- get(paste0(metric, "_vars"))

  covs <- if (metric == "th" && median_th_mpr_flag) {
    c("euler_mean_mpr")
  } else if (metric == "vol") {
    c("qc_var", "global_eTIV")
  } else if (metric == "sa") {
    c("qc_var")
  } else if (metric == "th") {
    c("qc_var")
  } else {
    NULL
  }
  suffix <- metric

  assign(paste0("ogdq_sp_int_models_", suffix),
         fit_bam_interactions(
           cent_all %>% filter(form_name %in% forms_totest),
           phenotypes = vars,
           predictor = "norm_devo_quotient_zscore",
           moderator = "form_name",
           interaction = "ordered",
           covariates = covs, 
           save = FALSE
         ))

  assign(paste0("ogdq_sp_int_table_", suffix),
         extract_bam_features_interaction(
           models = get(paste0("ogdq_sp_int_models_", suffix)),
           data = cent_all %>% filter(form_name %in% forms_totest),
           predictor = "norm_devo_quotient_zscore",
           moderator = "form_name",
           covariates = covs,
           interaction_type = "ordered"
         ))
  
  assign(paste0("ogdq_int_lrt_table_", suffix),
         test_interaction_lrt(data = cent_all %>% filter(form_name %in% forms_totest), 
                                phenotypes = vars,
                                predictor = "norm_devo_quotient_zscore",
                                moderator = "form_name",
                                interaction = "ordered",
                                covariates = covs, 
                                save = FALSE,
                              method = "ML"
         ))
}

for(suffix in c("vol", "sa", "th")){
    assign(paste0("ogdq_int_full_table_", suffix),
           merge(get(paste0("ogdq_sp_int_table_", suffix)) %>%
                   mutate(phenotype = rownames(.)),
                 get(paste0("ogdq_int_lrt_table_", suffix)) %>%
                   rename(lrt_p_int = p, 
                          lrt_p_int_adj = p_adj),
                 by = "phenotype")
    )
}

# Add label column that matches ggseg for plotting

## volume
cols_edit <- grepl("aseg(?!_CC)", ogdq_int_full_table_vol$phenotype, perl = TRUE)
ogdq_int_full_table_vol$label <- ogdq_int_full_table_vol$phenotype %>% 
  sub("aparc_GrayVol_", "", .) %>% 
  sub("aseg_", "", .) %>%
  sub("X", "x", .) %>%
  sub("rd_Ventricle", "rd_ventricle", .) %>%     sub("th_Ventricle", "th_ventricle", .) %>% sub("Thalamus", "Thalamus-Proper", .)
ogdq_int_full_table_vol$label[cols_edit] <- gsub("_", "-", ogdq_int_full_table_vol$label[cols_edit])

## surface area
ogdq_int_full_table_sa$label <- ogdq_int_full_table_sa$phenotype %>% 
  sub("aparc_SurfArea_", "", .)

## thickness
ogdq_int_full_table_th$label <- ogdq_int_full_table_th$phenotype %>% 
  sub("aparc_ThickAvg_", "", .)

```

### OG DQ visualize interactions ggseg - partial R2
using p-value from the ML-based LRT comparison table
```{r}
(ogdq_int_partialr2_dkt_vol <- plot_brain_map(
  table = ogdq_int_full_table_vol,
  value_col = "partial_r2",
  atlas = "dk",
  p_col = "lrt_p_int_adj",
  p_filter = TRUE,
  title = "OG DQ partial R2 vol",
  limits = c(0, 0.015)
))

(ogdq_int_partialr2_aseg_vol <- plot_brain_map(
  table = ogdq_int_full_table_vol,
  value_col = "partial_r2",
  atlas = "aseg",
  p_col = "lrt_p_int_adj",
  p_filter = TRUE,
  title = "OG DQ partial R2 vol",
  limits = c(0, 0.015)
))

(ogdq_int_partialr2_sa <- plot_brain_map(
  table = ogdq_int_full_table_sa,
  value_col = "partial_r2",
  atlas = "dk",
  p_col = "lrt_p_int_adj",
  p_filter = TRUE,
  title = "OG DQ partial R2 sa",
  limits = c(0, 0.015)
))

(ogdq_int_partialr2_th <- plot_brain_map(
  table = ogdq_int_full_table_th,
  value_col = "partial_r2",
  atlas = "dk",
  p_col = "lrt_p_int_adj",
  p_filter = TRUE,
  title = "OG DQ partial R2 th",
  limits = c(0, 0.015)
))
```

### GA-ADJ DQ fit interactions
```{r}
for(metric in c("vol", "sa", "th")){

  vars <- get(paste0(metric, "_vars"))

  covs <- if (metric == "th" && median_th_mpr_flag) {
    c("euler_mean_mpr")
  } else if (metric == "vol") {
    c("qc_var", "global_eTIV")
  } else if (metric == "sa") {
    c("qc_var")
  } else if (metric == "th") {
    c("qc_var")
  } else {
    NULL
  }
  suffix <- metric

  assign(paste0("adjdq_sp_int_models_", suffix),
         fit_bam_interactions(
           cent_all %>% filter(form_name %in% forms_totest),
           phenotypes = vars,
           predictor = "norm_devo_quotient_zscore",
           moderator = "form_name",
           interaction = "ordered",
           covariates = covs, 
           save = FALSE
         ))

  assign(paste0("adjdq_sp_int_table_", suffix),
         extract_bam_features_interaction(
           models = get(paste0("adjdq_sp_int_models_", suffix)),
           data = cent_all %>% filter(form_name %in% forms_totest),
           predictor = "norm_devo_quotient_zscore",
           moderator = "form_name",
           covariates = covs,
           interaction_type = "ordered"
         ))
  
  assign(paste0("adjdq_int_lrt_table_", suffix),
         test_interaction_lrt(data = cent_all %>% filter(form_name %in% forms_totest), 
                                phenotypes = vars,
                                predictor = "norm_devo_quotient_zscore",
                                moderator = "form_name",
                                interaction = "ordered",
                                covariates = covs, 
                                save = FALSE,
                              method = "ML"
         ))
}

for(suffix in c("vol", "sa", "th")){
    assign(paste0("adjdq_int_full_table_", suffix),
           merge(get(paste0("adjdq_sp_int_table_", suffix)) %>%
                   mutate(phenotype = rownames(.)),
                 get(paste0("adjdq_int_lrt_table_", suffix)) %>%
                   rename(lrt_p_int = p, 
                          lrt_p_int_adj = p_adj),
                 by = "phenotype")
    )
}

# Add label column that matches ggseg for plotting

## volume
cols_edit <- grepl("aseg(?!_CC)", adjdq_int_full_table_vol$phenotype, perl = TRUE)
adjdq_int_full_table_vol$label <- adjdq_int_full_table_vol$phenotype %>% 
  sub("aparc_GrayVol_", "", .) %>% 
  sub("aseg_", "", .) %>%
  sub("X", "x", .) %>%
  sub("rd_Ventricle", "rd_ventricle", .) %>%     sub("th_Ventricle", "th_ventricle", .) %>% sub("Thalamus", "Thalamus-Proper", .)
adjdq_int_full_table_vol$label[cols_edit] <- gsub("_", "-", adjdq_int_full_table_vol$label[cols_edit])

## surface area
adjdq_int_full_table_sa$label <- adjdq_int_full_table_sa$phenotype %>% 
  sub("aparc_SurfArea_", "", .)

## thickness
adjdq_int_full_table_th$label <- adjdq_int_full_table_th$phenotype %>% 
  sub("aparc_ThickAvg_", "", .)
```

### GA-ADJ DQ visualize interactions ggseg - partial R2
using p-value from the ML-based LRT comparison table
```{r}
(adjdq_int_partialr2_dkt_vol <- plot_brain_map(
  table = adjdq_int_full_table_vol,
  value_col = "partial_r2",
  atlas = "dk",
  p_col = "lrt_p_int_adj",
  p_filter = TRUE,
  title = "GA-ADJ DQ partial R2 vol",
  limits = c(0, 0.015)
))

(adjdq_int_partialr2_aseg_vol <- plot_brain_map(
  table = adjdq_int_full_table_vol,
  value_col = "partial_r2",
  atlas = "aseg",
  p_col = "lrt_p_int_adj",
  p_filter = TRUE,
  title = "GA-ADJ DQ partial R2 vol",
  limits = c(0, 0.015)
))

(adjdq_int_partialr2_sa <- plot_brain_map(
  table = adjdq_int_full_table_sa,
  value_col = "partial_r2",
  atlas = "dk",
  p_col = "lrt_p_int_adj",
  p_filter = TRUE,
  title = "GA-ADJ DQ partial R2 sa",
  limits = c(0, 0.015)
))

(adjdq_int_partialr2_th <- plot_brain_map(
  table = adjdq_int_full_table_th,
  value_col = "partial_r2",
  atlas = "dk",
  p_col = "lrt_p_int_adj",
  p_filter = TRUE,
  title = "GA-ADJ DQ partial R2 th",
  limits = c(0, 0.015)
))
```


# Relationships with Birth Factors (in this subsample)

## Median SWYC Scores
```{r}
og_cor_mat <- cor(df.zscore_avg[,c("median_dq", "ga_adj_median_dq", "gestational_age", "birth_weight_kg", "bwp_fen")], use = "pairwise.complete.obs")

og_p_mat <- cor_pmat(df.zscore_avg[,c("median_dq", "ga_adj_median_dq", "gestational_age", "birth_weight_kg", "bwp_fen")], use = "pairwise.complete.obs")

(swyc_ogDQ_bf_corrs_plot <- ggcorrplot(
  og_cor_mat,
  p.mat = og_p_mat,
  type = "full",
  insig = "blank",
  lab = TRUE
))
```
## Form-Based
### Correlations Grouped (not matrix)
```{r}
bf_vars <- c(
  "gestational_age_zscore",
  "birth_weight_kg_zscore",
  "bwp_fen"
)

ogDQ_bf_cor_results <- cent_all %>%
  filter(form_name %in% forms_totest) %>%
  pivot_longer(
    cols = all_of(bf_vars),
    names_to = "bf_type",
    values_to = "bf_score"
  ) %>%
  group_by(form_name, bf_type) %>%
  group_modify(~{
    ct <- cor.test(
      .x$norm_devo_quotient_zscore,
      .x$bf_score,
      method = "pearson"
    )
    tibble(
      n = nrow(.x),
      r = unname(ct$estimate),
      lower = ct$conf.int[1],
      upper = ct$conf.int[2],
      p = ct$p.value
    )
  }) %>%
  ungroup()

(swyc_ogDQ_forms_bf_corrs_plot <- ggplot(ogDQ_bf_cor_results,
       aes(x = r,
           y = form_name,
           colour = bf_type)) +
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "grey60") +
  geom_errorbarh(
    aes(xmin = lower, xmax = upper),
    height = 0.2
  ) +
  geom_point(size = 3) +
  facet_wrap(~bf_type) +
  scale_x_continuous(
    limits = c(-1, 1),
    breaks = seq(-1, 1, 0.25)
  ) +
  labs(
    x = "Pearson correlation (r)",
    y = NULL,
    title = "OG DQ",
    colour = "bf measure"
  ) +
  theme_classic())

ga_adjDQ_bf_cor_results <- cent_all %>%
  filter(form_name %in% forms_totest) %>%
  pivot_longer(
    cols = all_of(bf_vars),
    names_to = "bf_type",
    values_to = "bf_score"
  ) %>%
  group_by(form_name, bf_type) %>%
  group_modify(~{
    ct <- cor.test(
      .x$ga_adj_norm_devo_quotient,
      .x$bf_score,
      method = "pearson"
    )
    tibble(
      n = nrow(.x),
      r = unname(ct$estimate),
      lower = ct$conf.int[1],
      upper = ct$conf.int[2],
      p = ct$p.value
    )
  }) %>%
  ungroup()

(swyc_ga_adjDQ_forms_bf_corrs_plot <- ggplot(ga_adjDQ_bf_cor_results,
       aes(x = r,
           y = form_name,
           colour = bf_type)) +
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "grey60") +
  geom_errorbarh(
    aes(xmin = lower, xmax = upper),
    height = 0.2
  ) +
  geom_point(size = 3) +
  facet_wrap(~bf_type) +
  scale_x_continuous(
    limits = c(-1, 1),
    breaks = seq(-1, 1, 0.25)
  ) +
  labs(
    x = "Pearson correlation (r)",
    y = NULL,
    title = "GA-ADJ DQ",
    colour = "bf measure"
  ) +
  theme_classic())

if(save_figs == TRUE){
  panel_save(swyc_ogDQ_forms_bf_corrs_plot, "swyc_ogDQ_bf_cors", "svg", plot_output_folder)
  panel_save(swyc_ga_adjDQ_forms_bf_corrs_plot, "swyc_ga_adjDQ_bf_cors", "svg", plot_output_folder)
}
```

## Get brain vars corr with birth factors
```{r}
var_sets <- list(
  cmd    = cmd_vars,
  global = global_vars_totest,
  vol    = vol_vars,
  sa     = sa_vars,
  th     = th_vars
)

outcomes <- c("gestational_age", "bwp_fen")

results <- lapply(names(var_sets), function(group) {
  
  vars <- var_sets[[group]]
  
  do.call(rbind, lapply(outcomes, function(outcome) {
    
    do.call(rbind, lapply(vars, function(var) {
      
      x <- df.zscore_avg[[var]]
      y <- df.zscore_avg[[outcome]]
      
      ok <- complete.cases(x, y)
      
      test <- cor.test(x[ok], y[ok], method = "pearson")
      
      data.frame(
        group = group,
        outcome = outcome,
        variable = var,
        r = unname(test$estimate),
        p = test$p.value,
        n = sum(ok)
      )
    }))
  }))
})

results <- do.call(rbind, results)

# BH correction separately within each variable group AND outcome
results$p_adj <- ave(
  results$p,
  results$group,
  results$outcome,
  FUN = function(x) p.adjust(x, method = "BH")
)

# Significant correlations
sig_results <- results[results$p_adj < 0.05, ]
```

# Mediation (covar: partial r2 and RESI)

## Models for covars function
```{r}
fit_models_covars <- function(outcome,
                              predictor,
                              covar){
  
  if(is.null(covar)){
    model <- fit_bam_models(
      data = df.zscore_avg,
      "median_dq",
      predictor = "gestational_age",
      save = FALSE
    )
    
    label <- "No covariate"
    
  } else {
    model <- fit_bam_models(
      data = df.zscore_avg,
      phenotypes = outcome,
      predictor = predictor,
      covar = covar,
      save = FALSE
    )
    
    label <- covar
  }
  
  table <- extract_bam_features_main(
    models = model,
    data = df.zscore_avg,
    predictor = predictor,
    covar = covar
  )
  
  table_resi <- extract_bam_resi(
    models = model,
    term = glue("s({predictor})")
  )
  
  # if extraction works, use it
  data.frame(
    covar = label,
    partial_r2 = table$r2_full_vs_none,
    p_adj = table$full_p_adj,
    avg_first_deriv = table$avg_first_deriv,
    resi = table_resi$RESI
  )
}
```

## NULL model values
```{r}
# manually fit no-covariate model (because of extract_bam_features)
ga_model_null <- fit_bam_models(
  data = df.zscore_avg,
  "median_dq",
  predictor = "gestational_age",
  save = FALSE
)

# extract R2 from bam directly
ga_null_r2 <- summary(ga_model_null[[1]]$full)$r.sq

ga_null_table <- extract_bam_table(ga_model_null,
                                   "gestational_age")

ga_null_resi <- extract_bam_resi(
  ga_model_null,
  term = "s(gestational_age)"
)
```

## Get shared brain vars
```{r}
cmd_vars_shared <- intersect(
  sig_results %>% filter(group == "cmd" & outcome == "gestational_age" & p_adj < 0.05) %>%
  pull(variable), 
  rownames(ogdq_sp_table_cmd %>% filter(full_p_adj < 0.05)))

global_vars_shared <- intersect(
  sig_results %>% filter(group == "global" & outcome == "gestational_age" & p_adj < 0.05) %>%
  pull(variable), 
  rownames(ogdq_sp_table_global %>% filter(full_p_adj < 0.05)))

vol_vars_shared <- intersect(
  sig_results %>% filter(group == "vol" & outcome == "gestational_age" & p_adj < 0.05) %>%
  pull(variable), 
  rownames(ogdq_sp_table_vol %>% filter(full_p_adj < 0.05)))

sa_vars_shared <- intersect(
  sig_results %>% filter(group == "sa" & outcome == "gestational_age" & p_adj < 0.05) %>%
  pull(variable), 
  rownames(ogdq_sp_table_sa %>% filter(full_p_adj < 0.05)))

th_vars_shared <- intersect(
  sig_results %>% filter(group == "th" & outcome == "gestational_age" & p_adj < 0.05) %>%
  pull(variable), 
  rownames(ogdq_sp_table_th %>% filter(full_p_adj < 0.05)))
```

## Shared CMD and Global Covars

### Calculate partial r2 and RESI for different covariates
```{r}
# fit models
global_covars <- c(NULL, cmd_vars_shared, global_vars_shared)

# run covariate models separately
ga_global_covars_list <- lapply(
  global_covars[-1],
  function(.x) {
    fit_models_covars(
      outcome = "median_dq",
      predictor = "gestational_age",
      covar = .x
    )
  }
)

ga_global_covars_table <- rbind(
  data.frame(
    covar = "No covariate",
    partial_r2 = ga_null_r2,
    p_adj = ga_null_table$full_p_adj,
    avg_first_deriv = NA,
    resi = ga_null_resi$RESI
),
  do.call(rbind, ga_global_covars_list)
)

knitr::kable(ga_global_covars_table)
```

### plot them
```{r}
## RESI of GA with different brain covar
resi_covar_plot <- plot_lollipop(
  table_list = ga_covars_table, 
         value_col = "resi",
         label_col = "covar") +
  labs(
    x = "RESI",
    y = "Covariate"
  ) +
  theme(
    axis.title = element_text(size = 16)
  )

resi_covar_plot + 
  geom_vline(
    xintercept = ga_covars_table %>% 
      filter(covar == "No covariate") %>% 
      pull("resi"),
    linetype = "dashed",
    color = "#B86F6F",
    linewidth = 0.8
  )

## partial R2 of GA with different brain covar
partialr2_covar_plot <- plot_lollipop(
  table_list = ga_global_covars_table, 
         value_col = "partial_r2",
         label_col = "covar") +
  labs(
    x = expression("Partial " * R^2),
    y = "Covariate"
  )  +
  theme(
    axis.title = element_text(size = 16)
  )

partialr2_covar_plot + 
  geom_vline(
    xintercept = ga_global_covars_table %>% 
      filter(covar == "No covariate") %>% 
      pull("partial_r2"),
    linetype = "dashed",
    color = "#B86F6F",
    linewidth = 0.8
  )
```

## Shared Vol Covars

### Calculate partial r2 and RESI for different covariates
```{r}
# fit models
vol_covars <- c(NULL, vol_vars_shared)

# run covariate models separately
ga_vol_covars_list <- lapply(
  vol_covars[-1],
  function(.x) {
    fit_models_covars(
      outcome = "median_dq",
      predictor = "gestational_age",
      covar = .x
    )
  }
)

ga_vol_covars_table <- rbind(
  data.frame(
    covar = "No covariate",
    partial_r2 = ga_null_r2,
    p_adj = ga_null_table$full_p_adj,
    avg_first_deriv = NA,
    resi = ga_null_resi$RESI
),
  do.call(rbind, ga_vol_covars_list)
)

knitr::kable(ga_vol_covars_table)
```

### plot them
```{r}
## RESI of GA with different brain covar
resi_covar_plot <- plot_lollipop(
  table_list = ga_vol_covars_table, 
         value_col = "resi",
         label_col = "covar") +
  labs(
    x = "RESI",
    y = "Covariate"
  ) +
  theme(
    axis.title = element_text(size = 16)
  )

resi_covar_plot + 
  geom_vline(
    xintercept = ga_vol_covars_table %>% 
      filter(covar == "No covariate") %>% 
      pull("resi"),
    linetype = "dashed",
    color = "#B86F6F",
    linewidth = 0.8
  )

## partial R2 of GA with different brain covar
partialr2_covar_plot <- plot_lollipop(
  table_list = ga_vol_covars_table, 
         value_col = "partial_r2",
         label_col = "covar") +
  labs(
    x = expression("Partial " * R^2),
    y = "Covariate"
  )  +
  theme(
    axis.title = element_text(size = 16)
  )

partialr2_covar_plot + 
  geom_vline(
    xintercept = ga_vol_covars_table %>% 
      filter(covar == "No covariate") %>% 
      pull("partial_r2"),
    linetype = "dashed",
    color = "#B86F6F",
    linewidth = 0.8
  )
```



## Shared SA Covars

### Calculate partial r2 and RESI for different covariates
```{r}
# fit models
sa_covars <- c(NULL, sa_vars_shared)

# run covariate models separately
ga_sa_covars_list <- lapply(
  sa_covars[-1],
  function(.x) {
    fit_models_covars(
      outcome = "median_dq",
      predictor = "gestational_age",
      covar = .x
    )
  }
)

ga_sa_covars_table <- rbind(
  data.frame(
    covar = "No covariate",
    partial_r2 = ga_null_r2,
    p_adj = ga_null_table$full_p_adj,
    avg_first_deriv = NA,
    resi = ga_null_resi$RESI
),
  do.call(rbind, ga_sa_covars_list)
)

knitr::kable(ga_sa_covars_table)
```

### plot them
```{r}
## RESI of GA with different brain covar
resi_covar_plot <- plot_lollipop(
  table_list = ga_sa_covars_table, 
         value_col = "resi",
         label_col = "covar") +
  labs(
    x = "RESI",
    y = "Covariate"
  ) +
  theme(
    axis.title = element_text(size = 16)
  )

resi_covar_plot + 
  geom_vline(
    xintercept = ga_sa_covars_table %>% 
      filter(covar == "No covariate") %>% 
      pull("resi"),
    linetype = "dashed",
    color = "#B86F6F",
    linewidth = 0.8
  )

## partial R2 of GA with different brain covar
partialr2_covar_plot <- plot_lollipop(
  table_list = ga_sa_covars_table, 
         value_col = "partial_r2",
         label_col = "covar") +
  labs(
    x = expression("Partial " * R^2),
    y = "Covariate"
  )  +
  theme(
    axis.title = element_text(size = 16)
  )

partialr2_covar_plot + 
  geom_vline(
    xintercept = ga_sa_covars_table %>% 
      filter(covar == "No covariate") %>% 
      pull("partial_r2"),
    linetype = "dashed",
    color = "#B86F6F",
    linewidth = 0.8
  )
```

# Save all models and stats tables
```{r}
ogdq_stats <- mget(ls(pattern = "^ogdq_sp(_int)?_table"))
adjdq_stats <- mget(ls(pattern = "^adjdq_sp(_int)?_table"))

# Edit this to save relevant tables/models here

# save(ga_stats,
#      file = file.path(sp_stat_folder,
#                       "ga_nocovar_all_stats.RData"))
# save(bw_stats,
#      file = file.path(sp_stat_folder,
#                       "bw_nocovar_all_stats.RData"))
# save(bwp_stats,
#      file = file.path(sp_stat_folder,
#                       "bwp_nocovar_all_stats.RData"))
# 
# ga_models <- mget(ls(pattern = "^ga_sp(_int)?_models"))
# bw_models <- mget(ls(pattern = "^bw_sp(_int)?_models"))
# bwp_models <- mget(ls(pattern = "^bwp_sp(_int)?_models"))
# 
# save(ga_models,
#      file = file.path(models_folder,
#                       "ga_nocovar_all_models.RData"))
# save(bw_models,
#      file = file.path(models_folder,
#                       "bw_nocovar_all_models.RData"))
# save(bwp_models,
#      file = file.path(models_folder,
#                       "bwp_nocovar_all_models.RData"))save(ga_stats,
#      file = file.path(sp_stat_folder,
#                       "ga_nocovar_all_stats.RData"))
# save(bw_stats,
#      file = file.path(sp_stat_folder,
#                       "bw_nocovar_all_stats.RData"))
# save(bwp_stats,
#      file = file.path(sp_stat_folder,
#                       "bwp_nocovar_all_stats.RData"))
# 
# ga_models <- mget(ls(pattern = "^ga_sp(_int)?_models"))
# bw_models <- mget(ls(pattern = "^bw_sp(_int)?_models"))
# bwp_models <- mget(ls(pattern = "^bwp_sp(_int)?_models"))
# 
# save(ga_models,
#      file = file.path(models_folder,
#                       "ga_nocovar_all_models.RData"))
# save(bw_models,
#      file = file.path(models_folder,
#                       "bw_nocovar_all_models.RData"))
# save(bwp_models,
#      file = file.path(models_folder,
#                       "bwp_nocovar_all_models.RData"))
```


# Spin test export
```{r spin-export-fct, echo = TRUE}
save_stat_for_spin <- function(df, group, value_col, p_col,
                                region_type, atlas = "dk", output_folder) {

  table <- build_ggseg_table(df,  group, value_col, p_col, region_type, atlas)

  table$hemisphere <- dplyr::case_when(
    grepl("^lh_", table$label) ~ "L",
    grepl("^rh_", table$label) ~ "R",
    TRUE ~ NA_character_
  )
  table$label <- sub("^lh_|^rh_", "", table$label)

  names(table)[names(table) == "val"] <- value_col
  names(table)[names(table) == "p"]   <- p_col

  filename <- glue("{output_folder}swyc_{group}_{value_col}_spin.csv")
  write.csv(table, filename, row.names = FALSE)
  invisible(table)
}
```

```{r spin-export-run, echo = TRUE}
save_stat_for_spin(ogdq_sp_tables_regional, "vol", "r2_full_vs_none", "full_p_adj",
                    region_type = "vol", atlas = "dk", output_folder = spin_folder)

save_stat_for_spin(ogdq_sp_tables_regional, "sa", "r2_full_vs_none", "full_p_adj",
                    region_type = "sa", atlas = "dk", output_folder = spin_folder)

save_stat_for_spin(ogdq_sp_tables_regional, "th", "r2_full_vs_none", "full_p_adj",
                    region_type = "th", atlas = "dk", output_folder = spin_folder)

save_stat_for_spin(ogdq_sp_tables_regional, "vol", "signed_partial_r2", "full_p_adj",
                    region_type = "vol", atlas = "dk", output_folder = spin_folder)

save_stat_for_spin(ogdq_sp_tables_regional, "sa", "signed_partial_r2", "full_p_adj",
                    region_type = "sa", atlas = "dk", output_folder = spin_folder)

save_stat_for_spin(ogdq_sp_tables_regional, "th", "signed_partial_r2", "full_p_adj",
                    region_type = "th", atlas = "dk", output_folder = spin_folder)
```

# Test Mediation

## Get all variable names 

## Brain covars with BW
```{r}
lm_swyc_bw_globalcmd <- lm(median_dq ~ birth_weight_kg + global.cmd, data = swyc_avg_centiles)
lm_swyc_bw_thcmd <- lm(median_dq ~ birth_weight_kg + th.cmd, data = swyc_avg_centiles)
lm_swyc_bw_subcort <- lm(median_dq ~ birth_weight_kg + global_SubCortGrayVol, data = swyc_avg_centiles)
lm_swyc_bw_vent <- lm(median_dq ~ birth_weight_kg + global_VentricleChoroidVol, data = swyc_avg_centiles)
lm_swyc_bw_allcmd <- lm(median_dq ~ birth_weight_kg + global.cmd + vol.cmd + sa.cmd + th.cmd, data = swyc_avg_centiles)

lm_swyc_bw_globalcmd_euler <- lm(median_dq ~ birth_weight_kg + global.cmd + median_euler_mean, data = swyc_avg_centiles)
lm_swyc_bw_thcmd_euler <- lm(median_dq ~ birth_weight_kg + th.cmd + euler_mean_mpr, data = swyc_avg_centiles)
lm_swyc_bw_subcort_euler <- lm(median_dq ~ birth_weight_kg + global_SubCortGrayVol + median_euler_mean, data = swyc_avg_centiles)
lm_swyc_bw_vent_euler <- lm(median_dq ~ birth_weight_kg + global_VentricleChoroidVol + median_euler_mean, data = swyc_avg_centiles)
lm_swyc_bw_allcmd_euler <- lm(median_dq ~ birth_weight_kg + + global.cmd + vol.cmd + sa.cmd + th.cmd + median_euler_mean, data = swyc_avg_centiles)
lm_swyc_bw_euler <- lm(median_dq ~ birth_weight_kg + median_euler_mean, data = swyc_avg_centiles)

#plot
mediation_swycbw <- data.frame(
  model = c("no covar", "global.cmd", "th.cmd", "subcort.vol", "vent.vol", "all.cmd", "global.cmd-euler", "th.cmd-euler", "subcort.vol-euler", "vent.vol-euler", "all.cmd-euler", "only euler"),
  estimate = c(summary(lm_swyc_bw)$coefficients[2, "Estimate"],
               summary(lm_swyc_bw_globalcmd)$coefficients[2, "Estimate"],
               summary(lm_swyc_bw_thcmd)$coefficients[2, "Estimate"],
               summary(lm_swyc_bw_subcort)$coefficients[2, "Estimate"],
               summary(lm_swyc_bw_vent)$coefficients[2, "Estimate"],
               summary(lm_swyc_bw_allcmd)$coefficients[2, "Estimate"],
               summary(lm_swyc_bw_globalcmd_euler)$coefficients[2, "Estimate"],
               summary(lm_swyc_bw_thcmd_euler)$coefficients[2, "Estimate"],
               summary(lm_swyc_bw_subcort_euler)$coefficients[2, "Estimate"],
               summary(lm_swyc_bw_vent_euler)$coefficients[2, "Estimate"],
               summary(lm_swyc_bw_allcmd_euler)$coefficients[2, "Estimate"],
               summary(lm_swyc_bw_euler)$coefficients[2, "Estimate"]),
  se = c(summary(lm_swyc_bw)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bw_globalcmd)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bw_thcmd)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bw_subcort)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bw_vent)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bw_allcmd)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bw_globalcmd_euler)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bw_thcmd_euler)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bw_subcort_euler)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bw_vent_euler)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bw_allcmd_euler)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bw_euler)$coefficients[2, "Std. Error"])
) %>%
  mutate(
    ymin = estimate - se,
    ymax = estimate + se
  )

beta_swyc_bw_nocov = mediation_swycbw[mediation_swycbw$model == "no covar", "estimate"]
sehigh_swyc_bw_nocov = mediation_swycbw[mediation_swycbw$model == "no covar", "ymax"]
selow_swyc_bw_nocov = mediation_swycbw[mediation_swycbw$model == "no covar", "ymin"]

ggplot(mediation_swycbw, aes(x = model, y = estimate)) +
  geom_pointrange(aes(ymin = ymin, ymax = ymax), linewidth = 0.8) +
  geom_hline(yintercept = beta_swyc_bw_nocov, linetype = "dashed", color = "#7A7A7A", linewidth = 0.5) +
  geom_hline(yintercept = sehigh_swyc_bw_nocov, linetype = "dashed", color = "#BDBDBD", linewidth = 0.5) +
  geom_hline(yintercept = selow_swyc_bw_nocov, linetype = "dashed", color = "#BDBDBD", linewidth = 0.5) +
  scale_x_discrete(limits = c("no covar", "global.cmd", "th.cmd", "subcort.vol", "vent.vol", "all.cmd", "global.cmd-euler", "th.cmd-euler", "subcort.vol-euler", "vent.vol-euler", "all.cmd-euler", "only euler")) +
  labs(y = "Beta Estimate with Std Error", x = "LM Model covariates", title = "SWYC-DQ ~ Birth Weight") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
```

## Brain covars with BWp
```{r}
lm_swyc_bwp_subcort_euler <- lm(median_dq_zscore ~ bwp_fen +  global_SubCortGrayVol + median_euler_mean, data = swyc_avg_centiles)
lm_swyc_bwp_vent_euler <- lm(median_dq_zscore ~ bwp_fen + global_VentricleChoroidVol + median_euler_mean, data = swyc_avg_centiles)
lm_swyc_bwp_cortex_euler <- lm(median_dq_zscore ~ bwp_fen + global_CortexVol + median_euler_mean, data = swyc_avg_centiles)
lm_swyc_bwp_cbWM_euler <- lm(median_dq_zscore ~ bwp_fen + global_CerebralWhiteMatterVol + median_euler_mean, data = swyc_avg_centiles)

lm_swyc_bwp_allvol_euler <- lm(median_dq_zscore ~ bwp_fen + global_CortexVol + global_CerebralWhiteMatterVol + global_SubCortGrayVol + global_VentricleChoroidVol + median_euler_mean, data = swyc_avg_centiles)

lm_swyc_bwp_allcmd_euler <- lm(median_dq_zscore ~ bwp_fen + + global.cmd + vol.cmd + sa.cmd + th.cmd + subcort.cmd + median_euler_mean, data = swyc_avg_centiles)
lm_swyc_bwp_euler <- lm(median_dq_zscore ~ bwp_fen + median_euler_mean, data = swyc_avg_centiles)

#plot
mediation_swycbwp <- data.frame(
  model = c("no covar", "subcort.vol-euler", "vent.vol-euler", "cortex.vol-euler", "cbWM.vol-euler", "all.vol-euler", "all.cmd-euler", "only euler"),
  estimate = c(summary(lm_swyc_bwp)$coefficients[2, "Estimate"],
               summary(lm_swyc_bwp_subcort_euler)$coefficients[2, "Estimate"],
               summary(lm_swyc_bwp_vent_euler)$coefficients[2, "Estimate"],
               summary(lm_swyc_bwp_cortex_euler)$coefficients[2, "Estimate"],
               summary(lm_swyc_bwp_cbWM_euler)$coefficients[2, "Estimate"],
               summary(lm_swyc_bwp_allvol_euler)$coefficients[2, "Estimate"],
               summary(lm_swyc_bwp_allcmd_euler)$coefficients[2, "Estimate"],
               summary(lm_swyc_bwp_euler)$coefficients[2, "Estimate"]),
  se = c(summary(lm_swyc_bwp)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bwp_subcort_euler)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bwp_vent_euler)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bwp_cortex_euler)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bwp_cbWM_euler)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bwp_allvol_euler)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bwp_allcmd_euler)$coefficients[2, "Std. Error"],
               summary(lm_swyc_bwp_euler)$coefficients[2, "Std. Error"])
) %>%
  mutate(
    ymin = estimate - se,
    ymax = estimate + se
  )

beta_swyc_bwp_nocov = mediation_swycbwp[mediation_swycbwp$model == "no covar", "estimate"]
sehigh_swyc_bwp_nocov = mediation_swycbwp[mediation_swycbwp$model == "no covar", "ymax"]
selow_swyc_bwp_nocov = mediation_swycbwp[mediation_swycbwp$model == "no covar", "ymin"]

ggplot(mediation_swycbwp, aes(x = model, y = estimate)) +
  geom_pointrange(aes(ymin = ymin, ymax = ymax), linewidth = 0.8) +
  geom_hline(yintercept = beta_swyc_bwp_nocov, linetype = "dashed", color = "#7A7A7A", linewidth = 0.5) +
  geom_hline(yintercept = sehigh_swyc_bwp_nocov, linetype = "dashed", color = "#BDBDBD", linewidth = 0.5) +
  geom_hline(yintercept = selow_swyc_bwp_nocov, linetype = "dashed", color = "#BDBDBD", linewidth = 0.5) +
  scale_x_discrete(limits = c("no covar", "subcort.vol-euler", "vent.vol-euler", "cortex.vol-euler", "cbWM.vol-euler", "all.vol-euler", "all.cmd-euler", "only euler")) +
  labs(y = "Beta Estimate with Std Error", x = "LM Model covariates", title = "SWYC-DQ ~ Birth Weight Percentile") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
```

## Brain covars with GA
```{r}
# lm_swyc_ga <- lm(median_dq_zscore ~ gestational_age, data = swyc_avg_centiles)
# lm_swyc_ga_subcort_euler <- lm(median_dq_zscore ~ gestational_age +  global_SubCortGrayVol + median_euler_mean, data = swyc_avg_centiles)
# lm_swyc_ga_vent_euler <- lm(median_dq_zscore ~ gestational_age + global_VentricleChoroidVol + median_euler_mean, data = swyc_avg_centiles)
# lm_swyc_ga_cortex_euler <- lm(median_dq_zscore ~ gestational_age + global_CortexVol + median_euler_mean, data = swyc_avg_centiles)
# lm_swyc_ga_cbWM_euler <- lm(median_dq_zscore ~ gestational_age + global_CerebralWhiteMatterVol + median_euler_mean, data = swyc_avg_centiles)
# 
# lm_swyc_ga_allvol_euler <- lm(median_dq_zscore ~ gestational_age + global_CortexVol + global_CerebralWhiteMatterVol + global_SubCortGrayVol + global_VentricleChoroidVol + median_euler_mean, data = swyc_avg_centiles)
# 
# lm_swyc_ga_allcmd_euler <- lm(median_dq_zscore ~ gestational_age + + global.cmd + vol.cmd + sa.cmd + th.cmd + subcort.cmd + median_euler_mean, data = swyc_avg_centiles)
# lm_swyc_ga_euler <- lm(median_dq_zscore ~ gestational_age + median_euler_mean, data = swyc_avg_centiles)

lm_swyc_ga <- lm(ga_adj_median_dq_zscore ~ gestational_age, data = swyc_avg_centiles)
lm_swyc_ga_subcort_euler <- lm(ga_adj_median_dq_zscore ~ gestational_age +  global_SubCortGrayVol , data = swyc_avg_centiles)
lm_swyc_ga_vent_euler <- lm(ga_adj_median_dq_zscore ~ gestational_age + global_VentricleChoroidVol , data = swyc_avg_centiles)
lm_swyc_ga_cortex_euler <- lm(ga_adj_median_dq_zscore ~ gestational_age + global_CortexVol , data = swyc_avg_centiles)
lm_swyc_ga_cbWM_euler <- lm(ga_adj_median_dq_zscore ~ gestational_age + global_CerebralWhiteMatterVol , data = swyc_avg_centiles)

lm_swyc_ga_allvol_euler <- lm(ga_adj_median_dq_zscore ~ gestational_age + global_CortexVol + global_CerebralWhiteMatterVol + global_SubCortGrayVol + global_VentricleChoroidVol , data = swyc_avg_centiles)

lm_swyc_ga_allcmd_euler <- lm(ga_adj_median_dq_zscore ~ gestational_age + + global.cmd + vol.cmd + sa.cmd + th.cmd + subcort.cmd , data = swyc_avg_centiles)
lm_swyc_ga_euler <- lm(ga_adj_median_dq_zscore ~ gestational_age , data = swyc_avg_centiles)

#plot
mediation_swycga <- data.frame(
  model = c("no covar", "subcort.vol-euler", "vent.vol-euler", "cortex.vol-euler", "cbWM.vol-euler", "all.vol-euler", "all.cmd-euler", "only euler"),
  estimate = c(summary(lm_swyc_ga)$coefficients[2, "Estimate"],
               summary(lm_swyc_ga_subcort_euler)$coefficients[2, "Estimate"],
               summary(lm_swyc_ga_vent_euler)$coefficients[2, "Estimate"],
               summary(lm_swyc_ga_cortex_euler)$coefficients[2, "Estimate"],
               summary(lm_swyc_ga_cbWM_euler)$coefficients[2, "Estimate"],
               summary(lm_swyc_ga_allvol_euler)$coefficients[2, "Estimate"],
               summary(lm_swyc_ga_allcmd_euler)$coefficients[2, "Estimate"],
               summary(lm_swyc_ga_euler)$coefficients[2, "Estimate"]),
  se = c(summary(lm_swyc_ga)$coefficients[2, "Std. Error"],
               summary(lm_swyc_ga_subcort_euler)$coefficients[2, "Std. Error"],
               summary(lm_swyc_ga_vent_euler)$coefficients[2, "Std. Error"],
               summary(lm_swyc_ga_cortex_euler)$coefficients[2, "Std. Error"],
               summary(lm_swyc_ga_cbWM_euler)$coefficients[2, "Std. Error"],
               summary(lm_swyc_ga_allvol_euler)$coefficients[2, "Std. Error"],
               summary(lm_swyc_ga_allcmd_euler)$coefficients[2, "Std. Error"],
               summary(lm_swyc_ga_euler)$coefficients[2, "Std. Error"])
) %>%
  mutate(
    ymin = estimate - se,
    ymax = estimate + se
  )

beta_swyc_ga_nocov = mediation_swycga[mediation_swycga$model == "no covar", "estimate"]
sehigh_swyc_ga_nocov = mediation_swycga[mediation_swycga$model == "no covar", "ymax"]
selow_swyc_ga_nocov = mediation_swycga[mediation_swycga$model == "no covar", "ymin"]

ggplot(mediation_swycga, aes(x = model, y = estimate)) +
  geom_pointrange(aes(ymin = ymin, ymax = ymax), linewidth = 0.8) +
  geom_hline(yintercept = beta_swyc_ga_nocov, linetype = "dashed", color = "#7A7A7A", linewidth = 0.5) +
  geom_hline(yintercept = sehigh_swyc_ga_nocov, linetype = "dashed", color = "#BDBDBD", linewidth = 0.5) +
  geom_hline(yintercept = selow_swyc_ga_nocov, linetype = "dashed", color = "#BDBDBD", linewidth = 0.5) +
  scale_x_discrete(limits = c("no covar", "subcort.vol-euler", "vent.vol-euler", "cortex.vol-euler", "cbWM.vol-euler", "all.vol-euler", "all.cmd-euler", "only euler")) +
  labs(y = "Beta Estimate with Std Error", x = "LM Model covariates", title = "SWYC-DQ ~ Gestational Age") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
```

## Test via bootstrapping (code from Kevin)
He used it to test if the inclusion of the PFN decreased the effect of PRS_1 on p-factor
From rotation with Ran, who used lavaan to do mediation --- Kevin said maybe this is a more appropriate method

### Gestational Age
```{r}
bootga <- bootstrap_model(lm_swyc_ga, iterations = 100)
bootga$model <- "no_covar"

bootga_globalcmd_euler <- bootstrap_model(lm_swyc_ga_globalcmd_euler, iterations = 100)
bootga_globalcmd_euler$model <- "global.cmd_euler"
bootga_thcmd_euler <- bootstrap_model(lm_swyc_ga_thcmd_euler, iterations = 100)
bootga_thcmd_euler$model <- "th.cmd_euler"
bootga_subcort_euler <- bootstrap_model(lm_swyc_ga_subcort_euler, iterations = 100)
bootga_subcort_euler$model <- "subcort_euler"
bootga_vent_euler <- bootstrap_model(lm_swyc_ga_vent_euler, iterations = 100)
bootga_vent_euler$model <- "vent_euler"

t.test(bootga$gestational_age,bootga_thcmd_euler$gestational_age,paired=FALSE)
t.test(bootga$gestational_age,bootga_globalcmd_euler$gestational_age,paired=FALSE)
t.test(bootga$gestational_age,bootga_subcort_euler$gestational_age,paired=FALSE)
t.test(bootga$gestational_age,bootga_vent_euler$gestational_age,paired=FALSE)

models_tested_ga <- list(global.cmd_euler = bootga_globalcmd_euler %>% select(gestational_age, model),
                        th.cmd_euler = bootga_thcmd_euler %>% select(gestational_age, model),
                        subcort_euler = bootga_subcort_euler %>% select(gestational_age, model),
                        vent_euler = bootga_vent_euler %>% select(gestational_age, model))

bootga_plot_df <- do.call(rbind,
  lapply(names(models_tested_ga), function(facet_name) {
    df <- models_tested_ga[[facet_name]]
    df <- rbind(df, bootga %>% select(gestational_age, model))
    df$facet <- facet_name
    df
    }))

ggplot(bootga_plot_df) +
  geom_density(aes(x = gestational_age, fill = model), alpha = 0.5) +
  facet_wrap(~facet) +
  labs(
    x = "Beta Estimates",
    y = "",
    title = "Bootstrap Betas, Gestational Age "
  ) +
  scale_fill_manual(
    name = "Model",
    values = c("no_covar" = "#7A7A7A", 
               "global.cmd_euler" = "#D08C8A",
               "th.cmd_euler" = "#C9A87C",
               "subcort_euler" = "#A6C48A",
               "vent_euler" = "#90A0B5") ) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5),
    legend.position = "top",
    strip.text = element_blank())

bootga_plot_summary_df <- rbind(do.call(rbind, models_tested_ga), bootga %>% select(gestational_age, model)) %>%
  group_by(model) %>%
  summarize(ga_beta_mean = mean(gestational_age),
            ga_lower_CI = t.test(gestational_age)$conf.int[1],
            ga_upper_CI = t.test(gestational_age)$conf.int[2])

ggplot(bootga_plot_summary_df, aes(x = model, y = ga_beta_mean)) +
  geom_pointrange(aes(ymin = ga_lower_CI, ymax = ga_upper_CI), linewidth = 0.8) +
  scale_x_discrete(limits = c("no_covar", "global.cmd_euler", "th.cmd_euler", "subcort_euler", "vent_euler")) +
  labs(y = "Bootstrap Beta Mean with 95% CI", x = "LM Model covariates", title = "SWYC-DQ ~ Gestational Age") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

hist(bootga$gestational_age)
mean(bootga$gestational_age)
median(bootga$gestational_age)

hist(bootga_globalcmd_euler$gestational_age)
mean(bootga_globalcmd_euler$gestational_age)
median(bootga_globalcmd_euler$gestational_age)
shapiro.test(bootga_globalcmd_euler$gestational_age)
```

### Birth Weight
```{r}
bootbw <- bootstrap_model(lm_swyc_bw, iterations = 100)
bootbw$model <- "no_covar"

bootbw_globalcmd_euler <- bootstrap_model(lm_swyc_bw_globalcmd_euler, iterations = 100)
bootbw_globalcmd_euler$model <- "global.cmd_euler"
bootbw_thcmd_euler <- bootstrap_model(lm_swyc_bw_thcmd_euler, iterations = 100)
bootbw_thcmd_euler$model <- "th.cmd_euler"
bootbw_subcort_euler <- bootstrap_model(lm_swyc_bw_subcort_euler, iterations = 100)
bootbw_subcort_euler$model <- "subcort_euler"
bootbw_vent_euler <- bootstrap_model(lm_swyc_bw_vent_euler, iterations = 100)
bootbw_vent_euler$model <- "vent_euler"

t.test(bootbw$birth_weight_kg,bootbw_thcmd_euler$birth_weight_kg,paired=FALSE)
t.test(bootbw$birth_weight_kg,bootbw_globalcmd_euler$birth_weight_kg,paired=FALSE)
t.test(bootbw$birth_weight_kg,bootbw_subcort_euler$birth_weight_kg,paired=FALSE)
t.test(bootbw$birth_weight_kg,bootbw_vent_euler$birth_weight_kg,paired=FALSE)

models_tested_bw <- list(global.cmd_euler = bootbw_globalcmd_euler %>% select(birth_weight_kg, model),
                        th.cmd_euler = bootbw_thcmd_euler %>% select(birth_weight_kg, model),
                        subcort_euler = bootbw_subcort_euler %>% select(birth_weight_kg, model),
                        vent_euler = bootbw_vent_euler %>% select(birth_weight_kg, model))

bootbw_plot_df <- do.call(rbind,
  lapply(names(models_tested_bw), function(facet_name) {
    df <- models_tested_bw[[facet_name]]
    df <- rbind(df, bootbw %>% select(birth_weight_kg, model))
    df$facet <- facet_name
    df
    }))

ggplot(bootbw_plot_df) +
  geom_density(aes(x = birth_weight_kg, fill = model), alpha = 0.5) +
  facet_wrap(~facet) +
  labs(
    x = "Beta Estimates",
    y = "",
    title = "Bootstrap Betas, Birth Weight "
  ) +
  scale_fill_manual(
    name = "Model",
    values = c("no_covar" = "#7A7A7A", 
               "global.cmd_euler" = "#D08C8A",
               "th.cmd_euler" = "#C9A87C",
               "subcort_euler" = "#A6C48A",
               "vent_euler" = "#90A0B5") ) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5),
    legend.position = "top",
    strip.text = element_blank())
```

### Birth Weight Percentile
```{r}
bootbwp <- bootstrap_model(lm_swyc_bwp, iterations = 100)
bootbwp$model <- "no_covar"

bootbwp_globalcmd_euler <- bootstrap_model(lm_swyc_bwp_globalcmd_euler, iterations = 100)
bootbwp_globalcmd_euler$model <- "global.cmd_euler"
bootbwp_thcmd_euler <- bootstrap_model(lm_swyc_bwp_thcmd_euler, iterations = 100)
bootbwp_thcmd_euler$model <- "th.cmd_euler"
bootbwp_subcort_euler <- bootstrap_model(lm_swyc_bwp_subcort_euler, iterations = 100)
bootbwp_subcort_euler$model <- "subcort_euler"
bootbwp_vent_euler <- bootstrap_model(lm_swyc_bwp_vent_euler, iterations = 100)
bootbwp_vent_euler$model <- "vent_euler"

t.test(bootbwp$bwp_fen,bootbwp_thcmd_euler$bwp_fen,paired=FALSE)
t.test(bootbwp$bwp_fen,bootbwp_globalcmd_euler$bwp_fen,paired=FALSE)
t.test(bootbwp$bwp_fen,bootbwp_subcort_euler$bwp_fen,paired=FALSE)
t.test(bootbwp$bwp_fen,bootbwp_vent_euler$bwp_fen,paired=FALSE)

models_tested_bwp <- list(global.cmd_euler = bootbwp_globalcmd_euler %>% select(bwp_fen, model),
                        th.cmd_euler = bootbwp_thcmd_euler %>% select(bwp_fen, model),
                        subcort_euler = bootbwp_subcort_euler %>% select(bwp_fen, model),
                        vent_euler = bootbwp_vent_euler %>% select(bwp_fen, model))

bootbwp_plot_df <- do.call(rbind,
  lapply(names(models_tested_bwp), function(facet_name) {
    df <- models_tested_bwp[[facet_name]]
    df <- rbind(df, bootbwp %>% select(bwp_fen, model))
    df$facet <- facet_name
    df
    }))

ggplot(bootbwp_plot_df) +
  geom_density(aes(x = bwp_fen, fill = model), alpha = 0.5) +
  facet_wrap(~facet) +
  labs(
    x = "Beta Estimates",
    y = "",
    title = "Bootstrap Betas, Birth Weight Percentile"
  ) +
  scale_fill_manual(
    name = "Model",
    values = c("no_covar" = "#7A7A7A", 
               "global.cmd_euler" = "#D08C8A",
               "th.cmd_euler" = "#C9A87C",
               "subcort_euler" = "#A6C48A",
               "vent_euler" = "#90A0B5") ) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5),
    legend.position = "top",
    strip.text = element_blank())
```

## Test via partial R (code in python from Briana, translated to R) 
[need to check code]
```{r}
# y = standardized outcome
y <- scale(df[[feat]])[, 1]  # scale() returns a matrix; [,1] converts to vector

# full model
X_full <- as.data.frame(scale(df[predictors]))
X_full$const <- 1            # equivalent to sm.add_constant
model_full <- lm(y ~ . + 0, data = X_full)  # +0 because const column added manually

# reduced model (exclude target_cov)
reduced_predictors <- setdiff(predictors, target_cov)
X_reduced <- as.data.frame(scale(df[reduced_predictors]))
X_reduced$const <- 1
model_reduced <- lm(y ~ . + 0, data = X_reduced)

# compute partial r^2
model_red <- lm(median_dq ~ global.cmd + median_euler_mean, data = swyc_avg_centiles)
r2_full <- summary(lm_swyc_ga_globalcmd_euler)$r.squared
r2_red  <- summary(model_red)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(lm_swyc_ga_globalcmd_euler)["gestational_age"])
partial_r <- sign_coef * sqrt(abs(partial_r2))

r2_full <- summary(lm_swyc_ga_globalcmd)$r.squared
r2_red  <- summary(lm_swyc_ga)$r.squared
partial_r2 <- (r2_full - r2_red) / (1 - r2_red)

# sign from coefficient
sign_coef <- sign(coef(lm_swyc_ga_globalcmd_euler)["global.cmd"])
partial_r <- sign_coef * sqrt(abs(partial_r2))
```
## Test mediation per swyc form (not median dq)

### Regressions for Global
```{r}
vars_totest <- c("global_eTIV", "global_SubCortGrayVol","global_CortexVol", "global_CerebralWhiteMatterVol", "global_VentricleChoroidVol","global_SurfaceArea", "global_MeanThickness") 
swyc_forms_totest <- c("Month-6", "Month-9", "Month-15", "Month-18", "Month-24", "Month-30")

grid <- expand.grid(
  pheno = vars_totest,
  swyc_form = swyc_forms_totest,
  ga_adj = c(TRUE, FALSE),
  stringsAsFactors = FALSE
)

df <- swyc_avg_centiles.swyc.all.cent %>% 
  select(pat_id, gestational_age, birth_weight_kg_qc, vars_totest, 
         form_friendly_name, norm_devo_quotient, ga_adj_norm_devo_quotient_round, 
         median_euler_mean, euler_mean_mpr,complete_centile_vars) %>%
  rename(swyc_form = form_friendly_name) %>%
  mutate(swyc_form = recode(swyc_form,
                      "SWYC : 04 Month" = "Month-4",
                      "SWYC : 06 Month" = "Month-6",
                      "SWYC : 09 Month" = "Month-9",
                      "SWYC : 12 Month" = "Month-12",
                      "SWYC : 15 Month" = "Month-15",
                      "SWYC : 18 Month" = "Month-18",
                      "SWYC : 24 Month" = "Month-24",
                      "SWYC : 30 Month" = "Month-30",
                      "SWYC : 36 Month" = "Month-36",
                      "SWYC : 48 Month" = "Month-48",
                      "SWYC : 60 Month" = "Month-60")) %>%
  pivot_longer(cols = c(ga_adj_norm_devo_quotient_round, norm_devo_quotient),
               names_to = "ga_adj",
               values_to = "dq") %>%
  mutate(ga_adj = grepl("ga_adj", ga_adj))
  
results <- do.call(rbind, lapply(seq_len(nrow(grid)), function(i) {
  
  pheno_i <- grid$pheno[i]
  form_i  <- grid$swyc_form[i]
  ga_i    <- grid$ga_adj[i]
  
  df_sub <- df %>%
    dplyr::filter(swyc_form == form_i, ga_adj == ga_i) %>%
    dplyr::select(all_of(c("dq", pheno_i, "median_euler_mean"))) %>%
    na.omit()
  
  if (nrow(df_sub) < 3) {
    return(data.frame(
      pheno = pheno_i,
      swyc_type = form_i,
      ga_adj = ga_i,
      N = nrow(df_sub),
      beta = NA,
      p = NA,
      se = NA
    ))
  }
  
  fit <- lm(reformulate(c(pheno_i, "median_euler_mean"), response = "dq"),
            data = df_sub)
  
  coef_tab <- summary(fit)$coefficients
  
  data.frame(
    pheno = pheno_i,
    swyc_form = form_i,
    ga_adj = ga_i,
    N = nrow(df_sub),
    beta = coef_tab[pheno_i, "Estimate"],
    p = coef_tab[pheno_i, "Pr(>|t|)"],
    se = coef_tab[pheno_i, "Std. Error"],
    p_adj_pheno = p.adjust(coef_tab[pheno_i, "Pr(>|t|)"], method = "bonferroni", n = length(vars_totest)),
    p_adj_phenoform = p.adjust(coef_tab[pheno_i, "Pr(>|t|)"], method = "bonferroni", n = length(vars_totest)*length(swyc_forms_totest))
  )
}))

# arrange by swyc form and add columns for plotting
results <- results %>%
  mutate(swyc_form = factor(swyc_form, levels = c("Month-6", "Month-9", "Month-15", "Month-18", "Month-24", "Month-30"))) %>% arrange(swyc_form) %>%
  mutate(sig = p_adj_phenoform < 0.05,
          color = case_when(
            sig & beta > 0 ~ "#c87a7a",   #  muted red
            sig & beta < 0 ~ "#7293b0",   #  muted blue
            TRUE ~ "gray70"),
         y_labels = paste(swyc_form, N, sep = ", N="))


(cent_allforms_global_plot <- ggplot(results, aes(x = beta, y = swyc_form, 
                                                  color = ga_adj, alpha = p_adj_pheno <= 0.05)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbarh(aes(xmin = beta - se, xmax = beta + se),
                 height = 0.2,
                 position = position_dodge(width = 0.5)) +
  facet_wrap(~pheno, scales = "free_x") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "#B0B0B0") +
  theme_classic() +
  theme(
        panel.grid = element_blank(),
        legend.position = "bottom",
        legend.title = element_blank(),
        panel.border = element_blank(),
        axis.text.y  = element_text(color = "black"),
        axis.line.x  = element_line(color = "black"),
        axis.line.y  = element_blank(),
        legend.justification = c("left", "top")
    ) +
    xlab("Std Beta DQ ~ Global Centiles, Across Forms > N = 40") +
    ylab(""))


# if(save_figs == TRUE){
# panel_save(swyc_bf_all_plot, "swyc-bf-all", "pdf", plot_output_folder)
# panel_save(swyc_bf_median_plot, "swyc-bf-median", "pdf", plot_output_folder)
# panel_save(swyc_bf_allforms_ga, "swyc-bf-allforms-ga", "pdf", plot_output_folder)
# }
```

### Regressions for CMD
```{r}
vars_totest <- c("global.cmd", "vol.cmd", "sa.cmd", "th.cmd") 
swyc_forms_totest <- c("Month-6", "Month-9", "Month-15", "Month-18", "Month-24", "Month-30")

cmd.grid <- expand.grid(
  pheno = vars_totest,
  swyc_form = swyc_forms_totest,
  ga_adj = c(TRUE, FALSE),
  stringsAsFactors = FALSE
)

df <- swyc_avg_centiles.swyc.all.cent %>% 
  select(pat_id, gestational_age, birth_weight_kg_qc, vars_totest, 
         form_friendly_name, norm_devo_quotient, ga_adj_norm_devo_quotient_round, 
         median_euler_mean, euler_mean_mpr,complete_centile_vars,
         global.cmd, vol.cmd, sa.cmd, th.cmd) %>%
  rename(swyc_form = form_friendly_name) %>%
  mutate(swyc_form = recode(swyc_form,
                      "SWYC : 04 Month" = "Month-4",
                      "SWYC : 06 Month" = "Month-6",
                      "SWYC : 09 Month" = "Month-9",
                      "SWYC : 12 Month" = "Month-12",
                      "SWYC : 15 Month" = "Month-15",
                      "SWYC : 18 Month" = "Month-18",
                      "SWYC : 24 Month" = "Month-24",
                      "SWYC : 30 Month" = "Month-30",
                      "SWYC : 36 Month" = "Month-36",
                      "SWYC : 48 Month" = "Month-48",
                      "SWYC : 60 Month" = "Month-60")) %>%
  pivot_longer(cols = c(ga_adj_norm_devo_quotient_round, norm_devo_quotient),
               names_to = "ga_adj",
               values_to = "dq") %>%
  mutate(ga_adj = grepl("ga_adj", ga_adj))
  
cmd.results <- do.call(rbind, lapply(seq_len(nrow(cmd.grid)), function(i) {
  
  pheno_i <- cmd.grid$pheno[i]
  form_i  <- cmd.grid$swyc_form[i]
  ga_i    <- cmd.grid$ga_adj[i]
  
  df_sub <- df %>%
    dplyr::filter(swyc_form == form_i, ga_adj == ga_i) %>%
    dplyr::select(all_of(c("dq", pheno_i, "median_euler_mean"))) %>%
    na.omit()
  
  if (nrow(df_sub) < 3) {
    return(data.frame(
      pheno = pheno_i,
      swyc_type = form_i,
      ga_adj = ga_i,
      N = nrow(df_sub),
      beta = NA,
      p = NA,
      se = NA
    ))
  }
  
  fit <- lm(reformulate(c(pheno_i, "median_euler_mean"), response = "dq"),
            data = df_sub)
  
  coef_tab <- summary(fit)$coefficients
  
  data.frame(
    pheno = pheno_i,
    swyc_form = form_i,
    ga_adj = ga_i,
    N = nrow(df_sub),
    beta = coef_tab[pheno_i, "Estimate"],
    p = coef_tab[pheno_i, "Pr(>|t|)"],
    se = coef_tab[pheno_i, "Std. Error"],
    p_adj_pheno = p.adjust(coef_tab[pheno_i, "Pr(>|t|)"], method = "bonferroni", n = length(vars_totest)),
    p_adj_phenoform = p.adjust(coef_tab[pheno_i, "Pr(>|t|)"], method = "bonferroni", n = length(vars_totest)*length(swyc_forms_totest))
  )
}))


# arrange by swyc form and add columns for plotting
cmd.results <- cmd.results %>%
  mutate(swyc_form = factor(swyc_form, levels = c("Month-6", "Month-9", "Month-15", "Month-18", "Month-24", "Month-30"))) %>% arrange(swyc_form) %>%
  mutate(sig = p_adj_phenoform < 0.05,
          color = case_when(
            sig & beta > 0 ~ "#c87a7a",   #  muted red
            sig & beta < 0 ~ "#7293b0",   #  muted blue
            TRUE ~ "gray70"),
         y_labels = paste(swyc_form, pheno, ga_adj, N, sep = "_"))


(cent_allforms_cmd_plot <- ggplot(cmd.results, aes(x = beta, y = swyc_form, 
                                                  color = ga_adj, alpha = p_adj_pheno <= 0.05)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_text(aes(label = N),
          vjust = -0.5) +
  geom_errorbarh(aes(xmin = beta - se, xmax = beta + se),
                 height = 0.2,
                 position = position_dodge(width = 0.5)) +
  facet_wrap(~pheno, scales = "free_x") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "#B0B0B0") +
  scale_alpha_manual(values = c("FALSE" = 0.35, "TRUE" = 0.8),
    labels = c("FALSE" = "nonsig",
               "TRUE" = "p < 0.05"), 
    name = "Bonferroni corrected across phenotypes (N = 4)") +
  scale_color_manual( values = c("FALSE" = "#4F85B5", "TRUE" = "#B34F6B"),
    labels = c("FALSE" = "Original DQ",
               "TRUE" = "GA-ADJ DQ"), name = "DQ CALCULATION") +
  theme_classic() +
  theme(
        panel.grid = element_blank(),
        legend.position = "bottom",
        panel.border = element_blank(),
        axis.text.y  = element_text(color = "black"),
        axis.line.x  = element_line(color = "black"),
        axis.line.y  = element_blank(),
        legend.justification = c("left", "top")
    ) +
    xlab("Std Beta DQ ~ CMDs, Across Forms > N = 40") +
    ylab(""))


# if(save_figs == TRUE){
# panel_save(swyc_bf_all_plot, "swyc-bf-all", "pdf", plot_output_folder)
# panel_save(swyc_bf_median_plot, "swyc-bf-median", "pdf", plot_output_folder)
# panel_save(swyc_bf_allforms_ga, "swyc-bf-allforms-ga", "pdf", plot_output_folder)
# }
```
# Data Availability - All (includes repeated scan measures) (repeat analyses in "temporally restricted" sample)

## First-Pass plots for scan-swyc timing
```{r}
centiles_all$participant_id_modified <- sub("sub-", "", centiles_all$participant_id)
swyc_subset_all <- swyc %>% filter(pat_id %in% centiles_all$participant_id_modified)
swyc_subset_all <- merge(swyc_subset_all, centiles_all %>% select(c("participant_id_modified", "age_at_scan", "gestational_age", "birth_weight_kg")), 
                     by.x = "pat_id", by.y = "participant_id_modified", all.x = TRUE, all.y = FALSE)
swyc_subset_all$age_diff <- swyc_subset_all$age_at_scan - swyc_subset_all$effective_age

if (save_processed == TRUE){
  write.csv(swyc_subset_all, glue("{csv_tables_folder}swyc_overlap_centiles_all_2511_gabwinc.csv"))
}

hist(swyc_subset_all$age_diff)
summary(swyc_subset_all$age_diff)
sum(swyc_subset_all$age_diff <= 0)
(swyc_overlap_plot <- ggplot(swyc_subset_all, aes(y = pat_id, group = pat_id)) +
  geom_point(aes(x = effective_age, color = form_friendly_name), size = 0.2) +
  geom_line(aes(x = effective_age, color = form_friendly_name),size = 0.1) + 
  geom_point(data = swyc_subset_all %>% 
              filter(age_at_scan <= max(effective_age) | age_at_scan >= min(effective_age)), 
            aes(x = age_at_scan), color = "black", size = 0.1, alpha = 0.5) +
  theme(axis.text.y  = element_blank())) 

(swyc_overlap_scanb4swyc_plot <- ggplot(swyc_subset_all %>% filter(age_diff <= 0),  aes(y = pat_id, group = pat_id)) +
  geom_point(aes(x = effective_age, color = form_friendly_name), size = 0.2) +
  geom_line(aes(x = effective_age, color = form_friendly_name),size = 0.1) + 
  geom_point(aes(x = age_at_scan), color = "black", size = 0.1, alpha = 0.5) +
  theme(axis.text.y  = element_blank())) 
```

## Determine availability of SWYC based on scan date

### adjust age at scan and effective age (swyc) for GA and convert to months
```{r}
# adjust effective age and age at scan for prematurity
swyc_subset_all$effective_age_months <- swyc_subset_all$effective_age/30.437
swyc_subset_all$adj_effective_age_months <- ifelse(!is.na(swyc_subset_all$gestational_age),
                                                 swyc_subset_all$effective_age_months + swyc_subset_all$gestational_age/4.35,
                                                 swyc_subset_all$effective_age_months + 40/4.35)

swyc_subset_all$age_at_scan_months <- swyc_subset_all$age_at_scan/30.437
swyc_subset_all$adj_age_at_scan_months <- ifelse(!is.na(swyc_subset_all$gestational_age),
                                                 swyc_subset_all$age_at_scan_months + swyc_subset_all$gestational_age/4.35,
                                                 swyc_subset_all$age_at_scan_months + 40/4.35)

swyc_subset_all$age_diff_months <- swyc_subset_all$age_at_scan_months - swyc_subset_all$adj_effective_age_months
```

### swyc within 10% of scan in either direction
```{r, echo = FALSE}
swyc_subset_all_matched <- swyc_subset_all %>% filter(abs(age_diff_months) < 0.1*age_at_scan_months)
length(unique(swyc_subset_all_matched$pat_id))
dim(swyc_subset_all_matched)

(cent_all_matched_plot <- ggplot(swyc_subset_all_matched, aes(y = pat_id, group = pat_id)) +
  geom_point(aes(x = adj_effective_age_months, color = form_friendly_name), size = 0.2) +
  geom_line(aes(x = adj_effective_age_months, color = form_friendly_name),size = 0.1) + 
  geom_point(aes(x = adj_age_at_scan_months), color = "black", size = 0.1, alpha = 0.5) +
  theme(axis.text.y  = element_blank(),
        panel.grid = element_blank(),
        panel.background = element_blank()))
```



### scan before SWYC, take average of SWYCs
```{r, echo = FALSE}
swyc_subset_all_matched2 <- swyc_subset_all %>% filter(age_diff_months < 0)
length(unique(swyc_subset_all_matched2$pat_id))
dim(swyc_subset_all_matched2)
dim(swyc_subset_all_matched2[duplicated(swyc_subset_all_matched2) | duplicated(swyc_subset_all_matched2, fromLast = TRUE), ])
swyc_subset_all_matched2 <- swyc_subset_all_matched2[!duplicated(swyc_subset_all_matched2),]
dim(swyc_subset_all_matched2)


(cent_all_matched2_plot <- ggplot(swyc_subset_all_matched2, aes(y = pat_id, group = pat_id)) +
  geom_point(aes(x = adj_effective_age_months, color = form_friendly_name), size = 0.2) +
  geom_line(aes(x = adj_effective_age_months, color = form_friendly_name),size = 0.1) + 
  geom_point(aes(x = adj_age_at_scan_months), color = "black", size = 0.1, alpha = 0.5) +
  theme(axis.text.y  = element_blank(),
        panel.grid = element_blank(),
        panel.background = element_blank()))

# take average of SWYCs
swyc_subset_matched2_avg <- swyc_subset_all_matched2 %>% 
  group_by(pat_id) %>%
  summarise(
    mean_dq = mean(norm_devo_quotient),
    median_dq = median(norm_devo_quotient)
  )
dim(swyc_subset_matched2_avg)
```

### scan closest to SWYC, remove if > 24 months, choose the earliest scan
```{r, echo = FALSE}
swyc_subset_all_matched3 <- swyc_subset_all %>% 
  group_by(pat_id) %>%
  filter(age_diff_months == min(age_diff_months) & age_diff_months <= 24) %>%
  filter(age_at_scan == min(age_at_scan)) %>%
  ungroup() %>%
  mutate(pat_age = paste0(pat_id, effective_age)) %>%
  all(pat_age, .keep_all = TRUE)

length(unique(swyc_subset_all_matched3$pat_id))
dim(swyc_subset_all_matched3)
dim(swyc_subset_all_matched3[duplicated(swyc_subset_all_matched3) | duplicated(swyc_subset_all_matched3, fromLast = TRUE), ])
swyc_subset_all_matched3 <- swyc_subset_all_matched3[!duplicated(swyc_subset_all_matched3),]
dim(swyc_subset_all_matched3)

(cent_all_matched3_plot <- ggplot(swyc_subset_all_matched3, aes(y = pat_id, group = pat_id)) +
  geom_point(aes(x = adj_effective_age_months, color = form_friendly_name), size = 0.2) +
  geom_line(aes(x = adj_effective_age_months, color = form_friendly_name),size = 0.1) + 
  geom_point(aes(x = adj_age_at_scan_months), color = "black", size = 0.1, alpha = 0.5) +
  theme(axis.text.y  = element_blank(),
        panel.grid = element_blank(),
        panel.background = element_blank()))
hist(swyc_subset_all_matched3$age_diff_months)
```

# Repeat Analyses with Date-Matched Subset

## Prep Data for Analysis

### Merge swyc and centiles df
```{r}
swyc_subset_matched_merged <- merge(swyc_subset_all_matched3, centiles_all %>% select(all_of(c("participant_id_modified",
                                                                                                   "age_at_scan", "preterm",
                                                                        "birth_weight_kg", "bwp_fen",
                                                                        "sex", "adjusted_age_in_days","median_euler_mean",
                                                                        full_vars))), 
                     by.x = c("pat_id", "age_at_scan"), by.y = c("participant_id_modified", "age_at_scan"), all.x = TRUE, all.y = FALSE)
```

### keep one row per participant id (for now)
```{r}
swyc_subset_matched_merged <- swyc_subset_matched_merged %>%
  all(pat_id, .keep_all = TRUE)
```

### zscore variables of interest
```{r, echo = FALSE}
# zscore centiles and other variables of interest
df_matched.zscore <- swyc_subset_matched_merged %>%
  mutate(across(all_of(c(full_vars, "bwp_fen")), qnorm)) %>%
  mutate(across(all_of(c("gestational_age", "birth_weight_kg", "adjusted_age_in_days", 
                         "median_euler_mean", "norm_devo_quotient")), scale))
# set sex and preterm as factor
df_matched.zscore$sex <- as.factor(df_matched.zscore$sex)
df_matched.zscore$preterm <- as.factor(df_matched.zscore$preterm)
```

### find Inf cells, make a table of N per phenotype, then convert to NaN
```{r, echo = FALSE}
inf_nan_table <- as.data.frame(t(sapply(df_matched.zscore[c(full_vars, "bwp_fen")], function(x){
  c(inf = sum(is.infinite(x)), nan = sum(is.nan(x)))
})))
max(inf_nan_table$inf)
max(inf_nan_table$nan)

inf_nan_table %>% filter(inf > 0)
```

### set Inf cells to NaN
```{r, echo = FALSE}
df_matched.zscore <- df_matched.zscore %>% 
  mutate(across(all_of(c(full_vars, "bwp_fen")), ~ replace(., is.infinite(.), NaN)))
```

## Global variables

### Define global vars of interest
```{r, echo = TRUE}
global_vars_totest <- c("global_eTIV", "global_SubCortGrayVol","global_CortexVol", "global_CerebralWhiteMatterVol", "global_VentricleChoroidVol","global_SurfaceArea", "global_MeanThickness")
```

### SWYC ~ global without QC
```{r, echo = FALSE}
swyc_matched_reg_table_global <- as.data.frame(t(sapply(df_matched.zscore[, which(names(df_matched.zscore) %in% global_vars_totest)], function(x) {
  lm <- lm(x ~ df_matched.zscore$norm_devo_quotient)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2])
})))

swyc_matched_reg_table_global$swyc_p_adj <- p.adjust(swyc_matched_reg_table_global$swyc_p, method = "BH")

swyc_matched_reg_table_global %>% filter(swyc_p_adj < 0.05)
range(swyc_matched_reg_table_global %>% filter(swyc_p_adj < 0.05) %>% select(swyc_beta))
```

### SWYC ~ global with QC
```{r, echo = FALSE}
swyc.qc_reg_table_global <- as.data.frame(t(sapply(df_matched.zscore[, which(names(df_matched.zscore) %in% global_vars_totest)], function(x) {
  lm <- lm(x ~ df_matched.zscore$norm_devo_quotient + df_matched.zscore$median_euler_mean)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2])
})))

swyc.qc_reg_table_global$swyc_p_adj <- p.adjust(swyc.qc_reg_table_global$swyc_p, method = "BH")

swyc.qc_reg_table_global %>% filter(swyc_p_adj < 0.05)
range(swyc.qc_reg_table_global %>% filter(swyc_p_adj < 0.05) %>% select(swyc_beta))
```

### Visualize/Figure Panel without QC
```{r, echo = FALSE}
#  Add phenotype column from rownames
swyc_matched_reg_table_global$phenotype <- rownames(swyc_matched_reg_table_global)

#  Add model identifiers + rename to common column names
swyc_matched_reg_table_global <- swyc_matched_reg_table_global %>%
  mutate(model = "SWYC",
         beta = swyc_beta,
         se   = swyc_se,
         p_adj    = swyc_p_adj) %>%
    mutate(sig = p_adj < 0.05,
          color = case_when(
            sig & beta > 0 ~ "#c87a7a",   #  muted red
            sig & beta < 0 ~ "#7293b0",   #  muted blue
            TRUE ~ "gray70"))

#  Combine and arrange by specific order of global phenotypes
df <- swyc_matched_reg_table_global %>%
  mutate(phenotype = factor(phenotype, levels = c("global_eTIV", "global_CortexVol", "global_CerebralWhiteMatterVol",
                                                  "global_SubCortGrayVol", "global_VentricleChoroidVol",
                                                  "global_SurfaceArea", "global_MeanThickness"))) %>% arrange(phenotype)

#  Labeling for phenotypes on plot
phenotype_labs <- c(
 global_eTIV = "Intracranial Volume",
 global_MeanThickness = "Mean Cortical Thickness",
 global_VentricleChoroidVol = "Ventricular Volume",
 global_SurfaceArea = "Total Surface Area",
 global_CortexVol = "Cortical Gray Matter Volume",
 global_CerebralWhiteMatterVol = "Cerebral White Matter Volume",
 global_SubCortGrayVol = "Subcortical Gray Matter Volume"
)

#  Plot
(global_betas_noqc_plot <- ggplot(df, aes(x = beta, y = phenotype)) +
    geom_point(aes(shape = model, color = color), size = 3) +
    geom_errorbarh(aes(xmin = beta - se, xmax = beta + se, color = color), height = 0.2) +
    scale_shape_manual(values = c("SWYC" = 8),
                       name = "") +
    scale_color_identity() +
    scale_x_continuous(limits = c(-0.25, 0.35)) +
    scale_y_discrete(limits = rev, labels = phenotype_labs) +
    geom_vline(xintercept = 0, linetype = "dotted", color = "darkgray") +
    theme_minimal() +
    theme_bw(base_size = 16) +
    theme(
        panel.grid = element_blank(),
        legend.position = "bottom",
        legend.title = element_blank(),
        panel.border = element_blank(),
        axis.text.y  = element_text(color = "black"),
        axis.line.x  = element_line(color = "black"),
        axis.line.y  = element_blank(),
        legend.justification = c("left", "top")
    ) +
    xlab("Standardized Beta Estimate, without QC covar") +
    ylab(""))
```

### Visualize/Figure Panel with QC
```{r, echo = FALSE}
#  Add phenotype column from rownames
swyc.qc_reg_table_global$phenotype <- rownames(swyc.qc_reg_table_global)

swyc.qc_reg_table_global <- swyc.qc_reg_table_global %>%
  mutate(model = "SWYC",
         beta = swyc_beta,
         se   = swyc_se,
         p_adj    = swyc_p_adj) %>%
    mutate(sig = p_adj < 0.05,
          color = case_when(
            sig & beta > 0 ~ "#c87a7a",   #  muted red
            sig & beta < 0 ~ "#7293b0",   #  muted blue
            TRUE ~ "gray70"))

#  Combine and arrange by specific order of global phenotypes
df <- swyc.qc_reg_table_global %>%
  mutate(phenotype = factor(phenotype, levels = c("global_eTIV", "global_CortexVol", "global_CerebralWhiteMatterVol",
                                                  "global_SubCortGrayVol", "global_VentricleChoroidVol",
                                                  "global_SurfaceArea", "global_MeanThickness"))) %>% arrange(phenotype)

#  Labeling for phenotypes on plot
phenotype_labs <- c(
 global_eTIV = "Intracranial Volume",
 global_MeanThickness = "Mean Cortical Thickness",
 global_VentricleChoroidVol = "Ventricular Volume",
 global_SurfaceArea = "Total Surface Area",
 global_CortexVol = "Cortical Gray Matter Volume",
 global_CerebralWhiteMatterVol = "Cerebral White Matter Volume",
 global_SubCortGrayVol = "Subcortical Gray Matter Volume"
)

#  Plot
(global_betas_withqc_plot <- ggplot(df, aes(x = beta, y = phenotype)) +
    geom_point(aes(shape = model, color = color), size = 3) +
    geom_errorbarh(aes(xmin = beta - se, xmax = beta + se, color = color), height = 0.2) +
    scale_shape_manual(values = c("SWYC" = 8),
                       name = "") +
    scale_color_identity() +
    scale_x_continuous(limits = c(-0.25, 0.35)) +
    scale_y_discrete(limits = rev, labels = phenotype_labs) +
    geom_vline(xintercept = 0, linetype = "dotted", color = "darkgray") +
    theme_minimal() +
    theme_bw(base_size = 16) +
    theme(
        panel.grid = element_blank(),
        legend.position = "bottom",
        legend.title = element_blank(),
        panel.border = element_blank(),
        axis.text.y  = element_text(color = "black"),
        axis.line.x  = element_line(color = "black"),
        axis.line.y  = element_blank(),
        legend.justification = c("left", "top")
    ) +
    xlab("Standardized Beta Estimate, with QC covar") +
    ylab(""))
```

### Individual Scatterplots for SWYC
```{r}
swyc_matched_N <- sum(!is.na(df_matched.zscore$norm_devo_quotient) & !is.na(df_matched.zscore$global_VentricleChoroidVol))

#Ventricles
swyc_matched_vent_beta <- swyc.qc_reg_table_global %>% 
  filter(phenotype == "global_VentricleChoroidVol") %>% 
  select(swyc_beta) %>% pull(swyc_beta)
swyc_matched_vent_p_adj <- swyc.qc_reg_table_global %>% 
  filter(phenotype == "global_VentricleChoroidVol") %>% 
  select(swyc_p_adj) %>% pull(swyc_p_adj)
(swyc_matched_vent_plot <- ggplot(df_matched.zscore, aes(x = norm_devo_quotient, y = global_VentricleChoroidVol)) +
  geom_point(color = "#7293b0", alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE, color = "#7293b0") +
  annotate("label",
           x = 0.5, y = 3.2,
           label = paste0("std beta=", signif(swyc_matched_vent_beta,3), "; p-adj=", signif(swyc_matched_vent_p_adj,3))) +
  labs(
    x = paste0("Median SWYC Score (z-scored), N = ", swyc_matched_N),
    y = "Ventricle Volume Centile (z-scored)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    plot.title = element_blank()
  ))

#Subcortical GMV
swyc_matched_scgmv_beta <- swyc.qc_reg_table_global %>% 
  filter(phenotype == "global_SubCortGrayVol") %>% 
  select(swyc_beta) %>% pull(swyc_beta)
swyc_matched_scgmv_p_adj <- swyc.qc_reg_table_global %>% 
  filter(phenotype == "global_SubCortGrayVol") %>% 
  select(swyc_p_adj) %>% pull(swyc_p_adj)
(swyc_matched_scgmv_plot <- ggplot(df_matched.zscore, aes(x = norm_devo_quotient, y = global_SubCortGrayVol)) +
  geom_point(color = "#c87a7a", alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE, color = "#c87a7a") +
  annotate("label",
           x = 0.5, y = 3.2,
           label = paste0("std beta=", signif(swyc_matched_scgmv_beta,3), "; p-adj=", signif(swyc_matched_scgmv_p_adj,3))) +
  labs(
    x = paste0("Median SWYC Score (z-scored), N = ", swyc_matched_N),
    y = "Subcortical Gray Volume Centile (z-scored)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    plot.title = element_blank()
  ))


#Cortical GMV
swyc_matched_ctxgmv_beta <- swyc.qc_reg_table_global %>% 
  filter(phenotype == "global_CortexVol") %>% 
  select(swyc_beta) %>% pull(swyc_beta)
swyc_matched_ctxgmv_p_adj <- swyc.qc_reg_table_global %>% 
  filter(phenotype == "global_CortexVol") %>% 
  select(swyc_p_adj) %>% pull(swyc_p_adj)
(swyc_matched_ctxgmv_plot <- ggplot(df_matched.zscore, aes(x = norm_devo_quotient, y = global_CortexVol)) +
  geom_point(color = "#c87a7a", alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE, color = "#c87a7a") +
  annotate("label",
           x = 0.5, y = 3.2,
           label = paste0("std beta=", signif(swyc_matched_ctxgmv_beta,3), "; p-adj=", signif(swyc_matched_ctxgmv_p_adj,3))) +
  labs(
    x = paste0("Median SWYC ctxore (z-ctxored), N = ", swyc_matched_N),
    y = "Cortical Gray Volume Centile (z-ctxored)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    plot.title = element_blank()
  ))

#Total Surface Area
swyc_matched_sa_beta <- swyc.qc_reg_table_global %>% 
  filter(phenotype == "global_SurfaceArea") %>% 
  select(swyc_beta) %>% pull(swyc_beta)
swyc_matched_sa_p_adj <- swyc.qc_reg_table_global %>% 
  filter(phenotype == "global_SurfaceArea") %>% 
  select(swyc_p_adj) %>% pull(swyc_p_adj)
(swyc_matched_sa_plot <- ggplot(df_matched.zscore, aes(x = norm_devo_quotient, y = global_SurfaceArea)) +
  geom_point(color = "#c87a7a", alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE, color = "#c87a7a") +
  annotate("label",
           x = 0.5, y = 3.2,
           label = paste0("std beta=", signif(swyc_matched_sa_beta,3), "; p-adj=", signif(swyc_matched_sa_p_adj,3))) +
  labs(
    x = paste0("Median SWYC ctxore (z-scored), N = ", swyc_matched_N),
    y = "Total Surface Area Centile (z-scored)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    plot.title = element_blank()
  ))
```

## CMDs

### Calculate CMDs
```{r}
vol_vars_cortical <- vol_vars[grep("aparc", vol_vars)]
length_names <- length(c(vol_vars_cortical, sa_vars, th_vars))
names.global <- c("global_SubCortGrayVol","global_CortexVol", "global_CerebralWhiteMatterVol", "global_VentricleChoroidVol")
# Ensure all names are in the same order
names_base <- names(df_matched.zscore)[grep("aparc_GrayVol", names(df_matched.zscore))] %>% sub("aparc_GrayVol_","",.) %>%
  gsub("_Z", "", .)
names.vol <- paste0("aparc_GrayVol_",names_base)
names.sa <- paste0("aparc_SurfArea_",names_base)
names.th <- paste0("aparc_ThickAvg_",names_base)
# CMD Calculation
## vol
vol.center <- colMeans(df_matched.zscore %>% select(all_of(names.vol)))
vol.cov_matrix <- cov(df_matched.zscore %>% select(all_of(names.vol)))
vol.md <- sqrt(mahalanobis(df_matched.zscore %>% select(all_of(names.vol)), vol.center, vol.cov_matrix))
df_matched.zscore$vol.cmd <- vol.md
## sa
sa.center <- colMeans(df_matched.zscore %>% select(all_of(names.sa)), na.rm = TRUE)
sa.cov_matrix <- cov(df_matched.zscore %>% select(all_of(names.sa)), use = "complete.obs")
sa.md <- sqrt(mahalanobis(df_matched.zscore %>% select(all_of(names.sa)), sa.center, sa.cov_matrix))
df_matched.zscore$sa.cmd <- sa.md
## th
th.center <- colMeans(df_matched.zscore %>% select(all_of(names.th)), na.rm = TRUE)
th.cov_matrix <- cov(df_matched.zscore %>% select(all_of(names.th)), use = "complete.obs")
th.md <- sqrt(mahalanobis(df_matched.zscore %>% select(all_of(names.th)), th.center, th.cov_matrix))
df_matched.zscore$th.cmd <- th.md
## global
global.center <- colMeans(df_matched.zscore %>% select(all_of(names.global)), na.rm = TRUE)
global.cov_matrix <- cov(df_matched.zscore %>% select(all_of(names.global)), use = "complete.obs")
global.md <- sqrt(mahalanobis(df_matched.zscore %>% select(all_of(names.global)), global.center, global.cov_matrix))
df_matched.zscore$global.cmd <- global.md
```

### Cor.test CMDs
```{r, echo = FALSE}
# Exploratory
cor.test(df_matched.zscore$vol.cmd, df_matched.zscore$gestational_age)
cor.test(df_matched.zscore$vol.cmd, df_matched.zscore$birth_weight_kg)
cor.test(df_matched.zscore$vol.cmd, df_matched.zscore$norm_devo_quotient)

cor.test(df_matched.zscore$sa.cmd, df_matched.zscore$gestational_age)
cor.test(df_matched.zscore$sa.cmd, df_matched.zscore$birth_weight_kg)
cor.test(df_matched.zscore$sa.cmd, df_matched.zscore$norm_devo_quotient)

cor.test(df_matched.zscore$th.cmd, df_matched.zscore$gestational_age)
cor.test(df_matched.zscore$th.cmd, df_matched.zscore$birth_weight_kg)
cor.test(df_matched.zscore$th.cmd, df_matched.zscore$bwp_fen)
cor.test(df_matched.zscore$th.cmd, df_matched.zscore$norm_devo_quotient)

cor.test(df_matched.zscore$global.cmd, df_matched.zscore$gestational_age)
cor.test(df_matched.zscore$global.cmd, df_matched.zscore$birth_weight_kg)
cor.test(df_matched.zscore$global.cmd, df_matched.zscore$bwp_fen)
cor.test(df_matched.zscore$global.cmd, df_matched.zscore$norm_devo_quotient)
```

### CMD plots
```{r}
# Vol CMD
cor_volCMD_swyc.matched <- cor.test(df_matched.zscore$vol.cmd, df_matched.zscore$norm_devo_quotient)
## plot
volCMD_swyc.matched_r <- signif(cor_volCMD_swyc.matched$estimate, 3)
volCMD_swyc.matched_p <- signif(cor_volCMD_swyc.matched$p.value, 3)

(cor_volCMD_swyc.matched_plot <- ggplot(df_matched.zscore, aes(norm_devo_quotient, vol.cmd)) +
  geom_point(color = "#8C7FAF", alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE, color = "#8C7FAF") +
  annotate("label",
           x = 2, y = 10,
           label.size = 0,
           label = paste0("r=",volCMD_swyc.matched_r, "; p=", volCMD_swyc.matched_p)) +
  labs(
    x = paste0("SWYC DQ (age within 10% of scan), N = ", bw_N),
    y = "Volume CMD"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    plot.title = element_blank()
  ))

# SA CMD
cor_saCMD_swyc.matched <- cor.test(df_matched.zscore$sa.cmd, df_matched.zscore$norm_devo_quotient)
## plot
saCMD_swyc.matched_r <- signif(cor_saCMD_swyc.matched$estimate, 3)
saCMD_swyc.matched_p <- signif(cor_saCMD_swyc.matched$p.value, 3)

(cor_saCMD_swyc.matched_plot <- ggplot(df_matched.zscore, aes(norm_devo_quotient, sa.cmd)) +
  geom_point(color = "#8C7FAF", alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE, color = "#8C7FAF") +
  annotate("label",
           x = 2, y = 10,
           label.size = 0,
           label = paste0("r=",saCMD_swyc.matched_r, "; p=", saCMD_swyc.matched_p)) +
  labs(
    x = paste0("SWYC DQ (age within 10% of scan), N = ", bw_N),
    y = "Surface Area CMD"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    plot.title = element_blank()
  ))

# Thickness CMD
cor_thCMD_swyc.matched <- cor.test(df_matched.zscore$th.cmd, df_matched.zscore$norm_devo_quotient)
## plot
thCMD_swyc.matched_r <- signif(cor_thCMD_swyc.matched$estimate, 3)
thCMD_swyc.matched_p <- signif(cor_thCMD_swyc.matched$p.value, 3)

(cor_thCMD_swyc.matched_plot <- ggplot(df_matched.zscore, aes(norm_devo_quotient, th.cmd)) +
  geom_point(color = "#8C7FAF", alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE, color = "#8C7FAF") +
  annotate("label",
           x = 2, y = 10,
           label.size = 0,
           label = paste0("r=",thCMD_swyc.matched_r, "; p=", thCMD_swyc.matched_p)) +
  labs(
    x = paste0("SWYC DQ (age within 10% of scan), N = ", bw_N),
    y = "Cortical Thickness CMD"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    plot.title = element_blank()
  ))

#Global CMD
cor_globalCMD_swyc.matched <- cor.test(df_matched.zscore$global.cmd, df_matched.zscore$norm_devo_quotient)
## plot
globalCMD_swyc.matched_r <- signif(cor_globalCMD_swyc.matched$estimate, 3)
globalCMD_swyc.matched_p <- signif(cor_globalCMD_swyc.matched$p.value, 3)

(cor_globalCMD_swyc.matched_plot <- ggplot(df_matched.zscore, aes(norm_devo_quotient, global.cmd)) +
  geom_point(color = "#8C7FAF", alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE, color = "#8C7FAF") +
  annotate("label",
           x = 2, y = 10,
           label.size = 0,
           label = paste0("r=",globalCMD_swyc.matched_r, "; p=", globalCMD_swyc.matched_p)) +
  labs(
    x = paste0("SWYC DQ (age within 10% of scan), N = ", bw_N),
    y = "Global CMD"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    plot.title = element_blank()
  ))
```

# Regional Variables

## Set limits for ggseg visualization
```{r}
limits <- c(-0.3, 0.3)
low_color = "#4878CF"
mid_color = "white"
high_color = "#D65F5F"
```

## Volume

### simple linear regressions

#### compute regressions
```{r, echo = FALSE}
# Regional -- including cortical and subcortical
swyc_matched_vol <- as.data.frame(t(sapply(df_matched.zscore[, which(names(df_matched.zscore) %in% vol_vars)], function(x) {
  lm <- lm(x ~ norm_devo_quotient, data = df_matched.zscore)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2])
})))

swyc_matched_vol$swyc_p_adj <- p.adjust(swyc_matched_vol$swyc_p, method = "BH")
print(paste0("Number of significant IDPs after FDR correction: ", dim(swyc_matched_vol %>% filter(swyc_p_adj < 0.05))[1]))
```

#### add label column to match to ggseg
```{r, echo = FALSE}
cols_edit <- grepl("aseg(?!_CC)", rownames(swyc_matched_vol), perl = TRUE)

swyc_matched_vol$label <- rownames(swyc_matched_vol) %>% 
  sub("aparc_GrayVol_", "", .) %>% 
  sub("aseg_", "", .) %>%
  sub("X", "x", .) %>%
  sub("rd_Ventricle", "rd_ventricle", .) %>% 
  sub("th_Ventricle", "th_ventricle", .) %>% 
  sub("Thalamus", "Thalamus-Proper", .)

swyc_matched_vol$label[cols_edit] <- gsub("_", "-", swyc_matched_vol$label[cols_edit])

swyc_matched_vol_sig_MC <- swyc_matched_vol %>% filter(swyc_p_adj < 0.05)

print(paste0("Beta range of significant IDPs: ", range(swyc_matched_vol_sig_MC$swyc_beta)))
```

#### save betas
```{r, echo = FALSE}
save_table_full <- swyc_matched_vol
name <- "swyc_matched_median_volume.csv"
# Save dkt cortical betas for spin test
if(save_tables == TRUE){
  output_folder <-dkt_folder
  save_table_full$hemisphere <- ifelse(grepl("lh", save_table_full$label), "L", 
                         ifelse(grepl("rh", save_table_full$label), "R", NA))
  save_table_full$label <- sub("^rh_|lh_", "", save_table_full$label)
  filename_all <- paste0(output_folder, name)
  write.csv(save_table_full, filename_all)
}
# Save full betas
if(save_tables == TRUE){
output_folder <- betas_folder
filename <- paste0(output_folder, name)
write.csv(save_table_full, filename)
}
```

#### visualization of betas on surface maps
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(swyc_matched_vol_sig_MC, ggseg::dk, by = "label")
plot_data_aseg <- merge(swyc_matched_vol_sig_MC, ggseg::aseg, by = "label")

# Regions not matched in atlases
(swyc_matched_vol_noplot <- swyc_matched_vol_sig_MC[!(swyc_matched_vol_sig_MC$label %in% plot_data_dkt$label) & !(swyc_matched_vol_sig_MC$label %in% plot_data_aseg$label),])

# plots
(swyc_matched_dkt_vol_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = swyc_beta)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = limits) +
  theme_void()+
  labs(title = "SWYC Volume"))

(swyc_matched_aseg_vol_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = swyc_beta)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = limits) +
  theme_void() +
  labs(title = "SWYC Volume"))

if(save_figs == TRUE){
panel_save(swyc_matched_dkt_vol_plot, "swyc_matched_median-dkt-vol", "pdf", plot_output_folder)
panel_save(swyc_matched_aseg_vol_plot , "swyc_matched_median-aseg-vol", "pdf", plot_output_folder)
}
```

#### Scatterplot for top betas
```{r, echo = FALSE}
if (dim(swyc_matched_vol_sig_MC)[1] > 0){
top_n = if(dim(swyc_matched_vol_sig_MC)[1] > 10) {10} else {dim(swyc_matched_vol_sig_MC)[1]}
top <- swyc_matched_vol_sig_MC[order(swyc_matched_vol_sig_MC$swyc_beta, decreasing = TRUE)[1:top_n],]
scatterplots_swyc_matched_vol <- vector(mode = "list", length = top_n)
for(i in 1:top_n){
  pheno <- rownames(top)[i]
  title <- top$label[i]
  names(scatterplots_swyc_matched_vol[i]) <- pheno
  scatterplots_swyc_matched_vol[[i]] <- ggplot(df_matched.zscore, aes(x = norm_devo_quotient, y = !!sym(pheno))) +
    geom_point() +
    geom_smooth(method = "lm") +
    theme_minimal() +
    labs(x = "SWYC (z-scored)", y = "Centile (z-scored)", title = title)
}

nCol <- 2
nRow <- 2
breaks <- c(seq(from = 0, to = top_n , by = (nCol*nRow)), top_n) %>% unique(.)
for(k in 1:(length(breaks)-1)){
  start <- breaks[k]+1
  end <- breaks[k+1]
  do.call("grid.arrange", c(scatterplots_swyc_matched_vol[start:end], ncol=nCol))
}
}
```

### SWYC X age interaction

#### Regression and results
```{r, echo = FALSE}
swycbyage.qc_reg_table_vol <- as.data.frame(t(sapply(df_matched.zscore[, which(names(df_matched.zscore) %in% vol_vars)], function(x) {
  lm <- lm(x ~ df_matched.zscore$norm_devo_quotient*df_matched.zscore$adjusted_age_in_days + df_matched.zscore$median_euler_mean)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2],
   age_beta = summary$coefficients[3,1], age_p = summary$coefficients[3,4], age_se = summary$coefficients[3,2],
   swycbyage_beta = summary$coefficients[5,1], swycbyage_p = summary$coefficients[5,4], swycbyage_se = summary$coefficients[5,2])
})))

swycbyage.qc_reg_table_vol$swyc_p_adj <- p.adjust(swycbyage.qc_reg_table_vol$swyc_p, method = "BH")
swycbyage.qc_reg_table_vol$age_p_adj <- p.adjust(swycbyage.qc_reg_table_vol$age_p, method = "BH")
swycbyage.qc_reg_table_vol$swycbyage_p_adj <- p.adjust(swycbyage.qc_reg_table_vol$swycbyage_p, method = "BH")

print(paste0("Age Main significant results: ", dim(swycbyage.qc_reg_table_vol %>% filter(age_p_adj < 0.05))[1]))
print(paste0("Number of significant SWYC x Age IDPs - NO correction: ", dim(swycbyage.qc_reg_table_vol %>% filter(swycbyage_p < 0.05))[1]))
print(paste0("Number of significant SWYC x Age IDPs - AFTER FDR correction: ", dim(swycbyage.qc_reg_table_vol %>% filter(swycbyage_p_adj < 0.05))[1]))

range(swycbyage.qc_reg_table_vol %>% filter(swycbyage_p_adj < 0.05) %>% select(swycbyage_beta))

if(dim(swycbyage.qc_reg_table_vol %>% filter(age_p_adj < 0.05))[1] > 0){
  age_reg_table_vol <- as.data.frame(t(sapply(df_matched.zscore[, which(names(df_matched.zscore) %in% vol_vars)], function(x) {
  lm <- lm(x ~ df_matched.zscore$adjusted_age_in_days)
  summary <- summary(lm)
  c(age_beta = summary$coefficients[2,1], age_p = summary$coefficients[2,4], age_se = summary$coefficients[2,2])
})))
  age_reg_table_vol$age_p_adj <- p.adjust(age_reg_table_vol$age_p, method = "BH")
  print(paste0("Age Main significant results (no other covariates): ", dim(age_reg_table_vol %>% filter(age_p_adj < 0.05))[1]))
  
  
  full_age_reg_table_vol <- as.data.frame(t(sapply(df_matched.zscore_fullcentall[, which(names(df_matched.zscore_fullcentall) %in% vol_vars)], function(x) {
  lm <- lm(x ~ df_matched.zscore_fullcentall$adjusted_age_in_days)
  summary <- summary(lm)
  c(age_beta = summary$coefficients[2,1], age_p = summary$coefficients[2,4], age_se = summary$coefficients[2,2])
})))
  full_age_reg_table_vol$age_p_adj <- p.adjust(full_age_reg_table_vol$age_p, method = "BH")
  print(paste0("Age Main significant results full all sample (no other covariates): ", dim(full_age_reg_table_vol %>% filter(age_p_adj < 0.05))[1]))
}
```

## Surface Area

###  simple linear regressions

#### compute regressions
```{r, echo = FALSE}
sum(!is.na(df_matched.zscore$norm_devo_quotient))
# Regional -- including cortical and subcortical
swyc_matched_sa <- as.data.frame(t(sapply(df_matched.zscore[, which(names(df_matched.zscore) %in% sa_vars)], function(x) {
  lm <- lm(x ~ norm_devo_quotient, data = df_matched.zscore)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2])
})))

swyc_matched_sa$swyc_p_adj <- p.adjust(swyc_matched_sa$swyc_p, method = "BH")
print(paste0("Number of significant IDPs after FDR correction: ", dim(swyc_matched_sa %>% filter(swyc_p_adj < 0.05))[1]))
```

#### add label column to match to ggseg
```{r, echo = FALSE}
swyc_matched_sa$label <- rownames(swyc_matched_sa) %>% 
  sub("aparc_SurfArea_", "", .)

swyc_matched_sa_sig_MC <- swyc_matched_sa %>% filter(swyc_p_adj < 0.05)

print(paste0("Beta range of significant IDPs: ", range(swyc_matched_sa_sig_MC$swyc_beta)))
```

#### save betas
```{r, echo = FALSE}
save_table_full <- swyc_matched_sa
name <- "swyc_matched_median_surfArea.csv"
# Save dkt cortical betas for spin test
if(save_tables == TRUE){
  output_folder <-dkt_folder
  save_table_full$hemisphere <- ifelse(grepl("lh", save_table_full$label), "L", 
                         ifelse(grepl("rh", save_table_full$label), "R", NA))
  save_table_full$label <- sub("^rh_|lh_", "", save_table_full$label)
  filename_all <- paste0(output_folder, name)
  write.csv(save_table_full, filename_all)
}
# Save full betas
if(save_tables == TRUE){
output_folder <- betas_folder
filename <- paste0(output_folder, name)
write.csv(save_table_full, filename)
}
```

#### visualization of betas on surface maps
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(swyc_matched_sa_sig_MC, ggseg::dk, by = "label")

# Regions not matched in atlases
(swyc_matched_sa_noplot <- swyc_matched_sa_sig_MC[!(swyc_matched_sa_sig_MC$label %in% plot_data_dkt$label) & !(swyc_matched_sa_sig_MC$label %in% plot_data_aseg$label),])

# plots
(swyc_matched_dkt_sa_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = swyc_beta)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = limits) +
  theme_void()+
  labs(title = "SWYC Surface Area"))

if(save_figs == TRUE){
panel_save(swyc_matched_dkt_sa_plot, "swyc_matched_median-dkt-sa", "pdf", plot_output_folder)
}
```

#### Scatterplot for top betas
```{r, echo = FALSE}
if (dim(swyc_matched_sa_sig_MC)[1] > 0){
top_n = if(dim(swyc_matched_sa_sig_MC)[1] > 10) {10} else {dim(swyc_matched_sa_sig_MC)[1]}
top <- swyc_matched_sa_sig_MC[order(swyc_matched_sa_sig_MC$swyc_beta, decreasing = TRUE)[1:top_n],]
scatterplots_swyc_matched_sa <- vector(mode = "list", length = top_n)
for(i in 1:top_n){
  pheno <- rownames(top)[i]
  title <- top$label[i]
  names(scatterplots_swyc_matched_sa[i]) <- pheno
  scatterplots_swyc_matched_sa[[i]] <- ggplot(df_matched.zscore, aes(x = norm_devo_quotient, y = !!sym(pheno))) +
    geom_point() +
    geom_smooth(method = "lm") +
    theme_minimal() +
    labs(x = "SWYC (z-scored)", y = "Centile (z-scored)", title = title)
}

nCol <- 2
nRow <- 2
breaks <- c(seq(from = 0, to = top_n , by = (nCol*nRow)), top_n) %>% unique(.)
for(k in 1:(length(breaks)-1)){
  start <- breaks[k]+1
  end <- breaks[k+1]
  do.call("grid.arrange", c(scatterplots_swyc_matched_sa[start:end], ncol=nCol))
}
}
```

### SWYC X age interaction

#### Regression and results
```{r, echo = FALSE}
swycbyage.qc_reg_table_sa <- as.data.frame(t(sapply(df_matched.zscore[, which(names(df_matched.zscore) %in% sa_vars)], function(x) {
  lm <- lm(x ~ df_matched.zscore$norm_devo_quotient*df_matched.zscore$adjusted_age_in_days + df_matched.zscore$median_euler_mean)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2],
   age_beta = summary$coefficients[3,1], age_p = summary$coefficients[3,4], age_se = summary$coefficients[3,2],
   swycbyage_beta = summary$coefficients[5,1], swycbyage_p = summary$coefficients[5,4], swycbyage_se = summary$coefficients[5,2])
})))

swycbyage.qc_reg_table_sa$swyc_p_adj <- p.adjust(swycbyage.qc_reg_table_sa$swyc_p, method = "BH")
swycbyage.qc_reg_table_sa$age_p_adj <- p.adjust(swycbyage.qc_reg_table_sa$age_p, method = "BH")
swycbyage.qc_reg_table_sa$swycbyage_p_adj <- p.adjust(swycbyage.qc_reg_table_sa$swycbyage_p, method = "BH")

print(paste0("Age Main significant results: ", dim(swycbyage.qc_reg_table_sa %>% filter(age_p_adj < 0.05))[1]))
print(paste0("Number of significant SWYC x Age IDPs - NO correction: ", dim(swycbyage.qc_reg_table_sa %>% filter(swycbyage_p < 0.05))[1]))
print(paste0("Number of significant SWYC x Age IDPs - AFTER FDR correction: ", dim(swycbyage.qc_reg_table_sa %>% filter(swycbyage_p_adj < 0.05))[1]))

range(swycbyage.qc_reg_table_sa %>% filter(swycbyage_p_adj < 0.05) %>% select(swycbyage_beta))

if(dim(swycbyage.qc_reg_table_sa %>% filter(age_p_adj < 0.05))[1] > 0){
  age_reg_table_sa <- as.data.frame(t(sapply(df_matched.zscore[, which(names(df_matched.zscore) %in% sa_vars)], function(x) {
  lm <- lm(x ~ df_matched.zscore$adjusted_age_in_days)
  summary <- summary(lm)
  c(age_beta = summary$coefficients[2,1], age_p = summary$coefficients[2,4], age_se = summary$coefficients[2,2])
})))
  age_reg_table_sa$age_p_adj <- p.adjust(age_reg_table_sa$age_p, method = "BH")
  print(paste0("Age Main significant results (no other covariates): ", dim(age_reg_table_sa %>% filter(age_p_adj < 0.05))[1]))
  
  
  full_age_reg_table_sa <- as.data.frame(t(sapply(df_matched.zscore_fullcentall[, which(names(df_matched.zscore_fullcentall) %in% sa_vars)], function(x) {
  lm <- lm(x ~ df_matched.zscore_fullcentall$adjusted_age_in_days)
  summary <- summary(lm)
  c(age_beta = summary$coefficients[2,1], age_p = summary$coefficients[2,4], age_se = summary$coefficients[2,2])
})))
  full_age_reg_table_sa$age_p_adj <- p.adjust(full_age_reg_table_sa$age_p, method = "BH")
  print(paste0("Age Main significant results full all sample (no other covariates): ", dim(full_age_reg_table_sa %>% filter(age_p_adj < 0.05))[1]))
}
```

## Thickness

###  simple linear regressions

#### compute regressions
```{r, echo = FALSE}
sum(!is.na(df_matched.zscore$aparc_ThickAvg_lh_bankssts))
# Regional -- including cortical and subcortical
swyc_matched_th <- as.data.frame(t(sapply(df_matched.zscore[, which(names(df_matched.zscore) %in% th_vars)], function(x) {
  lm <- lm(x ~ norm_devo_quotient, data = df_matched.zscore)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2])
})))

swyc_matched_th$swyc_p_adj <- p.adjust(swyc_matched_th$swyc_p, method = "BH")
print(paste0("Number of significant IDPs after FDR correction: ", dim(swyc_matched_th %>% filter(swyc_p_adj < 0.05))[1]))
```

#### add label column to match to ggseg
```{r, echo = FALSE}
swyc_matched_th$label <- rownames(swyc_matched_th) %>% 
  sub("aparc_ThickAvg_", "", .)

swyc_matched_th_sig_MC <- swyc_matched_th %>% filter(swyc_p_adj < 0.05)

print(paste0("Beta range of significant IDPs: ", range(swyc_matched_th_sig_MC$swyc_beta)))
```

#### save betas
```{r, echo = FALSE}
save_table_full <- swyc_matched_th
name <- "swyc_matched_thick.csv"
# Save dkt cortical betas for spin test
if(save_tables == TRUE){
  output_folder <-dkt_folder
  save_table_full$hemisphere <- ifelse(grepl("lh", save_table_full$label), "L", 
                         ifelse(grepl("rh", save_table_full$label), "R", NA))
  save_table_full$label <- sub("^rh_|lh_", "", save_table_full$label)
  filename_all <- paste0(output_folder, name)
  write.csv(save_table_full, filename_all)
}
# Save full betas
if(save_tables == TRUE){
output_folder <- betas_folder
filename <- paste0(output_folder, name)
write.csv(save_table_full, filename)
}
```

#### visualization of betas on surface maps
```{r, echo = FALSE}
# merge with atlas data
plot_data_dkt <- merge(swyc_matched_th_sig_MC, ggseg::dk, by = "label")

# Regions not matched in atlases
(swyc_matched_th_noplot <- swyc_matched_th_sig_MC[!(swyc_matched_th_sig_MC$label %in% plot_data_dkt$label) & !(swyc_matched_th_sig_MC$label %in% plot_data_aseg$label),])

# plots
(swyc_matched_dkt_th_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = swyc_beta)) +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, limits = limits) +
  theme_void()+
  labs(title = "SWYC Cortical Thickness"))

if(save_figs == TRUE){
panel_save(swyc_matched_dkt_th_plot, "swyc_matched_median-dkt-th", "pdf", plot_output_folder)
}
```

#### Scatterplot for top betas
```{r, echo = FALSE}
if (dim(swyc_matched_th_sig_MC)[1] > 0){
top_n = if(dim(swyc_matched_th_sig_MC)[1] > 10) {10} else {dim(swyc_matched_th_sig_MC)[1]}
top <- swyc_matched_th_sig_MC[order(abs(swyc_matched_th_sig_MC$swyc_beta), decreasing = TRUE)[1:top_n],]
scatterplots_swyc_matched_th <- vector(mode = "list", length = top_n)
for(i in 1:top_n){
  print(i)
  pheno <- rownames(top)[i]
  print(pheno)
  title <- top$label[i]
  names(scatterplots_swyc_matched_th)[i] <- pheno
  scatterplots_swyc_matched_th[[i]] <- ggplot(df_matched.zscore, aes(x = norm_devo_quotient, y = !!sym(pheno))) +
    geom_point() +
    geom_smooth(method = "lm") +
    theme_minimal() +
    labs(x = "SWYC (z-scored)", y = "Centile (z-scored)", title = title)
}

nCol <- 2
nRow <- 2
breaks <- c(seq(from = 0, to = top_n , by = (nCol*nRow)), top_n) %>% unique(.)
for(k in 1:(length(breaks)-1)){
  start <- breaks[k]+1
  end <- breaks[k+1]
  do.call("grid.arrange", c(scatterplots_swyc_matched_th[start:end], ncol=nCol))
}
}
```

### SWYC X age interaction

#### Regression and results
```{r, echo = FALSE}
swycbyage.qc_reg_table_th <- as.data.frame(t(sapply(df_matched.zscore[, which(names(df_matched.zscore) %in% th_vars)], function(x) {
  lm <- lm(x ~ df_matched.zscore$norm_devo_quotient*df_matched.zscore$adjusted_age_in_days + df_matched.zscore$median_euler_mean)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2],
   age_beta = summary$coefficients[3,1], age_p = summary$coefficients[3,4], age_se = summary$coefficients[3,2],
   swycbyage_beta = summary$coefficients[5,1], swycbyage_p = summary$coefficients[5,4], swycbyage_se = summary$coefficients[5,2])
})))

swycbyage.qc_reg_table_th$swyc_p_adj <- p.adjust(swycbyage.qc_reg_table_th$swyc_p, method = "BH")
swycbyage.qc_reg_table_th$age_p_adj <- p.adjust(swycbyage.qc_reg_table_th$age_p, method = "BH")
swycbyage.qc_reg_table_th$swycbyage_p_adj <- p.adjust(swycbyage.qc_reg_table_th$swycbyage_p, method = "BH")

print(paste0("Age Main significant results: ", dim(swycbyage.qc_reg_table_th %>% filter(age_p_adj < 0.05))[1]))
print(paste0("Number of significant SWYC x Age IDPs - NO correction: ", dim(swycbyage.qc_reg_table_th %>% filter(swycbyage_p < 0.05))[1]))
print(paste0("Number of significant SWYC x Age IDPs - AFTER FDR correction: ", dim(swycbyage.qc_reg_table_th %>% filter(swycbyage_p_adj < 0.05))[1]))

range(swycbyage.qc_reg_table_th %>% filter(swycbyage_p_adj < 0.05) %>% select(swycbyage_beta))
```

## Test Mediation

### Brain covars with bw
```{r}
sum(!is.na(df_matched.zscore$birth_weight_kg))

lm_swyc_matched_bw_matched <- lm(norm_devo_quotient ~ birth_weight_kg, data = df_matched.zscore)
lm_swyc_matched_bw_matched_stats <- summary(lm_swyc_matched_bw_matched)
lm_swyc_matched_bw_thcmd <- lm(norm_devo_quotient ~ birth_weight_kg + th.cmd, data = df_matched.zscore)
lm_swyc_matched_bw_ctxvol <- lm(norm_devo_quotient ~ birth_weight_kg + global_CortexVol, data = df_matched.zscore)
lm_swyc_matched_bw_thcmd_stats <- summary(lm_swyc_matched_bw_thcmd)
lm_swyc_matched_bw_ctxvol_stats <- summary(lm_swyc_matched_bw_ctxvol)
lm_swyc_matched_bw_matched_stats
lm_swyc_matched_bw_thcmd_stats
lm_swyc_matched_bw_ctxvol_stats

#plot

df_plot_bw <- data.frame(
  model = c("no covar BW", "thickness cmd covar", "ctxvol covar"),
  estimate = c(summary(lm_swyc_matched_bw_matched)$coefficients[2, "Estimate"],
               summary(lm_swyc_matched_bw_thcmd)$coefficients[2, "Estimate"],
               summary(lm_swyc_matched_bw_ctxvol)$coefficients[2, "Estimate"]),
  se = c(summary(lm_swyc_matched_bw_matched)$coefficients[2, "Std. Error"],
         summary(lm_swyc_matched_bw_thcmd)$coefficients[2, "Std. Error"],
         summary(lm_swyc_matched_bw_ctxvol)$coefficients[2, "Std. Error"])
) %>%
  mutate(
    ymin = estimate - se,
    ymax = estimate + se
  )

ggplot(df_plot_bw, aes(x = model, y = estimate)) +
  geom_pointrange(aes(ymin = ymin, ymax = ymax), linewidth = 0.8) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_classic()

```

### Brain covars with bw
```{r}
sum(!is.na(df_matched.zscore$gestational_age))

lm_swyc_matched_ga_matched <- lm(norm_devo_quotient ~ gestational_age, data = df_matched.zscore)
lm_swyc_matched_ga_matched_stats <- summary(lm_swyc_matched_ga_matched)
lm_swyc_matched_ga_thcmd <- lm(norm_devo_quotient ~ gestational_age + th.cmd, data = df_matched.zscore)
lm_swyc_matched_ga_ctxvol <- lm(norm_devo_quotient ~ gestational_age + global_CortexVol, data = df_matched.zscore)
lm_swyc_matched_ga_thcmd_stats <- summary(lm_swyc_matched_ga_thcmd)
lm_swyc_matched_ga_ctxvol_stats <- summary(lm_swyc_matched_ga_ctxvol)
lm_swyc_matched_ga_matched_stats
lm_swyc_matched_ga_thcmd_stats
lm_swyc_matched_ga_ctxvol_stats

#plot

df_plot_ga <- data.frame(
  model = c("no covar ga", "thickness cmd covar", "ctxvol covar"),
  estimate = c(summary(lm_swyc_matched_ga_matched)$coefficients[2, "Estimate"],
               summary(lm_swyc_matched_ga_thcmd)$coefficients[2, "Estimate"],
               summary(lm_swyc_matched_ga_ctxvol)$coefficients[2, "Estimate"]),
  se = c(summary(lm_swyc_matched_ga_matched)$coefficients[2, "Std. Error"],
         summary(lm_swyc_matched_ga_thcmd)$coefficients[2, "Std. Error"],
         summary(lm_swyc_matched_ga_ctxvol)$coefficients[2, "Std. Error"])
) %>%
  mutate(
    ymin = estimate - se,
    ymax = estimate + se
  )

ggplot(df_plot_ga, aes(x = model, y = estimate)) +
  geom_pointrange(aes(ymin = ymin, ymax = ymax), linewidth = 0.8) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_classic()

```


# LMSz Centiles

## Prep SWYC-LMSz centiles data

### merge and take z-scores of centiles
```{r}
lmsz_centiles$participant_id_modified <- sub("sub-", "", lmsz_centiles$participant_id)

swyc_lmsz <- merge(swyc_subset_avg %>% select(all_of(c("pat_id", "mean_dq", "median_dq", "preterm",
                                                   "birth_weight_kg","gestational_age", "bwp_fen",
                                                   "sex", "adjusted_age_in_days", "median_euler_mean"))),
                   lmsz_centiles, by.x = c("pat_id", "sex"), by.y = c("participant_id_modified", "sex"), all.x = FALSE, all.y = FALSE)
#swyc_lmsz <- merge(swyc_lmsz, swyc_avg_centiles %>% select(pat_id, grep("cmd", names(.))), 
                   #by = "pat_id", all.x = TRUE, all.y = FALSE)

table(swyc_lmsz$preterm)
# zscore centiles and other variables of interest
lmsz_vars <- names(swyc_lmsz)[grep(".lmsz", names(swyc_lmsz))]
swyc_avg_centiles_lmsz <- swyc_lmsz %>%
  mutate(across(all_of(c(full_vars, lmsz_vars, "bwp_fen")), qnorm)) %>%
  mutate(across(all_of(c("gestational_age", "birth_weight_kg", 
                         #"adjusted_age_in_days", 
                         "age_at_scan",
                         "median_euler_mean", "mean_dq", "median_dq")), scale))
# set sex and preterm as factor
swyc_avg_centiles_lmsz$sex <- as.factor(swyc_avg_centiles_lmsz$sex)
swyc_avg_centiles_lmsz$preterm <- as.factor(swyc_avg_centiles_lmsz$preterm)
```

### find Inf cells, make a table of N per phenotype, then convert to NaN
```{r, echo = FALSE}
inf_nan_table <- as.data.frame(t(sapply(swyc_avg_centiles_lmsz[c(full_vars, "bwp_fen")], function(x){
  c(inf = sum(is.infinite(x)), nan = sum(is.nan(x)))
})))
max(inf_nan_table$inf)
max(inf_nan_table$nan)

inf_nan_table %>% filter(inf > 0)
```

## Global Variables

### Run regressions with original and lmsz centiles
```{r}
swyc_orig_reg_table_global <- as.data.frame(t(sapply(swyc_avg_centiles_lmsz[, which(names(swyc_avg_centiles_lmsz) %in% global_vars_totest)], function(x) {
  lm <- lm(x ~ swyc_avg_centiles_lmsz$median_dq)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2])
})))

swyc_orig_reg_table_global$swyc_p_adj <- p.adjust(swyc_orig_reg_table_global$swyc_p, method = "BH")

global_vars_totest_lmsz <- paste0(global_vars_totest, ".lmsz")

swyc_lmsz_reg_table_global <- as.data.frame(t(sapply(swyc_avg_centiles_lmsz[, which(names(swyc_avg_centiles_lmsz) %in% global_vars_totest_lmsz)], function(x) {
  lm <- lm(x ~ swyc_avg_centiles_lmsz$median_dq)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2])
})))

swyc_lmsz_reg_table_global$swyc_p_adj <- p.adjust(swyc_lmsz_reg_table_global$swyc_p, method = "BH")
```

### Calculate SE bars and combine into one df
```{r}
#orig centiles
swyc_orig_reg_table_global$swyc_lower <- (swyc_orig_reg_table_global$swyc_beta - 1.96*swyc_orig_reg_table_global$swyc_se)
swyc_orig_reg_table_global$swyc_upper <- (swyc_orig_reg_table_global$swyc_beta + 1.96*swyc_orig_reg_table_global$swyc_se)
swyc_orig_reg_table_global <- swyc_orig_reg_table_global %>% mutate(type = "orig")

#lmsz centiles
swyc_lmsz_reg_table_global$swyc_lower <- (swyc_lmsz_reg_table_global$swyc_beta - 1.96*swyc_lmsz_reg_table_global$swyc_se)
swyc_lmsz_reg_table_global$swyc_upper <- (swyc_lmsz_reg_table_global$swyc_beta + 1.96*swyc_lmsz_reg_table_global$swyc_se)
swyc_lmsz_reg_table_global <- swyc_lmsz_reg_table_global %>% mutate(type = "lmsz")

#combine tables
comb_swyc_lmsz_reg_table_global <- rbind(swyc_orig_reg_table_global, swyc_lmsz_reg_table_global)
comb_swyc_lmsz_reg_table_global$type <- as.factor(comb_swyc_lmsz_reg_table_global$type)
comb_swyc_lmsz_reg_table_global$label <- rownames(comb_swyc_lmsz_reg_table_global) %>% sub(".lmsz", "", .)
```

### Plot betas from both models with SE bars
```{r}
ggplot(comb_swyc_lmsz_reg_table_global, aes(y = label, x = swyc_beta, color = type)) +
  geom_point(size = 2, position = position_dodge(width = 0.3)) +
  geom_errorbar(aes(xmin = swyc_upper, xmax = swyc_lower),
                width = 0,
                position = position_dodge(width = 0.3)) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 15, margin = margin(r = 10)),  # space labels
        axis.title.y = element_text(margin = margin(r = 15)),             # space title
        panel.grid.major.x = element_line(), # x major grids visible
        panel.grid.major.y = element_blank(), # y grids gone
        panel.grid.minor.x = element_blank() ) + # x minor grids gone
  scale_y_discrete(expand = expansion(mult = c(0.1, 0.1))) +
  scale_color_manual(values = c("orig" = "grey", "lmsz" = "#B56576")) +
  labs(x = "Beta Estimate", y = "Phenotype", color = "Centile Model") + 
  geom_vline(xintercept = 0, linetype = "dashed", color = "black")
```


## Volume Variables

### Run regressions with original and lmsz centiles
```{r}
swyc_orig_reg_table_vol <- as.data.frame(t(sapply(swyc_avg_centiles_lmsz[, which(names(swyc_avg_centiles_lmsz) %in% vol_vars)], function(x) {
  lm <- lm(x ~ swyc_avg_centiles_lmsz$median_dq)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2])
})))

swyc_orig_reg_table_vol$swyc_p_adj <- p.adjust(swyc_orig_reg_table_vol$swyc_p, method = "BH")

vol_vars_lmsz <- paste0(vol_vars, ".lmsz")

swyc_lmsz_reg_table_vol <- as.data.frame(t(sapply(swyc_avg_centiles_lmsz[, which(names(swyc_avg_centiles_lmsz) %in% vol_vars_lmsz)], function(x) {
  lm <- lm(x ~ swyc_avg_centiles_lmsz$median_dq)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2])
})))

swyc_lmsz_reg_table_vol$swyc_p_adj <- p.adjust(swyc_lmsz_reg_table_vol$swyc_p, method = "BH")
```

### Calculate SE bars and combine into one df
```{r}
#orig centiles
swyc_orig_reg_table_vol$swyc_lower <- (swyc_orig_reg_table_vol$swyc_beta - 1.96*swyc_orig_reg_table_vol$swyc_se)
swyc_orig_reg_table_vol$swyc_upper <- (swyc_orig_reg_table_vol$swyc_beta + 1.96*swyc_orig_reg_table_vol$swyc_se)
swyc_orig_reg_table_vol <- swyc_orig_reg_table_vol %>% mutate(type = "orig")

#lmsz centiles
swyc_lmsz_reg_table_vol$swyc_lower <- (swyc_lmsz_reg_table_vol$swyc_beta - 1.96*swyc_lmsz_reg_table_vol$swyc_se)
swyc_lmsz_reg_table_vol$swyc_upper <- (swyc_lmsz_reg_table_vol$swyc_beta + 1.96*swyc_lmsz_reg_table_vol$swyc_se)
swyc_lmsz_reg_table_vol <- swyc_lmsz_reg_table_vol %>% mutate(type = "lmsz")

#combine tables
comb_swyc_lmsz_reg_table_vol <- rbind(swyc_orig_reg_table_vol, swyc_lmsz_reg_table_vol)
comb_swyc_lmsz_reg_table_vol$type <- as.factor(comb_swyc_lmsz_reg_table_vol$type)
comb_swyc_lmsz_reg_table_vol$label <- rownames(comb_swyc_lmsz_reg_table_vol) %>% sub(".lmsz", "", .)
```

### Plot betas from both models with SE bars
```{r}
# Right Hemisphere
ggplot(comb_swyc_lmsz_reg_table_vol %>% filter(grepl("aparc", label) & grepl("rh", label)), aes(y = sub("aparc_GrayVol_rh_", "", label), x = swyc_beta, color = type)) +
  geom_point(size = 2, position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(xmin = swyc_upper, xmax = swyc_lower),
                width = 0,
                position = position_dodge(width = 0.5)) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 10, margin = margin(r = 10)),  # space labels
        axis.title.y = element_text(margin = margin(r = 15)),             # space title
        panel.grid.major.x = element_line(), # x major grids visible
        panel.grid.major.y = element_blank(), # y grids gone
        panel.grid.minor.x = element_blank() ) + # x minor grids gone
  scale_y_discrete(expand = expansion(mult = c(0.1, 0.1))) +
  scale_color_manual(values = c("orig" = "grey", "lmsz" = "#B56576")) +
  labs(x = "Beta Estimate", y = "Right Cortical Vol", color = "Centile Model") + 
  geom_vline(xintercept = 0, linetype = "dashed", color = "black")

#Left Hemisphere
ggplot(comb_swyc_lmsz_reg_table_vol %>% filter(grepl("aparc", label) & grepl("lh", label)), aes(y = sub("aparc_GrayVol_lh_", "", label), x = swyc_beta, color = type)) +
  geom_point(size = 2, position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(xmin = swyc_upper, xmax = swyc_lower),
                width = 0,
                position = position_dodge(width = 0.5)) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 10, margin = margin(r = 10)),  # space labels
        axis.title.y = element_text(margin = margin(r = 15)),             # space title
        panel.grid.major.x = element_line(), # x major grids visible
        panel.grid.major.y = element_blank(), # y grids gone
        panel.grid.minor.x = element_blank() ) + # x minor grids gone
  scale_y_discrete(expand = expansion(mult = c(0.1, 0.1))) +
  scale_color_manual(values = c("orig" = "grey", "lmsz" = "#B56576")) +
  labs(x = "Beta Estimate", y = "Left Cortical Vol", color = "Centile Model") + 
  geom_vline(xintercept = 0, linetype = "dashed", color = "black")

#Subcortical
ggplot(comb_swyc_lmsz_reg_table_vol %>% filter(grepl("aseg", label)), aes(y = sub("aseg_", "", label), x = swyc_beta, color = type)) +
  geom_point(size = 2, position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(xmin = swyc_upper, xmax = swyc_lower),
                width = 0,
                position = position_dodge(width = 0.5)) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 10, margin = margin(r = 10)),  # space labels
        axis.title.y = element_text(margin = margin(r = 15)),             # space title
        panel.grid.major.x = element_line(), # x major grids visible
        panel.grid.major.y = element_blank(), # y grids gone
        panel.grid.minor.x = element_blank() ) + # x minor grids gone
  scale_y_discrete(expand = expansion(mult = c(0.1, 0.1))) +
  scale_color_manual(values = c("orig" = "grey", "lmsz" = "#B56576")) +
  labs(x = "Beta Estimate", y = "Subcortical Vol", color = "Centile Model") + 
  geom_vline(xintercept = 0, linetype = "dashed", color = "black")


```

## Surface Area Variables

### Run regressions with original and lmsz centiles
```{r}
swyc_orig_reg_table_sa <- as.data.frame(t(sapply(swyc_avg_centiles_lmsz[, which(names(swyc_avg_centiles_lmsz) %in% sa_vars)], function(x) {
  lm <- lm(x ~ swyc_avg_centiles_lmsz$median_dq)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2])
})))

swyc_orig_reg_table_sa$swyc_p_adj <- p.adjust(swyc_orig_reg_table_sa$swyc_p, method = "BH")

sa_vars_lmsz <- paste0(sa_vars, ".lmsz")

swyc_lmsz_reg_table_sa <- as.data.frame(t(sapply(swyc_avg_centiles_lmsz[, which(names(swyc_avg_centiles_lmsz) %in% sa_vars_lmsz)], function(x) {
  lm <- lm(x ~ swyc_avg_centiles_lmsz$median_dq)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2])
})))

swyc_lmsz_reg_table_sa$swyc_p_adj <- p.adjust(swyc_lmsz_reg_table_sa$swyc_p, method = "BH")
```

### Calculate SE bars and combine into one df
```{r}
#orig centiles
swyc_orig_reg_table_sa$swyc_lower <- (swyc_orig_reg_table_sa$swyc_beta - 1.96*swyc_orig_reg_table_sa$swyc_se)
swyc_orig_reg_table_sa$swyc_upper <- (swyc_orig_reg_table_sa$swyc_beta + 1.96*swyc_orig_reg_table_sa$swyc_se)
swyc_orig_reg_table_sa <- swyc_orig_reg_table_sa %>% mutate(type = "orig")

#lmsz centiles
swyc_lmsz_reg_table_sa$swyc_lower <- (swyc_lmsz_reg_table_sa$swyc_beta - 1.96*swyc_lmsz_reg_table_sa$swyc_se)
swyc_lmsz_reg_table_sa$swyc_upper <- (swyc_lmsz_reg_table_sa$swyc_beta + 1.96*swyc_lmsz_reg_table_sa$swyc_se)
swyc_lmsz_reg_table_sa <- swyc_lmsz_reg_table_sa %>% mutate(type = "lmsz")

#combine tables
comb_swyc_lmsz_reg_table_sa <- rbind(swyc_orig_reg_table_sa, swyc_lmsz_reg_table_sa)
comb_swyc_lmsz_reg_table_sa$type <- as.factor(comb_swyc_lmsz_reg_table_sa$type)
comb_swyc_lmsz_reg_table_sa$label <- rownames(comb_swyc_lmsz_reg_table_sa) %>% sub(".lmsz", "", .)
```

### Plot betas from both models with SE bars
```{r}
# Right Hemisphere
ggplot(comb_swyc_lmsz_reg_table_sa %>% filter(grepl("rh", label)), aes(y = sub("aparc_SurfArea_rh_", "", label), x = swyc_beta, color = type)) +
  geom_point(size = 2, position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(xmin = swyc_upper, xmax = swyc_lower),
                width = 0,
                position = position_dodge(width = 0.5)) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 10, margin = margin(r = 10)),  # space labels
        axis.title.y = element_text(margin = margin(r = 15)),             # space title
        panel.grid.major.x = element_line(), # x major grids visible
        panel.grid.major.y = element_blank(), # y grids gone
        panel.grid.minor.x = element_blank() ) + # x minor grids gone
  scale_y_discrete(expand = expansion(mult = c(0.1, 0.1))) +
  scale_color_manual(values = c("orig" = "grey", "lmsz" = "#B56576")) +
  labs(x = "Beta Estimate", y = "Right Cortical SA", color = "Centile Model") + 
  geom_vline(xintercept = 0, linetype = "dashed", color = "black")

#Left Hemisphere
ggplot(comb_swyc_lmsz_reg_table_sa %>% filter(grepl("lh", label)), aes(y = sub("aparc_SurfArea_lh_", "", label), x = swyc_beta, color = type)) +
  geom_point(size = 2, position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(xmin = swyc_upper, xmax = swyc_lower),
                width = 0,
                position = position_dodge(width = 0.5)) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 10, margin = margin(r = 10)),  # space labels
        axis.title.y = element_text(margin = margin(r = 15)),             # space title
        panel.grid.major.x = element_line(), # x major grids visible
        panel.grid.major.y = element_blank(), # y grids gone
        panel.grid.minor.x = element_blank() ) + # x minor grids gone
  scale_y_discrete(expand = expansion(mult = c(0.1, 0.1))) +
  scale_color_manual(values = c("orig" = "grey", "lmsz" = "#B56576")) +
  labs(x = "Beta Estimate", y = "Left Cortical SA", color = "Centile Model") + 
  geom_vline(xintercept = 0, linetype = "dashed", color = "black")

```

## Cortical Thickness Variables

### Run regressions with original and lmsz centiles
```{r}
swyc_orig_reg_table_th <- as.data.frame(t(sapply(swyc_avg_centiles_lmsz[, which(names(swyc_avg_centiles_lmsz) %in% th_vars)], function(x) {
  lm <- lm(x ~ swyc_avg_centiles_lmsz$median_dq)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2])
})))

swyc_orig_reg_table_th$swyc_p_adj <- p.adjust(swyc_orig_reg_table_th$swyc_p, method = "BH")

th_vars_lmsz <- paste0(th_vars, ".lmsz")

swyc_lmsz_reg_table_th <- as.data.frame(t(sapply(swyc_avg_centiles_lmsz[, which(names(swyc_avg_centiles_lmsz) %in% th_vars_lmsz)], function(x) {
  lm <- lm(x ~ swyc_avg_centiles_lmsz$median_dq)
  summary <- summary(lm)
  c(swyc_beta = summary$coefficients[2,1], swyc_p = summary$coefficients[2,4], swyc_se = summary$coefficients[2,2])
})))

swyc_lmsz_reg_table_th$swyc_p_adj <- p.adjust(swyc_lmsz_reg_table_th$swyc_p, method = "BH")
```

### Calculate SE bars and combine into one df
```{r}
#orig centiles
swyc_orig_reg_table_th$swyc_lower <- (swyc_orig_reg_table_th$swyc_beta - 1.96*swyc_orig_reg_table_th$swyc_se)
swyc_orig_reg_table_th$swyc_upper <- (swyc_orig_reg_table_th$swyc_beta + 1.96*swyc_orig_reg_table_th$swyc_se)
swyc_orig_reg_table_th <- swyc_orig_reg_table_th %>% mutate(type = "orig")

#lmsz centiles
swyc_lmsz_reg_table_th$swyc_lower <- (swyc_lmsz_reg_table_th$swyc_beta - 1.96*swyc_lmsz_reg_table_th$swyc_se)
swyc_lmsz_reg_table_th$swyc_upper <- (swyc_lmsz_reg_table_th$swyc_beta + 1.96*swyc_lmsz_reg_table_th$swyc_se)
swyc_lmsz_reg_table_th <- swyc_lmsz_reg_table_th %>% mutate(type = "lmsz")

#combine tables
comb_swyc_lmsz_reg_table_th <- rbind(swyc_orig_reg_table_th, swyc_lmsz_reg_table_th)
comb_swyc_lmsz_reg_table_th$type <- as.factor(comb_swyc_lmsz_reg_table_th$type)
comb_swyc_lmsz_reg_table_th$label <- rownames(comb_swyc_lmsz_reg_table_th) %>% sub(".lmsz", "", .)
```

### Plot betas from both models with SE bars
```{r}
# Right Hemisphere
ggplot(comb_swyc_lmsz_reg_table_th %>% filter(grepl("rh", label)), aes(y = sub("aparc_ThickAvg_rh_", "", label), x = swyc_beta, color = type)) +
  geom_point(size = 2, position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(xmin = swyc_upper, xmax = swyc_lower),
                width = 0,
                position = position_dodge(width = 0.5)) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 10, margin = margin(r = 10)),  # space labels
        axis.title.y = element_text(margin = margin(r = 15)),             # space title
        panel.grid.major.x = element_line(), # x major grids visible
        panel.grid.major.y = element_blank(), # y grids gone
        panel.grid.minor.x = element_blank() ) + # x minor grids gone
  scale_y_discrete(expand = expansion(mult = c(0.1, 0.1))) +
  scale_color_manual(values = c("orig" = "grey", "lmsz" = "#B56576")) +
  labs(x = "Beta Estimate", y = "Right Cortical Thick", color = "Centile Model") + 
  geom_vline(xintercept = 0, linetype = "dashed", color = "black")

#Left Hemisphere
ggplot(comb_swyc_lmsz_reg_table_th %>% filter(grepl("lh", label)), aes(y = sub("aparc_ThickAvg_lh_", "", label), x = swyc_beta, color = type)) +
  geom_point(size = 2, position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(xmin = swyc_upper, xmax = swyc_lower),
                width = 0,
                position = position_dodge(width = 0.5)) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 10, margin = margin(r = 10)),  # space labels
        axis.title.y = element_text(margin = margin(r = 15)),             # space title
        panel.grid.major.x = element_line(), # x major grids visible
        panel.grid.major.y = element_blank(), # y grids gone
        panel.grid.minor.x = element_blank() ) + # x minor grids gone
  scale_y_discrete(expand = expansion(mult = c(0.1, 0.1))) +
  scale_color_manual(values = c("orig" = "grey", "lmsz" = "#B56576")) +
  labs(x = "Beta Estimate", y = "Left Cortical Thick", color = "Centile Model") + 
  geom_vline(xintercept = 0, linetype = "dashed", color = "black")
```



<!-- # Remove BW outliers, and corresponding BWP -->
<!-- ```{r} -->
<!-- mu_bw <- mean(centiles_all %>% distinct(participant_id, .keep_all = TRUE) %>% select(birth_weight_kg) %>% pull(.), na.rm = TRUE) -->
<!-- sd_bw <- sd(centiles_all %>% distinct(participant_id, .keep_all = TRUE) %>% select(birth_weight_kg) %>% pull(.), na.rm = TRUE) -->

<!-- rm_dist = 5 -->

<!-- nrow(centiles_all %>% filter( -->
<!--   birth_weight_kg < mu_bw-sd_bw*rm_dist | birth_weight_kg > mu_bw+sd_bw*rm_dist) ) -->

<!-- centiles_all <- centiles_all %>% -->
<!--   mutate(birth_weight_kg_inc_out = birth_weight_kg, -->
<!--     birth_weight_kg = case_when( -->
<!--         (birth_weight_kg < mu_bw-sd_bw*rm_dist | birth_weight_kg > mu_bw+sd_bw*rm_dist) ~ NA, -->
<!--         TRUE ~ birth_weight_kg), -->
<!--     bwp_fen_inc_out = bwp_fen, -->
<!--     bwp_fen = case_when( -->
<!--         (birth_weight_kg < mu_bw-sd_bw*rm_dist | birth_weight_kg > mu_bw+sd_bw*rm_dist) ~ NA, -->
<!--         TRUE ~ bwp_fen)) -->
<!-- ``` -->
