---
title: "vars-interest-distributions"
author: "Eren Kafadar"
date: "2025-11-18"
output: html_document
editor_options: 
  chunk_output_type: console
params:
  output_folder : "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/vars_dist/"
  centiles_distinct_path: "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/data_clean/csv/centiles_distinct_median_2026-04-24.csv"
  centiles_all_path: "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/data_clean/csv/centiles_all_median_2026-04-24.csv"
  type: "median"
---

# Define params if they do not exist
```{r, echo = FALSE}
if (!exists("params")) {
  params <- list(
  output_folder = "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/vars_dist/",
  centiles_distinct_path = "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/data_clean/csv/centiles_distinct_median_2026-04-24.csv",
  centiles_all_path = "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/SLIP_25_11_new_models/data_clean/csv/centiles_all_median_2026-04-24.csv",
  type = "median"
  )
}
``` 

# Load Packages
```{r, echo = FALSE}
#Load libraries
require(dplyr)
require(magrittr)
require(readr)
require(stringr)
require(tidyr)
require(gamlssTools) #Margaret's package
require(tibble)
#For figures/plotting
require(ggplot2)
require(gridExtra)
require(ggsegDKT)
require(ggseg)
require(glue)
#require(ggpubr)
require(ggcorrplot)
require(cowplot)
require(egg)
require(interactions)
require(growthstandards)
require(gtsummary)
require(svglite)
```


# Controls
```{r, echo = TRUE}
#set command
save_tables <- TRUE
plot <- TRUE
save_figs <- TRUE
set.seed(42)

output_folder <- params$output_folder
centiles_distinct_path <- params$centiles_distinct_path
centiles_all_path <- params$centiles_all_path
type <- params$type
print(centiles_distinct_path)
print(centiles_all_path)
print(glue("CENTILE TYPE: {type}"))
```

# Define Paths
```{r, echo = TRUE}
#set folder paths
code_path <- "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/"
vars_path <-  "/mnt/isilon/bgdlab_processing/Eren/slip_premie_wip/output/vars_2511_new/"

## output/save paths
if (!dir.exists(output_folder)) dir.create(output_folder)
csv_save_path <- glue("{output_folder}csv/")
if (!dir.exists(csv_save_path)) dir.create(csv_save_path)
plot_save_path <- paste0(output_folder, "plots/")
if (!dir.exists(plot_save_path)) dir.create(plot_save_path)
```

# Load custom functions
```{r, echo = FALSE}
panel_save <- function(plot, filename, extension, folder, plot_width = 8, plot_height = 10){
  save_name <- paste0(folder, filename, ".", extension)
  ggsave(save_name, plot = plot, dpi = 300, width = plot_width, height = plot_height)
}
```

# Plots specifications
```{r, echo = TRUE}
limits = c(-0.3, 0.3)
low_color = "#4878CF"
mid_color = "white"
high_color = "#D65F5F"
```

# Read in data files and vars
```{r, echo = FALSE}
full_vars <- read.csv(paste0(vars_path,"all_vars.csv")) %>% pull(x)
global_vars <- read.csv(paste0(vars_path,"global_vars.csv")) %>% pull(x)
vol_vars <- read.csv(paste0(vars_path,"vol_vars.csv")) %>% pull(x)
sa_vars <- read.csv(paste0(vars_path,"sa_vars.csv")) %>% pull(x)
th_vars <- read.csv(paste0(vars_path,"th_vars.csv")) %>% pull(x)

cent_distinct <- read.csv(centiles_distinct_path) %>% select(-("X"))
cent_all <- read.csv(centiles_all_path) %>% select(-("X"))
```

# Age Distribution Stats

## Cent-All
```{r, echo = FALSE}
N_repeat <- (nrow(cent_all) - length(unique(cent_all$participant_id)))

print(glue("Range of adjusted age in centiles_all {range(cent_all$age_years_adj)}"))
print(glue("IQR of adjusted age in centiles_all {IQR(cent_all$age_years_adj)}"))
adj_age_quartiles_all <- quantile(cent_all$age_years_adj, probs = c(0.25, 0.75))
print(glue("25th centile of adjusted age in centiles_all {adj_age_quartiles_all[1]}"))
print(glue("75th centile of adjusted age in centiles_all {adj_age_quartiles_all[2]}"))
print(glue("Median of adjusted age in centiles_all {median(cent_all$age_years_adj)}"))

print(glue("Range of non-adj age in centiles_all {range(cent_all$age_at_scan_years)}"))
print(glue("IQR of non-adj age in centiles_all {IQR(cent_all$age_at_scan_years)}"))
unadj_age_quartiles_all <- quantile(cent_all$age_at_scan_years, probs = c(0.25, 0.75))
print(glue("25th centile of non-adj age in centiles_all {unadj_age_quartiles_all[1]}"))
print(glue("75th centile of non-adj age in centiles_all {unadj_age_quartiles_all[2]}"))
print(glue("Median of non-adj age in centiles_all {median(cent_all$age_at_scan_years)}"))

(adj_age_dist_all <- ggplot(cent_all, aes(x = age_years_adj)) +
  geom_histogram(bins = 30, fill = "#dea1c5", color = "white") +
  labs(x = "Adjusted Age (yrs)", y = "", title = glue("All, N = {nrow(cent_all)} Data Points, {N_repeat} repeat ids")) +
  theme_classic())

(unadj_age_dist_all <- ggplot(cent_all, aes(x = age_at_scan_years)) +
  geom_histogram(bins = 30, fill = "#dea1c5", color = "white") +
  labs(x = "Age At Scan, UnAdj (yrs)", y = "", title = glue("All, N = {nrow(cent_all)} Data Points, {N_repeat} repeat ids")) +
  theme_classic())

if(save_figs == TRUE){
panel_save(adj_age_dist_all, glue("adj-age-dist-all-{type}"), "svg", plot_save_path, plot_width= 14, plot_height = 11)
panel_save(unadj_age_dist_all, glue("unadj-age-dist-all-{type}"), "svg", plot_save_path, plot_width= 14, plot_height = 11)
}
```

## Cent-Distinct (first scan for each subject)
```{r, echo = FALSE}
print(glue("Range of adjusted age in centiles_distinct {range(cent_distinct$age_years_adj)}"))
print(glue("IQR of adjusted age in centiles_distinct {IQR(cent_distinct$age_years_adj)}"))
adj_age_quartiles_distinct <- quantile(cent_distinct$age_years_adj, probs = c(0.25, 0.75))
print(glue("25th centile of adjusted age in centiles_distinct {adj_age_quartiles_distinct[1]}"))
print(glue("75th centile of adjusted age in centiles_distinct {adj_age_quartiles_distinct[2]}"))
print(glue("Median of adjusted age in centiles_distinct {median(cent_distinct$age_years_adj)}"))

print(glue("Range of non-adj age in centiles_distinct {range(cent_distinct$age_at_scan_years)}"))
print(glue("IQR of non-adj age in centiles_distinct {IQR(cent_distinct$age_at_scan_years)}"))
unadj_age_quartiles_distinct <- quantile(cent_distinct$age_at_scan_years, probs = c(0.25, 0.75))
print(glue("25th centile of non-adj age in centiles_distinct {unadj_age_quartiles_distinct[1]}"))
print(glue("75th centile of non-adj age in centiles_distinct {unadj_age_quartiles_distinct[2]}"))
print(glue("Median of non-adj age in centiles_distinct {median(cent_distinct$age_at_scan_years)}"))

(adj_age_dist_distinct <- ggplot(cent_distinct, aes(x = age_years_adj)) +
  geom_histogram(bins = 30, fill = "#dea1c5", color = "white") +
  labs(x = "Adjusted Age (yrs)", y = "", title = glue("Distinct, N = {nrow(cent_distinct)} Data Points, first scan per id")) +
  theme_classic())

(unadj_age_dist_distinct <- ggplot(cent_distinct, aes(x = age_at_scan_years)) +
  geom_histogram(bins = 30, fill = "#dea1c5", color = "white") +
  labs(x = "Age At Scan, UnAdj (yrs)", y = "", title = glue("Distinct, N = {nrow(cent_distinct)} Data Points, first scan per id")) +
  theme_classic())

if(save_figs == TRUE){
panel_save(adj_age_dist_distinct, glue("adj-age-dist-distinct-{type}"), "svg", plot_save_path, plot_width= 14, plot_height = 11)
panel_save(unadj_age_dist_distinct, glue("unadj-age-dist-distinct-{type}"), "svg", plot_save_path, plot_width= 14, plot_height = 11)
}
```

# GA/BW numbers
In Cent-Distinct, one row per subject.
```{r, echo = FALSE}
print(glue("Number of Participants with gestational age (post-QC) {sum(!is.na(cent_distinct$gestational_age))}"))
print(glue("Number of Participants with birth weight (post-QC) {sum(!is.na(cent_distinct$birth_weight_kg))}"))
print(glue("Number of Participants with BOTH birth weight and gestational age {sum(!is.na(cent_distinct$birth_weight_kg) & !is.na(cent_distinct$gestational_age))}"))
```

# Sex Distribution in Sample and Sub-samples
In Cent-Distinct, one row per subject.
```{r, echo = FALSE}
print("Whole Sample")
prop.table(table(cent_distinct$sex))
print("GA Available")
prop.table(table(cent_distinct[which(!is.na(cent_distinct$gestational_age)),]$sex))
print("BW Available")
prop.table(table(cent_distinct[which(!is.na(cent_distinct$birth_weight_kg)),]$sex))
print("Both GA and BW Available")
prop.table(table(cent_distinct[which(!is.na(cent_distinct$birth_weight_kg) & !is.na(cent_distinct$gestational_age)),]$sex))
```

# Plot age distribution and birth factor availability

## Cent-All
```{r, echo = FALSE}
#Whole sample (one session per participant)
N_total <- nrow(cent_all)
N_ga <- sum(!is.na(cent_all$gestational_age))
N_bw <- sum(!is.na(cent_all$birth_weight_kg))
N_ga_only <- sum(!is.na(cent_all$gestational_age) & is.na(cent_all$birth_weight_kg))
N_bw_only <- sum(!is.na(cent_all$birth_weight_kg) & is.na(cent_all$gestational_age))
N_gabw <- sum(!is.na(cent_all$gestational_age) & !is.na(cent_all$birth_weight_kg))
N_none <- sum(is.na(cent_all$gestational_age) & is.na(cent_all$birth_weight_kg))


(age_dist2_plot_all <- ggplot(cent_all, aes(x = age_years_adj)) +
  geom_histogram(binwidth= 0.1, fill = "#B3B3B3", color = "white", lineplot_width= 0.1) +
  labs(
    x = paste0("Adjusted Age (yrs),  N = ", N_total),
    y = NULL,
    title = "All Participants for Growth Charts"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(hjust = 0.5)
  ))


if(save_figs == TRUE){
panel_save(age_dist2_plot_all, "age-dist2-all", "svg", plot_save_path, plot_width= 14, plot_height = 11)
}

df_age_all <- cent_all %>%
  mutate(
    ga_known  = !is.na(gestational_age),
    bw_known  = !is.na(birth_weight_kg),
    group = case_when(
      ga_known & bw_known ~ "GA & BW Known",
      ga_known & !bw_known ~ "Only GA Known",
      bw_known & !ga_known ~ "Only BW Known",
      TRUE ~ "None"
    )) %>%
  mutate(group = factor(group, levels = c(
    "Only GA Known",
    "Only BW Known",
    "GA & BW Known",
    "None"
  )))

(age_dist_gabw_plot_all <- ggplot(df_age_all, aes(x = age_years_adj, fill = group)) +
  geom_histogram(binwidth= 0.1, color = "white", lineplot_width= 0.1) +
  scale_fill_manual(
    values = c(
      "Only GA Known" = "#9EC3D2",
      "Only BW Known" = "#D6B8A8",
      "GA & BW Known" = "#6A5A78",
      "None" = "#E0E0E0"
    ),
    labels = c(
      paste0("Only GA Known, N=", N_ga_only),
      paste0("Only BW Known, N=", N_bw_only),
      paste0("GA & BW Known, N=", N_gabw),
      paste0("Neither Known, N=", N_none)
    ),name = NULL) +
  labs(
    x = paste0("Adjusted Age (years),  N = ", N_total),
    y = NULL
  ) +
  theme_minimal(base_size = 16) +
  theme_bw(base_size = 32) +
theme(
  legend.position = "bottom",
  legend.direction = "horizontal",
  legend.box = "horizontal",
  legend.key.plot_width= unit(1.5, "cm"),
  legend.box.just = "center",
  panel.grid = element_blank()
) +
guides(fill = guide_legend(nrow = 2)))

if(save_figs == TRUE){
panel_save(age_dist_gabw_plot_all, glue("age-dist-gabw-all-{type}"), "svg", plot_save_path, plot_width= 14, plot_height = 11)
}

#GA available
N_age_ga <- sum(!is.na(cent_all$gestational_age))

(age_dist_ga_plot_all <- ggplot(cent_all %>% filter(!is.na(gestational_age)), aes(x = age_years_adj)) +
  geom_histogram(binwidth= 0.1, fill = "#9EC3D2", color = "white", lineplot_width= 0.1) +
  labs(
    x = paste0("Adjusted Age (years),  N = ", N_age_ga),
    y = NULL,
    title = "Gestational Age Known"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(hjust = 0.5)
  ))

if(save_figs == TRUE){
panel_save(age_dist_ga_plot_all, "age-dist-ga-all", "svg", plot_save_path, plot_width= 14, plot_height = 11)
}

#BW available
N_age_bw <- sum(!is.na(cent_all$birth_weight_kg))

(age_dist_bw_plot_all <- ggplot(cent_all %>% filter(!is.na(birth_weight_kg)), aes(x = age_years_adj)) +
  geom_histogram(binwidth= 0.1, fill = "#D6B8A8", color = "white", lineplot_width= 0.1) +
  labs(
    x = paste0("Adjusted Age (years),  N = ", N_age_bw),
    y = NULL,
    title = "Birthweight Known"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(hjust = 0.5)
  ))

if(save_figs == TRUE){
panel_save(age_dist_bw_plot_all, glue("age-dist-bw-all-{type}"), "svg", plot_save_path, plot_width= 14, plot_height = 11)
}
```

## Cent-Distinct
```{r, echo = FALSE}
#Whole sample (one session per participant)
N_total <- nrow(cent_distinct)
N_ga <- sum(!is.na(cent_distinct$gestational_age))
N_bw <- sum(!is.na(cent_distinct$birth_weight_kg))
N_ga_only <- sum(!is.na(cent_distinct$gestational_age) & is.na(cent_distinct$birth_weight_kg))
N_bw_only <- sum(!is.na(cent_distinct$birth_weight_kg) & is.na(cent_distinct$gestational_age))
N_gabw <- sum(!is.na(cent_distinct$gestational_age) & !is.na(cent_distinct$birth_weight_kg))
N_none <- sum(is.na(cent_distinct$gestational_age) & is.na(cent_distinct$birth_weight_kg))


(age_dist2_plot_distinct <- ggplot(cent_distinct, aes(x = age_years_adj)) +
  geom_histogram(binwidth= 0.1, fill = "#B3B3B3", color = "white", lineplot_width= 0.1) +
  labs(
    x = paste0("Adjusted Age (years),  N = ", N_total),
    y = NULL,
    title = "All Participants for Growth Charts, distinct id rows"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(hjust = 0.5)
  ))


if(save_figs == TRUE){
panel_save(age_dist2_plot_distinct, glue("age-dist2-distinct-{type}"), "svg", plot_save_path, plot_width= 14, plot_height = 11)
}

df_age_distinct <- cent_distinct %>%
  mutate(
    ga_known  = !is.na(gestational_age),
    bw_known  = !is.na(birth_weight_kg),
    group = case_when(
      ga_known & bw_known ~ "GA & BW Known",
      ga_known & !bw_known ~ "Only GA Known",
      bw_known & !ga_known ~ "Only BW Known",
      TRUE ~ "None"
    )) %>%
  mutate(group = factor(group, levels = c(
    "Only GA Known",
    "Only BW Known",
    "GA & BW Known",
    "None"
  )))

(age_dist_gabw_plot_distinct <- ggplot(df_age_distinct, aes(x = age_years_adj, fill = group)) +
  geom_histogram(binwidth= 0.1, color = "white", lineplot_width= 0.1) +
  scale_fill_manual(
    values = c(
      "Only GA Known" = "#9EC3D2",
      "Only BW Known" = "#D6B8A8",
      "GA & BW Known" = "#6A5A78",
      "None" = "#E0E0E0"
    ),
    labels = c(
      paste0("Only GA Known, N=", N_ga_only),
      paste0("Only BW Known, N=", N_bw_only),
      paste0("GA & BW Known, N=", N_gabw),
      paste0("Neither Known, N=", N_none)
    ),name = NULL) +
  labs(
    x = paste0("Adjusted Age (years),  N = ", N_total),
    y = NULL,
    title = "Distinct id rows"
  ) +
  theme_minimal(base_size = 16) +
theme(
  legend.position = "bottom",
  legend.direction = "horizontal",
  legend.box = "horizontal",
  legend.text = element_text(size = 14),
  legend.key.plot_width= unit(1.5, "cm"),
  legend.box.just = "center",
  panel.grid = element_blank()
) +
guides(fill = guide_legend(nrow = 2)))

if(save_figs == TRUE){
panel_save(age_dist_gabw_plot_distinct, glue("age-dist-gabw-distinct-{type}"), "svg", plot_save_path, plot_width= 14, plot_height = 11)
}

#GA available
N_age_ga <- sum(!is.na(cent_distinct$gestational_age))

(age_dist_ga_plot_distinct <- ggplot(cent_distinct %>% filter(!is.na(gestational_age)), aes(x = age_years_adj)) +
  geom_histogram(binwidth= 0.1, fill = "#9EC3D2", color = "white", lineplot_width= 0.1) +
  labs(
    x = paste0("Adjusted Age (years),  N = ", N_age_ga),
    y = NULL,
    title = "Gestational Age Known, distinct id rows"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(hjust = 0.5)
  ))

if(save_figs == TRUE){
panel_save(age_dist_ga_plot_distinct, "age-dist-ga-distinct", "svg", plot_save_path, plot_width= 14, plot_height = 11)
}

#BW available
N_age_bw <- sum(!is.na(cent_distinct$birth_weight_kg))

(age_dist_bw_plot_distinct <- ggplot(cent_distinct %>% filter(!is.na(birth_weight_kg)), aes(x = age_years_adj)) +
  geom_histogram(binwidth= 0.1, fill = "#D6B8A8", color = "white", lineplot_width= 0.1) +
  labs(
    x = paste0("Adjusted Age (years),  N = ", N_age_bw),
    y = NULL,
    title = "Birthweight Known, distinct id rows"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(hjust = 0.5)
  ))

if(save_figs == TRUE){
panel_save(age_dist_bw_plot_distinct, glue("age-dist-bw-distinct-{type}"), "svg", plot_save_path, plot_width= 14, plot_height = 11)
}
```

# Gestational Age Plots
Within Distinct, so each subject is represented only once.
```{r, echo = FALSE}
table(cent_distinct$preterm)

ga_ptb <- cent_distinct %>%
  filter(!is.na(gestational_age)) %>%
  mutate(group = ifelse(gestational_age_weeks_complete < 37, "Preterm", "Term"))

pct_preterm <- round(mean(ga_ptb$group == "Preterm") * 100, 1)

ga_breaks_weeks <- seq(23, 43, by = 2)
ga_breaks_days  <- ga_breaks_weeks

(ga_preterm_dist_plot <- ggplot(ga_ptb, aes(x = gestational_age_weeks_complete, fill = group)) +
  geom_histogram(binwidth = 1, color = "white") +
  scale_fill_manual(values = c("Preterm" = "#F4A475", "Term" = "#A7D9A0")) +
    scale_x_continuous(breaks = ga_breaks_days, labels = ga_breaks_weeks) +
  labs(
    x = paste0("Gestational Age (complete weeks), \n % Preterm(<37wks) = ", pct_preterm),
    y = NULL
  ) +
  theme_minimal(base_size = 16) +
  theme(
    panel.grid = element_blank(),
    legend.position = "none"
  ))

if(save_figs == TRUE){
panel_save(ga_preterm_dist_plot, glue("ga-preterm-dist-plot-{type}"), "svg", plot_save_path, plot_width= 10, plot_height = 7)
}

cent_distinct <- cent_distinct %>%
  mutate(preterm = factor(preterm, levels = c("VPM", "LPM", "Term")))

(ga_preterm_age_plot <- ggplot(cent_distinct %>% filter(!is.na(preterm)), aes(x = age_years_adj, fill = preterm)) +
  geom_histogram(position = "fill", binwidth= 0.5, color = "white") +
  scale_fill_manual(
    values = c(
      "VPM"  = "#E8A0A8",
      "LPM"  = "#A8C4E8",
      "Term" = "#A7D9A0"
    )
  ) +
  labs(
    x = "Adjusted Age (years)",
    y = "Proportion",
    fill = "Preterm Status"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    panel.grid = element_blank(),
    legend.position = "bottom",
    legend.direction = "horizontal"
  ))

if(save_figs == TRUE){
panel_save(ga_preterm_age_plot, glue("ga-preterm-byage-plot-{type}"), "svg", plot_save_path, plot_width= 10, plot_height = 7)
}
```

# Birth Weight Plots
Within Distinct, so each subject is represented only once.
```{r, echo = FALSE}
cent_distinct$bw_category <- as.factor(cut(cent_distinct$birth_weight_kg, breaks = c(-Inf, 1.5, 2.5, Inf), labels = c("VLBW", "LBW", "Normal BW"), include.lowest = TRUE))

bw_low <- cent_distinct %>%
  filter(!is.na(birth_weight_kg)) %>%
  mutate(group = ifelse(birth_weight_kg < 2.5, "Low BW", "Normal BW"))

pct_lowbw <- round(mean(bw_low$group == "Low BW") * 100, 1)

(bw_low_dist_plot <- ggplot(bw_low, aes(x = birth_weight_kg, fill = group)) +
    geom_histogram(
    binwidth = 0.1,
    boundary = 0,
    color = "white"
  ) +
  scale_fill_manual(values = c("Low BW" = "#C76E6A", "Normal BW" = "#7FAF8A")) +
  labs(
    x = paste0("Birth Weight(kg),\n % Low(<2.5kg) = ", pct_lowbw),
    y = NULL
  ) +
  theme_minimal(base_size = 16) +
  theme(
    panel.grid = element_blank(),
    legend.position = "none"
  ))

if(save_figs == TRUE){
panel_save(bw_low_dist_plot, glue("bw-low-dist-plot-{type}"), "svg", plot_save_path, plot_width= 10, plot_height = 7)
}

(bw_low_age_plot <- ggplot(cent_distinct %>% filter(!is.na(bw_category)), aes(x = age_years_adj, fill = bw_category)) +
  geom_histogram(position = "fill", binwidth= 0.5, color = "white") +
  scale_fill_manual(
    values = c(
      "VLBW"  = "#B9836A",
      "LBW"  = "#6A8F9F",
      "Normal BW" = "#7FAF8A"
    )
  ) +
  labs(
    x = "Adjusted Age (years)",
    y = "Proportion",
    fill = "Birthweight"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    panel.grid = element_blank(),
    legend.position = "bottom",
    legend.direction = "horizontal"
  ))

if(save_figs == TRUE){
panel_save(bw_low_age_plot, glue("bw-low-byage-plot-{type}"), "svg", plot_save_path, plot_width= 10, plot_height = 7)
}

cent_distinct$age_group <- cut(cent_distinct$age_years_adj, breaks = 80, include.lowest = TRUE, labels = FALSE)

cent_distinct <- cent_distinct %>%
  group_by(age_group) %>%
  arrange(birth_weight_kg, .by_group = TRUE) %>%
  ungroup()

#for x axis labeling
age_labels <- c(0, 5, 10, 15, 20)
group_width<- (max(cent_distinct$age_years_adj) - min(cent_distinct$age_years_adj)) / 80
x_breaks <- floor((age_labels - min(cent_distinct$age_years_adj)) / group_width) + 1
#fix position of 0
x_breaks[1] <- 1

(bw_cont_age_plot <- ggplot(cent_distinct %>% filter(!is.na(birth_weight_kg)), aes(x = age_group, y = 1, fill = birth_weight_kg)) + geom_bar(stat = "identity", plot_width= 0.8) +
    scale_fill_gradient2(low = "#F2EAD3", mid = "#E6E2C8", high = "#7FAF8A", limits = range(cent_distinct$birth_weight_kg, na.rm = TRUE)) +
  labs(
    x = "Adjusted Age (years)",
    y = NULL,
    fill = "Birthweight"
  ) +
  scale_x_continuous(
      breaks = x_breaks,
      labels = age_labels
    ) +
  theme_minimal(base_size = 16) +
  theme(
    panel.grid = element_blank(),
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.title = element_text(margin = margin(r = 10))
  ) +
  guides(fill = guide_colorbar(
    barwidth= 10,
    barheight = 1,
    title.position = "left",
    title.vjust = 1
  )))

if(save_figs == TRUE){
panel_save(bw_cont_age_plot, glue("bw-cont-byage-plot-{type}"), "svg", plot_save_path, plot_width= 10, plot_height = 7)
}
```

# Birthweight Percentile Plots
Within Distinct, so each subject is represented only once.
Using Fenton-calculated ones (new)
```{r, echo = FALSE}
if(!("age_group" %in% names(cent_distinct))){
cent_distinct$age_group <- cut(cent_distinct$age_years_adj, breaks = 80, include.lowest = TRUE, labels = FALSE)
}

cent_distinct <- cent_distinct %>%
  group_by(age_group) %>%
  arrange(bwp_fen, .by_group = TRUE) %>%
  ungroup()

#for x axis labeling
age_labels <- c(0, 5, 10, 15, 20)
group_width<- (max(cent_distinct$age_years_adj) - min(cent_distinct$age_years_adj)) / 80
x_breaks <- floor((age_labels - min(cent_distinct$age_years_adj)) / group_width) + 1
#fix position of 0
x_breaks[1] <- 1

(bwp_cont_age_plot <- ggplot(cent_distinct %>% filter(!is.na(bwp_fen)), aes(x = age_group, y = 1, fill = bwp_fen)) + geom_bar(stat = "identity", plot_width= 0.8) +
    scale_fill_gradient2(low = "#F2EAD3", mid = "#E6E2C8", high = "#7FAF8A", limits = c(0,1)) +
  labs(
    x = "Adjusted Age (years)",
    y = NULL,
    fill = "Birthweight Percentile"
  ) +
  scale_x_continuous(
      breaks = x_breaks,
      labels = age_labels
    ) +
  theme_minimal(base_size = 16) +
  theme(
    panel.grid = element_blank(),
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.title = element_text(margin = margin(r = 10))
  ) +
  guides(fill = guide_colorbar(
    barwidth= 10,
    barheight = 1,
    title.position = "left",
    title.vjust = 1
  )))

if(save_figs == TRUE){
panel_save(bwp_cont_age_plot, glue("bwp-cont-byage-plot-{type}"), "svg", plot_save_path, plot_width= 10, plot_height = 7)
}
```

# Ages of repeat measurements plot
```{r, echo = FALSE}
cent_all <- cent_all %>%
  group_by(participant_id) %>%
  mutate(age_at_first_scan_years = min(age_at_scan_years, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(participant_id = reorder(participant_id, age_at_first_scan_years))

(repeat_scans_plot <- ggplot(cent_all, aes(y = participant_id, group = participant_id)) +
   geom_point(aes(x = age_years_adj, color = preterm), linewidth = 0.1) +
   geom_line(aes(x = age_years_adj, color = preterm), linewidth = 0.075) +
   scale_color_manual(values = c(
      "VPM"  = "#E8A0A8",
      "LPM"  = "#A8C4E8",
      "Term" = "#A7D9A0"),
      name = "") +
   theme_classic() +
   labs(x = "Adjusted Age at Scan (yrs)", y = "Participant") +
   theme(legend.position = "bottom",
         axis.ticks.y = element_blank(),
         axis.text.y = element_blank()))

if(save_figs == TRUE){
panel_save(repeat_scans_plot, glue("repeat-scans-plot-{type}"), "svg", plot_save_path, plot_width= 14, plot_height = 11)
}
```
